---
title: "Figure 6J: Strongest Path Valence Pattern Comparison Across Modalities"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 10, fig.height = 8)
source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel J: Valence Pattern Comparison — Compares neurotransmitter sequence patterns of the single strongest path per (modality, target) pair. Extracts E/I identity of each hop to form valence pattern strings (e.g. E_I_E) and asks whether different modalities reach the same target via the same or different neurotransmitter sequences."
panel_legend <- "(J) Neurotransmitter valence patterns along strongest paths. For each modality-target pair, the single strongest path is decomposed into its hop-by-hop excitatory/inhibitory (E/I) sequence. Heatmaps show the frequency of each valence pattern across modalities, revealing conserved or divergent signalling motifs."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_J_description", "txt"), "panel_J_description.txt"))
```

## Overview

This panel compares the **neurotransmitter sequence patterns** of the strongest
single path from each sensory modality to each 3rd-order target neuron. For
each (modality, target) pair we take only the single strongest path, extract
the E/I identity of each hop (excitatory or inhibitory based on presynaptic NT),
and form a "valence pattern" string (e.g. `E_I_E`). The key question is:
**do different modalities reach the same target neuron via the same or different
neurotransmitter sequences?**

## 1. Load row order from Panel B

```{r load-row-order}
# Use Panel B row order for consistent ordering across panels
row_order_file <- file.path(fig6_dir, "panel_B_row_order.rds")
use_ref_order <- file.exists(row_order_file)

if (use_ref_order) {
  row_order_ref <- readRDS(row_order_file)
  cat("Loaded Panel B row order:", length(row_order_ref), "neurons\n")
} else {
  cat("Panel B row order not found -- will cluster independently.\n")
}

# Row normalization: scale each row by its max absolute value -> [-1, 1]
row_normalize <- function(mat) {
  row_max <- apply(abs(mat), 1, max, na.rm = TRUE)
  row_max[row_max == 0] <- 1
  sweep(mat, 1, row_max, "/")
}
```

## 2. Derive neurotransmitter sequence for each path

```{r derive-nt-sequence}
# Build NT lookup from mba
nt_lookup <- mba %>%
  select(type, consensus_nt) %>%
  distinct() %>%
  group_by(type) %>%
  summarise(nt = first(consensus_nt), .groups = "drop")
nt_by_type <- setNames(nt_lookup$nt, nt_lookup$type)

inhibitory_nts <- c("gaba", "glycine", "histamine")

# Extract valence pattern from a path string
# For path "A -> B -> C -> D": NT of A,B,C determines E/I at each hop
get_valence_pattern <- function(path_str) {
  nodes <- strsplit(path_str, " -> ", fixed = TRUE)[[1]]
  transmitting <- nodes[-length(nodes)]
  nts <- nt_by_type[transmitting]
  pattern <- ifelse(nts %in% inhibitory_nts, "I", "E")
  paste(pattern, collapse = "_")
}

add_valence_pattern <- function(df) {
  df %>%
    filter(!is.na(path)) %>%
    rowwise() %>%
    mutate(valence_pattern = get_valence_pattern(path)) %>%
    ungroup()
}

paths_patterned <- add_valence_pattern(paths.aspf_aspg)
cat(sprintf("Total paths with patterns: %d\n", nrow(paths_patterned)))
cat(sprintf("Unique valence patterns: %d\n", n_distinct(paths_patterned$valence_pattern)))
```

## 3. Select absolute strongest single path per (modality, target)

```{r select-strongest}
# Strongest per (modality, end, valence) — for signed strength later
strongest_by_valence <- paths_patterned %>%
  group_by(modality, end, valence) %>%
  slice_max(strength, n = 1, with_ties = FALSE) %>%
  ungroup()

# For each (modality, end) pair, keep only THE single strongest path
absolute_strongest <- paths_patterned %>%
  group_by(modality, end) %>%
  slice_max(strength, n = 1, with_ties = FALSE) %>%
  ungroup()

cat(sprintf("Absolute strongest paths: %d\n", nrow(absolute_strongest)))
cat(sprintf("  Excitatory: %d (%.0f%%)\n",
            sum(absolute_strongest$valence == "excitatory"),
            100 * mean(absolute_strongest$valence == "excitatory")))
cat(sprintf("  Inhibitory: %d (%.0f%%)\n",
            sum(absolute_strongest$valence == "inhibitory"),
            100 * mean(absolute_strongest$valence == "inhibitory")))
```

---

# Cross-modality pattern comparison

## 4. Valence tile map: target neuron x modality

For each target neuron, is the strongest path from each modality overall
excitatory or inhibitory? Red = excitatory, blue = inhibitory.

```{r valence-tile, fig.width=12, fig.height=14}
valence_colors <- c("excitatory" = "#B2182B", "inhibitory" = "#2166AC", "none" = "grey95")

# Build binary valence matrix for clustering (1 = excitatory, 0 = inhibitory)
valence_wide <- absolute_strongest %>%
  distinct(end, modality, .keep_all = TRUE) %>%
  mutate(val_num = ifelse(valence == "excitatory", 1, 0)) %>%
  select(end, modality, val_num) %>%
  pivot_wider(names_from = modality, values_from = val_num, values_fill = 0.5) %>%
  column_to_rownames("end")
valence_wide <- as.matrix(valence_wide[, intersect(modality_order, colnames(valence_wide))])

# Cluster rows and columns
row_hc <- hclust(dist(valence_wide, method = "binary"), method = "ward.D2")
col_hc <- hclust(dist(t(valence_wide), method = "binary"), method = "ward.D2")

target_order <- rownames(valence_wide)[row_hc$order]
mod_order_clust <- colnames(valence_wide)[col_hc$order]

# Convert to long for ggplot with clustered order
valence_long <- absolute_strongest %>%
  select(end, modality, valence, strength) %>%
  mutate(
    end      = factor(end, levels = target_order),
    modality = factor(modality, levels = mod_order_clust)
  )

p_tile <- ggplot(valence_long, aes(x = modality, y = end, fill = valence)) +
  geom_tile(color = "white", linewidth = 0.3) +
  scale_fill_manual(values = valence_colors, name = "Overall\nvalence") +
  labs(
    title = "Overall valence of strongest path: target neuron x modality",
    subtitle = "Red = excitatory, Blue = inhibitory (clustered by row & column)",
    x = "Sensory modality",
    y = "Target neuron"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.y = element_text(size = 5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    panel.grid = element_blank(),
    legend.position = "right"
  )

p_tile

save_fig6_plot(p_tile, "panel_J_valence_tile_map_aspf_aspg", width = 12, height = 14)
```

## 5. Valence conservation across modalities

For each target neuron: do all modalities reach it with the same overall
valence (all excitatory or all inhibitory), or is it mixed?

```{r valence-conservation, fig.width=10, fig.height=6}
conservation <- absolute_strongest %>%
  group_by(end) %>%
  summarise(
    n_modalities = n_distinct(modality),
    n_excitatory = sum(valence == "excitatory"),
    n_inhibitory = sum(valence == "inhibitory"),
    pct_excitatory = round(100 * mean(valence == "excitatory"), 1),
    category = case_when(
      all(valence == "excitatory") ~ "all excitatory",
      all(valence == "inhibitory") ~ "all inhibitory",
      TRUE ~ "mixed"
    ),
    .groups = "drop"
  )

# Bar chart of categories
cat_counts <- conservation %>% count(category)

p_cons <- ggplot(cat_counts, aes(x = category, y = n, fill = category)) +
  geom_col(color = "grey30", linewidth = 0.3) +
  scale_fill_manual(values = c("all excitatory" = "#B2182B",
                                "all inhibitory" = "#2166AC",
                                "mixed" = "#999999"),
                    guide = "none") +
  geom_text(aes(label = n), vjust = -0.3, size = 5) +
  labs(
    title = "Valence conservation: do all modalities agree on excitatory vs inhibitory?",
    x = NULL, y = "Number of target neurons"
  ) +
  theme_minimal(base_size = 13)

p_cons

save_fig6_plot(p_cons, "panel_J_valence_conservation_aspf_aspg", width = 8, height = 5)

cat(sprintf("All excitatory: %d (%.0f%%)\n",
            sum(conservation$category == "all excitatory"),
            100 * mean(conservation$category == "all excitatory")))
cat(sprintf("All inhibitory: %d (%.0f%%)\n",
            sum(conservation$category == "all inhibitory"),
            100 * mean(conservation$category == "all inhibitory")))
cat(sprintf("Mixed: %d (%.0f%%)\n",
            sum(conservation$category == "mixed"),
            100 * mean(conservation$category == "mixed")))
```

```{r mixed-detail, fig.width=10, fig.height=6}
# For mixed targets: what fraction of modalities are excitatory?
mixed_targets <- conservation %>% filter(category == "mixed")

if (nrow(mixed_targets) > 0) {
  p_mixed <- ggplot(mixed_targets, aes(x = pct_excitatory)) +
    geom_histogram(binwidth = 10, fill = "#999999", color = "white", alpha = 0.8) +
    labs(
      title = "Mixed-valence targets: % of modalities that are excitatory",
      x = "% modalities excitatory",
      y = "Number of target neurons"
    ) +
    theme_minimal(base_size = 12)

  p_mixed
  save_fig6_plot(p_mixed, "panel_J_mixed_valence_detail_aspf_aspg", width = 10, height = 5)
}
```

## 6. Valence agreement between modalities

For each pair of modalities, what fraction of shared target neurons have the
**same** overall valence? High agreement means those modalities drive the same
targets in the same direction.

```{r valence-similarity, fig.width=8, fig.height=7}
mod_list <- modality_order
sim_mat <- matrix(NA, nrow = length(mod_list), ncol = length(mod_list),
                  dimnames = list(mod_list, mod_list))

valence_by_target <- absolute_strongest %>%
  select(end, modality, valence) %>%
  pivot_wider(names_from = modality, values_from = valence)

for (i in seq_along(mod_list)) {
  for (j in seq_along(mod_list)) {
    m1 <- mod_list[i]; m2 <- mod_list[j]
    if (m1 %in% names(valence_by_target) && m2 %in% names(valence_by_target)) {
      both_present <- !is.na(valence_by_target[[m1]]) & !is.na(valence_by_target[[m2]])
      if (sum(both_present) > 0) {
        sim_mat[i, j] <- mean(valence_by_target[[m1]][both_present] ==
                              valence_by_target[[m2]][both_present])
      }
    }
  }
}

pheatmap(
  sim_mat, cluster_rows = TRUE, cluster_cols = TRUE,
  color = colorRampPalette(c("#F7F7F7", "#D1E5F0", "#4393C3", "#2166AC"))(100),
  border_color = "grey80", display_numbers = TRUE,
  number_format = "%.2f", fontsize_number = 9,
  fontsize_row = 11, fontsize_col = 11,
  main = "Valence agreement between modalities\n(fraction of shared targets with same overall valence)"
)
```

```{r save-similarity}
for (ext in c("png", "pdf")) {
  f <- file.path(get(paste0("plot_dir_", ext)),
                 paste0("panel_J_valence_agreement_matrix.", ext))
  if (ext == "png") png(f, width = 8, height = 7, units = "in", res = 300)
  else              pdf(f, width = 8, height = 7)
  pheatmap(sim_mat, cluster_rows = TRUE, cluster_cols = TRUE,
           color = colorRampPalette(c("#F7F7F7", "#D1E5F0", "#4393C3", "#2166AC"))(100),
           border_color = "grey80", display_numbers = TRUE,
           number_format = "%.2f", fontsize_number = 9,
           fontsize_row = 11, fontsize_col = 11,
           main = "Valence agreement between modalities")
  dev.off()
}
```

## 7. Valence x strength tile map

Same as section 4 but with strength encoded as opacity.

```{r valence-strength-tile, fig.width=12, fig.height=14}
p_tile_str <- ggplot(valence_long, aes(x = modality, y = end)) +
  geom_tile(aes(fill = valence, alpha = strength),
            color = "white", linewidth = 0.3) +
  scale_fill_manual(values = valence_colors, name = "Overall\nvalence") +
  scale_alpha_continuous(range = c(0.3, 1), name = "Strength") +
  labs(
    title = "Valence x strength: target neuron x modality",
    subtitle = "Color = excitatory/inhibitory; opacity = path strength",
    x = "Sensory modality",
    y = "Target neuron"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.y = element_text(size = 5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    panel.grid = element_blank()
  )

p_tile_str

save_fig6_plot(p_tile_str, "panel_J_valence_strength_tile_aspf_aspg", width = 12, height = 14)
```

---

# Valence pattern frequency & distribution

## 8. Valence pattern frequency (bar chart)

```{r valence-pattern-freq, fig.width=12, fig.height=6}
pattern_counts <- absolute_strongest %>%
  group_by(valence_pattern) %>%
  summarise(
    count = n(),
    mean_strength = mean(strength),
    median_strength = median(strength),
    .groups = "drop"
  ) %>%
  arrange(desc(count))

p_freq <- ggplot(pattern_counts, aes(x = reorder(valence_pattern, -count), y = count)) +
  geom_col(aes(fill = log10(mean_strength)), color = "grey30", linewidth = 0.3) +
  scale_fill_viridis_c(option = "plasma", name = "log10(mean\nstrength)") +
  labs(
    title = "Valence pattern frequency (strongest single path per modality-target pair)",
    x = "Valence pattern (E = excitatory hop, I = inhibitory hop)",
    y = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

p_freq

save_fig6_plot(p_freq, "panel_J_valence_pattern_frequency_aspf_aspg", width = 12, height = 6)
```

## 9. Valence pattern by modality (stacked bar)

```{r pattern-by-modality, fig.width=12, fig.height=6}
pattern_by_mod <- absolute_strongest %>%
  group_by(modality, valence_pattern) %>%
  summarise(count = n(), mean_strength = mean(strength), .groups = "drop")

p_mod_pattern <- ggplot(pattern_by_mod,
    aes(x = factor(modality, levels = modality_order), y = count, fill = valence_pattern)) +
  geom_col(position = "stack", color = "white", linewidth = 0.2) +
  scale_fill_brewer(palette = "Set3", name = "Valence\npattern") +
  labs(title = "Detailed valence patterns by modality",
       x = "Modality", y = "Number of target neurons") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_mod_pattern

save_fig6_plot(p_mod_pattern, "panel_J_pattern_by_modality_aspf_aspg", width = 12, height = 6)
```

## 10. Modality x valence pattern heatmap (mean signed strength)

```{r mod-pattern-heatmap, fig.width=12, fig.height=6}
diverging_palette <- colorRampPalette(
  c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
    "#F7F7F7",
    "#FDDBC7", "#F4A582", "#D6604D", "#B2182B")
)(100)

pattern_heatmap_df <- absolute_strongest %>%
  mutate(signed_str = ifelse(valence == "excitatory", strength, -strength)) %>%
  group_by(modality, valence_pattern) %>%
  summarise(mean_signed = mean(signed_str), count = n(), .groups = "drop")

pattern_mat <- pattern_heatmap_df %>%
  select(modality, valence_pattern, mean_signed) %>%
  pivot_wider(names_from = valence_pattern, values_from = mean_signed, values_fill = 0) %>%
  column_to_rownames("modality")
pattern_mat <- as.matrix(pattern_mat[intersect(modality_order, rownames(pattern_mat)), ])

col_order <- names(sort(colSums(abs(pattern_mat)), decreasing = TRUE))
pattern_mat <- pattern_mat[, col_order, drop = FALSE]

abs_max_p <- max(abs(pattern_mat), na.rm = TRUE)
if (abs_max_p == 0) abs_max_p <- 1
breaks_p <- seq(-abs_max_p, abs_max_p, length.out = 101)

pheatmap(
  pattern_mat, cluster_rows = FALSE, cluster_cols = FALSE,
  color = diverging_palette, breaks = breaks_p,
  border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
  main = "Mean signed strength by modality x valence pattern",
  angle_col = 45, display_numbers = TRUE,
  number_format = "%.4f", fontsize_number = 7
)
```

```{r save-mod-pattern-heatmap}
for (ext in c("png", "pdf")) {
  f <- file.path(get(paste0("plot_dir_", ext)),
                 paste0("panel_J_modality_x_valence_pattern_aspf_aspg.", ext))
  if (ext == "png") png(f, width = 12, height = 6, units = "in", res = 300)
  else              pdf(f, width = 12, height = 6)
  pheatmap(pattern_mat, cluster_rows = FALSE, cluster_cols = FALSE,
           color = diverging_palette, breaks = breaks_p,
           border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
           main = "Mean signed strength by modality x valence pattern",
           angle_col = 45, display_numbers = TRUE,
           number_format = "%.4f", fontsize_number = 7)
  dev.off()
}
```

## 11. Modality x valence pattern count heatmap

```{r mod-pattern-count-heatmap, fig.width=12, fig.height=6}
count_mat <- pattern_heatmap_df %>%
  select(modality, valence_pattern, count) %>%
  pivot_wider(names_from = valence_pattern, values_from = count, values_fill = 0) %>%
  column_to_rownames("modality")
count_mat <- as.matrix(count_mat[intersect(modality_order, rownames(count_mat)), ])
count_mat <- count_mat[, intersect(col_order, colnames(count_mat)), drop = FALSE]

pheatmap(
  count_mat, cluster_rows = FALSE, cluster_cols = FALSE,
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(50),
  border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
  main = "Count of strongest paths by modality x valence pattern",
  angle_col = 45, display_numbers = TRUE,
  number_format = "%.0f", fontsize_number = 8
)
```

```{r save-mod-pattern-count}
for (ext in c("png", "pdf")) {
  f <- file.path(get(paste0("plot_dir_", ext)),
                 paste0("panel_J_modality_x_pattern_count_aspf_aspg.", ext))
  if (ext == "png") png(f, width = 12, height = 6, units = "in", res = 300)
  else              pdf(f, width = 12, height = 6)
  pheatmap(count_mat, cluster_rows = FALSE, cluster_cols = FALSE,
           color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(50),
           border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
           main = "Count of strongest paths by modality x valence pattern",
           angle_col = 45, display_numbers = TRUE,
           number_format = "%.0f", fontsize_number = 8)
  dev.off()
}
```

---

# Signed strength heatmaps

## 12. Signed strength heatmap: modality x target neuron (row-normalized)

Using strongest excitatory and strongest inhibitory path per (modality, target).
Row-normalized so each target neuron's values span [-1, 1].
Uses Panel B row order if available.

```{r signed-strength-heatmap, fig.width=10, fig.height=12}
signed_df <- strongest_by_valence %>%
  mutate(signed_strength = ifelse(valence == "excitatory", strength, -strength)) %>%
  group_by(modality, end) %>%
  summarise(net_signed = sum(signed_strength), .groups = "drop")

signed_mat <- signed_df %>%
  pivot_wider(names_from = modality, values_from = net_signed, values_fill = 0) %>%
  column_to_rownames("end")

for (mod in modality_order) {
  if (!mod %in% colnames(signed_mat)) signed_mat[[mod]] <- 0
}
signed_mat <- as.matrix(signed_mat[, intersect(modality_order, colnames(signed_mat))])

# Apply Panel B row order
cluster_rows_flag <- !use_ref_order
if (use_ref_order) {
  shared_rows <- intersect(row_order_ref, rownames(signed_mat))
  extra_rows  <- setdiff(rownames(signed_mat), row_order_ref)
  signed_mat  <- signed_mat[c(shared_rows, extra_rows), ]
}

# Row-normalize: each target neuron scaled to [-1, 1]
signed_mat_norm <- row_normalize(signed_mat)

breaks_signed <- seq(-1, 1, length.out = 101)

aspf_ends <- unique(paths.aspf$end)
aspg_ends <- unique(paths.aspg$end)
row_annot <- data.frame(
  `aSP-f` = ifelse(rownames(signed_mat_norm) %in% aspf_ends, "yes", "no"),
  `aSP-g` = ifelse(rownames(signed_mat_norm) %in% aspg_ends, "yes", "no"),
  row.names = rownames(signed_mat_norm), check.names = FALSE
)
col_annot <- data.frame(
  modality_group = modality_group[colnames(signed_mat_norm)],
  row.names = colnames(signed_mat_norm)
)
ann_colors <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  modality_group = modality_group_colors
)

draw_heatmap_fixed_annot(
  signed_mat_norm, row_annot, col_annot, ann_colors,
  cluster_rows = TRUE, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_signed,
  border_color = NA, fontsize_row = 6, fontsize_col = 10,
  main = "Signed strength (row-normalized, strongest exc - inh path)\nRed = excitatory, Blue = inhibitory",
  angle_col = 45
)
```

```{r save-signed-heatmap}
save_heatmap_both_orientations(
  mat = signed_mat_norm, row_annot = row_annot, col_annot = col_annot,
  ann_colors = ann_colors, name = "panel_J_signed_strength_aspf_aspg",
  width = 8, height = 10,
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_signed,
  border_color = NA, fontsize_row = 6, fontsize_col = 10,
  main = "Signed strength (row-normalized): strongest single path",
  angle_col = 45
)
```

## 13. Per-target valence profile (absolute strongest path, row-normalized)

```{r per-target-valence, fig.width=12, fig.height=10}
valence_profile <- absolute_strongest %>%
  mutate(signed_str = ifelse(valence == "excitatory", strength, -strength)) %>%
  select(end, modality, signed_str) %>%
  pivot_wider(names_from = modality, values_from = signed_str, values_fill = 0) %>%
  column_to_rownames("end")

for (mod in modality_order) {
  if (!mod %in% colnames(valence_profile)) valence_profile[[mod]] <- 0
}
valence_profile <- as.matrix(valence_profile[, intersect(modality_order, colnames(valence_profile))])

# Apply Panel B row order
if (use_ref_order) {
  shared_vp <- intersect(row_order_ref, rownames(valence_profile))
  extra_vp  <- setdiff(rownames(valence_profile), row_order_ref)
  valence_profile <- valence_profile[c(shared_vp, extra_vp), ]
}

# Row-normalize
valence_profile_norm <- row_normalize(valence_profile)
breaks_vp <- seq(-1, 1, length.out = 101)

row_annot_vp <- data.frame(
  `aSP-f` = ifelse(rownames(valence_profile_norm) %in% aspf_ends, "yes", "no"),
  `aSP-g` = ifelse(rownames(valence_profile_norm) %in% aspg_ends, "yes", "no"),
  row.names = rownames(valence_profile_norm), check.names = FALSE
)
col_annot_vp <- data.frame(
  modality_group = modality_group[colnames(valence_profile_norm)],
  row.names = colnames(valence_profile_norm)
)

draw_heatmap_fixed_annot(
  valence_profile_norm, row_annot_vp, col_annot_vp, ann_colors,
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_vp,
  border_color = NA, fontsize_row = 6, fontsize_col = 10,
  main = "Per-target valence profile (row-normalized): absolute strongest path\nRed = excitatory, Blue = inhibitory",
  angle_col = 45
)
```

```{r save-valence-profile}
save_heatmap_both_orientations(
  mat = valence_profile_norm, row_annot = row_annot_vp, col_annot = col_annot_vp,
  ann_colors = ann_colors, name = "panel_J_valence_profile_aspf_aspg",
  width = 8, height = 10,
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_vp,
  border_color = NA, fontsize_row = 6, fontsize_col = 10,
  main = "Valence profile (row-normalized): strongest single path",
  angle_col = 45
)
```

---

# Network diagrams

## 14. Network diagram: modality -> target (igraph)

```{r network-diagram, fig.width=14, fig.height=14}
edges <- absolute_strongest %>%
  select(modality, end, strength, valence, valence_pattern) %>%
  rename(from = modality, to = end)

g_net <- graph_from_data_frame(edges, directed = TRUE)

V(g_net)$type <- ifelse(V(g_net)$name %in% modality_order, "modality", "target")
V(g_net)$color <- ifelse(V(g_net)$type == "modality",
                          modality_colors[V(g_net)$name], "grey85")
V(g_net)$size <- ifelse(V(g_net)$type == "modality", 12, 5)
V(g_net)$label.cex <- ifelse(V(g_net)$type == "modality", 1.0, 0.5)
V(g_net)$frame.color <- ifelse(V(g_net)$type == "modality", "grey30", "grey60")

E(g_net)$color <- ifelse(E(g_net)$valence == "excitatory",
                          adjustcolor("#B2182B", alpha.f = 0.4),
                          adjustcolor("#2166AC", alpha.f = 0.4))
E(g_net)$width <- scales::rescale(E(g_net)$strength, to = c(0.3, 3))
E(g_net)$arrow.size <- 0.3

mod_nodes <- which(V(g_net)$type == "modality")
tar_nodes <- which(V(g_net)$type == "target")

layout_mat <- matrix(0, nrow = vcount(g_net), ncol = 2)
layout_mat[mod_nodes, 1] <- seq_along(mod_nodes)
layout_mat[mod_nodes, 2] <- 1
layout_mat[tar_nodes, 1] <- seq(0.5, length(mod_nodes) + 0.5, length.out = length(tar_nodes))
layout_mat[tar_nodes, 2] <- 0

plot(g_net, layout = layout_mat,
     vertex.label = V(g_net)$name, edge.curved = 0.1,
     main = "Modality -> target: strongest single path\nRed = excitatory, Blue = inhibitory")
legend("bottomright", legend = c("Excitatory", "Inhibitory"),
       col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n", cex = 0.9)
```

```{r save-network}
for (ext in c("png", "pdf")) {
  f <- file.path(get(paste0("plot_dir_", ext)),
                 paste0("panel_J_network_modality_target.", ext))
  if (ext == "png") png(f, width = 14, height = 14, units = "in", res = 300)
  else              pdf(f, width = 14, height = 14)
  plot(g_net, layout = layout_mat,
       vertex.label = V(g_net)$name, edge.curved = 0.1,
       main = "Modality -> target: strongest single path\nRed = excitatory, Blue = inhibitory")
  legend("bottomright", legend = c("Excitatory", "Inhibitory"),
         col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n", cex = 0.9)
  dev.off()
}
```

## 15. Simplified network (top 50% connections)

```{r simplified-network, fig.width=14, fig.height=10}
net_summary <- absolute_strongest %>%
  group_by(modality, end) %>%
  summarise(total_strength = sum(strength),
            dominant_valence = valence[which.max(strength)], .groups = "drop")

strength_threshold <- quantile(net_summary$total_strength, 0.5)
net_filtered <- net_summary %>% filter(total_strength >= strength_threshold)

g_simple <- graph_from_data_frame(
  net_filtered %>% select(from = modality, to = end, everything()), directed = TRUE
)

V(g_simple)$type <- ifelse(V(g_simple)$name %in% modality_order, "modality", "target")
V(g_simple)$color <- ifelse(V(g_simple)$type == "modality",
                             modality_colors[V(g_simple)$name], "grey85")
V(g_simple)$size <- ifelse(V(g_simple)$type == "modality", 15, 8)
V(g_simple)$label.cex <- ifelse(V(g_simple)$type == "modality", 1.1, 0.65)
V(g_simple)$frame.color <- ifelse(V(g_simple)$type == "modality", "grey30", "grey60")

E(g_simple)$color <- ifelse(E(g_simple)$dominant_valence == "excitatory",
                             adjustcolor("#B2182B", alpha.f = 0.5),
                             adjustcolor("#2166AC", alpha.f = 0.5))
E(g_simple)$width <- scales::rescale(E(g_simple)$total_strength, to = c(0.5, 4))
E(g_simple)$arrow.size <- 0.4

mod_n2 <- which(V(g_simple)$type == "modality")
tar_n2 <- which(V(g_simple)$type == "target")
layout2 <- matrix(0, nrow = vcount(g_simple), ncol = 2)
layout2[mod_n2, 1] <- seq_along(mod_n2)
layout2[mod_n2, 2] <- 1
layout2[tar_n2, 1] <- seq(0.5, length(mod_n2) + 0.5, length.out = length(tar_n2))
layout2[tar_n2, 2] <- 0

plot(g_simple, layout = layout2,
     vertex.label = V(g_simple)$name, edge.curved = 0.15,
     main = "Top 50% strongest modality-target connections\nRed = excitatory, Blue = inhibitory")
legend("bottomright", legend = c("Excitatory", "Inhibitory"),
       col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n")
```

```{r save-simplified-network}
for (ext in c("png", "pdf")) {
  f <- file.path(get(paste0("plot_dir_", ext)),
                 paste0("panel_J_network_simplified.", ext))
  if (ext == "png") png(f, width = 14, height = 10, units = "in", res = 300)
  else              pdf(f, width = 14, height = 10)
  plot(g_simple, layout = layout2,
       vertex.label = V(g_simple)$name, edge.curved = 0.15,
       main = "Top 50% strongest modality-target connections\nRed = excitatory, Blue = inhibitory")
  legend("bottomright", legend = c("Excitatory", "Inhibitory"),
         col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n")
  dev.off()
}
```

## 16. ggraph network (publication-quality)

```{r ggraph-network, fig.width=14, fig.height=10}
library(ggraph)
library(tidygraph)

node_df <- data.frame(
  name = c(modality_order, setdiff(unique(absolute_strongest$end), modality_order)),
  stringsAsFactors = FALSE
) %>%
  mutate(
    node_type = ifelse(name %in% modality_order, "modality", "target"),
    node_color = ifelse(node_type == "modality", unname(modality_colors[name]), "grey75"),
    type = (node_type == "modality")
  )

edge_df <- absolute_strongest %>%
  select(from = modality, to = end, strength, valence, valence_pattern)

tg <- tbl_graph(nodes = node_df, edges = edge_df, directed = TRUE)

p_net <- ggraph(tg, layout = "bipartite") +
  geom_edge_fan(
    aes(color = valence, width = strength, alpha = strength),
    arrow = grid::arrow(length = unit(2, "mm"), type = "closed"),
    end_cap = circle(3, "mm")
  ) +
  scale_edge_color_manual(values = c("excitatory" = "#B2182B", "inhibitory" = "#2166AC")) +
  scale_edge_width(range = c(0.2, 2), guide = "none") +
  scale_edge_alpha(range = c(0.15, 0.6), guide = "none") +
  geom_node_point(aes(color = node_color, size = node_type)) +
  scale_color_identity() +
  scale_size_manual(values = c("modality" = 8, "target" = 3), guide = "none") +
  geom_node_text(aes(label = name, y = y + ifelse(node_type == "modality", 0.08, -0.08)),
                 size = 3, fontface = "bold") +
  theme_void() +
  labs(title = "Modality -> 3rd-order neuron: strongest path per pair",
       subtitle = "Red = excitatory, Blue = inhibitory") +
  theme(plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 11))

p_net

save_fig6_plot(p_net, "panel_J_ggraph_network_aspf_aspg", width = 14, height = 10)
```

---

## 17. Summary statistics

```{r summary-table}
summary_tab <- absolute_strongest %>%
  group_by(modality) %>%
  summarise(
    n_targets = n_distinct(end),
    n_excitatory = sum(valence == "excitatory"),
    n_inhibitory = sum(valence == "inhibitory"),
    pct_excitatory = round(100 * mean(valence == "excitatory"), 1),
    mean_strength = mean(strength),
    n_unique_patterns = n_distinct(valence_pattern),
    top_pattern = names(sort(table(valence_pattern), decreasing = TRUE))[1],
    .groups = "drop"
  )

cat("Summary per modality:\n")
print(as.data.frame(summary_tab), row.names = FALSE)

cat(sprintf("\nOverall excitatory fraction: %.1f%%\n",
            100 * mean(absolute_strongest$valence == "excitatory")))
cat(sprintf("Most common pattern: %s (%d occurrences)\n",
            pattern_counts$valence_pattern[1], pattern_counts$count[1]))
cat(sprintf("\nValence conservation: %.0f%% all excitatory, %.0f%% all inhibitory, %.0f%% mixed\n",
            100 * mean(conservation$category == "all excitatory"),
            100 * mean(conservation$category == "all inhibitory"),
            100 * mean(conservation$category == "mixed")))
```
