---
title: "Figure 6B variant: Modality Composition Heatmap -- vAB3 & PPN1 targets (top 50 paths)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel B (vAB3/PPN1 variant): Modality Composition Heatmap — Clustered heatmap of column-normalized modality input strength to vAB3 and PPN1 downstream targets, using the top 50 strongest paths per modality-target pair. Row annotations mark target set membership (vAB3, PPN1); column annotations mark modality group."
panel_legend <- "(B, vAB3/PPN1 variant) Column-normalized heatmap of modality input strength to vAB3 and PPN1 downstream target neurons. Only the 50 strongest paths per modality-target pair are retained to match the aSP-f/aSP-g analysis depth. Rows are clustered by Ward's method; annotations indicate target set (vAB3 vs. PPN1) and modality group."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_B_vab3_ppn1_description", "txt"), "panel_B_vab3_ppn1_description.txt"))
```

## Overview

This is a Panel B variant focusing on the **vAB3 and PPN1 path data**.
The original evaluation notebook computed 1000 strongest paths per modality
to vAB3 and PPN1. Here we keep only the **top 50 per (modality, end) pair**
to match the aspf/aspg analysis, then build the same clustered heatmap.

## 1. Inspect available path data

```{r inspect-data}
cat("=== vAB3 paths ===\n")
cat(sprintf("  Total rows: %d\n", nrow(paths.vab3)))
cat(sprintf("  Unique end types: %s\n", paste(unique(paths.vab3$end), collapse = ", ")))
cat(sprintf("  Modalities: %s\n", paste(unique(paths.vab3$modality), collapse = ", ")))

cat("\n=== PPN1 paths ===\n")
cat(sprintf("  Total rows: %d\n", nrow(paths.ppn1)))
cat(sprintf("  Unique end types: %s\n", paste(unique(paths.ppn1$end), collapse = ", ")))
cat(sprintf("  Modalities: %s\n", paste(unique(paths.ppn1$modality), collapse = ", ")))
```

## 2. Filter to top 50 paths per (modality, end) combination

```{r filter-top50}
# For each (end, modality) pair, keep only the 50 strongest paths
filter_top_n <- function(df, n = 50) {
  df %>%
    group_by(end, modality) %>%
    slice_max(order_by = strength, n = n, with_ties = FALSE) %>%
    ungroup()
}

paths.vab3.top50 <- filter_top_n(paths.vab3, 50)
paths.ppn1.top50 <- filter_top_n(paths.ppn1, 50)

# Combine
paths.vab3_ppn1.top50 <- bind_rows(paths.vab3.top50, paths.ppn1.top50)

cat(sprintf("vAB3 top-50: %d paths, %d unique ends\n",
            nrow(paths.vab3.top50), length(unique(paths.vab3.top50$end))))
cat(sprintf("PPN1 top-50: %d paths, %d unique ends\n",
            nrow(paths.ppn1.top50), length(unique(paths.ppn1.top50$end))))
cat(sprintf("Combined: %d paths, %d unique ends\n",
            nrow(paths.vab3_ppn1.top50), length(unique(paths.vab3_ppn1.top50$end))))
```

## 3. Compute strength matrix

```{r strength-matrix}
mat_vp <- compute_strength_matrix(paths.vab3_ppn1.top50, normalize = FALSE)

# Column-normalize
col_maxes <- apply(mat_vp, 2, max, na.rm = TRUE)
col_maxes[col_maxes == 0] <- 1
mat_vp_norm <- sweep(mat_vp, 2, col_maxes, "/")

cat(sprintf("Strength matrix: %d target neurons x %d modalities\n",
            nrow(mat_vp_norm), ncol(mat_vp_norm)))
```

## 4. Build row annotation (vAB3 vs PPN1 target set)

```{r row-annotation}
# Multiple color bands: one column per target set (yes/no)
all_ends_vp <- rownames(mat_vp_norm)

vab3_ends <- unique(paths.vab3.top50$end)
ppn1_ends <- unique(paths.ppn1.top50$end)

row_annot <- data.frame(
  `vAB3` = ifelse(all_ends_vp %in% vab3_ends, "yes", "no"),
  `PPN1` = ifelse(all_ends_vp %in% ppn1_ends, "yes", "no"),
  row.names = all_ends_vp,
  check.names = FALSE
)

# Column annotation
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_vp_norm)],
  row.names      = colnames(mat_vp_norm)
)

ann_colors <- list(
  `vAB3`         = c("yes" = "#4DAF4A", "no" = "white"),
  `PPN1`         = c("yes" = "#FF7F00", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 5. Draw heatmap

```{r heatmap, fig.width=8, fig.height=10}
hm <- pheatmap(
  mat_vp_norm,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  annotation_row           = row_annot,
  annotation_col           = col_annot,
  annotation_colors        = ann_colors,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to vAB3 & PPN1 targets (top 50 paths)",
  angle_col                = 45,
  silent                   = TRUE
)

hm
```

## 6. Save

```{r save}
save_heatmap_both_orientations(
  mat        = mat_vp_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_vab3_ppn1_top50",
  width      = 8,
  height     = 10,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to vAB3 & PPN1 targets (top 50 paths)",
  angle_col                = 45
)
```

## 7. Summary statistics

```{r summary}
# Per target set: how many end neurons, which modalities strongest
cat("=== Per target set summary ===\n")
paths.vab3_ppn1.top50 %>%
  group_by(target_set, modality) %>%
  summarise(
    n_paths = n(),
    total_strength = sum(strength, na.rm = TRUE),
    mean_strength = mean(strength, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(target_set, desc(total_strength)) %>%
  print(n = 20)

cat(sprintf("\n=== Unique end types per target set ===\n"))
cat(sprintf("vAB3: %d unique end types\n",
            length(unique(paths.vab3.top50$end))))
cat(sprintf("PPN1: %d unique end types\n",
            length(unique(paths.ppn1.top50$end))))

# Top end neurons by total strength across modalities
cat("\n=== Top 20 target neurons by total path strength ===\n")
paths.vab3_ppn1.top50 %>%
  group_by(end, target_set) %>%
  summarise(total_strength = sum(strength, na.rm = TRUE),
            n_modalities = n_distinct(modality),
            .groups = "drop") %>%
  arrange(desc(total_strength)) %>%
  head(20) %>%
  print()
```
