---
title: "Figure 6B: Modality Composition Heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel B: Modality Composition Heatmap — Clustered heatmap of column-normalized modality input strength to 3rd-order courtship neurons. Rows are target neuron types (aSP-f, aSP-g); columns are the 7 sensory modalities. Row annotations mark target set membership; column annotations mark modality group (olfactory, auditory, visual, contact)."
panel_legend <- "(B) Column-normalized heatmap of modality input strength to third-order target neurons. Each column (modality) is scaled to its maximum, enabling comparison of relative target distributions across modalities. Rows are clustered by Ward's method; annotations indicate target set (aSP-f vs. aSP-g) and modality group."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_B_description", "txt"), "panel_B_description.txt"))
```

## Overview

Panel B is a clustered heatmap of modality input strength to third-order
courtship neurons. Columns are the 7 sensory modalities; rows are target
neuron types. The matrix is column-normalized so that each modality's maximum
is 1, making the relative distribution across targets comparable between
modalities of different absolute strengths.

Row annotations indicate the target set (aSP-f vs aSP-g).
Column annotations indicate the modality group (olfactory / auditory / visual / contact).

## 1. Prepare strength matrix (aSP-f + aSP-g targets)

```{r strength-matrix-aspf-aspg}
# ---------------------------------------------------------------------------
# Column-normalize the pre-computed strength matrix
# Each column (modality) is scaled so its maximum value = 1
# ---------------------------------------------------------------------------
mat <- strength.aspf_aspg

# Column-normalize: divide each column by its max
col_maxes <- apply(mat, 2, max, na.rm = TRUE)
col_maxes[col_maxes == 0] <- 1
mat_norm <- sweep(mat, 2, col_maxes, "/")

cat(sprintf("Matrix dimensions: %d target neurons x %d modalities\n",
            nrow(mat_norm), ncol(mat_norm)))
```

## 2. Build row annotation (target set membership)

Each target neuron may appear in aSP-f paths, aSP-g paths, or both. We assign
primary membership based on which target set contributes greater total strength.

```{r row-annotation}
# ---------------------------------------------------------------------------
# Determine target_set membership for each target neuron
# ---------------------------------------------------------------------------
# Separate binary bands per target set; pheatmap annotation widths will be
# adjusted after creation so the total annotation width stays fixed.
all_ends <- rownames(mat_norm)
aspf_ends <- unique(paths.aspf$end)
aspg_ends <- unique(paths.aspg$end)

row_annot <- data.frame(
  `aSP-f` = ifelse(all_ends %in% aspf_ends, "yes", "no"),
  `aSP-g` = ifelse(all_ends %in% aspg_ends, "yes", "no"),
  row.names = all_ends,
  check.names = FALSE
)
```

## 3. Build column annotation (modality group)

```{r col-annotation}
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_norm)],
  row.names      = colnames(mat_norm)
)
```

## 4. Define annotation colours and helper

```{r annotation-colors}
ann_colors <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  modality_group = modality_group_colors
)
# draw_heatmap_fixed_annot() and save_heatmap_fixed_annot() are in setup.R
```

## 5. Draw heatmap (aSP-f + aSP-g targets)

```{r heatmap-aspf-aspg, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Primary heatmap: modality strength to aSP-f / aSP-g targets
# ---------------------------------------------------------------------------
hm <- draw_heatmap_fixed_annot(
  mat_norm, row_annot, col_annot, ann_colors,
  total_annot_cm           = 1.5,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons",
  angle_col                = 45
)
```

## 6. Save primary heatmap

```{r save-primary}
save_heatmap_both_orientations(
  mat        = mat_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_aspf_aspg",
  width      = 8,
  height     = 10,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons",
  angle_col                = 45
)
```

## 7. Extended version: include vAB3 and PPN1 downstream targets

Recompute the strength matrix using `paths.all` to include targets reached
via vAB3 and PPN1 paths, providing a broader view of multimodal convergence.

```{r strength-matrix-all}
# ---------------------------------------------------------------------------
# Strength matrix from all paths (aspf + aspg + vAB3 + PPN1)
# ---------------------------------------------------------------------------
mat_all <- compute_strength_matrix(paths.all, normalize = FALSE)

# Column-normalize
col_maxes_all <- apply(mat_all, 2, max, na.rm = TRUE)
col_maxes_all[col_maxes_all == 0] <- 1
mat_all_norm <- sweep(mat_all, 2, col_maxes_all, "/")

cat(sprintf("Extended matrix: %d target neurons x %d modalities\n",
            nrow(mat_all_norm), ncol(mat_all_norm)))
```

## 8. Row annotation for extended matrix

```{r row-annotation-all}
# ---------------------------------------------------------------------------
# Target set membership for all targets
# ---------------------------------------------------------------------------
# 4 binary bands (one per target set), fixed total width via helper
all_ends_ext <- rownames(mat_all_norm)

row_annot_all <- data.frame(
  `aSP-f` = ifelse(all_ends_ext %in% unique(paths.aspf$end), "yes", "no"),
  `aSP-g` = ifelse(all_ends_ext %in% unique(paths.aspg$end), "yes", "no"),
  `vAB3`  = ifelse(all_ends_ext %in% unique(paths.vab3$end), "yes", "no"),
  `PPN1`  = ifelse(all_ends_ext %in% unique(paths.ppn1$end), "yes", "no"),
  row.names = all_ends_ext,
  check.names = FALSE
)

col_annot_all <- data.frame(
  modality_group = modality_group[colnames(mat_all_norm)],
  row.names      = colnames(mat_all_norm)
)

ann_colors_all <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  `vAB3`         = c("yes" = "#4DAF4A", "no" = "white"),
  `PPN1`         = c("yes" = "#FF7F00", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 9. Draw extended heatmap

```{r heatmap-all, fig.width=8, fig.height=14}
# ---------------------------------------------------------------------------
# Extended heatmap: all target sets
# ---------------------------------------------------------------------------
# 4 bands share 1.5 cm total = 0.375 cm each
hm_all <- draw_heatmap_fixed_annot(
  mat_all_norm, row_annot_all, col_annot_all, ann_colors_all,
  total_annot_cm           = 1.5,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (all target sets)",
  angle_col                = 45
)
```

## 10. Save extended heatmap

```{r save-extended}
save_heatmap_both_orientations(
  mat        = mat_all_norm,
  row_annot  = row_annot_all,
  col_annot  = col_annot_all,
  ann_colors = ann_colors_all,
  name       = "panel_B_heatmap_all_targets",
  width      = 8,
  height     = 14,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (all target sets)",
  angle_col                = 45
)
```

## 11. Export row clustering order for use in Panel C

Save the dendrogram row order so Panel C can use the same neuron ordering
for consistent visual comparison.

```{r export-row-order}
# ---------------------------------------------------------------------------
# Extract row order from the primary heatmap's dendrogram
# ---------------------------------------------------------------------------
row_order_aspf_aspg <- rownames(mat_norm)[hm$tree_row$order]

# Save to an RDS file for Panel C to load
saveRDS(
  row_order_aspf_aspg,
  file = file.path(fig6_dir, "panel_B_row_order.rds")
)

cat("Row order saved. First 10 neurons:\n")
head(row_order_aspf_aspg, 10)

# ---------------------------------------------------------------------------
# Also save the all-targets row order for Panel C extended sections
# ---------------------------------------------------------------------------
row_order_all <- rownames(mat_all_norm)[hm_all$tree_row$order]

saveRDS(
  row_order_all,
  file = file.path(fig6_dir, "panel_B_all_row_order.rds")
)

cat(sprintf("\nAll-targets row order saved: %d neurons\n", length(row_order_all)))
```
