---
title: "Figure 6H: Spotlight Examples — Individual Integration Nodes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel H: Spotlight Integration Nodes — Spotlights 3-4 top multimodal integration neurons with individual radar/spider plots showing their modality input profiles. Excitatory and inhibitory inputs are overlaid as separate polygons. Stacked bar charts show absolute and proportional modality breakdown per neuron."
panel_legend <- "(H) Spotlight profiles of top multimodal integration nodes. Radar plots show the normalized modality input profile of individual hub neurons, with excitatory (solid) and inhibitory (dashed) contributions overlaid. Stacked bar charts display absolute and proportional modality composition."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_H_description", "txt"), "panel_H_description.txt"))
```

## Overview

Panel H spotlights 3-4 specific neurons that serve as top multimodal
integration nodes. Each neuron receives input from multiple sensory
modalities, and we visualize its input profile using radar/spider plots
and stacked bar charts broken down by excitatory vs. inhibitory valence.

## Step 1: Compute entropy and rank neurons

Shannon entropy across the 7 modality channels quantifies how evenly a
neuron integrates across modalities. Higher entropy = more multimodal.

```{r compute-entropy}
# ---------------------------------------------------------------------------
# Compute Shannon entropy for each neuron (row) in the combined strength matrix
# ---------------------------------------------------------------------------
entropy_vals <- apply(strength.aspf_aspg, 1, shannon_entropy)

entropy_df <- data.frame(
  neuron    = names(entropy_vals),
  entropy   = entropy_vals,
  n_modalities = apply(strength.aspf_aspg > 0, 1, sum),
  total_strength = rowSums(strength.aspf_aspg),
  stringsAsFactors = FALSE
) %>%
  arrange(desc(entropy))

# Show the top 20 most multimodal neurons
head(entropy_df, 20)
```

## Step 2: Identify spotlight neurons

Select the top-entropy neurons. Also look for known neurons from the
literature (AL-AST1, LH008/LH002-containing aSP-f subtypes, aSP-g
subtypes) to provide biologically meaningful examples.

```{r select-spotlight}
# ---------------------------------------------------------------------------
# Inspect what neuron names exist in the strength matrix
# ---------------------------------------------------------------------------
all_neurons <- rownames(strength.aspf_aspg)
message("Total neurons in combined matrix: ", length(all_neurons))

# Check for known neuron types from the literature
known_patterns <- c("AL-AST1", "LH008", "LH002", "aSP-f", "aSP-g",
                     "SMP", "SIP", "SLP", "LH")
known_hits <- lapply(known_patterns, function(pat) {
  grep(pat, all_neurons, value = TRUE, ignore.case = TRUE)
})
names(known_hits) <- known_patterns

# Print matches for each pattern
for (pat in known_patterns) {
  if (length(known_hits[[pat]]) > 0) {
    message(sprintf("  Pattern '%s': %s", pat,
                    paste(known_hits[[pat]], collapse = ", ")))
  }
}

# ---------------------------------------------------------------------------
# Build a candidate list: top-entropy neurons + known types if present
# ---------------------------------------------------------------------------
# Start with top 4 by entropy
top_entropy <- entropy_df %>%
  filter(n_modalities >= 3) %>%     # require input from at least 3 modalities
  head(6) %>%
  pull(neuron)

# Add known neurons if they are present and have reasonable entropy
known_candidates <- c(
  grep("AL-AST1", all_neurons, value = TRUE, ignore.case = TRUE),
  grep("LH008", all_neurons, value = TRUE, ignore.case = TRUE),
  grep("LH002", all_neurons, value = TRUE, ignore.case = TRUE)
)
known_candidates <- intersect(known_candidates, entropy_df$neuron)

# Combine and deduplicate, then take 3-4
spotlight_candidates <- unique(c(top_entropy, known_candidates))

# Final selection: pick top 4 by entropy from the candidate pool
spotlight_neurons <- entropy_df %>%
  filter(neuron %in% spotlight_candidates) %>%
  arrange(desc(entropy)) %>%
  head(4) %>%
  pull(neuron)

message("Spotlight neurons: ", paste(spotlight_neurons, collapse = ", "))

# Look up synonyms/descriptions from mba for annotation
spotlight_info <- mba %>%
  filter(type %in% spotlight_neurons) %>%
  select(type, synonyms) %>%
  distinct(type, .keep_all = TRUE)

spotlight_info
```

## Step 3: Prepare per-neuron data for plotting

Extract strength values (total, excitatory, inhibitory) for each
spotlight neuron across all 7 modalities.

```{r prepare-neuron-data}
# ---------------------------------------------------------------------------
# Build a long-format dataframe for spotlight neurons
# ---------------------------------------------------------------------------

# Total strength
spotlight_total <- as.data.frame(strength.aspf_aspg[spotlight_neurons, , drop = FALSE]) %>%
  rownames_to_column("neuron") %>%
  pivot_longer(cols = all_of(modality_order), names_to = "modality",
               values_to = "strength") %>%
  mutate(valence = "total")

# Excitatory strength (handle neurons that may be absent from exc/inh matrices)
exc_neurons <- intersect(spotlight_neurons, rownames(strength.aspf_aspg.exc))
inh_neurons <- intersect(spotlight_neurons, rownames(strength.aspf_aspg.inh))

spotlight_exc <- if (length(exc_neurons) > 0) {
  as.data.frame(strength.aspf_aspg.exc[exc_neurons, , drop = FALSE]) %>%
    rownames_to_column("neuron") %>%
    pivot_longer(cols = all_of(modality_order), names_to = "modality",
                 values_to = "strength") %>%
    mutate(valence = "excitatory")
} else {
  data.frame(neuron = character(), modality = character(),
             strength = numeric(), valence = character())
}

spotlight_inh <- if (length(inh_neurons) > 0) {
  as.data.frame(strength.aspf_aspg.inh[inh_neurons, , drop = FALSE]) %>%
    rownames_to_column("neuron") %>%
    pivot_longer(cols = all_of(modality_order), names_to = "modality",
                 values_to = "strength") %>%
    mutate(valence = "inhibitory")
} else {
  data.frame(neuron = character(), modality = character(),
             strength = numeric(), valence = character())
}

# Combine all
spotlight_long <- bind_rows(spotlight_total, spotlight_exc, spotlight_inh) %>%
  mutate(modality = factor(modality, levels = modality_order))
```

## Step 4: Individual radar plots (total strength)

One radar per spotlight neuron showing total input strength across the 7
modalities.

```{r radar-individual, fig.width=14, fig.height=4}
# ---------------------------------------------------------------------------
# Build a custom radar plot for each spotlight neuron
# ---------------------------------------------------------------------------
make_neuron_radar <- function(neuron_name, strength_mat, fill_color = "steelblue",
                               fill_alpha = 0.3, line_color = "steelblue") {
  vals <- strength_mat[neuron_name, modality_order]
  # Normalize to 0-1 for comparable radar shape
  max_val <- max(vals, na.rm = TRUE)
  if (max_val == 0) max_val <- 1
  vals_norm <- vals / max_val

  radar_df <- data.frame(
    modality = factor(modality_order, levels = modality_order),
    value    = as.numeric(vals_norm),
    raw      = as.numeric(vals)
  )

  # Close the polygon by appending the first point
  radar_closed <- bind_rows(radar_df, radar_df[1, ])

  # Look up synonym if available
  syn <- mba %>%
    filter(type == neuron_name) %>%
    pull(synonyms) %>%
    unique() %>%
    na.omit()
  subtitle_text <- if (length(syn) > 0 && nchar(syn[1]) > 0) {
    paste0("(", substr(syn[1], 1, 60), ")")
  } else {
    ""
  }

  # Entropy for this neuron
  ent <- entropy_df %>% filter(neuron == neuron_name) %>% pull(entropy)
  ent_label <- sprintf("H = %.2f", ent)

  ggplot(radar_df, aes(x = modality, y = value, group = 1)) +
    # Grid circles at 25%, 50%, 75%, 100%
    annotate("path",
             x = rep(1:7, each = 100) + rep(seq(0, 1, length.out = 100) - 0.5, 7),
             y = rep(0.25, 700), alpha = 0, color = NA) +
    geom_polygon(fill = alpha(fill_color, fill_alpha), color = line_color,
                 linewidth = 0.8) +
    geom_point(aes(color = modality), size = 3, show.legend = FALSE) +
    scale_color_manual(values = modality_colors) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05), breaks = c(0.25, 0.5, 0.75, 1.0),
                       labels = c("25%", "50%", "75%", "100%")) +
    labs(
      title    = neuron_name,
      subtitle = paste(subtitle_text, ent_label)
    ) +
    theme_minimal(base_size = 11) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_text(size = 7, color = "grey50"),
      axis.text.x       = element_text(size = 9, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 12),
      plot.subtitle      = element_text(hjust = 0.5, size = 8, color = "grey40")
    )
}

# Generate individual radar plots
radar_plots <- lapply(spotlight_neurons, function(n) {
  make_neuron_radar(n, strength.aspf_aspg)
})
names(radar_plots) <- spotlight_neurons

# Arrange in a single row
radar_combined <- plot_grid(
  plotlist = radar_plots,
  nrow = 1,
  labels = LETTERS[1:length(radar_plots)],
  label_size = 12
)
radar_combined
```

```{r save-radar-individual}
save_fig6_plot(radar_combined, "panel_H_radar_individual_aspf_aspg",
               width = 3.5 * length(spotlight_neurons), height = 4)
```

## Step 5: Excitatory vs. inhibitory overlaid radars

For each spotlight neuron, overlay excitatory (solid) and inhibitory
(dashed) radar polygons on the same chart.

```{r radar-valence, fig.width=14, fig.height=4.5}
# ---------------------------------------------------------------------------
# Dual-valence radar: excitatory (solid) + inhibitory (dashed) overlaid
# ---------------------------------------------------------------------------
make_valence_radar <- function(neuron_name) {
  # Get excitatory values
  exc_vals <- if (neuron_name %in% rownames(strength.aspf_aspg.exc)) {
    strength.aspf_aspg.exc[neuron_name, modality_order]
  } else {
    setNames(rep(0, 7), modality_order)
  }

  # Get inhibitory values
  inh_vals <- if (neuron_name %in% rownames(strength.aspf_aspg.inh)) {
    strength.aspf_aspg.inh[neuron_name, modality_order]
  } else {
    setNames(rep(0, 7), modality_order)
  }

  # Normalize both to the same scale (max across both)
  global_max <- max(c(exc_vals, inh_vals), na.rm = TRUE)
  if (global_max == 0) global_max <- 1

  radar_df <- data.frame(
    modality = factor(rep(modality_order, 2), levels = modality_order),
    value    = c(as.numeric(exc_vals) / global_max,
                 as.numeric(inh_vals) / global_max),
    valence  = rep(c("excitatory", "inhibitory"), each = 7)
  )

  ggplot(radar_df, aes(x = modality, y = value, group = valence)) +
    geom_polygon(aes(fill = valence, linetype = valence, color = valence),
                 alpha = 0.2, linewidth = 0.9) +
    geom_point(aes(color = valence), size = 2) +
    scale_fill_manual(values = c("excitatory" = "#2166AC",
                                  "inhibitory" = "#B2182B")) +
    scale_color_manual(values = c("excitatory" = "#2166AC",
                                   "inhibitory" = "#B2182B")) +
    scale_linetype_manual(values = c("excitatory" = "solid",
                                      "inhibitory" = "dashed")) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05)) +
    labs(title = neuron_name) +
    theme_minimal(base_size = 10) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_blank(),
      axis.text.x       = element_text(size = 8, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 11),
      legend.position    = "bottom",
      legend.title       = element_blank(),
      legend.text        = element_text(size = 8)
    )
}

valence_radars <- lapply(spotlight_neurons, make_valence_radar)
names(valence_radars) <- spotlight_neurons

valence_combined <- plot_grid(
  plotlist = valence_radars,
  nrow = 1,
  labels = LETTERS[1:length(valence_radars)],
  label_size = 12
)
valence_combined
```

```{r save-radar-valence}
save_fig6_plot(valence_combined, "panel_H_radar_valence_aspf_aspg",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 6: Input breakdown stacked bar (excitatory/inhibitory by modality)

For each spotlight neuron, a stacked bar chart showing the proportion of
each modality split by valence.

```{r stacked-bar, fig.width=14, fig.height=5}
# ---------------------------------------------------------------------------
# Build stacked bar data for spotlight neurons
# ---------------------------------------------------------------------------
bar_data <- spotlight_long %>%
  filter(valence %in% c("excitatory", "inhibitory"),
         neuron %in% spotlight_neurons) %>%
  # Compute proportion within each neuron
  group_by(neuron) %>%
  mutate(
    total_all = sum(strength, na.rm = TRUE),
    proportion = ifelse(total_all > 0, strength / total_all, 0)
  ) %>%
  ungroup() %>%
  # Add modality group for ordering
  mutate(
    mod_group = modality_group[as.character(modality)],
    modality  = factor(modality, levels = modality_order)
  )

# ---------------------------------------------------------------------------
# Stacked bar: faceted by neuron, x = modality, fill = valence
# ---------------------------------------------------------------------------
p_stacked <- ggplot(bar_data,
                     aes(x = modality, y = strength, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ neuron, scales = "free_y", nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC",
                                "inhibitory" = "#B2182B"),
                     name = "Valence") +
  labs(
    x     = "Modality",
    y     = "Summed path strength",
    title = "Input breakdown by modality and valence"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 10),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )

p_stacked
```

```{r save-stacked-bar}
save_fig6_plot(p_stacked, "panel_H_stacked_bar_aspf_aspg",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 6b: Proportional stacked bar (normalized to 100%)

Same breakdown but normalized so each neuron sums to 100%, making the
modality proportions directly comparable.

```{r stacked-bar-proportional, fig.width=14, fig.height=5}
p_stacked_prop <- ggplot(bar_data,
                          aes(x = modality, y = proportion, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ neuron, nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC",
                                "inhibitory" = "#B2182B"),
                     name = "Valence") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x     = "Modality",
    y     = "Proportion of total input",
    title = "Proportional input breakdown by modality and valence"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 10),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )

p_stacked_prop
```

```{r save-stacked-prop}
save_fig6_plot(p_stacked_prop, "panel_H_stacked_bar_proportional_aspf_aspg",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 7: Neuron annotation summary table

Print a summary table for the spotlight neurons with their entropy,
number of modalities, total strength, and synonyms.

```{r annotation-table}
# ---------------------------------------------------------------------------
# Build annotation table
# ---------------------------------------------------------------------------
annotation_table <- entropy_df %>%
  filter(neuron %in% spotlight_neurons) %>%
  left_join(
    mba %>%
      select(type, synonyms) %>%
      distinct(type, .keep_all = TRUE),
    by = c("neuron" = "type")
  ) %>%
  mutate(
    synonyms = ifelse(is.na(synonyms) | synonyms == "", "—", synonyms),
    # Truncate long synonym strings
    synonyms = ifelse(nchar(synonyms) > 80,
                       paste0(substr(synonyms, 1, 77), "..."),
                       synonyms)
  ) %>%
  arrange(desc(entropy))

knitr::kable(annotation_table, digits = 3,
             caption = "Spotlight neuron summary",
             col.names = c("Neuron type", "Entropy (bits)",
                           "N modalities", "Total strength", "Synonyms"))
```

## Step 8: Combined final panel

Assemble the main panel figure: radars on top, stacked bars on bottom.

```{r combined-panel, fig.width=14, fig.height=9}
panel_H <- plot_grid(
  radar_combined,
  valence_combined,
  p_stacked,
  ncol    = 1,
  rel_heights = c(1, 1, 1),
  labels  = c("i", "ii", "iii"),
  label_size = 14
)
panel_H
```

```{r save-combined}
save_fig6_plot(panel_H, "panel_H_combined_aspf_aspg", width = 14, height = 12)
```
