---
title: "Contact sensory neurons to their downstream partners"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", cache = FALSE)
```

## Setup

`rgl` cannot initialise headless on this machine, so `rgl.useNULL` is set before
anything that might pull it in.

```{r libraries}
suppressMessages({
  library(neuprintr); library(malecns); library(nat)
  options(malecns.dataset = "male-cns:v1.0")   # malecns helpers ignore MCNS_DATASET
  library(coconatfly); library(coconat); library(fafbseg); library(bancr)
  library(dplyr); library(tibble); library(igraph)
  library(pheatmap); library(RColorBrewer); library(viridisLite); library(grid)
})
options(rgl.useNULL = TRUE)

source("/Users/fkampf/Documents/pheromone.paper/analyses/an_investigation/R/_paths.R")
neuprint_login(server = MCNS_SERVER, dataset = MCNS_DATASET)
```

Annotation tables are expensive to fetch, so they are cached in the global
environment for the session.

```{r load_once}
load_once <- function(name, expr, envir = .GlobalEnv) {
  # eval() here evaluates a quote()d expression written literally in this file -
  # it is a lazy-evaluation idiom for "only run this fetch if not already cached",
  # not a parser of external input.
  if (!exists(name, envir = envir, inherits = FALSE)) {
    assign(name, eval(expr), envir = envir)
  }
  get(name, envir = envir)
}

fct <- load_once("fct", quote(banc_edgelist(source = "cave")))
mba <- load_once("mba", quote(mcns_body_annotations()))
fba <- load_once("fba", quote(banc_codex_annotations()))
```

## The sensory pool

Leg and wing gustatory bristle neurons. Wider than the receptor-typed set on
purpose: the generic taste bristles act as a built-in negative control, so if the
pheromone candidates fail to separate from them that is itself a result.

`!is.na(type)` drops the handful of unannotated bodies, which would otherwise
collapse into a single `NA` row downstream.

```{r pool}
contact_pool <- mba |>
  filter(class == "gustatory",
         superclass == "vnc_sensory",
         !is.na(type),
         subclass %in% c("leg bristle", "wing bristle")) |>
  pull(bodyid)

length(contact_pool)
```

## Edge table

`mcns_connection_table(..., 'o')` returns `bodyid` = PRE and `partner` = POST.
Its `type` / `superclass` / `somaSide` columns describe the **partner**, not the
query body, so the pre-side type has to be joined explicitly on `bodyid` alone —
a bare `left_join(mba)` would join on all the shared columns at once and match
almost nothing.

```{r edges}
contact2ans <- mcns_connection_table(contact_pool, "o")
```

The denominator for normalisation is each target's total input, from every
source. Taking it from a connection table rather than `neuprint_get_meta()$post`
keeps numerator and denominator like-for-like — both are edge weights — so
`norm_input` sums to 1 across all presynaptic types, matching the
column-normalised `adj.matrix.rds` convention. This is the slow call.

```{r denominator}
post_in_conn <- load_once("post_in_conn", quote(
  mcns_connection_table(unique(contact2ans$partner), "i") |>
    summarise(post_total_in = sum(weight), .by = bodyid)
))
```

`filter(!post_type %in% pre_type)` removes sensory types from the post axis, so
what remains is genuinely downstream rather than sensory-to-sensory wiring.

```{r sn_an}
sn_an <- contact2ans |>
  rename(post_type = type) |>
  left_join(mba |> select(bodyid, pre_type = type), by = "bodyid") |>
  filter(!post_type %in% pre_type) |>
  left_join(post_in_conn, by = c("partner" = "bodyid")) |>
  mutate(post_total_in = sum(distinct(pick(partner, post_total_in))$post_total_in,
                             na.rm = TRUE),
         .by = post_type) |>
  summarise(weight     = sum(weight),
            norm_input = sum(weight) / first(post_total_in),
            n_pre      = n_distinct(bodyid),
            n_post     = n_distinct(partner),
            .by        = c(pre_type, post_type)) |>
  arrange(desc(weight))

head(sn_an, 20)
```

## Matrix

`MIN_WEIGHT` thresholds on the **group total** per post type, not per row: a post
type qualifies on its summed input from the whole pool, and then keeps its full
profile. Truncating individual edges instead would give the clustering
incomplete vectors to compare.

`xtabs` builds the matrix, fills absent pairs with 0, and drops NA levels, which
is why no `pivot_wider` / `values_fill` / `!is.na()` is needed.

```{r matrix}
MIN_WEIGHT <- 500

m <- sn_an |>
  filter(sum(weight) >= MIN_WEIGHT, .by = post_type) |>
  xtabs(norm_input ~ pre_type + post_type, data = _)

dim(m)
```

`MIN_WEIGHT` only thresholds the post axis, so every pool type gets a row however
little it contributes. The `SNch*` / `SNxxxx` types clear a weight threshold —
`SNch05` really does make ~670 synapses — but spread so thinly that they never
exceed ~0.8% of any target's input, which renders as the same dark floor as a
true zero. Pruning on peak impact rather than volume is what removes them:

```{r prune_rows}
MIN_IMPACT <- 0.01   # a pre type must supply this fraction of some target's input

dropped <- rownames(m)[apply(m, 1, max) < MIN_IMPACT]
m <- m[apply(m, 1, max) >= MIN_IMPACT, , drop = FALSE]

dropped
```

Dropping rows can leave a post type with nothing above the floor, so the columns
are re-checked afterwards.

```{r prune_cols}
m <- m[, colSums(m) > 0, drop = FALSE]
dim(m)
```

## Column families

Post types differing only in their trailing letter (`AN05B023a`–`d`) form one
family. Families with more than one member get a colour; singletons stay grey and
collect into a trailing `other` block.

`fam` and `blk` are recomputed after the reorder — colouring with the pre-reorder
vector would put every colour on the wrong column.

```{r families}
fam   <- sub("_?[a-z]$", "", colnames(m))
multi <- sort(names(which(table(fam) > 1)))
blk   <- factor(ifelse(fam %in% multi, fam, "other"), levels = c(multi, "other"))

m <- m[, order(blk, colnames(m)), drop = FALSE]

fam <- sub("_?[a-z]$", "", colnames(m))
blk <- factor(ifelse(fam %in% multi, fam, "other"), levels = c(multi, "other"))

pal     <- setNames(brewer.pal(max(3, length(multi)), "Set1")[seq_along(multi)], multi)
lab_col <- ifelse(fam %in% multi, pal[fam], "grey35")
gaps    <- head(cumsum(rle(as.character(blk))$lengths), -1)

levels(blk)
```

## Raw synapse matrix

The same edges as `m`, but unnormalised. Built from the full `sn_an` and then
indexed by `m`'s dimnames, so it inherits every filter and the family ordering —
`gaps` and `lab_col` stay valid for it without being recomputed.

```{r matrix_raw}
mw <- sn_an |>
  xtabs(weight ~ pre_type + post_type, data = _)

mw <- mw[rownames(m), colnames(m), drop = FALSE]

range(mw)
```

Raw counts span several orders of magnitude, so a linear viridis scale shows the
few largest edges and floors everything else. `log10(x + 1)` is the readable
version; the untransformed one is kept because it is the honest picture of how
lopsided the weights are.

```{r matrix_raw_log}
mw_log <- log10(mw + 1)
```

## Plotting helper

`pheatmap` has no option for coloured tick labels, and `angle_col` only accepts
0/45/90/270/315, so the `col_names` grob is recoloured and rotated after the
fact. That is why the call uses `silent = TRUE` and draws explicitly.

Note the `tree_col` branch: when columns are clustered, `pheatmap` reorders them,
so `lab_col` has to be permuted into dendrogram order or every colour lands on
the wrong label. `gaps_col` is ignored under clustering, so it is only passed
when clustering is off.

```{r draw_hm}
draw_hm <- function(mat, title, cluster_cols = FALSE, rot = 90) {
  ph <- pheatmap(mat, color = viridis(100),
                 cluster_rows = FALSE, cluster_cols = cluster_cols,
                 gaps_col = if (cluster_cols) NULL else gaps,
                 border_color = NA,
                 fontsize_row = 10, fontsize_col = 9,
                 main = title, silent = TRUE)

  # pheatmap returns tree_col = FALSE (not NULL) when clustering is off
  cols <- if (inherits(ph$tree_col, "hclust")) lab_col[ph$tree_col$order] else lab_col

  g <- ph$gtable
  i <- which(g$layout$name == "col_names")
  g$grobs[[i]]$gp$col <- cols
  g$grobs[[i]]$rot    <- rot
  g$grobs[[i]]$hjust  <- 1
  g$grobs[[i]]$vjust  <- 1
  grid.newpage(); grid.draw(g)
}
```

## Heatmaps

Fraction of each post type's **total** input that comes from each sensory type.
Absolute scale, so weakly-innervated targets stay visibly weak.

```{r hm_raw, fig.width = 20, fig.height = 5}
draw_hm(m, "Sensory input onto downstream types (fraction of each type's total input)")
```

The same matrix renormalised so each column sums to 1 — composition rather than
magnitude, which makes profile *shape* comparable between a target receiving 70%
sensory input and one receiving 5%. Margin 2 because post types are the columns.

```{r hm_composition, fig.width = 20, fig.height = 5}
draw_hm(prop.table(m, 2), "Sensory input composition (each column sums to 1)")
```

Clustered on the post axis, to let the families be judged against the data rather
than against their names. Correlation distance groups targets by profile shape;
Euclidean would group them by how much input they get.

```{r hm_clustered, fig.width = 20, fig.height = 5}
draw_hm(prop.table(m, 2), "Clustered on post type (composition)", cluster_cols = TRUE)
```

The same clustering on the unnormalised matrix. Because `norm_input` is already a
fraction of each target's total input, columns here differ in magnitude as well
as shape — a target drawing 70% of its input from the pool has a longer vector
than one drawing 5%, even if they sample the same sensory types in the same
proportions. Correlation distance is scale-free, so the two clusterings should
largely agree; where they do not, the split is driven by how strongly a target is
innervated rather than by which types innervate it.

```{r hm_clustered_raw, fig.width = 20, fig.height = 5}
draw_hm(m, "Clustered on post type (fraction of total input)", cluster_cols = TRUE)
```

## Raw synapse counts

Unnormalised weights. Dominated by the handful of very strong edges, which is the
point of showing it: normalisation hides that `LgLG1a` makes an order of
magnitude more synapses than most of the pool.

```{r hm_weight, fig.width = 20, fig.height = 5}
draw_hm(mw, "Raw synapse count")
```

Log scale, so the mid-range edges become visible.

```{r hm_weight_log, fig.width = 20, fig.height = 5}
draw_hm(mw_log, "Raw synapse count, log10(x + 1)")
```

Clustered on the log matrix. Unlike the normalised pair, this can genuinely
disagree with the composition clustering — correlation over raw counts is still
scale-free per column, but the zeros and the dynamic range weight the correlation
differently once the values are logged.

```{r hm_weight_clustered, fig.width = 20, fig.height = 5}
draw_hm(mw_log, "Raw synapse count, clustered on post type", cluster_cols = TRUE)
```

## What the families look like

Dominant sensory input per post type, for the multi-member families.

```{r family_table}
sn_an |>
  filter(post_type %in% colnames(m),
         sub("_?[a-z]$", "", post_type) %in% multi) |>
  slice_max(norm_input, n = 1, by = post_type) |>
  arrange(post_type) |>
  select(post_type, pre_type, norm_input, weight, n_pre)
```

Pre types contributing almost nothing survive the filter, because `MIN_WEIGHT`
only thresholds the post axis. They are not empty rows, just far below the colour
scale — worth checking before reading a row as absent.

```{r weak_rows}
tibble(pre_type = rownames(m),
       max_norm = apply(m, 1, max),
       total    = rowSums(m)) |>
  arrange(max_norm)
```
