```{r}
library(malecns)
library(coconatfly)
library(fafbseg)
library(neuprintr)
library(ggplot2)
library(dplyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(stringr)
library(igraph)
library(tidygraph)
library(ggraph)
library(jsonlite)
library(ggplot2)
library(reshape2)
library(Matrix)
library(ggplot2)
library(ggpubr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  fig.width = 8,
  fig.height = 6
)
```


## 1 Load data

### 1.1 Load male boday annotations
```{r}
synapse_threshold=5
mba<-mcns_body_annotations()
mba.static<-mcns_body_annotations()
mba<-mba%>%mutate(type=ifelse(type=='',NA,type))


ppk23.types <- mba%>%filter(receptor_type=='putative_ppk23')%>%pull(type)%>%unique()
ppk25.types <- mba%>%filter(receptor_type=='putative_ppk25')%>%pull(type)%>%unique()
IR52a.types <- mba%>%filter(receptor_type=='putative_IR52a')%>%pull(type)%>%unique()
IR52b.types <- mba%>%filter(receptor_type=='putative_IR52b')%>%pull(type)%>%unique()

mba.dimorphism <- mba %>%
  count(type, dimorphism) %>%
  group_by(type) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(type, dimorphism)
```






### 1.2 Function to load connectivity
```{r}
fetch_connectivity <- function(synapse_threshold = 5) {
  mba <- mcns_body_annotations()%>%
    mutate(type=coalesce(type,flywire_type,manc_type,hemibrain_type))%>%
    mutate(cachero.type = str_extract(synonyms, "(?<=Cachero 2010:)[^;]+") %>%
                            str_trim())
  connectivity <- cf_partners(cf_ids(malecns = mba %>% pull(bodyid)),
                              partners = 'o',
                              threshold = synapse_threshold)
  connectivity <- connectivity %>%
    left_join(mba %>%
                select("type", "bodyid",'fru_dsx','consensus_nt','flywire_type','synonyms','receptor_type','cachero.type','dimorphism') %>%
                mutate(bodyid=as.integer(bodyid))%>%
                rename(pre_type = type,pre_fru_dsx=fru_dsx,pre_nt=consensus_nt,pre_fw_type=flywire_type,pre_synonyms=synonyms,pre_receptor_type=receptor_type,pre.dimorphism=dimorphism,pre.cachero.type=cachero.type),
              by = c('pre_id'='bodyid'))%>%
    rename(post_type = type)%>%
    left_join(mba %>%
                select("bodyid",'fru_dsx','consensus_nt','flywire_type','synonyms','receptor_type','cachero.type','dimorphism') %>%
                mutate(bodyid=as.integer(bodyid))%>%
                rename(post_fru_dsx=fru_dsx,post_nt=consensus_nt,post_fw_type=flywire_type,post_synonyms=synonyms,post_receptor_type=receptor_type,post.dimorphism=dimorphism,post.cachero.type=cachero.type),
              by = c('post_id'='bodyid'))
  return(connectivity)
}
```

### 1.3 Load connectivity
```{r}
conn <-fetch_connectivity()
conn_save <- conn

conn <- conn %>%
  mutate(
    pre_type = case_when(
      grepl("putative_ppk23", pre_receptor_type) ~ "m-cell",
      grepl("putative_ppk25", pre_receptor_type) ~ "f-cell",
      TRUE ~ pre_type
    ),
    post_type = case_when(
      grepl("putative_ppk23", post_receptor_type) ~ "m-cell",
      grepl("putative_ppk25", post_receptor_type) ~ "f-cell",
      TRUE ~ post_type
    )
  )
```

## 2 Adjacency matrix
### 2.1 Function adjacency matrix
```{r}
calculate_normed_adj_matrix <-function(connectivity,cell.or.type='type',pre.or.post='pre'){
  colScale <- function(A, na.rm = TRUE) {
    scalefac <- 1 / Matrix::colSums(A)
    if (na.rm) scalefac[!is.finite(scalefac)] <- 0
    B <- A %*% Matrix::Diagonal(x = scalefac)
    B
    }
  rowScale <- function(A, na.rm = TRUE) {
    scalefac <- 1 / Matrix::rowSums(A)
    if (na.rm) scalefac[!is.finite(scalefac)] <- 0
    B <- Matrix::Diagonal(x = scalefac) %*% A
    B
    }
  if(cell.or.type=='cell'){
    unique.identifier <- union(as.character(connectivity$pre_id), as.character(connectivity$post_id))
    adj.matrix <- sparseMatrix(
      i = match(connectivity$pre_id, unique.identifier),
      j = match(connectivity$post_id, unique.identifier),
      x = connectivity$weight,
      dims = c(length(unique.identifier), length(unique.identifier)),
      dimnames = list(unique.identifier, unique.identifier))
  }else{
    unique.identifier <- union(connectivity$pre_type, connectivity$post_type)
    adj.matrix <- sparseMatrix(
      i = match(connectivity$pre_type, unique.identifier),
      j = match(connectivity$post_type, unique.identifier),
      x = connectivity$weight,
      dims = c(length(unique.identifier), length(unique.identifier)),
      dimnames = list(unique.identifier, unique.identifier))
  }
  if(pre.or.post=='pre'){
    adj.matrix.normed.pre <- rowScale(adj.matrix)
    colnames(adj.matrix.normed.pre) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.pre) <- rownames(adj.matrix)
    return(adj.matrix.normed.pre)
  }else if(pre.or.post=='avg'){
    adj.matrix.normed.post <- colScale(adj.matrix)
    colnames(adj.matrix.normed.post) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.post) <- rownames(adj.matrix)
    adj.matrix.normed.pre <- rowScale(adj.matrix)
    colnames(adj.matrix.normed.pre) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.pre) <- rownames(adj.matrix)
    
    return(avg_mat <- (adj.matrix.normed.post + adj.matrix.normed.post) / 2)
    
  }
  
  else{
    adj.matrix.normed.post <- colScale(adj.matrix)
    colnames(adj.matrix.normed.post) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.post) <- rownames(adj.matrix)
    return(adj.matrix.normed.post)
  }
}
```

### 2.1 Create adjacency matrix
```{r}
use.clone <- F
conn<-conn_save
conn<- conn%>%
  mutate(pre_type=coalesce(pre_type,pre_fw_type),post_type=coalesce(post_type,post_fw_type))%>%
  mutate(pre_type=if_else(is.na(pre_type),as.character(bodyid),pre_type),
         post_type=if_else(is.na(post_type),as.character(partner),post_type))%>%
  mutate(pre_type = gsub("[()]", "", pre_type))%>%
  mutate(post_type = gsub("[()]", "", post_type))




#nam <-calculate_normed_adj_matrix(conn.wo.gaba,cell.or.type='type',pre.or.post='avg') #avg
adj.output.normed <-calculate_normed_adj_matrix(conn,cell.or.type='type',pre.or.post='post') #avg

g.prout.mcns <- graph_from_adjacency_matrix(
  adj.output.normed,
  mode = "directed",
  weighted = TRUE,
  diag = FALSE
)
```



define kstrongest path function
```{r}
add_valence_flag <- function(df, nt_by_type, is_inhib_by_nt) {
  df %>%
    rowwise() %>%
    mutate(
      # split path and drop endpoint
      nodes = list(setdiff(strsplit(path, " -> ", fixed = TRUE)[[1]], end)),
      # lookup NTs
      nts   = list(unname(nt_by_type[unlist(nodes)])),
      # lookup inhibitory flags
      inhib = list(unname(is_inhib_by_nt[unlist(nts)])),
      # count inhibitors
      n_inhib = sum(unlist(inhib), na.rm = TRUE),
      # convert to +1 / -1
      d = ifelse(n_inhib %% 2 == 0, 1, -1)
    ) %>%
    ungroup()
}



find_k_strongest_paths <- function(
  g,
  starts,
  targets,
  mba,
  n_paths = 3,
  diversity = c("otheredges","othernodes","none"),
  ignore_pattern = NULL  # grepl() pattern; matched node names are excluded (start/target kept),
  
){
  
  type2nt <- mba %>%select(type,consensus_nt)%>%group_by(type)%>%distinct()%>%rename(nt=consensus_nt)
  nt_by_type <- setNames(type2nt$nt, type2nt$type)
  inhibitory_nts <- c("gaba","glycine","histamine")
  all_nts <- unique(type2nt$nt)
  is_inhib_by_nt <- setNames(all_nts %in% inhibitory_nts, all_nts)
    
  diversity <- match.arg(diversity)

  # optional: drop nodes matching ignore_pattern (but never drop starts/targets)
  if (!is.null(ignore_pattern)) {
    drop <- V(g)$name[grepl(ignore_pattern, V(g)$name)]
    keep <- unique(c(starts, targets))
    drop <- setdiff(drop, keep)
    if (length(drop)) g <- igraph::delete_vertices(g, drop)
  }

  # weights → log-costs; guard zeros/NAs
  w <- E(g)$weight
  if (is.null(w)) stop("Edges must have numeric 'weight'.")
  w[!is.finite(w) | w <= 0] <- .Machine$double.eps
  log_w <- -log(w)

  results <- list()
  all_nodes <- character(0)

  for (s in starts) for (t in targets) {
    if (!(s %in% V(g)$name) || !(t %in% V(g)$name)) {
      results[[paste(s,t,sep="->")]] <- list(list(
        start = s, end = t, path = NA_character_, strength = NA_real_, hops = NA_integer_
      ))
      next
    }

    g_tmp <- g
    E(g_tmp)$log_weight <- log_w
    pair_paths <- list()

    for (i in seq_len(n_paths)) {
      sp <- igraph::shortest_paths(
        g_tmp, from = s, to = t,
        weights = E(g_tmp)$log_weight,
        output = "both"
      )
      if (length(sp$vpath[[1]]) == 0) break

      vpath <- sp$vpath[[1]]
      epath <- sp$epath[[1]]

      final_strength <- exp(-sum(E(g_tmp)[epath]$log_weight))
      hops <- length(epath)

      pair_paths[[paste0("path", i)]] <- list(
        start = s,
        end = t,
        path = paste(V(g)[vpath]$name, collapse = " -> "),
        strength = final_strength,
        hops = hops
      )

      all_nodes <- c(all_nodes, V(g)[vpath]$name)

      if (diversity == "othernodes") {
        inter <- setdiff(V(g)[vpath]$name, c(s, t))
        if (length(inter)) g_tmp <- igraph::delete_vertices(g_tmp, inter)
      } else if (diversity == "otheredges") {
        g_tmp <- igraph::delete_edges(g_tmp, epath)
      }
    }

    if (length(pair_paths) == 0) {
      pair_paths <- list(list(
        start = s, end = t, path = NA_character_, strength = NA_real_, hops = NA_integer_
      ))
    }

    results[[paste(s,t,sep="->")]] <- pair_paths
  }

  # flatten
  results_df <- do.call(rbind, lapply(results, function(pl) {
    do.call(rbind, lapply(pl, function(r)
      data.frame(start=r$start, end=r$end, path=r$path,
                 strength=r$strength, hops=r$hops,
                 stringsAsFactors=FALSE)))
  }))

  results_df_wo_na <- dplyr::filter(results_df, !is.na(strength))
  all_nodes <- unique(all_nodes)
  
  results_df_wo_na= add_valence_flag(results_df_wo_na, nt_by_type, is_inhib_by_nt)%>%
              mutate(valence = ifelse(d == 1L, "excitatory", "inhibitory"))
  
  list(df = results_df_wo_na, nodes = all_nodes, raw = results)
}

```



All Neurons downstream

```{r}

all.PPN1.ids <- mba%>%filter(grepl('PPN1',synonyms))%>%pull(bodyid)
all.vAB3.ids <- mba%>%filter(grepl('vAB3 Clowney',notes))%>%pull(bodyid)
target.vAB3.all <- cf_partner_summary(cf_ids(malecns=all.vAB3.ids),partners='out',normalise = F) %>% 
  group_by(type.post)%>%
  summarize(weight=sum(weight))%>%
  filter(!grepl('AN',type.post)) %>% 
  filter(!grepl('IN',type.post)) %>%
  arrange(desc(weight)) %>%
  filter(weight/sum(weight) >= 0.001) %>% 
  filter(!is.na(type.post)) %>% 
  pull(type.post)

target.PPN1.all <- cf_partner_summary(cf_ids(malecns=all.PPN1.ids),partners='out',normalise = F) %>% 
  group_by(type.post)%>%
  summarize(weight=sum(weight))%>%
  filter(!grepl('AN',type.post)) %>% 
  filter(!grepl('IN',type.post)) %>%
  arrange(desc(weight)) %>%
  filter(weight/sum(weight) >= 0.001) %>% 
  filter(!is.na(type.post)) %>% 
  pull(type.post)

target.vAB3.all <- c(target.vAB3.all[1:10],intersect(target.vAB3.all,target.PPN1.all)[1:3])
target.PPN1.all <- c(target.PPN1.all[1:10],intersect(target.vAB3.all,target.PPN1.all)[1:3])




start.types.DA1 <- c('ORN_DA1')
start.types.VA1v <- c('ORN_VA1v')
res.vAB3.DA1 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.DA1, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',mba=mba)
res.PPN1.DA1 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.DA1, targets=target.PPN1.all,n_paths=3,diversity='otheredges',mba=mba)
res.vAB3.VA1v <- find_k_strongest_paths(g.prout.mcns, starts=start.types.VA1v, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',mba=mba)
res.PPN1.VA1v <- find_k_strongest_paths(g.prout.mcns, starts=start.types.VA1v, targets=target.PPN1.all,n_paths=3,diversity='otheredges',mba=mba)

res.vAB3.DA1$df$dataset <- "vAB3"
res.PPN1.DA1$df$dataset <- "PPN1"
res.vAB3.VA1v$df$dataset <- "vAB3"
res.PPN1.VA1v$df$dataset <- "PPN1"
df_all.all.DA1 <- bind_rows(res.vAB3.DA1$df, res.PPN1.DA1$df)
df_all.all.VA1v <- bind_rows(res.vAB3.VA1v$df, res.PPN1.VA1v$df)


t_res.DA1  <- t.test(log(strength) ~ dataset, data = df_all.all.DA1)
pval.DA1   <- unname(t_res.DA1$p.value)
ann.DA1<- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval.DA1,
  y.position  = max(df_all.all.DA1$strength, na.rm = TRUE) * 1.12
)

t_res.VA1v  <- t.test(log(strength) ~ dataset, data = df_all.all.VA1v)
pval.VA1v   <- unname(t_res.VA1v$p.value)
ann.VA1v<- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval.VA1v,
  y.position  = max(df_all.all.VA1v$strength, na.rm = TRUE) * 1.12
)

ggplot(df_all.all.DA1, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength DA1") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann.DA1,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 

ggplot(df_all.all.VA1v, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength VA1v") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann.VA1v,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 


ggplot(df_all.all.DA1, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path DA1") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(df_all.all.VA1v, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path VA1v") +
  theme(plot.title = element_text(hjust = 0.5))



ann_strength.DA1 <- df_all.all.DA1 %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

ann_strength.VA1v <- df_all.all.VA1v %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

# strength boxplots, split by valence
ggplot(df_all.all.DA1, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (DA1) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength.DA1, label = "p", tip.length = 0)

ggplot(df_all.all.VA1v, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (VA1v) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength.VA1v, label = "p", tip.length = 0)


```

auditory
```{r}

#auditory
start.types.aud <- mba.static%>%filter(grepl('cb_sensory',superclass))%>%filter(grepl('auditory',subclass))%>%pull(type)%>%unique()
res.vAB3.aud <- find_k_strongest_paths(g.prout.mcns, starts=start.types.aud, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',mba=mba)
res.PPN1.aud <- find_k_strongest_paths(g.prout.mcns, starts=start.types.aud, targets=target.PPN1.all,n_paths=3,diversity='otheredges',mba=mba)

res.vAB3.aud$df$dataset <- "vAB3"
res.PPN1.aud$df$dataset <- "PPN1"
df_all.all.aud <- bind_rows(res.vAB3.aud$df, res.PPN1.aud$df)



t_res  <- t.test(log(strength) ~ dataset, data = df_all.all.aud)
pval   <- unname(t_res$p.value)
ann <- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval,
  y.position  = max(df_all.all.aud$strength, na.rm = TRUE) * 1.12
)

ggplot(df_all.all.aud, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength auditory") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 


ggplot(df_all.all.aud, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path auditory") +
  theme(plot.title = element_text(hjust = 0.5))



ann_strength <- df_all.all.aud %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

# strength boxplots, split by valence
ggplot(df_all.all.aud, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (auditory) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength, label = "p", tip.length = 0)

```


```{r}
#visual
start.types.vis <- mba%>%filter(superclass=='ol_sensory')%>%pull(type)%>%unique()
res.vAB3.vis <- find_k_strongest_paths(g.prout.mcns, starts=start.types.vis, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',mba=mba)
res.PPN1.vis <- find_k_strongest_paths(g.prout.mcns, starts=start.types.vis, targets=target.PPN1.all,n_paths=3,diversity='otheredges',mba=mba)

res.vAB3.vis$df$dataset <- "vAB3"
res.PPN1.vis$df$dataset <- "PPN1"
df_all.all.vis <- bind_rows(res.vAB3.vis$df, res.PPN1.vis$df)



t_res  <- t.test(log(strength) ~ dataset, data = df_all.all.vis)
pval   <- unname(t_res$p.value)
ann <- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval,
  y.position  = max(df_all.all.vis$strength, na.rm = TRUE) * 1.12
)

ggplot(df_all.all.vis, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength Visual") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 


ggplot(df_all.all.vis, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path Visual") +
  theme(plot.title = element_text(hjust = 0.5))



ann_strength <- df_all.all.vis %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

# strength boxplots, split by valence
ggplot(df_all.all.vis, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (visual) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength, label = "p", tip.length = 0)

```

```{r}


#contact chemosensor ppk23

start.types.ppk23 <- mba.static%>%filter(grepl('ppk23',receptor_type))%>%pull(type)%>%unique()
res.vAB3.ppk23 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk23, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',ignore_pattern='AN09B017',mba=mba)
res.PPN1.ppk23 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk23, targets=target.PPN1.all,n_paths=3,diversity='otheredges',ignore_pattern='AN05B102',mba=mba)
res.vAB3.ppk23.include <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk23, targets=target.vAB3.all ,n_paths=100,diversity='otheredges',mba=mba)
res.PPN1.ppk23.include <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk23, targets=target.PPN1.all,n_paths=100,diversity='otheredges',mba=mba)

res.vAB3.ppk23$df$dataset <- "vAB3"
res.PPN1.ppk23$df$dataset <- "PPN1"
res.vAB3.ppk23.include$df$dataset <- "vAB3"
res.PPN1.ppk23.include$df$dataset <- "PPN1"
df_all.all.ppk23 <- bind_rows(res.vAB3.ppk23$df, res.PPN1.ppk23$df)
df_all.all.ppk23.include <- bind_rows(res.vAB3.ppk23.include$df, res.PPN1.ppk23.include$df)


t_res  <- t.test(log(strength) ~ dataset, data = df_all.all.ppk23)
pval   <- unname(t_res$p.value)
ann <- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval,
  y.position  = max(df_all.all.ppk23$strength, na.rm = TRUE) * 1.12
)

ggplot(df_all.all.ppk23, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength PPK23") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 


ggplot(df_all.all.ppk23, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path PPK23") +
  theme(plot.title = element_text(hjust = 0.5))



ann_strength <- df_all.all.ppk23 %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

# strength boxplots, split by valence
ggplot(df_all.all.ppk23, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (PPK23) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength, label = "p", tip.length = 0)
```


```{r}
#contact chemosensor ppk25
start.types.ppk25 <- mba.static%>%filter(grepl('ppk25',receptor_type))%>%pull(type)%>%unique()
res.vAB3.ppk25 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk25, targets=target.vAB3.all ,n_paths=3,diversity='otheredges',ignore_pattern='AN09B017',mba=mba)
res.PPN1.ppk25 <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk25, targets=target.PPN1.all,n_paths=3,diversity='otheredges',ignore_pattern='AN05B102',mba=mba)
res.vAB3.ppk25.include <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk25, targets=target.vAB3.all ,n_paths=100,diversity='otheredges',mba=mba)
res.PPN1.ppk25.include <- find_k_strongest_paths(g.prout.mcns, starts=start.types.ppk25, targets=target.PPN1.all,n_paths=100,diversity='otheredges',mba=mba)


res.vAB3.ppk25$df$dataset <- "vAB3"
res.PPN1.ppk25$df$dataset <- "PPN1"
res.vAB3.ppk25.include$df$dataset <- "vAB3"
res.PPN1.ppk25.include$df$dataset <- "PPN1"
df_all.all.ppk25 <- bind_rows(res.vAB3.ppk25$df, res.PPN1.ppk25$df)
df_all.all.ppk25.include <- bind_rows(res.vAB3.ppk25.include$df, res.PPN1.ppk25.include$df)


t_res  <- t.test(log(strength) ~ dataset, data = df_all.all.ppk25)
pval   <- unname(t_res$p.value)
ann <- data.frame(
  group1      = "PPN1",
  group2      = "vAB3",
  p           = pval,
  y.position  = max(df_all.all.ppk25$strength, na.rm = TRUE) * 1.12
)

ggplot(df_all.all.ppk25, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  theme_bw() +
  labs(title = "Path strength ppk25") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_pvalue_manual(
    ann,
    label = "p",         # use the column 'p'
    tip.length = 0
  )+
  scale_y_log10() 


ggplot(df_all.all.ppk25, aes(x = dataset, y = hops)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
  scale_y_log10() +
  theme_bw()+
  labs(title = "Hops in strongest path ppk25") +
  theme(plot.title = element_text(hjust = 0.5))



ann_strength <- df_all.all.ppk25 %>%
  group_by(valence) %>%
  summarise(
    group1 = "PPN1",
    group2 = "vAB3",
    p = unname(t.test(log(strength) ~ dataset)$p.value),
    y.position = max(strength, na.rm = TRUE) * 1.12,
    .groups = "drop"
  )

# strength boxplots, split by valence
ggplot(df_all.all.ppk25, aes(x = dataset, y = strength)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4, color = "blue") +
  scale_y_log10() +
  theme_bw() +
  labs(title = "Path strength (ppk25) by valence") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ valence, nrow = 1) +
  stat_pvalue_manual(ann_strength, label = "p", tip.length = 0)
```




```{r}
library(scales)


# add modality labels (as you already do)
df_all.all.ppk25$modality <- 'female.touch'
df_all.all.ppk23$modality <- 'male.touch'
df_all.all.aud$modality   <- 'auditory'
df_all.all.DA1$modality   <- 'DA1'
df_all.all.VA1v$modality   <- 'VA1v'
df_all.all.vis$modality   <- 'visual'

all.modalities <- rbind(
  df_all.all.ppk25,
  df_all.all.ppk23,
  df_all.all.VA1v,
  df_all.all.DA1
)

pie_data <- function(df, dataset_name, valence_keep = NULL) {
  df %>%
    filter(dataset == dataset_name) %>%
    { if (!is.null(valence_keep)) filter(., valence %in% valence_keep) else . } %>%
    group_by(end, modality) %>%
    summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop_last") %>%
    mutate(prop = total_strength / sum(total_strength)) %>%
    ungroup() %>%
    filter(is.finite(prop))  # drop ends with all-zero strength
}

pie_plot <- function(pdat, title_txt) {
  ggplot(pdat, aes(x = "", y = prop, fill = modality)) +
    geom_col(width = 1, color = "white") +
    coord_polar(theta = "y") +
    facet_wrap(~ end) +
    theme_void() +
    labs(title = title_txt) +
    geom_text(aes(label = percent(prop, accuracy = 1)),
              position = position_stack(vjust = 0.5), size = 2)
}

## PPN1
ppn1_all <- pie_data(all.modalities, "PPN1", NULL)
ppn1_exc <- pie_data(all.modalities, "PPN1", c("excitatory", 1))     # handles either text or numeric
ppn1_inh <- pie_data(all.modalities, "PPN1", c("inhibitory", -1))

p_ppn1_all <- pie_plot(ppn1_all, "PPN1: modality contribution per target neuron (all)")
p_ppn1_exc <- pie_plot(ppn1_exc, "PPN1: modality contribution per target neuron (excitatory only)")
p_ppn1_inh <- pie_plot(ppn1_inh, "PPN1: modality contribution per target neuron (inhibitory only)")

p_ppn1_all; p_ppn1_exc; p_ppn1_inh

ppn1_all$connection.type <- 'all'
ppn1_exc$connection.type <- 'excitatory'
ppn1_inh$connection.type <- 'inhibitory'

connected.scatterplot.all.ppn1 <- rbind(ppn1_all,ppn1_exc,ppn1_inh)%>%
                          mutate(
                            x_pos = as.numeric(factor(end, levels = unique(end))),
                            x_shift = case_when(
                              connection.type == "inhibitory" ~ -0.2,
                              connection.type == "all" ~ 0,
                              connection.type == "excitatory" ~ 0.2
                            ),
                            x_jittered = x_pos + x_shift
                          )

ggplot(connected.scatterplot.all.ppn1, aes(x = x_jittered, y = prop, group = interaction(end, modality), color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +    # line inherits modality color
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(connected.scatterplot.all.ppn1$x_pos),
    labels = unique(connected.scatterplot.all.ppn1$end)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (excluding PPN1) to post PPN1 (all) by connection type",
    x = "Target neurons",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+ theme(axis.text.x = element_text(
    angle = 90, hjust = 1,
    color = ifelse(unique(connected.scatterplot.all.ppn1$end) %in% intersect(target.vAB3.all, target.PPN1.all), "red", "black")
  ))

## vAB3
vab3_all <- pie_data(all.modalities, "vAB3", NULL)
vab3_exc <- pie_data(all.modalities, "vAB3", c("excitatory", 1))
vab3_inh <- pie_data(all.modalities, "vAB3", c("inhibitory", -1))

p_vab3_all <- pie_plot(vab3_all, "vAB3: modality contribution per target neuron (all)")
p_vab3_exc <- pie_plot(vab3_exc, "vAB3: modality contribution per target neuron (excitatory only)")
p_vab3_inh <- pie_plot(vab3_inh, "vAB3: modality contribution per target neuron (inhibitory only)")


p_vab3_all; p_vab3_exc; p_vab3_inh

#connected scatter plot

vab3_all$connection.type <- 'all'
vab3_exc$connection.type <- 'excitatory'
vab3_inh$connection.type <- 'inhibitory'

connected.scatterplot.all.vab3 <- rbind(vab3_all,vab3_exc,vab3_inh)%>%
                          mutate(
                            x_pos = as.numeric(factor(end, levels = unique(end))),
                            x_shift = case_when(
                              connection.type == "inhibitory" ~ -0.2,
                              connection.type == "all" ~ 0,
                              connection.type == "excitatory" ~ 0.2
                            ),
                            x_jittered = x_pos + x_shift
                          )

ggplot(connected.scatterplot.all.vab3, aes(x = x_jittered, y = prop, group = interaction(end, modality), color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +    # line inherits modality color
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(connected.scatterplot.all.vab3$x_pos),
    labels = unique(connected.scatterplot.all.vab3$end)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (excluding vAB3) to post vAB3 (all) by connection type",
    x = "Target neurons",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+ theme(axis.text.x = element_text(
    angle = 90, hjust = 1,
    color = ifelse(unique(connected.scatterplot.all.vab3$end) %in% intersect(target.vAB3.all, target.PPN1.all), "red", "black")
  ))



```



```{r}


# add modality labels (as you already do)
df_all.all.ppk25.include$modality <- 'female.touch'
df_all.all.ppk23.include$modality <- 'male.touch'
df_all.all.aud$modality   <- 'auditory'
df_all.all.DA1$modality   <- 'DA1'
df_all.all.VA1v$modality   <- 'VA1v'
df_all.all.vis$modality   <- 'visual'

all.modalities <- rbind(  
  df_all.all.ppk25.include,
  df_all.all.ppk23.include,
  df_all.all.vis,
  df_all.all.aud,
  df_all.all.VA1v,
  df_all.all.DA1
)

all.modalities <- rbind(  
  df_all.all.ppk25.include,
  df_all.all.ppk23.include,
  df_all.all.VA1v,
  df_all.all.DA1
)

all.modalities <- rbind(  
  df_all.all.ppk25.include,
  df_all.all.ppk23.include
)


## PPN1
ppn1_all <- pie_data(all.modalities, "PPN1", NULL)
ppn1_exc <- pie_data(all.modalities, "PPN1", c("excitatory", 1))     # handles either text or numeric
ppn1_inh <- pie_data(all.modalities, "PPN1", c("inhibitory", -1))

p_ppn1_all <- pie_plot(ppn1_all, "PPN1: modality contribution per target neuron (all)")
p_ppn1_exc <- pie_plot(ppn1_exc, "PPN1: modality contribution per target neuron (excitatory only) Including PPN1")
p_ppn1_inh <- pie_plot(ppn1_inh, "PPN1: modality contribution per target neuron (inhibitory only) Including PPN1")

p_ppn1_all; p_ppn1_exc; p_ppn1_inh

ppn1_all$connection.type <- 'all'
ppn1_exc$connection.type <- 'excitatory'
ppn1_inh$connection.type <- 'inhibitory'

connected.scatterplot.all.include.ppn1 <- rbind(ppn1_all,ppn1_exc,ppn1_inh)%>%
                          mutate(
                            x_pos = as.numeric(factor(end, levels = unique(end))),
                            x_shift = case_when(
                              connection.type == "inhibitory" ~ -0.2,
                              connection.type == "all" ~ 0,
                              connection.type == "excitatory" ~ 0.2
                            ),
                            x_jittered = x_pos + x_shift
                          )

ggplot(connected.scatterplot.all.include.ppn1, aes(x = x_jittered, y = prop, group = interaction(end, modality), color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +    # line inherits modality color
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(connected.scatterplot.all.include.ppn1$x_pos),
    labels = unique(connected.scatterplot.all.include.ppn1$end)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (including PPN1) to post PPN1 (all) by connection type",
    x = "Target neurons",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+ theme(axis.text.x = element_text(
    angle = 90, hjust = 1,
    color = ifelse(unique(connected.scatterplot.all.ppn1$end) %in% intersect(target.vAB3.all, target.PPN1.all), "red", "black")
  ))

df <- connected.scatterplot.all.include.ppn1%>%
  group_by(connection.type,modality)%>%
  summarize(strength=sum(total_strength)) %>%
  group_by(connection.type) %>%
  mutate(prop = strength / sum(strength)) %>%
  ungroup() %>%
  mutate(
    x_pos = as.numeric(factor(modality, levels = unique(modality))),
    x_shift = case_when(
      connection.type == "inhibitory" ~ -0.2,
      connection.type == "all"        ~  0,
      connection.type == "excitatory" ~  0.2
    ),
    x_jittered = x_pos + x_shift
  )

ggplot(df , aes(x = x_jittered, y = prop, group = modality, color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(df$x_pos),
    labels = unique(df$modality)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (including PPN1) to post PPN1 (all) by connection type across neurons",
    x = "Modality",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(hjust = 0.5))+
  theme(
        plot.title = element_text(hjust = 0.5)
    )

ggplot(df%>% filter(connection.type != 'all'), aes(x = connection.type, y = prop, fill = modality)) +
    geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.2) +
    theme_bw() +
    labs(
        title = "Neurotransmitter valence of input (including PPN1) to post PPN1 (all) by connection type",
        x = "Modality",
        y = "Proportion",
        fill = "Connection type"
    )+
  theme(axis.text.x = element_text(hjust = 0.5))+
  theme(
        plot.title = element_text(hjust = 0.5)
    )


## vAB3
vab3_all <- pie_data(all.modalities, "vAB3", NULL)
vab3_exc <- pie_data(all.modalities, "vAB3", c("excitatory", 1))
vab3_inh <- pie_data(all.modalities, "vAB3", c("inhibitory", -1))

p_vab3_all <- pie_plot(vab3_all, "vAB3: modality contribution per target neuron (all)")
p_vab3_exc <- pie_plot(vab3_exc, "vAB3: modality contribution per target neuron (excitatory only) Including vAB3")
p_vab3_inh <- pie_plot(vab3_inh, "vAB3: modality contribution per target neuron (inhibitory only) Including vAB3")

# print any of these:
p_vab3_all; p_vab3_exc; p_vab3_inh

#connected scatter plot

vab3_all$connection.type <- 'all'
vab3_exc$connection.type <- 'excitatory'
vab3_inh$connection.type <- 'inhibitory'

connected.scatterplot.all.include.vab3 <- rbind(vab3_all,vab3_exc,vab3_inh)%>%
                          mutate(
                            x_pos = as.numeric(factor(end, levels = unique(end))),
                            x_shift = case_when(
                              connection.type == "inhibitory" ~ -0.2,
                              connection.type == "all" ~ 0,
                              connection.type == "excitatory" ~ 0.2
                            ),
                            x_jittered = x_pos + x_shift
                          )

ggplot(connected.scatterplot.all.include.vab3, aes(x = x_jittered, y = prop, group = interaction(end, modality), color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +    # line inherits modality color
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(connected.scatterplot.all.include.vab3$x_pos),
    labels = unique(connected.scatterplot.all.include.vab3$end)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (including vAB3) to post vAB3 (all) by connection type",
    x = "Target neurons",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+ theme(axis.text.x = element_text(
    angle = 90, hjust = 1,
    color = ifelse(unique(connected.scatterplot.all.vab3$end) %in% intersect(target.vAB3.all, target.PPN1.all), "red", "black")
  ))


df <- connected.scatterplot.all.include.vab3%>%
  group_by(connection.type,modality)%>%
  summarize(strength=sum(total_strength)) %>%
  group_by(connection.type) %>%
  mutate(prop = strength / sum(strength)) %>%
  ungroup() %>%
  mutate(
    x_pos = as.numeric(factor(modality, levels = unique(modality))),
    x_shift = case_when(
      connection.type == "inhibitory" ~ -0.2,
      connection.type == "all"        ~  0,
      connection.type == "excitatory" ~  0.2
    ),
    x_jittered = x_pos + x_shift
  )

ggplot(df, aes(x = x_jittered, y = prop, group = modality, color = modality)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(aes(shape = connection.type), size = 3, alpha = 0.9) +
  scale_x_continuous(
    breaks = unique(df$x_pos),
    labels = unique(df$modality)
  ) +
  theme_bw() +
  labs(
    title = "Fraction of input (including vAB3) to post vAB3 (all) by connection type across neurons",
    x = "Modality",
    y = "Proportion",
    color = "Modality",
    shape = "Connection type"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(df %>% filter(connection.type != 'all'), aes(x = connection.type, y = prop, fill = modality)) +
    geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.2) +
    theme_bw() +
    labs(
        title = "Neurotransmitter valence of input (including vAB3) to post vAB3 (all) by connection type",
        x = "Modality",
        y = "Proportion",
        fill = "Connection type"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5)
    )



```

stacked barplot showing how much the strongest paths from the modalities take of the synapse budget 


```{r}
library(dplyr)
library(stringr)
library(purrr)
  all.modalities.second_last <- all.modalities %>%
  mutate(second_last = map_chr(str_split(path, " -> "), ~ .x[length(.x) - 1]))

pairs <- all.modalities.second_last %>%
  select(second_last, end) %>%
  distinct() %>%
  mutate(weight = igraph::E(g.prout.mcns)$weight[
    igraph::get.edge.ids(g.prout.mcns, as.vector(t(as.matrix(cur_data()))), directed = TRUE, error = FALSE)
  ])

all.modalities.second_last <- all.modalities.second_last %>%
  left_join(pairs, by = c("second_last", "end"))

aaa <- all.modalities.second_last %>% 
  select(second_last, end, weight, modality, dataset) %>% 
  distinct() %>% 
  group_by(end, modality, dataset) %>% 
  summarise(weight = sum(weight) / 5, .groups = "drop")




```

test that code is correct
```{r}

ggplot(aaa, aes(x = end, y = weight, fill = modality)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ dataset, scales = "free_x") +
  theme_minimal() +
  labs(x = "Neuron (end)", y = "Weight", fill = "Modality") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```

Look at certain subnetworks in cytoscpae
```{r}
library(stringr)
library(igraph)

input<- all.modalities %>% filter(modality %in% c('male.touch','female.touch','DA1','VA1v'))
# 1) collect unique node names from the path strings
nodes <- unique(unlist(strsplit(input$path, " *-> *")))

# 2) keep only nodes present in the graph
nodes_in_g <- intersect(nodes, V(g.prout.mcns)$name)

# 3) induced subgraph on those vertices
g_sub_vertices <- induced_subgraph(g.prout.mcns, vids = V(g.prout.mcns)[name %in% nodes_in_g])


library(RCy3)

cytoscapePing()                                      # sanity check
net.suid <- createNetworkFromIgraph(g_sub_vertices,     # or g_sub_vertices
                                    title = "mcns subnetwork",
                                    collection = "mcns")

layoutNetwork("force-directed")
fitContent()

```



```{r}
# Edge-extracted subgraph: parse paths -> build edge list -> select matching edges -> subgraph

library(dplyr)
library(stringr)
library(igraph)
library(RCy3)

library(stringr)
library(purrr)
library(igraph)
input<- all.modalities %>% filter(modality %in% c('male.touch','female.touch','DA1','VA1v'))




paths <- input$path %>% as.character() %>% na.omit()
parts <- strsplit(paths, " *-+> *")

edge_pairs <- map_dfr(parts, function(x) {
  x <- trimws(x)
  if (length(x) < 2) return(NULL)
  data.frame(from = head(x, -1), to = tail(x, -1), stringsAsFactors = FALSE)
}) %>%
  unique()

eid <- igraph::get.edge.ids(g.prout.mcns,
                            as.vector(t(as.matrix(edge_pairs))),
                            directed = TRUE, error = FALSE)
eid <- unique(eid[eid > 0])

g_sub_edges <- subgraph.edges(g.prout.mcns, eids = eid, delete.vertices = TRUE)


# use `input` (already filtered) — switch to `all.modalities` if you want all
paths   <- input$path %>% as.character() %>% na.omit()
mods    <- input$modality[match(paths, input$path)]  # align modalities to paths
parts   <- strsplit(paths, " *-+> *")

nodes_by_mod <- map2_dfr(parts, mods, function(nodes, m) {
  data.frame(node = trimws(nodes), modality = m, stringsAsFactors = FALSE)
}) %>%
  filter(nzchar(node)) %>%
  distinct(node, modality)

modalities_by_node <- nodes_by_mod %>%
  group_by(node) %>%
  summarise(
    modalities     = paste(sort(unique(modality)), collapse = ", "),
    n_modalities   = n_distinct(modality),
    .groups = "drop"
  )


# attach to your subgraph vertices
mt <- match(V(g_sub_edges)$name, modalities_by_node$node)
V(g_sub_edges)$modalities   <- modalities_by_node$modalities[mt]
V(g_sub_edges)$n_modalities <- modalities_by_node$n_modalities[mt]

V(g_sub_edges)$is_start <- V(g_sub_edges)$name %in% unique(input$start)
V(g_sub_edges)$is_end <- V(g_sub_edges)$name %in% unique(input$end)

V(g_sub_edges)$consensus_nt <- {
  kb <- dplyr::bind_rows(
    dplyr::transmute(mba, name = as.character(bodyid),   consensus_nt),
    dplyr::transmute(mba, name = as.character(type),     consensus_nt),
    dplyr::transmute(mba, name = as.character(instance), consensus_nt)
  ) |>
  dplyr::filter(!is.na(name) & nzchar(name)) |>
  dplyr::group_by(name) |>
  dplyr::summarise(consensus_nt = dplyr::coalesce(consensus_nt[!is.na(consensus_nt)][1], NA_character_), .groups="drop")
  kb$consensus_nt[match(V(g_sub_edges)$name, kb$name)]
}


# 5) send to Cytoscape (all vertex/edge attributes preserved)
cytoscapePing()
net.suid <- createNetworkFromIgraph(g_sub_edges,
                                    title = "mcns subnetwork (edge-extracted)",
                                    collection = "mcns")
layoutNetwork("force-directed")
fitContent()

```

attempts to find mediator
```{r}
all.modalities %>%
  filter(modality == "male.touch",
         dataset == "PPN1",
         valence == "excitatory") %>%
  unnest_longer(nodes) %>%
  filter(!nodes %in% c(start.types.ppk25,
                       start.types.ppk23,
                       target.vAB3.all,
                       target.PPN1.all)) %>%
  count(nodes, sort = TRUE) %>% mutate(n_frac=n/max(n)) ->a

mba %>% 
  select(type, superclass, bodyid) %>% 
  filter(type %in% nodes & !grepl('sensory', superclass)) %>% 
  left_join(a, by = c('type' = 'nodes')) %>% 
  filter(!is.na(n_frac)) -> aaa

aaa <- aaa %>% arrange(desc(n_frac)) %>% filter(n_frac >= 0.3)

```




```{r}
# Example of CN receiving input from MBON and LHN

library(nat.ggplot)
library(ggplot2)
library(dplyr)

dir = '/Users/fkampf/Documents/pheromone.paper/strongest.path/morphology compared across modality valence path'
random_pastel <- function() {
  rgb(runif(1, 0.5, 1), runif(1, 0.5, 1), runif(1, 0.5, 1))
}


anterior = structure(c(1, 0, 0, 0, 
                       0, -1, 0.000000000000000122464679914735, 0, 
                       0, -0.000000000000000122464679914735, -1, 0, 
                       0, 0, 0, 1), 
                     dim = c(4L, 4L))

top = structure(c(1, 0, 0, 0,
                  0, 0, 1, 0,
                  0, -1, 0, 0,
                  0, 0, 0, 1),
                dim = c(4L, 4L))

view = top 

# Switch to cns just for plotting (V9 snapshot doesn't work)
choose_mcns_dataset("cns")
mc_ids <- c(aaa$bodyid) 
mcns.meshes=read_mcns_meshes(mc_ids)
names(mcns.meshes) <- aaa$type
key <- match(names(mcns.meshes), aaa$type)
alphas <- aaa$n_frac[key]
names(alphas) <- names(mcns.meshes)


p1 <- gganat +
  geom_neuron(malecns.surf, rotation_matrix=view, cols=c("grey95","grey85"), alpha=0.3) +
  geom_neuron(malecnsvnc.surf, rotation_matrix=view, cols=c("grey95","grey85"), alpha=0.3)

p2 <- gganat +
  geom_neuron(malecns.surf, rotation_matrix=anterior, cols=c("grey95","grey85"), alpha=0.3) +
  geom_neuron(malecnsvnc.surf, rotation_matrix=anterior, cols=c("grey95","grey85"), alpha=0.3)

for(i in seq_along(mcns.meshes)){
  p1 <- p1 + geom_neuron(mcns.meshes[[i]],
                       rotation_matrix=view,
                       cols=c(random_pastel()),
                       alpha=unname(alphas[i]))
  
  p2 <- p2 + geom_neuron(mcns.meshes[[i]],
                       rotation_matrix=anterior,
                       cols=c(random_pastel()),
                       alpha=unname(alphas[i]))
}


path = 'PPN1'
valence = 'excitatory'
modality = 'male.touch'


ggsave(
  file.path(dir,'dorsal',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path.png")),
  p1,
  width = 7, height = 7, dpi = 600, bg = "white"
)

ggsave(
  file.path(dir,'anterior',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path.png")),
  p2,
  width = 7, height = 7, dpi = 600, bg = "white"
)
```
```{r}

```


synapse
```{r}
for (modality in c('female.touch','male.touch')){
  for (path in c('PPN1','vAB3')){
    for (valence in c('excitatory','inhibitory')){
comp1 <-modality
comp2 <- path
comp3 <- valence
all.modalities %>%
  filter(modality == comp1) %>% 
  filter(dataset == comp2) %>% 
  filter(valence == comp3) %>%
  unnest_longer(nodes) %>%
  filter(!nodes %in% c(start.types.ppk25,
                       start.types.ppk23,
                       target.vAB3.all,
                       target.PPN1.all)) %>%
  count(nodes, sort = TRUE) %>% mutate(n_frac=n/max(n)) ->a

mba %>% 
  select(type, superclass, bodyid) %>% 
  filter(type %in% nodes & !grepl('sensory', superclass)) %>% 
  left_join(a, by = c('type' = 'nodes')) %>% 
  filter(!is.na(n_frac)) -> aaa

aaa <- aaa %>% arrange(desc(n_frac)) %>% filter(n_frac >= 0.3)




syn_list <- lapply(aaa$bodyid, function(x) {
  neuprint_get_synapses(x) %>%
    mutate(prepost = as.factor(ifelse(prepost == 1, "input", "output")))
})






# assumes: syn_list (list of synapse dfs), aaa$bodyid (vector of bodyids), mba (optional)

library(dplyr)
library(ggplot2)
library(patchwork)
library(grid)

ids <- aaa$bodyid
stopifnot(length(syn_list) == length(ids))

# bind all synapses with bodyid (+ optional type)
syn_all <- bind_rows(
  lapply(seq_along(syn_list), function(i) mutate(syn_list[[i]], bodyid = ids[i]))
)
if (exists("mba")) {
  syn_all <- syn_all %>% left_join(select(mba, bodyid, type), by = "bodyid")
}

syn_all <- syn_all %>%
  mutate(
    prepost = factor(prepost, levels = c("input", "output"))
  ) %>% 
  left_join(aaa %>% distinct(type,.keep_all = TRUE)%>% select(type,n_frac),by="type")

base_theme <- theme_minimal() +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())



#scatter plot
p_xy <- ggplot(syn_all, aes(x = x, y = z, color = prepost, alpha = n_frac)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = c(input = "blue", output = "red")) +
  scale_alpha_continuous(range = c(0.0001, 0.01)) +
  scale_y_reverse() +
  coord_fixed() +
  labs(title = "All synapses (XY)") +
  base_theme +
  guides(
    colour = guide_legend(override.aes = list(alpha = 1, size = 2)),
    alpha  = guide_legend(title = "n_frac")
  )

p_xz <- ggplot(syn_all, aes(x = x, y = y, color = prepost, alpha = n_frac)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = c(input = "blue", output = "red")) +
  scale_alpha_continuous(range = c(0.0001, 0.01)) +
  scale_y_reverse() +
  coord_fixed() +
  labs(title = "All synapses (XZ)") +
  base_theme +
  guides(
    colour = guide_legend(override.aes = list(alpha = 1, size = 2)),
    alpha  = guide_legend(title = "n_frac")
  )

combined <- p_xy / p_xz + plot_layout(nrow = 2, heights = c(1, 1))
g <- grid.grabExpr(print(combined))


ggsave(
  file.path(getwd(),'output.synapses.exploration','scatter',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  g,
  width = 7, height = 7, dpi = 900, bg = "white"
)



#heatmap

library(patchwork)
library(dplyr)

# compute common limits
common_limits <- range(
  ggplot_build(
    ggplot(syn_all, aes(x = x, y = y)) + stat_bin2d(bins = 100)
  )$data[[1]]$count,
  na.rm = TRUE
)

# define base heatmap function
heat_map_xz <- function(df, title, show_legend = TRUE) {
  ggplot(df, aes(x = x, y = y)) +
    stat_bin2d(bins = 100, drop = TRUE) +
    scale_fill_viridis_c(
      name = "count",
      trans = "log10",
      limits = common_limits
    ) +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = title, x = "x", y = "z") +
    theme_bw() +
    theme(legend.position = if (show_legend) "right" else "none",
          axis.text = element_blank(),
          axis.ticks = element_blank())
}

heat_map_xy <- function(df, title, show_legend = TRUE) {
  ggplot(df, aes(x = x, y = z)) +
    stat_bin2d(bins = 150, drop = TRUE) +
    scale_fill_viridis_c(
      name = "count",
      trans = "log10",
      limits = common_limits
    ) +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = title, x = "x", y = "y") +
    theme_bw() +
    theme(legend.position = if (show_legend) "right" else "none",
          axis.text = element_blank(),
          axis.ticks = element_blank())
}

p_in  <- heat_map_xz(filter(syn_all, prepost == "input"),  "Input",  FALSE)
p_out <- heat_map_xz(filter(syn_all, prepost == "output"), "Output", FALSE)
p_all <- heat_map_xz(syn_all,                              "All synapses", TRUE)
p_in_xy  <- heat_map_xy(filter(syn_all, prepost == "input"),  "Input",  FALSE)
p_out_xy <- heat_map_xy(filter(syn_all, prepost == "output"), "Output", FALSE)
p_all_xy <- heat_map_xy(syn_all,                              "All synapses", TRUE)



library(cowplot)

# one legend only
legend <- get_legend(p_all + theme(legend.position = "right"))
p_in  <- p_in  + theme(legend.position = "none")
p_out <- p_out + theme(legend.position = "none")
p_all <- p_all + theme(legend.position = "none")
p_in_xy  <- p_in_xy  + theme(legend.position = "none")
p_out_xy <- p_out_xy + theme(legend.position = "none")
p_all_xy <- p_all_xy + theme(legend.position = "none")

r1 <- plot_grid(p_in, p_out, p_all, nrow = 1, rel_widths = c(1,1,1), align = "h", axis = "tb")
r2 <- plot_grid(p_in_xy, p_out_xy, p_all_xy, nrow = 1, rel_widths = c(1.2,1.2,1.2), align = "h", axis = "tb")
grid <- plot_grid(r1, r2, ncol = 1, rel_heights = c(1,1))

ggg <- plot_grid(grid, legend, ncol = 2, rel_widths = c(1, 0.12))

ggsave(
  file.path(getwd(),'output.synapses.exploration','heatmap',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  ggg,
  width = 7, height = 7, dpi = 900, bg = "white"
)

#Density

density_xz <- ggplot(syn_all, aes(x = x, y = y)) +
    stat_density_2d(aes(fill = after_stat(level)), geom = "polygon", bins = 12, alpha = 0.6) +
    facet_wrap(~ prepost, ncol = 2) +
    scale_fill_viridis_c(name = "density level") +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = "Density (XZ)", x = "x", y = "z") +
    theme_bw()+
      theme(axis.text = element_blank(),
          axis.ticks = element_blank())

density_xy <- ggplot(syn_all, aes(x = x, y = z)) +
    stat_density_2d(aes(fill = after_stat(level)), geom = "polygon", bins = 12, alpha = 0.6) +
    facet_wrap(~ prepost, ncol = 2) +
    scale_fill_viridis_c(name = "density level") +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = "Density (XZ)", x = "x", y = "y") +
    theme_bw()+
    theme(axis.text = element_blank(),
          axis.ticks = element_blank())

combined_density <- density_xy + density_xz

combined_density

ggsave(
  file.path(getwd(),'output.synapses.exploration','density',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  combined_density,
  width = 7, height = 7, dpi = 900, bg = "white"
)
}}}

```



without the corresponding neurons
```{r}
for (modality in c('female.touch','male.touch')){
  for (path in c('PPN1','vAB3')){
    for (valence in c('excitatory','inhibitory')){
comp1 <-modality
comp2 <- path
comp3 <- valence
all.modalities %>%
  filter(modality == comp1) %>% 
  filter(dataset == comp2) %>% 
  filter(valence == comp3) %>%
  unnest_longer(nodes) %>%
  filter(!nodes %in% c(start.types.ppk25,
                       start.types.ppk23,
                       target.vAB3.all,
                       target.PPN1.all)) %>%
  filter(!grepl('AN09B017',nodes)) %>%
  filter(!grepl('AN05B102',nodes)) %>% 
  count(nodes, sort = TRUE) %>% mutate(n_frac=n/max(n)) ->a

mba %>% 
  select(type, superclass, bodyid) %>% 
  filter(type %in% nodes & !grepl('sensory', superclass)) %>% 
  left_join(a, by = c('type' = 'nodes')) %>% 
  filter(!is.na(n_frac)) -> aaa

aaa <- aaa %>% arrange(desc(n_frac)) %>% filter(n_frac >= 0.3)




syn_list <- lapply(aaa$bodyid, function(x) {
  neuprint_get_synapses(x) %>%
    mutate(prepost = as.factor(ifelse(prepost == 1, "input", "output")))
})






# assumes: syn_list (list of synapse dfs), aaa$bodyid (vector of bodyids), mba (optional)

library(dplyr)
library(ggplot2)
library(patchwork)
library(grid)

ids <- aaa$bodyid
stopifnot(length(syn_list) == length(ids))

# bind all synapses with bodyid (+ optional type)
syn_all <- bind_rows(
  lapply(seq_along(syn_list), function(i) mutate(syn_list[[i]], bodyid = ids[i]))
)
if (exists("mba")) {
  syn_all <- syn_all %>% left_join(select(mba, bodyid, type), by = "bodyid")
}

syn_all <- syn_all %>%
  mutate(
    prepost = factor(prepost, levels = c("input", "output"))
  ) %>% 
  left_join(aaa %>% distinct(type,.keep_all = TRUE)%>% select(type,n_frac),by="type")

base_theme <- theme_minimal() +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())



#scatter plot
p_xy <- ggplot(syn_all, aes(x = x, y = z, color = prepost, alpha = n_frac)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = c(input = "blue", output = "red")) +
  scale_alpha_continuous(range = c(0.0001, 0.01)) +
  scale_y_reverse() +
  coord_fixed() +
  labs(title = "All synapses (XY)") +
  base_theme +
  guides(
    colour = guide_legend(override.aes = list(alpha = 1, size = 2)),
    alpha  = guide_legend(title = "n_frac")
  )

p_xz <- ggplot(syn_all, aes(x = x, y = y, color = prepost, alpha = n_frac)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = c(input = "blue", output = "red")) +
  scale_alpha_continuous(range = c(0.0001, 0.01)) +
  scale_y_reverse() +
  coord_fixed() +
  labs(title = "All synapses (XZ)") +
  base_theme +
  guides(
    colour = guide_legend(override.aes = list(alpha = 1, size = 2)),
    alpha  = guide_legend(title = "n_frac")
  )

combined <- p_xy / p_xz + plot_layout(nrow = 2, heights = c(1, 1))
g <- grid.grabExpr(print(combined))


ggsave(
  file.path(getwd(),'output.synapses.exploration.wo.vAB3.and.PPN1','scatter',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  g,
  width = 7, height = 7, dpi = 900, bg = "white"
)



#heatmap

library(patchwork)
library(dplyr)

# compute common limits
common_limits <- range(
  ggplot_build(
    ggplot(syn_all, aes(x = x, y = y)) + stat_bin2d(bins = 100)
  )$data[[1]]$count,
  na.rm = TRUE
)

# define base heatmap function
heat_map_xz <- function(df, title, show_legend = TRUE) {
  ggplot(df, aes(x = x, y = y)) +
    stat_bin2d(bins = 100, drop = TRUE) +
    scale_fill_viridis_c(
      name = "count",
      trans = "log10",
      limits = common_limits
    ) +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = title, x = "x", y = "z") +
    theme_bw() +
    theme(legend.position = if (show_legend) "right" else "none",
          axis.text = element_blank(),
          axis.ticks = element_blank())
}

heat_map_xy <- function(df, title, show_legend = TRUE) {
  ggplot(df, aes(x = x, y = z)) +
    stat_bin2d(bins = 150, drop = TRUE) +
    scale_fill_viridis_c(
      name = "count",
      trans = "log10",
      limits = common_limits
    ) +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = title, x = "x", y = "y") +
    theme_bw() +
    theme(legend.position = if (show_legend) "right" else "none",
          axis.text = element_blank(),
          axis.ticks = element_blank())
}

p_in  <- heat_map_xz(filter(syn_all, prepost == "input"),  "Input",  FALSE)
p_out <- heat_map_xz(filter(syn_all, prepost == "output"), "Output", FALSE)
p_all <- heat_map_xz(syn_all,                              "All synapses", TRUE)
p_in_xy  <- heat_map_xy(filter(syn_all, prepost == "input"),  "Input",  FALSE)
p_out_xy <- heat_map_xy(filter(syn_all, prepost == "output"), "Output", FALSE)
p_all_xy <- heat_map_xy(syn_all,                              "All synapses", TRUE)



library(cowplot)

# one legend only
legend <- get_legend(p_all + theme(legend.position = "right"))
p_in  <- p_in  + theme(legend.position = "none")
p_out <- p_out + theme(legend.position = "none")
p_all <- p_all + theme(legend.position = "none")
p_in_xy  <- p_in_xy  + theme(legend.position = "none")
p_out_xy <- p_out_xy + theme(legend.position = "none")
p_all_xy <- p_all_xy + theme(legend.position = "none")

r1 <- plot_grid(p_in, p_out, p_all, nrow = 1, rel_widths = c(1,1,1), align = "h", axis = "tb")
r2 <- plot_grid(p_in_xy, p_out_xy, p_all_xy, nrow = 1, rel_widths = c(1.2,1.2,1.2), align = "h", axis = "tb")
grid <- plot_grid(r1, r2, ncol = 1, rel_heights = c(1,1))

ggg <- plot_grid(grid, legend, ncol = 2, rel_widths = c(1, 0.12))

ggsave(
  file.path(getwd(),'output.synapses.exploration.wo.vAB3.and.PPN1','heatmap',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  ggg,
  width = 7, height = 7, dpi = 900, bg = "white"
)

#Density

density_xz <- ggplot(syn_all, aes(x = x, y = y)) +
    stat_density_2d(aes(fill = after_stat(level)), geom = "polygon", bins = 12, alpha = 0.6) +
    facet_wrap(~ prepost, ncol = 2) +
    scale_fill_viridis_c(name = "density level") +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = "Density (XZ)", x = "x", y = "z") +
    theme_bw()+
      theme(axis.text = element_blank(),
          axis.ticks = element_blank())

density_xy <- ggplot(syn_all, aes(x = x, y = z)) +
    stat_density_2d(aes(fill = after_stat(level)), geom = "polygon", bins = 12, alpha = 0.6) +
    facet_wrap(~ prepost, ncol = 2) +
    scale_fill_viridis_c(name = "density level") +
    coord_fixed() +
    scale_y_reverse() +
    labs(title = "Density (XZ)", x = "x", y = "y") +
    theme_bw()+
    theme(axis.text = element_blank(),
          axis.ticks = element_blank())

combined_density <- density_xy + density_xz

combined_density

ggsave(
  file.path(getwd(),'output.synapses.exploration.wo.vAB3.and.PPN1','density',
             paste0(path, "_", valence, "_", modality, "_most common neuron in path synapses.png")),
  combined_density,
  width = 7, height = 7, dpi = 900, bg = "white"
)
}}}

```

