---
title: "aspg_targets_exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 6
)

# ============================================================================
# Library imports (consolidated from throughout the file)
# ============================================================================

# Data manipulation
library(arrow)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(stringr)
library(reshape2)

# Visualization
library(ggplot2)
library(ggpubr)
library(ggraph)
library(ggridges)
library(cowplot)
library(grid)
library(patchwork)
library(plotly)
library(pheatmap)
library(RColorBrewer)
library(viridisLite)
library(scales)
library(nat.ggplot)

# Network analysis
library(igraph)
library(tidygraph)
library(RCy3)

# Neuroscience-specific
library(coconatfly)
library(fafbseg)
library(malecns)
library(neuprintr)

# Statistical/ML
library(Rtsne)
library(dbscan)

# Other
library(Matrix)
library(jsonlite)

# ============================================================================
# Source custom functions from R/ directory
# ============================================================================
source("R/utils.R")
source("R/data_processing.R")
source("R/path_analysis.R")
source("R/visualization.R")
source("R/neuroglancer.R")
neuprint_login(server='https://neuprint-cns.janelia.org')
```


## 1. Data Loading and Preprocessing

### 1.1 Define save paths
```{r}
plot.path.png <- "/Users/fkampf/Documents/pheromone.paper/analyses/strongest_path_closer_evaluation/plots/mcns.strongest.path.closer.evaluation.hierarchical/png/"
plot.path.pdf <- "/Users/fkampf/Documents/pheromone.paper/analyses/strongest_path_closer_evaluation/plots/mcns.strongest.path.closer.evaluation.hierarchical/pdf/"
```


### 1.2 Load all data (mba, connectivity, adjacency matrices, graph)

All data loading is now handled by functions in `R/data_processing.R`:
- `load_mba()` - Load body annotations with ppk region suffixes
- `load_connectivity()` - Load connectivity data
- `load_adjacency_matrices()` - Load all three adjacency matrices (post, pre, raw)
- `load_graph()` - Load igraph with log weights
- `load_all_data()` - Complete pipeline in one call

```{r load-data}
choose_mcns()

# Load everything at once
data <- load_all_data(
  cache_dir = "/Users/fkampf/Documents/pheromone.paper/feather",
  force_recompute = FALSE,
  set_ppk25_glutamate = TRUE
)

# Extract to global environment for compatibility with rest of notebook
mba <- data$mba
conn <- data$conn
conn_save <- conn
adj.matrix <- data$adj.matrix
adj.matrix.pre <- data$adj.matrix.pre
adj.matrix.raw <- data$adj.matrix.raw
graph.general <- data$graph
```


```{r}
aspg.ids <- mba %>% filter(grepl('aSP-g',synonyms)) %>% pull(bodyid) %>% unique()
aspg.types <- mba %>% filter(grepl('aSP-g',synonyms)) %>% pull(type) %>% unique()



start.types.DA1 <- c("ORN_DA1")
start.types.VA1v <- c("ORN_VA1v")
start.types.VA1d <- c("ORN_VA1d")
start.types.aud <- "JO-B"
start.types.vis <- mba %>%
  filter(grepl("LC10", type)) %>%
  pull(type) %>%
  unique()

start.types.ppk23 <- mba %>%
  filter(grepl("ppk23", receptor_type)) %>%
  pull(type) %>%
  unique()
start.types.ppk23[grepl("Pro", start.types.ppk23)] -> start.types.ppk23.pro

start.types.ppk25 <- mba %>%
  filter(grepl("ppk25", receptor_type)) %>%
  pull(type) %>%
  unique()
start.types.ppk25[grepl("Pro", start.types.ppk25)] -> start.types.ppk25.pro

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.DA1.2.aspg.feather")) {
  res.aspg.DA1 <- find_k_strongest_paths_yen(
    graph.general, 
    starts = start.types.DA1, 
    targets = aspg.types, 
    n_paths = 50, 
    mba = mba)
  res.aspg.DA1$df$dataset <- "aspg"
  strongest.50.paths.DA1.2.aspg <- res.aspg.DA1$df
  write_feather(strongest.50.paths.DA1.2.aspg, "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.DA1.2.aspg.feather")
} else {
  strongest.50.paths.DA1.2.aspg <- read_feather("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.DA1.2.aspg.feather")
}

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1v.2.aspg.feather")) {
  res.aspg.VA1v <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.VA1v,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.VA1v$df$dataset <- "aspg"
  strongest.50.paths.VA1v.2.aspg <- res.aspg.VA1v$df
  write_feather(
    strongest.50.paths.VA1v.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1v.2.aspg.feather"
  )
} else {
  strongest.50.paths.VA1v.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1v.2.aspg.feather"
  )
}

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1d.2.aspg.feather")) {
  res.aspg.VA1d <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.VA1d,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.VA1d$df$dataset <- "aspg"
  strongest.50.paths.VA1d.2.aspg <- res.aspg.VA1d$df
  write_feather(
    strongest.50.paths.VA1d.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1d.2.aspg.feather"
  )
} else {
  strongest.50.paths.VA1d.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.VA1d.2.aspg.feather"
  )
}

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.aud.2.aspg.feather")) {
  res.aspg.aud <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.aud,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.aud$df$dataset <- "aspg"
  strongest.50.paths.aud.2.aspg <- res.aspg.aud$df
  write_feather(
    strongest.50.paths.aud.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.aud.2.aspg.feather"
  )
} else {
  strongest.50.paths.aud.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.aud.2.aspg.feather"
  )
}

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.vis.2.aspg.feather")) {
  res.aspg.vis <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.vis,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.vis$df$dataset <- "aspg"
  strongest.50.paths.vis.2.aspg <- res.aspg.vis$df
  write_feather(
    strongest.50.paths.vis.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.vis.2.aspg.feather"
  )
} else {
  strongest.50.paths.vis.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.vis.2.aspg.feather"
  )
}

if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk23.2.aspg.feather")) {
  res.aspg.ppk23 <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.ppk23,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.ppk23$df$dataset <- "aspg"
  strongest.50.paths.ppk23.2.aspg <- res.aspg.ppk23$df
  write_feather(
    strongest.50.paths.ppk23.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk23.2.aspg.feather"
  )
} else {
  strongest.50.paths.ppk23.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk23.2.aspg.feather"
  )
}
if (load_or_execute("/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk25.2.aspg.feather")) {
  res.aspg.ppk25 <- find_k_strongest_paths_yen(
    graph.general,
    starts   = start.types.ppk25,
    targets  = aspg.types,
    n_paths  = 50,
    mba      = mba
  )
  res.aspg.ppk25$df$dataset <- "aspg"
  strongest.50.paths.ppk25.2.aspg <- res.aspg.ppk25$df
  write_feather(
    strongest.50.paths.ppk25.2.aspg,
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk25.2.aspg.feather"
  )
} else {
  strongest.50.paths.ppk25.2.aspg <- read_feather(
    "/Users/fkampf/Documents/pheromone.paper/feather/strongest.paths/strongest.50.paths.ppk25.2.aspg.feather"
  )
}


```


```{r}
# Helper to drop problematic arrow_list columns before binding
drop_list_cols <- function(df) {
  df %>% dplyr::select(-nodes, -nts, -inhib)
}

strongest.50.paths.all.aspg <- dplyr::bind_rows(
  drop_list_cols(strongest.50.paths.DA1.2.aspg)   %>% dplyr::mutate(modality = "DA1"),
  drop_list_cols(strongest.50.paths.VA1v.2.aspg)  %>% dplyr::mutate(modality = "VA1v"),
  drop_list_cols(strongest.50.paths.VA1d.2.aspg)  %>% dplyr::mutate(modality = "VA1d"),
  drop_list_cols(strongest.50.paths.aud.2.aspg)   %>% dplyr::mutate(modality = "aud"),
  drop_list_cols(strongest.50.paths.vis.2.aspg)   %>% dplyr::mutate(modality = "vis"),
  drop_list_cols(strongest.50.paths.ppk23.2.aspg) %>% dplyr::mutate(modality = "ppk23"),
  drop_list_cols(strongest.50.paths.ppk25.2.aspg) %>% dplyr::mutate(modality = "ppk25")
)
```


```{r}
# Sum strength per modality per end neuron
strength_summary <- strongest.50.paths.all.aspg %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop")

ggplot(strength_summary, aes(x = end, y = total_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Total Strength",
    fill = "Modality",
    title = "Sum of Path Strength per Modality per Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Get total raw synapse input per target type from adjacency matrix
total_input_raw <- colSums(adj.matrix.raw, na.rm = TRUE)
total_input_df <- tibble(
  end = names(total_input_raw),
  total_raw_input = as.numeric(total_input_raw)
)

# Normalize by total raw synapse input to each neuron type
strength_summary_normalized <- strength_summary %>%
  dplyr::left_join(total_input_df, by = "end") %>%
  dplyr::mutate(normalized_strength = total_strength / total_raw_input)

ggplot(strength_summary_normalized, aes(x = end, y = normalized_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Total Raw Synapse Input",
    fill = "Modality",
    title = "Path Strength per Modality Normalized by Total Raw Synapse Input to Each asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Normalize so modalities sum to 1 per end neuron
strength_normalized <- strongest.50.paths.all.aspg %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(
    total_strength = sum(strength, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(end) %>%
  dplyr::mutate(
    proportion = total_strength / sum(total_strength)
  ) %>%
  dplyr::ungroup()

ggplot(strength_normalized, aes(x = end, y = proportion, fill = modality)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Input",
    fill = "Modality",
    title = "Distribution of Input per Modality to Each Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# ============================================================================
# EXCITATORY PATHS ONLY
# ============================================================================

# Filter for excitatory paths
excitatory_paths <- strongest.50.paths.all.aspg %>%
  dplyr::filter(valence == "excitatory")

# Sum strength per modality per end neuron (excitatory only)
strength_summary_exc <- excitatory_paths %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop")

ggplot(strength_summary_exc, aes(x = end, y = total_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Total Strength",
    fill = "Modality",
    title = "Sum of EXCITATORY Path Strength per Modality per Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Normalized by total raw synapse input (excitatory only)
strength_summary_exc_normalized <- strength_summary_exc %>%
  dplyr::left_join(total_input_df, by = "end") %>%
  dplyr::mutate(normalized_strength = total_strength / total_raw_input)

ggplot(strength_summary_exc_normalized, aes(x = end, y = normalized_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Total Raw Synapse Input",
    fill = "Modality",
    title = "EXCITATORY Path Strength Normalized by Total Raw Synapse Input"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Stacked proportion (excitatory only)
strength_normalized_exc <- excitatory_paths %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  dplyr::group_by(end) %>%
  dplyr::mutate(proportion = total_strength / sum(total_strength)) %>%
  dplyr::ungroup()

ggplot(strength_normalized_exc, aes(x = end, y = proportion, fill = modality)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Input",
    fill = "Modality",
    title = "Distribution of EXCITATORY Input per Modality to Each Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# ============================================================================
# INHIBITORY PATHS ONLY
# ============================================================================

# Filter for inhibitory paths
inhibitory_paths <- strongest.50.paths.all.aspg %>%
  dplyr::filter(valence == "inhibitory")

# Sum strength per modality per end neuron (inhibitory only)
strength_summary_inh <- inhibitory_paths %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop")

ggplot(strength_summary_inh, aes(x = end, y = total_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Total Strength",
    fill = "Modality",
    title = "Sum of INHIBITORY Path Strength per Modality per Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Normalized by total raw synapse input (inhibitory only)
strength_summary_inh_normalized <- strength_summary_inh %>%
  dplyr::left_join(total_input_df, by = "end") %>%
  dplyr::mutate(normalized_strength = total_strength / total_raw_input)

ggplot(strength_summary_inh_normalized, aes(x = end, y = normalized_strength, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Total Raw Synapse Input",
    fill = "Modality",
    title = "INHIBITORY Path Strength Normalized by Total Raw Synapse Input"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# Stacked proportion (inhibitory only)
strength_normalized_inh <- inhibitory_paths %>%
  dplyr::group_by(modality, end) %>%
  dplyr::summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  dplyr::group_by(end) %>%
  dplyr::mutate(proportion = total_strength / sum(total_strength)) %>%
  dplyr::ungroup()

ggplot(strength_normalized_inh, aes(x = end, y = proportion, fill = modality)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Proportion of Input",
    fill = "Modality",
    title = "Distribution of INHIBITORY Input per Modality to Each Target asp-g Neuron"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# ============================================================================
# NORMALIZED BY NUMBER OF CELLS PER TYPE
# ============================================================================

# Count unique cells per type from mba
cells_per_type <- mba %>%
  dplyr::group_by(type) %>%
  dplyr::summarise(n_cells = n_distinct(bodyid), .groups = "drop") %>%
  dplyr::rename(end = type)

# --- ALL PATHS ---
strength_summary_per_cell <- strength_summary %>%
  dplyr::left_join(cells_per_type, by = "end") %>%
  dplyr::mutate(strength_per_cell = total_strength / n_cells)

ggplot(strength_summary_per_cell, aes(x = end, y = strength_per_cell, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Strength per Cell",
    fill = "Modality",
    title = "Path Strength per Modality Normalized by Number of Cells per Type"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# --- EXCITATORY PATHS (per cell) ---
strength_summary_exc_per_cell <- strength_summary_exc %>%
  dplyr::left_join(cells_per_type, by = "end") %>%
  dplyr::mutate(strength_per_cell = total_strength / n_cells)

ggplot(strength_summary_exc_per_cell, aes(x = end, y = strength_per_cell, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Strength per Cell",
    fill = "Modality",
    title = "EXCITATORY Path Strength Normalized by Number of Cells per Type"
  ) +
  scale_fill_brewer(palette = "Set2")
```


```{r}
# --- INHIBITORY PATHS (per cell) ---
strength_summary_inh_per_cell <- strength_summary_inh %>%
  dplyr::left_join(cells_per_type, by = "end") %>%
  dplyr::mutate(strength_per_cell = total_strength / n_cells)

ggplot(strength_summary_inh_per_cell, aes(x = end, y = strength_per_cell, fill = modality)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Target Neuron",
    y = "Strength per Cell",
    fill = "Modality",
    title = "INHIBITORY Path Strength Normalized by Number of Cells per Type"
  ) +
  scale_fill_brewer(palette = "Set2")
```

