---
title: "Figure 5A — Supplementary: path-count confound checks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

## Why this supplement exists

Panel A plots the **sum of net path strengths** from each sensory modality to each mAL subtype. The cache stores different numbers of paths per modality:

| modality | paths / (modality, target) |
|---|---|
| DA1, VA1v, VA1d, auditory | 50 |
| visual | ~350 |
| ppk23, ppk25 | ~500 |

ppk and visual aggregate ~10× more paths per pair than DA1 simply because those modalities have more source neurons (each source contributes its own strongest-path set). This risks inflating ppk/visual values purely through sampling — not biology.

This supplement runs four alternative aggregation schemes and compares each re-clustered heatmap to the raw panel A. If the story survives, the main figure holds; if it doesn't, the main figure needs a methods-section caveat or a switch to one of these normalizations.

```{r out-dirs}
supp_png <- panel_plot_dir("panel_A", "png", "supp")
supp_pdf <- panel_plot_dir("panel_A", "pdf", "supp")
supp_txt <- panel_plot_dir("panel_A", "txt", "supp")
cat("Output dirs:\n  ", supp_png, "\n  ", supp_pdf, "\n")
```

```{r restrict-paths-to-active-subtypes}
# setup_mspecific.R shrinks mal_subtypes to male-specific only but does NOT
# filter mal_paths — so aggregations over mal_paths would still include
# non-male-specific targets. Restrict paths to whatever mal_subtypes is now.
n_before <- length(unique(mal_paths$end))
mal_paths <- mal_paths %>% filter(end %in% mal_subtypes)
n_after  <- length(unique(mal_paths$end))
cat(sprintf("mal_paths restricted: %d -> %d distinct target subtypes\n",
            n_before, n_after))
cat("Active mal_subtypes (", length(mal_subtypes), "):\n", sep = "")
cat(" ", paste(mal_subtypes, collapse = ", "), "\n")
```

## Diagnostic 0 — tail mass of the path-strength distribution

How much of each modality's total path-strength lives in the top-50-per-pair head vs the tail? If top-50 ≥ 90% the count-inflation concern is cosmetic; if top-50 ≤ 70% the current panel A is leaning on redundant sampling.

```{r tail-mass}
tail_mass <- mal_paths %>%
  group_by(modality, end) %>%
  arrange(desc(strength)) %>%
  mutate(rank = row_number()) %>%
  summarise(
    top50_sum = sum(strength[rank <= 50]),
    total_sum = sum(strength),
    n_paths   = n(),
    .groups = "drop"
  ) %>%
  group_by(modality) %>%
  summarise(
    top50_frac = sum(top50_sum) / sum(total_sum),
    mean_n_paths = mean(n_paths),
    .groups = "drop"
  ) %>%
  mutate(modality = factor(modality, levels = modality_order))

print(tail_mass)

p_tail <- ggplot(tail_mass, aes(x = modality, y = top50_frac, fill = modality)) +
  geom_col() +
  geom_hline(yintercept = 0.9, linetype = "dashed") +
  geom_hline(yintercept = 0.7, linetype = "dotted") +
  geom_text(aes(label = sprintf("%.2f\n(n=%.0f)", top50_frac, mean_n_paths)),
            vjust = -0.2, size = 3) +
  scale_y_continuous(limits = c(0, 1.05), expand = c(0, 0)) +
  labs(x = NULL, y = "Fraction of total path-strength in top 50/pair",
       title = "Does the top-50 head capture the signal?",
       subtitle = "Dashed = 0.9 (head-dominant), dotted = 0.7 (tail-substantial)") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1))

ggsave(file.path(supp_png, "panel_A_supp_tail_mass.png"), p_tail, width = 7, height = 4, dpi = 300)
ggsave(file.path(supp_pdf, "panel_A_supp_tail_mass.pdf"), p_tail, width = 7, height = 4)
writeLines(capture.output(print(tail_mass)), file.path(supp_txt, "panel_A_supp_tail_mass.txt"))
p_tail
```

## Helper: compute mat from a path set

```{r helpers}
compute_activation_matrix <- function(paths_df, cell_scale = NULL) {
  exc <- paths_df %>%
    filter(valence == "excitatory") %>%
    group_by(end, modality) %>%
    summarise(total = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = modality, values_from = total, values_fill = 0) %>%
    column_to_rownames("end") %>% as.matrix()
  inh <- paths_df %>%
    filter(valence == "inhibitory") %>%
    group_by(end, modality) %>%
    summarise(total = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = modality, values_from = total, values_fill = 0) %>%
    column_to_rownames("end") %>% as.matrix()

  # Ensure both matrices align on all mAL types x all modalities
  all_rows <- mal_subtypes
  all_cols <- intersect(modality_order, union(colnames(exc), colnames(inh)))
  fill <- function(m) {
    missing_r <- setdiff(all_rows, rownames(m))
    if (length(missing_r)) m <- rbind(m, matrix(0, length(missing_r), ncol(m),
                                                dimnames = list(missing_r, colnames(m))))
    missing_c <- setdiff(all_cols, colnames(m))
    if (length(missing_c)) m <- cbind(m, matrix(0, nrow(m), length(missing_c),
                                                dimnames = list(rownames(m), missing_c)))
    m[all_rows, all_cols, drop = FALSE]
  }
  mat <- fill(exc) - fill(inh)

  if (!is.null(cell_scale)) {
    # cell_scale is a named vector keyed by modality (applied per column)
    # OR a matrix of the same shape (applied elementwise)
    if (is.matrix(cell_scale)) {
      mat <- mat / cell_scale[rownames(mat), colnames(mat)]
      mat[!is.finite(mat)] <- 0
    } else {
      for (m in names(cell_scale)) if (m %in% colnames(mat)) {
        mat[, m] <- mat[, m] / cell_scale[m]
      }
    }
  }
  mat
}

plot_supp_heatmap <- function(mat, name, subtitle) {
  row_annot <- mal_annot[rownames(mat), c("neurotransmitter", "dimorphism"), drop = FALSE]
  col_annot <- data.frame(
    modality_group = modality_group[colnames(mat)],
    row.names = colnames(mat)
  )
  ann_colors <- list(
    neurotransmitter = nt_colors,
    dimorphism       = dimorphism_colors,
    modality_group   = modality_group_colors
  )
  n_colors <- 100
  max_abs <- max(abs(mat))
  if (max_abs == 0) max_abs <- 1
  breaks <- seq(-max_abs, max_abs, length.out = n_colors + 1)
  colors <- colorRampPalette(c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                                "white",
                                "#FDDBC7", "#F4A582", "#D6604D", "#B2182B"))(n_colors)

  png_file <- file.path(supp_png, paste0(name, ".png"))
  pdf_file <- file.path(supp_pdf, paste0(name, ".pdf"))

  pheatmap::pheatmap(
    mat,
    annotation_row = row_annot,
    annotation_col = col_annot,
    annotation_colors = ann_colors,
    breaks = breaks,
    color = colors,
    cluster_rows = TRUE,
    cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    fontsize_row = 6,
    fontsize_col = 10,
    border_color = NA,
    cellwidth = 30,
    cellheight = 14,
    gaps_col = c(3, 4, 5),
    main = subtitle,
    filename = png_file, width = 8, height = 10
  )
  pheatmap::pheatmap(
    mat,
    annotation_row = row_annot,
    annotation_col = col_annot,
    annotation_colors = ann_colors,
    breaks = breaks,
    color = colors,
    cluster_rows = TRUE,
    cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    fontsize_row = 6,
    fontsize_col = 10,
    border_color = NA,
    cellwidth = 30,
    cellheight = 14,
    gaps_col = c(3, 4, 5),
    main = subtitle,
    filename = pdf_file, width = 8, height = 10
  )
  cat("Saved:", name, "  range:", sprintf("[%.4f, %.4f]", min(mat), max(mat)), "\n")
  invisible(mat)
}
```

## Check 1 — cap at top-50 paths per (modality, target) pair

Equalizes the number of paths summed across modalities. If DA1 has 50 paths and ppk23 has 500, take only the top 50 by strength for each.

```{r check1}
paths_top50 <- mal_paths %>%
  group_by(end, modality) %>%
  slice_max(order_by = strength, n = 50, with_ties = FALSE) %>%
  ungroup()

cat("Path counts after cap:\n")
paths_top50 %>% group_by(modality) %>% summarise(n = n()) %>% print()

mat_c1 <- compute_activation_matrix(paths_top50)
plot_supp_heatmap(mat_c1, "panel_A_supp_check1_top50_per_pair",
                  "Check 1: top-50 paths per (modality, target) pair")
```

## Check 2 — normalize by number of source neurons per modality

Each modality has a different number of source sensory neurons (ppk23 has many sensilla; DA1 has 1 glomerulus-equivalent aggregate). Dividing each column by the source-neuron count puts every channel on a per-source-neuron footing.

```{r check2}
n_sources <- mal_paths %>%
  group_by(modality) %>%
  summarise(n_sources = n_distinct(start), .groups = "drop") %>%
  { setNames(.$n_sources, .$modality) }

cat("Source-neuron counts per modality:\n"); print(n_sources)

mat_c2 <- compute_activation_matrix(mal_paths, cell_scale = n_sources)
plot_supp_heatmap(mat_c2, "panel_A_supp_check2_per_source_neuron",
                  "Check 2: divide by n source neurons per modality")
```

## Check 3 — normalize by total paths per modality

Divide each column by the total number of paths aggregated across all mAL targets for that modality. Compares modalities on a "strength per available path" basis.

```{r check3}
n_paths_mod <- mal_paths %>%
  group_by(modality) %>%
  summarise(n_paths = n(), .groups = "drop") %>%
  { setNames(.$n_paths, .$modality) }

cat("Total paths per modality:\n"); print(n_paths_mod)

mat_c3 <- compute_activation_matrix(mal_paths, cell_scale = n_paths_mod)
plot_supp_heatmap(mat_c3, "panel_A_supp_check3_per_total_paths",
                  "Check 3: divide by total paths per modality")
```

## Check 5 — simulate upstream source aggregation (route-deduplication)

Mimics what would happen if each modality had a **single aggregate source node** (as DA1 does with `ORN_DA1`) instead of 10 separate sensilla. We strip the source from each path, dedupe routes that share an identical downstream sequence across sensilla (keeping the strongest), and then cap at top-50 per (modality, target). Two sensilla that both reach a target via the same intermediate neurons no longer double-count.

```{r check5}
paths_with_route <- mal_paths %>%
  mutate(downstream = sub("^[^>]+-> ", "", path))  # drop "<source> -> "

paths_dedup <- paths_with_route %>%
  group_by(end, modality, valence, downstream) %>%
  slice_max(order_by = strength, n = 1, with_ties = FALSE) %>%
  ungroup()

paths_merged <- paths_dedup %>%
  group_by(end, modality) %>%
  slice_max(order_by = strength, n = 50, with_ties = FALSE) %>%
  ungroup()

cat("Path counts per modality after source-aggregation simulation:\n")
paths_merged %>% group_by(modality) %>% summarise(n = n()) %>% print()

mat_c5 <- compute_activation_matrix(paths_merged)
plot_supp_heatmap(mat_c5, "panel_A_supp_check5_source_merge",
                  "Check 5: simulated upstream source merge (route-dedup + top-50)")
```

## Check 4 — normalize by paths per (modality, target) pair (→ mean path strength)

Divide each cell by the number of paths summed INTO that cell. Converts the sum to a mean path strength — directly comparable across modalities regardless of how many paths contributed.

```{r check4}
n_paths_pair <- mal_paths %>%
  group_by(end, modality) %>%
  summarise(n_paths = n(), .groups = "drop") %>%
  pivot_wider(names_from = modality, values_from = n_paths, values_fill = 0) %>%
  column_to_rownames("end") %>% as.matrix()
# align shape
n_paths_pair_full <- matrix(0, length(mal_subtypes), length(modality_order),
                            dimnames = list(mal_subtypes, modality_order))
shared_r <- intersect(rownames(n_paths_pair), rownames(n_paths_pair_full))
shared_c <- intersect(colnames(n_paths_pair), colnames(n_paths_pair_full))
n_paths_pair_full[shared_r, shared_c] <- n_paths_pair[shared_r, shared_c]

# Replace zeros with NA so division doesn't blow up; cells with no paths stay 0 below
scale_mat <- n_paths_pair_full
scale_mat[scale_mat == 0] <- NA_real_

mat_c4 <- compute_activation_matrix(mal_paths, cell_scale = scale_mat)
plot_supp_heatmap(mat_c4, "panel_A_supp_check4_mean_path_strength",
                  "Check 4: per-pair mean path strength")
```

## Cross-method comparison

Column-means of each matrix give a coarse "how strong does this channel look?" ranking under each scheme. If the rankings agree qualitatively, the main panel A is robust. If they disagree, the normalization choice is doing heavy lifting.

```{r comparison}
channel_summary <- bind_rows(
  tibble(method = "raw (panel A)",                modality = colnames(mal_activation_matrix),
         mean_abs = colMeans(abs(mal_activation_matrix))),
  tibble(method = "C1: top-50/pair cap",          modality = colnames(mat_c1),
         mean_abs = colMeans(abs(mat_c1))),
  tibble(method = "C2: per source neuron",        modality = colnames(mat_c2),
         mean_abs = colMeans(abs(mat_c2))),
  tibble(method = "C3: per total paths",          modality = colnames(mat_c3),
         mean_abs = colMeans(abs(mat_c3))),
  tibble(method = "C4: per-pair mean strength",   modality = colnames(mat_c4),
         mean_abs = colMeans(abs(mat_c4))),
  tibble(method = "C5: source-merge simulation",  modality = colnames(mat_c5),
         mean_abs = colMeans(abs(mat_c5)))
) %>%
  group_by(method) %>%
  mutate(mean_abs_rel = mean_abs / max(mean_abs)) %>%
  ungroup() %>%
  mutate(modality = factor(modality, levels = modality_order),
         method   = factor(method, levels = c("raw (panel A)", "C1: top-50/pair cap",
                                              "C2: per source neuron",
                                              "C3: per total paths",
                                              "C4: per-pair mean strength",
                                              "C5: source-merge simulation")))

p_cmp <- ggplot(channel_summary, aes(x = modality, y = mean_abs_rel, fill = method)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.8) +
  labs(x = NULL, y = "Mean |cell| per channel, normalized to each method's max",
       title = "Channel-strength ranking under different aggregation schemes",
       subtitle = "If all bars within a modality are similar, that channel's ranking is robust.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank())

ggsave(file.path(supp_png, "panel_A_supp_method_comparison.png"), p_cmp, width = 9, height = 5, dpi = 300)
ggsave(file.path(supp_pdf, "panel_A_supp_method_comparison.pdf"), p_cmp, width = 9, height = 5)

write.csv(channel_summary,
          file.path(supp_txt, "panel_A_supp_channel_summary.csv"),
          row.names = FALSE)

p_cmp
```

## Summary

```{r narrative, echo=FALSE}
summary_txt <- sprintf(
"Panel A supplement -- path-count confound checks
=================================================

Top-50 fraction per modality (Diagnostic 0):
%s

Interpretation thresholds:
  > 0.9  head-dominant -- count inflation is cosmetic.
  0.7-0.9  partial inflation -- consider reporting Check 1 or 4 alongside.
  < 0.7  tail-substantial -- raw panel A is leaning on redundant sampling.

Heatmaps produced (all in plots/panel_A/supp/):
  - panel_A_supp_check1_top50_per_pair            (equal K across modalities)
  - panel_A_supp_check2_per_source_neuron         (divide by n sensory sources)
  - panel_A_supp_check3_per_total_paths           (divide by total paths / modality)
  - panel_A_supp_check4_mean_path_strength        (per-pair mean strength)

Cross-method comparison bar plot:
  - panel_A_supp_method_comparison

If all four checks preserve the same qualitative pattern (same clustering groups,
same sign pattern for ppk23 sign reversal, same dominance of ppk25 in the
male-specific ensemble), the raw panel A is robust. If Check 1 or 4 reshuffles
the story, the methods section should use the corresponding normalized variant
(or show both).
",
paste(sprintf("  %s: %.3f (mean n per pair %.0f)",
              tail_mass$modality, tail_mass$top50_frac, tail_mass$mean_n_paths),
      collapse = "\n")
)
writeLines(summary_txt, file.path(supp_txt, "panel_A_supp_summary.txt"))
cat(summary_txt)
```
