---
title: "Figure 5L: mAL + P1 ppk selectivity in shared coordinate system"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

## What this is

Two overlay plots that put mAL and P1 subtypes on the same coordinate system,
directly answering "does ppk selectivity survive the synapse from mAL to P1?".

- **Path-based**: x = ppk23 drive, y = ppk25 drive.
  mAL ppk23/ppk25 = columns of `mal_activation_matrix`.
  P1 ppk23/ppk25 = `mal_activation_matrix[, ppk] %*% mal_p1_signed`.
- **Signal-flow**: x = ppk23-alone net input, y = (ppk23+ppk25) co-activation
  net input (same axes as `panel_L_*_sf_ppk_selectivity.pdf`).

Shape/color distinguishes the two populations.

```{r build-path}
p1_subtypes <- unique(p1_all_combos$target_type)

mal_rows_present <- intersect(mal_subtypes, rownames(adj.matrix))
p1_cols_present  <- intersect(p1_subtypes,  colnames(adj.matrix))
mal_p1_signed    <- matrix(0, length(mal_subtypes), length(p1_subtypes),
                           dimnames = list(mal_subtypes, p1_subtypes))
mal_p1_signed[mal_rows_present, p1_cols_present] <-
  as.matrix(adj.matrix[mal_rows_present, p1_cols_present])
sign_mult <- ifelse(mal_nt[mal_subtypes] == "GABA", -1, 1)
mal_p1_signed <- sweep(mal_p1_signed, 1, sign_mult, "*")

mal_path <- tibble(
  type  = mal_subtypes,
  pop   = "mAL",
  ppk23 = mal_activation_matrix[mal_subtypes, "ppk23"],
  ppk25 = mal_activation_matrix[mal_subtypes, "ppk25"]
)
p1_path <- tibble(
  type  = p1_subtypes,
  pop   = "P1",
  ppk23 = as.numeric(mal_activation_matrix[mal_subtypes, "ppk23"] %*% mal_p1_signed),
  ppk25 = as.numeric(mal_activation_matrix[mal_subtypes, "ppk25"] %*% mal_p1_signed)
)
path_df <- bind_rows(mal_path, p1_path) %>%
  filter(abs(ppk23) > 1e-6 | abs(ppk25) > 1e-6) %>%
  mutate(pop = factor(pop, levels = c("mAL", "P1")))

cat("Path-based combined: ", nrow(path_df), "subtypes total\n")
cat("ppk23 range:", range(path_df$ppk23), "\n")
cat("ppk25 range:", range(path_df$ppk25), "\n")
```

## Overlay scatter — path-based

```{r path-scatter, fig.width=9, fig.height=8}
pop_colors <- c("mAL" = "#1F77B4", "P1" = "#D62728")

# Label top-N per population by distance from origin
label_set <- path_df %>%
  mutate(d = sqrt(ppk23^2 + ppk25^2)) %>%
  group_by(pop) %>%
  slice_max(d, n = 12) %>%
  pull(type)

p_path <- ggplot(path_df, aes(x = ppk23, y = ppk25,
                              color = pop, shape = pop)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.3) +
  geom_point(size = 3, alpha = 0.85) +
  ggrepel::geom_text_repel(
    data = path_df %>% filter(type %in% label_set),
    aes(label = type), size = 2.8, color = "grey20",
    max.overlaps = 30, min.segment.length = 0
  ) +
  scale_color_manual(values = pop_colors, name = "Population") +
  scale_shape_manual(values = c("mAL" = 16, "P1" = 17), name = "Population") +
  coord_fixed(ratio = 1) +
  labs(x = "ppk23 path drive",
       y = "ppk25 path drive",
       title = "mAL and P1 ppk selectivity in shared path-based coordinate system") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot")

save_fig5_plot(p_path, "panel_L_mal_and_p1_ppk_selectivity", width = 9, height = 8)
print(p_path)
```

## Signal-flow version

```{r build-sf}
# x = ppk23 alone net_input;  y = ppk23 + ppk25 co-activation net_input.
mal_sf <- tibble(
  type  = rownames(mal_sf_net_input_matrix),
  pop   = "mAL",
  ppk23_sf       = mal_sf_net_input_matrix[, "ppk23"],
  ppk23_ppk25_sf = setNames(
    mal_all_combos %>%
      filter(combination == "ppk25",
             target_type %in% rownames(mal_sf_net_input_matrix)) %>%
      pull(net_input),
    mal_all_combos %>%
      filter(combination == "ppk25",
             target_type %in% rownames(mal_sf_net_input_matrix)) %>%
      pull(target_type)
  )[rownames(mal_sf_net_input_matrix)]
) %>% filter(type %in% mal_subtypes)

p1_sf <- tibble(
  type  = rownames(p1_sf_activation_matrix),
  pop   = "P1",
  ppk23_sf = p1_sf_activation_matrix[, "ppk23"],
  ppk23_ppk25_sf = setNames(
    p1_all_combos %>% filter(combination == "ppk25") %>% pull(net_input),
    p1_all_combos %>% filter(combination == "ppk25") %>% pull(target_type)
  )[rownames(p1_sf_activation_matrix)]
)

sf_df <- bind_rows(mal_sf, p1_sf) %>%
  filter(abs(ppk23_sf) > 1e-4 | abs(ppk23_ppk25_sf) > 1e-4) %>%
  mutate(pop = factor(pop, levels = c("mAL", "P1")))

cat("Signal-flow combined: ", nrow(sf_df), "subtypes total\n")
```

```{r sf-scatter, fig.width=9, fig.height=8}
label_set_sf <- sf_df %>%
  mutate(d = sqrt(ppk23_sf^2 + ppk23_ppk25_sf^2)) %>%
  group_by(pop) %>%
  slice_max(d, n = 12) %>%
  pull(type)

p_sf <- ggplot(sf_df, aes(x = ppk23_sf, y = ppk23_ppk25_sf,
                          color = pop, shape = pop)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.3) +
  geom_point(size = 3, alpha = 0.85) +
  ggrepel::geom_text_repel(
    data = sf_df %>% filter(type %in% label_set_sf),
    aes(label = type), size = 2.8, color = "grey20",
    max.overlaps = 30, min.segment.length = 0
  ) +
  scale_color_manual(values = pop_colors, name = "Population") +
  scale_shape_manual(values = c("mAL" = 16, "P1" = 17), name = "Population") +
  coord_fixed(ratio = 1) +
  labs(x = "ppk23 alone (signal flow net input)",
       y = "ppk25 alone (signal flow net input)",
       title = "mAL and P1 ppk selectivity in shared signal-flow coordinate system") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot")

save_fig5_plot(p_sf, "panel_L_mal_and_p1_sf_ppk_selectivity", width = 9, height = 8)
print(p_sf)
```
