---
title: "Figure 5A: max-flow raw-synapse version"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(pheatmap)
library(igraph)
```

## What this is

Max-flow analog of panel A. For each (modality, mAL subtype) pair, we:

1. Collect all cached K-strongest paths from that modality's sources to that mAL.
2. Split paths by valence (excitatory / inhibitory).
3. Within each valence, induce a subgraph from the UNION of edges in those paths. Edge capacity = raw synapse count (`adj.matrix.raw`).
4. Add a virtual super-source with infinite-capacity edges to every source sensillum of that modality.
5. Compute max-flow from super-source to the target mAL via `igraph::max_flow`.
6. Cell = `max_flow_excitatory - max_flow_inhibitory`.

Unlike the widest-path supp (which picks a single path), this aggregates total parallel signal capacity through the cached route set — shared bottleneck edges are counted once, and multiple independent routes add up.

```{r compute}
parse_path <- function(p) strsplit(p, " -> ", fixed = TRUE)[[1]]

all_edges_for_paths <- function(paths_df) {
  edges <- lapply(paths_df$path, function(p) {
    toks <- parse_path(p)
    if (length(toks) < 2) return(NULL)
    data.frame(from = toks[-length(toks)], to = toks[-1], stringsAsFactors = FALSE)
  })
  bind_rows(edges) %>% distinct(from, to)
}

compute_max_flow <- function(paths_df, sources, target) {
  if (nrow(paths_df) == 0) return(0)
  ed <- all_edges_for_paths(paths_df)
  if (nrow(ed) == 0) return(0)
  ed$raw <- adj.matrix.raw[cbind(ed$from, ed$to)]
  ed <- ed %>% filter(is.finite(raw), raw > 0)
  if (nrow(ed) == 0) return(0)

  # Add super-source edges
  present_sources <- intersect(sources, unique(c(ed$from, ed$to)))
  if (length(present_sources) == 0) return(0)
  ss <- data.frame(from = "__SS__", to = present_sources, raw = 1e9)
  ed_full <- rbind(ed, ss)

  g <- graph_from_data_frame(ed_full, directed = TRUE)
  if (!(target %in% V(g)$name) || !("__SS__" %in% V(g)$name)) return(0)
  flow <- max_flow(g, source = "__SS__", target = target, capacity = E(g)$raw)
  as.numeric(flow$value)
}

# Build exc/inh matrices
modalities <- modality_order
mat_exc <- matrix(0, length(mal_subtypes), length(modalities),
                  dimnames = list(mal_subtypes, modalities))
mat_inh <- mat_exc

cat("Computing max-flow for", length(mal_subtypes) * length(modalities), "pairs...\n")
t0 <- Sys.time()
for (mal in mal_subtypes) {
  for (mod in modalities) {
    srcs <- channel_neurons[[mod]]
    sub <- mal_paths %>% filter(modality == mod, end == mal)
    if (nrow(sub) == 0) next
    mat_exc[mal, mod] <- compute_max_flow(sub %>% filter(valence == "excitatory"),
                                          srcs, mal)
    mat_inh[mal, mod] <- compute_max_flow(sub %>% filter(valence == "inhibitory"),
                                          srcs, mal)
  }
}
cat("done in", as.numeric(difftime(Sys.time(), t0, units = "secs")), "s\n")
cat("exc range:", range(mat_exc), "  inh range:", range(mat_inh), "\n")

mat <- mat_exc - mat_inh
cat("net range:", range(mat), "\n")
```

## Heatmap

```{r heatmap, fig.width=8, fig.height=10}
row_annot <- mal_annot[rownames(mat), c("neurotransmitter", "dimorphism"), drop = FALSE]
col_annot <- data.frame(modality_group = modality_group[colnames(mat)],
                        row.names = colnames(mat))
ann_colors <- list(
  neurotransmitter = nt_colors,
  dimorphism       = dimorphism_colors,
  modality_group   = modality_group_colors
)

max_abs <- max(abs(mat)); if (!is.finite(max_abs) || max_abs == 0) max_abs <- 1
breaks <- seq(-max_abs, max_abs, length.out = 101)
colors <- colorRampPalette(c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                              "white",
                              "#FDDBC7", "#F4A582", "#D6604D", "#B2182B"))(100)

for (ext in c("png", "pdf")) {
  pheatmap(
    mat,
    annotation_row = row_annot, annotation_col = col_annot,
    annotation_colors = ann_colors,
    breaks = breaks, color = colors,
    cluster_rows = TRUE, cluster_cols = FALSE,
    clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
    fontsize_row = 7, fontsize_col = 10,
    border_color = NA, cellwidth = 40, cellheight = 18,
    gaps_col = c(3, 4, 5),
    display_numbers = TRUE, number_format = "%.0f", fontsize_number = 6,
    filename = file.path(panel_plot_dir("panel_A_input_profiles_max_flow", ext),
                         paste0("panel_A_input_profiles_max_flow.", ext)),
    width = 8, height = 10, silent = TRUE
  )
}
cat("saved panel_A_input_profiles_max_flow\n")
```
