---
title: "Figure 5I: P1 mean path vs mean signal-flow (analog of subpanel b)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(patchwork)
```

## What this is

P1 analog of subpanel (b) in `panel_I_scenario_discriminability.pdf`
(= `panel_I_spread_scatter.pdf`). For each P1 subtype we compute its
activation under the three encounter scenarios under BOTH models and plot
the means against each other, with point size encoding the 2D (path, sf)
spread across scenarios.

- x = mean path-based activation across the 3 scenarios
      (path-based drive = sum of signed mAL -> P1 weights applied to the
       scenario's channel drives, as computed in panel J)
- y = mean signal-flow net input across the 3 scenarios (from p1_all_combos)
- size = spread_2d = sqrt(var_across_scenarios(path) + var_across_scenarios(sf))

```{r path-and-sf}
scenarios <- c(
  "Female (ppk25)" = "ppk25",
  "cVA + male (cVA+ppk23)"       = "DA1+ppk23",
  "cVA + female (cVA+ppk25)"     = "DA1+ppk25"
)

# ---------- Signed mAL -> P1 matrix (mirrors panel_J.Rmd construction) ----
p1_subtypes <- unique(p1_all_combos$target_type)

mal_rows_present <- intersect(mal_subtypes,  rownames(adj.matrix))
p1_cols_present  <- intersect(p1_subtypes,   colnames(adj.matrix))
mal_p1_signed    <- matrix(0, length(mal_subtypes), length(p1_subtypes),
                           dimnames = list(mal_subtypes, p1_subtypes))
mal_p1_signed[mal_rows_present, p1_cols_present] <-
  as.matrix(adj.matrix[mal_rows_present, p1_cols_present])
sign_mult <- ifelse(mal_nt[mal_subtypes] == "GABA", -1, 1)
mal_p1_signed <- sweep(mal_p1_signed, 1, sign_mult, "*")

# ---------- Path-based drive: scenario x mAL, then project to P1 --------
path_scenario_mat <- matrix(0, nrow = length(scenarios), ncol = length(mal_subtypes),
                            dimnames = list(names(scenarios), mal_subtypes))
for (i in seq_along(scenarios)) {
  channels <- strsplit(scenarios[i], "\\+")[[1]]
  for (ch in channels) {
    if (ch %in% colnames(mal_activation_matrix)) {
      path_scenario_mat[i, ] <- path_scenario_mat[i, ] +
        mal_activation_matrix[mal_subtypes, ch]
    }
  }
}
p1_path_mat <- path_scenario_mat %*% mal_p1_signed    # scenarios x P1

# ---------- Signal-flow drive ----------------------------------------------
p1_sf_long <- bind_rows(lapply(names(scenarios), function(nm) {
  p1_all_combos %>%
    filter(channels == gsub("\\+", ",", scenarios[[nm]])) %>%
    transmute(P1 = target_type, scenario = nm, sf = net_input)
}))
p1_sf_mat <- p1_sf_long %>%
  pivot_wider(names_from = P1, values_from = sf, values_fill = 0) %>%
  column_to_rownames("scenario") %>% as.matrix()

# Align columns
common_p1 <- intersect(colnames(p1_path_mat), colnames(p1_sf_mat))
p1_path_mat <- p1_path_mat[names(scenarios), common_p1, drop = FALSE]
p1_sf_mat   <- p1_sf_mat[names(scenarios), common_p1, drop = FALSE]

# Per-P1 summaries
summary_df <- tibble(
  P1         = common_p1,
  mean_path  = colMeans(p1_path_mat),
  mean_sf    = colMeans(p1_sf_mat),
  var_path   = apply(p1_path_mat, 2, var),
  var_sf     = apply(p1_sf_mat,   2, var),
  spread_2d  = sqrt(var_path + var_sf)
)

cat("Top 10 P1 subtypes by joint (path, sf) spread:\n")
print(summary_df %>% arrange(desc(spread_2d)) %>% head(10))
```

## Scatter plot

```{r scatter, fig.width=9, fig.height=7}
p <- ggplot(summary_df, aes(x = mean_path, y = mean_sf,
                            size = spread_2d)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(alpha = 0.75, color = "#455A64") +
  ggrepel::geom_text_repel(aes(label = P1), size = 2.6, color = "grey30",
                            max.overlaps = 25, min.segment.length = 0) +
  scale_size_continuous(range = c(1.5, 9), name = "Scenario spread\nsqrt(var_path + var_sf)") +
  labs(x = "Mean path-based P1 drive (across scenarios)",
       y = "Mean signal-flow P1 drive (across scenarios)") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right")

save_fig5_plot(p, "panel_I_p1_spread_scatter", width = 9, height = 7)
print(p)
```

```{r narrative, echo=FALSE}
out <- panel_plot_dir("panel_I", "txt")
writeLines(sprintf(
"P1 mean-path vs mean-signal-flow scatter (subpanel b analog)
=============================================================
For each of %d P1 subtypes, mean activation across the 3 scenarios is
computed under both path-based (scenario_mat x mal_p1_signed) and signal-flow
(p1_all_combos$net_input) models. Point size = sqrt(var_path + var_sf) =
joint spread of the 3-scenario trajectory in 2D.

Top 10 P1s by joint spread:
%s
", nrow(summary_df),
paste(capture.output(print(summary_df %>% arrange(desc(spread_2d)) %>% head(10))),
      collapse = "\n")
), file.path(out, "panel_I_p1_spread_scatter.txt"))
```
