---
title: "Figure 5B supp v3: AN09B017 input composition — annotated quadrants"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(scatterpie)
library(ggplot2)
library(dplyr)
```

## What this is (v3)

Per-mAL pie-chart scatter with interpretive quadrant backgrounds, guide lines, brighter
AN09B017 variant palette, and an inline example annotation.

- **x** = ppk23 - ppk25 path-based drive (male minus female contact). Positive = ppk23-dominant (male-contact), negative = ppk25-dominant (female-contact).
- **y** = % of mAL's total input contributed by all AN09B017 variants combined.
- **pie slice area** = fraction of AN09B017 input from each variant (a-g).
- Horizontal guide at `median(y)` marks the threshold for *strong AN09B017 coupling*.
- Vertical guide at `x = 0` marks *ppk balanced*.

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
# Attach per-variant fractions as columns
for (v in an09_present) {
  df_wide[[v]] <- ifelse(colSums(w) > 0, w[v, ] / colSums(w), 0)
}
print(df_wide)
```

## Pie scatter with quadrant annotations

```{r plot, fig.width=12, fig.height=9}
# --- Brighter, more distinct palette (Dark2 -> Set2 fallback) ----------------
base_pal <- RColorBrewer::brewer.pal(8, "Dark2")
if (length(an09_present) > length(base_pal)) {
  base_pal <- c(base_pal, RColorBrewer::brewer.pal(8, "Set2"))
}
an09_palette <- setNames(base_pal[seq_along(an09_present)], an09_present)

# --- Axis geometry -----------------------------------------------------------
x_pad   <- diff(range(df_wide$delta_ppk)) * 0.08
y_pad   <- diff(range(df_wide$total_pct)) * 0.08
x_min   <- min(df_wide$delta_ppk) - x_pad
x_max   <- max(df_wide$delta_ppk) + x_pad
y_min   <- min(0, min(df_wide$total_pct) - y_pad)
y_max   <- max(df_wide$total_pct) + y_pad
y_med   <- median(df_wide$total_pct)
x_range <- x_max - x_min
y_range <- y_max - y_min
pie_r   <- x_range * 0.035

# --- Quadrant background tiles ----------------------------------------------
quad_df <- tibble(
  xmin = c(0,     x_min, x_min, 0),
  xmax = c(x_max, 0,     0,     x_max),
  ymin = c(y_med, y_med, y_min, y_min),
  ymax = c(y_max, y_max, y_med, y_med),
  fill = c("#FDE9D9", "#D9E8FD", "#F2F2F2", "#F2F2F2")  # TR warm, TL cool, bot grey
)

quad_labels <- tibble(
  x = c(x_max - x_pad * 0.5,
        x_min + x_pad * 0.5,
        x_max - x_pad * 0.5,
        x_min + x_pad * 0.5),
  y = c(y_max - y_pad * 0.2,
        y_max - y_pad * 0.2,
        y_min + y_pad * 0.2,
        y_min + y_pad * 0.2),
  hjust = c(1, 0, 1, 0),
  label = c(
    "male-contact-driven mAL\nstrong AN09B017 input",
    "female-contact-driven mAL\nstrong AN09B017 input",
    "weak AN09B017 involvement\n(ppk23-dominant)",
    "weak AN09B017 involvement\n(ppk25-dominant)"
  ),
  col = c("#B45F06", "#1F4E79", "grey45", "grey45")
)

# --- Identify example a-d dominant dot --------------------------------------
ad_cols <- intersect(c("AN09B017a","AN09B017b","AN09B017c","AN09B017d"), an09_present)
if (length(ad_cols) > 0) {
  df_wide$ad_share <- rowSums(df_wide[, ad_cols, drop = FALSE])
  ex_idx <- which.max(df_wide$ad_share * (df_wide$total_pct > y_med))
  if (length(ex_idx) == 0 || df_wide$ad_share[ex_idx] < 0.5) {
    ex_idx <- which.max(df_wide$ad_share)
  }
  example_row <- df_wide[ex_idx, ]
  ex_label <- sprintf(
    "e.g. %s: %.0f%% of its AN09B017\ninput is from a-d variants",
    example_row$mAL, example_row$ad_share * 100
  )
} else {
  example_row <- df_wide[1, ]
  ex_label <- ""
}

# --- Build plot --------------------------------------------------------------
p <- ggplot() +
  # quadrant tinted backgrounds
  geom_rect(data = quad_df,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = quad_df$fill, alpha = 0.55, inherit.aes = FALSE) +
  # guide lines
  geom_hline(yintercept = y_med,
             linewidth = 0.45, color = "grey40", linetype = "dashed") +
  geom_vline(xintercept = 0,
             linewidth = 0.45, color = "grey40", linetype = "dashed") +
  # inline guide labels
  annotate("text", x = 0, y = y_max - y_pad * 0.05,
           label = "ppk balanced", angle = 90,
           hjust = 1, vjust = -0.4, size = 3, color = "grey30",
           fontface = "italic") +
  annotate("text", x = x_max, y = y_med,
           label = sprintf("median y = %.1f%%  (strong-coupling threshold)", y_med),
           hjust = 1, vjust = -0.4, size = 3, color = "grey30",
           fontface = "italic") +
  # quadrant annotation labels
  geom_text(data = quad_labels,
            aes(x = x, y = y, label = label, hjust = hjust),
            vjust = 1, size = 3.4, fontface = "bold",
            color = quad_labels$col, lineheight = 0.95) +
  # pies
  geom_scatterpie(data = df_wide,
                  aes(x = delta_ppk, y = total_pct, r = pie_r),
                  cols = an09_present,
                  color = "grey25", linewidth = 0.2) +
  coord_equal(ratio = x_range / y_range,
              xlim = c(x_min, x_max), ylim = c(y_min, y_max),
              expand = FALSE) +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  ggrepel::geom_text_repel(data = df_wide,
                           aes(x = delta_ppk, y = total_pct, label = mAL),
                           size = 2.9, color = "grey20",
                           max.overlaps = 30, min.segment.length = 0,
                           box.padding = 0.7, point.padding = pie_r,
                           segment.color = "grey55", segment.size = 0.25) +
  # example interpretation annotation
  annotate("curve",
           x = example_row$delta_ppk + x_range * 0.18,
           y = example_row$total_pct + y_range * 0.12,
           xend = example_row$delta_ppk + pie_r * 1.1,
           yend = example_row$total_pct + pie_r * 0.5,
           curvature = -0.25, arrow = arrow(length = unit(0.15, "cm")),
           color = "grey25", linewidth = 0.35) +
  annotate("label",
           x = example_row$delta_ppk + x_range * 0.18,
           y = example_row$total_pct + y_range * 0.13,
           label = ex_label, hjust = 0, vjust = 0,
           size = 3.1, color = "grey10",
           label.size = 0.25, fill = "white",
           fontface = "italic", lineheight = 0.95) +
  labs(
    x = expression(ppk23 - ppk25 ~ "(male - female contact drive)"),
    y = "Total AN09B017 input to mAL (% of mAL's total input)",
    title = "AN09B017 input composition per mAL vs ppk asymmetry"
  ) +
  theme_pub(base_size = 11) +
  theme(legend.position = "right",
        plot.title.position = "plot",
        panel.grid.major = element_line(linewidth = 0.2, color = "grey92"),
        panel.grid.minor = element_blank())

save_fig5_plot(p, "panel_B_an09_dominance_vs_ppk_asymmetry_v3",
               width = 12, height = 9)

# Also drop the PDF directly into the curated "single panels" folder
single_panels_dir <- file.path(
  fig5_dir,
  "my_selection_that_makes_halfway_sense_mspecific",
  "single panels"
)
dir.create(single_panels_dir, recursive = TRUE, showWarnings = FALSE)
ggsave(file.path(single_panels_dir, "panel_B_an09_dominance_vs_ppk_asymmetry_v3.pdf"),
       plot = p + theme_pub(base_size = 11),
       width = 12, height = 9)

print(p)
```
