---
title: "Figure 5I: mAL scenario discriminability (cVA+male vs cVA+female only)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(patchwork)
```

## What this is

Restricted version of the mAL scenario-discriminability plots with only the
two cVA-containing encounters:

- cVA + male:   DA1 + ppk23
- cVA + female: DA1 + ppk23 + ppk25

Both scenarios share DA1 (cVA) and ppk23 (male contact) — only ppk25 (female
contact) toggles. This isolates the specific question: "given that cVA is in
the air, which mAL subtypes respond differently when the fly additionally
contacts a female?" Spread here = `drive(cVA+female) - drive(cVA+male)` in
absolute value. Argmax coloring distinguishes the two.

```{r drive}
scenarios <- c(
  "cVA + male (cVA+ppk23)"   = "DA1,ppk23",
  "cVA + female (cVA+ppk25)" = "DA1,ppk25"
)
scenario_colors <- c(
  "cVA + male (cVA+ppk23)"   = "#8E24AA",
  "cVA + female (cVA+ppk25)" = "#F57C00"
)

drive_long <- bind_rows(lapply(names(scenarios), function(nm) {
  mal_all_combos %>%
    filter(channels == scenarios[[nm]], target_type %in% mal_subtypes) %>%
    transmute(mAL = target_type, scenario = nm, drive = net_input)
})) %>%
  mutate(scenario = factor(scenario, levels = names(scenarios)))

discrim <- drive_long %>%
  group_by(mAL) %>%
  summarise(
    spread     = max(drive) - min(drive),   # = |cVA+female - cVA+male|
    mean_drive = mean(drive),
    argmax     = scenario[which.max(drive)],
    delta      = drive[scenario == "cVA + female (cVA+ppk25)"] - drive[scenario == "cVA + male (cVA+ppk23)"],
    .groups = "drop"
  ) %>%
  arrange(desc(spread))

cat("Top 10 most ppk25-sensitive mAL subtypes (under cVA background):\n")
print(head(discrim, 10))
cat("\nArgmax counts:\n"); print(discrim %>% count(argmax))
```

## Bar chart: spread (|female+cVA - male+cVA|) per mAL

```{r bar, fig.width=12, fig.height=4}
discrim_ord <- discrim %>% mutate(mAL = factor(mAL, levels = mAL))

p_bar <- ggplot(discrim_ord, aes(x = mAL, y = spread, fill = argmax)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = scenario_colors, name = "Higher-drive scenario") +
  labs(x = "mAL subtype (sorted by |cVA+female - cVA+male|)",
       y = "ppk25 sensitivity\n|cVA+female drive - cVA+male drive|") +
  theme_pub(base_size = 10) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 7),
        legend.position = "top")

save_fig5_plot(p_bar, "panel_I_mal_cva_mvf_bar", width = 12, height = 4)
print(p_bar)
```

## Scatter: mean drive vs spread

```{r scatter, fig.width=8, fig.height=6}
p_scatter <- ggplot(discrim, aes(x = mean_drive, y = spread,
                                 size = spread, color = argmax)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(alpha = 0.85) +
  ggrepel::geom_text_repel(aes(label = mAL), size = 2.6, color = "grey30",
                            max.overlaps = 25, min.segment.length = 0) +
  scale_color_manual(values = scenario_colors,
                     name = "Higher-drive scenario",
                     guide = guide_legend(override.aes = list(size = 4))) +
  scale_size_continuous(range = c(1.5, 8), name = "|delta|") +
  labs(x = "Mean drive across cVA+male and cVA+female",
       y = "|cVA+female drive - cVA+male drive|") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right")

save_fig5_plot(p_scatter, "panel_I_mal_cva_mvf_scatter", width = 8, height = 6)
print(p_scatter)
```

## Trajectories: top-15 most ppk25-sensitive subtypes

```{r traj, fig.width=10, fig.height=6}
top_n <- 15
top_mal <- head(discrim$mAL, top_n)
traj <- drive_long %>% filter(mAL %in% top_mal) %>%
  mutate(mAL = factor(mAL, levels = top_mal))

p_traj <- ggplot(traj, aes(x = scenario, y = drive, group = mAL, color = mAL)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_line(linewidth = 0.75, alpha = 0.85) +
  geom_point(size = 2.2) +
  ggrepel::geom_text_repel(
    data = traj %>% filter(scenario == levels(scenario)[2]),
    aes(label = mAL), size = 2.7, hjust = -0.1, direction = "y",
    nudge_x = 0.15, segment.size = 0.2, segment.color = "grey60"
  ) +
  scale_color_viridis_d(option = "C", end = 0.9, guide = "none") +
  expand_limits(x = length(scenarios) + 0.8) +
  labs(x = NULL, y = "Signal flow net input",
       title = sprintf("Top %d most ppk25-sensitive mAL subtypes (under cVA)", top_n)) +
  theme_pub(base_size = 10) +
  theme(plot.title = element_text(size = rel(1.05)))

save_fig5_plot(p_traj, "panel_I_mal_cva_mvf_trajectories", width = 10, height = 6)
print(p_traj)
```

```{r narrative, echo=FALSE}
out <- panel_plot_dir("panel_I", "txt")
top10 <- head(discrim, 10)
writeLines(sprintf(
"mAL cVA+male vs cVA+female (male vs female under shared cVA background)
========================================================================
For each of %d mAL subtypes, drive under cVA+male (DA1+ppk23) and cVA+female
(DA1+ppk23+ppk25) from the signal-flow model (mal_all_combos$net_input).
Spread = |drive(cVA+female) - drive(cVA+male)| isolates ppk25 sensitivity
under a shared cVA+ppk23 background.

Top 10 most ppk25-sensitive:
%s

Argmax counts (tie between the two scenarios at the 2-scenario resolution):
%s
",
nrow(discrim),
paste(sprintf("  %-12s |delta|=%.4f  higher=%s  delta(F-M)=%+.4f",
              top10$mAL, top10$spread, top10$argmax, top10$delta),
      collapse = "\n"),
paste(capture.output(discrim %>% count(argmax)), collapse = "\n")
), file.path(out, "panel_I_mal_cva_mvf.txt"))
```
