---
title: "Figure 5B supp v4: lollipop/callout — dots on top, pies below"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(scatterpie)
library(patchwork)
```

## What this is

Two-row infographic-style layout:

- **Top panel**: scatter at (delta_ppk, total_pct) with plain dots (no pies).
  Dot color = argmax AN09B017 variant. Dot size scales with total_pct.
- **Bottom panel**: one pie per mAL along the x-axis, sorted by delta_ppk.
  Each pie shows the AN09B017 variant composition for that mAL.
- **Connectors**: thin vertical lines link each top-panel dot to its bottom-panel pie
  (lollipop / callout pattern).

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
for (v in an09_present) {
  df_wide[[v]] <- ifelse(colSums(w) > 0, w[v, ] / colSums(w), 0)
}

# argmax variant per mAL (NA if no AN09B017 input)
frac_mat <- as.matrix(df_wide[, an09_present])
argmax_idx <- apply(frac_mat, 1, function(r) {
  if (all(r == 0) || all(is.na(r))) NA_integer_ else which.max(r)
})
df_wide$argmax_variant <- ifelse(is.na(argmax_idx),
                                 NA_character_,
                                 an09_present[argmax_idx])

# Rank ordering by delta_ppk for the bottom row
df_wide <- df_wide %>% arrange(delta_ppk)
df_wide$rank_x <- seq_len(nrow(df_wide))
df_wide$pie_y  <- 0  # constant y for the bottom row of pies
print(df_wide)
```

## Lollipop plot

```{r plot, fig.width=12, fig.height=9}
an09_palette <- setNames(
  RColorBrewer::brewer.pal(max(3, length(an09_present)), "Set1")[seq_along(an09_present)],
  an09_present
)

# Extend palette to include NA-safe grey for mALs without AN09B017 input
an09_palette_dot <- c(an09_palette, "none" = "grey70")
df_wide$argmax_for_color <- ifelse(is.na(df_wide$argmax_variant),
                                   "none", df_wide$argmax_variant)

# ---- TOP: scatter of dots -------------------------------------------------
p_top <- ggplot(df_wide, aes(x = delta_ppk, y = total_pct)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(aes(color = argmax_for_color, size = total_pct),
             alpha = 0.9, stroke = 0.3) +
  ggrepel::geom_text_repel(aes(label = mAL),
                           size = 2.6, color = "grey25",
                           max.overlaps = 30, min.segment.length = 0,
                           box.padding = 0.4, point.padding = 0.2,
                           segment.size = 0.2, segment.color = "grey60") +
  scale_color_manual(values = an09_palette_dot,
                     name = "argmax AN09B017 variant",
                     na.value = "grey70") +
  scale_size_area(max_size = 8, name = "total AN09B017 input (%)") +
  labs(
    x = expression(ppk23 - ppk25 ~ "(male - female contact drive)"),
    y = "Total AN09B017 input to mAL (% of mAL's total input)"
  ) +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot",
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(6, 6, 2, 6))

# ---- CONNECTORS: top dot (delta_ppk, total_pct) -> bottom pie (rank_x, 1) ---
# We use a shared x-scale trick: bottom panel x is rank_x, top is delta_ppk.
# We render connectors as their own middle panel with coord shared by rank_x.
# Simpler: draw connectors inside the bottom panel extending upward from the pie
# to a normalized y reference; and separately draw lines in top panel going down
# to x-axis. Use a dedicated middle "bridge" panel that maps delta_ppk -> rank_x.

# Build bridge data: each mAL has top position (delta_ppk) and bottom position (rank_x)
bridge_df <- df_wide %>%
  transmute(mAL, argmax_for_color,
            x_top = delta_ppk, x_bot = rank_x)

# Normalize top x onto the bottom x scale for visual alignment of connectors
x_top_range <- range(df_wide$delta_ppk)
x_bot_range <- range(df_wide$rank_x)
rescale_x <- function(x) {
  (x - x_top_range[1]) / diff(x_top_range) *
    diff(x_bot_range) + x_bot_range[1]
}
bridge_df$x_top_rescaled <- rescale_x(bridge_df$x_top)

p_bridge <- ggplot(bridge_df) +
  geom_segment(aes(x = x_top_rescaled, xend = x_bot,
                   y = 1, yend = 0,
                   color = argmax_for_color),
               linewidth = 0.3, alpha = 0.6) +
  scale_color_manual(values = an09_palette_dot, guide = "none",
                     na.value = "grey70") +
  scale_x_continuous(limits = x_bot_range, expand = c(0.02, 0.02)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = margin(0, 6, 0, 6))

# ---- BOTTOM: row of pies, sorted by delta_ppk -----------------------------
pie_r_bot <- 0.42  # radius in rank-x units (<0.5 so pies don't overlap)

p_bot <- ggplot() +
  geom_scatterpie(data = df_wide,
                  aes(x = rank_x, y = pie_y, r = pie_r_bot),
                  cols = an09_present,
                  color = "grey30", linewidth = 0.15) +
  coord_equal() +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  scale_x_continuous(breaks = df_wide$rank_x,
                     labels = df_wide$mAL,
                     limits = x_bot_range + c(-0.6, 0.6),
                     expand = c(0, 0)) +
  scale_y_continuous(limits = c(-0.6, 0.6), expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_pub(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.9)),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "right",
        plot.margin = margin(2, 6, 6, 6))

# ---- Stack ---------------------------------------------------------------
p <- p_top / p_bridge / p_bot +
  plot_layout(heights = c(4, 0.8, 2.2))

# Save via mspecific-aware helper -> plots_mspecific/panel_B/pdf/
save_fig5_plot(p,
               "panel_B_an09_dominance_vs_ppk_asymmetry_v4",
               width = 12, height = 9)

# Also copy to the curated single-panels directory
single_dir <- file.path(fig5_dir,
                        "my_selection_that_makes_halfway_sense_mspecific",
                        "single panels")
dir.create(single_dir, recursive = TRUE, showWarnings = FALSE)
src_pdf <- file.path(panel_plot_dir("panel_B_an09_dominance_vs_ppk_asymmetry_v4", "pdf"),
                     "panel_B_an09_dominance_vs_ppk_asymmetry_v4.pdf")
file.copy(src_pdf,
          file.path(single_dir, "panel_B_an09_dominance_vs_ppk_asymmetry_v4.pdf"),
          overwrite = TRUE)

print(p)
```
