---
title: "Figure 5I: Signal Flow Model vs Path-Based Predictions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

```{r panel-description, echo=FALSE}
description <- "
Panel I: Two independent models converge on the same population predictions
============================================================================
The path-based analysis (Panel F) and signal flow model make independent
predictions about how mAL subtypes respond to pheromone encounter scenarios.
This panel compares their predictions to assess robustness.

Left:  Signal flow scenario heatmap (3 scenarios x mAL subtypes)
Right: Scatter comparing path-based vs signal flow activation per subtype,
       faceted by scenario.

Key finding: The two models agree on the major features of the population
code, particularly which subtypes are activated vs inhibited by each scenario.
"
pdir <- panel_plot_dir("panel_I_description", "txt")
writeLines(description, file.path(pdir, "panel_I_description.txt"))
cat(description)
```

## Extract scenario data from signal flow model

```{r scenario-data}
# Define the 3 scenarios (same as Panel F)
# Lone male encounter excluded: male contact always co-occurs with cVA
scenarios <- c(
  "Female (ppk25)" = "ppk25",
  "cVA + male (cVA+ppk23)"      = "DA1+ppk23",
  "cVA + female (cVA+ppk25)"    = "DA1+ppk25"
)

scenario_colors <- c(
  "Female (ppk25)" = "#D73027",
  "cVA + male (cVA+ppk23)"      = "#8E24AA",
  "cVA + female (cVA+ppk25)"    = "#F57C00"
)

# Signal flow model now covers all mAL types via mAL_all target group
cat(sprintf("Signal flow model covers %d mAL types\n", length(mal_subtypes)))

# Signal flow model predictions (use net_input = pre-nonlinearity, same as Panel F)
scenario_data_sf <- list()
for (scenario_name in names(scenarios)) {
  combo_name <- scenarios[[scenario_name]]
  dat <- mal_all_combos %>%
    filter(combination == combo_name, target_type %in% mal_subtypes) %>%
    select(target_type, net_input) %>%
    arrange(target_type)
  scenario_data_sf[[scenario_name]] <- dat
}

# Build signal flow scenario matrix
sf_scenario_mat <- do.call(rbind, lapply(names(scenario_data_sf), function(nm) {
  d <- scenario_data_sf[[nm]]
  vals <- setNames(d$net_input, d$target_type)
  vals[mal_subtypes]
}))
rownames(sf_scenario_mat) <- names(scenarios)
sf_scenario_mat[is.na(sf_scenario_mat)] <- 0

# Path-based scenario matrix
path_scenario_mat <- matrix(0, nrow = length(scenarios), ncol = length(mal_subtypes),
                            dimnames = list(names(scenarios), mal_subtypes))
for (i in seq_along(scenarios)) {
  channels <- strsplit(scenarios[i], "\\+")[[1]]
  for (ch in channels) {
    if (ch %in% colnames(mal_activation_matrix)) {
      path_scenario_mat[i, ] <- path_scenario_mat[i, ] + mal_activation_matrix[mal_subtypes, ch]
    }
  }
}

cat("Signal flow scenario matrix:\n")
print(round(sf_scenario_mat, 3))
cat("\nPath-based scenario matrix:\n")
print(round(path_scenario_mat, 3))
```

## Signal flow scenario heatmap

```{r sf-heatmap, fig.width=12, fig.height=5}
max_abs <- max(abs(sf_scenario_mat))
if (max_abs == 0) max_abs <- 1
breaks <- seq(-max_abs, max_abs, length.out = 101)
colors <- colorRampPalette(c("#2166AC", "white", "#B2182B"))(100)

row_annot_sc <- data.frame(
  encounter = names(scenarios),
  row.names = names(scenarios)
)

col_annot_sc <- data.frame(
  row.names = colnames(sf_scenario_mat)
)

ann_colors_sc <- list(
  encounter = scenario_colors
)

# Dynamic cell sizing
n_sf_cols <- ncol(sf_scenario_mat)
cw_sf <- min(25, max(12, 400 / n_sf_cols))
fig_w_sf <- max(10, n_sf_cols * cw_sf / 72 + 4)

save_heatmap_fixed_annot(
  mat        = sf_scenario_mat,
  row_annot  = row_annot_sc,
  col_annot  = col_annot_sc,
  ann_colors = ann_colors_sc,
  filename   = file.path(panel_plot_dir("panel_I_scenario_heatmap", "png"),
                         "panel_I_scenario_heatmap.png"),
  width      = fig_w_sf, height = 5,
  breaks     = breaks, color = colors,
  cluster_rows = FALSE, cluster_cols = TRUE,
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  fontsize_row = 10, fontsize_col = 6,
  border_color = NA, cellwidth = cw_sf, cellheight = 25
)

save_heatmap_fixed_annot(
  mat        = sf_scenario_mat,
  row_annot  = row_annot_sc,
  col_annot  = col_annot_sc,
  ann_colors = ann_colors_sc,
  filename   = file.path(panel_plot_dir("panel_I_scenario_heatmap", "pdf"),
                         "panel_I_scenario_heatmap.pdf"),
  width      = fig_w_sf, height = 5,
  breaks     = breaks, color = colors,
  cluster_rows = FALSE, cluster_cols = TRUE,
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  fontsize_row = 10, fontsize_col = 6,
  border_color = NA, cellwidth = cw_sf, cellheight = 25
)
```

## Model comparison: path-based vs signal flow

```{r model-comparison, fig.width=14, fig.height=10}
# Build comparison dataframe
comparison <- data.frame()
for (sc in names(scenarios)) {
  for (st in mal_subtypes) {
    path_val <- path_scenario_mat[sc, st]
    sf_val   <- sf_scenario_mat[sc, st]
    comparison <- rbind(comparison, data.frame(
      scenario = sc,
      subtype = st,
      path_activation = path_val,
      sf_activation = sf_val,
      dimorphism = mal_dimorphism[st],
      nt = mal_nt[st],
      stringsAsFactors = FALSE
    ))
  }
}

comparison$scenario <- factor(comparison$scenario, levels = names(scenarios))

# Compute per-scenario Spearman correlations for annotation
cor_labels <- comparison %>%
  group_by(scenario) %>%
  summarise(
    rho = cor(path_activation, sf_activation, method = "spearman"),
    p = cor.test(path_activation, sf_activation, method = "spearman")$p.value,
    .groups = "drop"
  ) %>%
  mutate(label = sprintf("rho = %.2f\np = %s",
                          rho, ifelse(p < 0.001, "<0.001", format(p, digits = 2))))

# Scatter: path vs signal flow, faceted by scenario, colored by dimorphism
p_compare <- ggplot(comparison, aes(x = path_activation, y = sf_activation,
                                     color = dimorphism)) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.3,
              linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(size = 2.5, alpha = 0.8) +
  ggrepel::geom_text_repel(aes(label = subtype), size = 1.8,
                            max.overlaps = 15, min.segment.length = 0,
                            color = "grey30") +
  geom_text(data = cor_labels, aes(label = label),
            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2,
            size = 3, color = "black", inherit.aes = FALSE) +
  facet_wrap(~ scenario, ncol = 2) +
  scale_color_manual(values = dimorphism_colors, name = "Dimorphism") +
  coord_fixed(ratio = 1,
              xlim = range(c(comparison$path_activation, comparison$sf_activation)),
              ylim = range(c(comparison$path_activation, comparison$sf_activation))) +
  labs(x = "Path-based activation (exc - inh)",
       y = "Signal flow net input") +
  theme_pub(base_size = 10) +
  theme(legend.position = "top",
        strip.text = element_text(size = rel(1.1)))

save_fig5_plot(p_compare, "panel_I_model_comparison", width = 14, height = 10)
print(p_compare)
```

## Scenario discriminability: which subtypes shift most across encounters

```{r scenario-discriminability, fig.width=14, fig.height=12}
# For each subtype, compute how much its (path, sf) position varies across scenarios.
# High spread = this subtype responds differently to different encounters.
# Low spread = this subtype responds the same regardless of encounter type.

discrim <- comparison %>%
  group_by(subtype) %>%
  summarise(
    path_sd = sd(path_activation),
    sf_sd = sd(sf_activation),
    path_range = diff(range(path_activation)),
    sf_range = diff(range(sf_activation)),
    # Total spread in 2D (path, sf) space
    spread_2d = sqrt(var(path_activation) + var(sf_activation)),
    # Sum of pairwise Euclidean distances between scenario positions
    total_distance = {
      pts <- cbind(path_activation, sf_activation)
      sum(dist(pts))
    },
    dimorphism = first(dimorphism),
    nt = first(nt),
    .groups = "drop"
  ) %>%
  arrange(desc(spread_2d))

cat("Top 10 most discriminating subtypes:\n")
print(discrim %>% head(10) %>% select(subtype, spread_2d, path_range, sf_range, dimorphism))

# --- Bar plot: spread per subtype ---
discrim$subtype <- factor(discrim$subtype, levels = discrim$subtype)

p_spread_bar <- ggplot(discrim, aes(x = subtype, y = spread_2d, fill = dimorphism)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = dimorphism_colors, name = "Dimorphism") +
  labs(x = "mAL subtype", y = "Scenario discriminability\n(spread in path x signal flow space)") +
  theme_pub(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 6),
    legend.position = "top"
  )

# --- Scatter: mean position per subtype, size = spread ---
# Plot in 2D mean(path) vs mean(sf) space, point size encodes scenario spread
subtype_means <- comparison %>%
  group_by(subtype, dimorphism) %>%
  summarise(
    mean_path = mean(path_activation),
    mean_sf = mean(sf_activation),
    .groups = "drop"
  ) %>%
  left_join(discrim %>% select(subtype, spread_2d), by = "subtype")

p_spread_scatter <- ggplot(subtype_means, aes(x = mean_path, y = mean_sf,
                                                size = spread_2d, color = dimorphism)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(alpha = 0.7) +
  ggrepel::geom_text_repel(aes(label = subtype), size = 2, color = "grey30",
                            max.overlaps = 20, min.segment.length = 0) +
  scale_size_continuous(range = c(1.5, 14), name = "Scenario spread") +
  scale_color_manual(values = dimorphism_colors, name = "Dimorphism") +
  labs(x = "Mean path-based activation",
       y = "Mean signal flow net input") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right")

# --- Arrow plot: show each subtype's trajectory across scenarios ---
# Arrows connecting scenario positions for each subtype
scenario_order <- names(scenarios)
arrow_data <- comparison %>%
  arrange(subtype, scenario) %>%
  group_by(subtype) %>%
  mutate(
    path_next = lead(path_activation),
    sf_next = lead(sf_activation),
    scenario_next = lead(as.character(scenario))
  ) %>%
  filter(!is.na(path_next)) %>%
  ungroup() %>%
  left_join(discrim %>% select(subtype, spread_2d), by = "subtype")

# Only show top 15 most discriminating subtypes for readability
top_discrim <- discrim$subtype[1:min(15, nrow(discrim))]
arrow_filtered <- arrow_data %>% filter(subtype %in% top_discrim)
points_filtered <- comparison %>% filter(subtype %in% top_discrim)

p_trajectories <- ggplot() +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_segment(data = arrow_filtered,
               aes(x = path_activation, y = sf_activation,
                   xend = path_next, yend = sf_next, group = subtype),
               arrow = arrow(length = unit(0.08, "inches"), type = "closed"),
               linewidth = 0.3, alpha = 0.4, color = "grey40") +
  geom_point(data = points_filtered,
             aes(x = path_activation, y = sf_activation, color = scenario),
             size = 2.5, alpha = 0.8) +
  ggrepel::geom_text_repel(
    data = points_filtered %>% filter(scenario == "cVA + male (cVA+ppk23)"),
    aes(x = path_activation, y = sf_activation, label = subtype),
    size = 2, color = "grey30", max.overlaps = 20, min.segment.length = 0) +
  scale_color_manual(values = scenario_colors, name = "Encounter") +
  labs(x = "Path-based activation",
       y = "Signal flow net input",
       subtitle = "Top 15 most scenario-discriminating subtypes") +
  theme_pub(base_size = 10) +
  theme(legend.position = "top",
        plot.subtitle = element_text(size = 9, face = "italic"))

# Combined
p_discrim_combined <- (p_spread_bar) / (p_spread_scatter | p_trajectories) +
  plot_annotation(tag_levels = "a")

save_fig5_plot(p_discrim_combined, "panel_I_scenario_discriminability",
               width = 14, height = 12)
save_fig5_plot(p_spread_bar, "panel_I_spread_bar", width = 12, height = 5)
save_fig5_plot(p_spread_scatter, "panel_I_spread_scatter", width = 10, height = 8)
save_fig5_plot(p_trajectories, "panel_I_trajectories", width = 10, height = 8)
print(p_discrim_combined)
```

## Per-scenario correlation statistics

```{r correlation-stats}
cat("Model agreement per scenario:\n")
cor_results <- list()
for (sc in names(scenarios)) {
  d <- comparison %>% filter(scenario == sc)
  ct <- cor.test(d$path_activation, d$sf_activation, method = "spearman")
  cor_results[[sc]] <- ct
  cat(sprintf("  %s: Spearman rho=%.3f, p=%s\n",
              sc, ct$estimate, format(ct$p.value, digits = 3)))
}

# Sign agreement
cat("\nSign agreement (both models agree on +/-):\n")
for (sc in names(scenarios)) {
  d <- comparison %>% filter(scenario == sc)
  agree <- sum(sign(d$path_activation) == sign(d$sf_activation))
  cat(sprintf("  %s: %d/%d subtypes agree (%.0f%%)\n",
              sc, agree, nrow(d), 100 * agree / nrow(d)))
}
```

## Where do models disagree?

```{r disagreements, fig.width=10, fig.height=6}
# Identify subtypes where models disagree on sign
disagreements <- comparison %>%
  filter(sign(path_activation) != sign(sf_activation),
         abs(path_activation) > 0.001 | abs(sf_activation) > 0.01) %>%
  mutate(
    path_sign = ifelse(path_activation > 0, "+", "-"),
    sf_sign   = ifelse(sf_activation > 0, "+", "-")
  )

if (nrow(disagreements) > 0) {
  cat(sprintf("\nSubtypes with sign disagreement: %d\n", nrow(disagreements)))
  print(disagreements %>%
          select(scenario, subtype, path_activation, sf_activation, path_sign, sf_sign) %>%
          arrange(scenario, subtype))

  # Plot disagreement magnitudes
  p_disagree <- ggplot(disagreements,
                       aes(x = reorder(subtype, abs(path_activation - sf_activation)),
                           y = path_activation - sf_activation,
                           fill = scenario)) +
    geom_col(width = 0.7, position = "dodge") +
    scale_fill_manual(values = scenario_colors) +
    labs(x = "mAL subtype", y = "Path - Signal flow\n(activation difference)",
         fill = "Scenario") +
    coord_flip() +
    theme_pub(base_size = 10)

  save_fig5_plot(p_disagree, "panel_I_disagreements", width = 10, height = 6)
  print(p_disagree)
} else {
  cat("\nNo sign disagreements between models.\n")
}
```

## Activation function used by the signal-flow model

```{r activation-fn, fig.width=5, fig.height=4}
# f(x) = max(0, tanh(beta * x))    (sigmoid_rectified in signal_flow.py)
beta <- 5
act_df <- tibble(
  x = seq(-1, 1, length.out = 401)
) %>%
  mutate(
    tanh_only         = tanh(beta * x),
    sigmoid_rectified = pmax(0, tanh(beta * x))
  )

p_act <- ggplot(act_df, aes(x = x)) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.3, color = "grey70") +
  geom_line(aes(y = tanh_only), linetype = "dashed",
            color = "grey55", linewidth = 0.6) +
  geom_line(aes(y = sigmoid_rectified), color = "#B2182B", linewidth = 1) +
  annotate("text", x = 0.6,  y = -0.15, label = "tanh(beta*x)",
           hjust = 0, size = 3.2, color = "grey40") +
  annotate("text", x = 0.3,  y = 1.07,  label = "max(0, tanh(beta*x))",
           hjust = 0, size = 3.2, color = "#B2182B") +
  scale_y_continuous(limits = c(-1.05, 1.15), breaks = c(-1, 0, 1)) +
  scale_x_continuous(limits = c(-1, 1), breaks = c(-1, -0.5, 0, 0.5, 1)) +
  labs(x = "net input",
       y = "output (firing rate)",
       title = sprintf("Activation fn: sigmoid_rectified (beta = %d)", beta)) +
  theme_pub(base_size = 10) +
  theme(plot.title = element_text(size = rel(1)))

save_fig5_plot(p_act, "panel_I_activation_fn", width = 5, height = 4)
print(p_act)
```

## Combined panel — 2x2 grid (3 scenario scatters + activation function)

```{r combined, fig.width=12, fig.height=12}
common_lim <- range(c(comparison$path_activation, comparison$sf_activation))

scenario_scatter <- function(sc) {
  d <- comparison %>% filter(scenario == sc)
  cl <- cor_labels %>% filter(scenario == sc)
  ggplot(d, aes(x = path_activation, y = sf_activation, color = dimorphism)) +
    geom_abline(slope = 1, intercept = 0, linewidth = 0.3,
                linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
    geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
    geom_point(size = 2.5, alpha = 0.85) +
    ggrepel::geom_text_repel(aes(label = subtype), size = 2,
                              max.overlaps = 15, min.segment.length = 0,
                              color = "grey30") +
    geom_text(data = cl, aes(label = label),
              x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2,
              size = 3, color = "black", inherit.aes = FALSE) +
    scale_color_manual(values = dimorphism_colors, name = "Dimorphism",
                       guide = "none") +
    coord_fixed(ratio = 1, xlim = common_lim, ylim = common_lim) +
    labs(title = sc,
         x = "Path-based activation (exc - inh)",
         y = "Signal flow net input") +
    theme_pub(base_size = 10) +
    theme(plot.title = element_text(size = rel(1.05), face = "bold"))
}

p_fem    <- scenario_scatter("Female (ppk25)")
p_cva_m  <- scenario_scatter("cVA + male (cVA+ppk23)")
p_cva_f  <- scenario_scatter("cVA + female (cVA+ppk25)")

p_combined <- ((p_fem | p_cva_m) / (p_cva_f | p_act)) +
  plot_annotation(tag_levels = "a")

save_fig5_plot(p_combined, "panel_I_combined", width = 12, height = 12)
print(p_combined)
```

## Panel narrative

```{r narrative, echo=FALSE}
n_subtypes <- length(mal_subtypes)
n_scenarios <- length(scenarios)

# Overall correlation
all_cor <- cor.test(comparison$path_activation, comparison$sf_activation,
                    method = "spearman")

# Per-scenario correlations
rho_vals <- sapply(cor_results, function(x) x$estimate)
p_vals   <- sapply(cor_results, function(x) x$p.value)

# Sign agreement overall
sign_agree_all <- sum(sign(comparison$path_activation) == sign(comparison$sf_activation))
sign_agree_pct <- 100 * sign_agree_all / nrow(comparison)

n_disagree <- nrow(disagreements)

narrative <- sprintf(
"PANEL I: Model Comparison -- Path Analysis vs Signal Flow
============================================================

WHAT THIS PANEL SHOWS
---------------------
Left: Signal flow model scenario heatmap showing predicted mAL population
responses to the same 3 encounter scenarios as Panel F.

Right: Scatter comparing path-based activation (x-axis) vs signal flow
net input (y-axis) for each subtype under each scenario. Points near
the diagonal indicate model agreement. Colored by dimorphism class.

Panel F uses the path-based model (K-strongest paths, excitatory minus
inhibitory strengths). Panel I uses the signal flow model (iterative
propagation through the full connectivity matrix with sigmoid activation).
Both models make independent predictions about the same biological scenarios.

KEY OBSERVATIONS
----------------
1. Overall correlation across all %d subtype-scenario pairs:
   Spearman rho = %.3f, p = %s.

2. Per-scenario correlations:
   %s

3. Sign agreement: %d / %d pairs (%.0f%%) agree on direction.

4. Sign disagreements: %d subtype-scenario pairs where the two models
   predict opposite response directions.

BIOLOGICAL INTERPRETATION
-------------------------
Strong agreement between two independent computational models increases
confidence in the population code predictions. Disagreements highlight
subtypes where the specific computation matters -- likely neurons where
multi-hop signal propagation (captured by signal flow) differs from
the dominant K-strongest paths.

The path-based model captures the strongest direct routes; the signal flow
model integrates all routes through the network. Agreement means the
strongest paths dominate; disagreement means distributed connectivity
matters for that subtype.
",
nrow(comparison),
all_cor$estimate, format(all_cor$p.value, digits = 3),
paste(sprintf("  %s: rho=%.3f (p=%s)", names(rho_vals), rho_vals,
              sapply(p_vals, format, digits = 3)), collapse = "\n"),
sign_agree_all, nrow(comparison), sign_agree_pct,
n_disagree
)

save_panel_narrative("panel_I", narrative)
cat(narrative)
```
