---
title: "Figure 5B supp: AN09B017 input composition as pies vs ppk asymmetry"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(scatterpie)
```

## What this is

Per-mAL pie-chart scatter.

- **x** = ppk23 - ppk25 path-based drive (male minus female contact). Positive = ppk23-dominant, negative = ppk25-dominant.
- **y** = % of mAL's total input contributed by all AN09B017 variants combined.
- **pie** at each mAL's (x, y) = composition of that mAL's AN09B017 input broken down by variant (a-g). Slice area = fraction of AN09B017 input that variant supplies.
- **pie radius** = constant (purely visual); use y-axis value to read total AN09B017 contribution instead of radius.

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
# Attach per-variant fractions as columns
for (v in an09_present) {
  df_wide[[v]] <- ifelse(colSums(w) > 0, w[v, ] / colSums(w), 0)
}
print(df_wide)
```

## Pie scatter

```{r plot, fig.width=10, fig.height=8}
an09_palette <- setNames(
  RColorBrewer::brewer.pal(max(3, length(an09_present)), "Set1")[seq_along(an09_present)],
  an09_present
)

# Pie radius chosen relative to x/y axis ranges; each pie shows fractions as slices
x_range <- diff(range(df_wide$delta_ppk))
y_range <- diff(range(df_wide$total_pct))
pie_r   <- x_range * 0.04    # ~4% of x axis span

p <- ggplot() +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_scatterpie(data = df_wide,
                  aes(x = delta_ppk, y = total_pct, r = pie_r),
                  cols = an09_present,
                  color = "grey30", linewidth = 0.15) +
  coord_equal(ratio = x_range / y_range) +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  ggrepel::geom_text_repel(data = df_wide,
                           aes(x = delta_ppk, y = total_pct, label = mAL),
                           size = 2.8, color = "grey25",
                           max.overlaps = 30, min.segment.length = 0,
                           box.padding = 0.6, point.padding = pie_r) +
  labs(
    x = expression(ppk23 - ppk25 ~ "(male - female contact drive)"),
    y = "Total AN09B017 input to mAL (% of mAL's total input)",
    title = "AN09B017 input composition per mAL vs ppk asymmetry"
  ) +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot")

save_fig5_plot(p, "panel_B_an09_dominance_vs_ppk_asymmetry", width = 10, height = 8)
print(p)
```
