---
title: "Figure 5B supp v2: AN09B017 input composition as pies vs ppk asymmetry, radius = total_pct"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(scatterpie)
```

## What this is (v2)

Per-mAL pie-chart scatter, but now **pie area encodes total AN09B017 contribution**.

- **x** = ppk23 - ppk25 path-based drive (male minus female contact). Positive = ppk23-dominant, negative = ppk25-dominant.
- **pie radius** = sqrt(total_pct) so that **pie area is proportional to total_pct** (total % of mAL's total input contributed by all AN09B017 variants combined).
- **pie slices** = per-variant AN09B017 fractions (a-g) of that mAL's AN09B017 input.
- Y-axis value is not independently displayed; total is read off the pie size directly.

Rationale: biologically, total AN09B017 contribution IS the "radius of influence" on that mAL, so coupling it to the visual radius conveys it more directly than a y-coordinate.

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
# Attach per-variant fractions as columns
for (v in an09_present) {
  df_wide[[v]] <- ifelse(colSums(w) > 0, w[v, ] / colSums(w), 0)
}

# Drop mALs with zero AN09B017 contribution (no pie to draw, fractions all 0)
df_wide <- df_wide %>% filter(total_pct > 0)
print(df_wide)
```

## Pie scatter (1-D radial layout along x-axis, pie area ~ total_pct)

```{r plot, fig.width=12, fig.height=5}
an09_palette <- setNames(
  RColorBrewer::brewer.pal(max(3, length(an09_present)), "Set1")[seq_along(an09_present)],
  an09_present
)

# Pie radius: area-proportional to total_pct so r = k * sqrt(total_pct).
# Choose k so max pie radius is a sensible fraction of the x-axis span.
x_range <- diff(range(df_wide$delta_ppk))
max_r_target <- x_range * 0.08     # biggest pie ~ 8% of x-axis span
k_scale <- max_r_target / sqrt(max(df_wide$total_pct))
df_wide$r <- k_scale * sqrt(df_wide$total_pct)

# y is cosmetic: small deterministic jitter so equal-x pies don't overlap.
set.seed(1)
df_wide <- df_wide %>%
  arrange(delta_ppk) %>%
  mutate(y = 0)

# Deterministic vertical offset to reduce horizontal overlap: stagger neighbours.
# If two consecutive (sorted by x) pies would overlap, push one up.
if (nrow(df_wide) > 1) {
  y_vals <- numeric(nrow(df_wide))
  for (i in 2:nrow(df_wide)) {
    dx <- df_wide$delta_ppk[i] - df_wide$delta_ppk[i - 1]
    overlap_thresh <- df_wide$r[i] + df_wide$r[i - 1]
    if (dx < overlap_thresh) {
      # Alternate sign, and stack if still overlapping with two-back
      sign_i <- if (y_vals[i - 1] <= 0) 1 else -1
      y_vals[i] <- sign_i * overlap_thresh * 0.9
    } else {
      y_vals[i] <- 0
    }
  }
  df_wide$y <- y_vals
}

y_range <- max(1e-9, diff(range(df_wide$y)) + 2 * max(df_wide$r))

p <- ggplot() +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_scatterpie(data = df_wide,
                  aes(x = delta_ppk, y = y, r = r),
                  cols = an09_present,
                  color = "grey30", linewidth = 0.15) +
  coord_equal() +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  ggrepel::geom_text_repel(data = df_wide,
                           aes(x = delta_ppk, y = y, label = mAL),
                           size = 2.8, color = "grey25",
                           max.overlaps = 40, min.segment.length = 0,
                           box.padding = 0.6,
                           point.padding = mean(df_wide$r)) +
  scale_y_continuous(breaks = NULL, name = NULL) +
  labs(
    x = expression(ppk23 - ppk25 ~ "(male - female contact drive)"),
    title = "AN09B017 input composition per mAL vs ppk asymmetry (pie area ~ total AN09B017 %)"
  ) +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

# Size legend for pie area: draw reference pies at chosen total_pct values.
ref_pcts <- pretty(df_wide$total_pct, n = 3)
ref_pcts <- ref_pcts[ref_pcts > 0]
if (length(ref_pcts) == 0) ref_pcts <- round(max(df_wide$total_pct), 1)
ref_r    <- k_scale * sqrt(ref_pcts)
x_legend <- max(df_wide$delta_ppk) + x_range * 0.05
y_legend_base <- min(df_wide$y) - max(df_wide$r) * 2
ref_df <- tibble(
  label = paste0(ref_pcts, "%"),
  x = x_legend,
  y = y_legend_base - cumsum(2 * ref_r + max(df_wide$r) * 0.2),
  r = ref_r
)
# (Legend pies drawn as additional scatterpies with a single grey slice.)
ref_df$ref <- 1
p <- p +
  geom_scatterpie(data = ref_df,
                  aes(x = x, y = y, r = r),
                  cols = "ref",
                  color = "grey40", linewidth = 0.2, alpha = 0.3) +
  geom_text(data = ref_df,
            aes(x = x + max(ref_r) * 1.3, y = y, label = label),
            hjust = 0, size = 3, color = "grey25")

save_fig5_plot(p, "panel_B_an09_dominance_vs_ppk_asymmetry_v2", width = 12, height = 5)
print(p)
```
