---
title: "Figure 5A — Supplementary: real upstream source merge (option 1)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(igraph)
```

## What this does

The main panel A and the post-hoc supplementary checks (1–5) all work with the **existing** cached paths (`feather/strongest.paths/strongest.50.paths.{mod}.2.mAL_all.feather`). In those caches, modalities with multiple source neurons (ppk23: 10 sensilla, ppk25: 10, visual: 7) contribute 50 paths **per source**, i.e. 500 / 500 / 350 paths per (modality, target) pair.

DA1/VA1*/auditory have a single aggregate source (`ORN_DA1`, etc.) already upstream, so they get 50 paths per target.

This script **actually modifies the graph** — it pools the multi-source modalities into one virtual aggregate node each (`ppk23_merged`, `ppk25_merged`, `visual_merged`), with pooled outgoing edges — then re-runs Yen's k-strongest-paths from these virtual sources. Each modality ends up with a single "source" in the path-finder's view, matching DA1's structure. This is the faithful version of Option 1, not an approximation.

The output is cached (`feather/strongest.paths.merged/...`) so the expensive path-finding only runs once.

```{r out-dirs}
supp_png <- panel_plot_dir("panel_A", "png", "supp")
supp_pdf <- panel_plot_dir("panel_A", "pdf", "supp")
supp_txt <- panel_plot_dir("panel_A", "txt", "supp")

merged_cache_dir <- file.path(feather_dir, "strongest.paths.merged")
dir.create(merged_cache_dir, recursive = TRUE, showWarnings = FALSE)
cat("Path cache (merged sources):\n  ", merged_cache_dir, "\n")
```

## Build merged-source graph

```{r build-merged-graph}
# Which modalities to merge (the ones with > 1 source in channel_neurons)
multi_source_mods <- names(channel_neurons)[sapply(channel_neurons, length) > 1]
cat("Multi-source modalities:\n")
for (mod in multi_source_mods) {
  cat("  ", mod, "(n=", length(channel_neurons[[mod]]), "sensilla)\n")
}

# Start from the full graph
g_merged <- graph.general

# For each multi-source modality, add a synthetic merged source node
# whose outgoing edges sum the sensilla's outgoing edges
merged_source_name <- function(mod) sprintf("%s_merged", mod)

for (mod in multi_source_mods) {
  sensilla <- channel_neurons[[mod]]
  sensilla <- sensilla[sensilla %in% V(g_merged)$name]
  if (length(sensilla) == 0) next

  merged_name <- merged_source_name(mod)
  # Skip if somehow already added
  if (merged_name %in% V(g_merged)$name) next

  # Find each sensillum's outgoing edges
  out_edges <- ego(g_merged, order = 1, nodes = sensilla, mode = "out", mindist = 1)
  neighbor_weight_list <- list()
  for (i in seq_along(sensilla)) {
    s <- sensilla[i]
    # get outgoing edges for this sensillum
    eids <- incident(g_merged, v = s, mode = "out")
    if (length(eids) == 0) next
    ends <- ends(g_merged, eids, names = TRUE)[, 2]
    wts  <- E(g_merged)$weight[eids]
    for (j in seq_along(ends)) {
      y <- ends[j]
      w <- wts[j]
      neighbor_weight_list[[y]] <- c(neighbor_weight_list[[y]], w)
    }
  }

  if (length(neighbor_weight_list) == 0) next

  # Aggregate: sum of edge weights per unique downstream neighbor
  # (biological rationale: pooled activation — action potentials in any sensillum
  #  contribute to postsynaptic drive at Y)
  agg_targets <- names(neighbor_weight_list)
  agg_weights <- sapply(neighbor_weight_list, sum)

  # Add the merged vertex
  g_merged <- add_vertices(g_merged, 1, attr = list(name = merged_name))
  # Add edges from merged vertex to aggregated downstream neighbors
  new_edges <- as.vector(rbind(rep(merged_name, length(agg_targets)), agg_targets))
  g_merged <- add_edges(g_merged, new_edges, attr = list(weight = agg_weights))

  cat(sprintf("  Added %s: %d outgoing edges (merged from %d sensilla)\n",
              merged_name, length(agg_targets), length(sensilla)))
}

cat("\nMerged graph: |V|=", vcount(g_merged), " |E|=", ecount(g_merged),
    " (original: |V|=", vcount(graph.general), " |E|=", ecount(graph.general), ")\n", sep = "")
```

## Run path-finder on merged graph (cached)

```{r run-path-finder}
run_merged_paths <- function(mod, force = FALSE) {
  mod_file_label <- modality_file_labels[[mod]]
  fname <- file.path(merged_cache_dir,
    sprintf("strongest.%d.paths.%s_merged.2.%s.feather",
            n_paths_mal, mod_file_label, target_name_mal))

  if (file.exists(fname) && !force) {
    df <- read_feather(fname)
    df <- drop_list_cols(df)
    message(sprintf("  Loaded cached merged paths for %s (%d rows)", mod, nrow(df)))
    return(df)
  }

  # Determine the merged source name for this modality
  merged_name <- if (mod %in% multi_source_mods) merged_source_name(mod) else {
    # Already single-source modality — use its original start
    channel_neurons[[mod]][1]
  }
  if (!merged_name %in% V(g_merged)$name) {
    message(sprintf("  SKIP %s: merged source %s not in graph", mod, merged_name))
    return(NULL)
  }
  starts <- merged_name
  targets <- mal_subtypes[mal_subtypes %in% V(g_merged)$name]

  # Prepare log-weight edge attribute (same as setup.R does for path-finding)
  g <- g_merged
  w <- E(g)$weight
  w[!is.finite(w) | w <= 0] <- .Machine$double.eps
  E(g)$log_weight <- -log(w)

  message(sprintf("  Running Yen K-strongest-paths: %s (%s) -> %d mAL targets ...",
                  mod, merged_name, length(targets)))
  t0 <- Sys.time()
  result <- find_k_strongest_paths_yen(
    g       = g,
    starts  = starts,
    targets = targets,
    mba     = mba,
    n_paths = n_paths_mal
  )
  t1 <- Sys.time()
  message(sprintf("    completed in %.1f s", as.numeric(difftime(t1, t0, units = "secs"))))

  df <- result$df
  if (nrow(df) == 0) {
    message(sprintf("    No paths found for %s", mod))
    return(NULL)
  }
  df$modality   <- mod
  df$target_set <- target_name_mal
  write_feather(df, fname)
  message(sprintf("    Cached %d paths to %s", nrow(df), basename(fname)))
  df
}

# Run for all 7 modalities (cheap for single-source ones; potentially slow for merged)
merged_paths_list <- list()
for (mod in modality_order) {
  merged_paths_list[[mod]] <- run_merged_paths(mod)
}
mal_paths_merged <- bind_rows(merged_paths_list)

cat("\nPath counts after actual merge:\n")
mal_paths_merged %>% count(modality) %>% print()
```

## Compute activation matrix from merged paths

```{r activation-merged}
# Restrict to the active mAL_subtypes (works for both regular and mspecific)
mal_paths_merged <- mal_paths_merged %>% filter(end %in% mal_subtypes)

compute_valence_strength_local <- function(paths_df, val) {
  paths_df %>%
    filter(valence == val) %>%
    group_by(end, modality) %>%
    summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = modality, values_from = total_strength, values_fill = 0) %>%
    column_to_rownames("end") %>%
    as.matrix()
}

mat_exc <- compute_valence_strength_local(mal_paths_merged, "excitatory")
mat_inh <- compute_valence_strength_local(mal_paths_merged, "inhibitory")

# Align
align_mat <- function(m) {
  missing_r <- setdiff(mal_subtypes, rownames(m))
  if (length(missing_r)) m <- rbind(m, matrix(0, length(missing_r), ncol(m),
                                              dimnames = list(missing_r, colnames(m))))
  missing_c <- setdiff(modality_order, colnames(m))
  if (length(missing_c)) m <- cbind(m, matrix(0, nrow(m), length(missing_c),
                                              dimnames = list(rownames(m), missing_c)))
  m[mal_subtypes, modality_order, drop = FALSE]
}
mat_exc <- align_mat(mat_exc)
mat_inh <- align_mat(mat_inh)
mat_merge <- mat_exc - mat_inh
cat("Merged-source activation matrix: ", nrow(mat_merge), "x", ncol(mat_merge),
    " range [", sprintf("%.4f, %.4f", min(mat_merge), max(mat_merge)), "]\n", sep = "")
```

## Plot heatmap

```{r heatmap-merged}
row_annot <- mal_annot[rownames(mat_merge), c("neurotransmitter", "dimorphism"), drop = FALSE]
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_merge)],
  row.names = colnames(mat_merge)
)
ann_colors <- list(
  neurotransmitter = nt_colors,
  dimorphism       = dimorphism_colors,
  modality_group   = modality_group_colors
)
n_colors <- 100
max_abs <- max(abs(mat_merge)); if (max_abs == 0) max_abs <- 1
breaks <- seq(-max_abs, max_abs, length.out = n_colors + 1)
colors <- colorRampPalette(c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                              "white",
                              "#FDDBC7", "#F4A582", "#D6604D", "#B2182B"))(n_colors)

png_file <- file.path(supp_png, "panel_A_supp_real_merge.png")
pdf_file <- file.path(supp_pdf, "panel_A_supp_real_merge.pdf")

pheatmap::pheatmap(
  mat_merge,
  annotation_row = row_annot, annotation_col = col_annot,
  annotation_colors = ann_colors,
  breaks = breaks, color = colors,
  cluster_rows = TRUE, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  fontsize_row = 6, fontsize_col = 10, border_color = NA,
  cellwidth = 30, cellheight = 14, gaps_col = c(3, 4, 5),
  main = "Real upstream source merge (re-run path-finder on pooled-source graph)",
  filename = png_file, width = 8, height = 10
)
pheatmap::pheatmap(
  mat_merge,
  annotation_row = row_annot, annotation_col = col_annot,
  annotation_colors = ann_colors,
  breaks = breaks, color = colors,
  cluster_rows = TRUE, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  fontsize_row = 6, fontsize_col = 10, border_color = NA,
  cellwidth = 30, cellheight = 14, gaps_col = c(3, 4, 5),
  main = "Real upstream source merge (re-run path-finder on pooled-source graph)",
  filename = pdf_file, width = 8, height = 10
)
cat("Saved:", basename(png_file), "  range: [", sprintf("%.4f, %.4f]", min(mat_merge), max(mat_merge)), "\n")
```

## Comparison vs raw panel A

```{r compare}
comparison <- tibble(
  modality = colnames(mal_activation_matrix),
  raw_mean_abs   = colMeans(abs(mal_activation_matrix)),
  merge_mean_abs = colMeans(abs(mat_merge))
) %>%
  mutate(modality = factor(modality, levels = modality_order),
         raw_rel   = raw_mean_abs   / max(raw_mean_abs),
         merge_rel = merge_mean_abs / max(merge_mean_abs))

print(comparison)
write.csv(comparison, file.path(supp_txt, "panel_A_supp_real_merge_comparison.csv"),
          row.names = FALSE)

p_cmp <- comparison %>%
  pivot_longer(c(raw_rel, merge_rel), names_to = "method", values_to = "rel_strength") %>%
  mutate(method = factor(method, levels = c("raw_rel", "merge_rel"),
                         labels = c("raw panel A", "real merge (option 1)"))) %>%
  ggplot(aes(x = modality, y = rel_strength, fill = method)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.8) +
  labs(x = NULL, y = "Mean |cell| per channel, normalized to each method's max",
       title = "Panel A raw vs real upstream source merge",
       subtitle = "Merged graph re-runs Yen K-strongest paths with pooled source nodes") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "bottom", legend.title = element_blank())

ggsave(file.path(supp_png, "panel_A_supp_real_merge_comparison.png"),
       p_cmp, width = 9, height = 5, dpi = 300)
ggsave(file.path(supp_pdf, "panel_A_supp_real_merge_comparison.pdf"),
       p_cmp, width = 9, height = 5)

p_cmp
```

```{r narrative, echo=FALSE}
txt <- sprintf(
"Panel A -- REAL upstream source merge (Option 1)
==================================================

Graph modification:
  Multi-source modalities (%s) have their sensilla pooled into a single
  virtual node '%s_merged', with outgoing edges = sum of the sensilla's
  outgoing edge weights to each shared downstream neighbor. Single-source
  modalities (DA1, VA1v, VA1d, auditory) are untouched.

Path cache:
  Re-run of Yen K-strongest paths (K=%d) from merged sources to all
  mAL subtypes. Results cached at feather/strongest.paths.merged/.

Each modality now contributes %d paths per (modality, target) pair.

Mean |cell| per channel (normalized to each method's max):
%s

Interpretation:
  Any remaining dominance of ppk channels is NOT driven by having
  more sensilla; it reflects the genuine graph structure after pooling.
",
paste(multi_source_mods, collapse = ", "),
multi_source_mods[1],
n_paths_mal, n_paths_mal,
paste(sprintf("  %-9s  raw=%.2f  merge=%.2f",
              comparison$modality, comparison$raw_rel, comparison$merge_rel),
      collapse = "\n")
)
writeLines(txt, file.path(supp_txt, "panel_A_supp_real_merge.txt"))
cat(txt)
```
