---
title: "Figure 5I: mAL male vs female contact (no cVA)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(patchwork)
```

## What this is

Two-scenario version with **no cVA (volatile DA1) at all**:

- Male encounter:   `ppk23`         (male contact pheromone only)
- Female encounter: `ppk23,ppk25`   (both contact pheromones)

Shared channel: ppk23. Toggled channel: ppk25. Spread = `drive(female) - drive(male)`
in absolute value - isolates how each mAL subtype changes when ppk25 (female
contact) is added on top of ppk23 alone.

```{r drive}
scenarios <- c(
  "Male (ppk23)"   = "ppk23",
  "Female (ppk25)" = "ppk25"
)
scenario_colors <- c(
  "Male (ppk23)"   = "#1B7837",
  "Female (ppk25)" = "#D73027"
)

drive_long <- bind_rows(lapply(names(scenarios), function(nm) {
  mal_all_combos %>%
    filter(channels == scenarios[[nm]], target_type %in% mal_subtypes) %>%
    transmute(mAL = target_type, scenario = nm, drive = net_input)
})) %>%
  mutate(scenario = factor(scenario, levels = names(scenarios)))

discrim <- drive_long %>%
  group_by(mAL) %>%
  summarise(
    spread     = max(drive) - min(drive),
    mean_drive = mean(drive),
    argmax     = scenario[which.max(drive)],
    delta      = drive[scenario == "Female (ppk25)"] - drive[scenario == "Male (ppk23)"],
    .groups = "drop"
  ) %>%
  arrange(desc(spread))

cat("Top 10 most ppk25-sensitive mAL subtypes (no cVA):\n")
print(head(discrim, 10))
cat("\nArgmax counts:\n"); print(discrim %>% count(argmax))
```

## Bar chart

```{r bar, fig.width=12, fig.height=4}
discrim_ord <- discrim %>% mutate(mAL = factor(mAL, levels = mAL))

p_bar <- ggplot(discrim_ord, aes(x = mAL, y = spread, fill = argmax)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = scenario_colors, name = "Higher-drive scenario") +
  labs(x = "mAL subtype (sorted by |female - male| drive)",
       y = "ppk25 sensitivity\n|female drive - male drive|") +
  theme_pub(base_size = 10) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 7),
        legend.position = "top")

save_fig5_plot(p_bar, "panel_I_mal_mvf_bar", width = 12, height = 4)
print(p_bar)
```

## Scatter

```{r scatter, fig.width=8, fig.height=6}
p_scatter <- ggplot(discrim, aes(x = mean_drive, y = spread,
                                 size = spread, color = argmax)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_point(alpha = 0.85) +
  ggrepel::geom_text_repel(aes(label = mAL), size = 2.6, color = "grey30",
                            max.overlaps = 25, min.segment.length = 0) +
  scale_color_manual(values = scenario_colors,
                     name = "Higher-drive scenario",
                     guide = guide_legend(override.aes = list(size = 4))) +
  scale_size_continuous(range = c(1.5, 8), name = "|delta|") +
  labs(x = "Mean drive across male and female",
       y = "|female drive - male drive|") +
  theme_pub(base_size = 10) +
  theme(legend.position = "right")

save_fig5_plot(p_scatter, "panel_I_mal_mvf_scatter", width = 8, height = 6)
print(p_scatter)
```

## Trajectories

```{r traj, fig.width=10, fig.height=6}
top_n <- 15
top_mal <- head(discrim$mAL, top_n)
traj <- drive_long %>% filter(mAL %in% top_mal) %>%
  mutate(mAL = factor(mAL, levels = top_mal))

p_traj <- ggplot(traj, aes(x = scenario, y = drive, group = mAL, color = mAL)) +
  geom_hline(yintercept = 0, linewidth = 0.2, color = "grey70") +
  geom_line(linewidth = 0.75, alpha = 0.85) +
  geom_point(size = 2.2) +
  ggrepel::geom_text_repel(
    data = traj %>% filter(scenario == levels(scenario)[2]),
    aes(label = mAL), size = 2.7, hjust = -0.1, direction = "y",
    nudge_x = 0.15, segment.size = 0.2, segment.color = "grey60"
  ) +
  scale_color_viridis_d(option = "C", end = 0.9, guide = "none") +
  expand_limits(x = length(scenarios) + 0.8) +
  labs(x = NULL, y = "Signal flow net input",
       title = sprintf("Top %d most ppk25-sensitive mAL subtypes (no cVA)", top_n)) +
  theme_pub(base_size = 10) +
  theme(plot.title = element_text(size = rel(1.05)))

save_fig5_plot(p_traj, "panel_I_mal_mvf_trajectories", width = 10, height = 6)
print(p_traj)
```

```{r narrative, echo=FALSE}
out <- panel_plot_dir("panel_I", "txt")
top10 <- head(discrim, 10)
writeLines(sprintf(
"mAL male-contact vs female-contact (no cVA)
============================================
For each of %d mAL subtypes, drive under male (ppk23) and female
(ppk23+ppk25) encounter scenarios from the signal-flow model
(mal_all_combos$net_input). Spread = |drive(female) - drive(male)|
isolates ppk25 sensitivity without cVA.

Top 10 most ppk25-sensitive:
%s

Argmax counts:
%s
",
nrow(discrim),
paste(sprintf("  %-12s |delta|=%.4f  higher=%s  delta(F-M)=%+.4f",
              top10$mAL, top10$spread, top10$argmax, top10$delta),
      collapse = "\n"),
paste(capture.output(discrim %>% count(argmax)), collapse = "\n")
), file.path(out, "panel_I_mal_mvf.txt"))
```
