---
title: "Figure 5A — Supplementary: intermediate-node variants collapsed"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

## What this does

The raw path cache distinguishes fine-grained neuron variants along the relay chain — e.g. `AN09B017a`, `AN09B017b`, ..., `AN09B017g` (7 variants), `AN05B102a/b/c`, etc. The mAL target column in panel A is already at its finest granularity (`mAL_m3a`, `mAL_m3b`, ...), and the user wants to keep those intact.

This supplement asks: **what if we treat each intermediate neuron group as a single aggregate node** (pool AN09B017a–g into AN09B017, AN05B102a–c into AN05B102, etc.), leaving only the sensory source (first hop) and the mAL target (last hop) at their original granularity?

Mechanic: each path is a node sequence. We strip trailing lowercase letters from every **intermediate** node. Paths that collapse to the same sequence get deduplicated — strengths are **summed** across the collapsed variants (biologically: the effective strongest path through "AN09B017" is the pooled contribution of its subvariants that share the same downstream route).

```{r out-dirs}
supp_png <- panel_plot_dir("panel_A", "png", "supp")
supp_pdf <- panel_plot_dir("panel_A", "pdf", "supp")
supp_txt <- panel_plot_dir("panel_A", "txt", "supp")
```

## Collapse intermediate nodes in each path

```{r collapse-paths}
collapse_intermediate <- function(path_string) {
  toks <- strsplit(path_string, " -> ", fixed = TRUE)[[1]]
  if (length(toks) <= 2) return(paste(toks, collapse = " -> "))
  # strip trailing single lowercase letter on the intermediate nodes
  toks[-c(1, length(toks))] <- sub("([a-z])$", "", toks[-c(1, length(toks))])
  paste(toks, collapse = " -> ")
}

collapse_paths <- function(paths_df) {
  paths_df %>%
    mutate(path_collapsed = vapply(path, collapse_intermediate, character(1))) %>%
    # Within each (end, modality, valence, collapsed path), the variants
    # collapse to one logical path. Sum their strengths.
    group_by(start, end, modality, valence, path_collapsed) %>%
    summarise(strength = sum(strength, na.rm = TRUE),
              hops     = first(hops),
              .groups  = "drop") %>%
    rename(path = path_collapsed)
}

paths_raw_collapsed <- collapse_paths(mal_paths)

# Path-count summary before vs after
before <- mal_paths %>% count(modality, name = "n_before")
after  <- paths_raw_collapsed %>% count(modality, name = "n_after")
cat("Path counts per modality (before vs after collapsing intermediates):\n")
print(full_join(before, after, by = "modality"))

# How many unique intermediate node names disappeared
unique_nodes_raw <- unique(unlist(strsplit(mal_paths$path, " -> ", fixed = TRUE)))
unique_nodes_collapsed <- unique(unlist(strsplit(paths_raw_collapsed$path, " -> ", fixed = TRUE)))
cat(sprintf("\nDistinct node labels in paths:  raw = %d   collapsed = %d   (reduction = %d)\n",
            length(unique_nodes_raw), length(unique_nodes_collapsed),
            length(unique_nodes_raw) - length(unique_nodes_collapsed)))
```

## Helper: activation matrix from any path set (filters to mal_subtypes)

```{r matrix-helper}
activation_matrix_from <- function(paths_df) {
  paths_df <- paths_df %>% filter(end %in% mal_subtypes)

  per_valence <- function(val) {
    paths_df %>%
      filter(valence == val) %>%
      group_by(end, modality) %>%
      summarise(total = sum(strength, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(names_from = modality, values_from = total, values_fill = 0) %>%
      column_to_rownames("end") %>% as.matrix()
  }
  exc <- per_valence("excitatory")
  inh <- per_valence("inhibitory")

  fill <- function(m) {
    mr <- setdiff(mal_subtypes, rownames(m))
    if (length(mr)) m <- rbind(m, matrix(0, length(mr), ncol(m),
                                         dimnames = list(mr, colnames(m))))
    mc <- setdiff(modality_order, colnames(m))
    if (length(mc)) m <- cbind(m, matrix(0, nrow(m), length(mc),
                                         dimnames = list(rownames(m), mc)))
    m[mal_subtypes, modality_order, drop = FALSE]
  }
  fill(exc) - fill(inh)
}

plot_supp_heatmap <- function(mat, name, subtitle) {
  row_annot <- mal_annot[rownames(mat), c("neurotransmitter", "dimorphism"), drop = FALSE]
  col_annot <- data.frame(modality_group = modality_group[colnames(mat)],
                          row.names = colnames(mat))
  ann_colors <- list(
    neurotransmitter = nt_colors,
    dimorphism       = dimorphism_colors,
    modality_group   = modality_group_colors
  )
  n_colors <- 100
  max_abs <- max(abs(mat)); if (max_abs == 0) max_abs <- 1
  breaks <- seq(-max_abs, max_abs, length.out = n_colors + 1)
  colors <- colorRampPalette(c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                                "white",
                                "#FDDBC7", "#F4A582", "#D6604D", "#B2182B"))(n_colors)

  png_file <- file.path(supp_png, paste0(name, ".png"))
  pdf_file <- file.path(supp_pdf, paste0(name, ".pdf"))

  for (f in c(png_file, pdf_file)) {
    pheatmap::pheatmap(
      mat,
      annotation_row = row_annot, annotation_col = col_annot,
      annotation_colors = ann_colors,
      breaks = breaks, color = colors,
      cluster_rows = TRUE, cluster_cols = FALSE,
      clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
      fontsize_row = 6, fontsize_col = 10, border_color = NA,
      cellwidth = 30, cellheight = 14, gaps_col = c(3, 4, 5),
      main = subtitle,
      filename = f, width = 8, height = 10
    )
  }
  cat(sprintf("Saved %s   range [%.4f, %.4f]\n", name, min(mat), max(mat)))
  invisible(mat)
}
```

## Heatmap 1 — from raw paths, intermediates collapsed

```{r lvl1-raw}
mat_raw <- activation_matrix_from(paths_raw_collapsed)
plot_supp_heatmap(mat_raw, "panel_A_supp_interm_collapse_raw",
                  "Intermediate-node variants collapsed (from raw paths)")
```

## Heatmap 2 — from merged-source paths, intermediates also collapsed

```{r lvl1-merged}
merged_cache_dir <- file.path(feather_dir, "strongest.paths.merged")
mat_merge <- NULL
if (dir.exists(merged_cache_dir) && length(list.files(merged_cache_dir)) > 0) {
  load_merged <- function() {
    out <- list()
    for (mod in modality_order) {
      lbl <- modality_file_labels[[mod]]
      fname <- file.path(merged_cache_dir,
        sprintf("strongest.%d.paths.%s_merged.2.%s.feather",
                n_paths_mal, lbl, target_name_mal))
      if (!file.exists(fname)) next
      df <- read_feather(fname)
      df <- drop_list_cols(df)
      df$modality <- mod
      out[[mod]] <- df
    }
    bind_rows(out)
  }
  mal_paths_merged <- load_merged()

  if (nrow(mal_paths_merged) > 0) {
    paths_merged_collapsed <- collapse_paths(mal_paths_merged)
    mat_merge <- activation_matrix_from(paths_merged_collapsed)
    plot_supp_heatmap(mat_merge, "panel_A_supp_interm_collapse_merged",
                      "Intermediate-node variants collapsed AND sources merged")
  } else {
    cat("No merged paths found; skipping combined heatmap.\n")
  }
} else {
  cat("Merged cache not present. Run panel_A_supp_real_merge.Rmd first.\n")
}
```

## Summary

```{r summary, echo=FALSE}
rows <- list(
  tibble(method = "raw (panel A)",           modality = colnames(mal_activation_matrix),
         mean_abs = colMeans(abs(mal_activation_matrix))),
  tibble(method = "intermediates collapsed", modality = colnames(mat_raw),
         mean_abs = colMeans(abs(mat_raw)))
)
if (!is.null(mat_merge)) {
  rows <- c(rows, list(
    tibble(method = "intermediates + sources merged", modality = colnames(mat_merge),
           mean_abs = colMeans(abs(mat_merge)))
  ))
}
summary_df <- bind_rows(rows) %>%
  group_by(method) %>% mutate(rel = mean_abs / max(mean_abs)) %>% ungroup() %>%
  mutate(modality = factor(modality, levels = modality_order))

write.csv(summary_df, file.path(supp_txt, "panel_A_supp_interm_collapse_summary.csv"),
          row.names = FALSE)
print(summary_df %>% select(method, modality, rel) %>%
      pivot_wider(names_from = method, values_from = rel))

txt <- sprintf(
"Panel A supplement -- intermediate-node variants collapsed
===========================================================

Collapse rule: for each cached path, strip trailing single lowercase letter
from every *intermediate* node (not the sensory source, not the mAL target).
Example: AN09B017a/b/c/d/e/f/g all collapse to AN09B017.

Paths that become identical after collapsing are deduplicated; their
strengths are summed (biologically: the effective aggregate AN09B017 neuron
contributes the pooled synaptic drive of its variants along the shared
downstream route).

Distinct node labels along paths:
  raw         : %d
  collapsed   : %d
  reduction   : %d

Heatmaps:
  - panel_A_supp_interm_collapse_raw       (from mal_paths, intermediates collapsed)
  - panel_A_supp_interm_collapse_merged    (from merged-source cache, intermediates also collapsed)

Note: mAL targets are kept at their finest granularity (mAL_m3a, mAL_m3b, ...)
so the heatmap is directly comparable to the main panel A.
",
length(unique_nodes_raw), length(unique_nodes_collapsed),
length(unique_nodes_raw) - length(unique_nodes_collapsed))
writeLines(txt, file.path(supp_txt, "panel_A_supp_interm_collapse.txt"))
cat(txt)
```
