---
title: "Figure 5B supp v1: AN09B017 input composition as pies vs ppk asymmetry (fixed geometry)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(scatterpie)
```

## What this is

Per-mAL pie-chart scatter (v1: normalized coords for truly circular pies).

- **x** = ppk23 - ppk25 path-based drive (male minus female contact). Positive = ppk23-dominant, negative = ppk25-dominant.
- **y** = % of mAL's total input contributed by all AN09B017 variants combined.
- **pie** at each mAL's (x, y) = composition of that mAL's AN09B017 input broken down by variant (a-g). Slice area = fraction of AN09B017 input that variant supplies.
- **pie radius** = constant (purely visual); use y-axis value to read total AN09B017 contribution instead of radius.

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
# Attach per-variant fractions as columns
for (v in an09_present) {
  df_wide[[v]] <- ifelse(colSums(w) > 0, w[v, ] / colSums(w), 0)
}
print(df_wide)
```

## Pie scatter (fixed geometry: both axes normalized to [0,1], coord_fixed(1))

```{r plot, fig.width=10, fig.height=8}
an09_palette <- setNames(
  RColorBrewer::brewer.pal(max(3, length(an09_present)), "Set1")[seq_along(an09_present)],
  an09_present
)

# --- Normalize both axes to [0, 1] so that coord_fixed(ratio = 1) yields
#     genuinely circular pies regardless of the raw x/y scales. ---
x_raw <- df_wide$delta_ppk
y_raw <- df_wide$total_pct

x_lo <- min(x_raw); x_hi <- max(x_raw)
y_lo <- min(y_raw); y_hi <- max(y_raw)

# Pad the ranges a bit so pies near the edges don't get clipped.
x_pad <- 0.08 * (x_hi - x_lo)
y_pad <- 0.08 * (y_hi - y_lo)
x_lo <- x_lo - x_pad; x_hi <- x_hi + x_pad
y_lo <- y_lo - y_pad; y_hi <- y_hi + y_pad

x_span <- x_hi - x_lo
y_span <- y_hi - y_lo

df_wide$x_n <- (df_wide$delta_ppk - x_lo) / x_span
df_wide$y_n <- (df_wide$total_pct - y_lo) / y_span

# --- Choose a pie radius in the normalized [0,1] unit system.
#     Large enough to read 7 slices, small enough that no two pies overlap.
#     Compute min pairwise distance among pie centres to bound the radius. ---
pts <- as.matrix(df_wide[, c("x_n", "y_n")])
d_mat <- as.matrix(dist(pts))
diag(d_mat) <- Inf
min_dist <- min(d_mat)

# Target radius: big, but strictly smaller than half the nearest-neighbour
# distance so pies never overlap. Cap at a comfortable absolute size.
pie_r <- min(0.45 * min_dist, 0.04)
pie_r <- max(pie_r, 0.018)  # never shrink below legibility threshold

# Axis breaks in original units, but placed on the [0,1] coord system.
x_breaks_raw <- pretty(c(min(x_raw), max(x_raw)), n = 5)
y_breaks_raw <- pretty(c(min(y_raw), max(y_raw)), n = 5)
x_breaks_n <- (x_breaks_raw - x_lo) / x_span
y_breaks_n <- (y_breaks_raw - y_lo) / y_span
# Keep only in-range breaks
keep_x <- x_breaks_n >= 0 & x_breaks_n <= 1
keep_y <- y_breaks_n >= 0 & y_breaks_n <= 1
x_breaks_n <- x_breaks_n[keep_x]; x_breaks_raw <- x_breaks_raw[keep_x]
y_breaks_n <- y_breaks_n[keep_y]; y_breaks_raw <- y_breaks_raw[keep_y]

# Zero lines in normalized coordinates
zero_x_n <- (0 - x_lo) / x_span
zero_y_n <- (0 - y_lo) / y_span

p <- ggplot() +
  { if (zero_y_n >= 0 && zero_y_n <= 1)
      geom_hline(yintercept = zero_y_n, linewidth = 0.2, color = "grey70") } +
  { if (zero_x_n >= 0 && zero_x_n <= 1)
      geom_vline(xintercept = zero_x_n, linewidth = 0.2, color = "grey70") } +
  geom_scatterpie(data = df_wide,
                  aes(x = x_n, y = y_n, r = pie_r),
                  cols = an09_present,
                  color = "grey30", linewidth = 0.1) +
  coord_fixed(ratio = 1, xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  scale_x_continuous(breaks = x_breaks_n, labels = x_breaks_raw) +
  scale_y_continuous(breaks = y_breaks_n, labels = y_breaks_raw) +
  ggrepel::geom_text_repel(data = df_wide,
                           aes(x = x_n, y = y_n, label = mAL),
                           size = 2.8, color = "grey25",
                           max.overlaps = 30, min.segment.length = 0,
                           box.padding = pie_r * 1.5 + 0.01,
                           point.padding = pie_r * 1.1) +
  labs(
    x = expression(ppk23 - ppk25 ~ "(male - female contact drive)"),
    y = "Total AN09B017 input to mAL (% of mAL's total input)",
    title = "AN09B017 input composition per mAL vs ppk asymmetry"
  ) +
  theme_pub(base_size = 10) +
  theme(legend.position = "right",
        plot.title.position = "plot")

save_fig5_plot(p, "panel_B_an09_dominance_vs_ppk_asymmetry_v1", width = 10, height = 8)
print(p)
```
