---
title: "Figure 5B supp v5: AN09B017 input composition as horizontal stacked bars"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(ggplot2)
library(dplyr)
library(tidyr)
```

## What this is (v5)

Replaces the v0 pie-scatter with **horizontal stacked bars** — one bar per mAL (rows
sorted by delta_ppk, most ppk23-biased at the top, most ppk25-biased at the bottom).
Bar length = total_pct (% of mAL's total input from all AN09B017 variants combined).
Segments stacked inside each bar = fraction of that AN09B017 input contributed by
each variant (a-g). A small coloured dot to the right of each bar encodes the sign
of delta_ppk (red = ppk23-biased, blue = ppk25-biased) along with its numeric value.

- **y** = mAL subtype (ordered by delta_ppk desc)
- **x** = total_pct (% of mAL input that comes from AN09B017)
- **fill** = AN09B017 variant (a-g), each segment width = variant's share of the
  AN09B017 input to that mAL, scaled so the full stack sums to total_pct
- **right-hand dot + label** = delta_ppk (ppk23 - ppk25)

```{r build}
an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")
an09_present <- intersect(an09_types, rownames(adj.matrix))
mal_present  <- intersect(mal_subtypes, colnames(adj.matrix))

w <- as.matrix(adj.matrix[an09_present, mal_present])

df_wide <- tibble(
  mAL       = mal_present,
  total_pct = colSums(w) * 100,
  delta_ppk = mal_activation_matrix[mal_present, "ppk23"] -
              mal_activation_matrix[mal_present, "ppk25"]
)
# Attach per-variant absolute contribution to total_pct (so segments sum to total_pct)
col_tot <- colSums(w)
for (v in an09_present) {
  df_wide[[v]] <- ifelse(col_tot > 0, (w[v, ] / col_tot) * df_wide$total_pct, 0)
}
print(df_wide)
```

## Horizontal stacked bars

```{r plot, fig.width=11, fig.height=9}
# --- Palette: bright, distinct ---------------------------------------------
base_pal <- RColorBrewer::brewer.pal(8, "Dark2")
if (length(an09_present) > length(base_pal)) {
  base_pal <- c(base_pal, RColorBrewer::brewer.pal(8, "Set2"))
}
an09_palette <- setNames(base_pal[seq_along(an09_present)], an09_present)

# --- Long form for stacked bars --------------------------------------------
mal_order <- df_wide %>% arrange(delta_ppk) %>% pull(mAL)
# (lowest delta at bottom of factor levels -> because ggplot plots factor level 1 at
#  bottom, and we want ppk23-biased at TOP, we keep ascending here so that the TOP
#  row is the max delta.)

df_long <- df_wide %>%
  select(mAL, all_of(an09_present)) %>%
  pivot_longer(cols = all_of(an09_present),
               names_to = "variant", values_to = "pct_share") %>%
  mutate(mAL = factor(mAL, levels = mal_order),
         variant = factor(variant, levels = an09_present))

df_tot <- df_wide %>%
  mutate(mAL = factor(mAL, levels = mal_order),
         delta_sign = ifelse(delta_ppk >= 0, "ppk23-biased", "ppk25-biased"))

# --- Dot + label x position (a bit past max bar length) ---------------------
x_bar_max <- max(df_wide$total_pct)
x_pad     <- x_bar_max * 0.08
x_dot     <- x_bar_max + x_pad * 0.6
x_label   <- x_bar_max + x_pad * 1.4
x_right   <- x_bar_max + x_pad * 3.0   # plot x upper limit

# --- Plot -------------------------------------------------------------------
p <- ggplot() +
  # stacked bars
  geom_col(data = df_long,
           aes(x = pct_share, y = mAL, fill = variant),
           width = 0.78, color = "grey25", linewidth = 0.15) +
  # delta_ppk sign dot
  geom_point(data = df_tot,
             aes(x = x_dot, y = mAL, color = delta_sign),
             size = 3.2) +
  # delta_ppk numeric label
  geom_text(data = df_tot,
            aes(x = x_label, y = mAL,
                label = sprintf("%+.2f", delta_ppk)),
            hjust = 0, size = 3.0, color = "grey15", family = "mono") +
  scale_fill_manual(values = an09_palette, name = "AN09B017 variant") +
  scale_color_manual(
    values = c("ppk23-biased" = "#D7301F", "ppk25-biased" = "#2166AC"),
    name   = expression(Delta == ppk23 - ppk25),
    breaks = c("ppk23-biased", "ppk25-biased"),
    labels = c("ppk23 > ppk25 (male contact)",
               "ppk25 > ppk23 (female contact)")
  ) +
  scale_x_continuous(
    name   = "AN09B017 input to mAL  (% of mAL's total input)",
    limits = c(0, x_right),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_discrete(name = NULL) +
  # axis annotation: which end is which
  annotate("text",
           x = x_bar_max * 0.5,
           y = length(mal_order) + 0.9,
           label = paste0("Rows sorted by ",
                          "\u0394ppk = ppk23 \u2212 ppk25  ",
                          "(top: ppk23-biased  \u2192  bottom: ppk25-biased)"),
           hjust = 0.5, vjust = 0.5,
           size = 3.2, fontface = "italic", color = "grey25") +
  coord_cartesian(clip = "off",
                  ylim = c(0.5, length(mal_order) + 1.4)) +
  labs(title = "AN09B017 variant composition per mAL, sorted by ppk asymmetry") +
  theme_pub(base_size = 11) +
  theme(
    legend.position  = "right",
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = rel(0.95)),
    axis.text.x        = element_text(angle = 0, hjust = 0.5),
    plot.title.position = "plot",
    plot.margin      = margin(14, 10, 6, 6)
  )

save_fig5_plot(p, "panel_B_an09_dominance_vs_ppk_asymmetry_v5",
               width = 11, height = 9)

# Also drop the PDF directly into the curated "single panels" folder
single_panels_dir <- file.path(
  fig5_dir,
  "my_selection_that_makes_halfway_sense_mspecific",
  "single panels"
)
dir.create(single_panels_dir, recursive = TRUE, showWarnings = FALSE)
ggsave(file.path(single_panels_dir, "panel_B_an09_dominance_vs_ppk_asymmetry_v5.pdf"),
       plot = p + theme_pub(base_size = 11),
       width = 11, height = 9)

print(p)
```
