---
title: "Figure 5J: raw-synapse mAL -> P1 heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

## What this is

Raw-synapse-count version of `panel_J_heatmap.pdf`. Rows = mAL subtypes,
columns = P1 subtypes, cell value = signed synapse count (GABAergic mAL ->
P1 shown as negative). The regular `panel_J_heatmap.pdf` uses
input-normalized weights (fraction of each P1's total input). This version
uses absolute synapse counts so the connection magnitudes are directly
readable; it is complementary to the normalized version, not a replacement.

```{r build}
p1_subtypes <- unique(p1_all_combos$target_type)

mal_rows <- intersect(mal_subtypes, rownames(adj.matrix.raw))
p1_cols  <- intersect(p1_subtypes,  colnames(adj.matrix.raw))
mat_raw  <- as.matrix(adj.matrix.raw[mal_rows, p1_cols])

# Signed: negate rows whose mAL type is GABAergic
sign_mult <- ifelse(mal_nt[mal_rows] == "GABA", -1, 1)
mat_signed <- sweep(mat_raw, 1, sign_mult, "*")

# Fill to full mAL_subtypes x p1_subtypes (zero-pad missing)
mat_full <- matrix(0, length(mal_subtypes), length(p1_subtypes),
                   dimnames = list(mal_subtypes, p1_subtypes))
mat_full[mal_rows, p1_cols] <- mat_signed

cat("Raw-synapse signed matrix range:",
    range(mat_full), "  non-zero cells:",
    sum(mat_full != 0), "\n")

# ppk23 response polarity for rows (same as panel_J)
ppk23_vec <- mal_activation_matrix[mal_subtypes, "ppk23"]
ppk23_group <- factor(
  ifelse(ppk23_vec > 0, "ppk23-excited",
         ifelse(ppk23_vec < 0, "ppk23-inhibited", "ppk23-none")),
  levels = c("ppk23-excited", "ppk23-inhibited", "ppk23-none")
)

row_ord <- order(ppk23_group, -abs(rowSums(mat_full)))
mat_ordered <- mat_full[row_ord, , drop = FALSE]
ppk23_ordered <- ppk23_group[row_ord]

row_annot <- data.frame(
  ppk23_group       = ppk23_ordered,
  neurotransmitter  = mal_nt[rownames(mat_ordered)],
  row.names         = rownames(mat_ordered)
)
```

## Heatmap

```{r heatmap, fig.width=14, fig.height=6}
ann_colors <- list(
  ppk23_group      = c("ppk23-excited"   = "#B2182B",
                       "ppk23-inhibited" = "#2166AC",
                       "ppk23-none"      = "grey80"),
  neurotransmitter = nt_colors
)

max_abs <- max(abs(mat_ordered)); if (max_abs == 0) max_abs <- 1
breaks  <- seq(-max_abs, max_abs, length.out = 101)
colors  <- colorRampPalette(c("#2166AC", "white", "#B2182B"))(100)

n_cols <- ncol(mat_ordered)
cw <- min(22, max(10, 500 / n_cols))
fig_w <- max(12, n_cols * cw / 72 + 4)

save_heatmap_fixed_annot(
  mat        = mat_ordered,
  row_annot  = row_annot,
  col_annot  = data.frame(row.names = colnames(mat_ordered)),
  ann_colors = ann_colors,
  filename   = file.path(panel_plot_dir("panel_J_heatmap_raw", "png"),
                          "panel_J_heatmap_raw.png"),
  width = fig_w, height = 6,
  breaks = breaks, color = colors,
  cluster_rows = FALSE, cluster_cols = TRUE,
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  fontsize_row = 8, fontsize_col = 7,
  border_color = NA, cellwidth = cw, cellheight = 15,
  display_numbers = (n_cols <= 30),
  number_format = "%.0f"
)
save_heatmap_fixed_annot(
  mat        = mat_ordered,
  row_annot  = row_annot,
  col_annot  = data.frame(row.names = colnames(mat_ordered)),
  ann_colors = ann_colors,
  filename   = file.path(panel_plot_dir("panel_J_heatmap_raw", "pdf"),
                          "panel_J_heatmap_raw.pdf"),
  width = fig_w, height = 6,
  breaks = breaks, color = colors,
  cluster_rows = FALSE, cluster_cols = TRUE,
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  fontsize_row = 8, fontsize_col = 7,
  border_color = NA, cellwidth = cw, cellheight = 15,
  display_numbers = (n_cols <= 30),
  number_format = "%.0f"
)
cat("Saved raw-synapse heatmap (", nrow(mat_ordered), "x", ncol(mat_ordered), ")\n")

# Report top-5 strongest connections
top <- which(abs(mat_full) > 0, arr.ind = TRUE)
if (nrow(top) > 0) {
  top_df <- data.frame(
    mAL = rownames(mat_full)[top[, 1]],
    P1  = colnames(mat_full)[top[, 2]],
    signed_synapses = mat_full[top]
  ) %>% arrange(desc(abs(signed_synapses))) %>% head(10)
  cat("\nTop 10 strongest connections:\n")
  print(top_df)
}
```
