---
title: "Figure 5J: cVA-gain scatter per P1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
```

## What this panel shows

For each P1 courtship command neuron, we quantify how much cVA (the volatile DA1 signal) adds to its drive in two contexts:

- **x-axis**: Δ drive when cVA is added to the **female contact** signal
  `drive(DA1 + ppk23 + ppk25) - drive(ppk23 + ppk25)`
- **y-axis**: Δ drive when cVA is added to the **male contact** signal
  `drive(DA1 + ppk23) - drive(ppk23)`

Drive is the signal-flow net input predicted by `p1_all_combos` (the nonlinear propagation model). Using the nonlinear model is what makes the x vs y information non-trivial: under a purely linear channel sum, both Δs would equal the DA1 contribution and every point would lie on the y = x diagonal.

Points above the diagonal: cVA has a **bigger** effect during male contact than during female contact (cVA-gain is male-selective at that P1).
Points below the diagonal: cVA-gain is female-selective.
Points near the origin: P1 subtypes whose cVA sensitivity is negligible in either context.

```{r compute-deltas}
pick <- function(channels_str) {
  p1_all_combos %>%
    filter(channels == channels_str) %>%
    select(target_type, net_input) %>%
    deframe()
}

drive_female         <- pick("ppk25")
drive_female_plus_cva <- pick("DA1,ppk25")
drive_male           <- pick("ppk23")
drive_male_plus_cva   <- pick("DA1,ppk23")

# Align on the P1 subtype set (all should be present, but be defensive)
p1_all <- Reduce(intersect, list(names(drive_female), names(drive_female_plus_cva),
                                 names(drive_male),   names(drive_male_plus_cva)))

delta_df <- tibble(
  P1              = p1_all,
  delta_cva_female = drive_female_plus_cva[p1_all] - drive_female[p1_all],
  delta_cva_male   = drive_male_plus_cva[p1_all]   - drive_male[p1_all]
)

# Summarize
cat("Per-P1 Δ range (female context): [",
    sprintf("%.4f, %.4f", min(delta_df$delta_cva_female), max(delta_df$delta_cva_female)), "]\n")
cat("Per-P1 Δ range (male context)  : [",
    sprintf("%.4f, %.4f", min(delta_df$delta_cva_male),   max(delta_df$delta_cva_male)),   "]\n")
cat("Spearman correlation x vs y: ",
    cor(delta_df$delta_cva_female, delta_df$delta_cva_male, method = "spearman"), "\n", sep = "")
cat("Mean vertical deviation from y = x: ",
    mean(delta_df$delta_cva_male - delta_df$delta_cva_female), "\n")
```

## Scatter: cVA gain per P1

```{r plot-scatter}
# Label a handful of outliers / interesting P1s
delta_df <- delta_df %>%
  mutate(
    diagonal_dev = delta_cva_male - delta_cva_female,
    abs_total    = abs(delta_cva_male) + abs(delta_cva_female)
  )

top_label <- delta_df %>%
  arrange(desc(abs(diagonal_dev))) %>%
  slice_head(n = 8) %>%
  pull(P1)
top_strong <- delta_df %>%
  arrange(desc(abs_total)) %>%
  slice_head(n = 8) %>%
  pull(P1)
# Also label the bottom-left cluster (cVA suppresses P1 in both contexts)
bottom_left <- delta_df %>%
  filter(delta_cva_female < 0 & delta_cva_male < 0) %>%
  pull(P1)
to_label <- Reduce(union, list(top_label, top_strong, bottom_left))

ax_lim <- range(c(delta_df$delta_cva_female, delta_df$delta_cva_male))
ax_pad <- diff(ax_lim) * 0.05
ax_lim <- ax_lim + c(-ax_pad, ax_pad)

p <- ggplot(delta_df, aes(x = delta_cva_female, y = delta_cva_male)) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "grey70") +
  geom_vline(xintercept = 0, linewidth = 0.3, color = "grey70") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 3, alpha = 0.85, color = "#455A64") +
  ggrepel::geom_text_repel(
    data = delta_df %>% filter(P1 %in% to_label),
    aes(label = P1), size = 3, max.overlaps = 20, min.segment.length = 0
  ) +
  scale_x_continuous(limits = ax_lim) +
  scale_y_continuous(limits = ax_lim) +
  coord_fixed(ratio = 1, xlim = ax_lim, ylim = ax_lim) +
  labs(
    x = "Δ drive when cVA added to female contact\n[drive(DA1+ppk25) - drive(ppk25)]",
    y = "Δ drive when cVA added to male contact\n[drive(DA1+ppk23) - drive(ppk23)]",
    title = "cVA gain per P1 courtship command neuron",
    subtitle = "Above diagonal: cVA gain is larger during male contact. Below: larger during female contact."
  ) +
  theme_pub(base_size = 10) +
  theme(plot.title = element_text(size = rel(1.2)))

save_fig5_plot(p, "panel_J_cva_delta_scatter", width = 8, height = 8)
print(p)
```

```{r narrative, echo=FALSE}
save_panel_narrative("panel_J_cva_delta", sprintf(
"Panel J (cVA-gain scatter) -- per-P1 characterization
======================================================
For each of %d P1 courtship command neurons, the scatter plots two quantities
from the signal-flow model:
  x = net_input[DA1+ppk23+ppk25] - net_input[ppk23+ppk25]
  y = net_input[DA1+ppk23]       - net_input[ppk23]

Dashed line: y = x. In a strictly linear channel-sum model both deltas would
be identical (equal to DA1's contribution to P1) and every point would lie
exactly on this line. Deviation above/below the diagonal reflects the
nonlinearity of the signal-flow propagation: cVA has a context-dependent
effect on P1 drive depending on whether male or female contact pheromone
signals are simultaneously active.

Range summaries:
  delta_cva_female: [%.4f, %.4f]
  delta_cva_male  : [%.4f, %.4f]
  Spearman rho(x, y) = %.3f
  Mean(y - x) = %.4f   (positive = male-context cVA gain dominates; negative = female-context dominates)
",
nrow(delta_df),
min(delta_df$delta_cva_female), max(delta_df$delta_cva_female),
min(delta_df$delta_cva_male),   max(delta_df$delta_cva_male),
cor(delta_df$delta_cva_female, delta_df$delta_cva_male, method = "spearman"),
mean(delta_df$delta_cva_male - delta_df$delta_cva_female)
))
```
