---
title: "Figure 5A: top-K widest paths (sum and mean bottleneck)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup.R")
library(pheatmap)
```

## What this is

Two additional raw-synapse aggregations of Panel A, sitting between the
"single widest path" (A-SUPP) and "max-flow" (A-SUPP-2) extremes:

- **Sum of top-K widest bottlenecks**: pick the K paths with highest
  bottleneck per valence, sum their bottleneck synapse counts. Treats routes
  as parallel pipes. Double-counts when paths share bottleneck edges.
- **Mean of top-K widest bottlenecks**: same K paths, mean instead of sum.
  Represents "typical capacity of the strongest K routes", robust to
  parallel-route redundancy.

K = 10 for both.

```{r compute}
K <- 10

parse_path <- function(p) strsplit(p, " -> ", fixed = TRUE)[[1]]
path_bottleneck <- function(p) {
  toks <- parse_path(p)
  if (length(toks) < 2) return(NA_real_)
  synapses <- adj.matrix.raw[cbind(toks[-length(toks)], toks[-1])]
  if (any(!is.finite(synapses))) return(NA_real_)
  min(synapses)
}

cat("Computing bottlenecks for", nrow(mal_paths), "paths...\n")
mp <- mal_paths %>%
  filter(end %in% mal_subtypes) %>%
  mutate(bn = vapply(path, path_bottleneck, numeric(1)))

agg <- function(fn, label) {
  exc <- mp %>% filter(valence == "excitatory") %>%
    group_by(end, modality) %>%
    arrange(desc(bn)) %>%
    slice_head(n = K) %>%
    summarise(v = fn(bn, na.rm = TRUE), .groups = "drop") %>%
    mutate(v = ifelse(is.finite(v), v, 0)) %>%
    pivot_wider(names_from = modality, values_from = v, values_fill = 0) %>%
    column_to_rownames("end") %>% as.matrix()
  inh <- mp %>% filter(valence == "inhibitory") %>%
    group_by(end, modality) %>%
    arrange(desc(bn)) %>%
    slice_head(n = K) %>%
    summarise(v = fn(bn, na.rm = TRUE), .groups = "drop") %>%
    mutate(v = ifelse(is.finite(v), v, 0)) %>%
    pivot_wider(names_from = modality, values_from = v, values_fill = 0) %>%
    column_to_rownames("end") %>% as.matrix()

  fill <- function(m) {
    mr <- setdiff(mal_subtypes, rownames(m))
    if (length(mr)) m <- rbind(m, matrix(0, length(mr), ncol(m), dimnames = list(mr, colnames(m))))
    mc <- setdiff(modality_order, colnames(m))
    if (length(mc)) m <- cbind(m, matrix(0, nrow(m), length(mc), dimnames = list(rownames(m), mc)))
    m[mal_subtypes, modality_order, drop = FALSE]
  }
  fill(exc) - fill(inh)
}

mat_sum  <- agg(sum,  "sum")
mat_mean <- agg(mean, "mean")
cat("sum  range:", range(mat_sum),  "\n")
cat("mean range:", range(mat_mean), "\n")
```

## Heatmaps

```{r plot}
make_plot <- function(mat, name, fmt = "%.0f") {
  row_annot <- mal_annot[rownames(mat), c("neurotransmitter", "dimorphism"), drop = FALSE]
  col_annot <- data.frame(modality_group = modality_group[colnames(mat)],
                          row.names = colnames(mat))
  ann_colors <- list(
    neurotransmitter = nt_colors,
    dimorphism       = dimorphism_colors,
    modality_group   = modality_group_colors
  )
  max_abs <- max(abs(mat)); if (!is.finite(max_abs) || max_abs == 0) max_abs <- 1
  breaks <- seq(-max_abs, max_abs, length.out = 101)
  colors <- colorRampPalette(c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                                "white",
                                "#FDDBC7", "#F4A582", "#D6604D", "#B2182B"))(100)
  for (ext in c("png", "pdf")) {
    pheatmap(
      mat,
      annotation_row = row_annot, annotation_col = col_annot,
      annotation_colors = ann_colors,
      breaks = breaks, color = colors,
      cluster_rows = TRUE, cluster_cols = FALSE,
      clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
      fontsize_row = 7, fontsize_col = 10,
      border_color = NA, cellwidth = 40, cellheight = 18,
      gaps_col = c(3, 4, 5),
      display_numbers = TRUE, number_format = fmt, fontsize_number = 5,
      filename = file.path(panel_plot_dir(name, ext), paste0(name, ".", ext)),
      width = 8, height = 10, silent = TRUE
    )
  }
  cat("saved", name, "\n")
}

make_plot(mat_sum,  "panel_A_input_profiles_topK_sum",  fmt = "%.0f")
make_plot(mat_mean, "panel_A_input_profiles_topK_mean", fmt = "%.0f")
```
