---
title: "Figure 6C: Excitatory/Inhibitory Diverging Heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel C: Excitatory/Inhibitory Valence Heatmap — Visualises the excitatory vs. inhibitory balance of modality input. Side-by-side heatmaps show excitatory and inhibitory path strengths separately; a diverging heatmap shows net valence (excitatory minus inhibitory) on a red–white–blue scale."
panel_legend <- "(C) Excitatory and inhibitory input balance. Left: separate heatmaps for excitatory and inhibitory path strengths. Right: diverging heatmap of net valence (excitatory minus inhibitory), with red indicating excitatory dominance and blue indicating inhibitory dominance."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_C_description", "txt"), "panel_C_description.txt"))
```

## Overview

Panel C visualises the excitatory vs. inhibitory balance of modality input to
third-order courtship neurons. Two complementary approaches are implemented:

- **Approach 1 (side-by-side):** Separate heatmaps for excitatory and
  inhibitory path strengths, enabling direct comparison of the two valences.
- **Approach 2 (diverging):** A single heatmap of net valence (excitatory
  minus inhibitory), with a red-white-blue diverging colour scale.

Both use the same row and column annotations as Panel B.

## 1. Load row order from Panel B

If Panel B has been run, use its dendrogram ordering for row consistency.
Otherwise, cluster independently.

```{r load-row-order}
# ---------------------------------------------------------------------------
# Attempt to load the row order saved by Panel B
# ---------------------------------------------------------------------------
row_order_file <- file.path(fig6_dir, "panel_B_row_order.rds")
use_panel_b_order <- file.exists(row_order_file)

if (use_panel_b_order) {
  row_order_b <- readRDS(row_order_file)
  cat("Loaded Panel B row order:", length(row_order_b), "neurons\n")
} else {
  cat("Panel B row order not found -- will cluster independently.\n")
}
```

## 2. Prepare annotations (same as Panel B)

```{r annotations}
# ---------------------------------------------------------------------------
# Row annotation: target set membership
# ---------------------------------------------------------------------------
# Multiple color bands for target set membership
aspf_ends_c <- unique(paths.aspf$end)
aspg_ends_c <- unique(paths.aspg$end)

make_row_annot <- function(mat) {
  ends <- rownames(mat)
  data.frame(
    `aSP-f` = ifelse(ends %in% aspf_ends_c, "yes", "no"),
    `aSP-g` = ifelse(ends %in% aspg_ends_c, "yes", "no"),
    row.names = ends,
    check.names = FALSE
  )
}

# ---------------------------------------------------------------------------
# Column annotation: modality group
# ---------------------------------------------------------------------------
make_col_annot <- function(mat) {
  data.frame(
    modality_group = modality_group[colnames(mat)],
    row.names      = colnames(mat)
  )
}

# ---------------------------------------------------------------------------
# Annotation colours
# ---------------------------------------------------------------------------
ann_colors <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 3. Prepare excitatory and inhibitory matrices

```{r prepare-matrices}
# ---------------------------------------------------------------------------
# Pre-computed matrices from setup.R
# ---------------------------------------------------------------------------
mat_exc <- strength.aspf_aspg.exc
mat_inh <- strength.aspf_aspg.inh

# Ensure both matrices have the same rows and columns
all_rows <- union(rownames(mat_exc), rownames(mat_inh))
all_cols <- modality_order

# Expand matrices to full dimensions, filling missing entries with 0
expand_matrix <- function(mat, rows, cols) {
  out <- matrix(0, nrow = length(rows), ncol = length(cols),
                dimnames = list(rows, cols))
  common_r <- intersect(rownames(mat), rows)
  common_c <- intersect(colnames(mat), cols)
  out[common_r, common_c] <- mat[common_r, common_c]
  out
}

mat_exc_full <- expand_matrix(mat_exc, all_rows, all_cols)
mat_inh_full <- expand_matrix(mat_inh, all_rows, all_cols)

cat(sprintf("Excitatory matrix: %d x %d\n", nrow(mat_exc_full), ncol(mat_exc_full)))
cat(sprintf("Inhibitory matrix: %d x %d\n", nrow(mat_inh_full), ncol(mat_inh_full)))
```

## 4. Column-normalize each matrix

```{r normalize}
# ---------------------------------------------------------------------------
# Column-normalize: each modality's max = 1
# ---------------------------------------------------------------------------
col_normalize <- function(mat) {
  cm <- apply(mat, 2, max, na.rm = TRUE)
  cm[cm == 0] <- 1
  sweep(mat, 2, cm, "/")
}

# ---------------------------------------------------------------------------
# Row-normalize: each target neuron's max absolute value = 1 -> [-1, 1]
# ---------------------------------------------------------------------------
row_normalize <- function(mat) {
  row_max <- apply(abs(mat), 1, max, na.rm = TRUE)
  row_max[row_max == 0] <- 1
  sweep(mat, 1, row_max, "/")
}

# ---------------------------------------------------------------------------
# Column-normalize (diverging): each modality's max |value| = 1 -> [-1, 1]
# ---------------------------------------------------------------------------
col_normalize_diverging <- function(mat) {
  cm <- apply(abs(mat), 2, max, na.rm = TRUE)
  cm[cm == 0] <- 1
  sweep(mat, 2, cm, "/")
}

mat_exc_norm <- col_normalize(mat_exc_full)
mat_inh_norm <- col_normalize(mat_inh_full)

# Joint column-normalize: E and I share the total (E+I) column maxes
# This preserves the E/I ratio -- E + I panels visually decompose Panel B
mat_total <- mat_exc_full + mat_inh_full
col_maxes_total <- apply(mat_total, 2, max, na.rm = TRUE)
col_maxes_total[col_maxes_total == 0] <- 1
mat_exc_joint_norm <- sweep(mat_exc_full, 2, col_maxes_total, "/")
mat_inh_joint_norm <- sweep(mat_inh_full, 2, col_maxes_total, "/")

# Row-normalize E/I jointly (preserves E/I balance per neuron)
row_maxes_ei <- pmax(
  apply(mat_exc_full, 1, max, na.rm = TRUE),
  apply(mat_inh_full, 1, max, na.rm = TRUE)
)
row_maxes_ei[row_maxes_ei == 0] <- 1
mat_exc_rownorm <- sweep(mat_exc_full, 1, row_maxes_ei, "/")
mat_inh_rownorm <- sweep(mat_inh_full, 1, row_maxes_ei, "/")
```

## 5. Optionally apply Panel B row ordering

```{r apply-row-order}
# ---------------------------------------------------------------------------
# If Panel B order is available, restrict and reorder rows to match
# ---------------------------------------------------------------------------
if (use_panel_b_order) {
  shared_rows <- intersect(row_order_b, all_rows)
  extra_rows  <- setdiff(all_rows, row_order_b)
  ordered_rows <- c(shared_rows, extra_rows)  # panel B order first, then extras

  mat_exc_norm <- mat_exc_norm[ordered_rows, ]
  mat_inh_norm <- mat_inh_norm[ordered_rows, ]
  mat_exc_full <- mat_exc_full[ordered_rows, ]
  mat_inh_full <- mat_inh_full[ordered_rows, ]
  mat_exc_joint_norm <- mat_exc_joint_norm[ordered_rows, ]
  mat_inh_joint_norm <- mat_inh_joint_norm[ordered_rows, ]
  mat_exc_rownorm    <- mat_exc_rownorm[ordered_rows, ]
  mat_inh_rownorm    <- mat_inh_rownorm[ordered_rows, ]
}
```

---

# Approach 1: Side-by-side excitatory and inhibitory heatmaps

## 6. Draw excitatory heatmap

```{r heatmap-excitatory, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Use Panel B row order if available (cluster_rows = FALSE), else cluster
# ---------------------------------------------------------------------------
cluster_rows_flag <- !use_panel_b_order

row_annot_exc <- make_row_annot(mat_exc_norm)
col_annot_exc <- make_col_annot(mat_exc_norm)

hm_exc <- pheatmap(
  mat_exc_norm,
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  annotation_row           = row_annot_exc,
  annotation_col           = col_annot_exc,
  annotation_colors        = ann_colors,
  color                    = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Excitatory path strength",
  angle_col                = 45,
  silent                   = TRUE
)

hm_exc
```

## 7. Draw inhibitory heatmap

```{r heatmap-inhibitory, fig.width=8, fig.height=10}
row_annot_inh <- make_row_annot(mat_inh_norm)
col_annot_inh <- make_col_annot(mat_inh_norm)

hm_inh <- pheatmap(
  mat_inh_norm,
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  annotation_row           = row_annot_inh,
  annotation_col           = col_annot_inh,
  annotation_colors        = ann_colors,
  color                    = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Inhibitory path strength",
  angle_col                = 45,
  silent                   = TRUE
)

hm_inh
```

## 8. Save side-by-side heatmaps

```{r save-side-by-side}
# ---------------------------------------------------------------------------
# Save excitatory heatmap
# ---------------------------------------------------------------------------
save_heatmap_both_orientations(
  mat        = mat_exc_norm,
  row_annot  = make_row_annot(mat_exc_norm),
  col_annot  = make_col_annot(mat_exc_norm),
  ann_colors = ann_colors,
  name       = "panel_C_excitatory_aspf_aspg",
  width      = 8,
  height     = 10,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Excitatory path strength",
  angle_col                = 45
)

# ---------------------------------------------------------------------------
# Save inhibitory heatmap
# ---------------------------------------------------------------------------
save_heatmap_both_orientations(
  mat        = mat_inh_norm,
  row_annot  = make_row_annot(mat_inh_norm),
  col_annot  = make_col_annot(mat_inh_norm),
  ann_colors = ann_colors,
  name       = "panel_C_inhibitory_aspf_aspg",
  width      = 8,
  height     = 10,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Inhibitory path strength",
  angle_col                = 45
)
```

## 8b. Joint column-normalized E/I heatmaps

E and I normalized by the same column maxes (from total strength), so E + I
visually decomposes Panel B. Preserves the E/I ratio per modality.

```{r save-joint-colnorm}
save_heatmap_both_orientations(
  mat = mat_exc_joint_norm, row_annot = make_row_annot(mat_exc_joint_norm),
  col_annot = make_col_annot(mat_exc_joint_norm), ann_colors = ann_colors,
  name = "panel_C_excitatory_aspf_aspg", width = 8, height = 10,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Excitatory path strength (joint col-normalized)", angle_col = 45
)

save_heatmap_both_orientations(
  mat = mat_inh_joint_norm, row_annot = make_row_annot(mat_inh_joint_norm),
  col_annot = make_col_annot(mat_inh_joint_norm), ann_colors = ann_colors,
  name = "panel_C_inhibitory_aspf_aspg", width = 8, height = 10,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Inhibitory path strength (joint col-normalized)", angle_col = 45
)

# Joint col-normalized diverging: net valence / total column maxes
mat_net_joint_norm <- sweep(mat_net, 2, col_maxes_total, "/")

save_heatmap_both_orientations(
  mat = mat_net_joint_norm, row_annot = make_row_annot(mat_net_joint_norm),
  col_annot = make_col_annot(mat_net_joint_norm), ann_colors = ann_colors,
  name = "panel_C_diverging_valence_aspf_aspg", width = 8, height = 10,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = diverging_palette, breaks = seq(-1, 1, length.out = 101),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Net valence (joint col-normalized)", angle_col = 45
)
```

## 8c. Row-normalized E/I heatmaps

```{r save-rownorm-ei}
save_heatmap_both_orientations(
  mat = mat_exc_rownorm, row_annot = make_row_annot(mat_exc_rownorm),
  col_annot = make_col_annot(mat_exc_rownorm), ann_colors = ann_colors,
  name = "panel_C_excitatory_aspf_aspg", width = 8, height = 10,
  subdir = "rownorm",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Excitatory path strength (row-normalized)", angle_col = 45
)

save_heatmap_both_orientations(
  mat = mat_inh_rownorm, row_annot = make_row_annot(mat_inh_rownorm),
  col_annot = make_col_annot(mat_inh_rownorm), ann_colors = ann_colors,
  name = "panel_C_inhibitory_aspf_aspg", width = 8, height = 10,
  subdir = "rownorm",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Inhibitory path strength (row-normalized)", angle_col = 45
)
```

## 8d. Raw (unnormalized) E/I heatmaps

```{r save-raw-ei}
save_heatmap_both_orientations(
  mat = mat_exc_full, row_annot = make_row_annot(mat_exc_full),
  col_annot = make_col_annot(mat_exc_full), ann_colors = ann_colors,
  name = "panel_C_excitatory_aspf_aspg", width = 8, height = 10,
  subdir = "raw",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Excitatory path strength (unnormalized)", angle_col = 45
)

save_heatmap_both_orientations(
  mat = mat_inh_full, row_annot = make_row_annot(mat_inh_full),
  col_annot = make_col_annot(mat_inh_full), ann_colors = ann_colors,
  name = "panel_C_inhibitory_aspf_aspg", width = 8, height = 10,
  subdir = "raw",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Inhibitory path strength (unnormalized)", angle_col = 45
)
```

---

# Approach 2: Diverging net-valence heatmap

## 9. Compute net valence matrix

Net valence = excitatory strength - inhibitory strength. Positive values
indicate excitatory dominance; negative values indicate inhibitory dominance.

```{r net-valence}
# ---------------------------------------------------------------------------
# Net valence: raw (unnormalized) excitatory - inhibitory
# ---------------------------------------------------------------------------
mat_net <- mat_exc_full - mat_inh_full

cat(sprintf("Net valence matrix: %d x %d\n", nrow(mat_net), ncol(mat_net)))
cat(sprintf("Range: [%.3f, %.3f]\n", min(mat_net), max(mat_net)))
```

## 10. Row-normalize for diverging colour map

Apply row normalization so each target neuron's most extreme value becomes
+/-1, highlighting within-neuron excitatory/inhibitory balance across modalities.

```{r scale-net-valence}
# ---------------------------------------------------------------------------
# Row-normalize: each neuron's max |value| = 1 -> values in [-1, 1]
# ---------------------------------------------------------------------------
mat_net_scaled <- row_normalize(mat_net)

cat(sprintf("Row-normalized range: [%.3f, %.3f]\n", min(mat_net_scaled), max(mat_net_scaled)))
```

## 11. Draw diverging heatmap

```{r heatmap-diverging, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Diverging colour palette: blue (inhibitory) - white (balanced) - red (excitatory)
# ---------------------------------------------------------------------------
diverging_palette <- colorRampPalette(
  c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
    "#F7F7F7",
    "#FDDBC7", "#F4A582", "#D6604D", "#B2182B")
)(100)

# Fixed breaks for row-normalized data: [-1, 1]
breaks <- seq(-1, 1, length.out = 101)

row_annot_net <- make_row_annot(mat_net_scaled)
col_annot_net <- make_col_annot(mat_net_scaled)

hm_div <- pheatmap(
  mat_net_scaled,
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  annotation_row           = row_annot_net,
  annotation_col           = col_annot_net,
  annotation_colors        = ann_colors,
  color                    = diverging_palette,
  breaks                   = breaks,
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Net valence, row-normalized",
  angle_col                = 45,
  silent                   = TRUE
)

hm_div
```

## 12. Save diverging heatmap

```{r save-diverging}
save_heatmap_both_orientations(
  mat        = mat_net_scaled,
  row_annot  = make_row_annot(mat_net_scaled),
  col_annot  = make_col_annot(mat_net_scaled),
  ann_colors = ann_colors,
  name       = "panel_C_diverging_valence_aspf_aspg",
  width      = 8,
  height     = 10,
  subdir     = "rownorm",
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = diverging_palette,
  breaks                   = breaks,
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Net valence, row-normalized",
  angle_col                = 45
)
```

## 12b. Column-normalized diverging heatmap

Column normalization: each modality's most extreme net value becomes +/-1,
highlighting which modalities carry the strongest excitatory or inhibitory
signal relative to their own range.

```{r col-norm-net-valence}
# ---------------------------------------------------------------------------
# Column-normalize: each modality's max |value| = 1 -> values in [-1, 1]
# ---------------------------------------------------------------------------
mat_net_col_scaled <- col_normalize_diverging(mat_net)

cat(sprintf("Column-normalized range: [%.3f, %.3f]\n", min(mat_net_col_scaled), max(mat_net_col_scaled)))
```

```{r heatmap-diverging-colnorm, fig.width=8, fig.height=10}
hm_div_col <- pheatmap(
  mat_net_col_scaled,
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  annotation_row           = make_row_annot(mat_net_col_scaled),
  annotation_col           = make_col_annot(mat_net_col_scaled),
  annotation_colors        = ann_colors,
  color                    = diverging_palette,
  breaks                   = breaks,
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Net valence, column-normalized",
  angle_col                = 45,
  silent                   = TRUE
)

hm_div_col
```

```{r save-diverging-colnorm}
save_heatmap_both_orientations(
  mat        = mat_net_col_scaled,
  row_annot  = make_row_annot(mat_net_col_scaled),
  col_annot  = make_col_annot(mat_net_col_scaled),
  ann_colors = ann_colors,
  name       = "panel_C_diverging_valence_colnorm_aspf_aspg",
  width      = 8,
  height     = 10,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_flag,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = diverging_palette,
  breaks                   = breaks,
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Net valence, column-normalized",
  angle_col                = 45
)
```

## 12c. Raw (unnormalized) diverging heatmap

```{r save-diverging-raw}
# Raw diverging breaks need to span the actual data range
breaks_raw_div <- seq(min(mat_net), max(mat_net), length.out = 101)

save_heatmap_both_orientations(
  mat = mat_net, row_annot = make_row_annot(mat_net),
  col_annot = make_col_annot(mat_net), ann_colors = ann_colors,
  name = "panel_C_diverging_valence_aspf_aspg", width = 8, height = 10,
  subdir = "raw",
  cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_raw_div,
  border_color = NA, fontsize_row = 7, fontsize_col = 10,
  main = "Net valence (unnormalized)", angle_col = 45
)
```

---

## 13. Summary statistics

Quick summary of the excitatory/inhibitory balance across modalities.

```{r summary-stats}
# ---------------------------------------------------------------------------
# Per-modality summary of net valence
# ---------------------------------------------------------------------------
net_summary <- data.frame(
  modality     = colnames(mat_net),
  mean_net     = colMeans(mat_net),
  median_net   = apply(mat_net, 2, median),
  n_exc_dominant = colSums(mat_net > 0),
  n_inh_dominant = colSums(mat_net < 0),
  n_balanced     = colSums(mat_net == 0)
)

cat("Per-modality net valence summary:\n")
print(net_summary, row.names = FALSE)

# ---------------------------------------------------------------------------
# Per-neuron: number of modalities with excitatory vs inhibitory dominance
# ---------------------------------------------------------------------------
neuron_summary <- data.frame(
  neuron         = rownames(mat_net),
  n_exc_modalities = rowSums(mat_net > 0),
  n_inh_modalities = rowSums(mat_net < 0),
  net_total        = rowSums(mat_net)
)

cat(sprintf("\nNeurons with purely excitatory input: %d\n",
            sum(neuron_summary$n_inh_modalities == 0)))
cat(sprintf("Neurons with purely inhibitory input: %d\n",
            sum(neuron_summary$n_exc_modalities == 0)))
cat(sprintf("Neurons with mixed valence input: %d\n",
            sum(neuron_summary$n_exc_modalities > 0 & neuron_summary$n_inh_modalities > 0)))
```

---

# Extended: All 4 target sets (aSP-f + aSP-g + vAB3 targets + PPN1 targets)

## 14. Prepare extended excitatory/inhibitory matrices

```{r extended-matrices}
mat_exc_all <- strength.all.exc
mat_inh_all <- strength.all.inh

# Expand to shared dimensions
all_rows_ext <- union(rownames(mat_exc_all), rownames(mat_inh_all))

mat_exc_all_full <- expand_matrix(mat_exc_all, all_rows_ext, modality_order)
mat_inh_all_full <- expand_matrix(mat_inh_all, all_rows_ext, modality_order)

mat_exc_all_norm <- col_normalize(mat_exc_all_full)
mat_inh_all_norm <- col_normalize(mat_inh_all_full)

# Joint column-normalize (extended)
mat_total_all <- mat_exc_all_full + mat_inh_all_full
col_maxes_total_all <- apply(mat_total_all, 2, max, na.rm = TRUE)
col_maxes_total_all[col_maxes_total_all == 0] <- 1
mat_exc_all_joint_norm <- sweep(mat_exc_all_full, 2, col_maxes_total_all, "/")
mat_inh_all_joint_norm <- sweep(mat_inh_all_full, 2, col_maxes_total_all, "/")

# Row-normalize (extended)
row_maxes_ei_all <- pmax(
  apply(mat_exc_all_full, 1, max, na.rm = TRUE),
  apply(mat_inh_all_full, 1, max, na.rm = TRUE)
)
row_maxes_ei_all[row_maxes_ei_all == 0] <- 1
mat_exc_all_rownorm <- sweep(mat_exc_all_full, 1, row_maxes_ei_all, "/")
mat_inh_all_rownorm <- sweep(mat_inh_all_full, 1, row_maxes_ei_all, "/")

# ---------------------------------------------------------------------------
# Load Panel B all-targets row order for consistent ordering
# ---------------------------------------------------------------------------
row_order_all_file <- file.path(fig6_dir, "panel_B_all_row_order.rds")
use_panel_b_all_order <- file.exists(row_order_all_file)

if (use_panel_b_all_order) {
  row_order_b_all <- readRDS(row_order_all_file)
  shared_ext <- intersect(row_order_b_all, all_rows_ext)
  extra_ext  <- setdiff(all_rows_ext, row_order_b_all)
  ordered_ext <- c(shared_ext, extra_ext)

  mat_exc_all_norm <- mat_exc_all_norm[ordered_ext, ]
  mat_inh_all_norm <- mat_inh_all_norm[ordered_ext, ]
  mat_exc_all_full <- mat_exc_all_full[ordered_ext, ]
  mat_inh_all_full <- mat_inh_all_full[ordered_ext, ]
  mat_exc_all_joint_norm <- mat_exc_all_joint_norm[ordered_ext, ]
  mat_inh_all_joint_norm <- mat_inh_all_joint_norm[ordered_ext, ]
  mat_exc_all_rownorm    <- mat_exc_all_rownorm[ordered_ext, ]
  mat_inh_all_rownorm    <- mat_inh_all_rownorm[ordered_ext, ]

  cat(sprintf("Using Panel B all-targets row order: %d shared, %d extra\n",
              length(shared_ext), length(extra_ext)))
} else {
  cat("Panel B all-targets row order not found -- will cluster independently.\n")
}

cluster_rows_ext <- !use_panel_b_all_order

cat(sprintf("Extended exc matrix: %d x %d\n", nrow(mat_exc_all_norm), ncol(mat_exc_all_norm)))
cat(sprintf("Extended inh matrix: %d x %d\n", nrow(mat_inh_all_norm), ncol(mat_inh_all_norm)))
```

## 15. Extended annotations (4-band)

```{r extended-annotations}
# 4 binary bands for all target sets
make_row_annot_all <- function(mat) {
  ends <- rownames(mat)
  data.frame(
    `aSP-f` = ifelse(ends %in% unique(paths.aspf$end), "yes", "no"),
    `aSP-g` = ifelse(ends %in% unique(paths.aspg$end), "yes", "no"),
    `vAB3`  = ifelse(ends %in% unique(paths.vab3$end), "yes", "no"),
    `PPN1`  = ifelse(ends %in% unique(paths.ppn1$end), "yes", "no"),
    row.names = ends,
    check.names = FALSE
  )
}

ann_colors_all <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  `vAB3`         = c("yes" = "#4DAF4A", "no" = "white"),
  `PPN1`         = c("yes" = "#FF7F00", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 16. Extended side-by-side heatmaps

```{r extended-heatmaps, fig.width=8, fig.height=14}
hm_exc_all <- draw_heatmap_fixed_annot(
  mat_exc_all_norm,
  make_row_annot_all(mat_exc_all_norm),
  make_col_annot(mat_exc_all_norm),
  ann_colors_all,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = cluster_rows_ext,
  color                    = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Excitatory path strength (all target sets)",
  angle_col                = 45
)
```

```{r extended-heatmap-inh, fig.width=8, fig.height=14}
hm_inh_all <- draw_heatmap_fixed_annot(
  mat_inh_all_norm,
  make_row_annot_all(mat_inh_all_norm),
  make_col_annot(mat_inh_all_norm),
  ann_colors_all,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = cluster_rows_ext,
  color                    = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Inhibitory path strength (all target sets)",
  angle_col                = 45
)
```

## 17. Extended diverging heatmap

```{r extended-diverging, fig.width=8, fig.height=14}
mat_net_all <- mat_exc_all_full - mat_inh_all_full

# Row-normalize: each neuron's max |value| = 1
mat_net_all_scaled <- row_normalize(mat_net_all)

breaks_all <- seq(-1, 1, length.out = 101)

hm_div_all <- draw_heatmap_fixed_annot(
  mat_net_all_scaled,
  make_row_annot_all(mat_net_all_scaled),
  make_col_annot(mat_net_all_scaled),
  ann_colors_all,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = cluster_rows_ext,
  color                    = diverging_palette,
  breaks                   = breaks_all,
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Net valence, row-normalized (all target sets)",
  angle_col                = 45
)
```

## 18. Save extended heatmaps

```{r save-extended}
save_heatmap_both_orientations(
  mat        = mat_exc_all_norm,
  row_annot  = make_row_annot_all(mat_exc_all_norm),
  col_annot  = make_col_annot(mat_exc_all_norm),
  ann_colors = ann_colors_all,
  name       = "panel_C_excitatory_all_targets",
  width      = 8,
  height     = 14,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_ext,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Excitatory path strength (all target sets)",
  angle_col                = 45
)

save_heatmap_both_orientations(
  mat        = mat_inh_all_norm,
  row_annot  = make_row_annot_all(mat_inh_all_norm),
  col_annot  = make_col_annot(mat_inh_all_norm),
  ann_colors = ann_colors_all,
  name       = "panel_C_inhibitory_all_targets",
  width      = 8,
  height     = 14,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_ext,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Inhibitory path strength (all target sets)",
  angle_col                = 45
)

save_heatmap_both_orientations(
  mat        = mat_net_all_scaled,
  row_annot  = make_row_annot_all(mat_net_all_scaled),
  col_annot  = make_col_annot(mat_net_all_scaled),
  ann_colors = ann_colors_all,
  name       = "panel_C_diverging_all_targets",
  width      = 8,
  height     = 14,
  subdir     = "rownorm",
  cluster_rows             = cluster_rows_ext,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = diverging_palette,
  breaks                   = breaks_all,
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Net valence, row-normalized (all target sets)",
  angle_col                = 45
)
```

## 18b. Extended column-normalized diverging heatmap

```{r extended-diverging-colnorm, fig.width=8, fig.height=14}
mat_net_all_col_scaled <- col_normalize_diverging(mat_net_all)

hm_div_all_col <- draw_heatmap_fixed_annot(
  mat_net_all_col_scaled,
  make_row_annot_all(mat_net_all_col_scaled),
  make_col_annot(mat_net_all_col_scaled),
  ann_colors_all,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = cluster_rows_ext,
  color                    = diverging_palette,
  breaks                   = breaks_all,
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Net valence, column-normalized (all target sets)",
  angle_col                = 45
)
```

```{r save-extended-colnorm}
save_heatmap_both_orientations(
  mat        = mat_net_all_col_scaled,
  row_annot  = make_row_annot_all(mat_net_all_col_scaled),
  col_annot  = make_col_annot(mat_net_all_col_scaled),
  ann_colors = ann_colors_all,
  name       = "panel_C_diverging_all_targets_colnorm",
  width      = 8,
  height     = 14,
  subdir     = "separate_colnorm",
  cluster_rows             = cluster_rows_ext,
  cluster_cols             = FALSE,
  clustering_distance_rows = "euclidean",
  clustering_method        = "ward.D2",
  color                    = diverging_palette,
  breaks                   = breaks_all,
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Net valence, column-normalized (all target sets)",
  angle_col                = 45
)
```

## 18c. Joint column-normalized E/I (extended)

```{r save-extended-joint-colnorm}
save_heatmap_both_orientations(
  mat = mat_exc_all_joint_norm, row_annot = make_row_annot_all(mat_exc_all_joint_norm),
  col_annot = make_col_annot(mat_exc_all_joint_norm), ann_colors = ann_colors_all,
  name = "panel_C_excitatory_all_targets", width = 8, height = 14,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Excitatory path strength (joint col-norm, all targets)", angle_col = 45
)
save_heatmap_both_orientations(
  mat = mat_inh_all_joint_norm, row_annot = make_row_annot_all(mat_inh_all_joint_norm),
  col_annot = make_col_annot(mat_inh_all_joint_norm), ann_colors = ann_colors_all,
  name = "panel_C_inhibitory_all_targets", width = 8, height = 14,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Inhibitory path strength (joint col-norm, all targets)", angle_col = 45
)

# Joint col-normalized diverging (extended)
mat_net_all_joint_norm <- sweep(mat_net_all, 2, col_maxes_total_all, "/")

save_heatmap_both_orientations(
  mat = mat_net_all_joint_norm, row_annot = make_row_annot_all(mat_net_all_joint_norm),
  col_annot = make_col_annot(mat_net_all_joint_norm), ann_colors = ann_colors_all,
  name = "panel_C_diverging_all_targets", width = 8, height = 14,
  subdir = "joint_colnorm",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = diverging_palette, breaks = seq(-1, 1, length.out = 101),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Net valence (joint col-normalized, all targets)", angle_col = 45
)
```

## 18d. Row-normalized E/I (extended)

```{r save-extended-rownorm-ei}
save_heatmap_both_orientations(
  mat = mat_exc_all_rownorm, row_annot = make_row_annot_all(mat_exc_all_rownorm),
  col_annot = make_col_annot(mat_exc_all_rownorm), ann_colors = ann_colors_all,
  name = "panel_C_excitatory_all_targets", width = 8, height = 14,
  subdir = "rownorm",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Excitatory path strength (row-normalized, all targets)", angle_col = 45
)
save_heatmap_both_orientations(
  mat = mat_inh_all_rownorm, row_annot = make_row_annot_all(mat_inh_all_rownorm),
  col_annot = make_col_annot(mat_inh_all_rownorm), ann_colors = ann_colors_all,
  name = "panel_C_inhibitory_all_targets", width = 8, height = 14,
  subdir = "rownorm",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Inhibitory path strength (row-normalized, all targets)", angle_col = 45
)
```

## 18e. Raw E/I + raw diverging (extended)

```{r save-extended-raw}
save_heatmap_both_orientations(
  mat = mat_exc_all_full, row_annot = make_row_annot_all(mat_exc_all_full),
  col_annot = make_col_annot(mat_exc_all_full), ann_colors = ann_colors_all,
  name = "panel_C_excitatory_all_targets", width = 8, height = 14,
  subdir = "raw",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Excitatory path strength (unnormalized, all targets)", angle_col = 45
)
save_heatmap_both_orientations(
  mat = mat_inh_all_full, row_annot = make_row_annot_all(mat_inh_all_full),
  col_annot = make_col_annot(mat_inh_all_full), ann_colors = ann_colors_all,
  name = "panel_C_inhibitory_all_targets", width = 8, height = 14,
  subdir = "raw",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = colorRampPalette(c("white", "#DEEBF7", "#9ECAE1", "#3182BD"))(100),
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Inhibitory path strength (unnormalized, all targets)", angle_col = 45
)

# Raw diverging (extended)
mat_net_all <- mat_exc_all_full - mat_inh_all_full
breaks_raw_div_all <- seq(min(mat_net_all), max(mat_net_all), length.out = 101)
save_heatmap_both_orientations(
  mat = mat_net_all, row_annot = make_row_annot_all(mat_net_all),
  col_annot = make_col_annot(mat_net_all), ann_colors = ann_colors_all,
  name = "panel_C_diverging_all_targets", width = 8, height = 14,
  subdir = "raw",
  cluster_rows = cluster_rows_ext, cluster_cols = FALSE,
  clustering_distance_rows = "euclidean", clustering_method = "ward.D2",
  color = diverging_palette, breaks = breaks_raw_div_all,
  border_color = NA, fontsize_row = 5, fontsize_col = 10,
  main = "Net valence (unnormalized, all targets)", angle_col = 45
)
```
