---
title: "Figure 6G: Shared Intermediate Nodes -- Hidden Hubs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel G: Shared Intermediate Nodes — Identifies hidden integration hubs: intermediate neurons that appear in the strongest paths from multiple sensory modalities. Ranks intermediates by total path strength and Shannon entropy, shows modality strength heatmaps, source vs. target diversity scatter, and co-innervation analysis. AN neurons (AN09B017, AN05B102 and variants) are excluded."
panel_legend <- "(G) Intermediate integration hubs shared across modalities. Neurons appearing in strongest paths from multiple sensory modalities are ranked by total path strength and Shannon entropy. Heatmaps show per-modality strength profiles; scatter plots display source diversity vs. target diversity, identifying convergence-divergence hubs."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_G_description", "txt"), "panel_G_description.txt"))
```

## Overview

Panel G identifies intermediate neurons that appear in the strongest paths
from multiple sensory modalities, revealing hidden integration hubs in the
circuit. These are neurons that are neither the sensory start nor the final
target, yet they sit at the crossroads of multiple modality-specific routes.
We parse all paths, extract intermediate nodes, count how many modalities
route through each node, and weight the analysis by path strength.

## 1. Define sensory neurons to exclude

Sensory start neurons appear as intermediates in some paths but are not
meaningful integration points. We exclude them from the analysis.

```{r sensory-exclusions}
# ---------------------------------------------------------------------------
# Sensory neuron types that should be excluded from intermediate analysis.
# These are the start nodes of the various modality paths.
# ---------------------------------------------------------------------------
sensory_patterns <- c(
  "^ORN_DA1",
  "^ORN_VA1v",
  "^ORN_VA1d",
  "^JO-B",
  "^LC10",
  "^ppk23",
  "^ppk25"
)

# Build a single regex for fast matching
sensory_regex <- paste(sensory_patterns, collapse = "|")
```

## 2. Parse intermediate nodes from all paths

Split each path string on " -> " and extract everything between the first
(start) and last (end) node.

```{r parse-intermediates}
# ---------------------------------------------------------------------------
# Extract intermediate nodes from each path, together with modality and
# strength metadata.
# ---------------------------------------------------------------------------
intermediates_raw <- paths.aspf_aspg %>%
  filter(!is.na(path), !is.na(strength), strength > 0) %>%
  rowwise() %>%
  mutate(
    nodes = list(strsplit(path, " -> ", fixed = TRUE)[[1]])
  ) %>%
  ungroup() %>%
  # Only keep paths with at least 3 nodes (start, intermediate(s), end)
  filter(lengths(nodes) >= 3) %>%
  mutate(
    intermediates = purrr::map(nodes, ~ .x[2:(length(.x) - 1)])
  ) %>%
  select(path, start, end, modality, strength, valence, target_set, intermediates) %>%
  tidyr::unnest(intermediates) %>%
  rename(intermediate = intermediates)

# ---------------------------------------------------------------------------
# Remove sensory neurons from the intermediate list
# ---------------------------------------------------------------------------
intermediates_clean <- intermediates_raw %>%
  filter(!grepl(sensory_regex, intermediate),
         !grepl("^AN09B017|^AN05B102", intermediate))

message(sprintf(
  "Extracted %s intermediate occurrences across %s unique nodes (after removing %s sensory occurrences)",
  format(nrow(intermediates_clean), big.mark = ","),
  format(n_distinct(intermediates_clean$intermediate), big.mark = ","),
  format(nrow(intermediates_raw) - nrow(intermediates_clean), big.mark = ",")
))
```

## 3. Compute per-intermediate summary

For each intermediate node, compute quantitative hub metrics: total path
strength, path count, source diversity (distinct start neurons routing
through), target diversity (distinct end neurons fed), and Shannon entropy
over the modality strength distribution.

```{r intermediate-summary}
# ---------------------------------------------------------------------------
# Per intermediate x modality: summed strength
# ---------------------------------------------------------------------------
inter_mod_strength <- intermediates_clean %>%
  group_by(intermediate, modality) %>%
  summarise(
    mod_strength = sum(strength, na.rm = TRUE),
    n_paths      = n(),
    .groups      = "drop"
  )

# ---------------------------------------------------------------------------
# Source diversity & target diversity per intermediate
# ---------------------------------------------------------------------------
inter_diversity <- intermediates_clean %>%
  group_by(intermediate) %>%
  summarise(
    source_diversity = n_distinct(start),
    target_diversity = n_distinct(end),
    .groups          = "drop"
  )

# ---------------------------------------------------------------------------
# Summary per intermediate node
# ---------------------------------------------------------------------------
inter_summary <- inter_mod_strength %>%
  group_by(intermediate) %>%
  summarise(
    n_modalities   = n_distinct(modality),
    total_strength = sum(mod_strength),
    total_paths    = sum(n_paths),
    .groups        = "drop"
  ) %>%
  left_join(inter_diversity, by = "intermediate") %>%
  arrange(desc(total_strength))

# ---------------------------------------------------------------------------
# Shannon entropy over modality strength distribution per intermediate
# ---------------------------------------------------------------------------
inter_entropy <- inter_mod_strength %>%
  group_by(intermediate) %>%
  summarise(
    entropy = shannon_entropy(mod_strength),
    .groups = "drop"
  )

inter_summary <- inter_summary %>%
  left_join(inter_entropy, by = "intermediate")

# ---------------------------------------------------------------------------
# Dominant modality per intermediate (modality with highest strength)
# ---------------------------------------------------------------------------
inter_dominant <- inter_mod_strength %>%
  group_by(intermediate) %>%
  slice_max(mod_strength, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(intermediate, dominant_modality = modality)

inter_summary <- inter_summary %>%
  left_join(inter_dominant, by = "intermediate")

# ---------------------------------------------------------------------------
# Wide-format per-modality strength (for heatmap)
# ---------------------------------------------------------------------------
inter_wide <- inter_mod_strength %>%
  select(intermediate, modality, mod_strength) %>%
  pivot_wider(
    names_from  = modality,
    values_from = mod_strength,
    values_fill = 0
  )

# Ensure all modalities present
for (mod in modality_order) {
  if (!mod %in% colnames(inter_wide)) inter_wide[[mod]] <- 0
}

inter_wide <- inter_wide %>%
  left_join(inter_summary %>% select(intermediate, total_strength, entropy,
                                      source_diversity, target_diversity),
            by = "intermediate")

message(sprintf(
  "Found %d unique intermediate nodes; top node has total strength = %.1f, entropy = %.2f",
  nrow(inter_summary),
  max(inter_summary$total_strength),
  inter_summary$entropy[1]
))
```

## 4. Annotate neuron types by name patterns

Classify intermediate nodes into broad neuron type categories using name
prefixes. Categories are ordered so more specific patterns match first.

```{r annotate-types}
# ---------------------------------------------------------------------------
# Neuron type patterns: name -> regex, ordered specific-first
# ---------------------------------------------------------------------------
neuron_patterns <- c(
  # Courtship command / integration
  "aSP-f"  = "^aSP-f",
  "aSP-g"  = "^aSP-g",
  "P1"     = "^P1_",
  "pC1"    = "^pC1",
  "pC2"    = "^pC2",
  "mAL"    = "^mAL",
  # Second-order relays
  "vAB3"   = "^vAB3",
  "PPN1"   = "^PPN1",
  # Lateral horn
  "LH"     = "^LH",
  # Projection neurons (olfactory, thermo, etc.)
  "PN"     = "PN",
  # Mushroom body output
  "MBON"   = "^MBON",
  # Kenyon cells
  "KC"     = "^KC",
  # Superior brain neuropils
  "SMP"    = "^SMP",
  "SIP"    = "^SIP",
  "SLP"    = "^SLP",
  # Ventrolateral neuropils
  "AVLP"   = "^AVLP",
  "PVLP"   = "^PVLP",
  "WED"    = "^WED",
  # Central complex
  "CX"     = "^(FB|EB|PB|NO|EPG|PFN|PFL|hDelta|FC|FR)",
  # Antennal lobe local
  "AL-local" = "^(lLN|vLN|LN|v2LN|l2LN)",
  # Lobula / optic lobe
  "LC"     = "^(LC|LPLC|LPC)",
  # Descending neurons
  "DN"     = "^DN",
  # Lateral accessory lobe
  "LAL"    = "^LAL",
  # Crepine
  "CRE"    = "^CRE",
  # AOTU / anterior optic tubercle
  "AOTU"   = "^(AOTU|TuBu)",
  # Suboesophageal zone
  "SEZ"    = "^(SEZ|GNG)"
)

# ---------------------------------------------------------------------------
# Match each intermediate node against the patterns (first match wins)
# ---------------------------------------------------------------------------
inter_summary <- inter_summary %>%
  mutate(
    neuron_class = {
      classes <- character(n())
      for (i in seq_along(neuron_patterns)) {
        mask <- grepl(neuron_patterns[i], intermediate)
        classes[mask & classes == ""] <- names(neuron_patterns)[i]
      }
      classes[classes == ""] <- "other"
      classes
    }
  )

message(sprintf("Neuron class breakdown: %s",
  paste(sort(table(inter_summary$neuron_class), decreasing = TRUE),
        names(sort(table(inter_summary$neuron_class), decreasing = TRUE)),
        sep = " ", collapse = ", ")))
```

## 5. Plot 1 -- Ranked dot plot of top intermediate nodes

Show the top 40 intermediate nodes ranked by total path strength, with
point color indicating Shannon entropy (multimodal balance).

```{r ranked-dot-plot, fig.width=10, fig.height=8}
# ---------------------------------------------------------------------------
# Select top N intermediates by total_strength for display
# ---------------------------------------------------------------------------
n_top <- min(40, nrow(inter_summary))

top_nodes <- inter_summary %>%
  filter(!intermediate %in% c("AN09B017", "AN05B102")) %>%
  slice_head(n = n_top) %>%
  mutate(
    intermediate = factor(intermediate,
                          levels = rev(intermediate))
  )

# ---------------------------------------------------------------------------
# Ranked dot plot: x = total_strength, color = entropy, shape = courtship
# ---------------------------------------------------------------------------
p_ranked <- ggplot(top_nodes, aes(
  x     = total_strength,
  y     = intermediate,
  color = entropy,
  shape = neuron_class != "other"
)) +
  geom_point(size = 3.5) +
  scale_color_viridis_c(
    option = "inferno",
    name   = "Shannon\nentropy"
  ) +
  scale_shape_manual(
    values = c(`FALSE` = 16, `TRUE` = 17),
    name   = "Known courtship\nneuron",
    labels = c(`FALSE` = "No", `TRUE` = "Yes")
  ) +
  scale_x_log10(labels = scales::comma) +
  labs(
    x     = "Total path strength (log scale)",
    y     = NULL,
    title = sprintf("Top %d intermediate nodes by total path strength", n_top)
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.2, color = "grey90"),
    axis.text.y        = element_text(size = 7)
  )

p_ranked
```

```{r save-ranked}
save_fig6_plot(p_ranked, "panel_G_ranked_intermediates_aspf_aspg", width = 10, height = 8)
```

## 6. Plot 2 -- Heatmap of modality strength through top intermediate nodes

Rows are the top 30 intermediate nodes (clustered by strength profile),
columns are the 7 modalities, and colour intensity represents the summed
path strength routing through that intermediate from that modality.

```{r heatmap, fig.width=8, fig.height=9}
# ---------------------------------------------------------------------------
# Prepare heatmap matrix from top 30 nodes
# ---------------------------------------------------------------------------
n_heatmap <- min(30, nrow(inter_summary))

heatmap_nodes <- inter_summary %>%
  slice_head(n = n_heatmap) %>%
  pull(intermediate)

heatmap_mat <- inter_wide %>%
  filter(intermediate %in% heatmap_nodes) %>%
  column_to_rownames("intermediate") %>%
  select(all_of(intersect(modality_order, colnames(.)))) %>%
  as.matrix()

# Log-transform for visual clarity (many orders of magnitude)
heatmap_mat_log <- log10(heatmap_mat + 1)

# ---------------------------------------------------------------------------
# Column annotation: modality group colors
# ---------------------------------------------------------------------------
col_annot <- data.frame(
  group = modality_group[colnames(heatmap_mat_log)],
  row.names = colnames(heatmap_mat_log)
)

annot_colors <- list(group = modality_group_colors)

# ---------------------------------------------------------------------------
# Row annotation: known courtship neuron flag
# ---------------------------------------------------------------------------
row_class <- inter_summary %>%
  filter(intermediate %in% heatmap_nodes) %>%
  select(intermediate, neuron_class) %>%
  column_to_rownames("intermediate")

annot_colors$neuron_class <- c(
  # Courtship command / integration
  "aSP-f"  = "#E41A1C", "aSP-g"  = "#377EB8", "P1"    = "#984EA3",
  "pC1"    = "#FF7F00", "pC2"    = "#A65628", "mAL"   = "#4DAF4A",
  # Relays
  "vAB3"   = "#F781BF", "PPN1"   = "#999999",
  # Sensory processing
  "LH"     = "#1B9E77", "PN"     = "#D95F02", "MBON"  = "#7570B3",
  "KC"     = "#E7298A", "AL-local" = "#B3DE69",
  # Superior brain
  "SMP"    = "#66C2A5", "SIP"    = "#8DD3C7", "SLP"   = "#BEBADA",
  # Ventrolateral / optic
  "AVLP"   = "#FB8072", "PVLP"   = "#FCCDE5", "WED"   = "#BC80BD",
  "LC"     = "#80B1D3", "AOTU"   = "#FDB462",
  # Central complex / motor
  "CX"     = "#B3DE69", "DN"     = "#FC8D62", "LAL"   = "#CCEBC5",
  # Other regions
  "CRE"    = "#FFED6F", "SEZ"    = "#D9D9D9",
  "other"  = "#F0F0F0"
)

# ---------------------------------------------------------------------------
# Draw heatmap
# ---------------------------------------------------------------------------
p_heat <- pheatmap(
  heatmap_mat_log,
  cluster_cols         = FALSE,
  cluster_rows         = TRUE,
  clustering_method    = "ward.D2",
  clustering_distance_rows = "euclidean",
  color                = colorRampPalette(c("white", "#FDE725", "#35B779", "#31688E", "#440154"))(100),
  border_color         = "grey85",
  annotation_col       = col_annot,
  annotation_row       = row_class,
  annotation_colors    = annot_colors,
  fontsize_row         = 7,
  fontsize_col         = 10,
  main                 = "Modality-specific strength through top intermediate nodes\n(log10 scale)",
  silent               = TRUE
)

p_heat
```

```{r save-heatmap}
# pheatmap returns a gtable; save with ggsave-compatible approach
png(file.path(panel_plot_dir("panel_G_heatmap_intermediates_aspf_aspg", "png"), "panel_G_heatmap_intermediates_aspf_aspg.png"),
    width = 8, height = 9, units = "in", res = 300)
print(p_heat)
dev.off()

pdf(file.path(panel_plot_dir("panel_G_heatmap_intermediates_aspf_aspg", "pdf"), "panel_G_heatmap_intermediates_aspf_aspg.pdf"),
    width = 8, height = 9)
print(p_heat)
dev.off()
```

## 7. Plot 3 -- Total strength vs Shannon entropy

Scatter plot where each point is an intermediate node. X-axis shows the
total path strength (how much signal flows through), y-axis shows Shannon
entropy (how balanced the modality inputs are). Top-right quadrant nodes
are the strongest multimodal integration hubs.

```{r strength-vs-entropy, fig.width=8, fig.height=6}
# ---------------------------------------------------------------------------
# Scatter: total_strength vs entropy, sized by source_diversity
# ---------------------------------------------------------------------------
# Filter out sensory neurons
inter_nonsensory <- inter_summary %>%
  filter(total_strength > 0, !grepl(sensory_regex, intermediate))

# --- Version 1: colored by neuron type ---
# Build a palette for neuron classes
neuron_classes <- sort(unique(inter_nonsensory$neuron_class))
neuron_type_colors <- setNames(
  scales::hue_pal()(length(neuron_classes)),
  neuron_classes
)
# Make "other" grey so it doesn't dominate
if ("other" %in% names(neuron_type_colors)) neuron_type_colors["other"] <- "grey70"

p_str_ent_type <- ggplot(
  inter_nonsensory,
  aes(x = total_strength, y = entropy,
      size = source_diversity, color = neuron_class)
) +
  geom_point(alpha = 0.65, size = 2, shape = 16) +
  scale_x_log10(labels = scales::comma) +
  scale_size_continuous(
    name   = "Source div.",
    range  = c(0.8, 4.5),
    breaks = scales::pretty_breaks(4)
  ) +
  scale_color_manual(
    values = neuron_type_colors,
    name   = "Type"
  ) +
  ggrepel::geom_text_repel(
    data = inter_nonsensory %>% slice_head(n = 15),
    aes(label = intermediate),
    size = 2, max.overlaps = 20, color = "grey35",
    segment.color = "grey75", segment.size = 0.25
  ) +
  labs(
    x     = "Total strength",
    y     = "Entropy (bits)",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm")
  )

p_str_ent_type

# --- Version 2: colored by dominant modality ---
p_str_ent_mod <- ggplot(
  inter_nonsensory,
  aes(x = total_strength, y = entropy,
      size = source_diversity, color = dominant_modality)
) +
  geom_point(alpha = 0.65, size = 2, shape = 16) +
  scale_x_log10(labels = scales::comma) +
  scale_size_continuous(
    name   = "Source div.",
    range  = c(0.8, 4.5),
    breaks = scales::pretty_breaks(4)
  ) +
  scale_color_manual(
    values = modality_colors,
    name   = "Modality"
  ) +
  ggrepel::geom_text_repel(
    data = inter_nonsensory %>% slice_head(n = 15),
    aes(label = intermediate),
    size = 2, max.overlaps = 20, color = "grey35",
    segment.color = "grey75", segment.size = 0.25
  ) +
  labs(
    x     = "Total strength",
    y     = "Entropy (bits)",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm")
  )

p_str_ent_mod
```

```{r save-strength-entropy}
save_fig6_plot(p_str_ent_type, "panel_G_strength_vs_entropy_neurontype_aspf_aspg", width = 7, height = 5)
save_fig6_plot(p_str_ent_mod, "panel_G_strength_vs_entropy_modality_aspf_aspg", width = 7, height = 5)
```

## 8. Supplementary: table of top hub neurons

Print the top hub neurons with their modality counts, total strength, and
neuron class annotation for reference.

```{r hub-table}
# ---------------------------------------------------------------------------
# Top 30 hubs with annotation
# ---------------------------------------------------------------------------
hub_table <- inter_summary %>%
  slice_head(n = 30) %>%
  left_join(
    inter_wide %>% select(intermediate, all_of(intersect(modality_order, colnames(inter_wide)))),
    by = "intermediate"
  )

# Look up type info from mba for the top hubs
hub_mba_info <- mba %>%
  filter(type %in% hub_table$intermediate) %>%
  group_by(type) %>%
  summarise(
    n_bodies     = n(),
    consensus_nt = first(consensus_nt),
    .groups      = "drop"
  ) %>%
  rename(intermediate = type)

hub_table <- hub_table %>%
  left_join(hub_mba_info, by = "intermediate")

hub_table %>%
  select(intermediate, total_strength, entropy, source_diversity,
         target_diversity, total_paths, neuron_class,
         consensus_nt, n_bodies, everything()) %>%
  knitr::kable(
    digits  = 2,
    caption = "Top 30 intermediate hub neurons by total path strength"
  )
```

## 9. Source diversity vs target diversity

How broadly each intermediate draws from distinct sensory start neurons
(source diversity) versus how many distinct target neurons it feeds
(target diversity). Nodes in the top-right are true convergence-divergence
hubs that bridge many inputs to many outputs.

```{r source-vs-target-diversity, fig.width=8, fig.height=7}
# ---------------------------------------------------------------------------
# Scatter: source_diversity vs target_diversity, size = total_strength,
#          color = entropy
# ---------------------------------------------------------------------------
p_div <- ggplot(
  inter_summary %>% filter(total_strength > 0),
  aes(x = source_diversity, y = target_diversity,
      size = total_strength, color = entropy)
) +
  geom_point(alpha = 0.65, size = 2, shape = 16) +
  scale_size_continuous(
    name   = "Strength",
    range  = c(0.8, 5.5),
    trans  = "log10",
    labels = scales::comma
  ) +
  scale_color_viridis_c(option = "inferno", name = "Entropy") +
  ggrepel::geom_text_repel(
    data = inter_summary %>%
      filter(total_strength > 0) %>%
      mutate(div_score = source_diversity * target_diversity) %>%
      slice_max(div_score, n = 15),
    aes(label = intermediate),
    size = 2, max.overlaps = 20, color = "grey35",
    segment.color = "grey75", segment.size = 0.25
  ) +
  labs(
    x     = "Source diversity",
    y     = "Target diversity",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm")
  )

p_div
```

```{r save-diversity}
save_fig6_plot(p_div, "panel_G_source_vs_target_diversity_aspf_aspg", width = 7, height = 5)
```

## 10. Combined summary: top 15 hubs labeled

A clean lollipop plot of the top 15 integration hubs ranked by total path
strength, with color encoding Shannon entropy (multimodal balance).

```{r top-hubs-labeled, fig.width=9, fig.height=6}
# ---------------------------------------------------------------------------
# Top 15 hubs: rank by total_strength (already sorted)
# ---------------------------------------------------------------------------
top15 <- inter_summary %>%
  filter(!grepl("^AN", intermediate)) %>%
  slice_head(n = 15) %>%
  mutate(
    label = intermediate,
    intermediate = factor(intermediate, levels = rev(intermediate))
  )

p_top15 <- ggplot(top15, aes(x = total_strength, y = intermediate)) +
  geom_segment(
    aes(x = 0, xend = total_strength, yend = intermediate),
    color    = "grey75",
    linewidth = 0.4
  ) +
  geom_point(
    aes(size = source_diversity, color = entropy),
    alpha = 0.85
  ) +
  scale_x_continuous(labels = scales::comma) +
  scale_size_continuous(
    name   = "Source\ndiversity",
    range  = c(2, 8),
    breaks = scales::pretty_breaks(4)
  ) +
  scale_color_viridis_c(
    option = "inferno",
    name   = "Shannon\nentropy"
  ) +
  labs(
    x     = "Total path strength",
    y     = NULL,
    title = "Top 15 integration hubs by path strength"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.15, color = "grey92"),
    axis.text.y        = element_text(size = 9, face = "bold")
  )

p_top15
```

```{r save-top15}
save_fig6_plot(p_top15, "panel_G_top15_hubs_aspf_aspg", width = 9, height = 6)
```

## 11. Modality co-innervation and opposition at intermediate nodes

Which modalities tend to converge on the same intermediates (co-innervation)
versus oppose each other? We compute:
1. A **signed modality co-innervation matrix** using per-intermediate
   excitatory/inhibitory valence profiles — modalities that consistently
   excite the same nodes cluster together, while those that oppose (one
   excites what the other inhibits) show negative correlation.
2. A **modality clustering dendrogram** derived from the signed correlation.

```{r modality-coinnervation, fig.width=8, fig.height=7}
# ---------------------------------------------------------------------------
# Build per-intermediate x modality signed strength matrix
# (positive = excitatory path strength, negative = inhibitory)
# ---------------------------------------------------------------------------
inter_mod_signed <- intermediates_clean %>%
  mutate(signed_strength = ifelse(valence == "inhibitory", -strength, strength)) %>%
  group_by(intermediate, modality) %>%
  summarise(
    net_strength = sum(signed_strength, na.rm = TRUE),
    .groups      = "drop"
  )

# Pivot to wide: rows = intermediates, cols = modalities
signed_wide <- inter_mod_signed %>%
  pivot_wider(names_from = modality, values_from = net_strength, values_fill = 0)

# Ensure all modalities present
for (mod in modality_order) {
  if (!mod %in% colnames(signed_wide)) signed_wide[[mod]] <- 0
}

signed_mat <- signed_wide %>%
  column_to_rownames("intermediate") %>%
  select(all_of(intersect(modality_order, colnames(.)))) %>%
  as.matrix()

# Keep only intermediates with nonzero variance across modalities
rv <- apply(signed_mat, 1, var, na.rm = TRUE)
signed_mat <- signed_mat[rv > 0, , drop = FALSE]

# ---------------------------------------------------------------------------
# Correlation between modality columns across all intermediates
# (positive = tend to co-innervate same nodes with same sign;
#  negative = tend to oppose — one excites what the other inhibits)
# ---------------------------------------------------------------------------
mod_cor <- cor(signed_mat, method = "pearson")

cor_pal <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)

ga <- data.frame(
  group = modality_group[colnames(mod_cor)],
  row.names = colnames(mod_cor)
)

p_cocor <- pheatmap(
  mod_cor,
  display_numbers  = TRUE,
  number_format    = "%.2f",
  fontsize_number  = 10,
  color            = cor_pal,
  breaks           = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col   = ga,
  annotation_row   = ga,
  annotation_colors = list(group = modality_group_colors),
  border_color     = "white",
  cellwidth        = 55,
  cellheight       = 55,
  main             = "Modality co-innervation at intermediate nodes\n(signed strength correlation)",
  fontsize         = 11,
  silent           = TRUE
)

# Save
png(file.path(panel_plot_dir("panel_G_modality_coinnervation_aspf_aspg", "png"), "panel_G_modality_coinnervation_aspf_aspg.png"),
    width = 8, height = 7, units = "in", res = 300)
grid::grid.draw(p_cocor$gtable)
dev.off()

pdf(file.path(panel_plot_dir("panel_G_modality_coinnervation_aspf_aspg", "pdf"), "panel_G_modality_coinnervation_aspf_aspg.pdf"),
    width = 8, height = 7)
grid::grid.draw(p_cocor$gtable)
dev.off()

grid::grid.newpage()
grid::grid.draw(p_cocor$gtable)
```

```{r modality-coinnervation-unsigned, fig.width=8, fig.height=7}
# ---------------------------------------------------------------------------
# Also show UNSIGNED co-innervation: which modalities route through the
# same intermediates regardless of valence (absolute strength correlation)
# ---------------------------------------------------------------------------
abs_mat <- abs(signed_mat)
cv <- apply(abs_mat, 2, var, na.rm = TRUE)
abs_mat <- abs_mat[, cv > 0, drop = FALSE]

mod_cor_abs <- cor(abs_mat, method = "pearson")

p_coabs <- pheatmap(
  mod_cor_abs,
  display_numbers  = TRUE,
  number_format    = "%.2f",
  fontsize_number  = 10,
  color            = colorRampPalette(c("white", "#FDE725", "#35B779", "#31688E", "#440154"))(100),
  clustering_method = "complete",
  annotation_col   = ga,
  annotation_row   = ga,
  annotation_colors = list(group = modality_group_colors),
  border_color     = "white",
  cellwidth        = 55,
  cellheight       = 55,
  main             = "Modality co-routing through shared intermediates\n(absolute strength correlation)",
  fontsize         = 11,
  silent           = TRUE
)

# Save
png(file.path(panel_plot_dir("panel_G_modality_corouting_aspf_aspg", "png"), "panel_G_modality_corouting_aspf_aspg.png"),
    width = 8, height = 7, units = "in", res = 300)
grid::grid.draw(p_coabs$gtable)
dev.off()

pdf(file.path(panel_plot_dir("panel_G_modality_corouting_aspf_aspg", "pdf"), "panel_G_modality_corouting_aspf_aspg.pdf"),
    width = 8, height = 7)
grid::grid.draw(p_coabs$gtable)
dev.off()

grid::grid.newpage()
grid::grid.draw(p_coabs$gtable)
```

## Session info

```{r session-info}
sessionInfo()
```
