---
title: "Figure 6E: Pairwise Modality Co-occurrence / Correlation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")

# Additional libraries for this panel
library(ggraph)
library(tidygraph)
library(grid)
```

```{r save-and-show-pheatmap-helper}
# ---------------------------------------------------------------------------
# Helper: build pheatmap once, save PNG + PDF, display inline
# ---------------------------------------------------------------------------
save_and_show_pheatmap <- function(..., png_path, pdf_path, save_width = 9, save_height = 8) {
  p <- pheatmap(..., silent = TRUE)

  png(png_path, width = save_width, height = save_height, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(pdf_path, width = save_width, height = save_height)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(p)
}
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel E: Modality Co-occurrence / Correlation — Examines which sensory modality pairs tend to co-target the same downstream neurons. Pairwise Pearson and Spearman correlations between modality strength vectors are shown as heatmaps and a network graph. High correlation = shared targeting; negative correlation = mutual exclusivity."
panel_legend <- "(E) Pairwise modality co-occurrence analysis. Pearson and Spearman correlation heatmaps quantify the tendency of modality pairs to co-target the same neurons. Network graph visualizes significant positive correlations as edges, with edge width proportional to correlation strength."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_E_description", "txt"), "panel_E_description.txt"))
```

## Overview

This panel examines which sensory modality pairs tend to co-target the same
downstream neurons. We compute pairwise correlations between modality strength
vectors (each vector = strengths across all target neurons for one modality)
and visualise the results as heatmaps and a network graph.

High positive correlation between two modalities means that neurons receiving
strong input from one modality also tend to receive strong input from the other.
Negative correlation implies mutual exclusivity of targeting.

```{r colour-palettes}
# Diverging colour palettes (defined once, used by all analyses)
cor_colors      <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)
spearman_colors <- colorRampPalette(c("#7B3294", "#F7F7F7", "#E66101"))(100)
cosine_colors   <- colorRampPalette(c("#01665E", "#F5F5F5", "#8C510A"))(100)
co_target_colors <- colorRampPalette(c("#F7FCF5", "#00441B"))(50)
jaccard_colors   <- colorRampPalette(c("#F2F0F7", "#54278F"))(50)

# Co-targeting threshold fraction
threshold_frac <- 0.05
```

```{r compute-net-strength}
# Compute net strength matrix: excitatory - inhibitory
# Aligns rows/columns from both matrices; missing entries treated as 0
compute_net_strength <- function(strength_exc, strength_inh) {
  all_mods <- union(colnames(strength_exc), colnames(strength_inh))
  all_neurons <- union(rownames(strength_exc), rownames(strength_inh))

  mat_exc <- matrix(0, nrow = length(all_neurons), ncol = length(all_mods),
                    dimnames = list(all_neurons, all_mods))
  mat_inh <- matrix(0, nrow = length(all_neurons), ncol = length(all_mods),
                    dimnames = list(all_neurons, all_mods))

  mat_exc[rownames(strength_exc), colnames(strength_exc)] <- strength_exc
  mat_inh[rownames(strength_inh), colnames(strength_inh)] <- strength_inh

  mat_exc - mat_inh
}
```

```{r valence-helpers}
# Helper to compute and plot correlation for a given strength matrix and label
plot_valence_correlation <- function(strength_mat, valence_label, subdir = NULL) {
  # Ensure modality order
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  # Skip if too few columns or rows
  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping %s: insufficient data (%d rows, %d modalities)",
                    valence_label, nrow(mat_v), ncol(mat_v)))
    return(invisible(NULL))
  }

  # Remove rows with zero variance (no information)
  row_var <- apply(mat_v, 1, var, na.rm = TRUE)
  mat_v <- mat_v[row_var > 0, , drop = FALSE]

  # Remove modality columns with zero variance
  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping %s: fewer than 2 variable modalities after filtering", valence_label))
    return(invisible(NULL))
  }

  cor_mat <- cor(mat_v, method = "pearson")

  # Annotation for remaining modalities
  ga <- data.frame(
    group = modality_group[colnames(cor_mat)],
    row.names = colnames(cor_mat)
  )

  title_str <- sprintf("Pearson correlation (%s paths only)", valence_label)

  # Build pheatmap once, save gtable
  p <- pheatmap(
    cor_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  # Save PNG
  png(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", valence_label), "png", subdir = subdir), sprintf("panel_E_pearson_%s.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Save PDF
  pdf(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", valence_label), "pdf", subdir = subdir), sprintf("panel_E_pearson_%s.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Display inline
  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(cor_mat)
}

# Co-targeting heatmaps per valence
compute_cotarget <- function(strength_mat, valence_label, subdir = NULL) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  mod_max_v <- apply(mat_v, 2, max, na.rm = TRUE)
  thresh_v <- mod_max_v * threshold_frac
  # Guard against zero max (would make threshold 0, flagging everything)
  thresh_v[thresh_v == 0] <- .Machine$double.eps

  mat_bin_v <- sweep(mat_v, 2, thresh_v, ">") * 1
  co_mat <- t(mat_bin_v) %*% mat_bin_v

  ga <- data.frame(
    group = modality_group[colnames(co_mat)],
    row.names = colnames(co_mat)
  )

  title_str <- sprintf("Co-targeting counts (%s, threshold: %.0f%%)", valence_label, threshold_frac * 100)

  # Build pheatmap once, save gtable
  p <- pheatmap(
    co_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  # Save PNG
  png(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s", valence_label), "png", subdir = subdir), sprintf("panel_E_cotarget_%s.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Save PDF
  pdf(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s", valence_label), "pdf", subdir = subdir), sprintf("panel_E_cotarget_%s.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Display inline
  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(co_mat)
}
```

```{r cosine-helper}
# Cosine similarity heatmap for a given strength matrix and label
plot_cosine <- function(strength_mat, valence_label, subdir = NULL) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping cosine %s: insufficient data", valence_label))
    return(invisible(NULL))
  }

  # Remove zero-variance columns
  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping cosine %s: fewer than 2 variable modalities", valence_label))
    return(invisible(NULL))
  }

  col_norms <- sqrt(colSums(mat_v^2))
  col_norms[col_norms == 0] <- 1
  mat_normed <- sweep(mat_v, 2, col_norms, "/")
  cosine_mat <- t(mat_normed) %*% mat_normed

  ga <- data.frame(
    group = modality_group[colnames(cosine_mat)],
    row.names = colnames(cosine_mat)
  )

  save_and_show_pheatmap(
    cosine_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cosine_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Cosine similarity (%s)", valence_label),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", valence_label), "png", subdir = subdir),
                         sprintf("panel_E_cosine_%s.png", valence_label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", valence_label), "pdf", subdir = subdir),
                         sprintf("panel_E_cosine_%s.pdf", valence_label)),
    save_width = 9,
    save_height = 8
  )

  invisible(cosine_mat)
}
```

```{r run-correlation-suite}
# ---------------------------------------------------------------------------
# run_correlation_suite: encapsulates ALL analysis for one strength matrix
# ---------------------------------------------------------------------------
run_correlation_suite <- function(mat, label, subdir,
                                  strength_exc = NULL, strength_inh = NULL,
                                  threshold_frac = 0.05,
                                  title_suffix = "") {

  # --- Group annotation from this matrix's columns ---
  group_annotation <- data.frame(
    group = modality_group[colnames(mat)],
    row.names = colnames(mat)
  )

  # ===== (a) Pearson correlation heatmap =====
  cor_pearson <- cor(mat, method = "pearson")

  save_and_show_pheatmap(
    cor_pearson,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Pearson correlation %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_pearson_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_pearson_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (b) Spearman rank correlation heatmap =====
  cor_spearman <- cor(mat, method = "spearman")

  save_and_show_pheatmap(
    cor_spearman,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = spearman_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Spearman rank correlation %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_spearman_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_spearman_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_spearman_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_spearman_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (b2) Cosine similarity heatmap =====
  # Unlike Pearson, cosine does NOT mean-center: sensitive to baseline level
  col_norms <- sqrt(colSums(mat^2))
  col_norms[col_norms == 0] <- 1
  mat_normed <- sweep(mat, 2, col_norms, "/")
  cosine_mat <- t(mat_normed) %*% mat_normed

  save_and_show_pheatmap(
    cosine_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cosine_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Cosine similarity %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cosine_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cosine_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (c) Co-targeting counts (threshold-based) =====
  mod_max <- apply(mat, 2, max, na.rm = TRUE)
  thresholds <- mod_max * threshold_frac
  thresholds[thresholds == 0] <- .Machine$double.eps

  mat_binary <- sweep(mat, 2, thresholds, ">") * 1
  co_target_mat <- t(mat_binary) %*% mat_binary

  cat(sprintf("Neurons with non-trivial input per modality (%s):\n", label))
  print(diag(co_target_mat))

  save_and_show_pheatmap(
    co_target_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting counts %s (threshold: %.0f%%)", title_suffix, threshold_frac * 100),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (c1b) Co-targeting counts (threshold: % of total strength) =====
  mod_sum <- colSums(abs(mat))
  thresholds_total <- mod_sum * threshold_frac
  thresholds_total[thresholds_total == 0] <- .Machine$double.eps

  mat_binary_total <- sweep(mat, 2, thresholds_total, ">") * 1
  co_target_mat_total <- t(mat_binary_total) %*% mat_binary_total

  cat(sprintf("Neurons above threshold (%%total) per modality (%s):\n", label))
  print(diag(co_target_mat_total))

  save_and_show_pheatmap(
    co_target_mat_total,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting counts %s (threshold: %.0f%% of total)", title_suffix, threshold_frac * 100),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (c1c) Column-normalized version of total-based co-targeting =====
  diag_counts_total <- diag(co_target_mat_total)
  diag_counts_total[diag_counts_total == 0] <- 1
  co_target_colnorm_total <- sweep(co_target_mat_total, 2, diag_counts_total, "/")

  save_and_show_pheatmap(
    co_target_colnorm_total,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting col-norm %s (threshold: %.0f%% of total)", title_suffix, threshold_frac * 100),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (c1d) Jaccard similarity (total-based threshold) =====
  n_mods_total <- ncol(mat_binary_total)
  jaccard_mat_total <- matrix(0, nrow = n_mods_total, ncol = n_mods_total,
                              dimnames = list(colnames(mat_binary_total), colnames(mat_binary_total)))

  for (i in seq_len(n_mods_total)) {
    for (j in seq_len(n_mods_total)) {
      both <- sum(mat_binary_total[, i] == 1 & mat_binary_total[, j] == 1)
      either <- sum(mat_binary_total[, i] == 1 | mat_binary_total[, j] == 1)
      jaccard_mat_total[i, j] <- ifelse(either > 0, both / either, 0)
    }
  }

  save_and_show_pheatmap(
    jaccard_mat_total,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = jaccard_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Jaccard similarity %s (threshold: %.0f%% of total)", title_suffix, threshold_frac * 100),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_jaccard_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_jaccard_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_jaccard_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_jaccard_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (c2) Column-normalized co-targeting (fraction of modality j's targets) =====
  diag_counts <- diag(co_target_mat)
  diag_counts[diag_counts == 0] <- 1  # avoid division by zero
  co_target_colnorm <- sweep(co_target_mat, 2, diag_counts, "/")

  save_and_show_pheatmap(
    co_target_colnorm,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting (column-normalized) %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (d) Jaccard similarity =====
  n_mods <- ncol(mat_binary)
  jaccard_mat <- matrix(0, nrow = n_mods, ncol = n_mods,
                        dimnames = list(colnames(mat_binary), colnames(mat_binary)))

  for (i in seq_len(n_mods)) {
    for (j in seq_len(n_mods)) {
      both <- sum(mat_binary[, i] == 1 & mat_binary[, j] == 1)
      either <- sum(mat_binary[, i] == 1 | mat_binary[, j] == 1)
      jaccard_mat[i, j] <- ifelse(either > 0, both / either, 0)
    }
  }

  save_and_show_pheatmap(
    jaccard_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = jaccard_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Jaccard similarity %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_jaccard_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_jaccard_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_jaccard_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_jaccard_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ===== (e) Correlation network graphs =====

  # Build edge list from upper triangle of Pearson correlation
  edge_df <- expand.grid(
    from = rownames(cor_pearson),
    to = colnames(cor_pearson),
    stringsAsFactors = FALSE
  ) %>%
    filter(from < to) %>%
    rowwise() %>%
    mutate(
      correlation = cor_pearson[from, to],
      abs_cor = abs(correlation),
      sign = ifelse(correlation >= 0, "positive", "negative")
    ) %>%
    ungroup()

  # Node metadata (built from this matrix's columns)
  mods_in_mat <- colnames(mat)
  node_df <- data.frame(
    name = mods_in_mat,
    group = modality_group[mods_in_mat],
    stringsAsFactors = FALSE
  )

  # Full network (all edges)
  graph <- tbl_graph(nodes = node_df, edges = edge_df, directed = FALSE)

  p_network <- ggraph(graph, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign, alpha = abs_cor)
    ) +
    scale_edge_width_continuous(range = c(0.3, 4), name = "|Correlation|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
    geom_node_point(aes(color = group), size = 12) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Modality correlation network %s (Pearson)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

  print(p_network)
  save_fig6_plot(p_network, sprintf("panel_E_correlation_network_%s", label),
                 width = 9, height = 8, subdir = subdir)

  # Thresholded network (|r| > 0.1)
  edge_df_thresh <- edge_df %>% filter(abs_cor > 0.1)

  graph_thresh <- tbl_graph(nodes = node_df, edges = edge_df_thresh, directed = FALSE)

  p_network_thresh <- ggraph(graph_thresh, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign),
      alpha = 0.8
    ) +
    geom_edge_link(
      aes(label = sprintf("%.2f", correlation)),
      angle_calc = "along",
      label_size = 2.5,
      label_dodge = unit(3, "mm"),
      alpha = 0
    ) +
    scale_edge_width_continuous(range = c(0.5, 4), name = "|Correlation|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    geom_node_point(aes(color = group), size = 14) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Modality correlation network %s (|r| > 0.1)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

  print(p_network_thresh)
  save_fig6_plot(p_network_thresh, sprintf("panel_E_correlation_network_%s_thresholded", label),
                 width = 9, height = 8, subdir = subdir)

  # ===== (e2) Averaged-metrics network (mean of Spearman + Cosine + Jaccard) =====
  avg_mat <- (cor_spearman + cosine_mat + jaccard_mat) / 3

  edge_df_avg <- expand.grid(
    from = rownames(avg_mat),
    to = colnames(avg_mat),
    stringsAsFactors = FALSE
  ) %>%
    filter(from < to) %>%
    rowwise() %>%
    mutate(
      correlation = avg_mat[from, to],
      abs_cor = abs(correlation),
      sign = ifelse(correlation >= 0, "positive", "negative")
    ) %>%
    ungroup()

  graph_avg <- tbl_graph(nodes = node_df, edges = edge_df_avg, directed = FALSE)

  p_network_avg <- ggraph(graph_avg, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign, alpha = abs_cor)
    ) +
    scale_edge_width_continuous(range = c(0.3, 4), name = "|Similarity|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
    geom_node_point(aes(color = group), size = 12) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Averaged similarity network %s\n(mean of Spearman + Cosine + Jaccard)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 12)
    )

  print(p_network_avg)
  save_fig6_plot(p_network_avg, sprintf("panel_E_avg_network_%s", label),
                 width = 9, height = 8, subdir = subdir)

  # Thresholded averaged network (score > 0.1)
  edge_df_avg_thresh <- edge_df_avg %>% filter(abs_cor > 0.1)

  graph_avg_thresh <- tbl_graph(nodes = node_df, edges = edge_df_avg_thresh, directed = FALSE)

  p_network_avg_thresh <- ggraph(graph_avg_thresh, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign),
      alpha = 0.8
    ) +
    geom_edge_link(
      aes(label = sprintf("%.2f", correlation)),
      angle_calc = "along",
      label_size = 2.5,
      label_dodge = unit(3, "mm"),
      alpha = 0
    ) +
    scale_edge_width_continuous(range = c(0.5, 4), name = "|Similarity|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    geom_node_point(aes(color = group), size = 14) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Averaged similarity network %s (score > 0.1)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 12)
    )

  print(p_network_avg_thresh)
  save_fig6_plot(p_network_avg_thresh, sprintf("panel_E_avg_network_%s_thresholded", label),
                 width = 9, height = 8, subdir = subdir)

  # ===== (f) Valence-split analysis (if provided) =====
  if (!is.null(strength_exc)) {
    mat_exc <- strength_exc[, intersect(modality_order, colnames(strength_exc))]
    plot_valence_correlation(mat_exc, sprintf("excitatory_%s", label), subdir = subdir)
    compute_cotarget(mat_exc, sprintf("excitatory_%s", label), subdir = subdir)
    plot_cosine(mat_exc, sprintf("excitatory_%s", label), subdir = subdir)
  }

  if (!is.null(strength_inh)) {
    mat_inh <- strength_inh[, intersect(modality_order, colnames(strength_inh))]
    plot_valence_correlation(mat_inh, sprintf("inhibitory_%s", label), subdir = subdir)
    compute_cotarget(mat_inh, sprintf("inhibitory_%s", label), subdir = subdir)
    plot_cosine(mat_inh, sprintf("inhibitory_%s", label), subdir = subdir)
  }

  # ===== Return summary objects =====
  invisible(list(
    cor_pearson  = cor_pearson,
    cor_spearman = cor_spearman,
    co_target_mat = co_target_mat
  ))
}
```

---

## 1-6. aSP-f + aSP-g targets

```{r aspf-aspg, fig.width=9, fig.height=8}
mat <- strength.aspf_aspg[, intersect(modality_order, colnames(strength.aspf_aspg))]
results_aspf_aspg <- run_correlation_suite(
  mat, label = "aspf_aspg", subdir = "aspf_aspg",
  strength_exc = strength.aspf_aspg.exc, strength_inh = strength.aspf_aspg.inh,
  title_suffix = "(aSP-f + aSP-g targets)"
)
```

```{r aspf-aspg-net-strength, fig.width=9, fig.height=8}
## Net strength (excitatory - inhibitory)

strength.aspf_aspg.net <- compute_net_strength(strength.aspf_aspg.exc, strength.aspf_aspg.inh)
mat_net <- strength.aspf_aspg.net[, intersect(modality_order, colnames(strength.aspf_aspg.net))]
results_aspf_aspg_net <- run_correlation_suite(
  mat_net, label = "aspf_aspg", subdir = "aspf_aspg/net_strength",
  title_suffix = "(aSP-f + aSP-g, net strength)"
)
```

### Summary statistics (aSP-f + aSP-g)

```{r aspf-aspg-summary}
print_correlation_summary <- function(results, tag) {
  cor_pearson  <- results$cor_pearson
  cor_spearman <- results$cor_spearman
  co_target_mat <- results$co_target_mat

  upper_idx <- upper.tri(cor_pearson)
  cor_vals <- cor_pearson[upper_idx]
  pair_labels <- outer(rownames(cor_pearson), colnames(cor_pearson), paste, sep = " - ")[upper_idx]

  cor_summary <- data.frame(
    pair = pair_labels,
    pearson = cor_vals,
    spearman = cor_spearman[upper_idx]
  ) %>%
    arrange(desc(abs(pearson)))

  cat(sprintf("\n=== Pearson Correlation Summary (%s) ===\n", tag))
  cat(sprintf("Number of modality pairs: %d\n", length(cor_vals)))
  cat(sprintf("Mean |Pearson r|: %.3f\n", mean(abs(cor_vals))))
  cat(sprintf("Max positive: %.3f\n", max(cor_vals)))
  cat(sprintf("Max negative: %.3f\n", min(cor_vals)))
  cat(sprintf("Pairs with |r| > 0.3: %d\n", sum(abs(cor_vals) > 0.3)))
  cat("\nAll pairwise correlations (sorted by |Pearson|):\n")
  print(knitr::kable(cor_summary, digits = 3))

  cat(sprintf("\n=== Co-targeting Summary (%s) ===\n", tag))
  cat(sprintf("Threshold: %.0f%% of modality max strength\n", threshold_frac * 100))
  cat("\nNeurons above threshold per modality:\n")
  print(diag(co_target_mat))

  co_upper <- co_target_mat[upper.tri(co_target_mat)]
  co_labels <- outer(rownames(co_target_mat), colnames(co_target_mat), paste, sep = " & ")[upper.tri(co_target_mat)]
  co_summary <- data.frame(pair = co_labels, shared_neurons = co_upper) %>%
    arrange(desc(shared_neurons))
  cat("\nTop co-targeting pairs:\n")
  print(knitr::kable(co_summary))
}

print_correlation_summary(results_aspf_aspg, "aSP-f + aSP-g")
print_correlation_summary(results_aspf_aspg_net, "aSP-f + aSP-g, net strength")
```

---

## 7. vAB3 + PPN1 targets

Repeat all correlation and co-targeting analyses for targets reached via
vAB3 and PPN1 paths only.

```{r vab3-ppn1-setup}
# Compute strength matrices for vAB3 + PPN1 targets
paths.vab3_ppn1 <- bind_rows(paths.vab3, paths.ppn1)

strength.vab3_ppn1     <- compute_strength_matrix(paths.vab3_ppn1)
strength.vab3_ppn1.exc <- compute_valence_strength(paths.vab3_ppn1, "excitatory")
strength.vab3_ppn1.inh <- compute_valence_strength(paths.vab3_ppn1, "inhibitory")

cat(sprintf("vAB3+PPN1 targets: %d neurons x %d modalities\n",
            nrow(strength.vab3_ppn1),
            ncol(strength.vab3_ppn1)))
```

```{r vab3-ppn1, fig.width=9, fig.height=8}
mat_vp <- strength.vab3_ppn1[, intersect(modality_order, colnames(strength.vab3_ppn1))]
results_vab3_ppn1 <- run_correlation_suite(
  mat_vp, label = "vab3_ppn1", subdir = "vab3_ppn1",
  strength_exc = strength.vab3_ppn1.exc, strength_inh = strength.vab3_ppn1.inh,
  title_suffix = "(vAB3+PPN1 targets)"
)
```

```{r vab3-ppn1-net-strength, fig.width=9, fig.height=8}
strength.vab3_ppn1.net <- compute_net_strength(strength.vab3_ppn1.exc, strength.vab3_ppn1.inh)
mat_vp_net <- strength.vab3_ppn1.net[, intersect(modality_order, colnames(strength.vab3_ppn1.net))]
results_vab3_ppn1_net <- run_correlation_suite(
  mat_vp_net, label = "vab3_ppn1", subdir = "vab3_ppn1/net_strength",
  title_suffix = "(vAB3+PPN1, net strength)"
)
```

### Summary statistics (vAB3 + PPN1)

```{r vab3-ppn1-summary}
print_correlation_summary(results_vab3_ppn1, "vAB3+PPN1")
print_correlation_summary(results_vab3_ppn1_net, "vAB3+PPN1, net strength")
```

---

## 8. All targets (aSP-f + aSP-g + vAB3 + PPN1)

Repeat all analyses using the full set of targets from all four path sets.

```{r all-targets-setup}
cat(sprintf("All targets: %d neurons x %d modalities\n", nrow(strength.all), ncol(strength.all)))
```

```{r all-targets, fig.width=9, fig.height=8}
mat_all <- strength.all[, intersect(modality_order, colnames(strength.all))]
results_all <- run_correlation_suite(
  mat_all, label = "all_targets", subdir = "all_targets",
  strength_exc = strength.all.exc, strength_inh = strength.all.inh,
  title_suffix = "(all targets)"
)
```

```{r all-targets-net-strength, fig.width=9, fig.height=8}
strength.all.net <- compute_net_strength(strength.all.exc, strength.all.inh)
mat_all_net <- strength.all.net[, intersect(modality_order, colnames(strength.all.net))]
results_all_net <- run_correlation_suite(
  mat_all_net, label = "all_targets", subdir = "all_targets/net_strength",
  title_suffix = "(all targets, net strength)"
)
```

### Summary statistics (all targets)

```{r all-targets-summary}
print_correlation_summary(results_all, "all targets")
print_correlation_summary(results_all_net, "all targets, net strength")
```

```{r session-info}
sessionInfo()
```
