---
title: "Figure 6J: Strongest Path Valence Pattern Comparison Across Modalities"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 10, fig.height = 8)
source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel J: Valence Pattern Comparison — Compares neurotransmitter sequence patterns of the single strongest path per (modality, target) pair. Extracts E/I identity of each hop to form valence pattern strings (e.g. E_I_E) and asks whether different modalities reach the same target via the same or different neurotransmitter sequences."
panel_legend <- "(J) Neurotransmitter valence patterns along strongest paths. For each modality-target pair, the single strongest path is decomposed into its hop-by-hop excitatory/inhibitory (E/I) sequence. Heatmaps show the frequency of each valence pattern across modalities, revealing conserved or divergent signalling motifs."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_J_description", "txt"), "panel_J_description.txt"))
```

## Overview

This panel compares the **neurotransmitter sequence patterns** of the strongest
single path from each sensory modality to each 3rd-order target neuron. For
each (modality, target) pair we take only the single strongest path, extract
the E/I identity of each hop (excitatory or inhibitory based on presynaptic NT),
and form a "valence pattern" string (e.g. `E_I_E`). The key question is:
**do different modalities reach the same target neuron via the same or different
neurotransmitter sequences?**

The analysis is repeated for each target set, with outputs saved into subfolders.

## 1. Load row order from Panel B

```{r load-row-order}
# Use Panel B row order for consistent ordering across panels
row_order_file <- file.path(fig6_dir, "panel_B_row_order.rds")
use_ref_order <- file.exists(row_order_file)

if (use_ref_order) {
  row_order_ref <- readRDS(row_order_file)
  cat("Loaded Panel B row order:", length(row_order_ref), "neurons\n")
} else {
  cat("Panel B row order not found -- will cluster independently.\n")
}

# Row normalization: scale each row by its max absolute value -> [-1, 1]
row_normalize <- function(mat) {
  row_max <- apply(abs(mat), 1, max, na.rm = TRUE)
  row_max[row_max == 0] <- 1
  sweep(mat, 1, row_max, "/")
}
```

## 2. Define helper functions and target sets

```{r define-helpers-and-targets}
# Build NT lookup from mba
nt_lookup <- mba %>%
  select(type, consensus_nt) %>%
  distinct() %>%
  group_by(type) %>%
  summarise(nt = first(consensus_nt), .groups = "drop")
nt_by_type <- setNames(nt_lookup$nt, nt_lookup$type)

inhibitory_nts <- c("gaba", "glycine", "histamine")

# Extract valence pattern from a path string
# For path "A -> B -> C -> D": NT of A,B,C determines E/I at each hop
get_valence_pattern <- function(path_str) {
  nodes <- strsplit(path_str, " -> ", fixed = TRUE)[[1]]
  transmitting <- nodes[-length(nodes)]
  nts <- nt_by_type[transmitting]
  pattern <- ifelse(nts %in% inhibitory_nts, "I", "E")
  paste(pattern, collapse = "_")
}

add_valence_pattern <- function(df) {
  df %>%
    filter(!is.na(path)) %>%
    rowwise() %>%
    mutate(valence_pattern = get_valence_pattern(path)) %>%
    ungroup()
}

# aSP-f / aSP-g end neurons (used for row annotations in sections 12-13)
aspf_ends <- unique(paths.aspf$end)
aspg_ends <- unique(paths.aspg$end)

# Define target sets to iterate over
target_sets <- list(
  "aspf_aspg"   = paths.aspf_aspg,
  "aspf"        = paths.aspf,
  "aspg"        = paths.aspg,
  "vab3"        = paths.vab3,
  "ppn1"        = paths.ppn1,
  "vab3_ppn1"   = bind_rows(paths.vab3, paths.ppn1),
  "all_targets" = paths.all
)
```

## 3-17. Looped analysis across target sets

Sections 3 through 17 are run once per target set. Outputs are saved into
subfolders named by target set (e.g. `aspf_aspg/`, `vab3/`, `all_targets/`).

```{r run-all-sets, fig.width=14, fig.height=14, results='asis'}
library(ggraph)
library(tidygraph)

for (suffix in names(target_sets)) {

  current_paths <- target_sets[[suffix]]

  cat(sprintf("\n\n---\n\n## Target set: %s\n\n", suffix))
  message(sprintf("=== Processing target set: %s ===", suffix))

  # --------------------------------------------------------------------------
  # 2. Derive neurotransmitter sequence for each path
  # --------------------------------------------------------------------------
  paths_patterned <- add_valence_pattern(current_paths)
  cat(sprintf("[%s] Total paths with patterns: %d\n", suffix, nrow(paths_patterned)))
  cat(sprintf("[%s] Unique valence patterns: %d\n", suffix, n_distinct(paths_patterned$valence_pattern)))

  # --------------------------------------------------------------------------
  # 3. Select absolute strongest single path per (modality, target)
  # --------------------------------------------------------------------------
  # Strongest per (modality, end, valence) -- for signed strength later
  strongest_by_valence <- paths_patterned %>%
    group_by(modality, end, valence) %>%
    slice_max(strength, n = 1, with_ties = FALSE) %>%
    ungroup()

  # For each (modality, end) pair, keep only THE single strongest path
  absolute_strongest <- paths_patterned %>%
    group_by(modality, end) %>%
    slice_max(strength, n = 1, with_ties = FALSE) %>%
    ungroup()

  cat(sprintf("[%s] Absolute strongest paths: %d\n", suffix, nrow(absolute_strongest)))
  cat(sprintf("[%s]   Excitatory: %d (%.0f%%)\n", suffix,
              sum(absolute_strongest$valence == "excitatory"),
              100 * mean(absolute_strongest$valence == "excitatory")))
  cat(sprintf("[%s]   Inhibitory: %d (%.0f%%)\n", suffix,
              sum(absolute_strongest$valence == "inhibitory"),
              100 * mean(absolute_strongest$valence == "inhibitory")))

  # --------------------------------------------------------------------------
  # 4. Valence tile map: target neuron x modality
  # --------------------------------------------------------------------------
  valence_colors <- c("excitatory" = "#B2182B", "inhibitory" = "#2166AC", "none" = "grey95")

  # Build binary valence matrix for clustering (1 = excitatory, 0 = inhibitory)
  valence_wide <- absolute_strongest %>%
    distinct(end, modality, .keep_all = TRUE) %>%
    mutate(val_num = ifelse(valence == "excitatory", 1, 0)) %>%
    select(end, modality, val_num) %>%
    pivot_wider(names_from = modality, values_from = val_num, values_fill = 0.5) %>%
    column_to_rownames("end")
  valence_wide <- as.matrix(valence_wide[, intersect(modality_order, colnames(valence_wide))])

  # Cluster rows and columns
  row_hc <- hclust(dist(valence_wide, method = "binary"), method = "ward.D2")
  col_hc <- hclust(dist(t(valence_wide), method = "binary"), method = "ward.D2")

  target_order <- rownames(valence_wide)[row_hc$order]
  mod_order_clust <- colnames(valence_wide)[col_hc$order]

  # Convert to long for ggplot with clustered order
  valence_long <- absolute_strongest %>%
    select(end, modality, valence, strength) %>%
    mutate(
      end      = factor(end, levels = target_order),
      modality = factor(modality, levels = mod_order_clust)
    )

  p_tile <- ggplot(valence_long, aes(x = modality, y = end, fill = valence)) +
    geom_tile(color = "white", linewidth = 0.3) +
    scale_fill_manual(values = valence_colors, name = "Overall\nvalence") +
    labs(
      title = sprintf("Overall valence of strongest path: target neuron x modality (%s)", suffix),
      subtitle = "Red = excitatory, Blue = inhibitory (clustered by row & column)",
      x = "Sensory modality",
      y = "Target neuron"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.y = element_text(size = 5),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
      panel.grid = element_blank(),
      legend.position = "right"
    )

  print(p_tile)
  save_fig6_plot(p_tile, "panel_J_valence_tile_map", width = 12, height = 14, subdir = suffix)

  # --------------------------------------------------------------------------
  # 5. Valence conservation across modalities
  # --------------------------------------------------------------------------
  conservation <- absolute_strongest %>%
    group_by(end) %>%
    summarise(
      n_modalities = n_distinct(modality),
      n_excitatory = sum(valence == "excitatory"),
      n_inhibitory = sum(valence == "inhibitory"),
      pct_excitatory = round(100 * mean(valence == "excitatory"), 1),
      category = case_when(
        all(valence == "excitatory") ~ "all excitatory",
        all(valence == "inhibitory") ~ "all inhibitory",
        TRUE ~ "mixed"
      ),
      .groups = "drop"
    )

  # Bar chart of categories
  cat_counts <- conservation %>% count(category)

  p_cons <- ggplot(cat_counts, aes(x = category, y = n, fill = category)) +
    geom_col(color = "grey30", linewidth = 0.3) +
    scale_fill_manual(values = c("all excitatory" = "#B2182B",
                                  "all inhibitory" = "#2166AC",
                                  "mixed" = "#999999"),
                      guide = "none") +
    geom_text(aes(label = n), vjust = -0.3, size = 5) +
    labs(
      title = sprintf("Valence conservation (%s)", suffix),
      x = NULL, y = "Number of target neurons"
    ) +
    theme_minimal(base_size = 13)

  print(p_cons)
  save_fig6_plot(p_cons, "panel_J_valence_conservation", width = 8, height = 5, subdir = suffix)

  cat(sprintf("[%s] All excitatory: %d (%.0f%%)\n", suffix,
              sum(conservation$category == "all excitatory"),
              100 * mean(conservation$category == "all excitatory")))
  cat(sprintf("[%s] All inhibitory: %d (%.0f%%)\n", suffix,
              sum(conservation$category == "all inhibitory"),
              100 * mean(conservation$category == "all inhibitory")))
  cat(sprintf("[%s] Mixed: %d (%.0f%%)\n", suffix,
              sum(conservation$category == "mixed"),
              100 * mean(conservation$category == "mixed")))

  # Mixed detail
  mixed_targets <- conservation %>% filter(category == "mixed")

  if (nrow(mixed_targets) > 0) {
    p_mixed <- ggplot(mixed_targets, aes(x = pct_excitatory)) +
      geom_histogram(binwidth = 10, fill = "#999999", color = "white", alpha = 0.8) +
      labs(
        title = sprintf("Mixed-valence targets (%s)", suffix),
        x = "% modalities excitatory",
        y = "Number of target neurons"
      ) +
      theme_minimal(base_size = 12)

    print(p_mixed)
    save_fig6_plot(p_mixed, "panel_J_mixed_valence_detail", width = 10, height = 5, subdir = suffix)
  }

  # --------------------------------------------------------------------------
  # 6. Valence agreement between modalities
  # --------------------------------------------------------------------------
  mod_list <- modality_order
  sim_mat <- matrix(NA, nrow = length(mod_list), ncol = length(mod_list),
                    dimnames = list(mod_list, mod_list))

  valence_by_target <- absolute_strongest %>%
    select(end, modality, valence) %>%
    pivot_wider(names_from = modality, values_from = valence)

  for (i in seq_along(mod_list)) {
    for (j in seq_along(mod_list)) {
      m1 <- mod_list[i]; m2 <- mod_list[j]
      if (m1 %in% names(valence_by_target) && m2 %in% names(valence_by_target)) {
        both_present <- !is.na(valence_by_target[[m1]]) & !is.na(valence_by_target[[m2]])
        if (sum(both_present) > 0) {
          sim_mat[i, j] <- mean(valence_by_target[[m1]][both_present] ==
                                valence_by_target[[m2]][both_present])
        }
      }
    }
  }

  pheatmap(
    sim_mat, cluster_rows = TRUE, cluster_cols = TRUE,
    color = colorRampPalette(c("#F7F7F7", "#D1E5F0", "#4393C3", "#2166AC"))(100),
    border_color = "grey80", display_numbers = TRUE,
    number_format = "%.2f", fontsize_number = 9,
    fontsize_row = 11, fontsize_col = 11,
    main = sprintf("Valence agreement between modalities (%s)", suffix)
  )

  # Save valence agreement heatmap
  for (ext in c("png", "pdf")) {
    sub_dir <- file.path(panel_plot_dir("panel_J_valence_agreement_matrix", ext, subdir = suffix))
    f <- file.path(sub_dir, paste0("panel_J_valence_agreement_matrix.", ext))
    if (ext == "png") png(f, width = 8, height = 7, units = "in", res = 300)
    else              pdf(f, width = 8, height = 7)
    pheatmap(sim_mat, cluster_rows = TRUE, cluster_cols = TRUE,
             color = colorRampPalette(c("#F7F7F7", "#D1E5F0", "#4393C3", "#2166AC"))(100),
             border_color = "grey80", display_numbers = TRUE,
             number_format = "%.2f", fontsize_number = 9,
             fontsize_row = 11, fontsize_col = 11,
             main = sprintf("Valence agreement between modalities (%s)", suffix))
    dev.off()
  }

  # --------------------------------------------------------------------------
  # 7. Valence x strength tile map
  # --------------------------------------------------------------------------
  p_tile_str <- ggplot(valence_long, aes(x = modality, y = end)) +
    geom_tile(aes(fill = valence, alpha = strength),
              color = "white", linewidth = 0.3) +
    scale_fill_manual(values = valence_colors, name = "Overall\nvalence") +
    scale_alpha_continuous(range = c(0.3, 1), name = "Strength") +
    labs(
      title = sprintf("Valence x strength (%s)", suffix),
      subtitle = "Color = excitatory/inhibitory; opacity = path strength",
      x = "Sensory modality",
      y = "Target neuron"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.y = element_text(size = 5),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
      panel.grid = element_blank()
    )

  print(p_tile_str)
  save_fig6_plot(p_tile_str, "panel_J_valence_strength_tile", width = 12, height = 14, subdir = suffix)

  # --------------------------------------------------------------------------
  # 8. Valence pattern frequency (bar chart)
  # --------------------------------------------------------------------------
  diverging_palette <- colorRampPalette(
    c("#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
      "#F7F7F7",
      "#FDDBC7", "#F4A582", "#D6604D", "#B2182B")
  )(100)

  pattern_counts <- absolute_strongest %>%
    group_by(valence_pattern) %>%
    summarise(
      count = n(),
      mean_strength = mean(strength),
      median_strength = median(strength),
      .groups = "drop"
    ) %>%
    arrange(desc(count))

  p_freq <- ggplot(pattern_counts, aes(x = reorder(valence_pattern, -count), y = count)) +
    geom_col(aes(fill = log10(mean_strength)), color = "grey30", linewidth = 0.3) +
    scale_fill_viridis_c(option = "plasma", name = "log10(mean\nstrength)") +
    labs(
      title = sprintf("Valence pattern frequency (%s)", suffix),
      x = "Valence pattern (E = excitatory hop, I = inhibitory hop)",
      y = "Count"
    ) +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

  print(p_freq)
  save_fig6_plot(p_freq, "panel_J_valence_pattern_frequency", width = 12, height = 6, subdir = suffix)

  # --------------------------------------------------------------------------
  # 9. Valence pattern by modality (stacked bar)
  # --------------------------------------------------------------------------
  pattern_by_mod <- absolute_strongest %>%
    group_by(modality, valence_pattern) %>%
    summarise(count = n(), mean_strength = mean(strength), .groups = "drop")

  p_mod_pattern <- ggplot(pattern_by_mod,
      aes(x = factor(modality, levels = modality_order), y = count, fill = valence_pattern)) +
    geom_col(position = "stack", color = "white", linewidth = 0.2) +
    scale_fill_brewer(palette = "Set3", name = "Valence\npattern") +
    labs(title = sprintf("Detailed valence patterns by modality (%s)", suffix),
         x = "Modality", y = "Number of target neurons") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  print(p_mod_pattern)
  save_fig6_plot(p_mod_pattern, "panel_J_pattern_by_modality", width = 12, height = 6, subdir = suffix)

  # --------------------------------------------------------------------------
  # 10. Modality x valence pattern heatmap (mean signed strength)
  # --------------------------------------------------------------------------
  pattern_heatmap_df <- absolute_strongest %>%
    mutate(signed_str = ifelse(valence == "excitatory", strength, -strength)) %>%
    group_by(modality, valence_pattern) %>%
    summarise(mean_signed = mean(signed_str), count = n(), .groups = "drop")

  pattern_mat <- pattern_heatmap_df %>%
    select(modality, valence_pattern, mean_signed) %>%
    pivot_wider(names_from = valence_pattern, values_from = mean_signed, values_fill = 0) %>%
    column_to_rownames("modality")
  pattern_mat <- as.matrix(pattern_mat[intersect(modality_order, rownames(pattern_mat)), ])

  col_order <- names(sort(colSums(abs(pattern_mat)), decreasing = TRUE))
  pattern_mat <- pattern_mat[, col_order, drop = FALSE]

  abs_max_p <- max(abs(pattern_mat), na.rm = TRUE)
  if (abs_max_p == 0) abs_max_p <- 1
  breaks_p <- seq(-abs_max_p, abs_max_p, length.out = 101)

  pheatmap(
    pattern_mat, cluster_rows = FALSE, cluster_cols = FALSE,
    color = diverging_palette, breaks = breaks_p,
    border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
    main = sprintf("Mean signed strength by modality x valence pattern (%s)", suffix),
    angle_col = 45, display_numbers = TRUE,
    number_format = "%.4f", fontsize_number = 7
  )

  # Save modality x valence pattern heatmap
  for (ext in c("png", "pdf")) {
    sub_dir <- file.path(panel_plot_dir("panel_J_modality_x_valence_pattern", ext, subdir = suffix))
    f <- file.path(sub_dir, paste0("panel_J_modality_x_valence_pattern.", ext))
    if (ext == "png") png(f, width = 12, height = 6, units = "in", res = 300)
    else              pdf(f, width = 12, height = 6)
    pheatmap(pattern_mat, cluster_rows = FALSE, cluster_cols = FALSE,
             color = diverging_palette, breaks = breaks_p,
             border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
             main = sprintf("Mean signed strength by modality x valence pattern (%s)", suffix),
             angle_col = 45, display_numbers = TRUE,
             number_format = "%.4f", fontsize_number = 7)
    dev.off()
  }

  # --------------------------------------------------------------------------
  # 11. Modality x valence pattern count heatmap
  # --------------------------------------------------------------------------
  count_mat <- pattern_heatmap_df %>%
    select(modality, valence_pattern, count) %>%
    pivot_wider(names_from = valence_pattern, values_from = count, values_fill = 0) %>%
    column_to_rownames("modality")
  count_mat <- as.matrix(count_mat[intersect(modality_order, rownames(count_mat)), ])
  count_mat <- count_mat[, intersect(col_order, colnames(count_mat)), drop = FALSE]

  pheatmap(
    count_mat, cluster_rows = FALSE, cluster_cols = FALSE,
    color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(50),
    border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
    main = sprintf("Count of strongest paths by modality x valence pattern (%s)", suffix),
    angle_col = 45, display_numbers = TRUE,
    number_format = "%.0f", fontsize_number = 8
  )

  # Save count heatmap
  for (ext in c("png", "pdf")) {
    sub_dir <- file.path(panel_plot_dir("panel_J_modality_x_pattern_count", ext, subdir = suffix))
    f <- file.path(sub_dir, paste0("panel_J_modality_x_pattern_count.", ext))
    if (ext == "png") png(f, width = 12, height = 6, units = "in", res = 300)
    else              pdf(f, width = 12, height = 6)
    pheatmap(count_mat, cluster_rows = FALSE, cluster_cols = FALSE,
             color = colorRampPalette(c("white", "#FEE0D2", "#FC9272", "#DE2D26"))(50),
             border_color = "grey80", fontsize_row = 11, fontsize_col = 9,
             main = sprintf("Count of strongest paths by modality x valence pattern (%s)", suffix),
             angle_col = 45, display_numbers = TRUE,
             number_format = "%.0f", fontsize_number = 8)
    dev.off()
  }

  # --------------------------------------------------------------------------
  # 12. Signed strength heatmap: modality x target neuron (row-normalized)
  # --------------------------------------------------------------------------
  signed_df <- strongest_by_valence %>%
    mutate(signed_strength = ifelse(valence == "excitatory", strength, -strength)) %>%
    group_by(modality, end) %>%
    summarise(net_signed = sum(signed_strength), .groups = "drop")

  signed_mat <- signed_df %>%
    pivot_wider(names_from = modality, values_from = net_signed, values_fill = 0) %>%
    column_to_rownames("end")

  for (mod in modality_order) {
    if (!mod %in% colnames(signed_mat)) signed_mat[[mod]] <- 0
  }
  signed_mat <- as.matrix(signed_mat[, intersect(modality_order, colnames(signed_mat))])

  # Apply Panel B row order
  cluster_rows_flag <- !use_ref_order
  if (use_ref_order) {
    shared_rows <- intersect(row_order_ref, rownames(signed_mat))
    extra_rows  <- setdiff(rownames(signed_mat), row_order_ref)
    signed_mat  <- signed_mat[c(shared_rows, extra_rows), ]
  }

  # Row-normalize: each target neuron scaled to [-1, 1]
  signed_mat_norm <- row_normalize(signed_mat)

  breaks_signed <- seq(-1, 1, length.out = 101)

  row_annot <- data.frame(
    `aSP-f` = ifelse(rownames(signed_mat_norm) %in% aspf_ends, "yes", "no"),
    `aSP-g` = ifelse(rownames(signed_mat_norm) %in% aspg_ends, "yes", "no"),
    row.names = rownames(signed_mat_norm), check.names = FALSE
  )
  col_annot <- data.frame(
    modality_group = modality_group[colnames(signed_mat_norm)],
    row.names = colnames(signed_mat_norm)
  )
  ann_colors <- list(
    `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
    `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
    modality_group = modality_group_colors
  )

  draw_heatmap_fixed_annot(
    signed_mat_norm, row_annot, col_annot, ann_colors,
    cluster_rows = TRUE, cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    color = diverging_palette, breaks = breaks_signed,
    border_color = NA, fontsize_row = 6, fontsize_col = 10,
    main = sprintf("Signed strength (row-normalized) (%s)", suffix),
    angle_col = 45
  )

  save_heatmap_both_orientations(
    mat = signed_mat_norm, row_annot = row_annot, col_annot = col_annot,
    ann_colors = ann_colors, name = "panel_J_signed_strength",
    width = 8, height = 10, subdir = suffix,
    cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    color = diverging_palette, breaks = breaks_signed,
    border_color = NA, fontsize_row = 6, fontsize_col = 10,
    main = sprintf("Signed strength (row-normalized): strongest single path (%s)", suffix),
    angle_col = 45
  )

  # --------------------------------------------------------------------------
  # 13. Per-target valence profile (absolute strongest path, row-normalized)
  # --------------------------------------------------------------------------
  valence_profile <- absolute_strongest %>%
    mutate(signed_str = ifelse(valence == "excitatory", strength, -strength)) %>%
    select(end, modality, signed_str) %>%
    pivot_wider(names_from = modality, values_from = signed_str, values_fill = 0) %>%
    column_to_rownames("end")

  for (mod in modality_order) {
    if (!mod %in% colnames(valence_profile)) valence_profile[[mod]] <- 0
  }
  valence_profile <- as.matrix(valence_profile[, intersect(modality_order, colnames(valence_profile))])

  # Apply Panel B row order
  if (use_ref_order) {
    shared_vp <- intersect(row_order_ref, rownames(valence_profile))
    extra_vp  <- setdiff(rownames(valence_profile), row_order_ref)
    valence_profile <- valence_profile[c(shared_vp, extra_vp), ]
  }

  # Row-normalize
  valence_profile_norm <- row_normalize(valence_profile)
  breaks_vp <- seq(-1, 1, length.out = 101)

  row_annot_vp <- data.frame(
    `aSP-f` = ifelse(rownames(valence_profile_norm) %in% aspf_ends, "yes", "no"),
    `aSP-g` = ifelse(rownames(valence_profile_norm) %in% aspg_ends, "yes", "no"),
    row.names = rownames(valence_profile_norm), check.names = FALSE
  )
  col_annot_vp <- data.frame(
    modality_group = modality_group[colnames(valence_profile_norm)],
    row.names = colnames(valence_profile_norm)
  )

  draw_heatmap_fixed_annot(
    valence_profile_norm, row_annot_vp, col_annot_vp, ann_colors,
    cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    color = diverging_palette, breaks = breaks_vp,
    border_color = NA, fontsize_row = 6, fontsize_col = 10,
    main = sprintf("Per-target valence profile (row-normalized) (%s)", suffix),
    angle_col = 45
  )

  save_heatmap_both_orientations(
    mat = valence_profile_norm, row_annot = row_annot_vp, col_annot = col_annot_vp,
    ann_colors = ann_colors, name = "panel_J_valence_profile",
    width = 8, height = 10, subdir = suffix,
    cluster_rows = cluster_rows_flag, cluster_cols = FALSE,
    clustering_distance_rows = "euclidean",
    clustering_method = "ward.D2",
    color = diverging_palette, breaks = breaks_vp,
    border_color = NA, fontsize_row = 6, fontsize_col = 10,
    main = sprintf("Valence profile (row-normalized): strongest single path (%s)", suffix),
    angle_col = 45
  )

  # --------------------------------------------------------------------------
  # 14. Network diagram: modality -> target (igraph)
  # --------------------------------------------------------------------------
  edges <- absolute_strongest %>%
    select(modality, end, strength, valence, valence_pattern) %>%
    rename(from = modality, to = end)

  g_net <- graph_from_data_frame(edges, directed = TRUE)

  V(g_net)$type <- ifelse(V(g_net)$name %in% modality_order, "modality", "target")
  V(g_net)$color <- ifelse(V(g_net)$type == "modality",
                            modality_colors[V(g_net)$name], "grey85")
  V(g_net)$size <- ifelse(V(g_net)$type == "modality", 12, 5)
  V(g_net)$label.cex <- ifelse(V(g_net)$type == "modality", 1.0, 0.5)
  V(g_net)$frame.color <- ifelse(V(g_net)$type == "modality", "grey30", "grey60")

  E(g_net)$color <- ifelse(E(g_net)$valence == "excitatory",
                            adjustcolor("#B2182B", alpha.f = 0.4),
                            adjustcolor("#2166AC", alpha.f = 0.4))
  E(g_net)$width <- scales::rescale(E(g_net)$strength, to = c(0.3, 3))
  E(g_net)$arrow.size <- 0.3

  mod_nodes <- which(V(g_net)$type == "modality")
  tar_nodes <- which(V(g_net)$type == "target")

  layout_mat <- matrix(0, nrow = vcount(g_net), ncol = 2)
  layout_mat[mod_nodes, 1] <- seq_along(mod_nodes)
  layout_mat[mod_nodes, 2] <- 1
  layout_mat[tar_nodes, 1] <- seq(0.5, length(mod_nodes) + 0.5, length.out = length(tar_nodes))
  layout_mat[tar_nodes, 2] <- 0

  plot(g_net, layout = layout_mat,
       vertex.label = V(g_net)$name, edge.curved = 0.1,
       main = sprintf("Modality -> target (%s)\nRed = excitatory, Blue = inhibitory", suffix))
  legend("bottomright", legend = c("Excitatory", "Inhibitory"),
         col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n", cex = 0.9)

  # Save network diagram
  for (ext in c("png", "pdf")) {
    sub_dir <- file.path(panel_plot_dir("panel_J_network_modality_target", ext, subdir = suffix))
    f <- file.path(sub_dir, paste0("panel_J_network_modality_target.", ext))
    if (ext == "png") png(f, width = 14, height = 14, units = "in", res = 300)
    else              pdf(f, width = 14, height = 14)
    plot(g_net, layout = layout_mat,
         vertex.label = V(g_net)$name, edge.curved = 0.1,
         main = sprintf("Modality -> target (%s)\nRed = excitatory, Blue = inhibitory", suffix))
    legend("bottomright", legend = c("Excitatory", "Inhibitory"),
           col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n", cex = 0.9)
    dev.off()
  }

  # --------------------------------------------------------------------------
  # 15. Simplified network (top 50% connections)
  # --------------------------------------------------------------------------
  net_summary <- absolute_strongest %>%
    group_by(modality, end) %>%
    summarise(total_strength = sum(strength),
              dominant_valence = valence[which.max(strength)], .groups = "drop")

  strength_threshold <- quantile(net_summary$total_strength, 0.5)
  net_filtered <- net_summary %>% filter(total_strength >= strength_threshold)

  g_simple <- graph_from_data_frame(
    net_filtered %>% select(from = modality, to = end, everything()), directed = TRUE
  )

  V(g_simple)$type <- ifelse(V(g_simple)$name %in% modality_order, "modality", "target")
  V(g_simple)$color <- ifelse(V(g_simple)$type == "modality",
                               modality_colors[V(g_simple)$name], "grey85")
  V(g_simple)$size <- ifelse(V(g_simple)$type == "modality", 15, 8)
  V(g_simple)$label.cex <- ifelse(V(g_simple)$type == "modality", 1.1, 0.65)
  V(g_simple)$frame.color <- ifelse(V(g_simple)$type == "modality", "grey30", "grey60")

  E(g_simple)$color <- ifelse(E(g_simple)$dominant_valence == "excitatory",
                               adjustcolor("#B2182B", alpha.f = 0.5),
                               adjustcolor("#2166AC", alpha.f = 0.5))
  E(g_simple)$width <- scales::rescale(E(g_simple)$total_strength, to = c(0.5, 4))
  E(g_simple)$arrow.size <- 0.4

  mod_n2 <- which(V(g_simple)$type == "modality")
  tar_n2 <- which(V(g_simple)$type == "target")
  layout2 <- matrix(0, nrow = vcount(g_simple), ncol = 2)
  layout2[mod_n2, 1] <- seq_along(mod_n2)
  layout2[mod_n2, 2] <- 1
  layout2[tar_n2, 1] <- seq(0.5, length(mod_n2) + 0.5, length.out = length(tar_n2))
  layout2[tar_n2, 2] <- 0

  plot(g_simple, layout = layout2,
       vertex.label = V(g_simple)$name, edge.curved = 0.15,
       main = sprintf("Top 50%% strongest connections (%s)\nRed = excitatory, Blue = inhibitory", suffix))
  legend("bottomright", legend = c("Excitatory", "Inhibitory"),
         col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n")

  # Save simplified network
  for (ext in c("png", "pdf")) {
    sub_dir <- file.path(panel_plot_dir("panel_J_network_simplified", ext, subdir = suffix))
    f <- file.path(sub_dir, paste0("panel_J_network_simplified.", ext))
    if (ext == "png") png(f, width = 14, height = 10, units = "in", res = 300)
    else              pdf(f, width = 14, height = 10)
    plot(g_simple, layout = layout2,
         vertex.label = V(g_simple)$name, edge.curved = 0.15,
         main = sprintf("Top 50%% strongest connections (%s)\nRed = excitatory, Blue = inhibitory", suffix))
    legend("bottomright", legend = c("Excitatory", "Inhibitory"),
           col = c("#B2182B", "#2166AC"), lwd = 2, bty = "n")
    dev.off()
  }

  # --------------------------------------------------------------------------
  # 16. ggraph network (publication-quality)
  # --------------------------------------------------------------------------
  node_df <- data.frame(
    name = c(modality_order, setdiff(unique(absolute_strongest$end), modality_order)),
    stringsAsFactors = FALSE
  ) %>%
    mutate(
      node_type = ifelse(name %in% modality_order, "modality", "target"),
      node_color = ifelse(node_type == "modality", unname(modality_colors[name]), "grey75"),
      type = (node_type == "modality")
    )

  edge_df <- absolute_strongest %>%
    select(from = modality, to = end, strength, valence, valence_pattern)

  tg <- tbl_graph(nodes = node_df, edges = edge_df, directed = TRUE)

  p_net <- ggraph(tg, layout = "bipartite") +
    geom_edge_fan(
      aes(color = valence, width = strength, alpha = strength),
      arrow = grid::arrow(length = unit(2, "mm"), type = "closed"),
      end_cap = circle(3, "mm")
    ) +
    scale_edge_color_manual(values = c("excitatory" = "#B2182B", "inhibitory" = "#2166AC")) +
    scale_edge_width(range = c(0.2, 2), guide = "none") +
    scale_edge_alpha(range = c(0.15, 0.6), guide = "none") +
    geom_node_point(aes(color = node_color, size = node_type)) +
    scale_color_identity() +
    scale_size_manual(values = c("modality" = 8, "target" = 3), guide = "none") +
    geom_node_text(aes(label = name, y = y + ifelse(node_type == "modality", 0.08, -0.08)),
                   size = 3, fontface = "bold") +
    theme_void() +
    labs(title = sprintf("Modality -> 3rd-order neuron (%s)", suffix),
         subtitle = "Red = excitatory, Blue = inhibitory") +
    theme(plot.title = element_text(size = 14, face = "bold"),
          plot.subtitle = element_text(size = 11))

  print(p_net)
  save_fig6_plot(p_net, "panel_J_ggraph_network", width = 14, height = 10, subdir = suffix)

  # --------------------------------------------------------------------------
  # 17. Summary statistics
  # --------------------------------------------------------------------------
  summary_tab <- absolute_strongest %>%
    group_by(modality) %>%
    summarise(
      n_targets = n_distinct(end),
      n_excitatory = sum(valence == "excitatory"),
      n_inhibitory = sum(valence == "inhibitory"),
      pct_excitatory = round(100 * mean(valence == "excitatory"), 1),
      mean_strength = mean(strength),
      n_unique_patterns = n_distinct(valence_pattern),
      top_pattern = names(sort(table(valence_pattern), decreasing = TRUE))[1],
      .groups = "drop"
    )

  cat(sprintf("\n[%s] Summary per modality:\n", suffix))
  print(as.data.frame(summary_tab), row.names = FALSE)

  cat(sprintf("\n[%s] Overall excitatory fraction: %.1f%%\n", suffix,
              100 * mean(absolute_strongest$valence == "excitatory")))
  cat(sprintf("[%s] Most common pattern: %s (%d occurrences)\n", suffix,
              pattern_counts$valence_pattern[1], pattern_counts$count[1]))
  cat(sprintf("\n[%s] Valence conservation: %.0f%% all excitatory, %.0f%% all inhibitory, %.0f%% mixed\n",
              suffix,
              100 * mean(conservation$category == "all excitatory"),
              100 * mean(conservation$category == "all inhibitory"),
              100 * mean(conservation$category == "mixed")))

  cat(sprintf("\n\n**[%s] analysis complete.**\n\n", suffix))
}
```

## Session info

```{r session-info}
sessionInfo()
```
