---
title: "Figure 6B: Modality Composition Heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel B: Modality Composition Heatmap — Clustered heatmap of column-normalized modality input strength to 3rd-order courtship neurons. Rows are target neuron types (aSP-f, aSP-g); columns are the 7 sensory modalities. Row annotations mark target set membership; column annotations mark modality group (olfactory, auditory, visual, contact)."
panel_legend <- "(B) Column-normalized heatmap of modality input strength to third-order target neurons. Each column (modality) is scaled to its maximum, enabling comparison of relative target distributions across modalities. Rows are clustered by Ward's method; annotations indicate target set (aSP-f vs. aSP-g) and modality group."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_B_description", "txt"), "panel_B_description.txt"))
```

## Overview

Panel B is a clustered heatmap of modality input strength to third-order
courtship neurons. Columns are the 7 sensory modalities; rows are target
neuron types. The matrix is column-normalized so that each modality's maximum
is 1, making the relative distribution across targets comparable between
modalities of different absolute strengths.

Row annotations indicate the target set (aSP-f vs aSP-g).
Column annotations indicate the modality group (olfactory / auditory / visual / contact).

## 1. Prepare strength matrix (aSP-f + aSP-g targets)

```{r strength-matrix-aspf-aspg}
# ---------------------------------------------------------------------------
# Column-normalize the pre-computed strength matrix
# Each column (modality) is scaled so its maximum value = 1
# ---------------------------------------------------------------------------
mat <- strength.aspf_aspg

# Column-normalize: divide each column by its max
col_maxes <- apply(mat, 2, max, na.rm = TRUE)
col_maxes[col_maxes == 0] <- 1
mat_norm <- sweep(mat, 2, col_maxes, "/")

cat(sprintf("Matrix dimensions: %d target neurons x %d modalities\n",
            nrow(mat_norm), ncol(mat_norm)))
```

## 2. Build row annotation (target set membership)

Each target neuron may appear in aSP-f paths, aSP-g paths, or both. We assign
primary membership based on which target set contributes greater total strength.

```{r row-annotation}
# ---------------------------------------------------------------------------
# Determine target_set membership for each target neuron
# ---------------------------------------------------------------------------
# Separate binary bands per target set; pheatmap annotation widths will be
# adjusted after creation so the total annotation width stays fixed.
all_ends <- rownames(mat_norm)
aspf_ends <- unique(paths.aspf$end)
aspg_ends <- unique(paths.aspg$end)

row_annot <- data.frame(
  `aSP-f` = ifelse(all_ends %in% aspf_ends, "yes", "no"),
  `aSP-g` = ifelse(all_ends %in% aspg_ends, "yes", "no"),
  row.names = all_ends,
  check.names = FALSE
)
```

## 3. Build column annotation (modality group)

```{r col-annotation}
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_norm)],
  row.names      = colnames(mat_norm)
)
```

## 4. Define annotation colours and helper

```{r annotation-colors}
ann_colors <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  modality_group = modality_group_colors
)
# draw_heatmap_fixed_annot() and save_heatmap_fixed_annot() are in setup.R
```

## 5. Draw heatmap (aSP-f + aSP-g targets)

```{r heatmap-aspf-aspg, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Primary heatmap: modality strength to aSP-f / aSP-g targets
# ---------------------------------------------------------------------------
hm <- draw_heatmap_fixed_annot(
  mat_norm, row_annot, col_annot, ann_colors,
  total_annot_cm           = 1.5,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons",
  angle_col                = 45
)
```

## 6. Save primary heatmap

```{r save-primary}
save_heatmap_both_orientations(
  mat        = mat_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_aspf_aspg",
  width      = 8,
  height     = 10,
  subdir     = "colnorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons",
  angle_col                = 45
)
```

## 6b. Row-normalized heatmap

```{r row-norm-matrix}
# Row-normalize: each neuron's max across modalities = 1
row_maxes <- apply(mat, 1, max, na.rm = TRUE)
row_maxes[row_maxes == 0] <- 1
mat_row_norm <- sweep(mat, 1, row_maxes, "/")
```

```{r save-row-norm}
save_heatmap_both_orientations(
  mat        = mat_row_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_aspf_aspg_rownorm",
  width      = 8,
  height     = 10,
  subdir     = "rownorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (row-normalized)",
  angle_col                = 45
)
```

## 6c. Unnormalized (raw strength) heatmap

```{r save-raw}
save_heatmap_both_orientations(
  mat        = mat,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_aspf_aspg_raw",
  width      = 8,
  height     = 10,
  subdir     = "raw",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (unnormalized)",
  angle_col                = 45
)
```

## 7. Extended version: include vAB3 and PPN1 downstream targets

Recompute the strength matrix using `paths.all` to include targets reached
via vAB3 and PPN1 paths, providing a broader view of multimodal convergence.

```{r strength-matrix-all}
# ---------------------------------------------------------------------------
# Strength matrix from all paths (aspf + aspg + vAB3 + PPN1)
# ---------------------------------------------------------------------------
mat_all <- compute_strength_matrix(paths.all, normalize = FALSE)

# Column-normalize
col_maxes_all <- apply(mat_all, 2, max, na.rm = TRUE)
col_maxes_all[col_maxes_all == 0] <- 1
mat_all_norm <- sweep(mat_all, 2, col_maxes_all, "/")

cat(sprintf("Extended matrix: %d target neurons x %d modalities\n",
            nrow(mat_all_norm), ncol(mat_all_norm)))
```

## 8. Row annotation for extended matrix

```{r row-annotation-all}
# ---------------------------------------------------------------------------
# Target set membership for all targets
# ---------------------------------------------------------------------------
# 4 binary bands (one per target set), fixed total width via helper
all_ends_ext <- rownames(mat_all_norm)

row_annot_all <- data.frame(
  `aSP-f` = ifelse(all_ends_ext %in% unique(paths.aspf$end), "yes", "no"),
  `aSP-g` = ifelse(all_ends_ext %in% unique(paths.aspg$end), "yes", "no"),
  `vAB3`  = ifelse(all_ends_ext %in% unique(paths.vab3$end), "yes", "no"),
  `PPN1`  = ifelse(all_ends_ext %in% unique(paths.ppn1$end), "yes", "no"),
  row.names = all_ends_ext,
  check.names = FALSE
)

col_annot_all <- data.frame(
  modality_group = modality_group[colnames(mat_all_norm)],
  row.names      = colnames(mat_all_norm)
)

ann_colors_all <- list(
  `aSP-f`        = c("yes" = "#E41A1C", "no" = "white"),
  `aSP-g`        = c("yes" = "#377EB8", "no" = "white"),
  `vAB3`         = c("yes" = "#4DAF4A", "no" = "white"),
  `PPN1`         = c("yes" = "#FF7F00", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 9. Draw extended heatmap

```{r heatmap-all, fig.width=8, fig.height=14}
# ---------------------------------------------------------------------------
# Extended heatmap: all target sets
# ---------------------------------------------------------------------------
# 4 bands share 1.5 cm total = 0.375 cm each
hm_all <- draw_heatmap_fixed_annot(
  mat_all_norm, row_annot_all, col_annot_all, ann_colors_all,
  total_annot_cm           = 1.5,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (all target sets)",
  angle_col                = 45
)
```

## 10. Save extended heatmap

```{r save-extended}
save_heatmap_both_orientations(
  mat        = mat_all_norm,
  row_annot  = row_annot_all,
  col_annot  = col_annot_all,
  ann_colors = ann_colors_all,
  name       = "panel_B_heatmap_all_targets",
  width      = 8,
  height     = 14,
  subdir     = "colnorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (all target sets)",
  angle_col                = 45
)
```

## 10a-rn. Row-normalized heatmap (all targets)

```{r all-targets-rownorm}
row_maxes_all <- apply(mat_all, 1, max, na.rm = TRUE)
row_maxes_all[row_maxes_all == 0] <- 1
mat_all_row_norm <- sweep(mat_all, 1, row_maxes_all, "/")

save_heatmap_both_orientations(
  mat        = mat_all_row_norm,
  row_annot  = row_annot_all,
  col_annot  = col_annot_all,
  ann_colors = ann_colors_all,
  name       = "panel_B_heatmap_all_targets_rownorm",
  width      = 8,
  height     = 14,
  subdir     = "rownorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (row-normalized, all target sets)",
  angle_col                = 45
)
```

## 10a-raw. Unnormalized heatmap (all targets)

```{r all-targets-raw}
save_heatmap_both_orientations(
  mat        = mat_all,
  row_annot  = row_annot_all,
  col_annot  = col_annot_all,
  ann_colors = ann_colors_all,
  name       = "panel_B_heatmap_all_targets_raw",
  width      = 8,
  height     = 14,
  subdir     = "raw",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 5,
  fontsize_col             = 10,
  main                     = "Modality input strength (unnormalized, all target sets)",
  angle_col                = 45
)
```

## 10b. vAB3 & PPN1 only heatmap (top 50 paths)

The evaluation notebook computed 1000 strongest paths per modality to vAB3 and
PPN1. Here we keep only the top 50 per (modality, end) pair to match the
aSP-f/aSP-g analysis depth, then build a standalone heatmap.

```{r vp-filter-top50}
filter_top_n <- function(df, n = 50) {
  df %>%
    group_by(end, modality) %>%
    slice_max(order_by = strength, n = n, with_ties = FALSE) %>%
    ungroup()
}

paths.vab3.top50 <- filter_top_n(paths.vab3, 50)
paths.ppn1.top50 <- filter_top_n(paths.ppn1, 50)
paths.vab3_ppn1.top50 <- bind_rows(paths.vab3.top50, paths.ppn1.top50)

cat(sprintf("vAB3 top-50: %d paths, %d unique ends\n",
            nrow(paths.vab3.top50), length(unique(paths.vab3.top50$end))))
cat(sprintf("PPN1 top-50: %d paths, %d unique ends\n",
            nrow(paths.ppn1.top50), length(unique(paths.ppn1.top50$end))))
cat(sprintf("Combined: %d paths, %d unique ends\n",
            nrow(paths.vab3_ppn1.top50), length(unique(paths.vab3_ppn1.top50$end))))
```

```{r vp-strength-matrix}
mat_vp <- compute_strength_matrix(paths.vab3_ppn1.top50, normalize = FALSE)

col_maxes_vp <- apply(mat_vp, 2, max, na.rm = TRUE)
col_maxes_vp[col_maxes_vp == 0] <- 1
mat_vp_norm <- sweep(mat_vp, 2, col_maxes_vp, "/")

all_ends_vp <- rownames(mat_vp_norm)
vab3_ends <- unique(paths.vab3.top50$end)
ppn1_ends <- unique(paths.ppn1.top50$end)

row_annot_vp <- data.frame(
  `vAB3` = ifelse(all_ends_vp %in% vab3_ends, "yes", "no"),
  `PPN1` = ifelse(all_ends_vp %in% ppn1_ends, "yes", "no"),
  row.names = all_ends_vp,
  check.names = FALSE
)

col_annot_vp <- data.frame(
  modality_group = modality_group[colnames(mat_vp_norm)],
  row.names      = colnames(mat_vp_norm)
)

ann_colors_vp <- list(
  `vAB3`         = c("yes" = "#4DAF4A", "no" = "white"),
  `PPN1`         = c("yes" = "#FF7F00", "no" = "white"),
  modality_group = modality_group_colors
)
```

```{r vp-save}
save_heatmap_both_orientations(
  mat        = mat_vp_norm,
  row_annot  = row_annot_vp,
  col_annot  = col_annot_vp,
  ann_colors = ann_colors_vp,
  name       = "panel_B_heatmap_vab3_ppn1_top50",
  width      = 8,
  height     = 10,
  subdir     = "colnorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to vAB3 & PPN1 targets (top 50 paths)",
  angle_col                = 45
)
```

```{r vp-save-rownorm}
row_maxes_vp <- apply(mat_vp, 1, max, na.rm = TRUE)
row_maxes_vp[row_maxes_vp == 0] <- 1
mat_vp_row_norm <- sweep(mat_vp, 1, row_maxes_vp, "/")

save_heatmap_both_orientations(
  mat        = mat_vp_row_norm,
  row_annot  = row_annot_vp,
  col_annot  = col_annot_vp,
  ann_colors = ann_colors_vp,
  name       = "panel_B_heatmap_vab3_ppn1_top50_rownorm",
  width      = 8,
  height     = 10,
  subdir     = "rownorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (row-normalized, vAB3 & PPN1 top 50)",
  angle_col                = 45
)
```

```{r vp-save-raw}
save_heatmap_both_orientations(
  mat        = mat_vp,
  row_annot  = row_annot_vp,
  col_annot  = col_annot_vp,
  ann_colors = ann_colors_vp,
  name       = "panel_B_heatmap_vab3_ppn1_top50_raw",
  width      = 8,
  height     = 10,
  subdir     = "raw",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (unnormalized, vAB3 & PPN1 top 50)",
  angle_col                = 45
)
```

## 10c. vAB3 & PPN1 summary statistics

```{r vp-summary}
cat("=== Per target set summary ===\n")
paths.vab3_ppn1.top50 %>%
  group_by(target_set, modality) %>%
  summarise(
    n_paths = n(),
    total_strength = sum(strength, na.rm = TRUE),
    mean_strength = mean(strength, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(target_set, desc(total_strength)) %>%
  print(n = 20)

cat(sprintf("\n=== Unique end types per target set ===\n"))
cat(sprintf("vAB3: %d unique end types\n", length(unique(paths.vab3.top50$end))))
cat(sprintf("PPN1: %d unique end types\n", length(unique(paths.ppn1.top50$end))))

cat("\n=== Top 20 target neurons by total path strength ===\n")
paths.vab3_ppn1.top50 %>%
  group_by(end, target_set) %>%
  summarise(total_strength = sum(strength, na.rm = TRUE),
            n_modalities = n_distinct(modality),
            .groups = "drop") %>%
  arrange(desc(total_strength)) %>%
  head(20) %>%
  print()
```

## 10d. Rank-transformed scatter (all targets)

PCA on within-column rank-transformed strength profiles. Ranking removes
magnitude effects while preserving ordinal structure — neurons that receive
the same relative ordering of modalities cluster together regardless of
their absolute strength.

```{r rank-scatter-all, fig.width=10, fig.height=8}
# ---------------------------------------------------------------------------
# Rank-transform each modality column, then PCA
# ---------------------------------------------------------------------------
mat_rank_all <- apply(mat_all, 2, rank, ties.method = "average")
pca_rank <- prcomp(mat_rank_all, center = TRUE, scale. = TRUE)

var_pct <- summary(pca_rank)$importance[2, ] * 100

rank_df <- data.frame(
  PC1    = pca_rank$x[, 1],
  PC2    = pca_rank$x[, 2],
  neuron = rownames(mat_all),
  stringsAsFactors = FALSE
)

# Assign primary target group for coloring
rank_df$target_group <- {
  is_aspf <- rank_df$neuron %in% unique(paths.aspf$end)
  is_aspg <- rank_df$neuron %in% unique(paths.aspg$end)
  is_vab3 <- rank_df$neuron %in% unique(paths.vab3$end)
  is_ppn1 <- rank_df$neuron %in% unique(paths.ppn1$end)
  ifelse(is_aspf & is_aspg, "aSP-f & aSP-g",
  ifelse(is_aspf, "aSP-f only",
  ifelse(is_aspg, "aSP-g only",
  ifelse(is_vab3 & is_ppn1, "vAB3 & PPN1",
  ifelse(is_vab3, "vAB3",
  ifelse(is_ppn1, "PPN1", "other"))))))
}

group_colors_scatter <- c(
  "aSP-f & aSP-g" = "#984EA3",
  "aSP-f only"    = "#E41A1C",
  "aSP-g only"    = "#377EB8",
  "vAB3 & PPN1"   = "#A65628",
  "vAB3"          = "#4DAF4A",
  "PPN1"          = "#FF7F00",
  "other"         = "grey60"
)

p_rank <- ggplot(rank_df, aes(x = PC1, y = PC2, color = target_group)) +
  geom_point(size = 2.5, alpha = 0.7) +
  scale_color_manual(values = group_colors_scatter) +
  labs(
    x = sprintf("PC1 (%.1f%% variance)", var_pct[1]),
    y = sprintf("PC2 (%.1f%% variance)", var_pct[2]),
    title = "PCA of rank-transformed modality profiles (all targets)",
    color = "Target group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right")

p_rank
save_fig6_plot(p_rank, "panel_B_rank_scatter_all", width = 10, height = 8)
```

## 10e. Correspondence analysis ordination (all targets)

Correspondence analysis is the natural ordination for compositional data:
neurons and modalities are projected into the same chi-squared metric space.
Neurons close to a modality arrow are enriched for that modality relative to
the average profile.

```{r ca-ordination-all, fig.width=10, fig.height=8}
# ---------------------------------------------------------------------------
# Manual Correspondence Analysis via SVD
# ---------------------------------------------------------------------------
mat_ca <- mat_all
mat_ca[mat_ca < 0] <- 0
mat_ca <- mat_ca + 1e-10   # avoid exact zeros

N_total <- sum(mat_ca)
P_ca    <- mat_ca / N_total
ri      <- rowSums(P_ca)
ci      <- colSums(P_ca)

# Standardized residuals
Z_ca <- diag(1 / sqrt(ri)) %*% (P_ca - outer(ri, ci)) %*% diag(1 / sqrt(ci))
sv   <- svd(Z_ca)

# Inertia explained per axis
total_inertia <- sum(sv$d^2)
inertia_pct   <- sv$d^2 / total_inertia * 100

# Row principal coordinates (neurons)
F_row <- diag(1 / sqrt(ri)) %*% sv$u[, 1:2] %*% diag(sv$d[1:2])
rownames(F_row) <- rownames(mat_ca)

# Column principal coordinates (modalities)
G_col <- diag(1 / sqrt(ci)) %*% sv$v[, 1:2] %*% diag(sv$d[1:2])
rownames(G_col) <- colnames(mat_ca)

# Build data frames
ca_neurons <- data.frame(
  Dim1 = F_row[, 1], Dim2 = F_row[, 2],
  neuron = rownames(F_row),
  stringsAsFactors = FALSE
)
ca_neurons$target_group <- rank_df$target_group[match(ca_neurons$neuron, rank_df$neuron)]

ca_modalities <- data.frame(
  Dim1 = G_col[, 1], Dim2 = G_col[, 2],
  modality = rownames(G_col),
  stringsAsFactors = FALSE
)

p_ca <- ggplot() +
  geom_point(data = ca_neurons,
             aes(x = Dim1, y = Dim2, color = target_group),
             size = 2, alpha = 0.6) +
  geom_segment(data = ca_modalities,
               aes(x = 0, y = 0, xend = Dim1, yend = Dim2),
               arrow = arrow(length = unit(0.25, "cm")),
               color = "grey30", linewidth = 0.8) +
  geom_text(data = ca_modalities,
            aes(x = Dim1 * 1.12, y = Dim2 * 1.12, label = modality),
            color = "grey20", size = 4, fontface = "bold") +
  scale_color_manual(values = group_colors_scatter) +
  labs(
    x = sprintf("CA Dim 1 (%.1f%% inertia)", inertia_pct[1]),
    y = sprintf("CA Dim 2 (%.1f%% inertia)", inertia_pct[2]),
    title = "Correspondence analysis: neurons and modalities (all targets)",
    color = "Target group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right") +
  coord_fixed()

p_ca
save_fig6_plot(p_ca, "panel_B_ca_ordination_all", width = 10, height = 8)
```

## 11. Export row clustering order for use in Panel C

Save the dendrogram row order so Panel C can use the same neuron ordering
for consistent visual comparison.

```{r export-row-order}
# ---------------------------------------------------------------------------
# Extract row order from the primary heatmap's dendrogram
# ---------------------------------------------------------------------------
row_order_aspf_aspg <- rownames(mat_norm)[hm$tree_row$order]

# Save to an RDS file for Panel C to load
saveRDS(
  row_order_aspf_aspg,
  file = file.path(fig6_dir, "panel_B_row_order.rds")
)

cat("Row order saved. First 10 neurons:\n")
head(row_order_aspf_aspg, 10)

# ---------------------------------------------------------------------------
# Also save the all-targets row order for Panel C extended sections
# ---------------------------------------------------------------------------
row_order_all <- rownames(mat_all_norm)[hm_all$tree_row$order]

saveRDS(
  row_order_all,
  file = file.path(fig6_dir, "panel_B_all_row_order.rds")
)

cat(sprintf("\nAll-targets row order saved: %d neurons\n", length(row_order_all)))
```
