---
title: "Figure 6A: Multimodal Circuit Schematic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel A: Circuit Schematic — Programmatic diagram showing 7 sensory modalities converging through 2nd-order relay neurons (vAB3, PPN1) onto 3rd-order courtship command neurons (aSP-f, aSP-g). Arrows indicate connectivity flow across three circuit layers."
panel_legend <- "(A) Schematic of the multimodal sensory-to-motor circuit. Seven sensory modalities (cVA, mAL, IR, Z4, auditory, visual, mechanosensory) converge through second-order relay neurons (vAB3, PPN1) onto third-order courtship command neurons (aSP-f, aSP-g). Arrow width represents aggregate path strength."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_A_description", "txt"), "panel_A_description.txt"))
```

## Overview

Panel A is a programmatic circuit schematic showing 7 sensory modalities
converging through second-order relay neurons (vAB3, PPN1) to third-order
courtship command neurons (aSP-f, aSP-g). The layout has three columns
representing circuit layers, with arrows indicating connectivity flow.

## Build node layout

Define positions for all nodes across three columns (layers): sensory inputs,
second-order relays, and third-order targets.

```{r node-layout}
# ---------------------------------------------------------------------------
# Column x-positions for the three circuit layers
# ---------------------------------------------------------------------------
x_sensory  <- 0
x_relay    <- 5
x_target   <- 10

# ---------------------------------------------------------------------------
# Sensory input nodes (left column)
# Ordered top-to-bottom: DA1, VA1v, VA1d, JO-B, LC10, ppk23, ppk25
# ---------------------------------------------------------------------------
sensory_labels <- c("DA1", "VA1v", "VA1d", "JO-B", "LC10", "ppk23", "ppk25")
sensory_keys   <- c("DA1", "VA1v", "VA1d", "auditory", "visual", "ppk23", "ppk25")

n_sensory <- length(sensory_labels)
sensory_y <- seq(from = n_sensory, to = 1, length.out = n_sensory)

nodes_sensory <- data.frame(
  label = sensory_labels,
  key   = sensory_keys,
  x     = x_sensory,
  y     = sensory_y,
  layer = "sensory",
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Relay nodes (middle column)
# ---------------------------------------------------------------------------
nodes_relay <- data.frame(
  label = c("vAB3", "PPN1"),
  key   = c("vAB3", "PPN1"),
  x     = x_relay,
  y     = c(5.5, 2.5),
  layer = "relay",
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Target nodes (right column)
# ---------------------------------------------------------------------------
nodes_target <- data.frame(
  label = c("aSP-f", "aSP-g", "..."),
  key   = c("aSP-f", "aSP-g", "other"),
  x     = x_target,
  y     = c(5.5, 4.0, 2.5),
  layer = "target",
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Combine all nodes
# ---------------------------------------------------------------------------
nodes <- bind_rows(nodes_sensory, nodes_relay, nodes_target)
```

## Build edges

Connect every sensory input to every relay neuron, and every relay neuron
to every target neuron, reflecting the convergence/divergence topology.

```{r edges}
# ---------------------------------------------------------------------------
# Sensory -> relay edges (all-to-all)
# ---------------------------------------------------------------------------
edges_s2r <- expand.grid(
  from = sensory_labels,
  to   = nodes_relay$label,
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Relay -> target edges (all-to-all)
# ---------------------------------------------------------------------------
edges_r2t <- expand.grid(
  from = nodes_relay$label,
  to   = nodes_target$label,
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Look up coordinates for edge endpoints
# ---------------------------------------------------------------------------
node_lookup <- nodes %>% select(label, x, y)

add_coords <- function(edges_df) {
  edges_df %>%
    left_join(node_lookup, by = c("from" = "label")) %>%
    rename(x_from = x, y_from = y) %>%
    left_join(node_lookup, by = c("to" = "label")) %>%
    rename(x_to = x, y_to = y)
}

edges_s2r <- add_coords(edges_s2r)
edges_r2t <- add_coords(edges_r2t)
edges_all <- bind_rows(edges_s2r, edges_r2t)
```

## Build group annotations

Label the sensory modality groups on the far left of the diagram.

```{r group-annotations}
# ---------------------------------------------------------------------------
# Map each sensory node to its group using modality_group from setup.R
# ---------------------------------------------------------------------------
group_mapping <- data.frame(
  label = sensory_labels,
  key   = sensory_keys,
  stringsAsFactors = FALSE
) %>%
  mutate(group = modality_group[as.character(key)])

# Compute y-position for each group label (mean y of its member nodes)
group_annot <- group_mapping %>%
  left_join(nodes_sensory %>% select(label, y), by = "label") %>%
  group_by(group) %>%
  summarise(
    y_mid = mean(y),
    y_min = min(y) - 0.35,
    y_max = max(y) + 0.35,
    .groups = "drop"
  ) %>%
  mutate(
    x = x_sensory - 2.8,
    group_label = str_to_title(group)
  )
```

## Draw schematic

Assemble the diagram with boxes, arrows, labels, and group annotations.

```{r schematic, fig.width=12, fig.height=7}
# ---------------------------------------------------------------------------
# Box dimensions
# ---------------------------------------------------------------------------
box_hw <- 0.8   # half-width
box_hh <- 0.35  # half-height
arrow_nudge_x <- box_hw + 0.05

# ---------------------------------------------------------------------------
# Pre-compute all colour columns on plain data.frames
# ---------------------------------------------------------------------------
nodes_df <- as.data.frame(nodes)
nodes_df$fill_color <- ifelse(
  nodes_df$layer == "sensory",
  unname(modality_colors[as.character(nodes_df$key)]),
  ifelse(nodes_df$layer == "relay", "#D9D9D9",
         ifelse(nodes_df$layer == "target", "#BDBDBD", "#F0F0F0"))
)

edges_df <- as.data.frame(edges_all)

group_df <- as.data.frame(group_annot)
group_df$group_color <- unname(modality_group_colors[as.character(group_df$group)])

layer_labels <- data.frame(
  label = c("Sensory\ninputs", "2nd-order\nneurons", "3rd-order\ntargets"),
  x     = c(x_sensory, x_relay, x_target),
  y     = max(sensory_y) + 1.0,
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Main plot
# ---------------------------------------------------------------------------
p <- ggplot() +
  geom_segment(
    data = edges_df,
    aes(x = x_from + arrow_nudge_x, xend = x_to - arrow_nudge_x,
        y = y_from, yend = y_to),
    arrow     = grid::arrow(length = grid::unit(0.12, "cm"), type = "closed"),
    linewidth = 0.35, color = "grey55", alpha = 0.5
  ) +
  geom_rect(
    data = nodes_df,
    aes(xmin = x - box_hw, xmax = x + box_hw,
        ymin = y - box_hh, ymax = y + box_hh,
        fill = fill_color),
    color = "grey30", linewidth = 0.4
  ) +
  scale_fill_identity() +
  geom_text(
    data = nodes_df,
    aes(x = x, y = y, label = label),
    color = "black", size = 3.5, fontface = "bold"
  ) +
  geom_segment(
    data = group_df,
    aes(x = x + 0.8, xend = x + 0.8, y = y_min, yend = y_max,
        color = group_color),
    linewidth = 0.6
  ) +
  scale_color_identity() +
  geom_text(
    data = group_df,
    aes(x = x + 0.35, y = y_mid, label = group_label, color = group_color),
    size = 3.2, fontface = "italic", hjust = 0.5
  ) +
  geom_text(
    data = layer_labels,
    aes(x = x, y = y, label = label),
    size = 3.8, fontface = "bold", color = "grey25", lineheight = 0.85
  ) +
  coord_fixed(ratio = 0.75, clip = "off") +
  xlim(x_sensory - 3.5, x_target + 1.5) +
  ylim(0.2, max(sensory_y) + 1.5) +
  theme_void() +
  theme(
    plot.margin = margin(10, 10, 10, 10, "pt"),
    plot.title  = element_text(hjust = 0.5, face = "bold", size = 14)
  ) +
  ggtitle("Multimodal sensory convergence onto courtship command neurons")

p
```

## Save

```{r save}
save_fig6_plot(p, "panel_A_circuit_schematic", width = 12, height = 7)
```
