---
title: "Figure 6D: Entropy / Multimodality Score"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel D: Multimodality / Entropy — Quantifies how multimodal each target neuron is using Shannon entropy over the 7-modality strength distribution. Scatter plots show total strength vs. entropy and dominance vs. entropy, with separate views for aSP-f/aSP-g, vAB3/PPN1, and all targets combined. Includes valence-split analysis."
panel_legend <- "(D) Multimodality quantified by Shannon entropy over the 7-modality strength distribution for each target neuron. High entropy indicates balanced multimodal input; low entropy indicates unimodal dominance. Scatter plots show strength vs. entropy and dominance vs. entropy for aSP-f/aSP-g, vAB3/PPN1, and all targets. Boxplots compare entropy distributions across target sets and valence classes."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_D_description", "txt"), "panel_D_description.txt"))
```

## Overview

This panel quantifies how multimodal each target neuron is using Shannon entropy
over the distribution of incoming path strengths across the 7 sensory modalities.
A neuron receiving perfectly balanced input from all modalities has maximum entropy
(log2(7) ~ 2.807); a neuron driven by a single modality has entropy 0.

We compute entropy, dominant modality, modality breadth (number of non-zero inputs),
and dominance (fraction of total strength from the dominant modality) for every
target neuron in the combined aSP-f / aSP-g path set.

## 1. Compute multimodality metrics

```{r compute-entropy-metrics}
# Maximum possible entropy for 7 modalities
max_entropy <- log2(length(modality_order))  # log2(7) ~ 2.807

# Build a tidy summary table from the raw strength matrix
# Each row = one target neuron; columns = per-modality strengths
entropy_df <- as.data.frame(strength.aspf_aspg) %>%

tibble::rownames_to_column("neuron") %>%
  rowwise() %>%
  mutate(
    # Shannon entropy across the 7 modality columns
    entropy = shannon_entropy(c_across(all_of(modality_order))),

    # Dominant modality: which column carries the most strength
    dominant_modality = modality_order[which.max(c_across(all_of(modality_order)))],

    # Breadth: how many modalities have non-zero input
    breadth = sum(c_across(all_of(modality_order)) > 0),

    # Total strength across all modalities
    total_strength = sum(c_across(all_of(modality_order))),

    # Dominance: fraction of total from the dominant modality
    dominance = max(c_across(all_of(modality_order))) / total_strength
  ) %>%
  ungroup() %>%
  # Normalised entropy (0-1 scale)
  mutate(
    entropy_norm = entropy / max_entropy,
    dominance = ifelse(is.finite(dominance), dominance, NA_real_)
  )

# Assign target_set based on which path set the neuron belongs to
aspf_neurons <- unique(paths.aspf$end)
aspg_neurons <- unique(paths.aspg$end)

entropy_df <- entropy_df %>%
  mutate(
    target_set = case_when(
      neuron %in% aspf_neurons & neuron %in% aspg_neurons ~ "both",
      neuron %in% aspf_neurons ~ "aSP-f",
      neuron %in% aspg_neurons ~ "aSP-g",
      TRUE ~ "unknown"
    )
  )

# Quick summary
message(sprintf("Total target neurons: %d", nrow(entropy_df)))
message(sprintf("Mean entropy: %.3f  (max possible: %.3f)", mean(entropy_df$entropy), max_entropy))
message(sprintf("Median breadth: %d modalities", median(entropy_df$breadth)))
```

```{r entropy-summary-table}
# Show the top 20 most multimodal neurons
entropy_df %>%
  arrange(desc(entropy)) %>%
  select(neuron, entropy, entropy_norm, dominant_modality, breadth, dominance, target_set) %>%
  head(20) %>%
  knitr::kable(digits = 3, caption = "Top 20 most multimodal target neurons (by entropy)")
```

## 2. Bar chart: neurons ranked by entropy

```{r bar-entropy-ranked, fig.width=14, fig.height=6}
# Order neurons by descending entropy
entropy_ranked <- entropy_df %>%
  arrange(desc(entropy)) %>%
  mutate(rank = row_number(),
         neuron_label = factor(neuron, levels = neuron))

p_bar <- ggplot(entropy_ranked, aes(x = neuron_label, y = entropy, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  annotate("text", x = nrow(entropy_ranked) * 0.95, y = max_entropy + 0.05,
           label = sprintf("max = log2(7) = %.2f", max_entropy),
           hjust = 1, size = 3, color = "grey40") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Shannon entropy (bits)",
    title = "Multimodality score across target neurons"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

print(p_bar)
save_fig6_plot(p_bar, "panel_D_entropy_bar_aspf_aspg", width = 14, height = 6)
```

## 3. Bar chart faceted by target set

```{r bar-entropy-faceted, fig.width=14, fig.height=8}
# Separate ranking within each target set
entropy_faceted <- entropy_df %>%
  group_by(target_set) %>%
  arrange(desc(entropy), .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup()

p_bar_faceted <- ggplot(entropy_faceted, aes(x = reorder(neuron, -entropy), y = entropy,
                                              fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  facet_wrap(~ target_set, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Shannon entropy (bits)",
    title = "Multimodality by target set (aSP-f vs aSP-g)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

print(p_bar_faceted)
save_fig6_plot(p_bar_faceted, "panel_D_entropy_bar_faceted_aspf_aspg", width = 14, height = 8)
```

```{r entropy-distribution-comparison, fig.width=2.5, fig.height=5}
# Statistical comparison of entropy between target sets
p_box <- ggplot(entropy_df, aes(x = target_set, y = entropy, fill = target_set)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.2, stroke = 0) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = "Target set",
    y = "Shannon entropy (bits)",
    title = "Entropy by target set"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_box)
save_fig6_plot(p_box, "panel_D_entropy_boxplot_targetset_aspf_aspg", width = 2.5, height = 5)
```

## 4. Scatter plot: total strength vs entropy

```{r scatter-strength-entropy, fig.width=10, fig.height=8}
# Identify neurons to label: those with extreme or interesting values
top_n_label <- 15
label_neurons <- entropy_df %>%
  arrange(desc(entropy)) %>%
  slice_head(n = top_n_label) %>%
  pull(neuron)

entropy_df <- entropy_df %>%
  mutate(show_label = neuron %in% label_neurons)

p_scatter <- ggplot(entropy_df, aes(x = total_strength, y = entropy,
                                    fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    data = filter(entropy_df, show_label),
    aes(label = neuron),
    color = "grey35",
    size = 2, max.overlaps = 20, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_brewer(palette = "Set1", name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Total path strength (sum across modalities)",
    y = "Shannon entropy (bits)",
    title = "Strength vs. entropy"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_scatter)
save_fig6_plot(p_scatter, "panel_D_strength_vs_entropy_aspf_aspg", width = 7, height = 5.5)
```

## 5. Valence-split analysis (excitatory vs inhibitory)

```{r compute-entropy-by-valence}
# Helper to compute entropy metrics from a strength matrix
compute_entropy_from_matrix <- function(mat, label) {
  df <- as.data.frame(mat) %>%
    tibble::rownames_to_column("neuron") %>%
    rowwise() %>%
    mutate(
      entropy = shannon_entropy(c_across(all_of(intersect(modality_order, colnames(mat))))),
      dominant_modality = {
        vals <- c_across(all_of(intersect(modality_order, colnames(mat))))
        mod_names <- intersect(modality_order, colnames(mat))
        if (all(vals == 0)) NA_character_ else mod_names[which.max(vals)]
      },
      breadth = sum(c_across(all_of(intersect(modality_order, colnames(mat)))) > 0),
      total_strength = sum(c_across(all_of(intersect(modality_order, colnames(mat))))),
      dominance = {
        vals <- c_across(all_of(intersect(modality_order, colnames(mat))))
        tot <- sum(vals)
        if (tot > 0) max(vals) / tot else NA_real_
      }
    ) %>%
    ungroup() %>%
    mutate(
      entropy_norm = entropy / max_entropy,
      valence = label
    )

  df
}

entropy_exc <- compute_entropy_from_matrix(strength.aspf_aspg.exc, "excitatory")
entropy_inh <- compute_entropy_from_matrix(strength.aspf_aspg.inh, "inhibitory")
entropy_combined <- compute_entropy_from_matrix(strength.aspf_aspg, "combined")

entropy_valence <- bind_rows(entropy_exc, entropy_inh, entropy_combined)
```

```{r bar-entropy-by-valence, fig.width=14, fig.height=10}
# Ranked bar charts per valence
p_valence_bars <- entropy_valence %>%
  group_by(valence) %>%
  arrange(desc(entropy), .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = rank, y = entropy, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey40", linewidth = 0.4) +
  facet_wrap(~ valence, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality", na.value = "grey80") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Shannon entropy (bits)",
    title = "Multimodality score split by path valence"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

print(p_valence_bars)
save_fig6_plot(p_valence_bars, "panel_D_entropy_bar_valence_aspf_aspg", width = 14, height = 10)
```

```{r boxplot-entropy-by-valence, fig.width=8, fig.height=5}
# Compare entropy distributions across valence types
p_valence_box <- ggplot(entropy_valence, aes(x = valence, y = entropy, fill = valence)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.2, stroke = 0) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(
    values = c("combined" = "#B3B3B3", "excitatory" = "#4DAF4A", "inhibitory" = "#E41A1C")
  ) +
  labs(
    x = "Path valence",
    y = "Shannon entropy (bits)",
    title = "Entropy by valence"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_valence_box)
save_fig6_plot(p_valence_box, "panel_D_entropy_boxplot_valence_aspf_aspg", width = 8, height = 5)
```

```{r scatter-dominance-entropy-valence, fig.width=14, fig.height=6}
# Dominance vs entropy, faceted by valence
p_scatter_valence <- entropy_valence %>%
  filter(is.finite(dominance)) %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = dominance, y = entropy, color = dominant_modality)) +
  geom_point(alpha = 0.65, size = 2) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  facet_wrap(~ valence) +
  scale_color_manual(values = modality_colors, name = "Dominant\nModality", na.value = "grey80") +
  labs(
    x = "Dominance (fraction of strength from top modality)",
    y = "Shannon entropy (bits)",
    title = "Dominance vs. entropy by valence"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

print(p_scatter_valence)
save_fig6_plot(p_scatter_valence, "panel_D_dominance_vs_entropy_valence_aspf_aspg", width = 10, height = 4.5)
```

## 6. Dominance vs entropy relationship

```{r dominance-vs-entropy, fig.width=8, fig.height=6}
# Dominance (fraction from top modality) vs entropy
# Expect strong negative correlation: high entropy = low dominance
p_dominance <- ggplot(entropy_df %>% filter(is.finite(dominance)),
                      aes(x = dominance, y = entropy,
                          fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_brewer(palette = "Set1", name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Dominance (fraction of strength from top modality)",
    y = "Shannon entropy (bits)",
    title = "Dominance vs. entropy"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_dominance)
save_fig6_plot(p_dominance, "panel_D_dominance_vs_entropy_aspf_aspg", width = 7, height = 5)
```

---

## 7. vAB3 + PPN1 targets

```{r entropy-vab3-ppn1}
# Compute strength matrix for vAB3 + PPN1 targets
paths.vab3_ppn1 <- bind_rows(paths.vab3, paths.ppn1)
strength.vab3_ppn1 <- compute_strength_matrix(paths.vab3_ppn1)

vab3_neurons <- unique(paths.vab3$end)
ppn1_neurons <- unique(paths.ppn1$end)

entropy_vp <- as.data.frame(strength.vab3_ppn1) %>%
  tibble::rownames_to_column("neuron") %>%
  rowwise() %>%
  mutate(
    entropy = shannon_entropy(c_across(all_of(intersect(modality_order, colnames(strength.vab3_ppn1))))),
    dominant_modality = {
      vals <- c_across(all_of(intersect(modality_order, colnames(strength.vab3_ppn1))))
      mod_names <- intersect(modality_order, colnames(strength.vab3_ppn1))
      if (all(vals == 0)) NA_character_ else mod_names[which.max(vals)]
    },
    total_strength = sum(c_across(all_of(intersect(modality_order, colnames(strength.vab3_ppn1))))),
    dominance = {
      vals <- c_across(all_of(intersect(modality_order, colnames(strength.vab3_ppn1))))
      tot <- sum(vals)
      if (tot > 0) max(vals) / tot else NA_real_
    }
  ) %>%
  ungroup() %>%
  mutate(
    entropy_norm = entropy / max_entropy,
    target_set = case_when(
      neuron %in% vab3_neurons & neuron %in% ppn1_neurons ~ "both",
      neuron %in% vab3_neurons ~ "vAB3",
      neuron %in% ppn1_neurons ~ "PPN1",
      TRUE ~ "unknown"
    )
  )

cat(sprintf("vAB3+PPN1 target neurons: %d\n", nrow(entropy_vp)))
```

```{r strength-vs-entropy-vp, fig.width=10, fig.height=8}
p_scatter_vp <- ggplot(entropy_vp, aes(x = total_strength, y = entropy,
                                        fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 2, max.overlaps = 20, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_manual(values = c("vAB3" = "#4DAF4A", "PPN1" = "#FF7F00", "both" = "#984EA3"),
                     name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Total path strength (sum across modalities)",
    y = "Shannon entropy (bits)",
    title = "Strength vs. entropy (vAB3+PPN1)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_scatter_vp)
save_fig6_plot(p_scatter_vp, "panel_D_strength_vs_entropy_vab3_ppn1", width = 7, height = 5.5)
```

```{r dominance-vs-entropy-vp, fig.width=8, fig.height=6}
p_dominance_vp <- ggplot(entropy_vp %>% filter(is.finite(dominance)),
                          aes(x = dominance, y = entropy,
                              fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_manual(values = c("vAB3" = "#4DAF4A", "PPN1" = "#FF7F00", "both" = "#984EA3"),
                     name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Dominance (fraction of strength from top modality)",
    y = "Shannon entropy (bits)",
    title = "Dominance vs. entropy (vAB3+PPN1)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_dominance_vp)
save_fig6_plot(p_dominance_vp, "panel_D_dominance_vs_entropy_vab3_ppn1", width = 7, height = 5)
```

```{r entropy-boxplot-vp, fig.width=2.5, fig.height=5}
p_box_vp <- ggplot(entropy_vp, aes(x = target_set, y = entropy, fill = target_set)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.2, stroke = 0) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = c("vAB3" = "#4DAF4A", "PPN1" = "#FF7F00", "both" = "#984EA3")) +
  labs(
    x = "Target set",
    y = "Shannon entropy (bits)",
    title = "Entropy by target set (vAB3+PPN1)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_box_vp)
save_fig6_plot(p_box_vp, "panel_D_entropy_boxplot_targetset_vab3_ppn1", width = 2.5, height = 5)
```

---

## 8. All targets (aSP-f + aSP-g + vAB3 + PPN1)

```{r entropy-all-targets}
entropy_all <- as.data.frame(strength.all) %>%
  tibble::rownames_to_column("neuron") %>%
  rowwise() %>%
  mutate(
    entropy = shannon_entropy(c_across(all_of(intersect(modality_order, colnames(strength.all))))),
    dominant_modality = {
      vals <- c_across(all_of(intersect(modality_order, colnames(strength.all))))
      mod_names <- intersect(modality_order, colnames(strength.all))
      if (all(vals == 0)) NA_character_ else mod_names[which.max(vals)]
    },
    total_strength = sum(c_across(all_of(intersect(modality_order, colnames(strength.all))))),
    dominance = {
      vals <- c_across(all_of(intersect(modality_order, colnames(strength.all))))
      tot <- sum(vals)
      if (tot > 0) max(vals) / tot else NA_real_
    }
  ) %>%
  ungroup() %>%
  mutate(
    entropy_norm = entropy / max_entropy,
    target_set = case_when(
      neuron %in% aspf_neurons & neuron %in% aspg_neurons ~ "aSP-f+g",
      neuron %in% aspf_neurons ~ "aSP-f",
      neuron %in% aspg_neurons ~ "aSP-g",
      neuron %in% vab3_neurons & neuron %in% ppn1_neurons ~ "vAB3+PPN1",
      neuron %in% vab3_neurons ~ "vAB3",
      neuron %in% ppn1_neurons ~ "PPN1",
      TRUE ~ "unknown"
    )
  )

cat(sprintf("All target neurons: %d\n", nrow(entropy_all)))
```

```{r strength-vs-entropy-all, fig.width=10, fig.height=8}
all_target_colors <- c("aSP-f" = "#E41A1C", "aSP-g" = "#377EB8", "aSP-f+g" = "#984EA3",
                        "vAB3" = "#4DAF4A", "PPN1" = "#FF7F00", "vAB3+PPN1" = "#A65628",
                        "unknown" = "grey50")

p_scatter_all <- ggplot(entropy_all, aes(x = total_strength, y = entropy,
                                          fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_manual(values = all_target_colors, name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Total path strength (sum across modalities)",
    y = "Shannon entropy (bits)",
    title = "Strength vs. entropy (all targets)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_scatter_all)
save_fig6_plot(p_scatter_all, "panel_D_strength_vs_entropy_all_targets", width = 7, height = 5.5)
```

```{r dominance-vs-entropy-all, fig.width=8, fig.height=6}
p_dominance_all <- ggplot(entropy_all %>% filter(is.finite(dominance)),
                           aes(x = dominance, y = entropy,
                               fill = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_color_manual(values = all_target_colors, name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.4)),
         color = guide_legend(override.aes = list(fill = "white", size = 2))) +
  labs(
    x = "Dominance (fraction of strength from top modality)",
    y = "Shannon entropy (bits)",
    title = "Dominance vs. entropy (all targets)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_dominance_all)
save_fig6_plot(p_dominance_all, "panel_D_dominance_vs_entropy_all_targets", width = 7, height = 5)
```

```{r entropy-boxplot-all, fig.width=4, fig.height=5}
p_box_all <- ggplot(entropy_all, aes(x = target_set, y = entropy, fill = target_set)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.2, stroke = 0) +
  geom_hline(yintercept = max_entropy, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = all_target_colors) +
  labs(
    x = "Target set",
    y = "Shannon entropy (bits)",
    title = "Entropy (all targets)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(p_box_all)
save_fig6_plot(p_box_all, "panel_D_entropy_boxplot_targetset_all", width = 4, height = 5)
```

---

## 9. Summary statistics

```{r summary-stats}
# Overall summary
cat("=== Entropy Summary ===\n")
cat(sprintf("Number of target neurons: %d\n", nrow(entropy_df)))
cat(sprintf("Max possible entropy: %.3f bits (log2(7))\n", max_entropy))
cat(sprintf("Mean entropy: %.3f bits\n", mean(entropy_df$entropy)))
cat(sprintf("Median entropy: %.3f bits\n", median(entropy_df$entropy)))
cat(sprintf("SD entropy: %.3f bits\n", sd(entropy_df$entropy)))
cat(sprintf("Mean breadth: %.1f modalities\n", mean(entropy_df$breadth)))
cat(sprintf("Neurons with all 7 modalities: %d (%.1f%%)\n",
            sum(entropy_df$breadth == 7),
            100 * mean(entropy_df$breadth == 7)))
cat(sprintf("Neurons with only 1 modality: %d (%.1f%%)\n",
            sum(entropy_df$breadth == 1),
            100 * mean(entropy_df$breadth == 1)))

# Per target set
cat("\n=== By Target Set ===\n")
entropy_df %>%
  group_by(target_set) %>%
  summarise(
    n = n(),
    mean_entropy = mean(entropy),
    median_entropy = median(entropy),
    mean_breadth = mean(breadth),
    mean_dominance = mean(dominance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  knitr::kable(digits = 3)
```

```{r session-info}
sessionInfo()
```
