---
title: "Figure 6F: Path Diversity -- Concentrated vs. Distributed Routes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel F: Path Diversity — Measures how concentrated or distributed the strongest paths are per modality. For each (target, modality) pair, computes the number of paths needed to accumulate 80%% of total strength. Dot plots and bar charts compare path diversity across modalities and between excitatory/inhibitory valence classes."
panel_legend <- "(F) Path diversity per modality. The number of paths required to accumulate 80%% of total strength measures how concentrated or distributed signal routing is. Dot plots show per-neuron values with median overlay; bar charts summarize means with standard errors. Faceted views split by excitatory and inhibitory valence."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_F_description", "txt"), "panel_F_description.txt"))
```

## Overview

Panel F examines how concentrated or distributed the strongest paths are for
each sensory modality reaching each target neuron. Some modalities may funnel
signal through a few dominant paths, while others spread it across many weaker
routes. We quantify this by computing the number of paths required to
accumulate 80% of total strength per (target, modality) pair, then visualise
the distribution across modalities and valence classes.

## 1. Compute cumulative strength and paths-to-80%

For each (end neuron, modality) combination, rank paths by descending strength,
compute the cumulative fraction of total strength, and identify the minimum
number of paths needed to capture 80% of total strength.

```{r cumulative-strength}
# ---------------------------------------------------------------------------
# Rank paths within each (end, modality) group and compute cumulative fraction
# ---------------------------------------------------------------------------
cumulative_df <- paths.aspf_aspg %>%
  filter(!is.na(strength), strength > 0) %>%
  group_by(end, modality) %>%
  arrange(desc(strength), .by_group = TRUE) %>%
  mutate(
    rank           = row_number(),
    total_strength = sum(strength),
    cum_strength   = cumsum(strength),
    cum_frac       = cum_strength / total_strength
  ) %>%
  ungroup()

# ---------------------------------------------------------------------------
# For each (end, modality), find paths needed to reach 80% of total strength
# ---------------------------------------------------------------------------
paths_to_80 <- cumulative_df %>%
  group_by(end, modality) %>%
  summarise(
    n_paths        = n(),
    total_strength = first(total_strength),
    paths_to_80pct = min(rank[cum_frac >= 0.80], na.rm = TRUE),
    .groups        = "drop"
  ) %>%
  # If all paths together still < 80% (rare edge case), set to total n_paths

  mutate(
    paths_to_80pct = ifelse(is.infinite(paths_to_80pct), n_paths, paths_to_80pct)
  )

# Enforce modality ordering
paths_to_80 <- paths_to_80 %>%
  mutate(modality = factor(modality, levels = modality_order))
```

## 2. Summary statistics by modality

```{r summary-stats}
# ---------------------------------------------------------------------------
# Mean, median, and SE of paths_to_80pct per modality
# ---------------------------------------------------------------------------
summary_80 <- paths_to_80 %>%
  group_by(modality) %>%
  summarise(
    n_targets   = n(),
    mean_p80    = mean(paths_to_80pct),
    median_p80  = median(paths_to_80pct),
    sd_p80      = sd(paths_to_80pct),
    se_p80      = sd(paths_to_80pct) / sqrt(n()),
    .groups     = "drop"
  )

summary_80
```

## 3. Plot 1 -- Dot plot: paths-to-80% by modality

Individual points per target neuron with jitter, overlaid with the modality
median.

```{r dot-plot, fig.width=9, fig.height=6}
# ---------------------------------------------------------------------------
# Dot plot with jitter + median overlay
# ---------------------------------------------------------------------------
p_dot <- ggplot(paths_to_80, aes(x = modality, y = paths_to_80pct)) +
  geom_jitter(
    aes(color = modality),
    width  = 0.2,
    alpha  = 0.4,
    size   = 1.2,
    stroke = 0
  ) +
  stat_summary(
    fun    = median,
    geom   = "point",
    aes(color = modality),
    size   = 3,
    shape  = 18
  ) +
  scale_color_manual(values = modality_colors) +
  labs(
    x     = "Sensory modality",
    y     = "Paths needed for 80% of total strength",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

p_dot
```

```{r save-dot}
save_fig6_plot(p_dot, "panel_F_dot_paths_to_80pct_aspf_aspg", width = 7, height = 5)
```

## 4. Plot 2 -- Cumulative strength curves (faceted by target)

Select a subset of target neurons that receive input from multiple modalities,
then plot cumulative strength curves coloured by modality.

```{r select-interesting-targets}
# ---------------------------------------------------------------------------
# Pick targets that are reached by at least 4 modalities and have
# reasonable total path counts, for readable facets
# ---------------------------------------------------------------------------
targets_multi <- paths_to_80 %>%
  group_by(end) %>%
  summarise(n_modalities = n_distinct(modality), .groups = "drop") %>%
  filter(n_modalities >= 4) %>%
  pull(end)

# If too many, take the top 12 by total path strength
if (length(targets_multi) > 12) {
  top_targets <- cumulative_df %>%
    filter(end %in% targets_multi) %>%
    group_by(end) %>%
    summarise(total = sum(strength), .groups = "drop") %>%
    slice_max(total, n = 12) %>%
    pull(end)
} else {
  top_targets <- targets_multi
}

# Fall back: if fewer than 4 multi-modality targets, just take top 9 by strength
if (length(top_targets) < 4) {
  top_targets <- cumulative_df %>%
    group_by(end) %>%
    summarise(total = sum(strength), .groups = "drop") %>%
    slice_max(total, n = 9) %>%
    pull(end)
}

message(sprintf("Showing cumulative curves for %d target neurons", length(top_targets)))
```

```{r cumulative-curves, fig.width=12, fig.height=8}
# ---------------------------------------------------------------------------
# Faceted cumulative strength curves
# ---------------------------------------------------------------------------
cum_subset <- cumulative_df %>%
  filter(end %in% top_targets) %>%
  mutate(modality = factor(modality, levels = modality_order))

p_cum <- ggplot(cum_subset, aes(x = rank, y = cum_frac, color = modality)) +
  geom_line(linewidth = 0.8, alpha = 0.85) +
  geom_hline(yintercept = 0.80, linetype = "dashed", color = "grey40", linewidth = 0.4) +
  facet_wrap(~end, scales = "free_x", ncol = 4) +
  scale_color_manual(values = modality_colors) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1.02)) +
  labs(
    x     = "Path rank (by descending strength)",
    y     = "Cumulative fraction of total strength",
    color = "Modality",
    title = "Cumulative path strength saturation per target neuron"
  ) +
  theme_bw(base_size = 11) +
  theme(
    strip.text       = element_text(size = 8, face = "bold"),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )

p_cum
```

```{r save-cumulative}
save_fig6_plot(p_cum, "panel_F_cumulative_curves_aspf_aspg", width = 12, height = 8)
```

## 5. Plot 3 -- Summary bar chart (publication figure)

Clean bar chart showing mean paths-to-80% per modality with standard error
bars. This is the primary panel for the publication figure.

```{r bar-chart, fig.width=7, fig.height=5}
# ---------------------------------------------------------------------------
# Summary bar chart: mean +/- SE
# ---------------------------------------------------------------------------
p_bar <- ggplot(summary_80, aes(x = modality, y = mean_p80, fill = modality)) +
  geom_col(width = 0.65, color = "grey30", linewidth = 0.3) +
  geom_errorbar(
    aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
    width    = 0.2,
    linewidth = 0.4,
    color    = "grey20"
  ) +
  scale_fill_manual(values = modality_colors) +
  labs(
    x     = "Sensory modality",
    y     = "Mean paths to 80% strength",
    title = "Path diversity by modality"
  ) +
  theme_bw(base_size = 13) +
  theme(
    legend.position  = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x      = element_text(angle = 30, hjust = 1)
  )

p_bar
```

```{r save-bar}
save_fig6_plot(p_bar, "panel_F_bar_paths_to_80pct_aspf_aspg", width = 7, height = 5)
```

## 6. Valence split -- excitatory vs. inhibitory path diversity

Repeat the analysis splitting paths by valence to test whether excitatory
and inhibitory routes differ in their concentration.

```{r valence-split}
# ---------------------------------------------------------------------------
# Compute paths_to_80pct separately for each valence class
# ---------------------------------------------------------------------------
cumulative_valence <- paths.aspf_aspg %>%
  filter(!is.na(strength), strength > 0, !is.na(valence)) %>%
  group_by(end, modality, valence) %>%
  arrange(desc(strength), .by_group = TRUE) %>%
  mutate(
    rank           = row_number(),
    total_strength = sum(strength),
    cum_strength   = cumsum(strength),
    cum_frac       = cum_strength / total_strength
  ) %>%
  ungroup()

paths_to_80_valence <- cumulative_valence %>%
  group_by(end, modality, valence) %>%
  summarise(
    n_paths        = n(),
    total_strength = first(total_strength),
    paths_to_80pct = min(rank[cum_frac >= 0.80], na.rm = TRUE),
    .groups        = "drop"
  ) %>%
  mutate(
    paths_to_80pct = ifelse(is.infinite(paths_to_80pct), n_paths, paths_to_80pct),
    modality       = factor(modality, levels = modality_order)
  )

# Summary by modality x valence
summary_valence <- paths_to_80_valence %>%
  group_by(modality, valence) %>%
  summarise(
    n_targets  = n(),
    mean_p80   = mean(paths_to_80pct),
    median_p80 = median(paths_to_80pct),
    se_p80     = sd(paths_to_80pct) / sqrt(n()),
    .groups    = "drop"
  )
```

```{r valence-bar, fig.width=9, fig.height=5}
# ---------------------------------------------------------------------------
# Grouped bar chart: valence side-by-side within each modality
# ---------------------------------------------------------------------------
p_valence <- ggplot(
  summary_valence,
  aes(x = modality, y = mean_p80, fill = modality, alpha = valence)
) +
  geom_col(
    position = position_dodge(width = 0.7),
    width    = 0.6,
    color    = "grey30",
    linewidth = 0.3
  ) +
  geom_errorbar(
    aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
    position  = position_dodge(width = 0.7),
    width     = 0.2,
    linewidth = 0.4,
    color     = "grey20"
  ) +
  scale_fill_manual(values = modality_colors) +
  scale_alpha_manual(
    values = c("excitatory" = 1.0, "inhibitory" = 0.5),
    name   = "Valence"
  ) +
  labs(
    x     = "Sensory modality",
    y     = "Mean paths to 80% strength",
    title = "Path diversity by modality and valence"
  ) +
  guides(fill = "none") +
  theme_bw(base_size = 13) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 30, hjust = 1),
    legend.position    = "top"
  )

p_valence
```

```{r save-valence}
save_fig6_plot(p_valence, "panel_F_valence_paths_to_80pct_aspf_aspg", width = 9, height = 5)
```

## 7. Dot plot split by valence

```{r valence-dot, fig.width=10, fig.height=5}
# ---------------------------------------------------------------------------
# Faceted dot plot: excitatory vs inhibitory
# ---------------------------------------------------------------------------
p_val_dot <- ggplot(
  paths_to_80_valence,
  aes(x = modality, y = paths_to_80pct)
) +
  geom_jitter(
    aes(color = modality),
    width  = 0.2,
    alpha  = 0.4,
    size   = 1.2,
    stroke = 0
  ) +
  stat_summary(
    fun    = median,
    geom   = "point",
    aes(color = modality),
    size   = 3,
    shape  = 18
  ) +
  facet_wrap(~valence) +
  scale_color_manual(values = modality_colors) +
  labs(
    x     = "Sensory modality",
    y     = "Paths needed for 80% of total strength",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

p_val_dot
```

```{r save-valence-dot}
save_fig6_plot(p_val_dot, "panel_F_valence_dot_paths_to_80pct_aspf_aspg", width = 8, height = 4.5)
```

## Session info

```{r session-info}
sessionInfo()
```
