---
title: "Figure 6G: Shared Intermediate Nodes -- Hidden Hubs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel G: Shared Intermediate Nodes — Identifies hidden integration hubs: intermediate neurons that appear in the strongest paths from multiple sensory modalities. Ranks intermediates by total path strength and Shannon entropy, shows modality strength heatmaps, source vs. target diversity scatter, and co-innervation analysis. AN neurons (AN09B017, AN05B102 and variants) are excluded."
panel_legend <- "(G) Intermediate integration hubs shared across modalities. Neurons appearing in strongest paths from multiple sensory modalities are ranked by total path strength and Shannon entropy. Heatmaps show per-modality strength profiles; scatter plots display source diversity vs. target diversity, identifying convergence-divergence hubs."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_G_description", "txt"), "panel_G_description.txt"))
```

## Overview

Panel G identifies intermediate neurons that appear in the strongest paths
from multiple sensory modalities, revealing hidden integration hubs in the
circuit. These are neurons that are neither the sensory start nor the final
target, yet they sit at the crossroads of multiple modality-specific routes.
We parse all paths, extract intermediate nodes, count how many modalities
route through each node, and weight the analysis by path strength.

## 1. Define sensory neurons to exclude

Sensory start neurons appear as intermediates in some paths but are not
meaningful integration points. We exclude them from the analysis.

```{r sensory-exclusions}
# ---------------------------------------------------------------------------
# Sensory neuron types that should be excluded from intermediate analysis.
# These are the start nodes of the various modality paths.
# ---------------------------------------------------------------------------
sensory_patterns <- c(
  "^ORN_DA1",
  "^ORN_VA1v",
  "^ORN_VA1d",
  "^JO-B",
  "^LC10",
  "^ppk23",
  "^ppk25"
)

# Build a single regex for fast matching
sensory_regex <- paste(sensory_patterns, collapse = "|")
```

## 2. Define analysis function and target sets

```{r define-path-sets}
# ---------------------------------------------------------------------------
# Define the three target sets to analyze
# ---------------------------------------------------------------------------
path_sets <- list(
  aspf_aspg   = list(paths = paths.aspf_aspg,   label = "aSP-f/aSP-g", suffix = "aspf_aspg"),
  vab3_ppn1   = list(paths = bind_rows(paths.vab3, paths.ppn1), label = "vAB3+PPN1", suffix = "vab3_ppn1"),
  all_targets = list(paths = paths.all,          label = "All targets",  suffix = "all_targets")
)

# Neuron type patterns (shared across all sets)
neuron_patterns <- c(
  "aSP-f"  = "^aSP-f",
  "aSP-g"  = "^aSP-g",
  "P1"     = "^pC1_",
  "pC1"    = "^pC1",
  "pC2"    = "^pC2",
  "mAL"    = "^mAL",
  "vAB3"   = "^vAB3",
  "PPN1"   = "^PPN1",
  "LH"     = "^LH",
  "PN"     = "PN",
  "MBON"   = "^MBON",
  "KC"     = "^KC",
  "SMP"    = "^SMP",
  "SIP"    = "^SIP",
  "SLP"    = "^SLP",
  "AVLP"   = "^AVLP",
  "PVLP"   = "^PVLP",
  "WED"    = "^WED",
  "CX"     = "^(FB|EB|PB|NO|EPG|PFN|PFL|hDelta|FC|FR)",
  "AL-local" = "^(lLN|vLN|LN|v2LN|l2LN)",
  "LC"     = "^(LC|LPLC|LPC)",
  "DN"     = "^DN",
  "LAL"    = "^LAL",
  "CRE"    = "^CRE",
  "AOTU"   = "^(AOTU|TuBu)",
  "SEZ"    = "^(SEZ|GNG)"
)

# Neuron class color palette (shared)
neuron_class_colors <- c(
  "aSP-f"  = "#E41A1C", "aSP-g"  = "#377EB8", "P1"    = "#984EA3",
  "pC1"    = "#FF7F00", "pC2"    = "#A65628", "mAL"   = "#4DAF4A",
  "vAB3"   = "#F781BF", "PPN1"   = "#999999",
  "LH"     = "#1B9E77", "PN"     = "#D95F02", "MBON"  = "#7570B3",
  "KC"     = "#E7298A", "AL-local" = "#B3DE69",
  "SMP"    = "#66C2A5", "SIP"    = "#8DD3C7", "SLP"   = "#BEBADA",
  "AVLP"   = "#FB8072", "PVLP"   = "#FCCDE5", "WED"   = "#BC80BD",
  "LC"     = "#80B1D3", "AOTU"   = "#FDB462",
  "CX"     = "#B3DE69", "DN"     = "#FC8D62", "LAL"   = "#CCEBC5",
  "CRE"    = "#FFED6F", "SEZ"    = "#D9D9D9",
  "other"  = "#F0F0F0"
)
```

```{r run-all-sets, fig.width=10, fig.height=8, results='asis'}
for (set_name in names(path_sets)) {

  ps       <- path_sets[[set_name]]
  paths_df <- ps$paths
  label    <- ps$label
  suffix   <- ps$suffix

  cat(sprintf("\n\n---\n\n## Target set: %s\n\n", label))

  # ---- Parse intermediates ----
  intermediates_raw <- paths_df %>%
    filter(!is.na(path), !is.na(strength), strength > 0) %>%
    rowwise() %>%
    mutate(nodes = list(strsplit(path, " -> ", fixed = TRUE)[[1]])) %>%
    ungroup() %>%
    filter(lengths(nodes) >= 3) %>%
    mutate(intermediates = purrr::map(nodes, ~ .x[2:(length(.x) - 1)])) %>%
    select(path, start, end, modality, strength, valence, target_set, intermediates) %>%
    tidyr::unnest(intermediates) %>%
    rename(intermediate = intermediates)

  intermediates_clean <- intermediates_raw %>%
    filter(!grepl(sensory_regex, intermediate),
           !grepl("^AN09B017|^AN05B102", intermediate))

  message(sprintf("[%s] %s intermediate occurrences, %s unique nodes",
    label, format(nrow(intermediates_clean), big.mark = ","),
    format(n_distinct(intermediates_clean$intermediate), big.mark = ",")))

  # ---- Per-intermediate summary ----
  inter_mod_strength <- intermediates_clean %>%
    group_by(intermediate, modality) %>%
    summarise(mod_strength = sum(strength, na.rm = TRUE), n_paths = n(), .groups = "drop")

  inter_diversity <- intermediates_clean %>%
    group_by(intermediate) %>%
    summarise(source_diversity = n_distinct(start), target_diversity = n_distinct(end), .groups = "drop")

  inter_summary <- inter_mod_strength %>%
    group_by(intermediate) %>%
    summarise(n_modalities = n_distinct(modality), total_strength = sum(mod_strength),
              total_paths = sum(n_paths), .groups = "drop") %>%
    left_join(inter_diversity, by = "intermediate") %>%
    arrange(desc(total_strength))

  inter_entropy <- inter_mod_strength %>%
    group_by(intermediate) %>%
    summarise(entropy = shannon_entropy(mod_strength), .groups = "drop")
  inter_summary <- inter_summary %>% left_join(inter_entropy, by = "intermediate")

  inter_dominant <- inter_mod_strength %>%
    group_by(intermediate) %>%
    slice_max(mod_strength, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(intermediate, dominant_modality = modality)
  inter_summary <- inter_summary %>% left_join(inter_dominant, by = "intermediate")

  inter_wide <- inter_mod_strength %>%
    select(intermediate, modality, mod_strength) %>%
    pivot_wider(names_from = modality, values_from = mod_strength, values_fill = 0)
  for (mod in modality_order) {
    if (!mod %in% colnames(inter_wide)) inter_wide[[mod]] <- 0
  }
  inter_wide <- inter_wide %>%
    left_join(inter_summary %>% select(intermediate, total_strength, entropy,
                                        source_diversity, target_diversity),
              by = "intermediate")

  # ---- Annotate neuron types ----
  inter_summary <- inter_summary %>%
    mutate(neuron_class = {
      classes <- character(n())
      for (i in seq_along(neuron_patterns)) {
        mask <- grepl(neuron_patterns[i], intermediate)
        classes[mask & classes == ""] <- names(neuron_patterns)[i]
      }
      classes[classes == ""] <- "other"
      classes
    })

  # ---- Plot 1: Ranked dot plot ----
  n_top <- min(40, nrow(inter_summary))
  top_nodes <- inter_summary %>%
    filter(!intermediate %in% c("AN09B017", "AN05B102")) %>%
    slice_head(n = n_top) %>%
    mutate(intermediate = factor(intermediate, levels = rev(intermediate)))

  p_ranked <- ggplot(top_nodes, aes(x = total_strength, y = intermediate,
                                     color = entropy, shape = neuron_class != "other")) +
    geom_point(size = 3.5) +
    scale_color_viridis_c(option = "inferno", name = "Shannon\nentropy") +
    scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 17),
                       name = "Known courtship\nneuron",
                       labels = c(`FALSE` = "No", `TRUE` = "Yes")) +
    scale_x_log10(labels = scales::comma) +
    labs(x = "Total path strength (log scale)", y = NULL,
         title = sprintf("Top %d intermediates (%s)", n_top, label)) +
    theme_bw(base_size = 11) +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.y = element_line(linewidth = 0.2, color = "grey90"),
          axis.text.y = element_text(size = 7))
  print(p_ranked)
  save_fig6_plot(p_ranked, sprintf("panel_G_ranked_intermediates_%s", suffix), width = 10, height = 8, subdir = suffix)

  # ---- Plot 2: Heatmap ----
  n_heatmap <- min(30, nrow(inter_summary))
  heatmap_nodes <- inter_summary %>% slice_head(n = n_heatmap) %>% pull(intermediate)
  heatmap_mat <- inter_wide %>%
    filter(intermediate %in% heatmap_nodes) %>%
    column_to_rownames("intermediate") %>%
    select(all_of(intersect(modality_order, colnames(.)))) %>%
    as.matrix()
  heatmap_mat_log <- log10(heatmap_mat + 1)

  col_annot <- data.frame(group = modality_group[colnames(heatmap_mat_log)],
                           row.names = colnames(heatmap_mat_log))
  annot_colors <- list(group = modality_group_colors)

  row_class <- inter_summary %>%
    filter(intermediate %in% heatmap_nodes) %>%
    select(intermediate, neuron_class) %>%
    column_to_rownames("intermediate")
  annot_colors$neuron_class <- neuron_class_colors

  p_heat <- pheatmap(heatmap_mat_log, cluster_cols = FALSE, cluster_rows = TRUE,
    clustering_method = "ward.D2", clustering_distance_rows = "euclidean",
    color = colorRampPalette(c("white", "#FDE725", "#35B779", "#31688E", "#440154"))(100),
    border_color = "grey85", annotation_col = col_annot, annotation_row = row_class,
    annotation_colors = annot_colors, fontsize_row = 7, fontsize_col = 10,
    main = sprintf("Modality strength through top intermediates (%s, log10)", label),
    silent = TRUE)

  hm_name <- sprintf("panel_G_heatmap_intermediates_%s", suffix)
  png(file.path(panel_plot_dir(hm_name, "png", subdir = suffix), paste0(hm_name, ".png")),
      width = 8, height = 9, units = "in", res = 300)
  print(p_heat); dev.off()
  pdf(file.path(panel_plot_dir(hm_name, "pdf", subdir = suffix), paste0(hm_name, ".pdf")),
      width = 8, height = 9)
  print(p_heat); dev.off()

  # ---- Plot 3: Strength vs entropy scatters ----
  inter_nonsensory <- inter_summary %>%
    filter(total_strength > 0, !grepl(sensory_regex, intermediate))
  neuron_classes <- sort(unique(inter_nonsensory$neuron_class))
  neuron_type_colors <- setNames(scales::hue_pal()(length(neuron_classes)), neuron_classes)
  if ("other" %in% names(neuron_type_colors)) neuron_type_colors["other"] <- "grey70"

  p_str_ent_type <- ggplot(inter_nonsensory,
    aes(x = total_strength, y = entropy, size = source_diversity, color = neuron_class)) +
    geom_point(alpha = 0.65, size = 2, shape = 16) +
    scale_x_log10(labels = scales::comma) +
    scale_size_continuous(name = "Source div.", range = c(0.8, 4.5), breaks = scales::pretty_breaks(4)) +
    scale_color_manual(values = neuron_type_colors, name = "Type") +
    ggrepel::geom_text_repel(data = inter_nonsensory %>% slice_head(n = 15),
      aes(label = intermediate), size = 2, max.overlaps = 20, color = "grey35",
      segment.color = "grey75", segment.size = 0.25) +
    labs(x = "Total strength", y = "Entropy (bits)", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(),
      legend.title = element_text(size = rel(0.85)),
      legend.text = element_text(size = rel(0.8)),
      legend.key.size = unit(0.4, "cm"))
  print(p_str_ent_type)
  save_fig6_plot(p_str_ent_type, sprintf("panel_G_strength_vs_entropy_neurontype_%s", suffix), width = 7, height = 5, subdir = suffix)

  p_str_ent_mod <- ggplot(inter_nonsensory,
    aes(x = total_strength, y = entropy, size = source_diversity, color = dominant_modality)) +
    geom_point(alpha = 0.65, size = 2, shape = 16) +
    scale_x_log10(labels = scales::comma) +
    scale_size_continuous(name = "Source div.", range = c(0.8, 4.5), breaks = scales::pretty_breaks(4)) +
    scale_color_manual(values = modality_colors, name = "Modality") +
    ggrepel::geom_text_repel(data = inter_nonsensory %>% slice_head(n = 15),
      aes(label = intermediate), size = 2, max.overlaps = 20, color = "grey35",
      segment.color = "grey75", segment.size = 0.25) +
    labs(x = "Total strength", y = "Entropy (bits)", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(),
      legend.title = element_text(size = rel(0.85)),
      legend.text = element_text(size = rel(0.8)),
      legend.key.size = unit(0.4, "cm"))
  print(p_str_ent_mod)
  save_fig6_plot(p_str_ent_mod, sprintf("panel_G_strength_vs_entropy_modality_%s", suffix), width = 7, height = 5, subdir = suffix)

  # ---- Hub table ----
  hub_table <- inter_summary %>%
    slice_head(n = 30) %>%
    left_join(inter_wide %>% select(intermediate, all_of(intersect(modality_order, colnames(inter_wide)))),
              by = "intermediate")
  hub_mba_info <- mba %>%
    filter(type %in% hub_table$intermediate) %>%
    group_by(type) %>%
    summarise(n_bodies = n(), consensus_nt = first(consensus_nt), .groups = "drop") %>%
    rename(intermediate = type)
  hub_table <- hub_table %>% left_join(hub_mba_info, by = "intermediate")
  cat("\n\n")
  print(knitr::kable(hub_table %>%
    select(intermediate, total_strength, entropy, source_diversity, target_diversity,
           total_paths, neuron_class, consensus_nt, n_bodies, everything()),
    digits = 2, caption = sprintf("Top 30 hub neurons (%s)", label)))
  cat("\n\n")

  # ---- Source vs target diversity ----
  p_div <- ggplot(inter_summary %>% filter(total_strength > 0),
    aes(x = source_diversity, y = target_diversity, size = total_strength, color = entropy)) +
    geom_point(alpha = 0.65, size = 2, shape = 16) +
    scale_size_continuous(name = "Strength", range = c(0.8, 5.5), trans = "log10", labels = scales::comma) +
    scale_color_viridis_c(option = "inferno", name = "Entropy") +
    ggrepel::geom_text_repel(data = inter_summary %>% filter(total_strength > 0) %>%
      mutate(div_score = source_diversity * target_diversity) %>% slice_max(div_score, n = 15),
      aes(label = intermediate), size = 2, max.overlaps = 20, color = "grey35",
      segment.color = "grey75", segment.size = 0.25) +
    labs(x = "Source diversity", y = "Target diversity", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(),
      legend.title = element_text(size = rel(0.85)),
      legend.text = element_text(size = rel(0.8)),
      legend.key.size = unit(0.4, "cm"))
  print(p_div)
  save_fig6_plot(p_div, sprintf("panel_G_source_vs_target_diversity_%s", suffix), width = 7, height = 5, subdir = suffix)

  # ---- Top 15 lollipop (hub integration score) ----
  # Hub score rewards multimodal integration & broad connectivity,
  # not raw path strength (which favours single-modality relays).
  inter_summary_scored <- inter_summary %>%
    filter(!grepl("^AN", intermediate)) %>%
    mutate(hub_score = entropy * sqrt(source_diversity * target_diversity)) %>%
    arrange(desc(hub_score))

  top15 <- inter_summary_scored %>%
    slice_head(n = 15) %>%
    mutate(intermediate = factor(intermediate, levels = rev(intermediate)))

  p_top15 <- ggplot(top15, aes(x = hub_score, y = intermediate)) +
    geom_segment(aes(x = 0, xend = hub_score, yend = intermediate),
                 color = "grey75", linewidth = 0.4) +
    geom_point(aes(size = total_strength, color = n_modalities), alpha = 0.85) +
    scale_size_continuous(name = "Total\nstrength", range = c(2, 8),
                          trans = "log10", labels = scales::comma) +
    scale_color_viridis_c(option = "inferno", name = "# modalities",
                          breaks = 1:7, limits = c(1, 7)) +
    labs(x = expression(Hub~score~~(entropy %*% sqrt(source~div. %*% target~div.))),
         y = NULL,
         title = sprintf("Top 15 integration hubs (%s)", label)) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.y = element_line(linewidth = 0.15, color = "grey92"),
          axis.text.y = element_text(size = 9, face = "bold"))
  print(p_top15)
  save_fig6_plot(p_top15, sprintf("panel_G_top15_hubs_%s", suffix), width = 9, height = 6, subdir = suffix)

  # ---- Modality co-innervation (signed) ----
  inter_mod_signed <- intermediates_clean %>%
    mutate(signed_strength = ifelse(valence == "inhibitory", -strength, strength)) %>%
    group_by(intermediate, modality) %>%
    summarise(net_strength = sum(signed_strength, na.rm = TRUE), .groups = "drop")

  signed_wide <- inter_mod_signed %>%
    pivot_wider(names_from = modality, values_from = net_strength, values_fill = 0)
  for (mod in modality_order) {
    if (!mod %in% colnames(signed_wide)) signed_wide[[mod]] <- 0
  }
  signed_mat <- signed_wide %>%
    column_to_rownames("intermediate") %>%
    select(all_of(intersect(modality_order, colnames(.)))) %>%
    as.matrix()
  rv <- apply(signed_mat, 1, var, na.rm = TRUE)
  signed_mat <- signed_mat[rv > 0, , drop = FALSE]

  mod_cor <- cor(signed_mat, method = "pearson")
  cor_pal <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)
  ga <- data.frame(group = modality_group[colnames(mod_cor)], row.names = colnames(mod_cor))

  p_cocor <- pheatmap(mod_cor, display_numbers = TRUE, number_format = "%.2f",
    fontsize_number = 10, color = cor_pal, breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete", annotation_col = ga, annotation_row = ga,
    annotation_colors = list(group = modality_group_colors), border_color = "white",
    cellwidth = 55, cellheight = 55,
    main = sprintf("Co-innervation (signed, %s)", label), fontsize = 11, silent = TRUE)

  co_name <- sprintf("panel_G_modality_coinnervation_%s", suffix)
  png(file.path(panel_plot_dir(co_name, "png", subdir = suffix), paste0(co_name, ".png")),
      width = 8, height = 7, units = "in", res = 300)
  grid::grid.draw(p_cocor$gtable); dev.off()
  pdf(file.path(panel_plot_dir(co_name, "pdf", subdir = suffix), paste0(co_name, ".pdf")),
      width = 8, height = 7)
  grid::grid.draw(p_cocor$gtable); dev.off()
  grid::grid.newpage(); grid::grid.draw(p_cocor$gtable)

  # ---- Unsigned co-routing ----
  abs_mat <- abs(signed_mat)
  cv <- apply(abs_mat, 2, var, na.rm = TRUE)
  abs_mat <- abs_mat[, cv > 0, drop = FALSE]
  mod_cor_abs <- cor(abs_mat, method = "pearson")

  p_coabs <- pheatmap(mod_cor_abs, display_numbers = TRUE, number_format = "%.2f",
    fontsize_number = 10,
    color = colorRampPalette(c("white", "#FDE725", "#35B779", "#31688E", "#440154"))(100),
    clustering_method = "complete", annotation_col = ga, annotation_row = ga,
    annotation_colors = list(group = modality_group_colors), border_color = "white",
    cellwidth = 55, cellheight = 55,
    main = sprintf("Co-routing (unsigned, %s)", label), fontsize = 11, silent = TRUE)

  cr_name <- sprintf("panel_G_modality_corouting_%s", suffix)
  png(file.path(panel_plot_dir(cr_name, "png", subdir = suffix), paste0(cr_name, ".png")),
      width = 8, height = 7, units = "in", res = 300)
  grid::grid.draw(p_coabs$gtable); dev.off()
  pdf(file.path(panel_plot_dir(cr_name, "pdf", subdir = suffix), paste0(cr_name, ".pdf")),
      width = 8, height = 7)
  grid::grid.draw(p_coabs$gtable); dev.off()
  grid::grid.newpage(); grid::grid.draw(p_coabs$gtable)

  cat(sprintf("\n\n**[%s] analysis complete.**\n\n", label))
}
```

## Session info

```{r session-info}
sessionInfo()
```
