---
title: "Figure 6E: Pairwise Modality Co-occurrence / Correlation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/fig6/setup.R")

# Additional libraries for this panel
library(ggraph)
library(tidygraph)
library(grid)
```

```{r save-and-show-pheatmap-helper}
# ---------------------------------------------------------------------------
# Helper: build pheatmap once, save PNG + PDF, display inline
# ---------------------------------------------------------------------------
save_and_show_pheatmap <- function(..., png_path, pdf_path, save_width = 9, save_height = 8) {
  p <- pheatmap(..., silent = TRUE)

  png(png_path, width = save_width, height = save_height, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(pdf_path, width = save_width, height = save_height)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(p)
}
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel E: Modality Co-occurrence / Correlation — Examines which sensory modality pairs tend to co-target the same downstream neurons. Pairwise Pearson and Spearman correlations between modality strength vectors are shown as heatmaps and a network graph. High correlation = shared targeting; negative correlation = mutual exclusivity."
panel_legend <- "(E) Pairwise modality co-occurrence analysis. Pearson and Spearman correlation heatmaps quantify the tendency of modality pairs to co-target the same neurons. Network graph visualizes significant positive correlations as edges, with edge width proportional to correlation strength."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_E_description", "txt"), "panel_E_description.txt"))
```

## Overview

This panel examines which sensory modality pairs tend to co-target the same
downstream neurons. We compute pairwise correlations between modality strength
vectors (each vector = strengths across all target neurons for one modality)
and visualise the results as heatmaps and a network graph.

High positive correlation between two modalities means that neurons receiving
strong input from one modality also tend to receive strong input from the other.
Negative correlation implies mutual exclusivity of targeting.

## 1. Pearson correlation heatmap

```{r pearson-correlation}
# Ensure modality order in the strength matrix
mat <- strength.aspf_aspg[, intersect(modality_order, colnames(strength.aspf_aspg))]

# Pearson correlation between modality columns
cor_pearson <- cor(mat, method = "pearson")

# Annotation: modality groups (olfactory / auditory / visual / contact)
group_annotation <- data.frame(
  group = modality_group[colnames(cor_pearson)],
  row.names = colnames(cor_pearson)
)

# Diverging colour palette
cor_colors <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)

save_and_show_pheatmap(
  cor_pearson,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = cor_colors,
  breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Pearson correlation (aSP-f + aSP-g targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_pearson_aspf_aspg", "png"), "panel_E_pearson_aspf_aspg.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_pearson_aspf_aspg", "pdf"), "panel_E_pearson_aspf_aspg.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 2. Spearman rank correlation heatmap

```{r spearman-correlation}
# Spearman: robust to outliers / non-linear monotonic relationships
cor_spearman <- cor(mat, method = "spearman")

save_and_show_pheatmap(
  cor_spearman,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = cor_colors,
  breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Spearman rank correlation (aSP-f + aSP-g targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_spearman_aspf_aspg", "png"), "panel_E_spearman_aspf_aspg.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_spearman_aspf_aspg", "pdf"), "panel_E_spearman_aspf_aspg.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 3. Co-targeting analysis

For each pair of modalities, count how many target neurons receive non-trivial
input from BOTH modalities. "Non-trivial" is defined as >5% of that modality's
maximum strength across all neurons.

```{r co-targeting-heatmap}
# Compute per-modality thresholds: 5% of each modality's max
mod_max <- apply(mat, 2, max, na.rm = TRUE)
threshold_frac <- 0.05
thresholds <- mod_max * threshold_frac

# Binary matrix: does neuron i receive non-trivial input from modality j?
mat_binary <- sweep(mat, 2, thresholds, ">") * 1

# Co-targeting count: for each pair (i,j), count neurons with both == 1
co_target_mat <- t(mat_binary) %*% mat_binary

# The diagonal = number of neurons with non-trivial input from that modality
cat("Neurons with non-trivial input per modality:\n")
print(diag(co_target_mat))

# Heatmap of co-targeting counts
co_target_colors <- colorRampPalette(c("#F7FCF5", "#00441B"))(50)

save_and_show_pheatmap(
  co_target_mat,
  display_numbers = TRUE,
  number_format = "%d",
  fontsize_number = 10,
  color = co_target_colors,
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = sprintf("Co-targeting counts (aSP-f + aSP-g, threshold: %.0f%%)", threshold_frac * 100),
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_counts_aspf_aspg", "png"), "panel_E_cotarget_counts_aspf_aspg.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_counts_aspf_aspg", "pdf"), "panel_E_cotarget_counts_aspf_aspg.pdf"),
  save_width = 9,
  save_height = 8
)
```

```{r co-targeting-jaccard}
# Jaccard index: co-target(i,j) / (count_i + count_j - co-target(i,j))
# More informative than raw counts since it accounts for marginal frequencies
n_mods <- ncol(mat_binary)
jaccard_mat <- matrix(0, nrow = n_mods, ncol = n_mods,
                      dimnames = list(colnames(mat_binary), colnames(mat_binary)))

for (i in seq_len(n_mods)) {
  for (j in seq_len(n_mods)) {
    both <- sum(mat_binary[, i] == 1 & mat_binary[, j] == 1)
    either <- sum(mat_binary[, i] == 1 | mat_binary[, j] == 1)
    jaccard_mat[i, j] <- ifelse(either > 0, both / either, 0)
  }
}

save_and_show_pheatmap(
  jaccard_mat,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = co_target_colors,
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Jaccard similarity (aSP-f + aSP-g targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_jaccard_aspf_aspg", "png"), "panel_E_cotarget_jaccard_aspf_aspg.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_jaccard_aspf_aspg", "pdf"), "panel_E_cotarget_jaccard_aspf_aspg.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 4. Network graph of modality correlations

Nodes represent modalities, edge thickness scales with correlation magnitude,
and edge color distinguishes positive (red) from negative (blue) correlations.

```{r correlation-network, fig.width=8, fig.height=8}
# Build edge list from the upper triangle of the Pearson correlation matrix
edge_df <- expand.grid(
  from = rownames(cor_pearson),
  to = colnames(cor_pearson),
  stringsAsFactors = FALSE
) %>%
  filter(from < to) %>%
  rowwise() %>%
  mutate(
    correlation = cor_pearson[from, to],
    abs_cor = abs(correlation),
    sign = ifelse(correlation >= 0, "positive", "negative")
  ) %>%
  ungroup()

# Node metadata
node_df <- data.frame(
  name = modality_order,
  group = modality_group[modality_order],
  stringsAsFactors = FALSE
)

# Create tidygraph object
graph <- tbl_graph(
  nodes = node_df,
  edges = edge_df,
  directed = FALSE
)

# Circular layout for clean publication figure
p_network <- ggraph(graph, layout = "circle") +
  geom_edge_link(
    aes(
      width = abs_cor,
      color = sign,
      alpha = abs_cor
    )
  ) +
  scale_edge_width_continuous(range = c(0.3, 4), name = "|Correlation|") +
  scale_edge_color_manual(
    values = c("positive" = "#B2182B", "negative" = "#2166AC"),
    name = "Direction"
  ) +
  scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
  geom_node_point(aes(color = group), size = 12) +
  geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
  scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
  labs(title = "Modality correlation network (aSP-f + aSP-g, Pearson)") +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

print(p_network)
save_fig6_plot(p_network, "panel_E_correlation_network_aspf_aspg", width = 9, height = 8)
```

```{r correlation-network-thresholded, fig.width=8, fig.height=8}
# Thresholded version: only show edges with |correlation| > 0.1
edge_df_thresh <- edge_df %>% filter(abs_cor > 0.1)

graph_thresh <- tbl_graph(
  nodes = node_df,
  edges = edge_df_thresh,
  directed = FALSE
)

p_network_thresh <- ggraph(graph_thresh, layout = "circle") +
  geom_edge_link(
    aes(
      width = abs_cor,
      color = sign
    ),
    alpha = 0.8
  ) +
  geom_edge_link(
    aes(label = sprintf("%.2f", correlation)),
    angle_calc = "along",
    label_size = 2.5,
    label_dodge = unit(3, "mm"),
    alpha = 0
  ) +
  scale_edge_width_continuous(range = c(0.5, 4), name = "|Correlation|") +
  scale_edge_color_manual(
    values = c("positive" = "#B2182B", "negative" = "#2166AC"),
    name = "Direction"
  ) +
  geom_node_point(aes(color = group), size = 14) +
  geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
  scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
  labs(title = "Modality correlation network (aSP-f + aSP-g, |r| > 0.1)") +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

print(p_network_thresh)
save_fig6_plot(p_network_thresh, "panel_E_correlation_network_aspf_aspg_thresholded", width = 9, height = 8)
```

## 5. Valence-split correlation analysis

Repeat the correlation heatmaps for excitatory-only and inhibitory-only paths
to determine whether co-occurrence patterns differ by valence.

```{r valence-split-correlations}
# Helper to compute and plot correlation for a given strength matrix and label
plot_valence_correlation <- function(strength_mat, valence_label) {
  # Ensure modality order
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  # Skip if too few columns or rows
  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping %s: insufficient data (%d rows, %d modalities)",
                    valence_label, nrow(mat_v), ncol(mat_v)))
    return(invisible(NULL))
  }

  # Remove rows with zero variance (no information)
  row_var <- apply(mat_v, 1, var, na.rm = TRUE)
  mat_v <- mat_v[row_var > 0, , drop = FALSE]

  # Remove modality columns with zero variance
  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping %s: fewer than 2 variable modalities after filtering", valence_label))
    return(invisible(NULL))
  }

  cor_mat <- cor(mat_v, method = "pearson")

  # Annotation for remaining modalities
  ga <- data.frame(
    group = modality_group[colnames(cor_mat)],
    row.names = colnames(cor_mat)
  )

  title_str <- sprintf("Pearson correlation (%s paths only)", valence_label)

  # Build pheatmap once, save gtable
  p <- pheatmap(
    cor_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  # Save PNG
  png(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", valence_label), "png"), sprintf("panel_E_pearson_%s.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Save PDF
  pdf(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", valence_label), "pdf"), sprintf("panel_E_pearson_%s.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Display inline
  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(cor_mat)
}
```

```{r valence-excitatory-heatmap, fig.width=9, fig.height=8}
cor_exc <- plot_valence_correlation(strength.aspf_aspg.exc, "excitatory_aspf_aspg")
```

```{r valence-inhibitory-heatmap, fig.width=9, fig.height=8}
cor_inh <- plot_valence_correlation(strength.aspf_aspg.inh, "inhibitory_aspf_aspg")
```

```{r valence-cotarget-comparison}
# Co-targeting heatmaps per valence
compute_cotarget <- function(strength_mat, valence_label) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  mod_max_v <- apply(mat_v, 2, max, na.rm = TRUE)
  thresh_v <- mod_max_v * threshold_frac
  # Guard against zero max (would make threshold 0, flagging everything)
  thresh_v[thresh_v == 0] <- .Machine$double.eps

  mat_bin_v <- sweep(mat_v, 2, thresh_v, ">") * 1
  co_mat <- t(mat_bin_v) %*% mat_bin_v

  ga <- data.frame(
    group = modality_group[colnames(co_mat)],
    row.names = colnames(co_mat)
  )

  title_str <- sprintf("Co-targeting counts (%s, threshold: %.0f%%)", valence_label, threshold_frac * 100)

  # Build pheatmap once, save gtable
  p <- pheatmap(
    co_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  # Save PNG
  png(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s", valence_label), "png"), sprintf("panel_E_cotarget_%s.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Save PDF
  pdf(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s", valence_label), "pdf"), sprintf("panel_E_cotarget_%s.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  # Display inline
  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(co_mat)
}
```

```{r cotarget-excitatory, fig.width=9, fig.height=8}
co_exc <- compute_cotarget(strength.aspf_aspg.exc, "excitatory_aspf_aspg")
```

```{r cotarget-inhibitory, fig.width=9, fig.height=8}
co_inh <- compute_cotarget(strength.aspf_aspg.inh, "inhibitory_aspf_aspg")
```

## 6. Summary statistics

```{r correlation-summary}
cat("=== Pearson Correlation Summary (combined) ===\n")
# Extract upper triangle values (excluding diagonal)
upper_idx <- upper.tri(cor_pearson)
cor_vals <- cor_pearson[upper_idx]
pair_labels <- outer(rownames(cor_pearson), colnames(cor_pearson), paste, sep = " - ")[upper_idx]

cor_summary <- data.frame(
  pair = pair_labels,
  pearson = cor_vals,
  spearman = cor_spearman[upper_idx]
) %>%
  arrange(desc(abs(pearson)))

cat(sprintf("Number of modality pairs: %d\n", length(cor_vals)))
cat(sprintf("Mean |Pearson r|: %.3f\n", mean(abs(cor_vals))))
cat(sprintf("Max positive: %.3f\n", max(cor_vals)))
cat(sprintf("Max negative: %.3f\n", min(cor_vals)))
cat(sprintf("Pairs with |r| > 0.3: %d\n", sum(abs(cor_vals) > 0.3)))

cat("\nAll pairwise correlations (sorted by |Pearson|):\n")
knitr::kable(cor_summary, digits = 3)
```

```{r co-targeting-summary}
cat("\n=== Co-targeting Summary ===\n")
cat(sprintf("Threshold: %.0f%% of modality max strength\n", threshold_frac * 100))
cat(sprintf("Total target neurons: %d\n", nrow(mat)))

# Diagonal = neurons per modality above threshold
cat("\nNeurons above threshold per modality:\n")
print(diag(co_target_mat))

# Off-diagonal co-targeting
co_upper <- co_target_mat[upper.tri(co_target_mat)]
co_labels <- outer(rownames(co_target_mat), colnames(co_target_mat), paste, sep = " & ")[upper.tri(co_target_mat)]

co_summary <- data.frame(
  pair = co_labels,
  shared_neurons = co_upper
) %>%
  arrange(desc(shared_neurons))

cat("\nTop co-targeting pairs:\n")
knitr::kable(co_summary)
```

---

## 7. vAB3 + PPN1 targets

Repeat all correlation and co-targeting analyses for targets reached via
vAB3 and PPN1 paths only.

```{r vab3-ppn1-setup}
# Compute strength matrices for vAB3 + PPN1 targets
paths.vab3_ppn1 <- bind_rows(paths.vab3, paths.ppn1)
strength.vab3_ppn1     <- compute_strength_matrix(paths.vab3_ppn1)
strength.vab3_ppn1.exc <- compute_valence_strength(paths.vab3_ppn1, "excitatory")
strength.vab3_ppn1.inh <- compute_valence_strength(paths.vab3_ppn1, "inhibitory")

mat_vp <- strength.vab3_ppn1[, intersect(modality_order, colnames(strength.vab3_ppn1))]
cat(sprintf("vAB3+PPN1 targets: %d neurons x %d modalities\n", nrow(mat_vp), ncol(mat_vp)))
```

```{r vab3-ppn1-pearson, fig.width=9, fig.height=8}
cor_pearson_vp <- cor(mat_vp, method = "pearson")

group_annotation_vp <- data.frame(
  group = modality_group[colnames(cor_pearson_vp)],
  row.names = colnames(cor_pearson_vp)
)

save_and_show_pheatmap(
  cor_pearson_vp,
  display_numbers = TRUE, number_format = "%.2f", fontsize_number = 9,
  color = cor_colors, breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation_vp, annotation_row = group_annotation_vp,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = "Pearson correlation (vAB3+PPN1 targets)", fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_pearson_vab3_ppn1", "png"), "panel_E_pearson_vab3_ppn1.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_pearson_vab3_ppn1", "pdf"), "panel_E_pearson_vab3_ppn1.pdf"),
  save_width = 9, save_height = 8
)
```

```{r vab3-ppn1-cotarget, fig.width=9, fig.height=8}
mod_max_vp <- apply(mat_vp, 2, max, na.rm = TRUE)
thresh_vp <- mod_max_vp * threshold_frac
thresh_vp[thresh_vp == 0] <- .Machine$double.eps
mat_bin_vp <- sweep(mat_vp, 2, thresh_vp, ">") * 1
co_target_vp <- t(mat_bin_vp) %*% mat_bin_vp

save_and_show_pheatmap(
  co_target_vp,
  display_numbers = TRUE, number_format = "%d", fontsize_number = 10,
  color = co_target_colors, cluster_cols = FALSE, clustering_method = "complete",
  annotation_col = group_annotation_vp, annotation_row = group_annotation_vp,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = sprintf("Co-targeting counts (vAB3+PPN1, threshold: %.0f%%)", threshold_frac * 100),
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_vab3_ppn1", "png"), "panel_E_cotarget_vab3_ppn1.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_vab3_ppn1", "pdf"), "panel_E_cotarget_vab3_ppn1.pdf"),
  save_width = 9, save_height = 8
)
```

```{r vab3-ppn1-jaccard, fig.width=9, fig.height=8}
n_mods_vp <- ncol(mat_bin_vp)
jaccard_vp <- matrix(0, nrow = n_mods_vp, ncol = n_mods_vp,
                     dimnames = list(colnames(mat_bin_vp), colnames(mat_bin_vp)))
for (i in seq_len(n_mods_vp)) {
  for (j in seq_len(n_mods_vp)) {
    both <- sum(mat_bin_vp[, i] == 1 & mat_bin_vp[, j] == 1)
    either <- sum(mat_bin_vp[, i] == 1 | mat_bin_vp[, j] == 1)
    jaccard_vp[i, j] <- ifelse(either > 0, both / either, 0)
  }
}

save_and_show_pheatmap(
  jaccard_vp,
  display_numbers = TRUE, number_format = "%.2f", fontsize_number = 9,
  color = co_target_colors, cluster_cols = FALSE, clustering_method = "complete",
  annotation_col = group_annotation_vp, annotation_row = group_annotation_vp,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = "Jaccard similarity (vAB3+PPN1 targets)", fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_jaccard_vab3_ppn1", "png"), "panel_E_jaccard_vab3_ppn1.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_jaccard_vab3_ppn1", "pdf"), "panel_E_jaccard_vab3_ppn1.pdf"),
  save_width = 9, save_height = 8
)
```

```{r vab3-ppn1-valence-exc, fig.width=9, fig.height=8}
plot_valence_correlation(strength.vab3_ppn1.exc, "excitatory_vab3_ppn1")
```

```{r vab3-ppn1-valence-inh, fig.width=9, fig.height=8}
plot_valence_correlation(strength.vab3_ppn1.inh, "inhibitory_vab3_ppn1")
```

```{r vab3-ppn1-cotarget-exc, fig.width=9, fig.height=8}
compute_cotarget(strength.vab3_ppn1.exc, "excitatory_vab3_ppn1")
```

```{r vab3-ppn1-cotarget-inh, fig.width=9, fig.height=8}
compute_cotarget(strength.vab3_ppn1.inh, "inhibitory_vab3_ppn1")
```

---

## 8. All targets (aSP-f + aSP-g + vAB3 + PPN1)

Repeat all analyses using the full set of targets from all four path sets.

```{r all-targets-setup}
mat_all <- strength.all[, intersect(modality_order, colnames(strength.all))]
cat(sprintf("All targets: %d neurons x %d modalities\n", nrow(mat_all), ncol(mat_all)))
```

```{r all-targets-pearson, fig.width=9, fig.height=8}
cor_pearson_all <- cor(mat_all, method = "pearson")

group_annotation_all <- data.frame(
  group = modality_group[colnames(cor_pearson_all)],
  row.names = colnames(cor_pearson_all)
)

save_and_show_pheatmap(
  cor_pearson_all,
  display_numbers = TRUE, number_format = "%.2f", fontsize_number = 9,
  color = cor_colors, breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation_all, annotation_row = group_annotation_all,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = "Pearson correlation (all targets)", fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_pearson_all_targets", "png"), "panel_E_pearson_all_targets.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_pearson_all_targets", "pdf"), "panel_E_pearson_all_targets.pdf"),
  save_width = 9, save_height = 8
)
```

```{r all-targets-spearman, fig.width=9, fig.height=8}
cor_spearman_all <- cor(mat_all, method = "spearman")

save_and_show_pheatmap(
  cor_spearman_all,
  display_numbers = TRUE, number_format = "%.2f", fontsize_number = 9,
  color = cor_colors, breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation_all, annotation_row = group_annotation_all,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = "Spearman rank correlation (all targets)", fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_spearman_all_targets", "png"), "panel_E_spearman_all_targets.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_spearman_all_targets", "pdf"), "panel_E_spearman_all_targets.pdf"),
  save_width = 9, save_height = 8
)
```

```{r all-targets-cotarget, fig.width=9, fig.height=8}
mod_max_all <- apply(mat_all, 2, max, na.rm = TRUE)
thresh_all <- mod_max_all * threshold_frac
thresh_all[thresh_all == 0] <- .Machine$double.eps
mat_bin_all <- sweep(mat_all, 2, thresh_all, ">") * 1
co_target_all <- t(mat_bin_all) %*% mat_bin_all

save_and_show_pheatmap(
  co_target_all,
  display_numbers = TRUE, number_format = "%d", fontsize_number = 10,
  color = co_target_colors, cluster_cols = FALSE, clustering_method = "complete",
  annotation_col = group_annotation_all, annotation_row = group_annotation_all,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = sprintf("Co-targeting counts (all targets, threshold: %.0f%%)", threshold_frac * 100),
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_all_targets", "png"), "panel_E_cotarget_all_targets.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_all_targets", "pdf"), "panel_E_cotarget_all_targets.pdf"),
  save_width = 9, save_height = 8
)
```

```{r all-targets-jaccard, fig.width=9, fig.height=8}
n_mods_all <- ncol(mat_bin_all)
jaccard_all <- matrix(0, nrow = n_mods_all, ncol = n_mods_all,
                      dimnames = list(colnames(mat_bin_all), colnames(mat_bin_all)))
for (i in seq_len(n_mods_all)) {
  for (j in seq_len(n_mods_all)) {
    both <- sum(mat_bin_all[, i] == 1 & mat_bin_all[, j] == 1)
    either <- sum(mat_bin_all[, i] == 1 | mat_bin_all[, j] == 1)
    jaccard_all[i, j] <- ifelse(either > 0, both / either, 0)
  }
}

save_and_show_pheatmap(
  jaccard_all,
  display_numbers = TRUE, number_format = "%.2f", fontsize_number = 9,
  color = co_target_colors, cluster_cols = FALSE, clustering_method = "complete",
  annotation_col = group_annotation_all, annotation_row = group_annotation_all,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white", cellwidth = 50, cellheight = 50,
  main = "Jaccard similarity (all targets)", fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_jaccard_all_targets", "png"), "panel_E_jaccard_all_targets.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_jaccard_all_targets", "pdf"), "panel_E_jaccard_all_targets.pdf"),
  save_width = 9, save_height = 8
)
```

```{r all-targets-network, fig.width=8, fig.height=8}
edge_df_all <- expand.grid(
  from = rownames(cor_pearson_all),
  to = colnames(cor_pearson_all),
  stringsAsFactors = FALSE
) %>%
  filter(from < to) %>%
  rowwise() %>%
  mutate(
    correlation = cor_pearson_all[from, to],
    abs_cor = abs(correlation),
    sign = ifelse(correlation >= 0, "positive", "negative")
  ) %>%
  ungroup() %>%
  filter(abs_cor > 0.1)

graph_all <- tbl_graph(
  nodes = node_df,
  edges = edge_df_all,
  directed = FALSE
)

p_network_all <- ggraph(graph_all, layout = "circle") +
  geom_edge_link(
    aes(width = abs_cor, color = sign),
    alpha = 0.8
  ) +
  geom_edge_link(
    aes(label = sprintf("%.2f", correlation)),
    angle_calc = "along", label_size = 2.5, label_dodge = unit(3, "mm"), alpha = 0
  ) +
  scale_edge_width_continuous(range = c(0.5, 4), name = "|Correlation|") +
  scale_edge_color_manual(
    values = c("positive" = "#B2182B", "negative" = "#2166AC"), name = "Direction"
  ) +
  geom_node_point(aes(color = group), size = 14) +
  geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
  scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
  labs(title = "Modality correlation network (all targets, |r| > 0.1)") +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

print(p_network_all)
save_fig6_plot(p_network_all, "panel_E_correlation_network_all_targets", width = 9, height = 8)
```

```{r all-targets-valence-exc, fig.width=9, fig.height=8}
plot_valence_correlation(strength.all.exc, "excitatory_all_targets")
```

```{r all-targets-valence-inh, fig.width=9, fig.height=8}
plot_valence_correlation(strength.all.inh, "inhibitory_all_targets")
```

```{r all-targets-cotarget-exc, fig.width=9, fig.height=8}
compute_cotarget(strength.all.exc, "excitatory_all_targets")
```

```{r all-targets-cotarget-inh, fig.width=9, fig.height=8}
compute_cotarget(strength.all.inh, "inhibitory_all_targets")
```

```{r session-info}
sessionInfo()
```
