---
title: "Figure 6F: Path Diversity -- Concentrated vs. Distributed Routes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel F: Path Diversity — Measures how concentrated or distributed the strongest paths are per modality. For each (target, modality) pair, computes the number of paths needed to accumulate 80%% of total strength. Dot plots and bar charts compare path diversity across modalities and between excitatory/inhibitory valence classes."
panel_legend <- "(F) Path diversity per modality. The number of paths required to accumulate 80%% of total strength measures how concentrated or distributed signal routing is. Dot plots show per-neuron values with median overlay; bar charts summarize means with standard errors. Faceted views split by excitatory and inhibitory valence."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_F_description", "txt"), "panel_F_description.txt"))
```

## Overview

Panel F examines how concentrated or distributed the strongest paths are for
each sensory modality reaching each target neuron. Some modalities may funnel
signal through a few dominant paths, while others spread it across many weaker
routes. We quantify this by computing the number of paths required to
accumulate 80% of total strength per (target, modality) pair, then visualise
the distribution across modalities and valence classes.

## 1. Define target sets

```{r define-path-sets}
path_sets <- list(
  aspf_aspg   = list(paths = paths.aspf_aspg,
                     label = "aSP-f/aSP-g", suffix = "aspf_aspg"),
  vab3_ppn1   = list(paths = bind_rows(paths.vab3, paths.ppn1),
                     label = "vAB3+PPN1",   suffix = "vab3_ppn1"),
  all_targets = list(paths = paths.all,
                     label = "All targets",  suffix = "all_targets")
)
```

## 2. Path diversity analysis for each target set

```{r run-all-sets, fig.width=10, fig.height=7, results='asis'}
for (set_name in names(path_sets)) {

  ps       <- path_sets[[set_name]]
  paths_df <- ps$paths
  label    <- ps$label
  suffix   <- ps$suffix

  cat(sprintf("\n\n---\n\n### Target set: %s\n\n", label))

  # ==== Cumulative strength and paths-to-80% ====
  cumulative_df <- paths_df %>%
    filter(!is.na(strength), strength > 0) %>%
    group_by(end, modality) %>%
    arrange(desc(strength), .by_group = TRUE) %>%
    mutate(
      rank           = row_number(),
      total_strength = sum(strength),
      cum_strength   = cumsum(strength),
      cum_frac       = cum_strength / total_strength
    ) %>%
    ungroup()

  paths_to_80 <- cumulative_df %>%
    group_by(end, modality) %>%
    summarise(
      n_paths        = n(),
      total_strength = first(total_strength),
      paths_to_80pct = min(rank[cum_frac >= 0.80], na.rm = TRUE),
      .groups        = "drop"
    ) %>%
    mutate(
      paths_to_80pct = ifelse(is.infinite(paths_to_80pct), n_paths, paths_to_80pct),
      modality       = factor(modality, levels = modality_order)
    )

  # ==== Summary statistics ====
  summary_80 <- paths_to_80 %>%
    group_by(modality) %>%
    summarise(
      n_targets  = n(),
      mean_p80   = mean(paths_to_80pct),
      median_p80 = median(paths_to_80pct),
      sd_p80     = sd(paths_to_80pct),
      se_p80     = sd(paths_to_80pct) / sqrt(n()),
      .groups    = "drop"
    )
  cat("\n\n")
  print(knitr::kable(summary_80, digits = 2,
        caption = sprintf("Summary: paths to 80%% (%s)", label)))
  cat("\n\n")

  # ==== Dot plot ====
  p_dot <- ggplot(paths_to_80, aes(x = modality, y = paths_to_80pct)) +
    geom_jitter(aes(color = modality), width = 0.2, alpha = 0.4, size = 1.2, stroke = 0) +
    stat_summary(fun = median, geom = "point", aes(color = modality), size = 3, shape = 18) +
    scale_color_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Paths needed for 80% of total strength", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(), legend.position = "none",
      axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_dot)
  save_fig6_plot(p_dot, sprintf("panel_F_dot_paths_to_80pct_%s", suffix), width = 4.5, height = 5)

  # ==== Cumulative curves ====
  targets_multi <- paths_to_80 %>%
    group_by(end) %>%
    summarise(n_modalities = n_distinct(modality), .groups = "drop") %>%
    filter(n_modalities >= 4) %>%
    pull(end)
  if (length(targets_multi) > 12) {
    top_targets <- cumulative_df %>%
      filter(end %in% targets_multi) %>%
      group_by(end) %>% summarise(total = sum(strength), .groups = "drop") %>%
      slice_max(total, n = 12) %>% pull(end)
  } else {
    top_targets <- targets_multi
  }
  if (length(top_targets) < 4) {
    top_targets <- cumulative_df %>%
      group_by(end) %>% summarise(total = sum(strength), .groups = "drop") %>%
      slice_max(total, n = 9) %>% pull(end)
  }

  cum_subset <- cumulative_df %>%
    filter(end %in% top_targets) %>%
    mutate(modality = factor(modality, levels = modality_order))

  p_cum <- ggplot(cum_subset, aes(x = rank, y = cum_frac, color = modality)) +
    geom_line(linewidth = 0.8, alpha = 0.85) +
    geom_hline(yintercept = 0.80, linetype = "dashed", color = "grey40", linewidth = 0.4) +
    facet_wrap(~end, scales = "free_x", ncol = 4) +
    scale_color_manual(values = modality_colors) +
    scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1.02)) +
    labs(x = "Path rank (by descending strength)", y = "Cumulative fraction of total strength",
         color = "Modality", title = sprintf("Cumulative path strength (%s)", label)) +
    theme_bw(base_size = 11) +
    theme(strip.text = element_text(size = 8, face = "bold"),
          legend.position = "bottom", panel.grid.minor = element_blank())
  print(p_cum)
  save_fig6_plot(p_cum, sprintf("panel_F_cumulative_curves_%s", suffix), width = 12, height = 8)

  # ==== Bar chart ====
  p_bar <- ggplot(summary_80, aes(x = modality, y = mean_p80, fill = modality)) +
    geom_col(width = 0.65, color = "grey30", linewidth = 0.3) +
    geom_errorbar(aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
                  width = 0.2, linewidth = 0.4, color = "grey20") +
    scale_fill_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Mean paths to 80% strength",
         title = sprintf("Path diversity (%s)", label)) +
    theme_bw(base_size = 13) +
    theme(legend.position = "none", panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_bar)
  save_fig6_plot(p_bar, sprintf("panel_F_bar_paths_to_80pct_%s", suffix), width = 4.5, height = 5)

  # ==== Valence split ====
  cumulative_valence <- paths_df %>%
    filter(!is.na(strength), strength > 0, !is.na(valence)) %>%
    group_by(end, modality, valence) %>%
    arrange(desc(strength), .by_group = TRUE) %>%
    mutate(rank = row_number(), total_strength = sum(strength),
           cum_strength = cumsum(strength), cum_frac = cum_strength / total_strength) %>%
    ungroup()

  paths_to_80_valence <- cumulative_valence %>%
    group_by(end, modality, valence) %>%
    summarise(n_paths = n(), total_strength = first(total_strength),
              paths_to_80pct = min(rank[cum_frac >= 0.80], na.rm = TRUE), .groups = "drop") %>%
    mutate(paths_to_80pct = ifelse(is.infinite(paths_to_80pct), n_paths, paths_to_80pct),
           modality = factor(modality, levels = modality_order))

  summary_valence <- paths_to_80_valence %>%
    group_by(modality, valence) %>%
    summarise(n_targets = n(), mean_p80 = mean(paths_to_80pct),
              median_p80 = median(paths_to_80pct),
              se_p80 = sd(paths_to_80pct) / sqrt(n()), .groups = "drop")

  p_valence <- ggplot(summary_valence,
    aes(x = modality, y = mean_p80, fill = modality, alpha = valence)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6,
             color = "grey30", linewidth = 0.3) +
    geom_jitter(data = paths_to_80_valence,
                aes(x = modality, y = paths_to_80pct, color = modality, alpha = valence),
                position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7),
                size = 0.8, stroke = 0, inherit.aes = FALSE) +
    geom_errorbar(aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
                  position = position_dodge(width = 0.7), width = 0.2,
                  linewidth = 0.4, color = "grey20") +
    scale_fill_manual(values = modality_colors) +
    scale_color_manual(values = modality_colors, guide = "none") +
    scale_alpha_manual(values = c("excitatory" = 0.7, "inhibitory" = 0.35), name = "Valence") +
    labs(x = "Sensory modality", y = "Mean paths to 80% strength",
         title = sprintf("Path diversity by valence (%s)", label)) +
    guides(fill = "none") +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "top")
  print(p_valence)
  save_fig6_plot(p_valence, sprintf("panel_F_valence_paths_to_80pct_%s", suffix), width = 6, height = 5)

  # Valence dot plot
  p_val_dot <- ggplot(paths_to_80_valence, aes(x = modality, y = paths_to_80pct)) +
    geom_jitter(aes(color = modality), width = 0.2, alpha = 0.4, size = 1.2, stroke = 0) +
    stat_summary(fun = median, geom = "point", aes(color = modality), size = 3, shape = 18) +
    facet_wrap(~valence) +
    scale_color_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Paths needed for 80% of total strength", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(), legend.position = "none",
      axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_val_dot)
  save_fig6_plot(p_val_dot, sprintf("panel_F_valence_dot_paths_to_80pct_%s", suffix), width = 5.5, height = 4.5)

  # ==== Start-type-scaled versions ====
  start_type_counts <- paths_df %>%
    filter(!is.na(strength), strength > 0) %>%
    group_by(modality) %>%
    summarise(n_start_types = n_distinct(start), .groups = "drop") %>%
    mutate(modality = factor(modality, levels = modality_order))

  paths_to_80_scaled <- paths_to_80 %>%
    left_join(start_type_counts, by = "modality") %>%
    mutate(paths_to_80pct_scaled = paths_to_80pct / n_start_types)

  summary_80_scaled <- paths_to_80_scaled %>%
    group_by(modality) %>%
    summarise(n_targets = n(), n_start_types = first(n_start_types),
              mean_p80 = mean(paths_to_80pct_scaled), median_p80 = median(paths_to_80pct_scaled),
              sd_p80 = sd(paths_to_80pct_scaled),
              se_p80 = sd(paths_to_80pct_scaled) / sqrt(n()), .groups = "drop")

  # Scaled dot plot
  p_dot_scaled <- ggplot(paths_to_80_scaled, aes(x = modality, y = paths_to_80pct_scaled)) +
    geom_jitter(aes(color = modality), width = 0.2, alpha = 0.4, size = 1.2, stroke = 0) +
    stat_summary(fun = median, geom = "point", aes(color = modality), size = 3, shape = 18) +
    scale_color_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Paths to 80% / # start types", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(), legend.position = "none",
      axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_dot_scaled)
  save_fig6_plot(p_dot_scaled, sprintf("panel_F_dot_paths_to_80pct_scaled_%s", suffix),
                 width = 4.5, height = 5, subdir = "scaled_by_start_types")

  # Scaled bar chart
  p_bar_scaled <- ggplot(summary_80_scaled, aes(x = modality, y = mean_p80, fill = modality)) +
    geom_col(width = 0.65, color = "grey30", linewidth = 0.3) +
    geom_errorbar(aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
                  width = 0.2, linewidth = 0.4, color = "grey20") +
    scale_fill_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Mean paths to 80% / # start types",
         title = sprintf("Path diversity scaled (%s)", label)) +
    theme_bw(base_size = 13) +
    theme(legend.position = "none", panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_bar_scaled)
  save_fig6_plot(p_bar_scaled, sprintf("panel_F_bar_paths_to_80pct_scaled_%s", suffix),
                 width = 4.5, height = 5, subdir = "scaled_by_start_types")

  # Scaled valence
  paths_to_80_valence_scaled <- paths_to_80_valence %>%
    left_join(start_type_counts, by = "modality") %>%
    mutate(paths_to_80pct_scaled = paths_to_80pct / n_start_types)

  summary_valence_scaled <- paths_to_80_valence_scaled %>%
    group_by(modality, valence) %>%
    summarise(n_targets = n(), mean_p80 = mean(paths_to_80pct_scaled),
              median_p80 = median(paths_to_80pct_scaled),
              se_p80 = sd(paths_to_80pct_scaled) / sqrt(n()), .groups = "drop")

  p_valence_scaled <- ggplot(summary_valence_scaled,
    aes(x = modality, y = mean_p80, fill = modality, alpha = valence)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6,
             color = "grey30", linewidth = 0.3) +
    geom_jitter(data = paths_to_80_valence_scaled,
                aes(x = modality, y = paths_to_80pct_scaled, color = modality, alpha = valence),
                position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7),
                size = 0.8, stroke = 0, inherit.aes = FALSE) +
    geom_errorbar(aes(ymin = mean_p80 - se_p80, ymax = mean_p80 + se_p80),
                  position = position_dodge(width = 0.7), width = 0.2,
                  linewidth = 0.4, color = "grey20") +
    scale_fill_manual(values = modality_colors) +
    scale_color_manual(values = modality_colors, guide = "none") +
    scale_alpha_manual(values = c("excitatory" = 0.7, "inhibitory" = 0.35), name = "Valence") +
    labs(x = "Sensory modality", y = "Mean paths to 80% / # start types",
         title = sprintf("Path diversity by valence, scaled (%s)", label)) +
    guides(fill = "none") +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "top")
  print(p_valence_scaled)
  save_fig6_plot(p_valence_scaled, sprintf("panel_F_valence_paths_to_80pct_scaled_%s", suffix),
                 width = 6, height = 5, subdir = "scaled_by_start_types")

  p_val_dot_scaled <- ggplot(paths_to_80_valence_scaled,
    aes(x = modality, y = paths_to_80pct_scaled)) +
    geom_jitter(aes(color = modality), width = 0.2, alpha = 0.4, size = 1.2, stroke = 0) +
    stat_summary(fun = median, geom = "point", aes(color = modality), size = 3, shape = 18) +
    facet_wrap(~valence) +
    scale_color_manual(values = modality_colors) +
    labs(x = "Sensory modality", y = "Paths to 80% / # start types", title = "") +
    theme_minimal(base_size = 11) +
    theme(text = element_text(color = "grey20"),
      axis.title = element_text(size = rel(0.9), color = "grey30"),
      axis.text = element_text(size = rel(0.85), color = "grey40"),
      axis.line = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid = element_blank(), legend.position = "none",
      axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_val_dot_scaled)
  save_fig6_plot(p_val_dot_scaled, sprintf("panel_F_valence_dot_paths_to_80pct_scaled_%s", suffix),
                 width = 5.5, height = 4.5, subdir = "scaled_by_start_types")

  cat(sprintf("\n\n**[%s] Panel F analysis complete.**\n\n", label))
}
```

## 3. Cross-target-set comparison: all target sets × valence

```{r cross-target-comparison, fig.width=14, fig.height=6}
# -- Define the 5 individual target sets --
comparison_sets <- list(
  "aSP-f"     = paths.aspf,
  "aSP-g"     = paths.aspg,
  "vAB3"      = paths.vab3,
  "PPN1"      = paths.ppn1,
  "vAB3+PPN1" = bind_rows(paths.vab3, paths.ppn1)
)

target_set_order <- names(comparison_sets)

target_set_colors_5 <- c(
  "aSP-f"     = "#E41A1C",
  "aSP-g"     = "#FF7F00",
  "vAB3"      = "#377EB8",
  "PPN1"      = "#4DAF4A",
  "vAB3+PPN1" = "#984EA3"
)

# -- Compute paths-to-80% by valence for each target set --
all_valence_raw    <- list()
all_valence_scaled <- list()

for (ts_name in names(comparison_sets)) {
  pdf <- comparison_sets[[ts_name]] %>%
    filter(!is.na(strength), strength > 0, !is.na(valence))

  cum_val <- pdf %>%
    group_by(end, modality, valence) %>%
    arrange(desc(strength), .by_group = TRUE) %>%
    mutate(rank = row_number(), total_strength = sum(strength),
           cum_strength = cumsum(strength),
           cum_frac = cum_strength / total_strength) %>%
    ungroup()

  p80_val <- cum_val %>%
    group_by(end, modality, valence) %>%
    summarise(n_paths = n(), total_strength = first(total_strength),
              paths_to_80pct = min(rank[cum_frac >= 0.80], na.rm = TRUE),
              .groups = "drop") %>%
    mutate(paths_to_80pct = ifelse(is.infinite(paths_to_80pct), n_paths, paths_to_80pct),
           modality   = factor(modality, levels = modality_order),
           target_set = ts_name)

  # Start-type counts for this target set
  stc <- pdf %>%
    group_by(modality) %>%
    summarise(n_start_types = n_distinct(start), .groups = "drop") %>%
    mutate(modality = factor(modality, levels = modality_order))

  p80_val_s <- p80_val %>%
    left_join(stc, by = "modality") %>%
    mutate(paths_to_80pct_scaled = paths_to_80pct / n_start_types)

  all_valence_raw[[ts_name]]    <- p80_val
  all_valence_scaled[[ts_name]] <- p80_val_s
}

combined_val_raw <- bind_rows(all_valence_raw) %>%
  mutate(target_set = factor(target_set, levels = target_set_order))

combined_val_scaled <- bind_rows(all_valence_scaled) %>%
  mutate(target_set = factor(target_set, levels = target_set_order))

# -- Summaries (mean ± SE) --
summary_cross_raw <- combined_val_raw %>%
  group_by(modality, valence, target_set) %>%
  summarise(n = n(), mean_p80 = mean(paths_to_80pct),
            se_p80 = sd(paths_to_80pct) / sqrt(n()), .groups = "drop")

summary_cross_scaled <- combined_val_scaled %>%
  group_by(modality, valence, target_set) %>%
  summarise(n = n(), mean_p80 = mean(paths_to_80pct_scaled),
            se_p80 = sd(paths_to_80pct_scaled) / sqrt(n()), .groups = "drop")

# -- Plot: raw paths-to-80%, 10 bars per modality --
p_cross_raw <- ggplot(summary_cross_raw,
    aes(x = modality, y = mean_p80, fill = target_set, alpha = valence)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.78,
           color = "grey30", linewidth = 0.15) +
  geom_jitter(data = combined_val_raw,
              aes(x = modality, y = paths_to_80pct, fill = target_set, alpha = valence),
              position = position_jitterdodge(jitter.width = 0.08, dodge.width = 0.85),
              size = 0.5, stroke = 0, shape = 21, color = "grey30", inherit.aes = FALSE) +
  geom_errorbar(aes(ymin = pmax(mean_p80 - se_p80, 0), ymax = mean_p80 + se_p80),
                position = position_dodge(width = 0.85), width = 0.12,
                linewidth = 0.3, color = "grey20") +
  scale_fill_manual(values = target_set_colors_5, name = "Target set") +
  scale_alpha_manual(values = c("excitatory" = 1, "inhibitory" = 0.45), name = "Valence") +
  labs(x = "Sensory modality", y = "Mean paths to 80% strength") +
  theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "top", legend.box = "horizontal") +
  guides(fill = guide_legend(order = 1, nrow = 1),
         alpha = guide_legend(order = 2, override.aes = list(fill = "grey50")))
print(p_cross_raw)
save_fig6_plot(p_cross_raw, "panel_F_cross_target_valence_raw",
               width = 14, height = 6)

# -- Plot: scaled by # start types, 10 bars per modality --
p_cross_scaled <- ggplot(summary_cross_scaled,
    aes(x = modality, y = mean_p80, fill = target_set, alpha = valence)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.78,
           color = "grey30", linewidth = 0.15) +
  geom_jitter(data = combined_val_scaled,
              aes(x = modality, y = paths_to_80pct_scaled, fill = target_set, alpha = valence),
              position = position_jitterdodge(jitter.width = 0.08, dodge.width = 0.85),
              size = 0.5, stroke = 0, shape = 21, color = "grey30", inherit.aes = FALSE) +
  geom_errorbar(aes(ymin = pmax(mean_p80 - se_p80, 0), ymax = mean_p80 + se_p80),
                position = position_dodge(width = 0.85), width = 0.12,
                linewidth = 0.3, color = "grey20") +
  scale_fill_manual(values = target_set_colors_5, name = "Target set") +
  scale_alpha_manual(values = c("excitatory" = 1, "inhibitory" = 0.45), name = "Valence") +
  labs(x = "Sensory modality", y = "Mean paths to 80% / # start types") +
  theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "top", legend.box = "horizontal") +
  guides(fill = guide_legend(order = 1, nrow = 1),
         alpha = guide_legend(order = 2, override.aes = list(fill = "grey50")))
print(p_cross_scaled)
save_fig6_plot(p_cross_scaled, "panel_F_cross_target_valence_scaled",
               width = 14, height = 6, subdir = "scaled_by_start_types")
```

## Session info

```{r session-info}
sessionInfo()
```
