---
title: "Figure 6H: Spotlight Examples — Best Integrator per Target Set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel H: Spotlight Integration Nodes — For each of 5 target sets (aSP-f, aSP-g, vAB3, PPN1, vAB3+PPN1), identifies the best multimodal integrator (highest entropy, >= 3 modalities). Radar plots show modality input profiles; valence radars overlay excitatory and inhibitory contributions; stacked bars show absolute and proportional modality breakdown."
panel_legend <- "(H) Best multimodal integrator per target set. For each target set, the neuron with the highest Shannon entropy across modalities is spotlighted. Radar plots show normalized modality profiles; excitatory (solid blue) and inhibitory (dashed red) contributions are overlaid; stacked bars display absolute and proportional modality composition."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_H_description", "txt"), "panel_H_description.txt"))
```

## Overview

Panel H spotlights the **best multimodal integrator** from each of 5 target
sets. For each target set we compute a strength matrix (targets × modalities),
calculate Shannon entropy per target neuron, and pick the neuron with the
highest entropy (requiring input from at least 3 modalities). Each spotlight
neuron is shown with radar plots and stacked bar charts.

## Step 1: Pick best integrator per target set

```{r pick-best-integrators}
# ---------------------------------------------------------------------------
# Define 5 target sets
# ---------------------------------------------------------------------------
target_sets <- list(
  "aSP-f"     = paths.aspf,
  "aSP-g"     = paths.aspg,
  "vAB3"      = paths.vab3,
  "PPN1"      = paths.ppn1,
  "vAB3+PPN1" = bind_rows(paths.vab3, paths.ppn1)
)

target_set_colors_h <- c(
  "aSP-f"     = "#E41A1C",
  "aSP-g"     = "#FF7F00",
  "vAB3"      = "#377EB8",
  "PPN1"      = "#4DAF4A",
  "vAB3+PPN1" = "#984EA3"
)

# ---------------------------------------------------------------------------
# Forced neuron picks per target set (NULL = auto-pick by entropy)
# A target set can have multiple forced neurons.
# ---------------------------------------------------------------------------
forced_picks <- list(
  "aSP-f"     = NULL,
  "aSP-g"     = NULL,
  "vAB3"      = "AVLP743m",
  "PPN1"      = "DNp103",
  "vAB3+PPN1" = c("LH004m", "AVLP597")
)

# ---------------------------------------------------------------------------
# For each target set: compute strength, entropy, assign spotlight neurons
# ---------------------------------------------------------------------------
# spotlight_data is a flat list keyed by "target_set::neuron"
spotlight_data  <- list()
spotlight_order <- character()

for (ts_name in names(target_sets)) {
  ts_paths <- target_sets[[ts_name]]

  str_total <- compute_strength_matrix(ts_paths)
  str_exc   <- compute_valence_strength(ts_paths, "excitatory")
  str_inh   <- compute_valence_strength(ts_paths, "inhibitory")

  ent_vals <- apply(str_total, 1, shannon_entropy)
  n_mods   <- apply(str_total > 0, 1, sum)

  edf <- data.frame(
    neuron         = names(ent_vals),
    entropy        = ent_vals,
    n_modalities   = n_mods,
    total_strength = rowSums(str_total),
    stringsAsFactors = FALSE
  ) %>% arrange(desc(entropy))

  # Determine neurons for this target set
  picks <- forced_picks[[ts_name]]
  if (is.null(picks)) {
    best <- edf %>% filter(n_modalities >= 3) %>% slice_head(n = 1)
    if (nrow(best) == 0) best <- edf %>% filter(n_modalities >= 2) %>% slice_head(n = 1)
    if (nrow(best) == 0) best <- edf %>% slice_head(n = 1)
    picks <- best$neuron
  }

  for (nn in picks) {
    nn_row <- edf %>% filter(neuron == nn)
    nn_ent <- if (nrow(nn_row) > 0) nn_row$entropy else 0
    nn_mod <- if (nrow(nn_row) > 0) nn_row$n_modalities else 0

    key <- paste0(ts_name, "::", nn)
    spotlight_data[[key]] <- list(
      neuron     = nn,
      target_set = ts_name,
      entropy    = nn_ent,
      n_mod      = nn_mod,
      str_total  = str_total,
      str_exc    = str_exc,
      str_inh    = str_inh,
      entropy_df = edf
    )
    spotlight_order <- c(spotlight_order, key)

    message(sprintf("  [%s] Spotlight: %s (H=%.2f, %d modalities)",
                    ts_name, nn, nn_ent, nn_mod))
  }
}

spotlight_neurons <- sapply(spotlight_data[spotlight_order], function(x) x$neuron)
n_spotlights <- length(spotlight_order)
```

## Step 2: Prepare per-neuron long-format data

```{r prepare-neuron-data}
# ---------------------------------------------------------------------------
# Build long-format data: for each spotlight neuron, use its target-set-specific
# strength matrices so the radar shows what reaches that neuron in context.
# ---------------------------------------------------------------------------
spotlight_long <- bind_rows(lapply(spotlight_order, function(key) {
  sd <- spotlight_data[[key]]
  nn <- sd$neuron
  ts_name <- sd$target_set

  # Total
  total_row <- if (nn %in% rownames(sd$str_total)) {
    sd$str_total[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  # Excitatory
  exc_row <- if (nn %in% rownames(sd$str_exc)) {
    sd$str_exc[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  # Inhibitory
  inh_row <- if (nn %in% rownames(sd$str_inh)) {
    sd$str_inh[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  bind_rows(
    data.frame(neuron = nn, modality = modality_order,
               strength = as.numeric(total_row), valence = "total",
               target_set = ts_name, stringsAsFactors = FALSE),
    data.frame(neuron = nn, modality = modality_order,
               strength = as.numeric(exc_row), valence = "excitatory",
               target_set = ts_name, stringsAsFactors = FALSE),
    data.frame(neuron = nn, modality = modality_order,
               strength = as.numeric(inh_row), valence = "inhibitory",
               target_set = ts_name, stringsAsFactors = FALSE)
  )
})) %>%
  mutate(modality = factor(modality, levels = modality_order),
         target_set = factor(target_set, levels = names(target_sets)))

# Construct a combined entropy_df for lookup
entropy_df <- bind_rows(lapply(spotlight_data, function(sd) {
  sd$entropy_df %>% mutate(target_set = sd$target_set)
}))
```

## Step 3: Individual radar plots (total strength)

```{r radar-individual, fig.width=21, fig.height=4}
# ---------------------------------------------------------------------------
# Radar plot function: uses per-neuron data from spotlight_data
# ---------------------------------------------------------------------------
make_neuron_radar <- function(key) {
  sd <- spotlight_data[[key]]
  nn <- sd$neuron
  ts <- sd$target_set

  vals <- if (nn %in% rownames(sd$str_total)) {
    sd$str_total[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  max_val <- max(vals, na.rm = TRUE)
  if (max_val == 0) max_val <- 1
  vals_norm <- vals / max_val

  radar_df <- data.frame(
    modality = factor(modality_order, levels = modality_order),
    value    = as.numeric(vals_norm),
    raw      = as.numeric(vals)
  )

  ent_label <- sprintf("H = %.2f  |  %d mod.", sd$entropy, sd$n_mod)

  ggplot(radar_df, aes(x = modality, y = value, group = 1)) +
    geom_polygon(fill = alpha(target_set_colors_h[[ts]], 0.3),
                 color = target_set_colors_h[[ts]], linewidth = 0.8) +
    geom_point(aes(color = modality), size = 3, show.legend = FALSE) +
    scale_color_manual(values = modality_colors) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05), breaks = c(0.25, 0.5, 0.75, 1.0),
                       labels = c("25%", "50%", "75%", "100%")) +
    labs(title = nn, subtitle = paste0(ts, " target  |  ", ent_label)) +
    theme_minimal(base_size = 11) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_text(size = 7, color = "grey50"),
      axis.text.x       = element_text(size = 9, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 12),
      plot.subtitle      = element_text(hjust = 0.5, size = 8, color = "grey40")
    )
}

radar_plots <- lapply(spotlight_order, make_neuron_radar)
names(radar_plots) <- spotlight_order

radar_combined <- plot_grid(
  plotlist = radar_plots, nrow = 1,
  labels = LETTERS[seq_along(radar_plots)], label_size = 12
)
radar_combined
```

```{r save-radar-individual}
save_fig6_plot(radar_combined, "panel_H_radar_individual",
               width = 3.5 * n_spotlights, height = 4)
```

## Step 4: Excitatory vs. inhibitory overlaid radars

```{r radar-valence, fig.width=21, fig.height=4.5}
# ---------------------------------------------------------------------------
# Dual-valence radar per spotlight neuron
# ---------------------------------------------------------------------------
make_valence_radar <- function(key) {
  sd <- spotlight_data[[key]]
  nn <- sd$neuron
  ts <- sd$target_set

  exc_vals <- if (nn %in% rownames(sd$str_exc)) {
    sd$str_exc[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  inh_vals <- if (nn %in% rownames(sd$str_inh)) {
    sd$str_inh[nn, modality_order]
  } else { setNames(rep(0, 7), modality_order) }

  global_max <- max(c(exc_vals, inh_vals), na.rm = TRUE)
  if (global_max == 0) global_max <- 1

  radar_df <- data.frame(
    modality = factor(rep(modality_order, 2), levels = modality_order),
    value    = c(as.numeric(exc_vals) / global_max,
                 as.numeric(inh_vals) / global_max),
    valence  = rep(c("excitatory", "inhibitory"), each = 7)
  )

  ggplot(radar_df, aes(x = modality, y = value, group = valence)) +
    geom_polygon(aes(fill = valence, linetype = valence, color = valence),
                 alpha = 0.2, linewidth = 0.9) +
    geom_point(aes(color = valence), size = 2) +
    scale_fill_manual(values = c("excitatory" = "#2166AC", "inhibitory" = "#B2182B")) +
    scale_color_manual(values = c("excitatory" = "#2166AC", "inhibitory" = "#B2182B")) +
    scale_linetype_manual(values = c("excitatory" = "solid", "inhibitory" = "dashed")) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05)) +
    labs(title = nn, subtitle = paste0(ts, " target")) +
    theme_minimal(base_size = 10) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_blank(),
      axis.text.x       = element_text(size = 8, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 11),
      plot.subtitle      = element_text(hjust = 0.5, size = 8, color = "grey40"),
      legend.position    = "bottom",
      legend.title       = element_blank(),
      legend.text        = element_text(size = 8)
    )
}

valence_radars <- lapply(spotlight_order, make_valence_radar)
names(valence_radars) <- spotlight_order

valence_combined <- plot_grid(
  plotlist = valence_radars, nrow = 1,
  labels = LETTERS[seq_along(valence_radars)], label_size = 12
)
valence_combined
```

```{r save-radar-valence}
save_fig6_plot(valence_combined, "panel_H_radar_valence",
               width = 3.5 * n_spotlights, height = 5)
```

## Step 5: Input breakdown stacked bar (excitatory/inhibitory by modality)

```{r stacked-bar, fig.width=21, fig.height=5}
# ---------------------------------------------------------------------------
# Stacked bar: faceted by neuron, coloured by target set border
# ---------------------------------------------------------------------------
# Build ordered facet labels matching spotlight_order
facet_levels <- sapply(spotlight_order, function(key) {
  sd <- spotlight_data[[key]]
  sprintf("%s\n(%s target)", sd$neuron, sd$target_set)
})

bar_data <- spotlight_long %>%
  filter(valence %in% c("excitatory", "inhibitory")) %>%
  group_by(neuron, target_set) %>%
  mutate(
    total_all  = sum(strength, na.rm = TRUE),
    proportion = ifelse(total_all > 0, strength / total_all, 0)
  ) %>%
  ungroup() %>%
  mutate(
    facet_label = sprintf("%s\n(%s target)", neuron, target_set),
    facet_label = factor(facet_label, levels = facet_levels)
  )

p_stacked <- ggplot(bar_data, aes(x = modality, y = strength, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ facet_label, scales = "free_y", nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC", "inhibitory" = "#B2182B"),
                     name = "Valence") +
  labs(x = "Modality", y = "Summed path strength",
       title = "Input breakdown by modality and valence") +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 9),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )
p_stacked
```

```{r save-stacked-bar}
save_fig6_plot(p_stacked, "panel_H_stacked_bar",
               width = 3.5 * n_spotlights, height = 5)
```

## Step 5b: Proportional stacked bar (normalized to 100%)

```{r stacked-bar-proportional, fig.width=21, fig.height=5}
p_stacked_prop <- ggplot(bar_data, aes(x = modality, y = proportion, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ facet_label, nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC", "inhibitory" = "#B2182B"),
                     name = "Valence") +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Modality", y = "Proportion of total input",
       title = "Proportional input breakdown") +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 9),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )
p_stacked_prop
```

```{r save-stacked-prop}
save_fig6_plot(p_stacked_prop, "panel_H_stacked_bar_proportional",
               width = 3.5 * n_spotlights, height = 5)
```

## Step 6: Annotation summary table

```{r annotation-table}
# ---------------------------------------------------------------------------
# Summary table for the 5 spotlight neurons
# ---------------------------------------------------------------------------
annotation_table <- bind_rows(lapply(spotlight_order, function(key) {
  sd <- spotlight_data[[key]]
  data.frame(
    neuron         = sd$neuron,
    target_set     = sd$target_set,
    entropy        = sd$entropy,
    n_modalities   = sd$n_mod,
    total_strength = sum(sd$str_total[sd$neuron, ]),
    stringsAsFactors = FALSE
  )
})) %>%
  left_join(
    mba %>% select(type, synonyms) %>% distinct(type, .keep_all = TRUE),
    by = c("neuron" = "type")
  ) %>%
  mutate(
    synonyms = ifelse(is.na(synonyms) | synonyms == "", "\u2014", synonyms),
    synonyms = ifelse(nchar(synonyms) > 60, paste0(substr(synonyms, 1, 57), "..."), synonyms)
  )

knitr::kable(annotation_table, digits = 3,
             caption = "Best integrator per target set",
             col.names = c("Neuron", "Target set", "Entropy",
                           "N modalities", "Total strength", "Synonyms"))
```

## Step 7: Combined final panel

```{r combined-panel, fig.width=21, fig.height=9}
panel_H <- plot_grid(
  radar_combined,
  valence_combined,
  p_stacked,
  ncol        = 1,
  rel_heights = c(1, 1, 1),
  labels      = c("i", "ii", "iii"),
  label_size  = 14
)
panel_H
```

```{r save-combined}
save_fig6_plot(panel_H, "panel_H_combined", width = 3.5 * n_spotlights, height = 12)
```
