---
title: "rewiring"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 6
)

# Shared dependencies for strongest-path analyses and visualisations
library(arrow)
library(coconatfly)
library(cowplot)
library(dplyr)
library(fafbseg)
library(ggraph)
library(ggplot2)
library(ggpubr)
library(grid)
library(igraph)
library(jsonlite)
library(malecns)
library(Matrix)
library(neuprintr)
library(nat.ggplot)
library(patchwork)
library(plotly)
library(purrr)
library(RCy3)
library(reshape2)
library(scales)
library(stringr)
library(tidygraph)
library(tidyr)
```

## Overview

This notebook assembles precomputed strongest-path summaries across male and female pheromone circuits, enriches them with morphological annotations, and visualises modality-specific connectivity patterns. Each section builds on the previous one: we begin by loading cached data, derive adjacency matrices and helper utilities, and then profile path strength statistics for different sensory inputs.


### 1 Define function that loads a file or alternatively if it doesnt exist exexutes a script
```{r}
load_or_execute <- function(path) {
  if (file.exists(path)) {
    file.exists(path)
    FALSE
  } else {
    TRUE
  }
}
```


### 1.1 Load male body annotations
```{r}
choose_mcns()
synapse_threshold <- 5
if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/mba.feather')) {
  mba <- mcns_body_annotations()
  mba %>% mutate(type = ifelse(type == "", NA, type))
  write_feather(mba, '/Users/fkampf/Documents/pheromone.paper/feather/mba.feather')

} else {
  mba <- read_feather('/Users/fkampf/Documents/pheromone.paper/feather/mba.feather')
}
```






### 1.2 Function to load connectivity
```{r}
fetch_connectivity <- function(synapse_threshold = 5) {
  mba <- mcns_body_annotations() %>%
    mutate(type = coalesce(type, flywire_type, manc_type, hemibrain_type)) %>%
    mutate(cachero.type = str_extract(synonyms, "(?<=Cachero 2010:)[^;]+") %>%
      str_trim())
  connectivity <- cf_partners(cf_ids(malecns = mba %>% pull(bodyid)),
    partners = "o",
    threshold = synapse_threshold
  )
  connectivity <- connectivity %>%
    left_join(
      mba %>%
        select("type", "bodyid", "fru_dsx", "consensus_nt", "flywire_type", "synonyms", "receptor_type", "cachero.type", "dimorphism") %>%
        mutate(bodyid = as.integer(bodyid)) %>%
        rename(pre_type = type, pre_fru_dsx = fru_dsx, pre_nt = consensus_nt, pre_fw_type = flywire_type, pre_synonyms = synonyms, pre_receptor_type = receptor_type, pre.dimorphism = dimorphism, pre.cachero.type = cachero.type),
      by = c("pre_id" = "bodyid")
    ) %>%
    rename(post_type = type) %>%
    left_join(
      mba %>%
        select("bodyid", "fru_dsx", "consensus_nt", "flywire_type", "synonyms", "receptor_type", "cachero.type", "dimorphism") %>%
        mutate(bodyid = as.integer(bodyid)) %>%
        rename(post_fru_dsx = fru_dsx, post_nt = consensus_nt, post_fw_type = flywire_type, post_synonyms = synonyms, post_receptor_type = receptor_type, post.dimorphism = dimorphism, post.cachero.type = cachero.type),
      by = c("post_id" = "bodyid")
    )
  return(connectivity)
}
```




### 1.3 Load connectivity
```{r}
if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/connectivity.feather')) {
  conn <- fetch_connectivity()
  write_feather(conn, '/Users/fkampf/Documents/pheromone.paper/feather/connectivity.feather')

} else {
  conn <- read_feather('/Users/fkampf/Documents/pheromone.paper/feather/connectivity.feather')
}

conn_save <- conn

conn <- conn %>%
  mutate(
    pre_type = case_when(
      grepl("putative_ppk23", pre_receptor_type) ~ "m-cell",
      grepl("putative_ppk25", pre_receptor_type) ~ "f-cell",
      TRUE ~ pre_type
    ),
    post_type = case_when(
      grepl("putative_ppk23", post_receptor_type) ~ "m-cell",
      grepl("putative_ppk25", post_receptor_type) ~ "f-cell",
      TRUE ~ post_type
    )
  )
```

## 2 Adjacency matrix

With the enriched connectivity, build a column/row-normalised adjacency matrix that becomes the weighted igraph backbone for all path searches.

### 2.1 Function adjacency matrix
```{r}
calculate_normed_adj_matrix <- function(connectivity, cell.or.type = "type", pre.or.post = "pre") {
  colScale <- function(A, na.rm = TRUE) {
    scalefac <- 1 / Matrix::colSums(A)
    if (na.rm) scalefac[!is.finite(scalefac)] <- 0
    B <- A %*% Matrix::Diagonal(x = scalefac)
    B
  }
  rowScale <- function(A, na.rm = TRUE) {
    scalefac <- 1 / Matrix::rowSums(A)
    if (na.rm) scalefac[!is.finite(scalefac)] <- 0
    B <- Matrix::Diagonal(x = scalefac) %*% A
    B
  }
  if (cell.or.type == "cell") {
    unique.identifier <- union(as.character(connectivity$pre_id), as.character(connectivity$post_id))
    adj.matrix <- sparseMatrix(
      i = match(connectivity$pre_id, unique.identifier),
      j = match(connectivity$post_id, unique.identifier),
      x = connectivity$weight,
      dims = c(length(unique.identifier), length(unique.identifier)),
      dimnames = list(unique.identifier, unique.identifier)
    )
  } else {
    unique.identifier <- union(connectivity$pre_type, connectivity$post_type)
    adj.matrix <- sparseMatrix(
      i = match(connectivity$pre_type, unique.identifier),
      j = match(connectivity$post_type, unique.identifier),
      x = connectivity$weight,
      dims = c(length(unique.identifier), length(unique.identifier)),
      dimnames = list(unique.identifier, unique.identifier)
    )
  }
  if (pre.or.post == "pre") {
    adj.matrix.normed.pre <- rowScale(adj.matrix)
    colnames(adj.matrix.normed.pre) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.pre) <- rownames(adj.matrix)
    return(adj.matrix.normed.pre)
  } else if (pre.or.post == "avg") {
    adj.matrix.normed.post <- colScale(adj.matrix)
    colnames(adj.matrix.normed.post) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.post) <- rownames(adj.matrix)
    adj.matrix.normed.pre <- rowScale(adj.matrix)
    colnames(adj.matrix.normed.pre) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.pre) <- rownames(adj.matrix)

    return(avg_mat <- (adj.matrix.normed.post + adj.matrix.normed.post) / 2)
  } else {
    adj.matrix.normed.post <- colScale(adj.matrix)
    colnames(adj.matrix.normed.post) <- colnames(adj.matrix)
    rownames(adj.matrix.normed.post) <- rownames(adj.matrix)
    return(adj.matrix.normed.post)
  }
}
```

### 2.2 Create adjacency matrix
```{r}
use.clone <- FALSE
conn <- conn_save
conn <- conn %>%
  mutate(pre_type = coalesce(pre_type, pre_fw_type), post_type = coalesce(post_type, post_fw_type)) %>%
  mutate(
    pre_type = if_else(is.na(pre_type), as.character(bodyid), pre_type),
    post_type = if_else(is.na(post_type), as.character(partner), post_type)
  ) %>%
  mutate(pre_type = gsub("[()]", "", pre_type)) %>%
  mutate(post_type = gsub("[()]", "", post_type))


if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/adj.matrix.feather')) {
  adj.matrix <- calculate_normed_adj_matrix(conn, cell.or.type = "type", pre.or.post = "post")
  saveRDS(adj.matrix, '/Users/fkampf/Documents/pheromone.paper/feather/adj.matrix.rds')

} else {
  adj.matrix <- readRDS('/Users/fkampf/Documents/pheromone.paper/feather/adj.matrix.rds')
}

if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/graph.general.rds')) {
  graph.general <- graph_from_adjacency_matrix(
    adj.matrix,
    mode = "directed",
    weighted = TRUE,
    diag = FALSE
  )
  saveRDS(graph.general, '/Users/fkampf/Documents/pheromone.paper/feather/graph.general.rds')

} else {
  graph.general <- readRDS('/Users/fkampf/Documents/pheromone.paper/feather/graph.general.rds')
}


```



## 2.2 Define kstrongest path function
```{r}
add_valence_flag <- function(df, nt_by_type, is_inhib_by_nt) {
  df %>%
    rowwise() %>%
    mutate(
      # split path and drop endpoint
      nodes = list(setdiff(strsplit(path, " -> ", fixed = TRUE)[[1]], end)),
      # lookup NTs
      nts = list(unname(nt_by_type[unlist(nodes)])),
      # lookup inhibitory flags
      inhib = list(unname(is_inhib_by_nt[unlist(nts)])),
      # count inhibitors
      n_inhib = sum(unlist(inhib), na.rm = TRUE),
      # convert to +1 / -1
      d = ifelse(n_inhib %% 2 == 0, 1, -1)
    ) %>%
    ungroup()
}


find_k_strongest_paths <- function(
  g,
  starts,
  targets,
  mba,
  n_paths = 3,
  diversity = c("otheredges", "othernodes", "none"),
  ignore_pattern = NULL # grepl() pattern; matched node names are excluded (start/target kept),
) {
  type2nt <- mba %>%
    select(type, consensus_nt) %>%
    group_by(type) %>%
    distinct() %>%
    rename(nt = consensus_nt)
  nt_by_type <- setNames(type2nt$nt, type2nt$type)
  inhibitory_nts <- c("gaba", "glycine", "histamine")
  all_nts <- unique(type2nt$nt)
  is_inhib_by_nt <- setNames(all_nts %in% inhibitory_nts, all_nts)

  diversity <- match.arg(diversity)

  # optional: drop nodes matching ignore_pattern (but never drop starts/targets)
  if (!is.null(ignore_pattern)) {
    drop <- V(g)$name[grepl(ignore_pattern, V(g)$name)]
    keep <- unique(c(starts, targets))
    drop <- setdiff(drop, keep)
    if (length(drop)) g <- igraph::delete_vertices(g, drop)
  }

  # weights → log-costs; guard zeros/NAs
  w <- E(g)$weight
  if (is.null(w)) stop("Edges must have numeric 'weight'.")
  w[!is.finite(w) | w <= 0] <- .Machine$double.eps
  log_w <- -log(w)

  results <- list()
  all_nodes <- character(0)

  for (s in starts) {
    for (t in targets) {
      if (!(s %in% V(g)$name) || !(t %in% V(g)$name)) {
        results[[paste(s, t, sep = "->")]] <- list(list(
          start = s, end = t, path = NA_character_, strength = NA_real_, hops = NA_integer_
        ))
        next
      }

      g_tmp <- g
      E(g_tmp)$log_weight <- log_w
      pair_paths <- list()

      for (i in seq_len(n_paths)) {
        sp <- igraph::shortest_paths(
          g_tmp,
          from = s, to = t,
          weights = E(g_tmp)$log_weight,
          output = "both"
        )
        if (length(sp$vpath[[1]]) == 0) break

        vpath <- sp$vpath[[1]]
        epath <- sp$epath[[1]]

        final_strength <- exp(-sum(E(g_tmp)[epath]$log_weight))
        hops <- length(epath)

        pair_paths[[paste0("path", i)]] <- list(
          start = s,
          end = t,
          path = paste(V(g)[vpath]$name, collapse = " -> "),
          strength = final_strength,
          hops = hops
        )

        all_nodes <- c(all_nodes, V(g)[vpath]$name)

        if (diversity == "othernodes") {
          inter <- setdiff(V(g)[vpath]$name, c(s, t))
          if (length(inter)) g_tmp <- igraph::delete_vertices(g_tmp, inter)
        } else if (diversity == "otheredges") {
          g_tmp <- igraph::delete_edges(g_tmp, epath)
        }
      }

      if (length(pair_paths) == 0) {
        pair_paths <- list(list(
          start = s, end = t, path = NA_character_, strength = NA_real_, hops = NA_integer_
        ))
      }

      results[[paste(s, t, sep = "->")]] <- pair_paths
    }
  }

  # flatten
  results_df <- do.call(rbind, lapply(results, function(pl) {
    do.call(rbind, lapply(pl, function(r) {
      data.frame(
        start = r$start, end = r$end, path = r$path,
        strength = r$strength, hops = r$hops,
        stringsAsFactors = FALSE
      )
    }))
  }))

  results_df_wo_na <- dplyr::filter(results_df, !is.na(strength))
  all_nodes <- unique(all_nodes)

  results_df_wo_na <- add_valence_flag(results_df_wo_na, nt_by_type, is_inhib_by_nt) %>%
    mutate(valence = ifelse(d == 1L, "excitatory", "inhibitory"))

  list(df = results_df_wo_na, nodes = all_nodes, raw = results)
}
```
## 3 Downstream target selection

Derive frequently targeted downstream neuron types for both vAB3 and PPN1 by summarising partner synapse weights. We drop broad projection categories (AN/IN) and keep targets contributing at least 0.1% of outgoing weight so subsequent plots focus on substantial partners.

```{r downstream-target-selection}

all.PPN1.ids <- mba %>%
  filter(grepl("PPN1", synonyms)) %>%
  pull(bodyid)
all.vAB3.ids <- mba %>%
  filter(grepl("vAB3", synonyms)) %>%
  pull(bodyid)

if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/target.vAB3.feather')) {
  target.vAB3 <- cf_partner_summary(cf_ids(malecns = all.vAB3.ids), partners = "out", normalise = F) %>%
    group_by(type.post) %>%
    summarize(weight = sum(weight)) %>%
    filter(!grepl("AN", type.post)) %>%
    filter(!grepl("IN", type.post)) %>%
    arrange(desc(weight)) %>%
    filter(weight / sum(weight) >= 0.001) %>%
    filter(!is.na(type.post)) %>%
    pull(type.post)
  saveRDS(target.vAB3, '/Users/fkampf/Documents/pheromone.paper/feather/target.vAB3.rds')

} else {
  target.vAB3 <- readRDS('/Users/fkampf/Documents/pheromone.paper/feather/target.vAB3.rds')
}

if (load_or_execute('/Users/fkampf/Documents/pheromone.paper/feather/target.PPN1.feather')) {
  target.PPN1 <- cf_partner_summary(cf_ids(malecns = all.PPN1.ids), partners = "out", normalise = F) %>%
    group_by(type.post) %>%
    summarize(weight = sum(weight)) %>%
    filter(!grepl("AN", type.post)) %>%
    filter(!grepl("IN", type.post)) %>%
    arrange(desc(weight)) %>%
    filter(weight / sum(weight) >= 0.001) %>%
    filter(!is.na(type.post)) %>%
    pull(type.post)
  saveRDS(target.vAB3, '/Users/fkampf/Documents/pheromone.paper/feather/target.PPN1.rds')

} else {
  target.PPN1 <- readRDS('/Users/fkampf/Documents/pheromone.paper/feather/target.PPN1.rds')
}


target.vAB3.all <- c(target.vAB3[1:10], intersect(target.vAB3, target.PPN1)[1:3])
target.PPN1.all <- c(target.PPN1[1:10], intersect(target.vAB3, target.PPN1)[1:3])

```


```{r}
library(igraph)

rewire_allow_loops_no_multiedges <- function(
  g,
  niter = ecount(g) * 50,
  report_every = 0.0001,
  seed = NULL
) {
  if (!is.null(seed)) set.seed(seed)


  el <- as_edgelist(g, names = FALSE)
  m  <- nrow(el)
  vnames <- V(g)$name

  # hash table of existing edges
  present <- new.env(hash = TRUE, parent = emptyenv())
  for (k in paste(el[,1], el[,2], sep=":")) present[[k]] <- TRUE

  step <- max(1L, as.integer(niter * report_every))

  for (t in seq_len(niter)) {
    i1 <- sample.int(m, 1L)
    i2 <- sample.int(m, 1L)
    if (i1 == i2) next

    a <- el[i1,1]; b <- el[i1,2]
    c <- el[i2,1]; d <- el[i2,2]

    # proposed new edges (loops allowed)
    k1_old <- paste(a,b,sep=":")
    k2_old <- paste(c,d,sep=":")
    k1_new <- paste(a,d,sep=":")
    k2_new <- paste(c,b,sep=":")

    # forbid creating duplicate A->B
    if (!is.null(present[[k1_new]])) next
    if (!is.null(present[[k2_new]])) next

    # update keys
    rm(list = c(k1_old, k2_old), envir = present)
    present[[k1_new]] <- TRUE
    present[[k2_new]] <- TRUE

    # perform swap
    el[i1,2] <- d
    el[i2,2] <- b

    # progress
    if (t %% step == 0L) {
      cat(sprintf("\r%.3f%%", 100 * t / niter))
      flush.console()
    }
  }
  cat("\r100%\n")

  h <- graph_from_edgelist(el, directed = TRUE)

  # restore vertex names
  if (!is.null(vnames)) V(h)$name <- vnames

  # degree sanity


  h
}

```




## Degree-preserving swap that forbids multiedges but allows loops
```{r Degree-preserving swap that forbids loops (keeps multi-edges)}
g_rw <- rewire_allow_loops_no_multiedges(graph.general, niter = ecount(graph.general)*50)

#nulls_swap <- replicate(2, rewire_multigraph_no_loops(graph.general), simplify = FALSE)

```


