---
title: "Figure 6E (Bella's targets): Pairwise Modality Co-occurrence / Correlation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")

# Additional libraries for this panel
library(ggraph)
library(tidygraph)
library(grid)
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel E: Modality Co-occurrence / Correlation (Bella's targets) — Pairwise Pearson and Spearman correlations between modality strength vectors shown as heatmaps and network graphs. Identifies which modality pairs co-target the same downstream neurons."
panel_legend <- "(E) Pairwise modality co-occurrence among Bella's targets. Pearson and Spearman correlation heatmaps and network graphs reveal which modality pairs tend to co-target the same downstream neurons."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_E_description", "txt"), "panel_E_description.txt"))
```

```{r save-and-show-pheatmap-helper, include=FALSE}
# ---------------------------------------------------------------------------
# Helper: build pheatmap once, save PNG + PDF, display inline
# ---------------------------------------------------------------------------
save_and_show_pheatmap <- function(..., png_path, pdf_path, save_width = 9, save_height = 8) {
  p <- pheatmap(..., silent = TRUE)

  png(png_path, width = save_width, height = save_height, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(pdf_path, width = save_width, height = save_height)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(p)
}
```

## Overview

This panel examines which sensory modality pairs tend to co-target the same
downstream neurons among Bella's targets (M cells, F cells, IR cells). We compute
pairwise correlations between modality strength vectors (each vector = strengths
across all target neurons for one modality) and visualise the results as heatmaps
and a network graph.

High positive correlation between two modalities means that neurons receiving
strong input from one modality also tend to receive strong input from the other.
Negative correlation implies mutual exclusivity of targeting.

```{r color-palettes-and-threshold, include=FALSE}
# ---------------------------------------------------------------------------
# Colour palettes and co-targeting threshold (defined once, used everywhere)
# ---------------------------------------------------------------------------
cor_colors      <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)
spearman_colors <- colorRampPalette(c("#7B3294", "#F7F7F7", "#E66101"))(100)
cosine_colors   <- colorRampPalette(c("#01665E", "#F5F5F5", "#8C510A"))(100)
co_target_colors <- colorRampPalette(c("#F7FCF5", "#00441B"))(50)
jaccard_colors   <- colorRampPalette(c("#F2F0F7", "#54278F"))(50)

threshold_frac <- 0.05
```

```{r compute-net-strength-helper, include=FALSE}
# ---------------------------------------------------------------------------
# Compute net strength matrix: excitatory - inhibitory
# Aligns rows/columns from both matrices; missing entries treated as 0
# ---------------------------------------------------------------------------
compute_net_strength <- function(strength_exc, strength_inh) {
  all_mods <- union(colnames(strength_exc), colnames(strength_inh))
  all_neurons <- union(rownames(strength_exc), rownames(strength_inh))

  mat_exc <- matrix(0, nrow = length(all_neurons), ncol = length(all_mods),
                    dimnames = list(all_neurons, all_mods))
  mat_inh <- matrix(0, nrow = length(all_neurons), ncol = length(all_mods),
                    dimnames = list(all_neurons, all_mods))

  mat_exc[rownames(strength_exc), colnames(strength_exc)] <- strength_exc
  mat_inh[rownames(strength_inh), colnames(strength_inh)] <- strength_inh

  mat_exc - mat_inh
}
```

```{r run-correlation-suite-helper, include=FALSE}
# ---------------------------------------------------------------------------
# run_correlation_suite(): full Pearson + Spearman + co-targeting + Jaccard +
#   network analysis for a given strength matrix.
# ---------------------------------------------------------------------------
run_correlation_suite <- function(mat, label, subdir,
                                  title_suffix = "(Bella's targets)",
                                  threshold_frac = 0.05) {

  # --- Modality-group annotation ---
  group_annotation <- data.frame(
    group = modality_group[colnames(mat)],
    row.names = colnames(mat)
  )

  # ---- 1. Pearson correlation heatmap ------------------------------------
  cor_pearson <- cor(mat, method = "pearson")

  save_and_show_pheatmap(
    cor_pearson,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Pearson correlation of modality strengths %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_pearson_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_pearson_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_pearson_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 2. Spearman rank correlation heatmap ------------------------------
  cor_spearman <- cor(mat, method = "spearman")

  save_and_show_pheatmap(
    cor_spearman,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = spearman_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Spearman rank correlation of modality strengths %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_spearman_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_spearman_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_spearman_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_spearman_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 2b. Cosine similarity heatmap -------------------------------------
  # Unlike Pearson, cosine does NOT mean-center: sensitive to baseline level
  col_norms <- sqrt(colSums(mat^2))
  col_norms[col_norms == 0] <- 1
  mat_normed <- sweep(mat, 2, col_norms, "/")
  cosine_mat <- t(mat_normed) %*% mat_normed

  save_and_show_pheatmap(
    cosine_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cosine_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Cosine similarity of modality strengths %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cosine_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cosine_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 3. Co-targeting counts --------------------------------------------
  mod_max <- apply(mat, 2, max, na.rm = TRUE)
  thresholds <- mod_max * threshold_frac
  thresholds[thresholds == 0] <- .Machine$double.eps

  mat_binary <- sweep(mat, 2, thresholds, ">") * 1
  co_target_mat <- t(mat_binary) %*% mat_binary

  cat(sprintf("Neurons with non-trivial input per modality %s:\n", title_suffix))
  print(diag(co_target_mat))

  save_and_show_pheatmap(
    co_target_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting counts (threshold: %.0f%% of modality max) %s",
                   threshold_frac * 100, title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 3b. Co-targeting counts (threshold: % of total strength) ----
  mod_sum <- colSums(abs(mat))
  thresholds_total <- mod_sum * threshold_frac
  thresholds_total[thresholds_total == 0] <- .Machine$double.eps

  mat_binary_total <- sweep(mat, 2, thresholds_total, ">") * 1
  co_target_mat_total <- t(mat_binary_total) %*% mat_binary_total

  cat(sprintf("Neurons above threshold (%%total) per modality %s:\n", title_suffix))
  print(diag(co_target_mat_total))

  save_and_show_pheatmap(
    co_target_mat_total,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting counts (threshold: %.0f%% of total) %s",
                   threshold_frac * 100, title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_counts_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_counts_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 3c. Column-normalized version of total-based co-targeting ----
  diag_counts_total <- diag(co_target_mat_total)
  diag_counts_total[diag_counts_total == 0] <- 1
  co_target_colnorm_total <- sweep(co_target_mat_total, 2, diag_counts_total, "/")

  save_and_show_pheatmap(
    co_target_colnorm_total,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting col-norm (threshold: %.0f%% of total) %s",
                   threshold_frac * 100, title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 3d. Jaccard similarity (total-based threshold) ----
  n_mods_total <- ncol(mat_binary_total)
  jaccard_mat_total <- matrix(0, nrow = n_mods_total, ncol = n_mods_total,
                              dimnames = list(colnames(mat_binary_total), colnames(mat_binary_total)))

  for (i in seq_len(n_mods_total)) {
    for (j in seq_len(n_mods_total)) {
      both <- sum(mat_binary_total[, i] == 1 & mat_binary_total[, j] == 1)
      either <- sum(mat_binary_total[, i] == 1 | mat_binary_total[, j] == 1)
      jaccard_mat_total[i, j] <- ifelse(either > 0, both / either, 0)
    }
  }

  save_and_show_pheatmap(
    jaccard_mat_total,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = jaccard_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Jaccard similarity (threshold: %.0f%% of total) %s",
                   threshold_frac * 100, title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_jaccard_total_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_jaccard_total_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_jaccard_total_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_jaccard_total_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 3e. Column-normalized co-targeting (fraction of modality j's targets) ----
  diag_counts <- diag(co_target_mat)
  diag_counts[diag_counts == 0] <- 1  # avoid division by zero
  co_target_colnorm <- sweep(co_target_mat, 2, diag_counts, "/")

  save_and_show_pheatmap(
    co_target_colnorm,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting (column-normalized) %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_colnorm_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_colnorm_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 4. Jaccard similarity ---------------------------------------------
  n_mods <- ncol(mat_binary)
  jaccard_mat <- matrix(0, nrow = n_mods, ncol = n_mods,
                        dimnames = list(colnames(mat_binary), colnames(mat_binary)))

  for (i in seq_len(n_mods)) {
    for (j in seq_len(n_mods)) {
      both   <- sum(mat_binary[, i] == 1 & mat_binary[, j] == 1)
      either <- sum(mat_binary[, i] == 1 | mat_binary[, j] == 1)
      jaccard_mat[i, j] <- ifelse(either > 0, both / either, 0)
    }
  }

  save_and_show_pheatmap(
    jaccard_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = jaccard_colors,
    clustering_method = "complete",
    annotation_col = group_annotation,
    annotation_row = group_annotation,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Co-targeting Jaccard similarity between modalities %s", title_suffix),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_jaccard_%s", label), "png", subdir = subdir),
                         sprintf("panel_E_cotarget_jaccard_%s.png", label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cotarget_jaccard_%s", label), "pdf", subdir = subdir),
                         sprintf("panel_E_cotarget_jaccard_%s.pdf", label)),
    save_width = 9,
    save_height = 8
  )

  # ---- 5. Correlation network (unthresholded) ----------------------------
  node_df <- data.frame(
    name = colnames(cor_pearson),
    group = modality_group[colnames(cor_pearson)],
    stringsAsFactors = FALSE
  )

  edge_df <- expand.grid(
    from = rownames(cor_pearson),
    to = colnames(cor_pearson),
    stringsAsFactors = FALSE
  ) %>%
    filter(from < to) %>%
    rowwise() %>%
    mutate(
      correlation = cor_pearson[from, to],
      abs_cor = abs(correlation),
      sign = ifelse(correlation >= 0, "positive", "negative")
    ) %>%
    ungroup()

  graph <- tbl_graph(nodes = node_df, edges = edge_df, directed = FALSE)

  p_network <- ggraph(graph, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign, alpha = abs_cor)
    ) +
    scale_edge_width_continuous(range = c(0.3, 4), name = "|Correlation|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
    geom_node_point(aes(color = group), size = 12) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Modality correlation network (Pearson) %s", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

  print(p_network)
  save_fig6_plot(p_network,
                 sprintf("panel_E_correlation_network_%s", label),
                 width = 9, height = 8, subdir = subdir)

  # ---- 6. Correlation network (thresholded |r| > 0.1) --------------------
  edge_df_thresh <- edge_df %>% filter(abs_cor > 0.1)

  graph_thresh <- tbl_graph(nodes = node_df, edges = edge_df_thresh, directed = FALSE)

  p_network_thresh <- ggraph(graph_thresh, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign),
      alpha = 0.8
    ) +
    geom_edge_link(
      aes(label = sprintf("%.2f", correlation)),
      angle_calc = "along",
      label_size = 2.5,
      label_dodge = unit(3, "mm"),
      alpha = 0
    ) +
    scale_edge_width_continuous(range = c(0.5, 4), name = "|Correlation|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    geom_node_point(aes(color = group), size = 14) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Modality correlation network (|r| > 0.1, Pearson) %s", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

  print(p_network_thresh)
  save_fig6_plot(p_network_thresh,
                 sprintf("panel_E_correlation_network_thresholded_%s", label),
                 width = 9, height = 8, subdir = subdir)

  # ---- 7. Averaged-metrics network (mean of Spearman + Cosine + Jaccard) ----
  avg_mat <- (cor_spearman + cosine_mat + jaccard_mat) / 3

  edge_df_avg <- expand.grid(
    from = rownames(avg_mat),
    to = colnames(avg_mat),
    stringsAsFactors = FALSE
  ) %>%
    filter(from < to) %>%
    rowwise() %>%
    mutate(
      correlation = avg_mat[from, to],
      abs_cor = abs(correlation),
      sign = ifelse(correlation >= 0, "positive", "negative")
    ) %>%
    ungroup()

  graph_avg <- tbl_graph(nodes = node_df, edges = edge_df_avg, directed = FALSE)

  p_network_avg <- ggraph(graph_avg, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign, alpha = abs_cor)
    ) +
    scale_edge_width_continuous(range = c(0.3, 4), name = "|Similarity|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
    geom_node_point(aes(color = group), size = 12) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Averaged similarity network %s\n(mean of Spearman + Cosine + Jaccard)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 12)
    )

  print(p_network_avg)
  save_fig6_plot(p_network_avg,
                 sprintf("panel_E_avg_network_%s", label),
                 width = 9, height = 8, subdir = subdir)

  # Thresholded averaged network (score > 0.1)
  edge_df_avg_thresh <- edge_df_avg %>% filter(abs_cor > 0.1)

  graph_avg_thresh <- tbl_graph(nodes = node_df, edges = edge_df_avg_thresh, directed = FALSE)

  p_network_avg_thresh <- ggraph(graph_avg_thresh, layout = "circle") +
    geom_edge_link(
      aes(width = abs_cor, color = sign),
      alpha = 0.8
    ) +
    geom_edge_link(
      aes(label = sprintf("%.2f", correlation)),
      angle_calc = "along",
      label_size = 2.5,
      label_dodge = unit(3, "mm"),
      alpha = 0
    ) +
    scale_edge_width_continuous(range = c(0.5, 4), name = "|Similarity|") +
    scale_edge_color_manual(
      values = c("positive" = "#B2182B", "negative" = "#2166AC"),
      name = "Direction"
    ) +
    geom_node_point(aes(color = group), size = 14) +
    geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
    scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
    labs(title = sprintf("Averaged similarity network %s (score > 0.1)", title_suffix)) +
    theme_void(base_size = 11) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 12)
    )

  print(p_network_avg_thresh)
  save_fig6_plot(p_network_avg_thresh,
                 sprintf("panel_E_avg_network_%s_thresholded", label),
                 width = 9, height = 8, subdir = subdir)

  # ---- Return computed matrices ------------------------------------------
  invisible(list(
    cor_pearson   = cor_pearson,
    cor_spearman  = cor_spearman,
    co_target_mat = co_target_mat,
    jaccard_mat   = jaccard_mat
  ))
}
```

```{r cosine-helper, include=FALSE}
# Cosine similarity heatmap for a given strength matrix and label
plot_cosine <- function(strength_mat, valence_label, subdir = NULL) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping cosine %s: insufficient data", valence_label))
    return(invisible(NULL))
  }

  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping cosine %s: fewer than 2 variable modalities", valence_label))
    return(invisible(NULL))
  }

  col_norms <- sqrt(colSums(mat_v^2))
  col_norms[col_norms == 0] <- 1
  mat_normed <- sweep(mat_v, 2, col_norms, "/")
  cosine_mat <- t(mat_normed) %*% mat_normed

  ga <- data.frame(
    group = modality_group[colnames(cosine_mat)],
    row.names = colnames(cosine_mat)
  )

  save_and_show_pheatmap(
    cosine_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cosine_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = sprintf("Cosine similarity (%s, Bella's targets)", valence_label),
    fontsize = 10,
    png_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s_bella", valence_label), "png", subdir = subdir),
                         sprintf("panel_E_cosine_%s_bella.png", valence_label)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_E_cosine_%s_bella", valence_label), "pdf", subdir = subdir),
                         sprintf("panel_E_cosine_%s_bella.pdf", valence_label)),
    save_width = 9,
    save_height = 8
  )

  invisible(cosine_mat)
}
```

```{r valence-helpers, include=FALSE}
# ---------------------------------------------------------------------------
# Valence-split helpers (accept subdir parameter)
# ---------------------------------------------------------------------------

# Helper to compute and plot correlation for a given strength matrix and label
plot_valence_correlation <- function(strength_mat, valence_label, subdir = NULL) {
  # Ensure modality order
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  # Skip if too few columns or rows
  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping %s: insufficient data (%d rows, %d modalities)",
                    valence_label, nrow(mat_v), ncol(mat_v)))
    return(invisible(NULL))
  }

  # Remove rows with zero variance (no information)
  row_var <- apply(mat_v, 1, var, na.rm = TRUE)
  mat_v <- mat_v[row_var > 0, , drop = FALSE]

  # Remove modality columns with zero variance
  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping %s: fewer than 2 variable modalities after filtering", valence_label))
    return(invisible(NULL))
  }

  cor_mat <- cor(mat_v, method = "pearson")

  # Annotation for remaining modalities
  ga <- data.frame(
    group = modality_group[colnames(cor_mat)],
    row.names = colnames(cor_mat)
  )

  title_str <- sprintf("Pearson correlation (%s paths only, Bella's targets)", valence_label)

  # Build once, save PNG + PDF, display inline
  p <- pheatmap(
    cor_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  png(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s_bella", valence_label), "png", subdir = subdir), sprintf("panel_E_pearson_%s_bella.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s_bella", valence_label), "pdf", subdir = subdir), sprintf("panel_E_pearson_%s_bella.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(cor_mat)
}

# Co-targeting heatmaps per valence
compute_cotarget <- function(strength_mat, valence_label, subdir = NULL) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  mod_max_v <- apply(mat_v, 2, max, na.rm = TRUE)
  thresh_v <- mod_max_v * threshold_frac
  # Guard against zero max (would make threshold 0, flagging everything)
  thresh_v[thresh_v == 0] <- .Machine$double.eps

  mat_bin_v <- sweep(mat_v, 2, thresh_v, ">") * 1
  co_mat <- t(mat_bin_v) %*% mat_bin_v

  ga <- data.frame(
    group = modality_group[colnames(co_mat)],
    row.names = colnames(co_mat)
  )

  title_str <- sprintf("Co-targeting counts (%s, threshold: %.0f%%, Bella's targets)", valence_label, threshold_frac * 100)

  # Build once, save PNG + PDF, display inline
  p <- pheatmap(
    co_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  png(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s_bella", valence_label), "png", subdir = subdir), sprintf("panel_E_cotarget_%s_bella.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s_bella", valence_label), "pdf", subdir = subdir), sprintf("panel_E_cotarget_%s_bella.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(co_mat)
}
```

## 1-4. Combined analysis

```{r combined, fig.width=10, fig.height=8, results='asis'}
mat <- strength.bella[, intersect(modality_order, colnames(strength.bella))]

results <- run_correlation_suite(
  mat, label = "bella", subdir = "combined",
  title_suffix = "(Bella's targets)"
)
```

## Net strength analysis (excitatory - inhibitory)

```{r combined-net-strength, fig.width=10, fig.height=8, results='asis'}
strength.bella.net <- compute_net_strength(strength.bella.exc, strength.bella.inh)
mat_net <- strength.bella.net[, intersect(modality_order, colnames(strength.bella.net))]

results_net <- run_correlation_suite(
  mat_net, label = "bella", subdir = "combined/net_strength",
  title_suffix = "(Bella's targets, net strength)"
)
```

## 5. Valence-split correlation analysis

Repeat the correlation heatmaps for excitatory-only and inhibitory-only paths
to determine whether co-occurrence patterns differ by valence.

```{r valence-exc-pearson, fig.width=9, fig.height=8}
plot_valence_correlation(strength.bella.exc, "excitatory", subdir = "valence")
```

```{r valence-inh-pearson, fig.width=9, fig.height=8}
plot_valence_correlation(strength.bella.inh, "inhibitory", subdir = "valence")
```

```{r valence-exc-cotarget, fig.width=9, fig.height=8}
compute_cotarget(strength.bella.exc, "excitatory", subdir = "valence")
```

```{r valence-inh-cotarget, fig.width=9, fig.height=8}
compute_cotarget(strength.bella.inh, "inhibitory", subdir = "valence")
```

```{r valence-exc-cosine, fig.width=9, fig.height=8}
plot_cosine(strength.bella.exc, "excitatory", subdir = "valence")
```

```{r valence-inh-cosine, fig.width=9, fig.height=8}
plot_cosine(strength.bella.inh, "inhibitory", subdir = "valence")
```

## 6. Summary statistics

```{r correlation-summary}
cat("=== Pearson Correlation Summary (combined, Bella's targets) ===\n")
cor_pearson  <- results$cor_pearson
cor_spearman <- results$cor_spearman
co_target_mat <- results$co_target_mat

# Extract upper triangle values (excluding diagonal)
upper_idx <- upper.tri(cor_pearson)
cor_vals <- cor_pearson[upper_idx]
pair_labels <- outer(rownames(cor_pearson), colnames(cor_pearson), paste, sep = " - ")[upper_idx]

cor_summary <- data.frame(
  pair = pair_labels,
  pearson = cor_vals,
  spearman = cor_spearman[upper_idx]
) %>%
  arrange(desc(abs(pearson)))

cat(sprintf("Number of modality pairs: %d\n", length(cor_vals)))
cat(sprintf("Mean |Pearson r|: %.3f\n", mean(abs(cor_vals))))
cat(sprintf("Max positive: %.3f\n", max(cor_vals)))
cat(sprintf("Max negative: %.3f\n", min(cor_vals)))
cat(sprintf("Pairs with |r| > 0.3: %d\n", sum(abs(cor_vals) > 0.3)))

cat("\nAll pairwise correlations (sorted by |Pearson|):\n")
knitr::kable(cor_summary, digits = 3)
```

```{r co-targeting-summary}
cat("\n=== Co-targeting Summary (Bella's targets) ===\n")
cat(sprintf("Threshold: %.0f%% of modality max strength\n", threshold_frac * 100))
cat(sprintf("Total target neurons: %d\n", nrow(mat)))

# Diagonal = neurons per modality above threshold
cat("\nNeurons above threshold per modality:\n")
print(diag(co_target_mat))

# Off-diagonal co-targeting
co_upper <- co_target_mat[upper.tri(co_target_mat)]
co_labels <- outer(rownames(co_target_mat), colnames(co_target_mat), paste, sep = " & ")[upper.tri(co_target_mat)]

co_summary <- data.frame(
  pair = co_labels,
  shared_neurons = co_upper
) %>%
  arrange(desc(shared_neurons))

cat("\nTop co-targeting pairs:\n")
knitr::kable(co_summary)
```

```{r net-strength-summary}
cat("=== Pearson Correlation Summary (net strength, Bella's targets) ===\n")
cor_pearson_net  <- results_net$cor_pearson
cor_spearman_net <- results_net$cor_spearman
co_target_mat_net <- results_net$co_target_mat

# Extract upper triangle values (excluding diagonal)
upper_idx_net <- upper.tri(cor_pearson_net)
cor_vals_net <- cor_pearson_net[upper_idx_net]
pair_labels_net <- outer(rownames(cor_pearson_net), colnames(cor_pearson_net), paste, sep = " - ")[upper_idx_net]

cor_summary_net <- data.frame(
  pair = pair_labels_net,
  pearson = cor_vals_net,
  spearman = cor_spearman_net[upper_idx_net]
) %>%
  arrange(desc(abs(pearson)))

cat(sprintf("Number of modality pairs: %d\n", length(cor_vals_net)))
cat(sprintf("Mean |Pearson r|: %.3f\n", mean(abs(cor_vals_net))))
cat(sprintf("Max positive: %.3f\n", max(cor_vals_net)))
cat(sprintf("Max negative: %.3f\n", min(cor_vals_net)))
cat(sprintf("Pairs with |r| > 0.3: %d\n", sum(abs(cor_vals_net) > 0.3)))

cat("\nAll pairwise correlations (sorted by |Pearson|):\n")
knitr::kable(cor_summary_net, digits = 3)
```

```{r net-strength-co-targeting-summary}
cat("\n=== Co-targeting Summary (net strength, Bella's targets) ===\n")
cat(sprintf("Threshold: %.0f%% of modality max net strength\n", threshold_frac * 100))
cat(sprintf("Total target neurons: %d\n", nrow(mat_net)))

# Diagonal = neurons per modality above threshold
cat("\nNeurons above threshold per modality:\n")
print(diag(co_target_mat_net))

# Off-diagonal co-targeting
co_upper_net <- co_target_mat_net[upper.tri(co_target_mat_net)]
co_labels_net <- outer(rownames(co_target_mat_net), colnames(co_target_mat_net), paste, sep = " & ")[upper.tri(co_target_mat_net)]

co_summary_net <- data.frame(
  pair = co_labels_net,
  shared_neurons = co_upper_net
) %>%
  arrange(desc(shared_neurons))

cat("\nTop co-targeting pairs:\n")
knitr::kable(co_summary_net)
```

```{r session-info}
sessionInfo()
```
