---
title: "Figure 6B (Bella's targets): Modality Composition Heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel B: Modality Composition Heatmap (Bella's targets) — Clustered heatmap of column-normalized modality input strength to Bella's target neurons. Rows are target neuron types; columns are the 7 sensory modalities. Row annotations mark target group (M, F, IR downstream)."
panel_legend <- "(B) Column-normalized heatmap of modality input strength to Bella's target neurons. Rows are clustered by Ward's method; annotations indicate target group membership (M-downstream, F-downstream, IR-downstream)."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_B_description", "txt"), "panel_B_description.txt"))
```

## Overview

Panel B is a clustered heatmap of modality input strength to Bella's third-order
courtship target neurons. Columns are the 7 sensory modalities; rows are target
neuron types. The matrix is column-normalized so that each modality's maximum
is 1, making the relative distribution across targets comparable between
modalities of different absolute strengths.

Row annotations indicate the target group (M_downstream, F_downstream,
IR_downstream). Column annotations indicate the modality group
(olfactory / auditory / visual / contact).

## 1. Prepare strength matrix

```{r strength-matrix}
# ---------------------------------------------------------------------------
# Column-normalize the pre-computed strength matrix
# Each column (modality) is scaled so its maximum value = 1
# ---------------------------------------------------------------------------
mat <- strength.bella

# Column-normalize: divide each column by its max
col_maxes <- apply(mat, 2, max, na.rm = TRUE)
col_maxes[col_maxes == 0] <- 1
mat_norm <- sweep(mat, 2, col_maxes, "/")

cat(sprintf("Matrix dimensions: %d target neurons x %d modalities\n",
            nrow(mat_norm), ncol(mat_norm)))
```

## 2. Build row annotation (target group membership)

Each target neuron belongs to one of Bella's three target groups:
M_downstream, F_downstream, or IR_downstream. We use `bella_groups_df`
to assign primary group membership.

```{r row-annotation}
# ---------------------------------------------------------------------------
# Determine target_group membership for each target neuron
# ---------------------------------------------------------------------------
# Multiple color bands: one column per target group (yes/no)
# Neurons belonging to multiple groups show all their memberships
all_ends <- rownames(mat_norm)

M_ends  <- unique(paths.M$end)
F_ends  <- unique(paths.F$end)
IR_ends <- unique(paths.IR$end)

row_annot <- data.frame(
  `M cells`  = ifelse(all_ends %in% M_ends,  "yes", "no"),
  `F cells`  = ifelse(all_ends %in% F_ends,  "yes", "no"),
  `IR cells` = ifelse(all_ends %in% IR_ends, "yes", "no"),
  row.names  = all_ends,
  check.names = FALSE
)
```

## 3. Build column annotation (modality group)

```{r col-annotation}
# ---------------------------------------------------------------------------
# Map each modality column to its sensory group
# ---------------------------------------------------------------------------
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_norm)],
  row.names      = colnames(mat_norm)
)
```

## 4. Define annotation colours

```{r annotation-colors}
# ---------------------------------------------------------------------------
# Annotation colour maps for pheatmap
# ---------------------------------------------------------------------------
ann_colors <- list(
  `M cells`      = c("yes" = "#E41A1C", "no" = "white"),
  `F cells`      = c("yes" = "#377EB8", "no" = "white"),
  `IR cells`     = c("yes" = "#4DAF4A", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 5. Draw heatmap

```{r heatmap-bella, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Primary heatmap: modality strength to Bella's target groups
# ---------------------------------------------------------------------------
hm <- pheatmap(
  mat_norm,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,       # keep modality order fixed
  cluster_rows             = TRUE,
  annotation_row           = row_annot,
  annotation_col           = col_annot,
  annotation_colors        = ann_colors,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons (Bella's targets)",
  angle_col                = 45,
  silent                   = TRUE
)

# Draw the heatmap
hm
```

## 6. Save heatmap

```{r save-heatmap}
save_heatmap_both_orientations(
  mat        = mat_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_bella",
  width      = 8,
  height     = 10,
  subdir     = "colnorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons (Bella's targets)",
  angle_col                = 45
)
```

## 6b. Row-normalized heatmap

```{r row-norm-matrix}
# Row-normalize: each neuron's max across modalities = 1
row_maxes <- apply(mat, 1, max, na.rm = TRUE)
row_maxes[row_maxes == 0] <- 1
mat_row_norm <- sweep(mat, 1, row_maxes, "/")
```

```{r save-row-norm}
save_heatmap_both_orientations(
  mat        = mat_row_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_bella_rownorm",
  width      = 8,
  height     = 10,
  subdir     = "rownorm",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (row-normalized, Bella's targets)",
  angle_col                = 45
)
```

## 6c. Unnormalized (raw strength) heatmap

```{r save-raw}
save_heatmap_both_orientations(
  mat        = mat,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_bella_raw",
  width      = 8,
  height     = 10,
  subdir     = "raw",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength (unnormalized, Bella's targets)",
  angle_col                = 45
)
```

## 6d. Per-target-group heatmaps (M, F, IR separately)

```{r per-group-heatmaps}
# Build separate heatmaps for each target group, with all 3 normalizations
group_list <- list(
  M  = list(paths = paths.M,  strength = strength.M,  label = "M-downstream"),
  F  = list(paths = paths.F,  strength = strength.F,  label = "F-downstream"),
  IR = list(paths = paths.IR, strength = strength.IR, label = "IR-downstream")
)

for (grp_name in names(group_list)) {
  grp <- group_list[[grp_name]]
  mat_grp <- grp$strength

  # Column-normalize
  cm <- apply(mat_grp, 2, max, na.rm = TRUE)
  cm[cm == 0] <- 1
  mat_grp_colnorm <- sweep(mat_grp, 2, cm, "/")

  # Row-normalize
  rm_grp <- apply(mat_grp, 1, max, na.rm = TRUE)
  rm_grp[rm_grp == 0] <- 1
  mat_grp_rownorm <- sweep(mat_grp, 1, rm_grp, "/")

  # Row annotation: just mark all as "yes" for this group
  grp_ends <- rownames(mat_grp)
  ra <- data.frame(
    group = rep("yes", length(grp_ends)),
    row.names = grp_ends,
    check.names = FALSE
  )
  names(ra) <- grp$label

  ca <- data.frame(
    modality_group = modality_group[colnames(mat_grp)],
    row.names      = colnames(mat_grp)
  )

  ac <- list(modality_group = modality_group_colors)
  ac[[grp$label]] <- c("yes" = target_set_colors[[paste0(grp_name, "_downstream")]], "no" = "white")

  hm_base <- sprintf("panel_B_heatmap_bella_%s_only", tolower(grp_name))
  hm_height <- max(6, nrow(mat_grp) * 0.12 + 2)

  # Common heatmap args
  hm_args <- list(
    row_annot  = ra,
    col_annot  = ca,
    ann_colors = ac,
    width      = 8,
    height     = hm_height,
    clustering_distance_rows = "euclidean",
    clustering_distance_cols = "euclidean",
    clustering_method        = "ward.D2",
    cluster_cols             = FALSE,
    cluster_rows             = TRUE,
    color                    = viridis(100),
    border_color             = NA,
    fontsize_row             = 7,
    fontsize_col             = 10,
    angle_col                = 45
  )

  # Column-normalized
  do.call(save_heatmap_both_orientations, c(list(
    mat    = mat_grp_colnorm,
    name   = hm_base,
    subdir = "colnorm",
    main   = sprintf("Modality input strength (%s only)", grp$label)
  ), hm_args))

  # Row-normalized
  do.call(save_heatmap_both_orientations, c(list(
    mat    = mat_grp_rownorm,
    name   = paste0(hm_base, "_rownorm"),
    subdir = "rownorm",
    main   = sprintf("Modality input strength (row-normalized, %s only)", grp$label)
  ), hm_args))

  # Raw (unnormalized)
  do.call(save_heatmap_both_orientations, c(list(
    mat    = mat_grp,
    name   = paste0(hm_base, "_raw"),
    subdir = "raw",
    main   = sprintf("Modality input strength (unnormalized, %s only)", grp$label)
  ), hm_args))

  cat(sprintf("%s: %d neurons x %d modalities (colnorm + rownorm + raw saved)\n",
              grp$label, nrow(mat_grp), ncol(mat_grp)))
}
```

## 6e. Rank-transformed scatter (all Bella's targets)

PCA on within-column rank-transformed strength profiles. Ranking removes
magnitude effects while preserving ordinal structure — neurons that receive
the same relative ordering of modalities cluster together regardless of
their absolute strength.

```{r rank-scatter-bella, fig.width=10, fig.height=8}
# ---------------------------------------------------------------------------
# Rank-transform each modality column, then PCA
# ---------------------------------------------------------------------------
mat_rank_bella <- apply(mat, 2, rank, ties.method = "average")
pca_rank <- prcomp(mat_rank_bella, center = TRUE, scale. = TRUE)

var_pct <- summary(pca_rank)$importance[2, ] * 100

rank_df <- data.frame(
  PC1    = pca_rank$x[, 1],
  PC2    = pca_rank$x[, 2],
  neuron = rownames(mat),
  stringsAsFactors = FALSE
)

# Assign primary target group for coloring
rank_df$target_group <- {
  is_M  <- rank_df$neuron %in% M_ends
  is_F  <- rank_df$neuron %in% F_ends
  is_IR <- rank_df$neuron %in% IR_ends
  n_grp <- is_M + is_F + is_IR
  ifelse(n_grp > 1, "multiple groups",
  ifelse(is_M,  "M-downstream",
  ifelse(is_F,  "F-downstream",
  ifelse(is_IR, "IR-downstream", "other"))))
}

group_colors_scatter <- c(
  "M-downstream"   = "#E41A1C",
  "F-downstream"   = "#377EB8",
  "IR-downstream"  = "#4DAF4A",
  "multiple groups" = "#984EA3",
  "other"          = "grey60"
)

p_rank <- ggplot(rank_df, aes(x = PC1, y = PC2, color = target_group)) +
  geom_point(size = 2.5, alpha = 0.7) +
  scale_color_manual(values = group_colors_scatter) +
  labs(
    x = sprintf("PC1 (%.1f%% variance)", var_pct[1]),
    y = sprintf("PC2 (%.1f%% variance)", var_pct[2]),
    title = "PCA of rank-transformed modality profiles (Bella's targets)",
    color = "Target group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right")

p_rank
save_fig6_plot(p_rank, "panel_B_rank_scatter_bella", width = 10, height = 8)
```

## 6f. Correspondence analysis ordination (all Bella's targets)

Correspondence analysis is the natural ordination for compositional data:
neurons and modalities are projected into the same chi-squared metric space.
Neurons close to a modality arrow are enriched for that modality relative to
the average profile.

```{r ca-ordination-bella, fig.width=10, fig.height=8}
# ---------------------------------------------------------------------------
# Manual Correspondence Analysis via SVD
# ---------------------------------------------------------------------------
mat_ca <- mat
mat_ca[mat_ca < 0] <- 0
mat_ca <- mat_ca + 1e-10   # avoid exact zeros

N_total <- sum(mat_ca)
P_ca    <- mat_ca / N_total
ri      <- rowSums(P_ca)
ci      <- colSums(P_ca)

# Standardized residuals
Z_ca <- diag(1 / sqrt(ri)) %*% (P_ca - outer(ri, ci)) %*% diag(1 / sqrt(ci))
sv   <- svd(Z_ca)

# Inertia explained per axis
total_inertia <- sum(sv$d^2)
inertia_pct   <- sv$d^2 / total_inertia * 100

# Row principal coordinates (neurons)
F_row <- diag(1 / sqrt(ri)) %*% sv$u[, 1:2] %*% diag(sv$d[1:2])
rownames(F_row) <- rownames(mat_ca)

# Column principal coordinates (modalities)
G_col <- diag(1 / sqrt(ci)) %*% sv$v[, 1:2] %*% diag(sv$d[1:2])
rownames(G_col) <- colnames(mat_ca)

# Build data frames
ca_neurons <- data.frame(
  Dim1 = F_row[, 1], Dim2 = F_row[, 2],
  neuron = rownames(F_row),
  stringsAsFactors = FALSE
)
ca_neurons$target_group <- rank_df$target_group[match(ca_neurons$neuron, rank_df$neuron)]

ca_modalities <- data.frame(
  Dim1 = G_col[, 1], Dim2 = G_col[, 2],
  modality = rownames(G_col),
  stringsAsFactors = FALSE
)

p_ca <- ggplot() +
  geom_point(data = ca_neurons,
             aes(x = Dim1, y = Dim2, color = target_group),
             size = 2, alpha = 0.6) +
  geom_segment(data = ca_modalities,
               aes(x = 0, y = 0, xend = Dim1, yend = Dim2),
               arrow = arrow(length = unit(0.25, "cm")),
               color = "grey30", linewidth = 0.8) +
  geom_text(data = ca_modalities,
            aes(x = Dim1 * 1.12, y = Dim2 * 1.12, label = modality),
            color = "grey20", size = 4, fontface = "bold") +
  scale_color_manual(values = group_colors_scatter) +
  labs(
    x = sprintf("CA Dim 1 (%.1f%% inertia)", inertia_pct[1]),
    y = sprintf("CA Dim 2 (%.1f%% inertia)", inertia_pct[2]),
    title = "Correspondence analysis: neurons and modalities (Bella's targets)",
    color = "Target group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right") +
  coord_fixed()

p_ca
save_fig6_plot(p_ca, "panel_B_ca_ordination_bella", width = 10, height = 8)
```

## 7. Export row clustering order for use in Panel C

Save the dendrogram row order so Panel C can use the same neuron ordering
for consistent visual comparison.

```{r export-row-order}
# ---------------------------------------------------------------------------
# Extract row order from the heatmap's dendrogram
# ---------------------------------------------------------------------------
row_order_bella <- rownames(mat_norm)[hm$tree_row$order]

# Save to an RDS file for Panel C to load
saveRDS(
  row_order_bella,
  file = file.path(fig6_dir, "panel_B_row_order.rds")
)

cat("Row order saved. First 10 neurons:\n")
head(row_order_bella, 10)
```
