---
title: "Figure 6I (Bella's targets): M vs. F vs. IR Downstream Comparison"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel I: Target Group Comparison (Bella's targets) — Compares modality input profiles between M-downstream, F-downstream, and IR-downstream neuron populations. Includes paired bar charts, radar overlays, entropy boxplots, and valence-split analyses."
panel_legend <- "(I) Comparison of modality input profiles across M-downstream, F-downstream, and IR-downstream target groups. Paired bar charts, radar overlays, entropy boxplots, and valence-split analyses reveal differences in sensory integration between target populations."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_I_description", "txt"), "panel_I_description.txt"))
```

## Overview

Panel I directly compares the modality input profiles of the three
Bella target populations: M cell downstream, F cell downstream, and
IR cell downstream neurons. These are distinct target sets receiving
input from M cells, F cells, and IR-expressing cells respectively.
We compare their mean profiles, individual neuron distributions, and
valence-split patterns in a 3-way analysis.

## Step 1: Prepare strength data for all three target sets

```{r prepare-data}
# ---------------------------------------------------------------------------
# Strength matrices are pre-computed in setup.R:
#   strength.M  - rows = neuron types targeting M cells, cols = 7 modalities
#   strength.F  - rows = neuron types targeting F cells, cols = 7 modalities
#   strength.IR - rows = neuron types targeting IR cells, cols = 7 modalities
# ---------------------------------------------------------------------------

# Ensure column order is consistent
strength.M  <- strength.M[, modality_order, drop = FALSE]
strength.F  <- strength.F[, modality_order, drop = FALSE]
strength.IR <- strength.IR[, modality_order, drop = FALSE]

message(sprintf("M downstream:  %d neurons x %d modalities", nrow(strength.M), ncol(strength.M)))
message(sprintf("F downstream:  %d neurons x %d modalities", nrow(strength.F), ncol(strength.F)))
message(sprintf("IR downstream: %d neurons x %d modalities", nrow(strength.IR), ncol(strength.IR)))

# ---------------------------------------------------------------------------
# Compute per-neuron normalized profiles (each neuron sums to 1)
# ---------------------------------------------------------------------------
normalize_rows <- function(mat) {
  rs <- rowSums(mat, na.rm = TRUE)
  rs[rs == 0] <- 1
  sweep(mat, 1, rs, "/")
}

norm.M  <- normalize_rows(strength.M)
norm.F  <- normalize_rows(strength.F)
norm.IR <- normalize_rows(strength.IR)

# ---------------------------------------------------------------------------
# Compute mean and SE profiles for each group
# ---------------------------------------------------------------------------
compute_mean_se <- function(norm_mat) {
  means <- colMeans(norm_mat, na.rm = TRUE)
  ses   <- apply(norm_mat, 2, function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))))
  list(mean = means, se = ses)
}

stats_M  <- compute_mean_se(norm.M)
stats_F  <- compute_mean_se(norm.F)
stats_IR <- compute_mean_se(norm.IR)

# Build a tidy summary dataframe
profile_summary <- data.frame(
  modality = factor(rep(modality_order, 3), levels = modality_order),
  mean     = c(stats_M$mean[modality_order],
               stats_F$mean[modality_order],
               stats_IR$mean[modality_order]),
  se       = c(stats_M$se[modality_order],
               stats_F$se[modality_order],
               stats_IR$se[modality_order]),
  group    = rep(c("M_downstream", "F_downstream", "IR_downstream"), each = 7),
  stringsAsFactors = FALSE
) %>%
  mutate(group_label = target_set_labels[group])

profile_summary
```

## Plot 1: Grouped bar chart (3-way)

Grouped bars comparing mean normalized strength per modality across
M, F, and IR downstream neurons, with standard error bars.

```{r plot1-grouped-bar, fig.width=10, fig.height=5.5}
# ---------------------------------------------------------------------------
# Grouped (dodge) bar chart: 3 groups
# ---------------------------------------------------------------------------
p1 <- ggplot(profile_summary,
             aes(x = modality, y = mean, fill = group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7,
           color = "white", linewidth = 0.3) +
  geom_errorbar(
    aes(ymin = pmax(mean - se, 0), ymax = mean + se),
    position = position_dodge(width = 0.8),
    width = 0.2, linewidth = 0.4
  ) +
  scale_fill_manual(
    values = target_set_colors,
    labels = target_set_labels,
    name   = "Target set"
  ) +
  labs(
    x     = "Modality",
    y     = "Mean normalized strength",
    title = "Modality input profiles: M vs. F vs. IR downstream"
  ) +
  theme_bw(base_size = 12) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    legend.position   = c(0.85, 0.85),
    legend.background = element_rect(fill = alpha("white", 0.8)),
    panel.grid.minor  = element_blank()
  )

p1
```

```{r save-plot1}
save_fig6_plot(p1, "panel_I_bella_grouped_bar", width = 10, height = 5.5)
```

## Plot 2: Faceted tornado chart (3-way)

Instead of a two-sided butterfly chart, we use a faceted horizontal
bar chart with one facet per target group for direct comparison.

```{r plot2-tornado, fig.width=10, fig.height=7}
# ---------------------------------------------------------------------------
# Faceted horizontal bar chart: one panel per group
# ---------------------------------------------------------------------------
tornado_df <- profile_summary %>%
  mutate(group_label = factor(group_label,
                               levels = c("M cells", "F cells", "IR cells")))

p2 <- ggplot(tornado_df,
             aes(x = modality, y = mean, fill = group)) +
  geom_col(width = 0.65, color = "white", linewidth = 0.3) +
  geom_errorbar(
    aes(ymin = pmax(mean - se, 0), ymax = mean + se),
    width = 0.2, linewidth = 0.4
  ) +
  coord_flip() +
  facet_wrap(~ group_label, nrow = 1) +
  scale_fill_manual(
    values = target_set_colors,
    labels = target_set_labels,
    guide  = "none"
  ) +
  labs(
    x     = "Modality",
    y     = "Mean normalized strength",
    title = "Faceted comparison: M vs. F vs. IR downstream"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.text        = element_text(face = "bold", size = 12),
    panel.grid.minor  = element_blank()
  )

p2
```

```{r save-plot2}
save_fig6_plot(p2, "panel_I_bella_tornado_faceted", width = 10, height = 7)
```

## Plot 3: Radar comparison (3 overlaid polygons)

Overlay three semi-transparent radar polygons (M=red, F=blue, IR=green)
on the same polar plot.

```{r plot3-radar, fig.width=7, fig.height=6}
# ---------------------------------------------------------------------------
# Radar overlay: M vs. F vs. IR mean profiles
# ---------------------------------------------------------------------------
# Normalize all profiles to 0-1 range (relative to the global max)
global_max <- max(c(stats_M$mean, stats_F$mean, stats_IR$mean), na.rm = TRUE)
if (global_max == 0) global_max <- 1

radar_comp <- data.frame(
  modality = factor(rep(modality_order, 3), levels = modality_order),
  value    = c(stats_M$mean[modality_order] / global_max,
               stats_F$mean[modality_order] / global_max,
               stats_IR$mean[modality_order] / global_max),
  group    = rep(c("M_downstream", "F_downstream", "IR_downstream"), each = 7),
  stringsAsFactors = FALSE
)

p3 <- ggplot(radar_comp,
             aes(x = modality, y = value, group = group,
                 fill = group, color = group)) +
  geom_polygon(alpha = 0.2, linewidth = 1) +
  geom_point(size = 3) +
  scale_fill_manual(values = target_set_colors, labels = target_set_labels) +
  scale_color_manual(values = target_set_colors, labels = target_set_labels) +
  coord_polar(start = 0) +
  scale_y_continuous(limits = c(0, 1.1)) +
  labs(
    title = "Radar comparison: M vs. F vs. IR mean modality profiles"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.title        = element_blank(),
    axis.text.y       = element_blank(),
    axis.text.x       = element_text(size = 10, face = "bold"),
    panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
    plot.title         = element_text(hjust = 0.5, face = "bold"),
    legend.position    = "bottom",
    legend.title       = element_blank()
  )

p3
```

```{r save-plot3}
save_fig6_plot(p3, "panel_I_bella_radar_comparison", width = 7, height = 6)
```

## Plot 4: Heatmap comparison (3 groups)

Stack the three strength matrices vertically (M on top, F in middle,
IR on bottom). Cluster within each group using pheatmap with row
annotations and gaps between groups.

```{r plot4-heatmap, fig.width=9, fig.height=12}
# ---------------------------------------------------------------------------
# Combine matrices and add group annotation
# ---------------------------------------------------------------------------
# Column-normalize for visualization (scale each modality to 0-1)
col_normalize <- function(mat) {
  cmax <- apply(mat, 2, max, na.rm = TRUE)
  cmax[cmax == 0] <- 1
  sweep(mat, 2, cmax, "/")
}

# Use unnormalized strength but column-scale for heatmap
combined_mat <- rbind(strength.M, strength.F, strength.IR)
combined_mat <- combined_mat[, modality_order, drop = FALSE]

# Build row annotation
row_annot <- data.frame(
  target_set = c(rep("M_downstream", nrow(strength.M)),
                 rep("F_downstream", nrow(strength.F)),
                 rep("IR_downstream", nrow(strength.IR))),
  row.names = rownames(combined_mat)
)

annot_colors <- list(
  target_set = target_set_colors
)

# Column-normalize for visual comparison
combined_mat_norm <- col_normalize(combined_mat)

# Define gaps between the three groups
gap_rows <- c(nrow(strength.M),
              nrow(strength.M) + nrow(strength.F))

# Draw heatmap
p4 <- pheatmap(
  combined_mat_norm,
  annotation_row    = row_annot,
  annotation_colors = annot_colors,
  cluster_cols      = FALSE,
  cluster_rows      = TRUE,
  clustering_method = "ward.D2",
  gaps_row          = gap_rows,
  color             = colorRampPalette(c("white", "#FEE08B", "#D73027"))(100),
  border_color      = NA,
  fontsize_row      = 6,
  fontsize_col      = 10,
  main              = "Modality strength: M (top) vs. F (middle) vs. IR (bottom)",
  angle_col         = 45,
  silent            = TRUE
)

# pheatmap returns a gtable, display it
grid::grid.newpage()
grid::grid.draw(p4$gtable)
```

```{r save-plot4}
# Save pheatmap output
png(file.path(panel_plot_dir("panel_I_bella_heatmap_comparison", "png"), "panel_I_bella_heatmap_comparison.png"),
    width = 9, height = 12, units = "in", res = 300)
grid::grid.draw(p4$gtable)
dev.off()

pdf(file.path(panel_plot_dir("panel_I_bella_heatmap_comparison", "pdf"), "panel_I_bella_heatmap_comparison.pdf"),
    width = 9, height = 12)
grid::grid.draw(p4$gtable)
dev.off()
```

## Plot 5: Entropy comparison (boxplot with pairwise tests)

Are M, F, or IR downstream neurons more or less multimodal?
Compare Shannon entropy distributions with pairwise Wilcoxon tests.

```{r plot5-entropy, fig.width=7, fig.height=5.5}
# ---------------------------------------------------------------------------
# Compute entropy for each neuron in each group
# ---------------------------------------------------------------------------
entropy_M <- data.frame(
  neuron  = rownames(strength.M),
  entropy = apply(strength.M, 1, shannon_entropy),
  group   = "M_downstream",
  stringsAsFactors = FALSE
)

entropy_F <- data.frame(
  neuron  = rownames(strength.F),
  entropy = apply(strength.F, 1, shannon_entropy),
  group   = "F_downstream",
  stringsAsFactors = FALSE
)

entropy_IR <- data.frame(
  neuron  = rownames(strength.IR),
  entropy = apply(strength.IR, 1, shannon_entropy),
  group   = "IR_downstream",
  stringsAsFactors = FALSE
)

entropy_all <- bind_rows(entropy_M, entropy_F, entropy_IR) %>%
  mutate(group_label = target_set_labels[group],
         group_label = factor(group_label, levels = c("M cells", "F cells", "IR cells")))

# Pairwise Wilcoxon rank-sum tests
pairs <- list(c("M_downstream", "F_downstream"),
              c("M_downstream", "IR_downstream"),
              c("F_downstream", "IR_downstream"))

pair_labels <- list(c("M cells", "F cells"),
                    c("M cells", "IR cells"),
                    c("F cells", "IR cells"))

pvals <- sapply(pairs, function(pair) {
  sub <- entropy_all %>% filter(group %in% pair)
  if (all(table(sub$group) >= 2)) {
    wt <- wilcox.test(entropy ~ group, data = sub)
    wt$p.value
  } else {
    NA_real_
  }
})

for (i in seq_along(pairs)) {
  message(sprintf("  %s vs %s: Wilcoxon p = %.3g",
                  pair_labels[[i]][1], pair_labels[[i]][2], pvals[i]))
}

# Build significance bracket y-positions
y_max <- max(entropy_all$entropy, na.rm = TRUE) * 1.05
y_step <- max(entropy_all$entropy, na.rm = TRUE) * 0.08

p5 <- ggplot(entropy_all, aes(x = group_label, y = entropy, fill = group)) +
  geom_boxplot(width = 0.5, outlier.shape = 21, outlier.size = 2, alpha = 0.7) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.2, stroke = 0, color = "grey30") +
  scale_fill_manual(values = target_set_colors, guide = "none") +
  # Add pairwise p-value annotations
  annotate("segment",
           x = 1, xend = 2,
           y = y_max, yend = y_max, linewidth = 0.4) +
  annotate("text", x = 1.5, y = y_max + y_step * 0.3,
           label = sprintf("p = %.3g", pvals[1]),
           size = 3, fontface = "italic") +
  annotate("segment",
           x = 1, xend = 3,
           y = y_max + y_step, yend = y_max + y_step, linewidth = 0.4) +
  annotate("text", x = 2, y = y_max + y_step + y_step * 0.3,
           label = sprintf("p = %.3g", pvals[2]),
           size = 3, fontface = "italic") +
  annotate("segment",
           x = 2, xend = 3,
           y = y_max + y_step * 2, yend = y_max + y_step * 2, linewidth = 0.4) +
  annotate("text", x = 2.5, y = y_max + y_step * 2 + y_step * 0.3,
           label = sprintf("p = %.3g", pvals[3]),
           size = 3, fontface = "italic") +
  labs(
    x     = "Target set",
    y     = "Shannon entropy (bits)",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

p5
```

```{r save-plot5}
save_fig6_plot(p5, "panel_I_bella_entropy_boxplot", width = 6, height = 5)
```

## Plot 6: Modality scatter (pairwise comparisons in facets)

For each modality, plot pairwise mean strength comparisons
(M vs F, M vs IR, F vs IR) in separate facets.

```{r plot6-scatter, fig.width=14, fig.height=5}
# ---------------------------------------------------------------------------
# Use the raw (unnormalized) mean strengths for the scatter
# ---------------------------------------------------------------------------
raw_mean_M  <- colMeans(strength.M, na.rm = TRUE)
raw_mean_F  <- colMeans(strength.F, na.rm = TRUE)
raw_mean_IR <- colMeans(strength.IR, na.rm = TRUE)

scatter_df <- bind_rows(
  data.frame(
    modality = modality_order,
    x_mean   = raw_mean_M[modality_order],
    y_mean   = raw_mean_F[modality_order],
    comparison = "M vs F",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality = modality_order,
    x_mean   = raw_mean_M[modality_order],
    y_mean   = raw_mean_IR[modality_order],
    comparison = "M vs IR",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality = modality_order,
    x_mean   = raw_mean_F[modality_order],
    y_mean   = raw_mean_IR[modality_order],
    comparison = "F vs IR",
    stringsAsFactors = FALSE
  )
) %>%
  mutate(comparison = factor(comparison, levels = c("M vs F", "M vs IR", "F vs IR")))

# Axis labels per facet
x_labels <- c("M vs F" = "Mean strength (M downstream)",
              "M vs IR" = "Mean strength (M downstream)",
              "F vs IR" = "Mean strength (F downstream)")
y_labels <- c("M vs F" = "Mean strength (F downstream)",
              "M vs IR" = "Mean strength (IR downstream)",
              "F vs IR" = "Mean strength (IR downstream)")

axis_max <- max(c(scatter_df$x_mean, scatter_df$y_mean), na.rm = TRUE) * 1.15

p6 <- ggplot(scatter_df,
             aes(x = x_mean, y = y_mean, color = modality, label = modality)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey75", linewidth = 0.3) +
  geom_point(alpha = 0.65, size = 2, shape = 16) +
  geom_text(nudge_x = axis_max * 0.04, nudge_y = axis_max * 0.03,
            size = 2.5, fontface = "bold", show.legend = FALSE) +
  facet_wrap(~ comparison, nrow = 1, scales = "free") +
  scale_color_manual(values = modality_colors, name = "Modality") +
  coord_fixed() +
  labs(
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    strip.text = element_text(face = "plain", size = 10, color = "grey30"),
    legend.position = "right"
  )

p6
```

```{r save-plot6}
save_fig6_plot(p6, "panel_I_bella_modality_scatter", width = 10, height = 4.5)
```

## Step 2: Valence-split comparison

Repeat the key comparisons separately for excitatory and inhibitory
inputs. This may reveal that excitatory profiles are similar but
inhibitory profiles differ between the three target sets.

```{r valence-split-data}
# ---------------------------------------------------------------------------
# Compute valence-specific strength matrices for each target set
# ---------------------------------------------------------------------------
strength.M.exc  <- compute_valence_strength(paths.M, "excitatory")
strength.M.inh  <- compute_valence_strength(paths.M, "inhibitory")
strength.F.exc  <- compute_valence_strength(paths.F, "excitatory")
strength.F.inh  <- compute_valence_strength(paths.F, "inhibitory")
strength.IR.exc <- compute_valence_strength(paths.IR, "excitatory")
strength.IR.inh <- compute_valence_strength(paths.IR, "inhibitory")

# Ensure column order for all valence matrices
for (nm in c("strength.M.exc", "strength.M.inh",
             "strength.F.exc", "strength.F.inh",
             "strength.IR.exc", "strength.IR.inh")) {
  mat <- get(nm)
  present_cols <- intersect(modality_order, colnames(mat))
  missing_cols <- setdiff(modality_order, colnames(mat))
  if (length(missing_cols) > 0) {
    for (mc in missing_cols) mat <- cbind(mat, setNames(data.frame(rep(0, nrow(mat))), mc))
  }
  assign(nm, as.matrix(mat[, modality_order, drop = FALSE]))
}

# Compute mean profiles per group and valence
build_valence_profile <- function(mat, group_label, valence_label) {
  norm_mat <- normalize_rows(mat)
  means <- colMeans(norm_mat, na.rm = TRUE)
  ses   <- apply(norm_mat, 2, function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))))
  data.frame(
    modality = factor(modality_order, levels = modality_order),
    mean     = means[modality_order],
    se       = ses[modality_order],
    group    = group_label,
    valence  = valence_label,
    stringsAsFactors = FALSE
  )
}

valence_profiles <- bind_rows(
  build_valence_profile(strength.M.exc,  "M_downstream",  "excitatory"),
  build_valence_profile(strength.M.inh,  "M_downstream",  "inhibitory"),
  build_valence_profile(strength.F.exc,  "F_downstream",  "excitatory"),
  build_valence_profile(strength.F.inh,  "F_downstream",  "inhibitory"),
  build_valence_profile(strength.IR.exc, "IR_downstream", "excitatory"),
  build_valence_profile(strength.IR.inh, "IR_downstream", "inhibitory")
)
```

### Valence-split grouped bar chart

```{r valence-grouped-bar, fig.width=14, fig.height=5.5}
p7 <- ggplot(valence_profiles,
             aes(x = modality, y = mean, fill = group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7,
           color = "white", linewidth = 0.3) +
  geom_errorbar(
    aes(ymin = pmax(mean - se, 0), ymax = mean + se),
    position = position_dodge(width = 0.8),
    width = 0.2, linewidth = 0.4
  ) +
  facet_wrap(~ valence, scales = "free_y") +
  scale_fill_manual(
    values = target_set_colors,
    labels = target_set_labels,
    name   = "Target set"
  ) +
  labs(
    x     = "Modality",
    y     = "Mean normalized strength",
    title = "Valence-split modality profiles: M vs. F vs. IR downstream"
  ) +
  theme_bw(base_size = 12) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    strip.text        = element_text(face = "bold", size = 12),
    legend.position   = "bottom",
    panel.grid.minor  = element_blank()
  )

p7
```

```{r save-plot7}
save_fig6_plot(p7, "panel_I_bella_valence_grouped_bar", width = 14, height = 5.5)
```

### Valence-split radar comparison

```{r valence-radar, fig.width=14, fig.height=5.5}
# ---------------------------------------------------------------------------
# One radar per valence, with M, F, IR overlaid
# ---------------------------------------------------------------------------
make_valence_comparison_radar <- function(valence_label, profiles_df) {
  sub_df <- profiles_df %>% filter(valence == valence_label)

  # Normalize to 0-1 within this valence
  vmax <- max(sub_df$mean, na.rm = TRUE)
  if (vmax == 0) vmax <- 1
  sub_df <- sub_df %>% mutate(value = mean / vmax)

  ggplot(sub_df, aes(x = modality, y = value, group = group,
                      fill = group, color = group)) +
    geom_polygon(alpha = 0.2, linewidth = 1) +
    geom_point(size = 2.5) +
    scale_fill_manual(values = target_set_colors, labels = target_set_labels) +
    scale_color_manual(values = target_set_colors, labels = target_set_labels) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.1)) +
    labs(title = paste0(str_to_title(valence_label), " inputs")) +
    theme_minimal(base_size = 11) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_blank(),
      axis.text.x       = element_text(size = 9, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold"),
      legend.position    = "bottom",
      legend.title       = element_blank()
    )
}

p8_exc <- make_valence_comparison_radar("excitatory", valence_profiles)
p8_inh <- make_valence_comparison_radar("inhibitory", valence_profiles)

p8 <- plot_grid(p8_exc, p8_inh, nrow = 1, labels = c("Exc", "Inh"),
                label_size = 12)
p8
```

```{r save-plot8}
save_fig6_plot(p8, "panel_I_bella_valence_radar", width = 14, height = 5.5)
```

### Valence-split entropy comparison

```{r valence-entropy, fig.width=10, fig.height=5.5}
# ---------------------------------------------------------------------------
# Entropy computed on excitatory and inhibitory matrices separately
# ---------------------------------------------------------------------------
ent_valence <- bind_rows(
  data.frame(
    neuron  = rownames(strength.M.exc),
    entropy = apply(strength.M.exc, 1, shannon_entropy),
    group   = "M_downstream", valence = "excitatory", stringsAsFactors = FALSE
  ),
  data.frame(
    neuron  = rownames(strength.M.inh),
    entropy = apply(strength.M.inh, 1, shannon_entropy),
    group   = "M_downstream", valence = "inhibitory", stringsAsFactors = FALSE
  ),
  data.frame(
    neuron  = rownames(strength.F.exc),
    entropy = apply(strength.F.exc, 1, shannon_entropy),
    group   = "F_downstream", valence = "excitatory", stringsAsFactors = FALSE
  ),
  data.frame(
    neuron  = rownames(strength.F.inh),
    entropy = apply(strength.F.inh, 1, shannon_entropy),
    group   = "F_downstream", valence = "inhibitory", stringsAsFactors = FALSE
  ),
  data.frame(
    neuron  = rownames(strength.IR.exc),
    entropy = apply(strength.IR.exc, 1, shannon_entropy),
    group   = "IR_downstream", valence = "excitatory", stringsAsFactors = FALSE
  ),
  data.frame(
    neuron  = rownames(strength.IR.inh),
    entropy = apply(strength.IR.inh, 1, shannon_entropy),
    group   = "IR_downstream", valence = "inhibitory", stringsAsFactors = FALSE
  )
) %>%
  mutate(group_label = factor(target_set_labels[group],
                               levels = c("M cells", "F cells", "IR cells")))

p9 <- ggplot(ent_valence, aes(x = group_label, y = entropy, fill = group)) +
  geom_boxplot(width = 0.5, outlier.shape = 21, outlier.size = 2, alpha = 0.7) +
  geom_jitter(width = 0.12, alpha = 0.4, size = 1.2, stroke = 0, color = "grey30") +
  facet_wrap(~ valence) +
  scale_fill_manual(values = target_set_colors, guide = "none") +
  labs(
    x     = "Target set",
    y     = "Shannon entropy (bits)",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

# Pairwise tests per facet
for (v in c("excitatory", "inhibitory")) {
  message(sprintf("--- %s ---", v))
  for (pair in pairs) {
    sub <- ent_valence %>% filter(valence == v, group %in% pair)
    if (all(table(sub$group) >= 2)) {
      wt <- wilcox.test(entropy ~ group, data = sub)
      message(sprintf("  %s vs %s: Wilcoxon p = %.3g",
                      target_set_labels[pair[1]],
                      target_set_labels[pair[2]], wt$p.value))
    }
  }
}

p9
```

```{r save-plot9}
save_fig6_plot(p9, "panel_I_bella_valence_entropy", width = 10, height = 5.5)
```

### Valence-split modality scatter (pairwise)

```{r valence-scatter, fig.width=14, fig.height=9}
# ---------------------------------------------------------------------------
# Scatter: pairwise mean per modality, split by valence
# ---------------------------------------------------------------------------
scatter_valence <- bind_rows(
  # Excitatory
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.M.exc, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.F.exc, na.rm = TRUE)[modality_order],
    comparison = "M vs F",
    valence    = "excitatory",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.M.exc, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.IR.exc, na.rm = TRUE)[modality_order],
    comparison = "M vs IR",
    valence    = "excitatory",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.F.exc, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.IR.exc, na.rm = TRUE)[modality_order],
    comparison = "F vs IR",
    valence    = "excitatory",
    stringsAsFactors = FALSE
  ),
  # Inhibitory
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.M.inh, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.F.inh, na.rm = TRUE)[modality_order],
    comparison = "M vs F",
    valence    = "inhibitory",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.M.inh, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.IR.inh, na.rm = TRUE)[modality_order],
    comparison = "M vs IR",
    valence    = "inhibitory",
    stringsAsFactors = FALSE
  ),
  data.frame(
    modality   = modality_order,
    x_mean     = colMeans(strength.F.inh, na.rm = TRUE)[modality_order],
    y_mean     = colMeans(strength.IR.inh, na.rm = TRUE)[modality_order],
    comparison = "F vs IR",
    valence    = "inhibitory",
    stringsAsFactors = FALSE
  )
) %>%
  mutate(
    comparison = factor(comparison, levels = c("M vs F", "M vs IR", "F vs IR")),
    valence    = factor(valence, levels = c("excitatory", "inhibitory"))
  )

p10 <- ggplot(scatter_valence,
              aes(x = x_mean, y = y_mean, color = modality, label = modality)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey75", linewidth = 0.3) +
  geom_point(alpha = 0.65, size = 2, shape = 16) +
  geom_text(nudge_x = 0.02, nudge_y = 0.02,
            size = 2, fontface = "bold", show.legend = FALSE) +
  facet_grid(valence ~ comparison, scales = "free") +
  scale_color_manual(values = modality_colors, name = "Modality") +
  coord_fixed() +
  labs(
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    strip.text = element_text(face = "plain", size = 10, color = "grey30"),
    legend.position = "right"
  )

p10
```

```{r save-plot10}
save_fig6_plot(p10, "panel_I_bella_valence_scatter", width = 14, height = 9)
```

## Summary figure assembly

Combine the key plots into a single composite panel.

```{r composite, fig.width=16, fig.height=20}
# ---------------------------------------------------------------------------
# Assemble composite: 4 rows
#   Row 1: grouped bar + radar side-by-side
#   Row 2: tornado + entropy + scatter (1 pair shown)
#   Row 3: valence-split bar + valence radar
#   Row 4: valence entropy + valence scatter
# ---------------------------------------------------------------------------
row1 <- plot_grid(p1, p3, nrow = 1, rel_widths = c(1.2, 1),
                  labels = c("A", "B"), label_size = 14)

row2 <- plot_grid(p2, p5, nrow = 1, rel_widths = c(1.2, 1),
                  labels = c("C", "D"), label_size = 14)

row3 <- plot_grid(p7, p8, nrow = 1, rel_widths = c(1.2, 1),
                  labels = c("E", "F"), label_size = 14)

row4 <- plot_grid(p9, p10, nrow = 1, rel_widths = c(0.8, 1.2),
                  labels = c("G", "H"), label_size = 14)

panel_I <- plot_grid(row1, row2, row3, row4, ncol = 1,
                     rel_heights = c(1, 1.2, 1, 1.2))
panel_I
```

```{r save-composite}
save_fig6_plot(panel_I, "panel_I_bella_composite", width = 16, height = 20)
```
