---
title: "Figure 6H (Bella's targets): Spotlight Examples -- Individual Integration Nodes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel H: Spotlight Integration Nodes (Bella's targets) — Individual radar/spider plots for top multimodal integration neurons, with excitatory/inhibitory overlays and stacked bar breakdowns."
panel_legend <- "(H) Spotlight profiles of top multimodal integration nodes for Bella's targets. Radar plots show normalized modality input with excitatory/inhibitory overlays; stacked bars show absolute and proportional composition."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_H_description", "txt"), "panel_H_description.txt"))
```

## Overview

Panel H spotlights 3-4 specific neurons that serve as top multimodal
integration nodes among the paths reaching Bella's target populations
(M cells, F cells, IR cells). Each neuron receives input from multiple
sensory modalities, and we visualize its input profile using radar/spider
plots and stacked bar charts broken down by excitatory vs. inhibitory
valence.

## Step 1: Compute entropy and rank neurons

Shannon entropy across the 7 modality channels quantifies how evenly a
neuron integrates across modalities. Higher entropy = more multimodal.

```{r compute-entropy}
# ---------------------------------------------------------------------------
# Compute Shannon entropy for each neuron (row) in the combined strength matrix
# ---------------------------------------------------------------------------
entropy_vals <- apply(strength.bella, 1, shannon_entropy)

entropy_df <- data.frame(
  neuron    = names(entropy_vals),
  entropy   = entropy_vals,
  n_modalities = apply(strength.bella > 0, 1, sum),
  total_strength = rowSums(strength.bella),
  stringsAsFactors = FALSE
) %>%
  arrange(desc(entropy))

# Show the top 20 most multimodal neurons
head(entropy_df, 20)
```

## Step 2: Identify spotlight neurons

Select the top-entropy neurons. Also look for known neurons from Bella's
target populations (mAL_m1, AVLP494m, P1_3c, LH004m, etc.) to provide
biologically meaningful examples.

```{r select-spotlight}
# ---------------------------------------------------------------------------
# Inspect what neuron names exist in the strength matrix
# ---------------------------------------------------------------------------
all_neurons <- rownames(strength.bella)
message("Total neurons in combined matrix: ", length(all_neurons))

# Check for known neuron types from Bella's targets
known_patterns <- c("mAL", "AVLP494", "pC1_3", "LH004",
                     "SMP", "SIP", "SLP", "LH", "pC1", "pC2")
known_hits <- lapply(known_patterns, function(pat) {
  grep(pat, all_neurons, value = TRUE, ignore.case = TRUE)
})
names(known_hits) <- known_patterns

# Print matches for each pattern
for (pat in known_patterns) {
  if (length(known_hits[[pat]]) > 0) {
    message(sprintf("  Pattern '%s': %s", pat,
                    paste(known_hits[[pat]], collapse = ", ")))
  }
}

# ---------------------------------------------------------------------------
# Build a candidate list: top-entropy neurons + known types if present
# ---------------------------------------------------------------------------
# Start with top 6 by entropy
top_entropy <- entropy_df %>%
  filter(n_modalities >= 3) %>%     # require input from at least 3 modalities
  head(6) %>%
  pull(neuron)

# Add known neurons if they are present and have reasonable entropy
known_candidates <- c(
  grep("mAL", all_neurons, value = TRUE, ignore.case = TRUE),
  grep("AVLP494", all_neurons, value = TRUE, ignore.case = TRUE),
  grep("pC1_3", all_neurons, value = TRUE, ignore.case = TRUE),
  grep("LH004", all_neurons, value = TRUE, ignore.case = TRUE)
)
known_candidates <- intersect(known_candidates, entropy_df$neuron)

# Combine and deduplicate, then take 3-4
spotlight_candidates <- unique(c(top_entropy, known_candidates))

# Final selection: pick top 4 by entropy from the candidate pool
spotlight_neurons <- entropy_df %>%
  filter(neuron %in% spotlight_candidates) %>%
  arrange(desc(entropy)) %>%
  head(4) %>%
  pull(neuron)

message("Spotlight neurons: ", paste(spotlight_neurons, collapse = ", "))

# Look up synonyms/descriptions from mba for annotation
spotlight_info <- mba %>%
  filter(type %in% spotlight_neurons) %>%
  select(type, synonyms) %>%
  distinct(type, .keep_all = TRUE)

spotlight_info
```

## Step 3: Prepare per-neuron data for plotting

Extract strength values (total, excitatory, inhibitory) for each
spotlight neuron across all 7 modalities.

```{r prepare-neuron-data}
# ---------------------------------------------------------------------------
# Build a long-format dataframe for spotlight neurons
# ---------------------------------------------------------------------------

# Total strength
spotlight_total <- as.data.frame(strength.bella[spotlight_neurons, , drop = FALSE]) %>%
  rownames_to_column("neuron") %>%
  pivot_longer(cols = all_of(modality_order), names_to = "modality",
               values_to = "strength") %>%
  mutate(valence = "total")

# Excitatory strength (handle neurons that may be absent from exc/inh matrices)
exc_neurons <- intersect(spotlight_neurons, rownames(strength.bella.exc))
inh_neurons <- intersect(spotlight_neurons, rownames(strength.bella.inh))

spotlight_exc <- if (length(exc_neurons) > 0) {
  as.data.frame(strength.bella.exc[exc_neurons, , drop = FALSE]) %>%
    rownames_to_column("neuron") %>%
    pivot_longer(cols = all_of(modality_order), names_to = "modality",
                 values_to = "strength") %>%
    mutate(valence = "excitatory")
} else {
  data.frame(neuron = character(), modality = character(),
             strength = numeric(), valence = character())
}

spotlight_inh <- if (length(inh_neurons) > 0) {
  as.data.frame(strength.bella.inh[inh_neurons, , drop = FALSE]) %>%
    rownames_to_column("neuron") %>%
    pivot_longer(cols = all_of(modality_order), names_to = "modality",
                 values_to = "strength") %>%
    mutate(valence = "inhibitory")
} else {
  data.frame(neuron = character(), modality = character(),
             strength = numeric(), valence = character())
}

# Combine all
spotlight_long <- bind_rows(spotlight_total, spotlight_exc, spotlight_inh) %>%
  mutate(modality = factor(modality, levels = modality_order))
```

## Step 4: Individual radar plots (total strength)

One radar per spotlight neuron showing total input strength across the 7
modalities.

```{r radar-individual, fig.width=14, fig.height=4}
# ---------------------------------------------------------------------------
# Build a custom radar plot for each spotlight neuron
# ---------------------------------------------------------------------------
make_neuron_radar <- function(neuron_name, strength_mat, fill_color = "steelblue",
                               fill_alpha = 0.3, line_color = "steelblue") {
  vals <- strength_mat[neuron_name, modality_order]
  # Normalize to 0-1 for comparable radar shape
  max_val <- max(vals, na.rm = TRUE)
  if (max_val == 0) max_val <- 1
  vals_norm <- vals / max_val

  radar_df <- data.frame(
    modality = factor(modality_order, levels = modality_order),
    value    = as.numeric(vals_norm),
    raw      = as.numeric(vals)
  )

  # Close the polygon by appending the first point
  radar_closed <- bind_rows(radar_df, radar_df[1, ])

  # Look up synonym if available
  syn <- mba %>%
    filter(type == neuron_name) %>%
    pull(synonyms) %>%
    unique() %>%
    na.omit()
  subtitle_text <- if (length(syn) > 0 && nchar(syn[1]) > 0) {
    paste0("(", substr(syn[1], 1, 60), ")")
  } else {
    ""
  }

  # Entropy for this neuron
  ent <- entropy_df %>% filter(neuron == neuron_name) %>% pull(entropy)
  ent_label <- sprintf("H = %.2f", ent)

  # Determine which target group(s) this neuron belongs to
  group_info <- bella_groups_df %>% filter(type == neuron_name)
  group_label <- if (nrow(group_info) > 0) {
    paste(target_set_labels[group_info$target_group], collapse = ", ")
  } else {
    ""
  }

  ggplot(radar_df, aes(x = modality, y = value, group = 1)) +
    geom_polygon(fill = alpha(fill_color, fill_alpha), color = line_color,
                 linewidth = 0.8) +
    geom_point(aes(color = modality), size = 3, show.legend = FALSE) +
    scale_color_manual(values = modality_colors) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05), breaks = c(0.25, 0.5, 0.75, 1.0),
                       labels = c("25%", "50%", "75%", "100%")) +
    labs(
      title    = neuron_name,
      subtitle = paste(group_label, ent_label, subtitle_text)
    ) +
    theme_minimal(base_size = 11) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_text(size = 7, color = "grey50"),
      axis.text.x       = element_text(size = 9, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 12),
      plot.subtitle      = element_text(hjust = 0.5, size = 8, color = "grey40")
    )
}

# Generate individual radar plots
radar_plots <- lapply(spotlight_neurons, function(n) {
  make_neuron_radar(n, strength.bella)
})
names(radar_plots) <- spotlight_neurons

# Arrange in a single row
radar_combined <- plot_grid(
  plotlist = radar_plots,
  nrow = 1,
  labels = LETTERS[1:length(radar_plots)],
  label_size = 12
)
radar_combined
```

```{r save-radar-individual}
save_fig6_plot(radar_combined, "panel_H_bella_radar_individual",
               width = 3.5 * length(spotlight_neurons), height = 4)
```

## Step 5: Excitatory vs. inhibitory overlaid radars

For each spotlight neuron, overlay excitatory (solid) and inhibitory
(dashed) radar polygons on the same chart.

```{r radar-valence, fig.width=14, fig.height=4.5}
# ---------------------------------------------------------------------------
# Dual-valence radar: excitatory (solid) + inhibitory (dashed) overlaid
# ---------------------------------------------------------------------------
make_valence_radar <- function(neuron_name) {
  # Get excitatory values
  exc_vals <- if (neuron_name %in% rownames(strength.bella.exc)) {
    strength.bella.exc[neuron_name, modality_order]
  } else {
    setNames(rep(0, 7), modality_order)
  }

  # Get inhibitory values
  inh_vals <- if (neuron_name %in% rownames(strength.bella.inh)) {
    strength.bella.inh[neuron_name, modality_order]
  } else {
    setNames(rep(0, 7), modality_order)
  }

  # Normalize both to the same scale (max across both)
  global_max <- max(c(exc_vals, inh_vals), na.rm = TRUE)
  if (global_max == 0) global_max <- 1

  radar_df <- data.frame(
    modality = factor(rep(modality_order, 2), levels = modality_order),
    value    = c(as.numeric(exc_vals) / global_max,
                 as.numeric(inh_vals) / global_max),
    valence  = rep(c("excitatory", "inhibitory"), each = 7)
  )

  ggplot(radar_df, aes(x = modality, y = value, group = valence)) +
    geom_polygon(aes(fill = valence, linetype = valence, color = valence),
                 alpha = 0.2, linewidth = 0.9) +
    geom_point(aes(color = valence), size = 2) +
    scale_fill_manual(values = c("excitatory" = "#2166AC",
                                  "inhibitory" = "#B2182B")) +
    scale_color_manual(values = c("excitatory" = "#2166AC",
                                   "inhibitory" = "#B2182B")) +
    scale_linetype_manual(values = c("excitatory" = "solid",
                                      "inhibitory" = "dashed")) +
    coord_polar(start = 0) +
    scale_y_continuous(limits = c(0, 1.05)) +
    labs(title = neuron_name) +
    theme_minimal(base_size = 10) +
    theme(
      axis.title        = element_blank(),
      axis.text.y       = element_blank(),
      axis.text.x       = element_text(size = 8, face = "bold"),
      panel.grid.major  = element_line(color = "grey85", linewidth = 0.3),
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 11),
      legend.position    = "bottom",
      legend.title       = element_blank(),
      legend.text        = element_text(size = 8)
    )
}

valence_radars <- lapply(spotlight_neurons, make_valence_radar)
names(valence_radars) <- spotlight_neurons

valence_combined <- plot_grid(
  plotlist = valence_radars,
  nrow = 1,
  labels = LETTERS[1:length(valence_radars)],
  label_size = 12
)
valence_combined
```

```{r save-radar-valence}
save_fig6_plot(valence_combined, "panel_H_bella_radar_valence",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 6: Input breakdown stacked bar (excitatory/inhibitory by modality)

For each spotlight neuron, a stacked bar chart showing the proportion of
each modality split by valence.

```{r stacked-bar, fig.width=14, fig.height=5}
# ---------------------------------------------------------------------------
# Build stacked bar data for spotlight neurons
# ---------------------------------------------------------------------------
bar_data <- spotlight_long %>%
  filter(valence %in% c("excitatory", "inhibitory"),
         neuron %in% spotlight_neurons) %>%
  # Compute proportion within each neuron
  group_by(neuron) %>%
  mutate(
    total_all = sum(strength, na.rm = TRUE),
    proportion = ifelse(total_all > 0, strength / total_all, 0)
  ) %>%
  ungroup() %>%
  # Add modality group for ordering
  mutate(
    mod_group = modality_group[as.character(modality)],
    modality  = factor(modality, levels = modality_order)
  )

# ---------------------------------------------------------------------------
# Stacked bar: faceted by neuron, x = modality, fill = valence
# ---------------------------------------------------------------------------
p_stacked <- ggplot(bar_data,
                     aes(x = modality, y = strength, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ neuron, scales = "free_y", nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC",
                                "inhibitory" = "#B2182B"),
                     name = "Valence") +
  labs(
    x     = "Modality",
    y     = "Summed path strength",
    title = "Input breakdown by modality and valence (Bella's targets)"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 10),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )

p_stacked
```

```{r save-stacked-bar}
save_fig6_plot(p_stacked, "panel_H_bella_stacked_bar",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 6b: Proportional stacked bar (normalized to 100%)

Same breakdown but normalized so each neuron sums to 100%, making the
modality proportions directly comparable.

```{r stacked-bar-proportional, fig.width=14, fig.height=5}
p_stacked_prop <- ggplot(bar_data,
                          aes(x = modality, y = proportion, fill = valence)) +
  geom_col(position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  facet_wrap(~ neuron, nrow = 1) +
  scale_fill_manual(values = c("excitatory" = "#2166AC",
                                "inhibitory" = "#B2182B"),
                     name = "Valence") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x     = "Modality",
    y     = "Proportion of total input",
    title = "Proportional input breakdown by modality and valence (Bella's targets)"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    strip.text       = element_text(face = "bold", size = 10),
    legend.position  = "bottom",
    panel.grid.minor = element_blank()
  )

p_stacked_prop
```

```{r save-stacked-prop}
save_fig6_plot(p_stacked_prop, "panel_H_bella_stacked_bar_proportional",
               width = 3.5 * length(spotlight_neurons), height = 5)
```

## Step 7: Neuron annotation summary table

Print a summary table for the spotlight neurons with their entropy,
number of modalities, total strength, and synonyms.

```{r annotation-table}
# ---------------------------------------------------------------------------
# Build annotation table
# ---------------------------------------------------------------------------

# Also note target group membership
group_membership <- bella_groups_df %>%
  filter(type %in% spotlight_neurons) %>%
  group_by(type) %>%
  summarise(target_groups = paste(target_set_labels[target_group], collapse = ", "),
            .groups = "drop")

annotation_table <- entropy_df %>%
  filter(neuron %in% spotlight_neurons) %>%
  left_join(
    mba %>%
      select(type, synonyms) %>%
      distinct(type, .keep_all = TRUE),
    by = c("neuron" = "type")
  ) %>%
  left_join(group_membership, by = c("neuron" = "type")) %>%
  mutate(
    synonyms = ifelse(is.na(synonyms) | synonyms == "", "\u2014", synonyms),
    target_groups = ifelse(is.na(target_groups), "\u2014", target_groups),
    # Truncate long synonym strings
    synonyms = ifelse(nchar(synonyms) > 80,
                       paste0(substr(synonyms, 1, 77), "..."),
                       synonyms)
  ) %>%
  arrange(desc(entropy))

knitr::kable(annotation_table, digits = 3,
             caption = "Spotlight neuron summary (Bella's targets)",
             col.names = c("Neuron type", "Entropy (bits)",
                           "N modalities", "Total strength",
                           "Synonyms", "Target group(s)"))
```

## Step 8: Combined final panel

Assemble the main panel figure: radars on top, stacked bars on bottom.

```{r combined-panel, fig.width=14, fig.height=9}
panel_H <- plot_grid(
  radar_combined,
  valence_combined,
  p_stacked,
  ncol    = 1,
  rel_heights = c(1, 1, 1),
  labels  = c("i", "ii", "iii"),
  label_size = 14
)
panel_H
```

```{r save-combined}
save_fig6_plot(panel_H, "panel_H_bella_combined", width = 14, height = 12)
```
