---
title: "Figure 6A (Bella's targets): Multimodal Circuit Schematic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 12,
  fig.height = 7
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel A: Circuit Schematic (Bella's targets) — Programmatic diagram showing 7 sensory modalities converging onto downstream targets of M cells, F cells, and IR-expressing cells. Arrows indicate connectivity flow across circuit layers."
panel_legend <- "(A) Schematic of the multimodal sensory circuit targeting neurons downstream of M cells, F cells, and IR-expressing cells. Seven sensory modalities converge through intermediate relay neurons onto Bella's target neuron populations."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_A_description", "txt"), "panel_A_description.txt"))
```

## Overview

Panel A is a programmatic circuit schematic showing 7 sensory modalities
converging through pathways to Bella's third-order target neuron groups:
M cells downstream, F cells downstream, and IR cells downstream. The layout
has two columns representing circuit layers (sensory inputs on the left,
target groups on the right), with arrows indicating direct connectivity flow.
Representative neuron type names are annotated beneath each target group box.

## Build node layout

Define positions for all nodes across two columns: sensory inputs (left)
and target groups (right).

```{r node-layout}
# ---------------------------------------------------------------------------
# Column x-positions for the two circuit layers
# ---------------------------------------------------------------------------
x_sensory <- 0
x_target  <- 8

# ---------------------------------------------------------------------------
# Sensory input nodes (left column)
# Ordered top-to-bottom: DA1, VA1v, VA1d, JO-B, LC10, ppk23, ppk25
# ---------------------------------------------------------------------------
sensory_labels <- c("DA1", "VA1v", "VA1d", "JO-B", "LC10", "ppk23", "ppk25")
sensory_keys   <- c("DA1", "VA1v", "VA1d", "auditory", "visual", "ppk23", "ppk25")

n_sensory <- length(sensory_labels)
sensory_y <- seq(from = n_sensory, to = 1, length.out = n_sensory)

nodes_sensory <- data.frame(
  label = sensory_labels,
  key   = sensory_keys,
  x     = x_sensory,
  y     = sensory_y,
  layer = "sensory",
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Target group nodes (right column)
# Three boxes for M, F, IR downstream groups
# ---------------------------------------------------------------------------
nodes_target <- data.frame(
  label = c("M cells\ndownstream", "F cells\ndownstream", "IR cells\ndownstream"),
  key   = c("M_downstream", "F_downstream", "IR_downstream"),
  x     = x_target,
  y     = c(6.0, 4.0, 2.0),
  layer = "target",
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Representative neuron types for each target group (shown as sub-labels)
# ---------------------------------------------------------------------------
representative_types <- data.frame(
  label = c("mAL_m1, AVLP494m",
            "P1_3c, AVLP743m",
            "LH004m, DNd02"),
  key   = c("M_downstream", "F_downstream", "IR_downstream"),
  x     = x_target,
  y     = c(6.0 - 0.65, 4.0 - 0.65, 2.0 - 0.65),
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Combine all nodes
# ---------------------------------------------------------------------------
nodes <- bind_rows(nodes_sensory, nodes_target)
```

## Build edges

Connect every sensory input directly to every target group, reflecting the
multimodal convergence topology (no relay intermediaries).

```{r edges}
# ---------------------------------------------------------------------------
# Sensory -> target edges (all-to-all)
# ---------------------------------------------------------------------------
edges_s2t <- expand.grid(
  from = sensory_labels,
  to   = nodes_target$label,
  stringsAsFactors = FALSE
)

# ---------------------------------------------------------------------------
# Look up coordinates for edge endpoints
# ---------------------------------------------------------------------------
node_lookup <- nodes %>% select(label, x, y)

add_coords <- function(edges_df) {
  edges_df %>%
    left_join(node_lookup, by = c("from" = "label")) %>%
    rename(x_from = x, y_from = y) %>%
    left_join(node_lookup, by = c("to" = "label")) %>%
    rename(x_to = x, y_to = y)
}

edges_all <- add_coords(edges_s2t)
```

## Build group annotations

Label the sensory modality groups on the far left of the diagram.

```{r group-annotations}
# ---------------------------------------------------------------------------
# Map each sensory node to its group using modality_group from setup.R
# ---------------------------------------------------------------------------
group_mapping <- data.frame(
  label = sensory_labels,
  key   = sensory_keys,
  stringsAsFactors = FALSE
) %>%
  mutate(group = modality_group[key])

# Compute y-position for each group label (mean y of its member nodes)
group_annot <- group_mapping %>%
  left_join(nodes_sensory %>% select(label, y), by = "label") %>%
  group_by(group) %>%
  summarise(
    y_mid = mean(y),
    y_min = min(y) - 0.35,
    y_max = max(y) + 0.35,
    .groups = "drop"
  ) %>%
  mutate(
    x = x_sensory - 2.8,
    group_label = str_to_title(group)
  )
```

## Draw schematic

Assemble the diagram with boxes, arrows, labels, and group annotations.

```{r schematic, fig.width=12, fig.height=7}
# ---------------------------------------------------------------------------
# Box dimensions
# ---------------------------------------------------------------------------
box_hw_sensory <- 0.8    # half-width for sensory nodes
box_hh_sensory <- 0.35   # half-height for sensory nodes
box_hw_target  <- 1.3    # half-width for target group boxes (wider)
box_hh_target  <- 0.45   # half-height for target group boxes (taller)

# Nudge arrow start/end so they originate from box edges, not centres
arrow_nudge_x_left  <- box_hw_sensory + 0.05
arrow_nudge_x_right <- box_hw_target + 0.05

# Assign fill colours: sensory nodes use modality_colors, targets use target_set_colors
nodes <- nodes %>%
  mutate(
    fill_color = case_when(
      layer == "sensory" ~ modality_colors[key],
      layer == "target"  ~ target_set_colors[key],
      TRUE               ~ "#F0F0F0"
    ),
    text_color = case_when(
      layer == "target" ~ "white",
      TRUE              ~ "black"
    ),
    hw = ifelse(layer == "target", box_hw_target, box_hw_sensory),
    hh = ifelse(layer == "target", box_hh_target, box_hh_sensory)
  )

# ---------------------------------------------------------------------------
# Column (layer) labels
# ---------------------------------------------------------------------------
layer_labels <- data.frame(
  label = c("Sensory\ninputs", "3rd-order\ntargets"),
  x     = c(x_sensory, x_target),
  y     = max(sensory_y) + 1.0
)

# ---------------------------------------------------------------------------
# Main plot
# ---------------------------------------------------------------------------
p <- ggplot() +

  # -- Edges (arrows) -------------------------------------------------------
  geom_segment(
    data = edges_all,
    aes(
      x    = x_from + arrow_nudge_x_left,
      xend = x_to   - arrow_nudge_x_right,
      y    = y_from,
      yend = y_to
    ),
    arrow     = grid::arrow(length = unit(0.12, "cm"), type = "closed"),
    linewidth = 0.35,
    color     = "grey55",
    alpha     = 0.5
  ) +

  # -- Node boxes -----------------------------------------------------------
  geom_rect(
    data = nodes,
    aes(
      xmin = x - hw,
      xmax = x + hw,
      ymin = y - hh,
      ymax = y + hh
    ),
    fill  = nodes$fill_color,
    color = "grey30",
    linewidth = 0.4
  ) +

  # -- Node labels ----------------------------------------------------------
  geom_text(
    data = nodes,
    aes(x = x, y = y, label = label),
    color    = nodes$text_color,
    size     = 3.5,
    fontface = "bold",
    lineheight = 0.85
  ) +

  # -- Representative neuron type names below target boxes -------------------
  geom_text(
    data = representative_types,
    aes(x = x, y = y, label = label),
    size     = 2.5,
    fontface = "italic",
    color    = "grey30"
  ) +

  # -- Group bracket lines (left side) --------------------------------------
  geom_segment(
    data = group_annot,
    aes(
      x    = x + 0.8,
      xend = x + 0.8,
      y    = y_min,
      yend = y_max
    ),
    linewidth = 0.6,
    color     = modality_group_colors[group_annot$group]
  ) +

  # -- Group labels ---------------------------------------------------------
  geom_text(
    data = group_annot,
    aes(x = x + 0.35, y = y_mid, label = group_label),
    size     = 3.2,
    fontface = "italic",
    hjust    = 0.5,
    color    = modality_group_colors[group_annot$group]
  ) +

  # -- Layer (column) headers -----------------------------------------------
  geom_text(
    data = layer_labels,
    aes(x = x, y = y, label = label),
    size     = 3.8,
    fontface = "bold",
    color    = "grey25",
    lineheight = 0.85
  ) +

  # -- Theme ----------------------------------------------------------------
  coord_fixed(ratio = 0.75, clip = "off") +
  xlim(x_sensory - 3.5, x_target + 2.0) +
  ylim(0.2, max(sensory_y) + 1.5) +
  theme_void() +
  theme(
    plot.margin = margin(10, 10, 10, 10, "pt"),
    plot.title  = element_text(hjust = 0.5, face = "bold", size = 14)
  ) +
  ggtitle("Multimodal sensory convergence onto third-order courtship neurons")

p
```

## Save

```{r save}
save_fig6_plot(p, "panel_A_circuit_schematic_bella", width = 12, height = 7)
```
