---
title: "Figure 6D (Bella's targets): Lifetime Sparseness / Multimodality Score"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/figures/fig6_bella_targets/setup.R")
```

```{r lifetime-sparseness-function}
lifetime_sparseness <- function(x) {
  x <- pmax(x, 0)  # set negative values to zero
  N <- length(x)
  if (N <= 1) return(0)
  mean_r <- mean(x)
  mean_r2 <- mean(x^2)
  if (mean_r2 == 0) return(0)
  (N / (N - 1)) * (1 - mean_r^2 / mean_r2)
}
```

```{r modality-shapes}
modality_shapes <- c(
  "DA1" = 16,      # filled circle
  "VA1v" = 15,     # filled square
  "VA1d" = 17,     # filled triangle up
  "auditory" = 18, # filled diamond
  "visual" = 3,    # plus
  "ppk23" = 4,     # cross
  "ppk25" = 8      # asterisk/star
)
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel D: Lifetime Sparseness (Bella's targets) — Lifetime sparseness (Rolls & Tovee 1995) over the 7-modality strength distribution quantifies how selective each target neuron is. Values near 1 indicate responses dominated by a single modality; values near 0 indicate balanced multimodal input. Scatter plots, boxplots, and valence-split analyses for M, F, and IR downstream target groups."
panel_legend <- "(D) Lifetime sparseness over the 7-modality strength distribution quantifies selectivity for each of Bella's target neurons. Scatter plots, boxplots, and valence-split analyses are shown for M-downstream, F-downstream, and IR-downstream groups."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_D_description", "txt"), "panel_D_description.txt"))
```

## Overview

This panel quantifies how multimodal each target neuron is using lifetime sparseness
(Rolls & Tovee 1995, Vinje & Gallant 2000) over the distribution of incoming path
strengths across the 7 sensory modalities.
A neuron driven by a single modality has sparseness 1 (maximally sparse); a neuron
receiving perfectly balanced input from all modalities has sparseness near 0.

We compute entropy, dominant modality, modality breadth (number of non-zero inputs),
and dominance (fraction of total strength from the dominant modality) for every
target neuron in the combined Bella target path set (M cells, F cells, IR cells).

## 1. Compute multimodality metrics

```{r compute-entropy-metrics}
# Maximum possible entropy for 7 modalities
max_entropy <- log2(length(modality_order))  # log2(7) ~ 2.807

# Build a tidy summary table from the raw strength matrix
# Each row = one target neuron; columns = per-modality strengths
entropy_df <- as.data.frame(strength.bella) %>%

tibble::rownames_to_column("neuron") %>%
  rowwise() %>%
  mutate(
    # Shannon entropy across the 7 modality columns
    entropy = shannon_entropy(c_across(all_of(modality_order))),

    # Dominant modality: which column carries the most strength
    dominant_modality = modality_order[which.max(c_across(all_of(modality_order)))],

    # Breadth: how many modalities have non-zero input
    breadth = sum(c_across(all_of(modality_order)) > 0),

    # Total strength across all modalities
    total_strength = sum(c_across(all_of(modality_order))),

    # Dominance: fraction of total from the dominant modality
    dominance = max(c_across(all_of(modality_order))) / total_strength,

    # Lifetime sparseness (Rolls & Tovee 1995)
    sparseness = lifetime_sparseness(c_across(all_of(modality_order)))
  ) %>%
  ungroup() %>%
  # Normalised entropy (0-1 scale)
  mutate(
    entropy_norm = entropy / max_entropy,
    eff_modalities = 2^entropy,
    dominance = ifelse(is.finite(dominance), dominance, NA_real_)
  )

# Assign target_set from paths.bella
M_neurons  <- unique(paths.M$end)
F_neurons  <- unique(paths.F$end)
IR_neurons <- unique(paths.IR$end)

entropy_df <- entropy_df %>%
  mutate(
    target_set = case_when(
      neuron %in% M_neurons  ~ "M_downstream",
      neuron %in% F_neurons  ~ "F_downstream",
      neuron %in% IR_neurons ~ "IR_downstream",
      TRUE ~ "other"
    )
  )

# Quick summary
message(sprintf("Total target neurons: %d", nrow(entropy_df)))
message(sprintf("Mean entropy: %.3f  (max possible: %.3f)", mean(entropy_df$entropy), max_entropy))
message(sprintf("Median breadth: %d modalities", median(entropy_df$breadth)))
```

```{r entropy-summary-table}
# Show the top 20 most multimodal neurons
entropy_df %>%
  arrange(desc(entropy)) %>%
  select(neuron, entropy, entropy_norm, dominant_modality, breadth, dominance, target_set) %>%
  head(20) %>%
  knitr::kable(digits = 3, caption = "Top 20 most multimodal target neurons (by entropy)")
```

## 2. Bar chart: neurons ranked by entropy

```{r bar-entropy-ranked, fig.width=14, fig.height=6}
# Order neurons by descending entropy
entropy_ranked <- entropy_df %>%
  arrange(desc(entropy)) %>%
  mutate(rank = row_number(),
         neuron_label = factor(neuron, levels = neuron))

p_bar <- ggplot(entropy_ranked, aes(x = neuron_label, y = sparseness, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  annotate("text", x = nrow(entropy_ranked) * 0.95, y = 1 + 0.02,
           label = "max = 1",
           hjust = 1, size = 3, color = "grey40") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Lifetime sparseness",
    title = "Multimodality score across Bella's target neurons"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

print(p_bar)
save_fig6_plot(p_bar, "panel_D_entropy_bar_bella", width = 14, height = 6, subdir = "combined/lifetime_sparseness")
```

```{r bar-entropy-ranked-effmod, fig.width=14, fig.height=6}
# Neurons ranked by entropy — effective modalities
p_bar_em <- ggplot(entropy_ranked, aes(x = neuron_label, y = eff_modalities, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  annotate("text", x = nrow(entropy_ranked) * 0.95, y = 7 + 0.05,
           label = "max = 7",
           hjust = 1, size = 3, color = "grey40") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(breaks = 1:7, limits = c(0, 7.5), expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Effective modalities",
    title = "Multimodality score across Bella's target neurons"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

print(p_bar_em)
save_fig6_plot(p_bar_em, "panel_D_entropy_bar_bella_effmod", width = 14, height = 6, subdir = "combined/effective_modalities")
```

## 3. Bar chart faceted by target set

```{r bar-entropy-faceted, fig.width=14, fig.height=10}
# Separate ranking within each target set (3 groups: M, F, IR)
entropy_faceted <- entropy_df %>%
  filter(target_set %in% names(target_set_labels)) %>%
  group_by(target_set) %>%
  arrange(desc(entropy), .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(target_set_label = target_set_labels[target_set])

p_bar_faceted <- ggplot(entropy_faceted, aes(x = reorder(neuron, -entropy), y = sparseness,
                                              fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  facet_wrap(~ target_set_label, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Lifetime sparseness",
    title = "Multimodality by target set (M cells vs F cells vs IR cells)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold", size = 12)
  )

print(p_bar_faceted)
save_fig6_plot(p_bar_faceted, "panel_D_entropy_bar_faceted_bella", width = 14, height = 10, subdir = "combined/lifetime_sparseness")
```

```{r bar-entropy-faceted-effmod, fig.width=14, fig.height=10}
# Faceted bar chart — effective modalities
p_bar_faceted_em <- ggplot(entropy_faceted, aes(x = reorder(neuron, -entropy), y = eff_modalities,
                                              fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  facet_wrap(~ target_set_label, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality") +
  scale_y_continuous(breaks = 1:7, limits = c(0, 7.5), expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Effective modalities",
    title = "Multimodality by target set (M cells vs F cells vs IR cells)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold", size = 12)
  )

print(p_bar_faceted_em)
save_fig6_plot(p_bar_faceted_em, "panel_D_entropy_bar_faceted_bella_effmod", width = 14, height = 10, subdir = "combined/effective_modalities")
```

```{r entropy-distribution-comparison, fig.width=8, fig.height=5}
# Statistical comparison of entropy between target sets
entropy_plot_df <- entropy_df %>%
  filter(target_set %in% names(target_set_labels)) %>%
  mutate(target_set_label = target_set_labels[target_set])

p_box <- ggplot(entropy_plot_df, aes(x = target_set_label, y = sparseness, fill = target_set)) +
  geom_boxplot(alpha = 0.35, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.2, stroke = 0) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = target_set_colors) +
  labs(
    x = "Target set",
    y = "Lifetime sparseness",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_box)
save_fig6_plot(p_box, "panel_D_entropy_boxplot_targetset_bella", width = 8, height = 5, subdir = "combined/lifetime_sparseness")
```

```{r entropy-distribution-comparison-effmod, fig.width=8, fig.height=5}
# Boxplot comparison of effective modalities between target sets
p_box_em <- ggplot(entropy_plot_df, aes(x = target_set_label, y = eff_modalities, fill = target_set)) +
  geom_boxplot(alpha = 0.35, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.2, stroke = 0) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = target_set_colors) +
  labs(
    x = "Target set",
    y = "Effective modalities",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_box_em)
save_fig6_plot(p_box_em, "panel_D_entropy_boxplot_targetset_bella_effmod", width = 8, height = 5, subdir = "combined/effective_modalities")
```

## 4. Scatter plot: total strength vs entropy

```{r scatter-strength-entropy, fig.width=10, fig.height=8}
# Identify neurons to label: those with extreme or interesting values
top_n_label <- 15
label_neurons <- entropy_df %>%
  arrange(desc(entropy)) %>%
  slice_head(n = top_n_label) %>%
  pull(neuron)

entropy_df <- entropy_df %>%
  mutate(show_label = neuron %in% label_neurons)

p_scatter <- ggplot(entropy_df %>% filter(target_set %in% names(target_set_colors)),
                    aes(x = total_strength, y = sparseness,
                        fill = dominant_modality, color = target_set)) +
  geom_point(size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    data = filter(entropy_df, show_label & target_set %in% names(target_set_colors)),
    aes(label = neuron),
    color = "grey35",
    size = 2, max.overlaps = 20, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = modality_colors, name = "Modality") +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.5)),
         color = guide_legend(override.aes = list(fill = "white", size = 3))) +
  labs(
    x = "Total path strength",
    y = "Lifetime sparseness",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_scatter)
save_fig6_plot(p_scatter, "panel_D_strength_vs_entropy_bella", width = 7, height = 5.5, subdir = "combined/lifetime_sparseness")
```

```{r scatter-strength-entropy-effmod, fig.width=10, fig.height=8}
p_scatter_em <- ggplot(entropy_df %>% filter(target_set %in% names(target_set_colors)),
                    aes(x = total_strength, y = eff_modalities,
                        fill = dominant_modality, color = target_set)) +
  geom_point(size = 2.2, shape = 21, stroke = 0.6) +
  ggrepel::geom_text_repel(
    data = filter(entropy_df, show_label & target_set %in% names(target_set_colors)),
    aes(label = neuron),
    color = "grey35",
    size = 2, max.overlaps = 20, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = modality_colors, name = "Modality") +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(fill = guide_legend(override.aes = list(color = "black", stroke = 0.5)),
         color = guide_legend(override.aes = list(fill = "white", size = 3))) +
  labs(
    x = "Total path strength",
    y = "Effective modalities",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_scatter_em)
save_fig6_plot(p_scatter_em, "panel_D_strength_vs_entropy_bella_effmod", width = 7, height = 5.5, subdir = "combined/effective_modalities")
```

## 5. Valence-split analysis (excitatory vs inhibitory)

```{r compute-entropy-by-valence}
# Helper to compute entropy metrics from a strength matrix
compute_entropy_from_matrix <- function(mat, label) {
  df <- as.data.frame(mat) %>%
    tibble::rownames_to_column("neuron") %>%
    rowwise() %>%
    mutate(
      entropy = shannon_entropy(c_across(all_of(intersect(modality_order, colnames(mat))))),
      dominant_modality = {
        vals <- c_across(all_of(intersect(modality_order, colnames(mat))))
        mod_names <- intersect(modality_order, colnames(mat))
        if (all(vals == 0)) NA_character_ else mod_names[which.max(vals)]
      },
      breadth = sum(c_across(all_of(intersect(modality_order, colnames(mat)))) > 0),
      total_strength = sum(c_across(all_of(intersect(modality_order, colnames(mat))))),
      dominance = {
        vals <- c_across(all_of(intersect(modality_order, colnames(mat))))
        tot <- sum(vals)
        if (tot > 0) max(vals) / tot else NA_real_
      },
      sparseness = lifetime_sparseness(c_across(all_of(intersect(modality_order, colnames(mat)))))
    ) %>%
    ungroup() %>%
    mutate(
      entropy_norm = entropy / max_entropy,
      eff_modalities = 2^entropy,
      valence = label
    )

  df
}

entropy_exc <- compute_entropy_from_matrix(strength.bella.exc, "excitatory")
entropy_inh <- compute_entropy_from_matrix(strength.bella.inh, "inhibitory")
entropy_combined <- compute_entropy_from_matrix(strength.bella, "combined")

entropy_valence <- bind_rows(entropy_exc, entropy_inh, entropy_combined)
```

```{r bar-entropy-by-valence, fig.width=14, fig.height=10}
# Ranked bar charts per valence
p_valence_bars <- entropy_valence %>%
  group_by(valence) %>%
  arrange(desc(entropy), .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = rank, y = sparseness, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.4) +
  facet_wrap(~ valence, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality", na.value = "grey80") +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Lifetime sparseness",
    title = "Multimodality score split by path valence (Bella's targets)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold", size = 12)
  )

print(p_valence_bars)
save_fig6_plot(p_valence_bars, "panel_D_entropy_bar_valence_bella", width = 14, height = 10, subdir = "valence/lifetime_sparseness")
```

```{r bar-entropy-by-valence-effmod, fig.width=14, fig.height=10}
# Ranked bar charts per valence — effective modalities
p_valence_bars_em <- entropy_valence %>%
  group_by(valence) %>%
  arrange(desc(entropy), .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = rank, y = eff_modalities, fill = dominant_modality)) +
  geom_col(width = 0.8) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey40", linewidth = 0.4) +
  facet_wrap(~ valence, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = modality_colors, name = "Dominant\nModality", na.value = "grey80") +
  scale_y_continuous(breaks = 1:7, limits = c(0, 7.5), expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Target neuron (ranked by entropy)",
    y = "Effective modalities",
    title = "Multimodality score split by path valence (Bella's targets)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold", size = 12)
  )

print(p_valence_bars_em)
save_fig6_plot(p_valence_bars_em, "panel_D_entropy_bar_valence_bella_effmod", width = 14, height = 10, subdir = "valence/effective_modalities")
```

```{r boxplot-entropy-by-valence, fig.width=8, fig.height=5}
# Compare entropy distributions across valence types
p_valence_box <- ggplot(entropy_valence, aes(x = valence, y = sparseness, fill = valence)) +
  geom_boxplot(alpha = 0.35, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.2, stroke = 0) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(
    values = c("combined" = "#B3B3B3", "excitatory" = "#4DAF4A", "inhibitory" = "#E41A1C")
  ) +
  labs(
    x = "Path valence",
    y = "Lifetime sparseness",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_valence_box)
save_fig6_plot(p_valence_box, "panel_D_entropy_boxplot_valence_bella", width = 8, height = 5, subdir = "valence/lifetime_sparseness")
```

```{r boxplot-entropy-by-valence-effmod, fig.width=8, fig.height=5}
# Compare effective modalities distributions across valence types
p_valence_box_em <- ggplot(entropy_valence, aes(x = valence, y = eff_modalities, fill = valence)) +
  geom_boxplot(alpha = 0.35, outlier.size = 1) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.2, stroke = 0) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(
    values = c("combined" = "#B3B3B3", "excitatory" = "#4DAF4A", "inhibitory" = "#E41A1C")
  ) +
  labs(
    x = "Path valence",
    y = "Effective modalities",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "none"
  )

print(p_valence_box_em)
save_fig6_plot(p_valence_box_em, "panel_D_entropy_boxplot_valence_bella_effmod", width = 8, height = 5, subdir = "valence/effective_modalities")
```

```{r scatter-dominance-entropy-valence, fig.width=14, fig.height=6}
# Dominance vs entropy, faceted by valence
p_scatter_valence <- entropy_valence %>%
  filter(is.finite(dominance)) %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = dominance, y = sparseness, shape = dominant_modality)) +
  geom_point(alpha = 0.65, size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  facet_wrap(~ valence) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  labs(
    x = "Dominance (top modality fraction)",
    y = "Lifetime sparseness",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

print(p_scatter_valence)
save_fig6_plot(p_scatter_valence, "panel_D_dominance_vs_entropy_valence_bella", width = 10, height = 4.5, subdir = "valence/lifetime_sparseness")
```

```{r scatter-dominance-entropy-valence-effmod, fig.width=14, fig.height=6}
# Dominance vs effective modalities, faceted by valence
p_scatter_valence_em <- entropy_valence %>%
  filter(is.finite(dominance)) %>%
  mutate(valence = factor(valence, levels = c("combined", "excitatory", "inhibitory"))) %>%
  ggplot(aes(x = dominance, y = eff_modalities, shape = dominant_modality)) +
  geom_point(alpha = 0.65, size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  facet_wrap(~ valence) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  labs(
    x = "Dominance (top modality fraction)",
    y = "Effective modalities",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right",
    strip.text = element_text(face = "plain", size = 10, color = "grey30")
  )

print(p_scatter_valence_em)
save_fig6_plot(p_scatter_valence_em, "panel_D_dominance_vs_entropy_valence_bella_effmod", width = 10, height = 4.5, subdir = "valence/effective_modalities")
```

## 6. Dominance vs entropy relationship

```{r dominance-vs-entropy, fig.width=8, fig.height=6}
# Dominance (fraction from top modality) vs entropy
# Expect strong negative correlation: high entropy = low dominance
p_dominance <- ggplot(entropy_df %>% filter(is.finite(dominance)),
                      aes(x = dominance, y = sparseness,
                          shape = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.5, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  labs(
    x = "Dominance (top modality fraction)",
    y = "Lifetime sparseness",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_dominance)
save_fig6_plot(p_dominance, "panel_D_dominance_vs_entropy_bella", width = 7, height = 5, subdir = "combined/lifetime_sparseness")
```

```{r dominance-vs-entropy-effmod, fig.width=8, fig.height=6}
# Dominance (fraction from top modality) vs effective modalities
p_dominance_em <- ggplot(entropy_df %>% filter(is.finite(dominance)),
                      aes(x = dominance, y = eff_modalities,
                          shape = dominant_modality, color = target_set)) +
  geom_point(alpha = 0.7, size = 2.5, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  labs(
    x = "Dominance (top modality fraction)",
    y = "Effective modalities",
    title = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.title = element_text(size = rel(0.85)),
    legend.text = element_text(size = rel(0.8)),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "right"
  )

print(p_dominance_em)
save_fig6_plot(p_dominance_em, "panel_D_dominance_vs_entropy_bella_effmod", width = 7, height = 5, subdir = "combined/effective_modalities")
```

```{r dominance-effmod-marginal-density-bella, fig.width=9, fig.height=8}
# Dominance vs lifetime sparseness — marginal densities on both axes
plot_dom <- entropy_df %>%
  filter(is.finite(dominance), target_set %in% names(target_set_colors))

p_dom_core <- ggplot(plot_dom, aes(x = dominance, y = sparseness)) +
  geom_point(aes(shape = dominant_modality, color = target_set),
             alpha = 0.7, size = 2.5, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(x = "Dominance (fraction from top modality)", y = "Lifetime sparseness") +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )

p_top_dens <- ggplot(plot_dom, aes(x = dominance, fill = target_set, color = target_set)) +
  geom_density(alpha = 0.12, linewidth = 0.4) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  theme_void() + theme(legend.position = "none")

p_right_dens <- ggplot(plot_dom, aes(x = sparseness, fill = target_set, color = target_set)) +
  geom_density(alpha = 0.12, linewidth = 0.4) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  coord_flip() + scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  theme_void() + theme(legend.position = "none")

p_dom_asm <- plot_grid(
  p_top_dens, NULL,
  p_dom_core + theme(legend.position = "none"), p_right_dens,
  ncol = 2, nrow = 2,
  rel_widths = c(4, 1), rel_heights = c(1, 4),
  align = "hv", axis = "tblr"
)
legend_dom <- get_legend(p_dom_core)
p_dom_density <- plot_grid(p_dom_asm, legend_dom, ncol = 1, rel_heights = c(1, 0.1))

print(p_dom_density)

pdir_png_dm <- panel_plot_dir("panel_D_dominance_marginal_bella", "png", subdir = "combined/lifetime_sparseness")
pdir_pdf_dm <- panel_plot_dir("panel_D_dominance_marginal_bella", "pdf", subdir = "combined/lifetime_sparseness")

ggsave(file.path(pdir_png_dm, "panel_D_dominance_marginal_bella_density.png"),
       p_dom_density, width = 9, height = 8, dpi = 300)
ggsave(file.path(pdir_pdf_dm, "panel_D_dominance_marginal_bella_density.pdf"),
       p_dom_density, width = 9, height = 8)
```

```{r dominance-effmod-marginal-boxplot-bella, fig.width=9, fig.height=8}
# Dominance vs lifetime sparseness — marginal boxplots on both axes, all points shown
p_top_box <- ggplot(plot_dom, aes(x = dominance, y = target_set, fill = target_set)) +
  geom_boxplot(alpha = 0.3, outlier.shape = NA, linewidth = 0.3) +
  geom_jitter(aes(color = target_set), height = 0.15, alpha = 0.5, size = 0.8, stroke = 0) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  theme_void() + theme(legend.position = "none")

p_right_box <- ggplot(plot_dom, aes(x = target_set, y = sparseness, fill = target_set)) +
  geom_boxplot(alpha = 0.3, outlier.shape = NA, linewidth = 0.3) +
  geom_jitter(aes(color = target_set), width = 0.15, alpha = 0.5, size = 0.8, stroke = 0) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  theme_void() + theme(legend.position = "none")

p_dom_asm_b <- plot_grid(
  p_top_box, NULL,
  p_dom_core + theme(legend.position = "none"), p_right_box,
  ncol = 2, nrow = 2,
  rel_widths = c(4, 1), rel_heights = c(1, 4),
  align = "hv", axis = "tblr"
)
p_dom_boxplot <- plot_grid(p_dom_asm_b, legend_dom, ncol = 1, rel_heights = c(1, 0.1))

print(p_dom_boxplot)

ggsave(file.path(pdir_png_dm, "panel_D_dominance_marginal_bella_boxplot.png"),
       p_dom_boxplot, width = 9, height = 8, dpi = 300)
ggsave(file.path(pdir_pdf_dm, "panel_D_dominance_marginal_bella_boxplot.pdf"),
       p_dom_boxplot, width = 9, height = 8)
```

```{r dominance-effmod-marginal-density-bella-effmod, fig.width=9, fig.height=8}
# Dominance vs effective modalities — marginal densities on both axes
p_dom_core_em <- ggplot(plot_dom, aes(x = dominance, y = eff_modalities)) +
  geom_point(aes(shape = dominant_modality, color = target_set),
             alpha = 0.7, size = 2.5, stroke = 0.6) +
  ggrepel::geom_text_repel(
    aes(label = neuron),
    color = "grey35",
    size = 1.5, max.overlaps = 30, segment.color = "grey75", segment.size = 0.25
  ) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_shape_manual(values = modality_shapes, name = "Dominant\nModality", na.value = 1) +
  scale_color_manual(values = target_set_colors, labels = target_set_labels, name = "Target set") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  scale_y_continuous(breaks = 1:7) +
  labs(x = "Dominance (fraction from top modality)", y = "Effective modalities") +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )

p_top_dens_em <- ggplot(plot_dom, aes(x = dominance, fill = target_set, color = target_set)) +
  geom_density(alpha = 0.12, linewidth = 0.4) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  theme_void() + theme(legend.position = "none")

p_right_dens_em <- ggplot(plot_dom, aes(x = eff_modalities, fill = target_set, color = target_set)) +
  geom_density(alpha = 0.12, linewidth = 0.4) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  coord_flip() + scale_x_continuous(breaks = 1:7) +
  theme_void() + theme(legend.position = "none")

p_dom_asm_em <- plot_grid(
  p_top_dens_em, NULL,
  p_dom_core_em + theme(legend.position = "none"), p_right_dens_em,
  ncol = 2, nrow = 2,
  rel_widths = c(4, 1), rel_heights = c(1, 4),
  align = "hv", axis = "tblr"
)
legend_dom_em <- get_legend(p_dom_core_em)
p_dom_density_em <- plot_grid(p_dom_asm_em, legend_dom_em, ncol = 1, rel_heights = c(1, 0.1))

print(p_dom_density_em)

pdir_png_dm_em <- panel_plot_dir("panel_D_dominance_marginal_bella", "png", subdir = "combined/effective_modalities")
pdir_pdf_dm_em <- panel_plot_dir("panel_D_dominance_marginal_bella", "pdf", subdir = "combined/effective_modalities")

ggsave(file.path(pdir_png_dm_em, "panel_D_dominance_marginal_bella_density_effmod.png"),
       p_dom_density_em, width = 9, height = 8, dpi = 300)
ggsave(file.path(pdir_pdf_dm_em, "panel_D_dominance_marginal_bella_density_effmod.pdf"),
       p_dom_density_em, width = 9, height = 8)
```

```{r dominance-effmod-marginal-boxplot-bella-effmod, fig.width=9, fig.height=8}
# Dominance vs effective modalities — marginal boxplots on both axes, all points shown
p_top_box_em <- ggplot(plot_dom, aes(x = dominance, y = target_set, fill = target_set)) +
  geom_boxplot(alpha = 0.3, outlier.shape = NA, linewidth = 0.3) +
  geom_jitter(aes(color = target_set), height = 0.15, alpha = 0.5, size = 0.8, stroke = 0) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  theme_void() + theme(legend.position = "none")

p_right_box_em <- ggplot(plot_dom, aes(x = target_set, y = eff_modalities, fill = target_set)) +
  geom_boxplot(alpha = 0.3, outlier.shape = NA, linewidth = 0.3) +
  geom_jitter(aes(color = target_set), width = 0.15, alpha = 0.5, size = 0.8, stroke = 0) +
  scale_fill_manual(values = target_set_colors) +
  scale_color_manual(values = target_set_colors) +
  scale_y_continuous(breaks = 1:7) +
  theme_void() + theme(legend.position = "none")

p_dom_asm_b_em <- plot_grid(
  p_top_box_em, NULL,
  p_dom_core_em + theme(legend.position = "none"), p_right_box_em,
  ncol = 2, nrow = 2,
  rel_widths = c(4, 1), rel_heights = c(1, 4),
  align = "hv", axis = "tblr"
)
p_dom_boxplot_em <- plot_grid(p_dom_asm_b_em, legend_dom_em, ncol = 1, rel_heights = c(1, 0.1))

print(p_dom_boxplot_em)

ggsave(file.path(pdir_png_dm_em, "panel_D_dominance_marginal_bella_boxplot_effmod.png"),
       p_dom_boxplot_em, width = 9, height = 8, dpi = 300)
ggsave(file.path(pdir_pdf_dm_em, "panel_D_dominance_marginal_bella_boxplot_effmod.pdf"),
       p_dom_boxplot_em, width = 9, height = 8)
```

```{r entropy-violin-bella, fig.width=5, fig.height=5}
# Violin plot of entropy by target set
entropy_violin_df <- entropy_df %>%
  filter(target_set %in% names(target_set_colors)) %>%
  mutate(target_set_label = target_set_labels[target_set])

p_violin <- ggplot(entropy_violin_df,
                   aes(x = target_set_label, y = sparseness, fill = target_set)) +
  geom_violin(alpha = 0.3, scale = "width", trim = FALSE) +
  geom_boxplot(width = 0.15, alpha = 0.4, outlier.shape = NA, fill = "white") +
  geom_jitter(width = 0.1, alpha = 0.5, size = 1, stroke = 0) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = target_set_colors) +
  labs(
    x = "Target set",
    y = "Lifetime sparseness"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "none"
  )

print(p_violin)
save_fig6_plot(p_violin, "panel_D_entropy_violin_targetset_bella", width = 5, height = 5, subdir = "combined/lifetime_sparseness")
```

```{r entropy-violin-bella-effmod, fig.width=5, fig.height=5}
# Violin plot of effective modalities by target set
p_violin_em <- ggplot(entropy_violin_df,
                   aes(x = target_set_label, y = eff_modalities, fill = target_set)) +
  geom_violin(alpha = 0.3, scale = "width", trim = FALSE) +
  geom_boxplot(width = 0.15, alpha = 0.4, outlier.shape = NA, fill = "white") +
  geom_jitter(width = 0.1, alpha = 0.5, size = 1, stroke = 0) +
  geom_hline(yintercept = 7, linetype = "dashed", color = "grey75", linewidth = 0.3) +
  scale_fill_manual(values = target_set_colors) +
  labs(
    x = "Target set",
    y = "Effective modalities"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    text = element_text(color = "grey20"),
    axis.title = element_text(size = rel(0.9), color = "grey30"),
    axis.text = element_text(size = rel(0.85), color = "grey40"),
    axis.line = element_line(linewidth = 0.3, color = "grey70"),
    axis.ticks = element_line(linewidth = 0.25, color = "grey70"),
    panel.grid = element_blank(),
    legend.position = "none"
  )

print(p_violin_em)
save_fig6_plot(p_violin_em, "panel_D_entropy_violin_targetset_bella_effmod", width = 5, height = 5, subdir = "combined/effective_modalities")
```

## 7. Summary statistics

```{r summary-stats}
# Overall summary
cat("=== Entropy Summary (Bella's targets) ===\n")
cat(sprintf("Number of target neurons: %d\n", nrow(entropy_df)))
cat(sprintf("Max possible entropy: %.3f bits (log2(7))\n", max_entropy))
cat(sprintf("Mean entropy: %.3f bits\n", mean(entropy_df$entropy)))
cat(sprintf("Median entropy: %.3f bits\n", median(entropy_df$entropy)))
cat(sprintf("SD entropy: %.3f bits\n", sd(entropy_df$entropy)))
cat(sprintf("Mean lifetime sparseness: %.3f\n", mean(entropy_df$sparseness)))
cat(sprintf("Mean effective modalities: %.1f\n", mean(entropy_df$eff_modalities)))
cat(sprintf("Mean breadth: %.1f modalities\n", mean(entropy_df$breadth)))
cat(sprintf("Neurons with all 7 modalities: %d (%.1f%%)\n",
            sum(entropy_df$breadth == 7),
            100 * mean(entropy_df$breadth == 7)))
cat(sprintf("Neurons with only 1 modality: %d (%.1f%%)\n",
            sum(entropy_df$breadth == 1),
            100 * mean(entropy_df$breadth == 1)))

# Per target set
cat("\n=== By Target Set ===\n")
entropy_df %>%
  filter(target_set %in% names(target_set_labels)) %>%
  mutate(target_set_label = target_set_labels[target_set]) %>%
  group_by(target_set_label) %>%
  summarise(
    n = n(),
    mean_entropy = mean(entropy),
    median_entropy = median(entropy),
    mean_breadth = mean(breadth),
    mean_dominance = mean(dominance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  knitr::kable(digits = 3)
```

```{r session-info}
sessionInfo()
```
