---
title: "Figure 6E (Bella's targets): Pairwise Modality Co-occurrence / Correlation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/fig6_bella_targets/setup.R")

# Additional libraries for this panel
library(ggraph)
library(tidygraph)
library(grid)
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel E: Modality Co-occurrence / Correlation (Bella's targets) — Pairwise Pearson and Spearman correlations between modality strength vectors shown as heatmaps and network graphs. Identifies which modality pairs co-target the same downstream neurons."
panel_legend <- "(E) Pairwise modality co-occurrence among Bella's targets. Pearson and Spearman correlation heatmaps and network graphs reveal which modality pairs tend to co-target the same downstream neurons."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_E_description", "txt"), "panel_E_description.txt"))
```

```{r save-and-show-pheatmap-helper, include=FALSE}
# ---------------------------------------------------------------------------
# Helper: build pheatmap once, save PNG + PDF, display inline
# ---------------------------------------------------------------------------
save_and_show_pheatmap <- function(..., png_path, pdf_path, save_width = 9, save_height = 8) {
  p <- pheatmap(..., silent = TRUE)

  png(png_path, width = save_width, height = save_height, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(pdf_path, width = save_width, height = save_height)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(p)
}
```

## Overview

This panel examines which sensory modality pairs tend to co-target the same
downstream neurons among Bella's targets (M cells, F cells, IR cells). We compute
pairwise correlations between modality strength vectors (each vector = strengths
across all target neurons for one modality) and visualise the results as heatmaps
and a network graph.

High positive correlation between two modalities means that neurons receiving
strong input from one modality also tend to receive strong input from the other.
Negative correlation implies mutual exclusivity of targeting.

## 1. Pearson correlation heatmap

```{r pearson-correlation}
# Ensure modality order in the strength matrix
mat <- strength.bella[, intersect(modality_order, colnames(strength.bella))]

# Pearson correlation between modality columns
cor_pearson <- cor(mat, method = "pearson")

# Annotation: modality groups (olfactory / auditory / visual / contact)
group_annotation <- data.frame(
  group = modality_group[colnames(cor_pearson)],
  row.names = colnames(cor_pearson)
)

# Diverging colour palette
cor_colors <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)

save_and_show_pheatmap(
  cor_pearson,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = cor_colors,
  breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Pearson correlation of modality strengths (Bella's targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_pearson_heatmap_bella", "png"), "panel_E_pearson_heatmap_bella.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_pearson_heatmap_bella", "pdf"), "panel_E_pearson_heatmap_bella.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 2. Spearman rank correlation heatmap

```{r spearman-correlation}
# Spearman: robust to outliers / non-linear monotonic relationships
cor_spearman <- cor(mat, method = "spearman")

save_and_show_pheatmap(
  cor_spearman,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = cor_colors,
  breaks = seq(-1, 1, length.out = 101),
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Spearman rank correlation of modality strengths (Bella's targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_spearman_heatmap_bella", "png"), "panel_E_spearman_heatmap_bella.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_spearman_heatmap_bella", "pdf"), "panel_E_spearman_heatmap_bella.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 3. Co-targeting analysis

For each pair of modalities, count how many target neurons receive non-trivial
input from BOTH modalities. "Non-trivial" is defined as >5% of that modality's
maximum strength across all neurons.

```{r co-targeting-heatmap}
# Compute per-modality thresholds: 5% of each modality's max
mod_max <- apply(mat, 2, max, na.rm = TRUE)
threshold_frac <- 0.05
thresholds <- mod_max * threshold_frac

# Binary matrix: does neuron i receive non-trivial input from modality j?
mat_binary <- sweep(mat, 2, thresholds, ">") * 1

# Co-targeting count: for each pair (i,j), count neurons with both == 1
co_target_mat <- t(mat_binary) %*% mat_binary

# The diagonal = number of neurons with non-trivial input from that modality
cat("Neurons with non-trivial input per modality:\n")
print(diag(co_target_mat))

# Heatmap of co-targeting counts
co_target_colors <- colorRampPalette(c("#F7FCF5", "#00441B"))(50)

save_and_show_pheatmap(
  co_target_mat,
  display_numbers = TRUE,
  number_format = "%d",
  fontsize_number = 10,
  color = co_target_colors,
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = sprintf("Co-targeting counts (threshold: %.0f%% of modality max, Bella's targets)", threshold_frac * 100),
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_counts_bella", "png"), "panel_E_cotarget_counts_bella.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_counts_bella", "pdf"), "panel_E_cotarget_counts_bella.pdf"),
  save_width = 9,
  save_height = 8
)
```

```{r co-targeting-jaccard}
# Jaccard index: co-target(i,j) / (count_i + count_j - co-target(i,j))
# More informative than raw counts since it accounts for marginal frequencies
n_mods <- ncol(mat_binary)
jaccard_mat <- matrix(0, nrow = n_mods, ncol = n_mods,
                      dimnames = list(colnames(mat_binary), colnames(mat_binary)))

for (i in seq_len(n_mods)) {
  for (j in seq_len(n_mods)) {
    both <- sum(mat_binary[, i] == 1 & mat_binary[, j] == 1)
    either <- sum(mat_binary[, i] == 1 | mat_binary[, j] == 1)
    jaccard_mat[i, j] <- ifelse(either > 0, both / either, 0)
  }
}

save_and_show_pheatmap(
  jaccard_mat,
  display_numbers = TRUE,
  number_format = "%.2f",
  fontsize_number = 9,
  color = co_target_colors,
  clustering_method = "complete",
  annotation_col = group_annotation,
  annotation_row = group_annotation,
  annotation_colors = list(group = modality_group_colors),
  border_color = "white",
  cellwidth = 50,
  cellheight = 50,
  main = "Co-targeting Jaccard similarity between modalities (Bella's targets)",
  fontsize = 10,
  png_path = file.path(panel_plot_dir("panel_E_cotarget_jaccard_bella", "png"), "panel_E_cotarget_jaccard_bella.png"),
  pdf_path = file.path(panel_plot_dir("panel_E_cotarget_jaccard_bella", "pdf"), "panel_E_cotarget_jaccard_bella.pdf"),
  save_width = 9,
  save_height = 8
)
```

## 4. Network graph of modality correlations

Nodes represent modalities, edge thickness scales with correlation magnitude,
and edge color distinguishes positive (red) from negative (blue) correlations.

```{r correlation-network, fig.width=8, fig.height=8}
# Build edge list from the upper triangle of the Pearson correlation matrix
edge_df <- expand.grid(
  from = rownames(cor_pearson),
  to = colnames(cor_pearson),
  stringsAsFactors = FALSE
) %>%
  filter(from < to) %>%
  rowwise() %>%
  mutate(
    correlation = cor_pearson[from, to],
    abs_cor = abs(correlation),
    sign = ifelse(correlation >= 0, "positive", "negative")
  ) %>%
  ungroup()

# Node metadata
node_df <- data.frame(
  name = modality_order,
  group = modality_group[modality_order],
  stringsAsFactors = FALSE
)

# Create tidygraph object
graph <- tbl_graph(
  nodes = node_df,
  edges = edge_df,
  directed = FALSE
)

# Circular layout for clean publication figure
p_network <- ggraph(graph, layout = "circle") +
  geom_edge_link(
    aes(
      width = abs_cor,
      color = sign,
      alpha = abs_cor
    )
  ) +
  scale_edge_width_continuous(range = c(0.3, 4), name = "|Correlation|") +
  scale_edge_color_manual(
    values = c("positive" = "#B2182B", "negative" = "#2166AC"),
    name = "Direction"
  ) +
  scale_edge_alpha_continuous(range = c(0.3, 1), guide = "none") +
  geom_node_point(aes(color = group), size = 12) +
  geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
  scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
  labs(title = "Modality correlation network (Pearson, Bella's targets)") +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

print(p_network)
save_fig6_plot(p_network, "panel_E_correlation_network_bella", width = 9, height = 8)
```

```{r correlation-network-thresholded, fig.width=8, fig.height=8}
# Thresholded version: only show edges with |correlation| > 0.1
edge_df_thresh <- edge_df %>% filter(abs_cor > 0.1)

graph_thresh <- tbl_graph(
  nodes = node_df,
  edges = edge_df_thresh,
  directed = FALSE
)

p_network_thresh <- ggraph(graph_thresh, layout = "circle") +
  geom_edge_link(
    aes(
      width = abs_cor,
      color = sign
    ),
    alpha = 0.8
  ) +
  geom_edge_link(
    aes(label = sprintf("%.2f", correlation)),
    angle_calc = "along",
    label_size = 2.5,
    label_dodge = unit(3, "mm"),
    alpha = 0
  ) +
  scale_edge_width_continuous(range = c(0.5, 4), name = "|Correlation|") +
  scale_edge_color_manual(
    values = c("positive" = "#B2182B", "negative" = "#2166AC"),
    name = "Direction"
  ) +
  geom_node_point(aes(color = group), size = 14) +
  geom_node_text(aes(label = name), size = 3.5, fontface = "bold", color = "white") +
  scale_color_manual(values = modality_group_colors, name = "Sensory\nGroup") +
  labs(title = "Modality correlation network (|r| > 0.1, Pearson, Bella's targets)") +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

print(p_network_thresh)
save_fig6_plot(p_network_thresh, "panel_E_correlation_network_thresholded_bella", width = 9, height = 8)
```

## 5. Valence-split correlation analysis

Repeat the correlation heatmaps for excitatory-only and inhibitory-only paths
to determine whether co-occurrence patterns differ by valence.

```{r valence-split-correlations}
# Helper to compute and plot correlation for a given strength matrix and label
plot_valence_correlation <- function(strength_mat, valence_label) {
  # Ensure modality order
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  # Skip if too few columns or rows
  if (ncol(mat_v) < 2 || nrow(mat_v) < 3) {
    message(sprintf("Skipping %s: insufficient data (%d rows, %d modalities)",
                    valence_label, nrow(mat_v), ncol(mat_v)))
    return(invisible(NULL))
  }

  # Remove rows with zero variance (no information)
  row_var <- apply(mat_v, 1, var, na.rm = TRUE)
  mat_v <- mat_v[row_var > 0, , drop = FALSE]

  # Remove modality columns with zero variance
  col_var <- apply(mat_v, 2, var, na.rm = TRUE)
  mat_v <- mat_v[, col_var > 0, drop = FALSE]

  if (ncol(mat_v) < 2) {
    message(sprintf("Skipping %s: fewer than 2 variable modalities after filtering", valence_label))
    return(invisible(NULL))
  }

  cor_mat <- cor(mat_v, method = "pearson")

  # Annotation for remaining modalities
  ga <- data.frame(
    group = modality_group[colnames(cor_mat)],
    row.names = colnames(cor_mat)
  )

  title_str <- sprintf("Pearson correlation (%s paths only, Bella's targets)", valence_label)

  # Build once, save PNG + PDF, display inline
  p <- pheatmap(
    cor_mat,
    display_numbers = TRUE,
    number_format = "%.2f",
    fontsize_number = 9,
    color = cor_colors,
    breaks = seq(-1, 1, length.out = 101),
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  png(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s_bella", valence_label), "png"), sprintf("panel_E_pearson_%s_bella.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(file.path(panel_plot_dir(sprintf("panel_E_pearson_%s_bella", valence_label), "pdf"), sprintf("panel_E_pearson_%s_bella.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(cor_mat)
}
```

```{r valence-excitatory-heatmap, fig.width=9, fig.height=8}
cor_exc <- plot_valence_correlation(strength.bella.exc, "excitatory")
```

```{r valence-inhibitory-heatmap, fig.width=9, fig.height=8}
cor_inh <- plot_valence_correlation(strength.bella.inh, "inhibitory")
```

```{r valence-cotarget-comparison}
# Co-targeting heatmaps per valence
compute_cotarget <- function(strength_mat, valence_label) {
  mods_present <- intersect(modality_order, colnames(strength_mat))
  mat_v <- strength_mat[, mods_present]

  mod_max_v <- apply(mat_v, 2, max, na.rm = TRUE)
  thresh_v <- mod_max_v * threshold_frac
  # Guard against zero max (would make threshold 0, flagging everything)
  thresh_v[thresh_v == 0] <- .Machine$double.eps

  mat_bin_v <- sweep(mat_v, 2, thresh_v, ">") * 1
  co_mat <- t(mat_bin_v) %*% mat_bin_v

  ga <- data.frame(
    group = modality_group[colnames(co_mat)],
    row.names = colnames(co_mat)
  )

  title_str <- sprintf("Co-targeting counts (%s, threshold: %.0f%%, Bella's targets)", valence_label, threshold_frac * 100)

  # Build once, save PNG + PDF, display inline
  p <- pheatmap(
    co_mat,
    display_numbers = TRUE,
    number_format = "%d",
    fontsize_number = 10,
    color = co_target_colors,
    clustering_method = "complete",
    annotation_col = ga,
    annotation_row = ga,
    annotation_colors = list(group = modality_group_colors),
    border_color = "white",
    cellwidth = 50,
    cellheight = 50,
    main = title_str,
    fontsize = 10,
    silent = TRUE
  )

  png(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s_bella", valence_label), "png"), sprintf("panel_E_cotarget_%s_bella.png", valence_label)),
      width = 9, height = 8, units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  pdf(file.path(panel_plot_dir(sprintf("panel_E_cotarget_%s_bella", valence_label), "pdf"), sprintf("panel_E_cotarget_%s_bella.pdf", valence_label)),
      width = 9, height = 8)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off()
  on.exit(NULL)

  grid::grid.newpage()
  grid::grid.draw(p$gtable)

  invisible(co_mat)
}
```

```{r cotarget-excitatory, fig.width=9, fig.height=8}
co_exc <- compute_cotarget(strength.bella.exc, "excitatory")
```

```{r cotarget-inhibitory, fig.width=9, fig.height=8}
co_inh <- compute_cotarget(strength.bella.inh, "inhibitory")
```

## 6. Summary statistics

```{r correlation-summary}
cat("=== Pearson Correlation Summary (combined, Bella's targets) ===\n")
# Extract upper triangle values (excluding diagonal)
upper_idx <- upper.tri(cor_pearson)
cor_vals <- cor_pearson[upper_idx]
pair_labels <- outer(rownames(cor_pearson), colnames(cor_pearson), paste, sep = " - ")[upper_idx]

cor_summary <- data.frame(
  pair = pair_labels,
  pearson = cor_vals,
  spearman = cor_spearman[upper_idx]
) %>%
  arrange(desc(abs(pearson)))

cat(sprintf("Number of modality pairs: %d\n", length(cor_vals)))
cat(sprintf("Mean |Pearson r|: %.3f\n", mean(abs(cor_vals))))
cat(sprintf("Max positive: %.3f\n", max(cor_vals)))
cat(sprintf("Max negative: %.3f\n", min(cor_vals)))
cat(sprintf("Pairs with |r| > 0.3: %d\n", sum(abs(cor_vals) > 0.3)))

cat("\nAll pairwise correlations (sorted by |Pearson|):\n")
knitr::kable(cor_summary, digits = 3)
```

```{r co-targeting-summary}
cat("\n=== Co-targeting Summary (Bella's targets) ===\n")
cat(sprintf("Threshold: %.0f%% of modality max strength\n", threshold_frac * 100))
cat(sprintf("Total target neurons: %d\n", nrow(mat)))

# Diagonal = neurons per modality above threshold
cat("\nNeurons above threshold per modality:\n")
print(diag(co_target_mat))

# Off-diagonal co-targeting
co_upper <- co_target_mat[upper.tri(co_target_mat)]
co_labels <- outer(rownames(co_target_mat), colnames(co_target_mat), paste, sep = " & ")[upper.tri(co_target_mat)]

co_summary <- data.frame(
  pair = co_labels,
  shared_neurons = co_upper
) %>%
  arrange(desc(shared_neurons))

cat("\nTop co-targeting pairs:\n")
knitr::kable(co_summary)
```

```{r session-info}
sessionInfo()
```
