---
title: "Figure 6K: Outlook -- Modality Integration Across Circuit Layers (Bella's Targets)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 8
)

source("/Users/fkampf/Documents/pheromone.paper/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel K: Outlook — Modality Integration Across Circuit Layers (Bella's targets) — How modality information converges at increasing circuit depth for Bella's target sets (M, F, IR, all). Modality-by-layer heatmaps, convergence curves, co-occurrence matrices, per-neuron heatmaps, dendrograms, and flow diagrams."
panel_legend <- "(K) Modality integration across circuit layers for Bella's target sets (M, F, IR, all combined). Heatmaps, convergence curves, co-occurrence matrices, per-neuron integration profiles, modality dendrograms, and flow diagrams at each circuit depth."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_K_description", "txt"), "panel_K_description.txt"))
```

## Overview

This panel examines how sensory modality information converges as signals
propagate deeper into the circuit. By assigning each node in a path a
**circuit order** based on its position (position 1 = sensory, position 2 =
2nd-order, etc.), we can ask: at which layer do modalities first share
neurons? Do modality groupings remain stable across layers?

## 1. Shared style and helper functions

```{r helpers}
# ---------------------------------------------------------------------------
# Cohesive minimal theme for all ggplots
# ---------------------------------------------------------------------------
theme_fig6 <- function(base_size = 11) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      text             = element_text(family = "Helvetica", color = "grey20"),
      plot.title       = element_text(size = rel(1.05), face = "plain",
                                      hjust = 0, margin = margin(b = 8)),
      axis.title       = element_text(size = rel(0.9), color = "grey30"),
      axis.text        = element_text(size = rel(0.85), color = "grey40"),
      axis.line        = element_line(linewidth = 0.3, color = "grey70"),
      axis.ticks       = element_line(linewidth = 0.25, color = "grey70"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.title     = element_text(size = rel(0.85), face = "plain"),
      legend.text      = element_text(size = rel(0.8)),
      legend.key.size  = unit(0.4, "cm"),
      plot.margin      = margin(8, 8, 8, 8)
    )
}

# ---------------------------------------------------------------------------
# Consistent heatmap palettes
# ---------------------------------------------------------------------------
hm_pal     <- colorRampPalette(c("#F7F7F7", "#D1E5F0", "#92C5DE",
                                  "#4393C3", "#2166AC"))(100)
hm_pal_exc <- colorRampPalette(c("#F7F7F7", "#FEE0D2", "#FC9272",
                                  "#DE2D26"))(100)
hm_pal_inh <- colorRampPalette(c("#F7F7F7", "#DEEBF7", "#9ECAE1",
                                  "#3182BD"))(100)
hm_pal_cor <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)
co_pal     <- colorRampPalette(c("#F7F7F7", "#FEE08B", "#FDAE61",
                                  "#D73027"))(100)

# ---------------------------------------------------------------------------
# Sensory neuron exclusion
# ---------------------------------------------------------------------------
sensory_patterns <- c(
  "^ORN_DA1", "^ORN_VA1v", "^ORN_VA1d",
  "^JO-B", "^LC10", "^ppk23", "^ppk25"
)
sensory_regex <- paste(sensory_patterns, collapse = "|")

# ---------------------------------------------------------------------------
# Neuron type classification
# ---------------------------------------------------------------------------
neuron_patterns <- c(
  "aSP-f" = "^aSP-f", "aSP-g" = "^aSP-g", "P1" = "^P1_",
  "pC1" = "^pC1", "pC2" = "^pC2", "mAL" = "^mAL",
  "vAB3" = "^vAB3", "PPN1" = "^PPN1",
  "LH" = "^LH", "PN" = "PN", "MBON" = "^MBON", "KC" = "^KC",
  "SMP" = "^SMP", "SIP" = "^SIP", "SLP" = "^SLP",
  "AVLP" = "^AVLP", "PVLP" = "^PVLP", "WED" = "^WED",
  "CX" = "^(FB|EB|PB|NO|EPG|PFN|PFL|hDelta|FC|FR)",
  "AL-local" = "^(lLN|vLN|LN|v2LN|l2LN)",
  "LC" = "^(LC|LPLC|LPC)", "DN" = "^DN", "LAL" = "^LAL",
  "CRE" = "^CRE", "AOTU" = "^(AOTU|TuBu)", "SEZ" = "^(SEZ|GNG)"
)

classify_neurons <- function(names_vec) {
  classes <- character(length(names_vec))
  for (i in seq_along(neuron_patterns)) {
    mask <- grepl(neuron_patterns[i], names_vec)
    classes[mask & classes == ""] <- names(neuron_patterns)[i]
  }
  classes[classes == ""] <- "other"
  classes
}

# ---------------------------------------------------------------------------
# Parse paths with circuit order
# ---------------------------------------------------------------------------
parse_paths_with_order <- function(paths_df) {
  paths_df %>%
    filter(!is.na(path), !is.na(strength), strength > 0) %>%
    rowwise() %>%
    mutate(nodes_list = list(strsplit(path, " -> ", fixed = TRUE)[[1]])) %>%
    ungroup() %>%
    mutate(path_id = row_number()) %>%
    tidyr::unnest_longer(nodes_list, indices_to = "position") %>%
    rename(node = nodes_list) %>%
    mutate(circuit_order = as.integer(position)) %>%
    filter(!grepl(sensory_regex, node), circuit_order >= 2)
}

# ---------------------------------------------------------------------------
# Safe pheatmap save (no device leak)
# ---------------------------------------------------------------------------
save_and_show_pheatmap <- function(..., png_path, pdf_path,
                                    save_width = 9, save_height = 8) {
  p <- pheatmap(..., silent = TRUE)
  png(png_path, width = save_width, height = save_height,
      units = "in", res = 300)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off(); on.exit(NULL)
  pdf(pdf_path, width = save_width, height = save_height)
  on.exit(try(dev.off(), silent = TRUE), add = TRUE)
  grid::grid.draw(p$gtable)
  dev.off(); on.exit(NULL)
  grid::grid.newpage(); grid::grid.draw(p$gtable)
  invisible(p)
}

col_normalize <- function(mat) {
  cm <- apply(mat, 2, max, na.rm = TRUE); cm[cm == 0] <- 1
  sweep(mat, 2, cm, "/")
}

# ---------------------------------------------------------------------------
# Dendrogram data extractor
# ---------------------------------------------------------------------------
dendro_data_from_hclust <- function(hc) {
  dend <- as.dendrogram(hc)
  segments <- list()
  extract_seg <- function(d, pm) {
    if (is.leaf(d)) return(invisible(NULL))
    ch <- as.list(d)
    cx <- vapply(ch, function(c) {
      if (is.leaf(c)) pm[[attr(c, "label")]] else attr(c, "midpoint_abs")
    }, numeric(1))
    ch_h <- vapply(ch, attr, numeric(1), "height")
    h <- attr(d, "height")
    segments[[length(segments) + 1]] <<- data.frame(
      x = min(cx), y = h, xend = max(cx), yend = h)
    for (i in seq_along(ch))
      segments[[length(segments) + 1]] <<- data.frame(
        x = cx[i], y = h, xend = cx[i], yend = ch_h[i])
    for (c in ch) extract_seg(c, pm)
  }
  add_mid <- function(d, pm) {
    if (is.leaf(d)) { attr(d, "midpoint_abs") <- pm[[attr(d, "label")]]; return(d) }
    ch <- lapply(as.list(d), add_mid, pm = pm)
    attr(d, "midpoint_abs") <- mean(vapply(ch, attr, numeric(1), "midpoint_abs"))
    for (i in seq_along(ch)) d[[i]] <- ch[[i]]
    d
  }
  ord <- hc$order; labs <- hc$labels[ord]
  pm <- setNames(as.list(seq_along(labs)), labs)
  dend <- add_mid(dend, pm); extract_seg(dend, pm)
  list(
    segments = do.call(rbind, segments),
    labels   = data.frame(x = as.numeric(pm[labs]), y = 0, label = labs,
                          stringsAsFactors = FALSE)
  )
}
```

```{r parse-all-sets}
# ---------------------------------------------------------------------------
# Target-set variants (Bella's targets)
# ---------------------------------------------------------------------------
path_sets <- list(
  bella = paths.bella,
  M     = paths.M,
  F     = paths.F,
  IR    = paths.IR
)
set_titles <- c(
  bella = "All Bella's targets",
  M     = "M downstream",
  F     = "F downstream",
  IR    = "IR downstream"
)
parsed <- lapply(path_sets, parse_paths_with_order)
```

## 2. Modality representation by circuit layer

```{r modality-by-layer, fig.width=7, fig.height=4.5, results='asis'}
for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]

  order_mod <- nd %>%
    group_by(circuit_order, modality) %>%
    summarise(total_strength = sum(strength, na.rm = TRUE),
              n_unique_nodes = n_distinct(node), .groups = "drop") %>%
    filter(circuit_order >= 2, circuit_order <= 5)

  mat_wide <- order_mod %>%
    mutate(order_lab = paste0("Order ", circuit_order)) %>%
    select(modality, order_lab, total_strength) %>%
    pivot_wider(names_from = order_lab, values_from = total_strength, values_fill = 0) %>%
    column_to_rownames("modality")
  mat <- as.matrix(mat_wide)
  present_mods <- intersect(modality_order, rownames(mat))
  mat <- mat[present_mods, , drop = FALSE]
  mat_norm <- col_normalize(mat)

  row_annot <- data.frame(row.names = rownames(mat_norm),
    group = unname(modality_group[rownames(mat_norm)]))

  save_and_show_pheatmap(
    mat_norm, cluster_rows = FALSE, cluster_cols = FALSE,
    color = hm_pal, border_color = NA,
    display_numbers = TRUE, number_format = "%.2f", number_color = "grey30",
    annotation_row = row_annot,
    annotation_colors = list(group = modality_group_colors),
    cellwidth = 55, cellheight = 35, fontsize = 10, fontsize_number = 8,
    main = "",
    png_path = file.path(panel_plot_dir(sprintf("panel_K_modality_by_layer_%s", set_name), "png"), sprintf("panel_K_modality_by_layer_%s.png", set_name)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_K_modality_by_layer_%s", set_name), "pdf"), sprintf("panel_K_modality_by_layer_%s.pdf", set_name)),
    save_width = 7, save_height = 4.5)

  save_and_show_pheatmap(
    mat, cluster_rows = FALSE, cluster_cols = FALSE,
    color = hm_pal, border_color = NA,
    display_numbers = TRUE, number_format = "%.1f", number_color = "grey30",
    annotation_row = row_annot,
    annotation_colors = list(group = modality_group_colors),
    cellwidth = 55, cellheight = 35, fontsize = 10, fontsize_number = 8,
    main = "",
    png_path = file.path(panel_plot_dir(sprintf("panel_K_modality_by_layer_raw_%s", set_name), "png"), sprintf("panel_K_modality_by_layer_raw_%s.png", set_name)),
    pdf_path = file.path(panel_plot_dir(sprintf("panel_K_modality_by_layer_raw_%s", set_name), "pdf"), sprintf("panel_K_modality_by_layer_raw_%s.pdf", set_name)),
    save_width = 7, save_height = 4.5)

  p_bar <- ggplot(
    order_mod %>% mutate(modality = factor(modality, levels = modality_order)),
    aes(x = factor(circuit_order), y = n_unique_nodes, fill = modality)
  ) +
    geom_col(position = "dodge", width = 0.75) +
    scale_fill_manual(values = modality_colors, name = NULL) +
    labs(x = "Circuit order", y = "Unique neurons") +
    theme_fig6()

  print(p_bar)
  save_fig6_plot(p_bar, sprintf("panel_K_neuron_count_by_layer_%s", set_name),
                 width = 7, height = 4)
}
```

## 3. Multimodal convergence by layer

```{r convergence, fig.width=6, fig.height=4, results='asis'}
for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]

  convergence <- nd %>%
    filter(circuit_order >= 2, circuit_order <= 5) %>%
    group_by(circuit_order, node) %>%
    summarise(n_mod = n_distinct(modality), .groups = "drop") %>%
    group_by(circuit_order) %>%
    summarise(n_nodes = n(), n_mm = sum(n_mod >= 2),
              pct_mm = 100 * n_mm / n_nodes, .groups = "drop")

  p_conv <- ggplot(convergence, aes(circuit_order, pct_mm)) +
    geom_line(linewidth = 0.8, color = "#4393C3") +
    geom_point(aes(size = n_nodes), color = "#4393C3", shape = 16) +
    scale_x_continuous(breaks = 2:5, labels = paste0(2:5, "nd/rd/th")) +
    scale_size_continuous(name = "n neurons", range = c(1.5, 5),
                          guide = guide_legend(override.aes = list(color = "grey50"))) +
    labs(x = "Circuit layer", y = "% multimodal neurons") +
    theme_fig6()

  print(p_conv)
  save_fig6_plot(p_conv, sprintf("panel_K_convergence_by_layer_%s", set_name),
                 width = 6, height = 4)
}
```

## 4. Modality co-occurrence by layer

```{r cooccurrence, fig.width=6, fig.height=5.5, results='asis'}
compute_cooccurrence <- function(nd, ord) {
  bdf <- nd %>% filter(circuit_order == ord) %>% distinct(node, modality) %>%
    mutate(v = 1L) %>%
    pivot_wider(names_from = modality, values_from = v, values_fill = 0L) %>%
    column_to_rownames("node")
  for (m in modality_order) if (!m %in% colnames(bdf)) bdf[[m]] <- 0L
  bm <- as.matrix(bdf[, intersect(modality_order, colnames(bdf))])
  t(bm) %*% bm
}

for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]
  for (ord in intersect(2:min(5, max(nd$circuit_order)), unique(nd$circuit_order))) {
    co <- compute_cooccurrence(nd, ord)
    if (nrow(co) < 2) next
    save_and_show_pheatmap(
      co, cluster_rows = TRUE, cluster_cols = TRUE,
      clustering_method = "ward.D2",
      color = co_pal, border_color = NA,
      display_numbers = TRUE, number_format = "%d", number_color = "grey25",
      cellwidth = 42, cellheight = 42, fontsize = 10, fontsize_number = 9,
      main = sprintf("Order %d", ord),
      png_path = file.path(panel_plot_dir(sprintf("panel_K_cooccurrence_order%d_%s", ord, set_name), "png"), sprintf("panel_K_cooccurrence_order%d_%s.png", ord, set_name)),
      pdf_path = file.path(panel_plot_dir(sprintf("panel_K_cooccurrence_order%d_%s", ord, set_name), "pdf"), sprintf("panel_K_cooccurrence_order%d_%s.pdf", ord, set_name)),
      save_width = 6, save_height = 5.5)
  }
}
```

## 5. Per-neuron modality profiles at 3rd and 4th order

```{r per-neuron-heatmaps, fig.width=8, fig.height=9, results='asis'}
build_order_matrix <- function(nd, ord, top_n = 40, val = NULL) {
  dat <- nd %>% filter(circuit_order == ord)
  if (!is.null(val)) dat <- dat %>% filter(valence == val)
  if (nrow(dat) == 0) return(NULL)
  mdf <- dat %>%
    group_by(node, modality) %>%
    summarise(s = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = modality, values_from = s, values_fill = 0) %>%
    column_to_rownames("node")
  for (m in modality_order) if (!m %in% colnames(mdf)) mdf[[m]] <- 0
  mdf <- as.matrix(mdf[, intersect(modality_order, colnames(mdf))])
  rs <- rowSums(mdf); n <- min(top_n, nrow(mdf))
  mdf[order(rs, decreasing = TRUE)[seq_len(n)], , drop = FALSE]
}

col_annot <- data.frame(row.names = modality_order,
  group = unname(modality_group[modality_order]))
annot_colors_base <- list(group = modality_group_colors)

for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]

  for (ord in c(3, 4)) {
    ol <- paste0(ord, ifelse(ord == 3, "rd", "th"))
    mat_raw <- build_order_matrix(nd, ord)
    if (is.null(mat_raw) || nrow(mat_raw) < 3) next

    ra <- data.frame(row.names = rownames(mat_raw),
      type = classify_neurons(rownames(mat_raw)))
    nc <- sort(unique(ra$type))
    nc_col <- setNames(scales::hue_pal(l = 65, c = 80)(length(nc)), nc)
    if ("other" %in% names(nc_col)) nc_col["other"] <- "#D0D0D0"
    ac <- c(annot_colors_base, list(type = nc_col))

    save_and_show_pheatmap(
      mat_raw, cluster_rows = TRUE, cluster_cols = FALSE,
      clustering_method = "ward.D2",
      color = hm_pal, border_color = NA,
      annotation_row = ra, annotation_col = col_annot, annotation_colors = ac,
      fontsize_row = 6.5, fontsize_col = 9, main = "",
      png_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_raw_%s", ol, set_name), "png"), sprintf("panel_K_heatmap_%s_order_raw_%s.png", ol, set_name)),
      pdf_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_raw_%s", ol, set_name), "pdf"), sprintf("panel_K_heatmap_%s_order_raw_%s.pdf", ol, set_name)),
      save_width = 8, save_height = 9)

    save_and_show_pheatmap(
      col_normalize(mat_raw), cluster_rows = TRUE, cluster_cols = FALSE,
      clustering_method = "ward.D2",
      color = hm_pal, border_color = NA,
      annotation_row = ra, annotation_col = col_annot, annotation_colors = ac,
      fontsize_row = 6.5, fontsize_col = 9, main = "",
      png_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_%s", ol, set_name), "png"), sprintf("panel_K_heatmap_%s_order_%s.png", ol, set_name)),
      pdf_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_%s", ol, set_name), "pdf"), sprintf("panel_K_heatmap_%s_order_%s.pdf", ol, set_name)),
      save_width = 8, save_height = 9)

    me <- build_order_matrix(nd, ord, val = "excitatory")
    if (!is.null(me) && nrow(me) >= 3) {
      re <- data.frame(row.names = rownames(me), type = classify_neurons(rownames(me)))
      ne <- sort(unique(re$type))
      ne_c <- setNames(scales::hue_pal(l = 65, c = 80)(length(ne)), ne)
      if ("other" %in% names(ne_c)) ne_c["other"] <- "#D0D0D0"
      save_and_show_pheatmap(
        me, cluster_rows = TRUE, cluster_cols = FALSE,
        clustering_method = "ward.D2",
        color = hm_pal_exc, border_color = NA,
        annotation_row = re, annotation_col = col_annot,
        annotation_colors = c(annot_colors_base, list(type = ne_c)),
        fontsize_row = 6.5, fontsize_col = 9, main = "",
        png_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_exc_%s", ol, set_name), "png"), sprintf("panel_K_heatmap_%s_order_exc_%s.png", ol, set_name)),
        pdf_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_exc_%s", ol, set_name), "pdf"), sprintf("panel_K_heatmap_%s_order_exc_%s.pdf", ol, set_name)),
        save_width = 8, save_height = 9)
    }

    mi <- build_order_matrix(nd, ord, val = "inhibitory")
    if (!is.null(mi) && nrow(mi) >= 3) {
      ri <- data.frame(row.names = rownames(mi), type = classify_neurons(rownames(mi)))
      ni <- sort(unique(ri$type))
      ni_c <- setNames(scales::hue_pal(l = 65, c = 80)(length(ni)), ni)
      if ("other" %in% names(ni_c)) ni_c["other"] <- "#D0D0D0"
      save_and_show_pheatmap(
        mi, cluster_rows = TRUE, cluster_cols = FALSE,
        clustering_method = "ward.D2",
        color = hm_pal_inh, border_color = NA,
        annotation_row = ri, annotation_col = col_annot,
        annotation_colors = c(annot_colors_base, list(type = ni_c)),
        fontsize_row = 6.5, fontsize_col = 9, main = "",
        png_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_inh_%s", ol, set_name), "png"), sprintf("panel_K_heatmap_%s_order_inh_%s.png", ol, set_name)),
        pdf_path = file.path(panel_plot_dir(sprintf("panel_K_heatmap_%s_order_inh_%s", ol, set_name), "pdf"), sprintf("panel_K_heatmap_%s_order_inh_%s.pdf", ol, set_name)),
        save_width = 8, save_height = 9)
    }
  }
}
```

## 6. Modality dendrogram

```{r dendrograms, fig.width=6, fig.height=4, results='asis'}
for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]

  mp <- nd %>%
    group_by(node, modality) %>%
    summarise(s = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = modality, values_from = s, values_fill = 0)
  for (m in modality_order) if (!m %in% colnames(mp)) mp[[m]] <- 0
  pmat <- mp %>% column_to_rownames("node") %>%
    select(all_of(intersect(modality_order, colnames(.)))) %>% as.matrix()

  if (nrow(pmat) >= 3) {
    mc <- cor(pmat, method = "pearson")
    hc <- hclust(as.dist(1 - mc), method = "ward.D2")
    dd <- dendro_data_from_hclust(hc)

    p_d <- ggplot() +
      geom_segment(data = dd$segments,
        aes(x = x, y = y, xend = xend, yend = yend),
        color = "grey50", linewidth = 0.5) +
      geom_text(data = dd$labels,
        aes(x = x, y = y - 0.015, label = label, color = label),
        hjust = 0.5, vjust = 1, size = 3.8, fontface = "bold") +
      scale_color_manual(values = modality_colors, guide = "none") +
      labs(y = "1 - r", x = NULL) +
      theme_fig6() +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            axis.line.x = element_blank())

    print(p_d)
    save_fig6_plot(p_d, sprintf("panel_K_modality_dendrogram_%s", set_name),
                   width = 6, height = 4)

    save_and_show_pheatmap(
      mc, cluster_rows = TRUE, cluster_cols = TRUE,
      clustering_method = "ward.D2",
      color = hm_pal_cor, breaks = seq(-1, 1, length.out = 101),
      border_color = NA,
      display_numbers = TRUE, number_format = "%.2f", number_color = "grey25",
      cellwidth = 42, cellheight = 42, fontsize = 10, fontsize_number = 8,
      main = "",
      png_path = file.path(panel_plot_dir(sprintf("panel_K_modality_correlation_%s", set_name), "png"), sprintf("panel_K_modality_correlation_%s.png", set_name)),
      pdf_path = file.path(panel_plot_dir(sprintf("panel_K_modality_correlation_%s", set_name), "pdf"), sprintf("panel_K_modality_correlation_%s.pdf", set_name)),
      save_width = 6, save_height = 5.5)
  }

  dps <- list()
  for (ord in intersect(2:min(5, max(nd$circuit_order)), unique(nd$circuit_order))) {
    od <- nd %>% filter(circuit_order == ord) %>%
      group_by(node, modality) %>%
      summarise(s = sum(strength, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(names_from = modality, values_from = s, values_fill = 0)
    for (m in modality_order) if (!m %in% colnames(od)) od[[m]] <- 0
    om <- od %>% column_to_rownames("node") %>%
      select(all_of(intersect(modality_order, colnames(.)))) %>% as.matrix()
    if (nrow(om) >= 3) {
      ohc <- hclust(as.dist(1 - cor(om)), method = "ward.D2")
      odd <- dendro_data_from_hclust(ohc)
      dps[[paste0("Order ", ord)]] <- ggplot() +
        geom_segment(data = odd$segments,
          aes(x = x, y = y, xend = xend, yend = yend),
          color = "grey50", linewidth = 0.4) +
        geom_text(data = odd$labels,
          aes(x = x, y = y - 0.015, label = label, color = label),
          hjust = 0.5, vjust = 1, size = 3, fontface = "bold") +
        scale_color_manual(values = modality_colors, guide = "none") +
        labs(title = paste0("Order ", ord), y = "1 - r", x = NULL) +
        theme_fig6(base_size = 9) +
        theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.line.x = element_blank(),
              plot.title = element_text(size = 9, hjust = 0.5))
    }
  }
  if (length(dps) >= 2) {
    pc <- wrap_plots(dps, nrow = 1)
    print(pc)
    save_fig6_plot(pc, sprintf("panel_K_modality_dendrogram_per_layer_%s", set_name),
                   width = 3.5 * length(dps), height = 4)
  }
}
```

## 7. Modality flow across layers

```{r modality-flow, fig.width=6, fig.height=4, results='asis'}
for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]

  flow <- nd %>%
    filter(circuit_order >= 2, circuit_order <= 5) %>%
    group_by(circuit_order, modality) %>%
    summarise(s = sum(strength, na.rm = TRUE), .groups = "drop") %>%
    mutate(modality = factor(modality, levels = modality_order))

  p_prop <- ggplot(flow, aes(factor(circuit_order), s, fill = modality)) +
    geom_col(position = "fill", width = 0.7) +
    scale_fill_manual(values = modality_colors, name = NULL) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Circuit order", y = "Proportion") +
    theme_fig6()

  print(p_prop)
  save_fig6_plot(p_prop, sprintf("panel_K_modality_flow_%s", set_name),
                 width = 6, height = 4)

  p_abs <- ggplot(flow, aes(factor(circuit_order), s, fill = modality)) +
    geom_col(position = "stack", width = 0.7) +
    scale_fill_manual(values = modality_colors, name = NULL) +
    scale_y_continuous(labels = scales::comma) +
    labs(x = "Circuit order", y = "Total strength") +
    theme_fig6()

  print(p_abs)
  save_fig6_plot(p_abs, sprintf("panel_K_modality_flow_raw_%s", set_name),
                 width = 6, height = 4)
}
```

## 8. Summary

```{r summary-table, results='asis'}
for (set_name in names(parsed)) {
  cat(sprintf("\n\n### %s\n\n", set_titles[set_name]))
  nd <- parsed[[set_name]]
  ls <- nd %>%
    filter(circuit_order >= 2, circuit_order <= 5) %>%
    group_by(circuit_order, node) %>%
    summarise(nm = n_distinct(modality), .groups = "drop") %>%
    group_by(circuit_order) %>%
    summarise(n = n(), multimodal = sum(nm >= 2),
              `% multimodal` = round(100 * multimodal / n, 1),
              `mean modalities` = round(mean(nm), 2), .groups = "drop")
  print(knitr::kable(ls))
}
```

```{r session-info}
sessionInfo()
```
