---
title: "Figure 6B (Bella's targets): Modality Composition Heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 12
)

source("/Users/fkampf/Documents/pheromone.paper/fig6_bella_targets/setup.R")
```

```{r panel-description, echo=FALSE}
panel_desc <- "Panel B: Modality Composition Heatmap (Bella's targets) — Clustered heatmap of column-normalized modality input strength to Bella's target neurons. Rows are target neuron types; columns are the 7 sensory modalities. Row annotations mark target group (M, F, IR downstream)."
panel_legend <- "(B) Column-normalized heatmap of modality input strength to Bella's target neurons. Rows are clustered by Ward's method; annotations indicate target group membership (M-downstream, F-downstream, IR-downstream)."
cat(panel_desc)
writeLines(c(panel_desc, "", paste(rep("_", 30), collapse = ""), "", "Figure legend:", "", panel_legend), file.path(panel_plot_dir("panel_B_description", "txt"), "panel_B_description.txt"))
```

## Overview

Panel B is a clustered heatmap of modality input strength to Bella's third-order
courtship target neurons. Columns are the 7 sensory modalities; rows are target
neuron types. The matrix is column-normalized so that each modality's maximum
is 1, making the relative distribution across targets comparable between
modalities of different absolute strengths.

Row annotations indicate the target group (M_downstream, F_downstream,
IR_downstream). Column annotations indicate the modality group
(olfactory / auditory / visual / contact).

## 1. Prepare strength matrix

```{r strength-matrix}
# ---------------------------------------------------------------------------
# Column-normalize the pre-computed strength matrix
# Each column (modality) is scaled so its maximum value = 1
# ---------------------------------------------------------------------------
mat <- strength.bella

# Column-normalize: divide each column by its max
col_maxes <- apply(mat, 2, max, na.rm = TRUE)
col_maxes[col_maxes == 0] <- 1
mat_norm <- sweep(mat, 2, col_maxes, "/")

cat(sprintf("Matrix dimensions: %d target neurons x %d modalities\n",
            nrow(mat_norm), ncol(mat_norm)))
```

## 2. Build row annotation (target group membership)

Each target neuron belongs to one of Bella's three target groups:
M_downstream, F_downstream, or IR_downstream. We use `bella_groups_df`
to assign primary group membership.

```{r row-annotation}
# ---------------------------------------------------------------------------
# Determine target_group membership for each target neuron
# ---------------------------------------------------------------------------
# Multiple color bands: one column per target group (yes/no)
# Neurons belonging to multiple groups show all their memberships
all_ends <- rownames(mat_norm)

M_ends  <- unique(paths.M$end)
F_ends  <- unique(paths.F$end)
IR_ends <- unique(paths.IR$end)

row_annot <- data.frame(
  `M cells`  = ifelse(all_ends %in% M_ends,  "yes", "no"),
  `F cells`  = ifelse(all_ends %in% F_ends,  "yes", "no"),
  `IR cells` = ifelse(all_ends %in% IR_ends, "yes", "no"),
  row.names  = all_ends,
  check.names = FALSE
)
```

## 3. Build column annotation (modality group)

```{r col-annotation}
# ---------------------------------------------------------------------------
# Map each modality column to its sensory group
# ---------------------------------------------------------------------------
col_annot <- data.frame(
  modality_group = modality_group[colnames(mat_norm)],
  row.names      = colnames(mat_norm)
)
```

## 4. Define annotation colours

```{r annotation-colors}
# ---------------------------------------------------------------------------
# Annotation colour maps for pheatmap
# ---------------------------------------------------------------------------
ann_colors <- list(
  `M cells`      = c("yes" = "#E41A1C", "no" = "white"),
  `F cells`      = c("yes" = "#377EB8", "no" = "white"),
  `IR cells`     = c("yes" = "#4DAF4A", "no" = "white"),
  modality_group = modality_group_colors
)
```

## 5. Draw heatmap

```{r heatmap-bella, fig.width=8, fig.height=10}
# ---------------------------------------------------------------------------
# Primary heatmap: modality strength to Bella's target groups
# ---------------------------------------------------------------------------
hm <- pheatmap(
  mat_norm,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,       # keep modality order fixed
  cluster_rows             = TRUE,
  annotation_row           = row_annot,
  annotation_col           = col_annot,
  annotation_colors        = ann_colors,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons (Bella's targets)",
  angle_col                = 45,
  silent                   = TRUE
)

# Draw the heatmap
hm
```

## 6. Save heatmap

```{r save-heatmap}
save_heatmap_both_orientations(
  mat        = mat_norm,
  row_annot  = row_annot,
  col_annot  = col_annot,
  ann_colors = ann_colors,
  name       = "panel_B_heatmap_bella",
  width      = 8,
  height     = 10,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method        = "ward.D2",
  cluster_cols             = FALSE,
  cluster_rows             = TRUE,
  color                    = viridis(100),
  border_color             = NA,
  fontsize_row             = 7,
  fontsize_col             = 10,
  main                     = "Modality input strength to 3rd-order courtship neurons (Bella's targets)",
  angle_col                = 45
)
```

## 7. Export row clustering order for use in Panel C

Save the dendrogram row order so Panel C can use the same neuron ordering
for consistent visual comparison.

```{r export-row-order}
# ---------------------------------------------------------------------------
# Extract row order from the heatmap's dendrogram
# ---------------------------------------------------------------------------
row_order_bella <- rownames(mat_norm)[hm$tree_row$order]

# Save to an RDS file for Panel C to load
saveRDS(
  row_order_bella,
  file = file.path(fig6_dir, "panel_B_row_order.rds")
)

cat("Row order saved. First 10 neurons:\n")
head(row_order_bella, 10)
```
