---
title: "[DEPRECATED] Figure 5 CE1: AN05B035 inhibitory routing fraction per mAL_m subtype (ppk23 only)"
output: html_document
---

> **DEPRECATED -- do not cite as a separate panel.**
>
> This figure is a 2-category, single-channel (ppk23-only) subset of the
> primary `panel_M_inhibitory_sources.Rmd`, which shows the full 4-source
> decomposition on BOTH ppk23 and ppk25 simultaneously. Retained for
> reproducibility; rendering still works and uses the shared palette +
> sort order.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
```

## Extract ppk23 paths ending at mAL_m subtypes

```{r ppk23-paths}
ppk23_paths <- mal_paths %>%
  filter(modality == "ppk23", end %in% mal_subtypes)

ppk23_paths <- ppk23_paths %>%
  mutate(
    via_an05b035 = grepl("AN05B035", path, fixed = TRUE),
    route = if_else(via_an05b035,
                    "AN05B035 (GABA)",
                    "non-AN05B035 (direct / AN09B017)")
  )
```

## Per-mAL routing totals

```{r routing-totals}
routing <- ppk23_paths %>%
  group_by(end, route) %>%
  summarise(total_strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  tidyr::complete(end = mal_subtypes, route,
                  fill = list(total_strength = 0)) %>%
  group_by(end) %>%
  mutate(
    grand_total = sum(total_strength),
    fraction    = if_else(grand_total > 0, total_strength / grand_total, 0)
  ) %>%
  ungroup()

inh_fraction <- routing %>%
  filter(route == "AN05B035 (GABA)") %>%
  select(end, inh_fraction = fraction, grand_total)

# Pass 2: use shared sort order when available
sort_path <- file.path(fig5_dir, "plots_mspecific", "panel_M",
                       "panel_M_mal_sort_order.rds")
if (file.exists(sort_path)) {
  ord <- readRDS(sort_path)
  ord <- ord[ord %in% mal_subtypes]
  ord <- c(ord, setdiff(mal_subtypes, ord))
  message("  Using shared sort order from panel_M_inhibitory_sources.")
} else {
  ord <- inh_fraction %>%
    arrange(desc(inh_fraction), desc(grand_total)) %>%
    pull(end)
}

routing <- routing %>%
  mutate(
    end   = factor(end, levels = rev(ord)),
    route = factor(route, levels = c("non-AN05B035 (direct / AN09B017)",
                                     "AN05B035 (GABA)"))
  )

dom_groups <- inh_fraction %>%
  mutate(group = case_when(
    inh_fraction >= 0.5  ~ "AN05B035-dominant",
    inh_fraction <= 0.1  ~ "AN09B017-dominant",
    TRUE                 ~ "mixed"
  ))
```

## Stacked horizontal bar chart

```{r bar-plot, fig.width=10, fig.height=7}
# Pass 1: shared palette -- AN05B035 = canonical deep blue, non-AN05B035
# collapsed into the grey complement (was red previously).
route_colors <- c(
  "non-AN05B035 (direct / AN09B017)" = "grey70",
  "AN05B035 (GABA)"                  = "#1F4E79"
)

# Pass 3: in-bar percentage labels where segment > 8%
routing <- routing %>%
  mutate(
    lbl   = ifelse(fraction > 0.08, sprintf("%.0f%%", 100 * fraction), ""),
    lbl_c = ifelse(route == "AN05B035 (GABA)", "white", "grey15")
  )

p_bar <- ggplot(routing, aes(x = fraction, y = end, fill = route)) +
  geom_col(width = 0.82, color = "white", linewidth = 0.2) +
  geom_text(aes(label = lbl, color = lbl_c),
            position = position_stack(vjust = 0.5),
            size = 3, show.legend = FALSE) +
  scale_color_identity() +
  geom_vline(xintercept = 0.5, linewidth = 0.4, linetype = "dashed",
             color = "grey40") +
  scale_fill_manual(values = route_colors, name = "Route from ppk23") +
  # Pass 6: explicit [0, 100%] range
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1.001),
                     breaks = c(0, 0.25, 0.5, 0.75, 1),
                     expand = expansion(mult = c(0, 0.01))) +
  labs(x = "Fraction of ppk23 -> mAL path strength",
       y = NULL,
       # Pass 8: informative title instead of generic
       caption = "DEPRECATED: see panel_M_inhibitory_sources for the full 4-source, 2-channel decomposition.") +
  theme_pub(base_size = 12) +
  theme(
    legend.position    = "top",
    # Pass 7
    axis.text.y        = element_text(size = rel(1.15), face = "bold"),
    axis.text.x        = element_text(angle = 0, hjust = 0.5),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.caption       = element_text(size = 9, color = "grey40",
                                      face = "italic", hjust = 0)
  )

# Dominance bracket annotations (retained, but simpler)
ord_top_to_bottom <- ord
n <- length(ord_top_to_bottom)
ann_df <- dom_groups %>% mutate(rank = match(end, ord_top_to_bottom))
an05_rows <- ann_df %>% filter(group == "AN05B035-dominant")
an09_rows <- ann_df %>% filter(group == "AN09B017-dominant")

if (nrow(an05_rows) > 0) {
  p_bar <- p_bar +
    annotate("segment",
             x = 1.03, xend = 1.03,
             y = n - min(an05_rows$rank) + 1 + 0.35,
             yend = n - max(an05_rows$rank) + 1 - 0.35,
             linewidth = 1.1, color = "#1F4E79") +
    annotate("text", x = 1.05,
             y = mean(c(n - min(an05_rows$rank) + 1,
                        n - max(an05_rows$rank) + 1)),
             label = "AN05B035-\ndominant", hjust = 0, size = 3.2,
             color = "#1F4E79", fontface = "bold", lineheight = 0.9)
}
if (nrow(an09_rows) > 0) {
  p_bar <- p_bar +
    annotate("segment",
             x = 1.03, xend = 1.03,
             y = n - min(an09_rows$rank) + 1 + 0.35,
             yend = n - max(an09_rows$rank) + 1 - 0.35,
             linewidth = 1.1, color = "grey45") +
    annotate("text", x = 1.05,
             y = mean(c(n - min(an09_rows$rank) + 1,
                        n - max(an09_rows$rank) + 1)),
             label = "non-AN05B035-\ndominant", hjust = 0, size = 3.2,
             color = "grey40", fontface = "bold", lineheight = 0.9)
}

p_bar <- p_bar + coord_cartesian(xlim = c(0, 1), clip = "off") +
  theme(plot.margin = margin(6, 100, 6, 6))

print(p_bar)
```

## Save outputs

```{r save}
# Keep old panel path name for backward compatibility but also write to the
# canonical panel_M folder so it groups with its siblings.
pdf_dir_main <- panel_plot_dir("panel_M", "pdf")
png_dir_main <- panel_plot_dir("panel_M", "png")

ggsave(file.path(pdf_dir_main, "panel_M_an05b035_fraction_bar.pdf"),
       plot = p_bar, width = 10, height = 7)
ggsave(file.path(png_dir_main, "panel_M_an05b035_fraction_bar.png"),
       plot = p_bar, width = 10, height = 7, dpi = 300)

write.csv(
  routing %>% mutate(end = as.character(end)) %>%
    select(end, route, total_strength, grand_total, fraction) %>%
    arrange(end, route),
  file.path(pdf_dir_main, "panel_M_an05b035_fraction_bar_data.csv"),
  row.names = FALSE
)
```

## Narrative

```{r narrative, echo=FALSE}
n_total <- nrow(dom_groups)
n_an05 <- sum(dom_groups$group == "AN05B035-dominant")
n_an09 <- sum(dom_groups$group == "AN09B017-dominant")
n_mix  <- sum(dom_groups$group == "mixed")

narrative <- sprintf(
"[DEPRECATED] PANEL CE1: AN05B035 as the GABAergic sign-inverter in ppk23 -> mAL
================================================================================

2-category, ppk23-only view. Superseded by panel_M_inhibitory_sources which
shows the full 4-source decomposition on BOTH ppk23 and ppk25.

Group counts (retained for reference):
  AN05B035-dominant (fraction >= 50%%):  %d / %d
  non-AN05B035-dominant (<= 10%%):        %d / %d
  mixed:                                 %d
",
n_an05, n_total,
n_an09, n_total,
n_mix
)

writeLines(narrative, file.path(pdf_dir_main,
                                "panel_M_an05b035_fraction_bar_narrative.txt"))
cat(narrative)
```
