---
title: "Figure 5 CE2: AN05B035-routed ppk23 input vs mAL E/I balance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggrepel)
```

## Goal

Test whether mAL_m subtypes whose ppk23 drive is routed more heavily through
the GABAergic sign-inverter AN05B035 exhibit a more inhibitory (more negative)
E/I balance. We compute, for each of the 16 male-specific mAL subtypes:

- x: fraction of total ppk23->mAL path strength (from `mal_paths`) that passes
  through AN05B035 (AN05B035 appears as an intermediate node in the path).
- y: ppk23 E/I balance index `(exc - inh)/(exc + inh)` from
  `mal_strength_exc` / `mal_strength_inh`.

## Compute AN05B035 routing fraction per mAL_m subtype

```{r compute-fraction}
an05_node <- "AN05B035"

# Restrict to ppk23 paths that end on a male-specific mAL
ppk23_paths <- mal_paths %>%
  filter(modality == "ppk23", end %in% mal_subtypes) %>%
  mutate(
    nodes_list = strsplit(path, " -> ", fixed = TRUE),
    has_an05   = vapply(nodes_list, function(x) an05_node %in% x, logical(1))
  )

cat(sprintf("ppk23 paths to mAL_m subtypes: %d total, %d via AN05B035\n",
            nrow(ppk23_paths), sum(ppk23_paths$has_an05)))

an05_frac <- ppk23_paths %>%
  group_by(end) %>%
  summarise(
    total_strength  = sum(strength, na.rm = TRUE),
    an05_strength   = sum(strength[has_an05], na.rm = TRUE),
    an05_fraction   = ifelse(total_strength > 0, an05_strength / total_strength, 0),
    .groups = "drop"
  )

# Ensure all 16 mAL_m subtypes present (fill zero if no ppk23 paths)
an05_frac <- tibble(end = mal_subtypes) %>%
  left_join(an05_frac, by = "end") %>%
  mutate(across(c(total_strength, an05_strength, an05_fraction),
                ~ replace_na(.x, 0)))

print(an05_frac)
```

## ppk23 E/I balance per subtype

```{r ei-balance}
exc <- mal_strength_exc[mal_subtypes, "ppk23"]
inh <- mal_strength_inh[mal_subtypes, "ppk23"]

ei_df <- tibble(
  subtype    = mal_subtypes,
  exc        = exc,
  inh        = inh,
  ei_balance = ifelse(exc + inh > 0, (exc - inh) / (exc + inh), 0),
  nt         = mal_nt[mal_subtypes]
)

plot_df <- ei_df %>%
  left_join(an05_frac, by = c("subtype" = "end"))

print(plot_df)
```

## Linear regression

```{r regression}
fit <- lm(ei_balance ~ an05_fraction, data = plot_df)
fit_summary <- summary(fit)
r2    <- fit_summary$r.squared
slope <- coef(fit)[["an05_fraction"]]
pval  <- if (nrow(fit_summary$coefficients) >= 2) fit_summary$coefficients[2, 4] else NA_real_

cat(sprintf("Linear fit: y = %.3f + %.3f * x\n", coef(fit)[["(Intercept)"]], slope))
cat(sprintf("R^2 = %.3f, p = %.3g, n = %d\n", r2, pval, nrow(plot_df)))
```

## Scatter plot

```{r scatter, fig.width=8, fig.height=7}
label_r2  <- sprintf("R^2 == %.3f", r2)
label_p   <- sprintf("italic(p) == %.3g", pval)

p_scatter <- ggplot(plot_df, aes(x = an05_fraction, y = ei_balance)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4, color = "grey40") +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE,
              color = "black", fill = "grey80", linewidth = 0.6, alpha = 0.3) +
  geom_point(aes(color = nt), size = 4.5) +
  ggrepel::geom_text_repel(aes(label = subtype), size = 3,
                            max.overlaps = Inf, min.segment.length = 0,
                            seed = 1) +
  scale_color_manual(values = nt_colors, name = "Neurotransmitter") +
  annotate("text", x = min(plot_df$an05_fraction, na.rm = TRUE),
           y = max(plot_df$ei_balance, na.rm = TRUE),
           label = label_r2, parse = TRUE, hjust = 0, vjust = 1, size = 4) +
  annotate("text", x = min(plot_df$an05_fraction, na.rm = TRUE),
           y = max(plot_df$ei_balance, na.rm = TRUE) - 0.08,
           label = label_p, parse = TRUE, hjust = 0, vjust = 1, size = 4) +
  labs(
    x = "Fraction of ppk23 -> mAL path strength routed via AN05B035",
    y = "ppk23 E/I balance  (exc - inh) / (exc + inh)"
  )

# Save via canonical plots_mspecific path
pdf_dir <- panel_plot_dir("panel_M_vs_balance", "pdf")
png_dir <- panel_plot_dir("panel_M_vs_balance", "png")
ggsave(file.path(pdf_dir, "panel_M_vs_balance.pdf"),
       plot = p_scatter + theme_pub(base_size = 14),
       width = 8, height = 7)
ggsave(file.path(png_dir, "panel_M_vs_balance.png"),
       plot = p_scatter + theme_pub(base_size = 14),
       width = 8, height = 7, dpi = 300)

print(p_scatter)
```

## Summary

```{r summary, echo=FALSE}
cat(sprintf(
"CE2 summary: for %d mAL_m subtypes, ppk23 E/I balance vs AN05B035-routed
ppk23 path fraction yields slope = %.3f, R^2 = %.3f, p = %.3g.
Negative slope is consistent with AN05B035 acting as a GABAergic sign-inverter
that pushes mAL subtypes toward inhibitory (negative) ppk23 E/I balance.\n",
nrow(plot_df), slope, r2, pval))
```
