---
title: "Figure 5 CE3: GABA sign-inverter layer - 3-layer bipartite network"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
```

```{r paths}
plot_name <- "panel_M_bipartite_network"
pdf_dir   <- panel_plot_dir(plot_name, "pdf")
png_dir   <- panel_plot_dir(plot_name, "png")
```

## Goal

Show the **GABA sign-inverter layer** as a distributed inhibitory relay
with a 3-layer bipartite network:

- **Top layer**: ppk23 AND ppk25 sensory neuron types, stacked as two
  rows (ppk23 above, ppk25 below) that feed into a shared middle SI
  layer. Source rows are color-coded by channel.
- **Middle layer**: individual GABAergic sign-inverter (SI) types drawn
  from the `gaba_sign_inverter_types` pool (all AN05B/IN05* GABA cells
  that appear as interior relays on ppk -> mAL_m inhibitory paths). The
  lowest-throughput SIs may be collapsed into "other GABA SIs" to keep
  the layout legible.
- **Bottom layer**: the male-specific mAL_m subtypes.

Edge thickness is proportional to raw synapse count (`adj.matrix.raw`).
All GABA-SI-incident edges are drawn in deep blue; any non-SI excitatory
relay edges (shown as a faint grey reference layer) are included only
when they help place ppk23 inputs that feed the SIs; the focus is the
GABA sign-inverter layer itself.

## Build edge list from raw synapse counts

```{r edges}
ppk23_srcs <- channel_neurons[["ppk23"]]
ppk25_srcs <- channel_neurons[["ppk25"]]

# GABA sign-inverter pool (from setup_mspecific.R)
gaba_si_pool <- gaba_sign_inverter_types

# mal_subtypes is already filtered to mAL_m* after setup_mspecific.R
mal_m_subtypes_used <- mal_subtypes

raw_rows <- rownames(adj.matrix.raw)
raw_cols <- colnames(adj.matrix.raw)

# ppk23 + ppk25 sources -> GABA SIs (layer 1 -> 2)
src_rows_23 <- intersect(ppk23_srcs, raw_rows)
src_rows_25 <- intersect(ppk25_srcs, raw_rows)
si_cols     <- intersect(gaba_si_pool, raw_cols)

e_top_23 <- expand.grid(from = src_rows_23, to = si_cols,
                        stringsAsFactors = FALSE) %>%
  mutate(weight = mapply(function(a, b) adj.matrix.raw[a, b],
                         from, to),
         src_channel = "ppk23") %>%
  filter(is.finite(weight) & weight > 0)

e_top_25 <- expand.grid(from = src_rows_25, to = si_cols,
                        stringsAsFactors = FALSE) %>%
  mutate(weight = mapply(function(a, b) adj.matrix.raw[a, b],
                         from, to),
         src_channel = "ppk25") %>%
  filter(is.finite(weight) & weight > 0)

e_top <- bind_rows(e_top_23, e_top_25)

# GABA SIs -> mAL_m subtypes (layer 2 -> 3)
si_rows   <- intersect(gaba_si_pool, raw_rows)
mal_cols  <- intersect(mal_m_subtypes_used, raw_cols)

e_bot <- expand.grid(from = si_rows, to = mal_cols,
                     stringsAsFactors = FALSE) %>%
  mutate(weight = mapply(function(a, b) adj.matrix.raw[a, b],
                         from, to)) %>%
  filter(is.finite(weight) & weight > 0)

# Inter-SI connections (cross-talk within the GABA SI layer)
e_inter <- expand.grid(from = si_rows, to = si_cols,
                       stringsAsFactors = FALSE) %>%
  filter(from != to) %>%
  mutate(weight = mapply(function(a, b) adj.matrix.raw[a, b],
                         from, to)) %>%
  filter(is.finite(weight) & weight > 0)

cat("n edges ppk23 -> GABA SIs:", nrow(e_top_23), "\n")
cat("n edges ppk25 -> GABA SIs:", nrow(e_top_25), "\n")
cat("n edges GABA SIs -> mAL_m:", nrow(e_bot), "\n")
cat("n edges SI <-> SI (inter-layer):", nrow(e_inter), "\n")

# Throughput per SI (inflow + outflow, used for sizing + collapsing)
si_throughput <- data.frame(
  name = gaba_si_pool,
  stringsAsFactors = FALSE
) %>%
  mutate(
    in_w  = vapply(name, function(x) sum(e_top$weight[e_top$to  == x]),
                   numeric(1)),
    out_w = vapply(name, function(x) sum(e_bot$weight[e_bot$from == x]),
                   numeric(1)),
    total = in_w + out_w
  ) %>%
  arrange(desc(total))

cat("\nSI throughput (ppk23 in + mAL_m out, raw synapse counts):\n")
print(si_throughput)

# Collapse lowest-throughput SIs if the layer has > 8 nodes: keep the
# top 7 individually and bin the rest.
N_KEEP <- 7L
collapse_rule <- nrow(si_throughput) > 8
if (collapse_rule) {
  keep_si <- si_throughput$name[seq_len(min(N_KEEP, nrow(si_throughput)))]
} else {
  keep_si <- si_throughput$name
}
binned_si <- setdiff(si_throughput$name, keep_si)
cat("\nKeeping as individual middle nodes:",
    paste(keep_si, collapse = ", "), "\n")
if (length(binned_si) > 0) {
  cat("Collapsed into 'other GABA SIs':",
      paste(binned_si, collapse = ", "), "\n")
}

collapse_name <- "other GABA SIs"
collapse_map <- function(x) {
  ifelse(x %in% keep_si, x,
         ifelse(x %in% binned_si, collapse_name, x))
}

e_top <- e_top %>%
  mutate(to = collapse_map(to)) %>%
  group_by(from, to, src_channel) %>%
  summarise(weight = sum(weight), .groups = "drop")

e_bot <- e_bot %>%
  mutate(from = collapse_map(from)) %>%
  group_by(from, to) %>% summarise(weight = sum(weight), .groups = "drop")

e_inter <- e_inter %>%
  mutate(from = collapse_map(from), to = collapse_map(to)) %>%
  filter(from != to) %>%
  group_by(from, to) %>% summarise(weight = sum(weight), .groups = "drop")

# Final middle-layer node order: keep_si in descending throughput, then
# collapse bin on the right.
mid_order <- c(keep_si,
               if (length(binned_si) > 0) collapse_name else character(0))

# Per-node total throughput (for sizing)
node_throughput <- data.frame(name = mid_order, stringsAsFactors = FALSE) %>%
  mutate(
    in_w  = vapply(name, function(x) sum(e_top$weight[e_top$to  == x]),
                   numeric(1)),
    out_w = vapply(name, function(x) sum(e_bot$weight[e_bot$from == x]),
                   numeric(1)),
    total = in_w + out_w
  )
cat("\nMiddle-layer nodes (throughput):\n"); print(node_throughput)
```

## Node positions (hand-placed, 3 horizontal layers)

```{r positions}
even_x <- function(labels, xmin = 0, xmax = 1) {
  n <- length(labels)
  if (n == 1) return(setNames((xmin + xmax) / 2, labels))
  setNames(seq(xmin, xmax, length.out = n), labels)
}

# Wider horizontal spread for all layers to reduce edge-crowding
# Stacked top layer: ppk23 on the upper row, ppk25 just below it.
top_x_23 <- even_x(ppk23_srcs, 0.04, 0.96)
top_x_25 <- even_x(ppk25_srcs, 0.04, 0.96)
mid_x    <- even_x(mid_order,  0.06, 0.94)
bot_x    <- even_x(mal_m_subtypes_used, 0.00, 1.00)

Y_PPK23 <- 3.6
Y_PPK25 <- 2.9
Y_SI    <- 2.0
Y_MAL   <- 0.8

# Size the middle-layer nodes by total throughput
size_scale <- scales::rescale(node_throughput$total, to = c(9, 18))
names(size_scale) <- node_throughput$name

nodes <- bind_rows(
  data.frame(name = ppk23_srcs,          layer = "ppk23",
             x = top_x_23[ppk23_srcs],   y = Y_PPK23,
             size_pt = 7,
             stringsAsFactors = FALSE),
  data.frame(name = ppk25_srcs,          layer = "ppk25",
             x = top_x_25[ppk25_srcs],   y = Y_PPK25,
             size_pt = 7,
             stringsAsFactors = FALSE),
  data.frame(name = mid_order,           layer = "GABA_SI",
             x = mid_x[mid_order],       y = Y_SI,
             size_pt = unname(size_scale[mid_order]),
             stringsAsFactors = FALSE),
  data.frame(name = mal_m_subtypes_used, layer = "mAL_m",
             x = bot_x[mal_m_subtypes_used], y = Y_MAL,
             size_pt = 7,
             stringsAsFactors = FALSE)
)

# Note: if a name appears in both ppk23 and ppk25 (unlikely but possible),
# node rows will duplicate. Join for edges uses per-channel y below.

attach_xy <- function(edges, nodes) {
  edges %>%
    left_join(nodes %>% select(name, x, y) %>%
                distinct(name, .keep_all = TRUE) %>%
                rename(x_from = x, y_from = y),
              by = c("from" = "name")) %>%
    left_join(nodes %>% select(name, x, y) %>%
                distinct(name, .keep_all = TRUE) %>%
                rename(x_to = x, y_to = y),
              by = c("to" = "name"))
}

# Top edges: resolve source coords per channel (ppk23 row vs ppk25 row).
top_nodes_23 <- nodes %>% filter(layer == "ppk23") %>%
  select(name, x_from = x, y_from = y) %>% mutate(src_channel = "ppk23")
top_nodes_25 <- nodes %>% filter(layer == "ppk25") %>%
  select(name, x_from = x, y_from = y) %>% mutate(src_channel = "ppk25")
top_src_lookup <- bind_rows(top_nodes_23, top_nodes_25)

mid_lookup <- nodes %>% filter(layer == "GABA_SI") %>%
  select(name, x_to = x, y_to = y)

e_top_xy <- e_top %>%
  left_join(top_src_lookup, by = c("from" = "name", "src_channel")) %>%
  left_join(mid_lookup,     by = c("to"   = "name")) %>%
  mutate(edge_kind = ifelse(src_channel == "ppk23",
                            "ppk23_to_GABA_SI",
                            "ppk25_to_GABA_SI"))

e_bot_xy <- attach_xy(e_bot, nodes) %>%
  mutate(edge_kind = "GABA_SI_out",
         src_channel = NA_character_)
e_inter_xy <- attach_xy(e_inter, nodes) %>%
  mutate(edge_kind = "GABA_SI_inter",
         src_channel = NA_character_,
         # offset slightly so mid-row curves are visible
         y_from = Y_SI + 0.05,
         y_to   = Y_SI - 0.05)

edges_all <- bind_rows(e_top_xy, e_bot_xy, e_inter_xy) %>%
  filter(!is.na(x_from), !is.na(x_to))
```

## Colors

```{r colors}
edge_colors <- c(
  "ppk23_to_GABA_SI" = unname(panel_M_palette["ppk23"]),   # ppk23 -> SI
  "ppk25_to_GABA_SI" = "#2BA89F",  # ppk25 -> SI (teal)
  "GABA_SI_out"      = unname(panel_M_palette["gaba_si"]), # GABA SI -> mAL_m (inhibitory)
  "GABA_SI_inter"    = "#3B6CAE"   # SI <-> SI cross-talk
)

node_fill <- c(
  "ppk23"   = unname(panel_M_palette["ppk23"]),    # shared palette (red)
  "ppk25"   = unname(panel_M_palette["ppk25"]),    # shared palette (orange)
  "GABA_SI" = unname(panel_M_palette["gaba_si"]),  # shared palette (dark blue)
  "mAL_m"   = "grey85"
)
```

## Plot

```{r plot, fig.width=14, fig.height=8}
# Edge widths on a shared scale
all_w <- edges_all$weight
edges_all$lw <- scales::rescale(all_w, to = c(0.18, 1.8))

# Alpha: main inflow/outflow edges solid; SI-SI cross-talk lighter
edges_all$alpha_e <- ifelse(edges_all$edge_kind == "GABA_SI_inter",
                            0.55, 0.75)

# Draw in two passes: straight edges first, inter-SI curves second
e_straight <- edges_all %>% filter(edge_kind != "GABA_SI_inter")
e_curve    <- edges_all %>% filter(edge_kind == "GABA_SI_inter")

# Halo behind each middle-layer node
mid_pos <- nodes %>% filter(layer == "GABA_SI")

p <- ggplot() +
  # Halo behind each GABA SI middle node
  geom_point(data = mid_pos,
             aes(x = x, y = y, size = size_pt * 1.8),
             shape = 21, fill = "#D0E2F5", color = NA,
             alpha = 0.50, show.legend = FALSE) +
  # Straight edges (ppk23 -> SI, SI -> mAL_m)
  geom_segment(data = e_straight,
               aes(x = x_from, y = y_from, xend = x_to, yend = y_to,
                   color = edge_kind, linewidth = lw, alpha = alpha_e),
               lineend = "round") +
  # Curved inter-SI cross-talk edges
  geom_curve(data = e_curve,
             aes(x = x_from, y = y_from, xend = x_to, yend = y_to,
                 color = edge_kind, linewidth = lw, alpha = alpha_e),
             curvature = -0.25, lineend = "round") +
  scale_color_manual(values = edge_colors,
                     breaks = c("ppk23_to_GABA_SI",
                                "ppk25_to_GABA_SI",
                                "GABA_SI_out",
                                "GABA_SI_inter"),
                     labels = c("ppk23 -> GABA SI (to inhibitor)",
                                "ppk25 -> GABA SI (to inhibitor)",
                                "GABA SI -> mAL_m (inhibitory)",
                                "GABA SI <-> GABA SI (cross-talk)"),
                     name = "Edge type") +
  scale_linewidth_identity() +
  scale_alpha_identity() +
  scale_size_identity() +
  # Nodes
  geom_point(data = nodes %>% filter(layer != "GABA_SI"),
             aes(x = x, y = y, fill = layer, size = size_pt),
             shape = 21, color = "black", stroke = 0.4) +
  geom_point(data = mid_pos,
             aes(x = x, y = y, size = size_pt),
             shape = 21, fill = unname(panel_M_palette["gaba_si"]),
             color = "black", stroke = 1.1) +
  # Labels
  geom_text(data = nodes %>% filter(layer == "ppk23"),
            aes(x = x, y = y + 0.20, label = name),
            angle = 30, hjust = 0, size = 3.2, color = "#8A6A1F") +
  geom_text(data = nodes %>% filter(layer == "ppk25"),
            aes(x = x, y = y - 0.20, label = name),
            angle = -30, hjust = 1, size = 3.2, color = "#1F6F68") +
  geom_text(data = mid_pos,
            aes(x = x, y = y + 0.34, label = name),
            size = 3.6, color = unname(panel_M_palette["gaba_si"]), fontface = "bold") +
  geom_text(data = nodes %>% filter(layer == "mAL_m"),
            aes(x = x, y = y - 0.28, label = name),
            angle = 30, hjust = 1, size = 3.2) +
  scale_fill_manual(values = node_fill, name = "Layer",
                    breaks = c("ppk23", "ppk25", "GABA_SI", "mAL_m"),
                    labels = c("ppk23 sensory",
                               "ppk25 sensory",
                               "GABA sign-inverter layer",
                               "mAL_m (male-specific)")) +
  coord_cartesian(xlim = c(-0.05, 1.05), ylim = c(0.15, 4.25), clip = "off") +
  labs(
    caption = paste(
      panel_M_pool_caption,
      "NODES: middle-layer = individual GABA sign-inverter types; size ~ total throughput (ppk23+ppk25 inflow + mAL_m outflow).",
      "EDGES: thickness ~ raw synapse count.",
      "COLORS: gold = ppk23 -> SI; teal = ppk25 -> SI; deep blue = SI -> mAL_m (inhibitory); lighter blue = SI <-> SI cross-talk.",
      sep = "\n"
    )
  ) +
  theme_void(base_size = 13) +
  theme(
    legend.position   = "right",
    legend.box        = "vertical",
    legend.title      = element_text(size = rel(1.0), face = "bold"),
    legend.text       = element_text(size = rel(0.9)),
    plot.margin       = margin(24, 28, 18, 28),
    plot.caption      = element_text(size = rel(0.85), hjust = 0.5,
                                     color = "grey25",
                                     margin = margin(t = 10))
  ) +
  guides(
    fill  = guide_legend(override.aes = list(size = 5), order = 1),
    color = guide_legend(override.aes = list(linewidth = 1.5), order = 2),
    size  = "none"
  )

# Landscape 14 x 8
ggsave(file.path(pdf_dir, paste0(plot_name, ".pdf")),
       plot = p, width = 14, height = 8)
ggsave(file.path(png_dir, paste0(plot_name, ".png")),
       plot = p, width = 14, height = 8, dpi = 300)

print(p)
```

## Fan-out summary

```{r fanout}
si_fanout <- e_bot %>%
  group_by(from) %>%
  summarise(n_mal = n(), total = sum(weight), .groups = "drop") %>%
  arrange(desc(total))

cat("GABA SI layer fan-out onto mAL_m:\n")
print(si_fanout)
cat("\nTotal GABA SI -> mAL_m synapses:", sum(e_bot$weight), "\n")
cat("Distinct mAL_m targets reached:",
    length(unique(e_bot$to)), "/", length(mal_m_subtypes_used), "\n")
```
