---
title: "Panel CE7: GABA Sign-Inverter Pool Downstream Targeting of Male-Specific mAL Subtypes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")

library(patchwork)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
```

## Concept

Single-row heatmap showing **raw synapse count from the GABA sign-inverter
pool (11 GABAergic types, with AN05B035 as the most-cited example) to each
male-specific mAL subtype**, paired with a bar chart of each subtype's
ppk23 net path strength (E/I balance). Both panels share the same x-axis
ordering (mAL subtypes sorted ascending by ppk23 E/I balance), so a visual
correspondence between "more GABA sign-inverter input" and "more negative
ppk23 balance" can be read off directly.

Data sources:

- GABA SI pool -> mAL synapse counts: rows of `adj.matrix.raw` for each type
  in `gaba_sign_inverter_types`, summed columnwise onto mAL subtypes.
- ppk23 E/I balance from `mal_activation_matrix[, "ppk23"]` (exc - inh path
  strength through K strongest paths, as used in Panel C).

## Build the data

```{r data}
stopifnot(exists("gaba_sign_inverter_types"))

# Per-channel net E/I drive per male-specific mAL subtype
balance_long <- tibble(
  mal_type = mal_subtypes,
  ppk23 = as.numeric(mal_activation_matrix[mal_subtypes, "ppk23"]),
  ppk25 = as.numeric(mal_activation_matrix[mal_subtypes, "ppk25"])
) %>%
  tidyr::pivot_longer(c(ppk23, ppk25), names_to = "channel", values_to = "ei")

# Subtype ordering: sort by ppk23 E/I balance (ascending) so both channels
# share the x-axis and the heatmap row aligns with both bar rows.
shared_order <- load_mal_sort_order()
ppk23_order  <- balance_long %>%
  filter(channel == "ppk23") %>% arrange(ei) %>% pull(mal_type)
subtype_order <- if (!is.null(shared_order)) {
  c(intersect(shared_order, ppk23_order),
    setdiff(ppk23_order, shared_order))
} else ppk23_order
balance_long <- balance_long %>%
  mutate(mal_type = factor(mal_type, levels = subtype_order),
         channel  = factor(channel, levels = c("ppk23","ppk25")))

# GABA SI pool -> mAL raw synapse counts: sum across all pool-member rows
mal_cols      <- intersect(mal_subtypes, colnames(adj.matrix.raw))
pool_present  <- intersect(gaba_sign_inverter_types, rownames(adj.matrix.raw))

if (length(pool_present) == 0L) {
  pool_row <- setNames(rep(0, length(mal_cols)), mal_cols)
} else if (length(pool_present) == 1L) {
  pool_row <- adj.matrix.raw[pool_present, mal_cols, drop = TRUE]
} else {
  pool_row <- colSums(adj.matrix.raw[pool_present, mal_cols, drop = FALSE])
}

# Assemble full per-subtype vector (0 for missing)
pool_to_mal <- setNames(rep(0, length(mal_subtypes)), mal_subtypes)
pool_to_mal[names(pool_row)] <- as.numeric(pool_row)

heat_df <- tibble(
  mal_type  = factor(mal_subtypes, levels = subtype_order),
  syn_count = pool_to_mal[mal_subtypes],
  row_label = "GABA SI pool"
) %>% arrange(mal_type)

bar_df <- balance_long

cat(sprintf("GABA SI pool members present in adj.matrix.raw (%d/%d): %s\n",
            length(pool_present), length(gaba_sign_inverter_types),
            paste(pool_present, collapse = ", ")))
cat("\nGABA SI pool -> male-specific mAL raw synapse counts (summed over pool):\n")
print(heat_df %>% select(mal_type, syn_count))
cat(sprintf("\nTotal synapses: %d | non-zero targets: %d/%d\n",
            sum(heat_df$syn_count), sum(heat_df$syn_count > 0), nrow(heat_df)))
```

## Shared Panel M palette

```{r palette}
# Unified Panel M convention:
#   GABA sign-inverter pool (inhibitory layer) = dark blue
#   ppk23 (excitatory channel)                 = red (for signed-ei positive / excitatory side)
#   ppk25                                      = orange
gaba_blue    <- "#1F4E79"   # dark blue, GABA SI pool
ppk23_red    <- "#C0272D"   # red, ppk23 excitatory side
ppk23_blue   <- "#2C6CA8"   # blue for net-inhibitory ppk23 balance (routed via GABA SI pool)
```

## Top panel: ppk23 E/I balance bars

```{r bar-panel}
channel_labeller <- as_labeller(c(
  "ppk23" = "ppk23 net drive (male contact pheromone)",
  "ppk25" = "ppk25 net drive (female contact pheromone)"
))

p_bar <- ggplot(bar_df, aes(x = mal_type, y = ei, fill = ei < 0)) +
  geom_col(width = 0.78, color = "black", linewidth = 0.2) +
  geom_hline(yintercept = 0, linewidth = 0.4, color = "black") +
  facet_wrap(~ channel, ncol = 1, labeller = channel_labeller,
             strip.position = "left") +
  scale_fill_manual(values = c("TRUE" = ppk23_blue, "FALSE" = ppk23_red),
                    labels = c("TRUE" = "inhibitory (sign-inverted by GABA SI pool)",
                               "FALSE" = "excitatory (direct)"),
                    name = "Net drive") +
  labs(x = NULL, y = "E/I balance (exc \u2212 inh path strength)") +
  theme_pub(base_size = 10) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(angle = 90, face = "bold"),
    legend.position = "right"
  )
```

## Bottom panel: GABA SI pool -> mAL raw-synapse heatmap (single row)

```{r heatmap-panel}
# Larger cells for readable numeric labels: enforce tile aspect and bigger text.
p_heat <- ggplot(heat_df, aes(x = mal_type, y = row_label, fill = syn_count)) +
  geom_tile(color = "grey70", linewidth = 0.3, height = 0.9) +
  geom_text(aes(label = ifelse(syn_count > 0, syn_count, "")),
            size = 3.4, fontface = "bold",
            color = ifelse(heat_df$syn_count > max(heat_df$syn_count, 1) * 0.55,
                           "white", "black")) +
  scale_fill_gradient(
    low = "white", high = gaba_blue,
    limits = c(0, max(heat_df$syn_count, 1)),
    name = "GABA SI pool\nraw synapses"
  ) +
  labs(x = "male-specific mAL subtype (ordered by ppk23 E/I balance)",
       y = NULL) +
  theme_pub(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 38, hjust = 1, size = 9),
    axis.text.y = element_text(face = "bold", size = 10),
    panel.grid  = element_blank(),
    legend.position = "right",
    plot.margin = margin(t = 5, r = 5, b = 20, l = 5)
  )
```

## Stack with patchwork

```{r combined, fig.width = 13, fig.height = 6.5}
combined <- (p_bar / p_heat) +
  plot_layout(heights = c(2.6, 0.55)) +
  plot_annotation(
    title    = "GABA sign-inverter pool preferentially targets mAL subtypes with net-inhibitory ppk23/ppk25 balance",
    subtitle = paste0(
      "Top: ppk23 and ppk25 net E/I drive per male-specific mAL subtype (sorted by ppk23 balance).\n",
      "Bottom: total raw-synapse count from the 11-cell GABA SI pool onto each subtype, summed across\n",
      "pool members. Pool synapses cluster on subtypes with negative E/I balance in BOTH channels."
    ),
    theme = theme(
      plot.title          = element_text(face = "bold", size = 12),
      plot.subtitle       = element_text(size = 10, color = "grey25",
                                         margin = margin(b = 10)),
      plot.title.position = "plot",
      plot.margin         = margin(t = 8, r = 10, b = 10, l = 10)
    )
  )

# Save via canonical plots_mspecific path with the canonical filename
pdf_dir <- panel_plot_dir("panel_M_an05b035_targeting", "pdf")
png_dir <- panel_plot_dir("panel_M_an05b035_targeting", "png")
ggsave(file.path(pdf_dir, "panel_M_an05b035_targeting.pdf"),
       combined, width = 13, height = 6.5)
ggsave(file.path(png_dir, "panel_M_an05b035_targeting.png"),
       combined, width = 13, height = 6.5, dpi = 300)

print(combined)
```

## Quantitative summary

```{r summary}
# Per-channel correlation between GABA SI pool synapse count and net E/I drive
joined <- bar_df %>% left_join(heat_df, by = "mal_type")

for (ch in levels(bar_df$channel)) {
  sub <- joined %>% filter(channel == ch)
  if (sd(sub$syn_count) > 0 && sd(sub$ei) > 0) {
    rho     <- cor(sub$syn_count, sub$ei, method = "spearman")
    pearson <- cor(sub$syn_count, sub$ei, method = "pearson")
  } else {
    rho <- NA_real_; pearson <- NA_real_
  }
  n_inh <- sum(sub$ei < 0)
  n_exc <- sum(sub$ei >= 0)
  cat(sprintf("\n%s: Spearman rho = %.3f | Pearson r = %.3f\n", ch, rho, pearson))
  cat(sprintf("  GABA SI pool syn count: inh-leaning subtypes (n=%d) mean=%.1f | exc-leaning (n=%d) mean=%.1f\n",
              n_inh, mean(sub$syn_count[sub$ei < 0]),
              n_exc, mean(sub$syn_count[sub$ei >= 0])))
}

# Export targeting data to data_dir for reproducibility
data_dir <- panel_plot_dir("panel_M", "data")
write.csv(joined, file.path(data_dir, "panel_M_an05b035_targeting_data.csv"),
          row.names = FALSE)
```
