---
title: "Figure 5 CE9: Example ppk23 -> mAL path traces (GABA sign-inverter pool)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggrepel)
```

## Goal

Concrete example traces of the routing differences underlying ppk23 sign
reversal. We pick three illustrative male-specific mAL subtypes:

- one with the strongest GABA-sign-inverter-routed inhibitory drive (most
  negative ppk23 E/I balance),
- one balanced,
- one with the strongest direct excitatory drive (most positive E/I balance).

For each, we draw the top 10 strongest ppk23 -> mAL paths as a small layered
network. Each path is drawn as a sequence of `geom_segment` edges between
nodes placed at their hop level. Edge color encodes the route type
(red = excitatory neighborhood; blue = path passes through any member of the
GABA sign-inverter pool). Edge width is proportional to path strength.

## Per-subtype stats: GABA SI fraction and ppk23 E/I balance

```{r subtype-stats}
ppk23_paths <- mal_paths %>%
  filter(modality == "ppk23", end %in% mal_subtypes) %>%
  mutate(
    nodes_list  = strsplit(path, " -> ", fixed = TRUE),
    via_gaba_si = vapply(nodes_list,
                         function(x) any(x %in% gaba_sign_inverter_types),
                         logical(1))
  )

subtype_stats <- ppk23_paths %>%
  group_by(end) %>%
  summarise(
    total_strength  = sum(strength, na.rm = TRUE),
    gaba_si_strength = sum(strength[via_gaba_si], na.rm = TRUE),
    gaba_si_fraction = ifelse(total_strength > 0, gaba_si_strength / total_strength, 0),
    n_paths         = n(),
    .groups = "drop"
  )

exc <- mal_strength_exc[mal_subtypes, "ppk23"]
inh <- mal_strength_inh[mal_subtypes, "ppk23"]
ei_df <- tibble(
  end = mal_subtypes,
  exc = exc,
  inh = inh,
  ei_balance = ifelse(exc + inh > 0, (exc - inh) / (exc + inh), 0)
)

stats_df <- subtype_stats %>%
  left_join(ei_df, by = "end") %>%
  filter(n_paths >= 5, total_strength > 0) %>%
  arrange(ei_balance)

print(stats_df)
```

## Pick three illustrative subtypes

```{r pick-examples}
# We rank by GABA-sign-inverter routing fraction -- the direct proxy for how
# much of the ppk23 drive is routed through the GABAergic sign-inverter pool.
# (E/I balance from mal_strength_exc / mal_strength_inh is reported above each
#  plot as the second stat.)

# Most ppk23-INHIBITED route: highest GABA SI fraction
sub_inh <- stats_df %>% arrange(desc(gaba_si_fraction)) %>% slice(1) %>% pull(end)

# Most ppk23-EXCITED route: lowest GABA SI fraction (paths bypass the
# sign-inverter pool and use direct / AN09B017 excitatory relays)
sub_exc <- stats_df %>% arrange(gaba_si_fraction)       %>% slice(1) %>% pull(end)

# Balanced: GABA SI fraction closest to the population median
med_frac <- median(stats_df$gaba_si_fraction)
sub_bal <- stats_df %>%
  filter(!end %in% c(sub_inh, sub_exc)) %>%
  mutate(d = abs(gaba_si_fraction - med_frac)) %>%
  arrange(d) %>%
  slice(1) %>% pull(end)

chosen <- c(sub_inh, sub_bal, sub_exc)
cat("Chosen example subtypes:\n")
print(stats_df %>% filter(end %in% chosen) %>%
        mutate(role = case_when(
          end == sub_inh ~ "most ppk23-inhibited (highest GABA SI fraction)",
          end == sub_exc ~ "most ppk23-excited (lowest GABA SI fraction)",
          TRUE           ~ "balanced (near-median GABA SI fraction)"
        )) %>% arrange(desc(gaba_si_fraction)))
```

## Build path-trace data for one subtype

```{r build-trace-fn}
# Helper: given a subtype, pull top-K strongest ppk23 paths, assign each
# unique node a canonical (x, y) position, and return an edge-level data
# frame ready for geom_segment.
build_trace <- function(subtype, topk = 10) {
  sub_paths <- ppk23_paths %>%
    filter(end == subtype) %>%
    arrange(desc(strength)) %>%
    slice_head(n = topk) %>%
    mutate(path_id = row_number())

  if (nrow(sub_paths) == 0) return(list(edges = NULL, nodes = NULL, paths = sub_paths))

  # -- Collect hop-level node info across all top-K paths
  max_hops <- max(vapply(sub_paths$nodes_list, length, integer(1)))

  node_long <- purrr::imap_dfr(sub_paths$nodes_list, function(nodes, i) {
    tibble(
      path_id = sub_paths$path_id[i],
      hop     = seq_along(nodes),
      node    = nodes
    )
  })

  # -- Canonical hop per unique node: most common hop across top-K paths,
  # ties broken by the smallest hop. This guarantees each unique node has
  # exactly ONE (x, y) in the final layout.
  node_hop_counts <- node_long %>%
    count(node, hop, name = "n_at_hop")

  canonical_hop <- node_hop_counts %>%
    group_by(node) %>%
    arrange(desc(n_at_hop), hop, .by_group = TRUE) %>%
    slice(1) %>%
    ungroup() %>%
    select(node, hop)

  # Node occurrence frequency (for size / y-ordering) -- count each unique
  # (path_id, node) pair so a node that appears twice in the same path
  # still counts once for that path.
  node_freq <- node_long %>%
    distinct(path_id, node) %>%
    count(node, name = "n_paths_with_node")

  node_layout <- canonical_hop %>%
    left_join(node_freq, by = "node") %>%
    group_by(hop) %>%
    arrange(desc(n_paths_with_node), node, .by_group = TRUE) %>%
    mutate(
      n_in_hop = n(),
      y_rank   = row_number(),   # 1 = most used within this hop
      y_pos    = {
        k <- y_rank
        # produce 0, +1, -1, +2, -2, ... pattern
        sgn <- ifelse(k %% 2 == 0, 1, -1)
        mag <- ceiling((k - 1) / 2)
        sgn * mag
      }
    ) %>%
    ungroup() %>%
    mutate(
      x_pos = hop - 1   # hop 1 (source) at x=0, increases rightward
    ) %>%
    select(hop, node, x_pos, y_pos, n_paths_with_node)

  # Normalize y_pos per hop so extrema sit in ~[-1, 1]
  node_layout <- node_layout %>%
    group_by(hop) %>%
    mutate(
      y_max = max(abs(y_pos)),
      y_pos = if_else(y_max > 0, y_pos / y_max, 0)
    ) %>%
    ungroup() %>%
    select(-y_max)

  # Edges: for each path, connect consecutive nodes; color by route; width
  # by strength. Use each node's CANONICAL (x_pos, y_pos) -- this may
  # occasionally produce visually backward edges when two canonical-same-hop
  # nodes connect, which is the correct deduped picture.
  edges <- purrr::map_dfr(seq_len(nrow(sub_paths)), function(i) {
    nodes <- sub_paths$nodes_list[[i]]
    if (length(nodes) < 2) return(tibble())
    from_nodes <- nodes[-length(nodes)]
    to_nodes   <- nodes[-1]

    via_gaba <- sub_paths$via_gaba_si[i]

    tibble(
      path_id  = sub_paths$path_id[i],
      strength = sub_paths$strength[i],
      from     = from_nodes,
      to       = to_nodes,
      route    = factor(if_else(via_gaba, "via GABA sign-inverter", "direct / other relay"),
                        levels = c("direct / other relay", "via GABA sign-inverter"))
    )
  })

  # Join canonical layout coordinates (lookup by node identity, not by hop)
  node_coords <- node_layout %>% select(node, x_pos, y_pos)
  edges <- edges %>%
    left_join(node_coords %>% rename(x = x_pos, y = y_pos),
              by = c("from" = "node")) %>%
    left_join(node_coords %>% rename(xend = x_pos, yend = y_pos),
              by = c("to" = "node"))

  list(edges = edges, nodes = node_layout, paths = sub_paths, max_hops = max_hops)
}
```

## Plotting helper

```{r plot-fn}
# Shorten long neuron names for readability
short_label <- function(x) {
  out <- sub("ORN_", "", x)
  out <- ifelse(nchar(out) > 12, paste0(substr(out, 1, 11), "..."), out)
  out
}

# Unified Panel M palette: GABA sign-inverter route = dark blue, direct = red.
route_colors <- c(
  "direct / other relay"     = "#C0272D",  # red (direct excitatory)
  "via GABA sign-inverter"   = "#1F4E79"   # dark blue (GABA sign-inverter pool)
)

plot_trace <- function(trace, subtype, ei_balance, gaba_si_fraction) {
  edges <- trace$edges
  nodes <- trace$nodes
  max_hops <- trace$max_hops

  # Determine a readable title
  title_lab <- sprintf(
    "%s  \u2014  ppk23 E/I = %+.2f, GABA SI frac = %.0f%%",
    subtype, ei_balance, 100 * gaba_si_fraction
  )

  # Point size: nodes that are on more paths get drawn larger
  nodes <- nodes %>% mutate(size_pt = 4 + 1.6 * sqrt(n_paths_with_node))

  # Simpler path thickness scale: two discrete tiers based on median split.
  if (nrow(edges) > 0) {
    s_med <- stats::median(edges$strength, na.rm = TRUE)
    edges <- edges %>%
      mutate(lw = ifelse(strength >= s_med, 1.8, 0.8))
  }

  p <- ggplot() +
    # Path edges (drawn before nodes so nodes sit on top)
    {
      if (!is.null(edges) && nrow(edges) > 0) {
        geom_segment(
          data = edges,
          aes(x = x, y = y, xend = xend, yend = yend, color = route),
          linewidth = edges$lw, alpha = 0.75, lineend = "round"
        )
      }
    } +
    geom_point(
      data = nodes,
      aes(x = x_pos, y = y_pos, size = size_pt),
      color = "grey25", fill = "white", shape = 21, stroke = 0.4
    ) +
    geom_text_repel(
      data = nodes,
      aes(x = x_pos, y = y_pos, label = short_label(node)),
      size = 3.2, min.segment.length = 0, seed = 1,
      segment.size = 0.25, max.overlaps = Inf,
      box.padding = 0.45, point.padding = 0.25,
      force = 2
    ) +
    scale_color_manual(values = route_colors, name = "Route", drop = FALSE) +
    scale_size_identity() +
    scale_x_continuous(
      breaks = 0:(max_hops - 1),
      labels = c("ppk23 source",
                 paste("hop", seq_len(max_hops - 2)),
                 "mAL target")[1:max_hops],
      expand = expansion(add = 0.25)
    ) +
    scale_y_continuous(expand = expansion(add = 0.55)) +
    labs(x = NULL, y = NULL, title = title_lab) +
    theme_pub(base_size = 10) +
    theme(
      plot.title     = element_text(size = rel(1), face = "bold", hjust = 0.5),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.text.y    = element_blank(),
      axis.text.x    = element_text(size = rel(0.9)),
      legend.position = "bottom"
    )

  p
}
```

## Build the three sub-plots

```{r build-plots, fig.width=16, fig.height=6}
subtype_roles <- list(
  list(subtype = sub_inh, role = "(a) Most GABA-SI-routed (ppk23-inhibited)"),
  list(subtype = sub_bal, role = "(b) Balanced (near-median GABA-SI fraction)"),
  list(subtype = sub_exc, role = "(c) Least GABA-SI-routed (ppk23-excited)")
)
# Dedupe in case pickers collapsed
seen <- c()
subtype_roles <- Filter(function(x) {
  keep <- !(x$subtype %in% seen)
  if (keep) seen <<- c(seen, x$subtype)
  keep
}, subtype_roles)

stats_df_lookup <- stats_df

plots <- lapply(subtype_roles, function(info) {
  subtype <- info$subtype
  st <- stats_df_lookup %>% filter(end == subtype)
  tr <- build_trace(subtype, topk = 10)
  p <- plot_trace(tr, subtype,
                  ei_balance = st$ei_balance,
                  gaba_si_fraction = st$gaba_si_fraction)
  # Add role label as a subtitle above the EI-balance title
  p + labs(subtitle = info$role) +
    theme(plot.subtitle = element_text(size = rel(0.95), face = "italic",
                                       hjust = 0.5, color = "grey30"))
})

# Arrange horizontally with shared legend
p_combined <- wrap_plots(plots, nrow = 1, guides = "collect") +
  plot_annotation(
    title    = "GABA sign-inverter (blue) vs direct (red) routes produce opposite ppk23 drive polarities",
    subtitle = "Top-10 ppk23 \u2192 mAL paths for three example male-specific mAL subtypes spanning the E/I axis.",
    theme = theme(
      plot.title    = element_text(face = "bold", size = 12, hjust = 0.5),
      plot.subtitle = element_text(size = 10, color = "grey25", hjust = 0.5)
    )
  ) & theme(legend.position = "bottom")

print(p_combined)
```

## Save outputs

```{r save}
pdf_dir <- panel_plot_dir("panel_M_example_traces", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
png_dir <- panel_plot_dir("panel_M_example_traces", "png")

ggsave(file.path(pdf_dir, "panel_M_example_traces.pdf"),
       plot = p_combined, width = 16, height = 6)
ggsave(file.path(png_dir, "panel_M_example_traces.png"),
       plot = p_combined, width = 16, height = 6, dpi = 300)

# Export the edge-level table for each chosen subtype
trace_tables <- purrr::imap_dfr(subtype_roles, function(info, .) {
  tr <- build_trace(info$subtype, topk = 10)
  if (is.null(tr$edges)) return(tibble())
  tr$edges %>% mutate(subtype = info$subtype, role = info$role)
})
write.csv(trace_tables,
          file.path(data_dir, "panel_M_example_traces_edges.csv"),
          row.names = FALSE)
```

## Narrative

```{r narrative, echo=FALSE}
row_inh <- stats_df_lookup %>% filter(end == sub_inh)
row_bal <- stats_df_lookup %>% filter(end == sub_bal)
row_exc <- stats_df_lookup %>% filter(end == sub_exc)

narrative <- sprintf(
"PANEL CE9: Example ppk23 -> mAL path traces showing GABA sign-inverter pool routing
==============================================================================

WHAT THIS PANEL SHOWS
---------------------
Three small example networks, one per mAL_m subtype. For each subtype we draw
the top 10 strongest ppk23 -> mAL paths (from `mal_paths`, K=%d per channel).
Nodes are placed at their canonical hop level along the x-axis (source ppk23
sensory neuron on the left, mAL target on the right); within each hop we
stagger in y to reduce edge crossings. Each unique node gets exactly ONE
(x, y) position (canonical hop = most common hop across top-K paths). Edges
are colored by route -- red when the path does NOT traverse any member of
the GABA sign-inverter pool (direct / AN09B017-type excitatory relays) and
blue when the path passes through at least one GABA sign-inverter
(GABA SI pool: %s).
Edge width scales with path strength; node size scales with how many of the
top-10 paths pass through that node.

CHOSEN EXAMPLES
---------------
* Most ppk23-INHIBITED:  %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)
* Balanced:              %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)
* Most ppk23-EXCITED:    %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)

These three subtypes were picked automatically from `mal_paths`/`mal_strength_*`
to span the extremes and midpoint of ppk23 E/I balance among male-specific
mAL subtypes.

INTERPRETATION
--------------
Visualizing only the top-10 paths per target makes the routing asymmetry
explicit: the ppk23-inhibited subtype's trace is dominated by BLUE edges
that converge on central GABA-sign-inverter nodes before reaching the mAL
target, while the ppk23-excited subtype's trace is almost entirely RED,
with paths sidestepping the GABA SI pool and flowing through direct /
AN09B017-type excitatory relays. The balanced subtype shows an intermediate
mix. This is the concrete per-subtype picture of the sign inversion that
panels C and E summarize in aggregate.

PLACE IN THE FIGURE 5 STORY
----------------------------
Panel CE9 is the mechanistic 'zoom-in' companion to CE1's aggregate routing
fractions and CE2's GABA-SI-fraction vs E/I regression: it shows WHICH
paths carry ppk23 drive into each of the three example mALs, and where in
that path the sign inversion happens.
",
n_paths_mal,
paste(gaba_sign_inverter_types, collapse = ", "),
row_inh$end, row_inh$ei_balance, 100 * row_inh$gaba_si_fraction,
row_bal$end, row_bal$ei_balance, 100 * row_bal$gaba_si_fraction,
row_exc$end, row_exc$ei_balance, 100 * row_exc$gaba_si_fraction
)

writeLines(narrative, file.path(data_dir, "panel_M_example_traces_narrative.txt"))
cat(narrative)
```
