---
title: "Panel M: Example ppk25 -> mAL path traces (GABA sign-inverter pool, F-cell channel)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggrepel)
```

## Goal

F-cell (ppk25, female contact pheromone) counterpart of `M9_example_traces`.
Same construction as the ppk23 panel but drawn from `mal_paths` filtered on
`modality == "ppk25"`. Three illustrative male-specific mAL subtypes are
chosen to span the ppk25 GABA-sign-inverter-routing axis:

- highest GABA SI fraction (most ppk25-inhibited),
- balanced (near-median fraction),
- lowest GABA SI fraction (most ppk25-excited).

For each, the top 10 strongest ppk25 -> mAL paths are drawn as a small
layered network. Edge color encodes route (red = direct/other relay;
blue = passes through any member of the GABA sign-inverter pool). Edge
width scales with path strength.

## Per-subtype stats: GABA SI fraction and ppk25 E/I balance

```{r subtype-stats}
ppk25_paths <- mal_paths %>%
  filter(modality == "ppk25", end %in% mal_subtypes) %>%
  mutate(
    nodes_list  = strsplit(path, " -> ", fixed = TRUE),
    via_gaba_si = vapply(nodes_list,
                         function(x) any(x %in% gaba_sign_inverter_types),
                         logical(1))
  )

subtype_stats <- ppk25_paths %>%
  group_by(end) %>%
  summarise(
    total_strength   = sum(strength, na.rm = TRUE),
    gaba_si_strength = sum(strength[via_gaba_si], na.rm = TRUE),
    gaba_si_fraction = ifelse(total_strength > 0, gaba_si_strength / total_strength, 0),
    n_paths          = n(),
    .groups = "drop"
  )

exc <- mal_strength_exc[mal_subtypes, "ppk25"]
inh <- mal_strength_inh[mal_subtypes, "ppk25"]
ei_df <- tibble(
  end = mal_subtypes,
  exc = exc,
  inh = inh,
  ei_balance = ifelse(exc + inh > 0, (exc - inh) / (exc + inh), 0)
)

stats_df <- subtype_stats %>%
  left_join(ei_df, by = "end") %>%
  filter(n_paths >= 5, total_strength > 0) %>%
  arrange(ei_balance)

print(stats_df)
```

## Pick three illustrative subtypes

```{r pick-examples}
sub_inh <- stats_df %>% arrange(desc(gaba_si_fraction)) %>% slice(1) %>% pull(end)
sub_exc <- stats_df %>% arrange(gaba_si_fraction)       %>% slice(1) %>% pull(end)

med_frac <- median(stats_df$gaba_si_fraction)
sub_bal <- stats_df %>%
  filter(!end %in% c(sub_inh, sub_exc)) %>%
  mutate(d = abs(gaba_si_fraction - med_frac)) %>%
  arrange(d) %>%
  slice(1) %>% pull(end)

chosen <- c(sub_inh, sub_bal, sub_exc)
cat("Chosen example subtypes (ppk25 channel):\n")
print(stats_df %>% filter(end %in% chosen) %>%
        mutate(role = case_when(
          end == sub_inh ~ "most ppk25-inhibited (highest GABA SI fraction)",
          end == sub_exc ~ "most ppk25-excited (lowest GABA SI fraction)",
          TRUE           ~ "balanced (near-median GABA SI fraction)"
        )) %>% arrange(desc(gaba_si_fraction)))
```

## Build path-trace data for one subtype

```{r build-trace-fn}
build_trace <- function(subtype, topk = 10) {
  sub_paths <- ppk25_paths %>%
    filter(end == subtype) %>%
    arrange(desc(strength)) %>%
    slice_head(n = topk) %>%
    mutate(path_id = row_number())

  if (nrow(sub_paths) == 0) return(list(edges = NULL, nodes = NULL, paths = sub_paths))

  max_hops <- max(vapply(sub_paths$nodes_list, length, integer(1)))

  node_long <- purrr::imap_dfr(sub_paths$nodes_list, function(nodes, i) {
    tibble(
      path_id = sub_paths$path_id[i],
      hop     = seq_along(nodes),
      node    = nodes
    )
  })

  # Canonical hop per unique node: most common hop across top-K paths,
  # ties broken by smallest hop -> each node appears at exactly one (x, y).
  node_hop_counts <- node_long %>%
    count(node, hop, name = "n_at_hop")

  canonical_hop <- node_hop_counts %>%
    group_by(node) %>%
    arrange(desc(n_at_hop), hop, .by_group = TRUE) %>%
    slice(1) %>%
    ungroup() %>%
    select(node, hop)

  node_freq <- node_long %>%
    distinct(path_id, node) %>%
    count(node, name = "n_paths_with_node")

  node_layout <- canonical_hop %>%
    left_join(node_freq, by = "node") %>%
    group_by(hop) %>%
    arrange(desc(n_paths_with_node), node, .by_group = TRUE) %>%
    mutate(
      n_in_hop = n(),
      y_rank   = row_number(),
      y_pos    = {
        k <- y_rank
        sgn <- ifelse(k %% 2 == 0, 1, -1)
        mag <- ceiling((k - 1) / 2)
        sgn * mag
      }
    ) %>%
    ungroup() %>%
    mutate(x_pos = hop - 1) %>%
    select(hop, node, x_pos, y_pos, n_paths_with_node)

  node_layout <- node_layout %>%
    group_by(hop) %>%
    mutate(
      y_max = max(abs(y_pos)),
      y_pos = if_else(y_max > 0, y_pos / y_max, 0)
    ) %>%
    ungroup() %>%
    select(-y_max)

  edges <- purrr::map_dfr(seq_len(nrow(sub_paths)), function(i) {
    nodes <- sub_paths$nodes_list[[i]]
    if (length(nodes) < 2) return(tibble())
    from_nodes <- nodes[-length(nodes)]
    to_nodes   <- nodes[-1]
    via_gaba <- sub_paths$via_gaba_si[i]

    tibble(
      path_id  = sub_paths$path_id[i],
      strength = sub_paths$strength[i],
      from     = from_nodes,
      to       = to_nodes,
      route    = factor(if_else(via_gaba, "via GABA sign-inverter", "direct / other relay"),
                        levels = c("direct / other relay", "via GABA sign-inverter"))
    )
  })

  # Join canonical layout by node identity (not per-path hop)
  node_coords <- node_layout %>% select(node, x_pos, y_pos)
  edges <- edges %>%
    left_join(node_coords %>% rename(x = x_pos, y = y_pos),
              by = c("from" = "node")) %>%
    left_join(node_coords %>% rename(xend = x_pos, yend = y_pos),
              by = c("to" = "node"))

  list(edges = edges, nodes = node_layout, paths = sub_paths, max_hops = max_hops)
}
```

## Plotting helper

```{r plot-fn}
short_label <- function(x) {
  out <- sub("ORN_", "", x)
  out <- ifelse(nchar(out) > 12, paste0(substr(out, 1, 11), "..."), out)
  out
}

route_colors <- c(
  "direct / other relay"   = "#C0272D",
  "via GABA sign-inverter" = "#1F4E79"
)

plot_trace <- function(trace, subtype, ei_balance, gaba_si_fraction) {
  edges <- trace$edges
  nodes <- trace$nodes
  max_hops <- trace$max_hops

  title_lab <- sprintf(
    "%s  \u2014  ppk25 E/I = %+.2f, GABA SI frac = %.0f%%",
    subtype, ei_balance, 100 * gaba_si_fraction
  )

  nodes <- nodes %>% mutate(size_pt = 4 + 1.6 * sqrt(n_paths_with_node))

  if (nrow(edges) > 0) {
    s_med <- stats::median(edges$strength, na.rm = TRUE)
    edges <- edges %>%
      mutate(lw = ifelse(strength >= s_med, 1.8, 0.8))
  }

  p <- ggplot() +
    {
      if (!is.null(edges) && nrow(edges) > 0) {
        geom_segment(
          data = edges,
          aes(x = x, y = y, xend = xend, yend = yend, color = route),
          linewidth = edges$lw, alpha = 0.75, lineend = "round"
        )
      }
    } +
    geom_point(
      data = nodes,
      aes(x = x_pos, y = y_pos, size = size_pt),
      color = "grey25", fill = "white", shape = 21, stroke = 0.4
    ) +
    geom_text_repel(
      data = nodes,
      aes(x = x_pos, y = y_pos, label = short_label(node)),
      size = 3.2, min.segment.length = 0, seed = 1,
      segment.size = 0.25, max.overlaps = Inf,
      box.padding = 0.45, point.padding = 0.25,
      force = 2
    ) +
    scale_color_manual(values = route_colors, name = "Route", drop = FALSE) +
    scale_size_identity() +
    scale_x_continuous(
      breaks = 0:(max_hops - 1),
      labels = c("ppk25 source",
                 paste("hop", seq_len(max_hops - 2)),
                 "mAL target")[1:max_hops],
      expand = expansion(add = 0.25)
    ) +
    scale_y_continuous(expand = expansion(add = 0.55)) +
    labs(x = NULL, y = NULL, title = title_lab) +
    theme_pub(base_size = 10) +
    theme(
      plot.title     = element_text(size = rel(1), face = "bold", hjust = 0.5),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.text.y    = element_blank(),
      axis.text.x    = element_text(size = rel(0.9)),
      legend.position = "bottom"
    )

  p
}
```

## Build the three sub-plots

```{r build-plots, fig.width=16, fig.height=6}
subtype_roles <- list(
  list(subtype = sub_inh, role = "(a) Most GABA-SI-routed (ppk25-inhibited)"),
  list(subtype = sub_bal, role = "(b) Balanced (near-median GABA-SI fraction)"),
  list(subtype = sub_exc, role = "(c) Least GABA-SI-routed (ppk25-excited)")
)
seen <- c()
subtype_roles <- Filter(function(x) {
  keep <- !(x$subtype %in% seen)
  if (keep) seen <<- c(seen, x$subtype)
  keep
}, subtype_roles)

stats_df_lookup <- stats_df

plots <- lapply(subtype_roles, function(info) {
  subtype <- info$subtype
  st <- stats_df_lookup %>% filter(end == subtype)
  tr <- build_trace(subtype, topk = 10)
  p <- plot_trace(tr, subtype,
                  ei_balance = st$ei_balance,
                  gaba_si_fraction = st$gaba_si_fraction)
  p + labs(subtitle = info$role) +
    theme(plot.subtitle = element_text(size = rel(0.95), face = "italic",
                                       hjust = 0.5, color = "grey30"))
})

p_combined <- wrap_plots(plots, nrow = 1, guides = "collect") +
  plot_annotation(
    title    = "ppk25 (female contact) channel: GABA sign-inverter (blue) vs direct (red) routes",
    subtitle = "Top-10 ppk25 \u2192 mAL paths for three example male-specific mAL subtypes spanning the E/I axis.",
    theme = theme(
      plot.title    = element_text(face = "bold", size = 12, hjust = 0.5),
      plot.subtitle = element_text(size = 10, color = "grey25", hjust = 0.5)
    )
  ) & theme(legend.position = "bottom")

print(p_combined)
```

## Save outputs

```{r save}
pdf_dir <- panel_plot_dir("panel_M_example_traces_ppk25", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
png_dir <- panel_plot_dir("panel_M_example_traces_ppk25", "png")

ggsave(file.path(pdf_dir, "panel_M_example_traces_ppk25.pdf"),
       plot = p_combined, width = 16, height = 6)
ggsave(file.path(png_dir, "panel_M_example_traces_ppk25.png"),
       plot = p_combined, width = 16, height = 6, dpi = 300)

trace_tables <- purrr::imap_dfr(subtype_roles, function(info, .) {
  tr <- build_trace(info$subtype, topk = 10)
  if (is.null(tr$edges)) return(tibble())
  tr$edges %>% mutate(subtype = info$subtype, role = info$role)
})
write.csv(trace_tables,
          file.path(data_dir, "panel_M_example_traces_ppk25_edges.csv"),
          row.names = FALSE)
```

## Narrative

```{r narrative, echo=FALSE}
row_inh <- stats_df_lookup %>% filter(end == sub_inh)
row_bal <- stats_df_lookup %>% filter(end == sub_bal)
row_exc <- stats_df_lookup %>% filter(end == sub_exc)

narrative <- sprintf(
"PANEL M: Example ppk25 -> mAL path traces (F-cell channel, GABA sign-inverter pool)
==============================================================================

WHAT THIS PANEL SHOWS
---------------------
F-cell counterpart of the ppk23 example traces. For three mAL_m subtypes we
draw the top 10 strongest ppk25 -> mAL paths. Nodes are placed at their
canonical hop level along x (ppk25 source left, mAL target right); each
unique node appears at exactly ONE (x, y) (canonical hop = most common hop
across top-K paths). Edges are colored by route: red = path does NOT
traverse any GABA sign-inverter; blue = path passes through at least one
member of the GABA sign-inverter pool (%s). Edge width scales with path
strength; node size scales with how many of the top-10 paths pass through
that node.

CHOSEN EXAMPLES
---------------
* Most ppk25-INHIBITED:  %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)
* Balanced:              %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)
* Most ppk25-EXCITED:    %s  (E/I = %+.2f, GABA SI fraction = %.0f%%)

INTERPRETATION
--------------
Together with the ppk23 traces, this panel shows that the GABA sign-inverter
pool sits on the sign-inverting route for BOTH pheromone channels. The
per-subtype distribution of GABA-SI routing differs between ppk23 and ppk25,
which is what produces the channel-specific E/I asymmetries summarized in
panels C and E.
",
paste(gaba_sign_inverter_types, collapse = ", "),
row_inh$end, row_inh$ei_balance, 100 * row_inh$gaba_si_fraction,
row_bal$end, row_bal$ei_balance, 100 * row_bal$gaba_si_fraction,
row_exc$end, row_exc$ei_balance, 100 * row_exc$gaba_si_fraction
)

writeLines(narrative, file.path(data_dir, "panel_M_example_traces_ppk25_narrative.txt"))
cat(narrative)
```
