---
title: "Figure 5 Panel M: AN05B035 morphology with ppk23/ppk25-biased synapses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")

suppressMessages({
  library(nat)
  library(neuprintr)
  library(rgl)
  library(dplyr)
  library(tibble)
  library(ggplot2)
  library(Matrix)
})

options(rgl.useNULL = TRUE)

out_root   <- "/Users/fkampf/Documents/pheromone.paper/figures/fig5/panel_M"
out_notes  <- file.path(out_root, "notes")
dir.create(out_notes, recursive = TRUE, showWarnings = FALSE)
out_pdf    <- panel_plot_dir("panel_M_morphology", "pdf")
out_png    <- panel_plot_dir("panel_M_morphology", "png")
```

## AN05B035 bodyids

```{r ids}
an05_meta <- mba %>%
  dplyr::filter(type == "AN05B035")
print(an05_meta %>% select(bodyid, type, instance, soma_side, status))
an05_ids <- an05_meta$bodyid
stopifnot(length(an05_ids) > 0)
```

## Fetch skeleton + synapses from neuprint (male-cns)

```{r fetch-neuron}
sk_list <- tryCatch(
  neuprint_read_skeletons(an05_ids, heal = TRUE),
  error = function(e) { message("skeleton fetch failed: ", conditionMessage(e)); NULL }
)

syn_tbl <- tryCatch(
  neuprint_get_synapses(an05_ids),
  error = function(e) { message("synapse fetch failed: ", conditionMessage(e)); NULL }
)

have_data <- !is.null(sk_list) && !is.null(syn_tbl) && nrow(syn_tbl) > 0
cat("skeleton OK:", !is.null(sk_list), "  synapses OK:",
    !is.null(syn_tbl) && nrow(syn_tbl) > 0, "\n")
if (have_data) cat("n synapses:", nrow(syn_tbl),
                   "  n presynaptic (outputs):", sum(syn_tbl$prepost == 0),
                   "  n postsynaptic (inputs):", sum(syn_tbl$prepost == 1), "\n")
```

## Compute ppk23/ppk25 sex-bias score for every partner type

For each connectome `type`, derive how strongly it is driven by ppk23 vs ppk25
contact-chemosensory channels, using the graph's adjacency matrix
(`adj.matrix`), summing incoming synapse weights from the ppk23 and ppk25
channel-neuron source sets.

```{r bias-per-type}
# adj.matrix[rownames -> colnames]: row = source, col = target (need to verify)
# standardize: make rows = pre, cols = post. From setup.R it comes from `load_all_data`.
# Inspect by checking a known channel neuron presence as a rowname.
is_row_source <- any(unlist(channel_neurons) %in% rownames(adj.matrix))
is_col_source <- any(unlist(channel_neurons) %in% colnames(adj.matrix))
cat("channel neurons in rows:", sum(unlist(channel_neurons) %in% rownames(adj.matrix)),
    "  in cols:", sum(unlist(channel_neurons) %in% colnames(adj.matrix)), "\n")

# Convenience getter: for a row vector indexed by type, sum across src_types
type_drive_from <- function(src_types) {
  src_types <- src_types[src_types %in% rownames(adj.matrix)]
  if (length(src_types) == 0) return(setNames(rep(0, ncol(adj.matrix)), colnames(adj.matrix)))
  if (length(src_types) == 1) return(as.numeric(adj.matrix[src_types, ]))
  Matrix::colSums(adj.matrix[src_types, , drop = FALSE])
}

drive_ppk23 <- type_drive_from(channel_neurons$ppk23)
drive_ppk25 <- type_drive_from(channel_neurons$ppk25)
names(drive_ppk23) <- colnames(adj.matrix)
names(drive_ppk25) <- colnames(adj.matrix)

# Also compute what ppk23/ppk25 RECEIVE from each source type (for presyn partners
# that are themselves upstream of AN05B035 but may be driving pheromone info
# downstream). We use direct drive as the bias for a partner type.
bias_tbl <- tibble::tibble(
  type        = names(drive_ppk23),
  drive_ppk23 = as.numeric(drive_ppk23),
  drive_ppk25 = as.numeric(drive_ppk25)
) %>%
  mutate(
    total = drive_ppk23 + drive_ppk25,
    bias  = ifelse(total > 0, (drive_ppk23 - drive_ppk25) / total, NA_real_)
  )

cat("types with any ppk23/ppk25 drive:", sum(!is.na(bias_tbl$bias)), "\n")
```

The `bias` value in `[-1, +1]`: **+1 = fully ppk23 (male-contact)**, **-1 = fully
ppk25 (female-contact)**, 0 = balanced, NA = no contact drive.

## Map every synapse partner bodyid -> type -> bias

```{r partner-bias}
# Partner bodyid -> type lookup via mba
bodyid_to_type <- setNames(mba$type, as.character(mba$bodyid))

if (have_data) {
  syn_df <- syn_tbl %>%
    as_tibble() %>%
    mutate(
      partner_type = bodyid_to_type[as.character(partner)],
      io           = ifelse(prepost == 1, "input", "output")   # 1 = postsynaptic (input to AN05B035), 0 = presynaptic (output)
    ) %>%
    left_join(bias_tbl %>% select(type, bias, drive_ppk23, drive_ppk25),
              by = c("partner_type" = "type")) %>%
    mutate(
      bias_class = dplyr::case_when(
        is.na(bias)            ~ "neither",
        bias >=  0.2           ~ "ppk23 (male)",
        bias <= -0.2           ~ "ppk25 (female)",
        TRUE                   ~ "balanced"
      )
    )

  # Sanity: what fraction of partners have a type, and a non-NA bias?
  cat("partners with known type:",
      mean(!is.na(syn_df$partner_type)) * 100, "%\n")
  cat("partners with non-NA bias:",
      mean(!is.na(syn_df$bias)) * 100, "%\n")

  cat("\nBias class breakdown (inputs):\n")
  print(syn_df %>% filter(io == "input")  %>% count(bias_class))
  cat("\nBias class breakdown (outputs):\n")
  print(syn_df %>% filter(io == "output") %>% count(bias_class))
}
```

## Render 2D morphology with coloured synapse markers

We project the skeleton to frontal (XY) and dorsal (XZ) views using
`nat::plot` style, then overlay synapse scatterpoints coloured by bias.

```{r render-plots, fig.width=12, fig.height=10}
render_panel <- function(sk_list, syn_df, view = c("XY","XZ"), title = "") {
  view <- match.arg(view)
  # Skeleton vertices across all neurons
  all_xyz <- do.call(rbind, lapply(sk_list, function(n) nat::xyzmatrix(n)))
  seg_list <- list()
  for (n in sk_list) {
    # build line segments from parent-child in swc-like data frame
    d <- n$d
    # parent column is named 'Parent' in neuron$d
    par_col <- intersect(c("Parent","parent"), colnames(d))[1]
    pt_col_x <- intersect(c("X","x"), colnames(d))[1]
    pt_col_y <- intersect(c("Y","y"), colnames(d))[1]
    pt_col_z <- intersect(c("Z","z"), colnames(d))[1]
    id_col   <- intersect(c("PointNo","pointno"), colnames(d))[1]
    valid <- d[[par_col]] != -1 & d[[par_col]] %in% d[[id_col]]
    if (!any(valid)) next
    child  <- d[valid, , drop = FALSE]
    parent <- d[match(child[[par_col]], d[[id_col]]), , drop = FALSE]
    seg <- data.frame(
      x0 = child[[pt_col_x]],  y0 = child[[pt_col_y]],  z0 = child[[pt_col_z]],
      x1 = parent[[pt_col_x]], y1 = parent[[pt_col_y]], z1 = parent[[pt_col_z]]
    )
    seg_list[[length(seg_list) + 1]] <- seg
  }
  segs <- dplyr::bind_rows(seg_list)
  if (view == "XY") {
    seg_df <- data.frame(a0 = segs$x0, b0 = segs$y0, a1 = segs$x1, b1 = segs$y1)
    syn_xy <- data.frame(a = syn_df$x, b = syn_df$y)
  } else {
    seg_df <- data.frame(a0 = segs$x0, b0 = segs$z0, a1 = segs$x1, b1 = segs$z1)
    syn_xy <- data.frame(a = syn_df$x, b = syn_df$z)
  }
  syn_xy$io         <- syn_df$io
  syn_xy$bias_class <- syn_df$bias_class
  syn_xy$bias       <- syn_df$bias

  bias_colors <- c(
    "ppk23 (male)"   = "#D62728",
    "ppk25 (female)" = "#1F77B4",
    "balanced"       = "#9E9E9E",
    "neither"        = "#DADADA"
  )

  # Draw inputs on one facet, outputs on another
  ggplot() +
    geom_segment(data = seg_df,
                 aes(x = a0, y = b0, xend = a1, yend = b1),
                 color = "grey25", linewidth = 0.12, alpha = 0.85) +
    geom_point(data = syn_xy,
               aes(x = a, y = b, color = bias_class),
               size = 0.45, alpha = 0.75) +
    scale_color_manual(values = bias_colors, name = "Partner ppk bias",
                       na.value = "#DADADA") +
    facet_wrap(~io, nrow = 1) +
    coord_fixed() +
    # EM voxel coords: y increases downward -> flip
    scale_y_reverse() +
    labs(title = title,
         subtitle = paste0("view = ", view,
                           "   (red = ppk23/male-contact-driven partner,",
                           "   blue = ppk25/female,",
                           "   grey = balanced/neither)")) +
    theme_void(base_size = 11) +
    theme(
      strip.text        = element_text(face = "bold"),
      plot.title        = element_text(face = "bold"),
      plot.subtitle     = element_text(size = 9, color = "grey30"),
      legend.position   = "bottom",
      panel.spacing     = unit(1, "lines")
    ) +
    guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))
}

if (have_data) {
  # Subsample synapses for scatter readability (keep all if <= 6000)
  max_pts_per_io <- 3500
  syn_plot <- syn_df %>%
    group_by(io) %>%
    group_modify(~ if (nrow(.x) > max_pts_per_io) slice_sample(.x, n = max_pts_per_io) else .x) %>%
    ungroup()

  p_xy <- render_panel(sk_list, syn_plot, view = "XY",
                       title = sprintf("AN05B035 (bodyids: %s) -- frontal view",
                                       paste(an05_ids, collapse = ", ")))
  p_xz <- render_panel(sk_list, syn_plot, view = "XZ",
                       title = "AN05B035 -- dorsal view")

  combined <- patchwork::wrap_plots(p_xy, p_xz, ncol = 1) +
    patchwork::plot_annotation(
      caption = paste(
        "SOURCE: skeleton + synapses from neuprint male-cns:v1.0.",
        "COLOR: synapses coloured by partner cell-type ppk23/ppk25 contact-pheromone drive.",
        "BIAS: (drive_ppk23 - drive_ppk25) / (drive_ppk23 + drive_ppk25).",
        "INPUTS: partner is presynaptic to AN05B035.",
        "OUTPUTS: partner is postsynaptic to AN05B035.",
        sep = "\n"
      )
    )
  print(combined)
} else {
  # -- Fallback: placeholder + Neuroglancer URL --
  ng_url <- tryCatch(
    paste0("https://clio-ng.janelia.org/#!",
           utils::URLencode(jsonlite::toJSON(make_mcns_scene(an05_ids), auto_unbox = TRUE),
                            reserved = TRUE)),
    error = function(e) NA_character_)

  combined <- ggplot() + theme_void() +
    annotate("text", x = 0, y = 0.6,
             label = "AN05B035 morphology -- data unavailable", size = 6, fontface = "bold") +
    annotate("text", x = 0, y = 0.3,
             label = "Could not fetch skeleton / synapses from neuprint.\nSee panel_M/notes/MORPHOLOGY_DATA_NOTES.md.\nNeuroglancer URL generated for manual inspection.",
             size = 4) +
    xlim(-1, 1) + ylim(0, 1)
  print(combined)
  cat("Neuroglancer URL:\n", ng_url, "\n")
}
```

## Save PDF and PNG

```{r save}
pdf_path <- file.path(out_pdf, "panel_M_morphology.pdf")
png_path <- file.path(out_png, "panel_M_morphology.png")

ggsave(pdf_path, plot = combined, width = 12, height = 14, device = cairo_pdf)
ggsave(png_path, plot = combined, width = 12, height = 14, dpi = 300)
cat("wrote:\n  ", pdf_path, "\n  ", png_path, "\n")

if (have_data) {
  # write synapse table for reproducibility
  write.csv(syn_df %>% select(connector_id, bodyid, partner, partner_type,
                              io, bias_class, bias, x, y, z),
            file.path(out_root, "M_an05b035_synapse_table.csv"),
            row.names = FALSE)
}
```

## Neuroglancer link (always generated)

```{r ng-url}
ng_url <- tryCatch(
  mcns_shortlink(an05_ids),
  error = function(e) {
    long <- paste0("https://clio-ng.janelia.org/#!",
                   utils::URLencode(jsonlite::toJSON(make_mcns_scene(an05_ids),
                                                     auto_unbox = TRUE),
                                    reserved = TRUE))
    long
  }
)
cat("Neuroglancer link:\n", ng_url, "\n")
writeLines(ng_url, file.path(out_root, "M_an05b035_neuroglancer_url.txt"))
```

## Narrative

```{r narrative, echo=FALSE}
narrative <- sprintf(
"PANEL M: AN05B035 morphology with ppk23/ppk25-biased synapses
==============================================================

WHAT THIS PANEL SHOWS
---------------------
Two 2D projections of the AN05B035 skeleton (both hemispheres,
bodyids %s), with input and output synapses plotted as points and
colour-coded by the ppk23/ppk25 contact-pheromone drive of the partner
cell type:
  RED   : partner type is predominantly driven by ppk23 (male cuticular contact)
  BLUE  : partner type is predominantly driven by ppk25 (female cuticular contact)
  GREY  : partner type is balanced or has no detectable ppk23/ppk25 drive

METHOD
------
Skeleton fetched via `neuprintr::neuprint_read_skeletons` (male-cns:v1.0,
healed). Synapses fetched via `neuprint_get_synapses`; prepost=1 -> input
(AN05B035 is postsynaptic), prepost=0 -> output (AN05B035 is presynaptic).
Each partner bodyid is mapped to its cell type via mba, and each type is
scored by bias = (drive_ppk23 - drive_ppk25) / (drive_ppk23 + drive_ppk25)
where drives are total synaptic input from the ppk23 vs ppk25 channel-neuron
sets defined in channel_neurons (fig5 setup.R). |bias| >= 0.2 classifies a
partner as male- or female-biased.

INTERPRETATION
--------------
This panel lets the reader see *spatially* where AN05B035 reads male- vs
female-contact information on its dendrite, and where it projects those
two streams on its axon. Colour clustering of inputs by bias highlights
whether the ppk23/ppk25 segregation is preserved at the level of synapse
locations on AN05B035, not just as bulk connectivity statistics.
",
paste(an05_ids, collapse = ", "))

writeLines(narrative, file.path(out_root, "M_an05b035_morphology_narrative.txt"))
cat(narrative)
```
