---
title: "Panel M supplement: GABA sign-inverter contribution to inhibition on ppk23 vs ppk25 (absolute strength)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")

library(dplyr)
library(tidyr)
library(ggplot2)
```

## Status

**Supplementary view** to the primary figure `panel_M_inhibitory_sources`.
This panel plots the same attribution in absolute (un-normalised) path
strength so readers can see subtype-level magnitude in addition to share.
Palette and mAL sort order are inherited from the primary figure for
visual consistency.

## Compute per-subtype GABA-sign-inverter share of inhibition on each channel

```{r compute}
channels <- c("ppk23", "ppk25")

inh_paths <- mal_paths %>%
  filter(modality %in% channels,
         end %in% mal_subtypes,
         valence == "inhibitory") %>%
  rowwise() %>%
  mutate(via_gaba_si = path_via_gaba_si(path)) %>%
  ungroup()

via_gaba_si_df <- inh_paths %>%
  filter(via_gaba_si) %>%
  group_by(end, modality) %>%
  summarise(inh_via_gaba_si = sum(strength, na.rm = TRUE), .groups = "drop")

stopifnot(all(channels %in% colnames(mal_strength_inh)))
inh_sub <- mal_strength_inh[mal_subtypes, channels, drop = FALSE]
tab <- as.data.frame(inh_sub) %>%
  tibble::rownames_to_column("mal") %>%
  pivot_longer(cols = all_of(channels), names_to = "channel",
               values_to = "inh_total") %>%
  as_tibble() %>%
  left_join(via_gaba_si_df, by = c("mal" = "end", "channel" = "modality")) %>%
  mutate(
    inh_via_gaba_si = ifelse(is.na(inh_via_gaba_si), 0, inh_via_gaba_si),
    inh_other       = pmax(inh_total - inh_via_gaba_si, 0),
    share           = ifelse(inh_total > 0, inh_via_gaba_si / inh_total, 0)
  )

wide <- tab %>%
  select(mal, channel, inh_total, inh_via_gaba_si, share) %>%
  pivot_wider(names_from = channel,
              values_from = c(inh_total, inh_via_gaba_si, share),
              values_fill = 0)

mean_share_ppk23 <- mean(wide$share_ppk23, na.rm = TRUE)
mean_share_ppk25 <- mean(wide$share_ppk25, na.rm = TRUE)

cat(sprintf("\nMean GABA-sign-inverter-attributable share across %d mAL_m subtypes:\n",
            nrow(wide)))
cat(sprintf("  ppk23: %.1f%%\n", 100 * mean_share_ppk23))
cat(sprintf("  ppk25: %.1f%%\n", 100 * mean_share_ppk25))
```

## Shared palette and sort order (Pass 1 + 2)

```{r shared}
# Same canonical palette as the primary figure.
# Here we only have two categories (GABA sign-inverter pool vs rest),
# so we collapse "mAL lateral" + "other" into a single grey group.
comp_colors <- c(
  "Via GABA sign-inverter"  = "#1F4E79",  # identical to primary
  "Other inhibitory paths"  = "grey70"    # collapsed non-GABA-SI
)

# Read the sort order exported by panel_M_inhibitory_sources if available;
# otherwise fall back to local ordering by combined GABA-SI contribution.
sort_path <- file.path(fig5_dir, "plots_mspecific", "panel_M",
                       "panel_M_mal_sort_order.rds")
if (file.exists(sort_path)) {
  shared_order <- readRDS(sort_path)
  # Keep only mAL_m subtypes present in this panel
  shared_order <- shared_order[shared_order %in% wide$mal]
  # Append any mAL missing from the shared order
  shared_order <- c(shared_order, setdiff(wide$mal, shared_order))
  message("  Using shared sort order from panel_M_inhibitory_sources.")
} else {
  shared_order <- wide %>%
    mutate(combined = inh_via_gaba_si_ppk23 + inh_via_gaba_si_ppk25) %>%
    arrange(desc(combined)) %>% pull(mal)
  message("  Shared sort order not found; using local fallback.")
}
```

## Plot: 2-column stacked bars (ppk23 | ppk25) per mAL_m subtype

```{r plot, fig.width=10, fig.height=8}
plot_df <- tab %>%
  select(mal, channel, inh_via_gaba_si, inh_other) %>%
  pivot_longer(cols = c(inh_via_gaba_si, inh_other),
               names_to = "component",
               values_to = "strength") %>%
  mutate(
    # Top = highest GABA-SI share
    mal       = factor(mal, levels = rev(shared_order)),
    channel   = factor(channel, levels = c("ppk23", "ppk25")),
    component = factor(component,
                       levels = c("inh_other", "inh_via_gaba_si"),
                       labels = c("Other inhibitory paths",
                                  "Via GABA sign-inverter"))
  )

# Pass 3: in-segment percentage labels where share > 8%
plot_df <- plot_df %>%
  group_by(mal, channel) %>%
  mutate(
    tot   = sum(strength),
    frac  = ifelse(tot > 0, strength / tot, 0),
    lbl   = ifelse(frac > 0.08, sprintf("%.0f%%", 100 * frac), ""),
    lbl_c = ifelse(component == "Via GABA sign-inverter", "white", "grey15")
  ) %>% ungroup()

# Pass 8: informative facet labels
channel_labeller <- as_labeller(c(
  "ppk23" = "ppk23 channel (male contact pheromone)",
  "ppk25" = "ppk25 channel (female contact pheromone)"
))

p <- ggplot(plot_df, aes(x = strength, y = mal, fill = component)) +
  geom_col(width = 0.82, color = "white", linewidth = 0.2) +
  geom_text(aes(label = lbl, color = lbl_c),
            position = position_stack(vjust = 0.5),
            size = 2.9, show.legend = FALSE) +
  scale_color_identity() +
  facet_wrap(~ channel, ncol = 2, scales = "free_x",
             labeller = channel_labeller) +
  scale_fill_manual(values = comp_colors, name = NULL) +
  labs(
    x = "Inhibitory path strength (sum)",
    y = NULL
  ) +
  theme_pub(base_size = 12) +
  theme(
    legend.position    = "bottom",
    # Pass 7: larger mAL labels
    axis.text.y        = element_text(size = rel(1.15), face = "bold"),
    axis.text.x        = element_text(angle = 0, hjust = 0.5),
    panel.grid.major.y = element_blank(),
    strip.text         = element_text(size = rel(1.05), face = "bold")
  )

print(p)
```

## Save outputs

```{r save}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
png_dir <- panel_plot_dir("panel_M", "png")

out_pdf <- file.path(pdf_dir, "panel_M_an05_share_per_channel.pdf")
out_png <- file.path(png_dir, "panel_M_an05_share_per_channel.png")

ggsave(out_pdf, plot = p, width = 10, height = 8)
ggsave(out_png, plot = p, width = 10, height = 8, dpi = 300)

write.csv(wide, file.path(dirname(pdf_dir),
                          "panel_M_an05_share_per_channel_data.csv"),
          row.names = FALSE)

cat("Saved:\n  ", out_pdf, "\n  ", out_png, "\n")
```

## Narrative

```{r narrative, echo=FALSE}
narrative <- sprintf(
"PANEL M (supplementary): GABA-sign-inverter-attributable inhibition, absolute strength
=====================================================================================

Companion to the primary figure panel_M_inhibitory_sources. That figure shows
share (normalised to 100%%); this supplement shows absolute summed path
strength on each channel so subtype magnitude is also readable.

For each of the %d male-specific mAL_m subtypes we compute the total
inhibitory path strength arriving from ppk23 and ppk25 (rows of
mal_strength_inh), and the subset of that strength carried by paths whose
interior nodes contain ANY member of the GABA sign-inverter pool
(gaba_sign_inverter_types, 11 AN05B / IN05 GABA relay types).

MEAN GABA SI SHARE:
  ppk23: %.1f%%
  ppk25: %.1f%%

Palette and row order are inherited from the primary figure.
",
nrow(wide),
100 * mean_share_ppk23, 100 * mean_share_ppk25
)

narr_path <- file.path(fig5_dir, "plots_mspecific", "panel_M",
                      "panel_M_an05_share_per_channel_narrative.txt")
dir.create(dirname(narr_path), recursive = TRUE, showWarnings = FALSE)
writeLines(narrative, narr_path)
cat(narrative)
```
