---
title: "[DEPRECATED] Panel M: AN05B035 inhibition attribution on ppk23 AND ppk25"
output: html_document
---

> **DEPRECATED -- do not cite as a separate panel.**
>
> This figure is a strict 2-category subset (AN05B035 vs "not via AN05B035")
> of the primary figure `panel_M_inhibitory_sources.Rmd`, which decomposes
> the same inhibition into four categories (AN05B035 / other GABA relays /
> mAL lateral / other). The primary figure also contains a matching
> absolute-strength supplement (`M_an05_share_per_channel.Rmd`).
>
> Kept in the repository for historical reproducibility only. Rendering it
> still works and produces an output with the shared palette and sort order,
> but it should not be promoted to the manuscript.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
```

```{r attribute-inhibition}
inh_paths <- mal_paths %>%
  filter(
    modality %in% c("ppk23", "ppk25"),
    end %in% mal_subtypes,
    valence == "inhibitory"
  ) %>%
  mutate(
    via_an05b035 = grepl("(^|[^A-Za-z0-9])AN05B035([^A-Za-z0-9]|$)", path)
  )

attrib <- inh_paths %>%
  mutate(attribution = ifelse(via_an05b035, "via_AN05B035", "other")) %>%
  group_by(end, modality, attribution) %>%
  summarise(strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  complete(end = mal_subtypes,
           modality = c("ppk23", "ppk25"),
           attribution = c("via_AN05B035", "other"),
           fill = list(strength = 0))

totals <- attrib %>%
  group_by(end, modality) %>%
  summarise(total_inh = sum(strength), .groups = "drop")

pct_tbl <- attrib %>%
  filter(attribution == "via_AN05B035") %>%
  select(end, modality, via = strength) %>%
  left_join(totals, by = c("end", "modality")) %>%
  mutate(pct = ifelse(total_inh > 0, 100 * via / total_inh, 0))

pct_wide <- pct_tbl %>%
  select(end, modality, pct) %>%
  pivot_wider(names_from = modality, values_from = pct, names_prefix = "pct_") %>%
  rename(target_type = end)

# Pass 2: use the shared sort order from the primary figure when available
sort_path <- file.path(fig5_dir, "plots_mspecific", "panel_M",
                       "panel_M_mal_sort_order.rds")
if (file.exists(sort_path)) {
  subtype_order <- readRDS(sort_path)
  subtype_order <- subtype_order[subtype_order %in% mal_subtypes]
  subtype_order <- c(subtype_order, setdiff(mal_subtypes, subtype_order))
  message("  Using shared sort order from panel_M_inhibitory_sources.")
} else {
  subtype_order <- attrib %>%
    filter(attribution == "via_AN05B035") %>%
    group_by(end) %>%
    summarise(an_contribution = sum(strength), .groups = "drop") %>%
    arrange(desc(an_contribution)) %>% pull(end)
}

attrib <- attrib %>%
  mutate(
    target_type = factor(end, levels = subtype_order),
    attribution = factor(attribution,
                         levels = c("other", "via_AN05B035"),
                         labels = c("not via AN05B035",
                                    "via AN05B035")),
    modality = factor(modality, levels = c("ppk23", "ppk25"))
  )

pct_wide <- pct_wide %>%
  mutate(target_type = factor(target_type, levels = subtype_order)) %>%
  arrange(target_type)
```

```{r mean-attribution}
mean_pct_ppk23 <- mean(pct_wide$pct_ppk23, na.rm = TRUE)
mean_pct_ppk25 <- mean(pct_wide$pct_ppk25, na.rm = TRUE)
median_pct_ppk23 <- median(pct_wide$pct_ppk23, na.rm = TRUE)
median_pct_ppk25 <- median(pct_wide$pct_ppk25, na.rm = TRUE)
```

```{r stacked-bars, fig.width=12, fig.height=8}
# Pass 1: shared palette -- AN05B035 uses the canonical deep-blue, "other"
# uses the grey reserved for the pooled non-AN05B035 complement.
fill_vals <- c("not via AN05B035" = "grey70",
               "via AN05B035"     = "#1F4E79")

channel_labeller <- as_labeller(c(
  "ppk23" = "ppk23 channel (male contact pheromone)",
  "ppk25" = "ppk25 channel (female contact pheromone)"
))

p_bars <- ggplot(attrib,
                 aes(x = target_type, y = strength,
                     fill = attribution)) +
  geom_col(position = position_stack(),
           width = 0.78, color = "grey25", linewidth = 0.15) +
  facet_wrap(~ modality, ncol = 1, scales = "free_y",
             strip.position = "right",
             labeller = channel_labeller) +
  scale_fill_manual(values = fill_vals, name = "Inhibitory path") +
  labs(
    x = NULL,
    y = "Inhibitory path strength",
    title = NULL
  ) +
  theme_pub(base_size = 12) +
  theme(
    legend.position = "top",
    # Pass 7: larger mAL labels
    axis.text.x = element_text(angle = 35, hjust = 1, size = rel(1.15),
                               face = "bold"),
    strip.text = element_text(face = "bold")
  )
```

```{r pct-panel, fig.width=12, fig.height=3}
pct_long <- pct_wide %>%
  pivot_longer(cols = starts_with("pct_"),
               names_to = "modality", values_to = "pct") %>%
  mutate(
    modality = sub("^pct_", "", modality),
    modality = factor(modality, levels = c("ppk23", "ppk25"))
  )

p_pct <- ggplot(pct_long,
                aes(x = target_type, y = modality,
                    fill = pct)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.0f%%", pct)),
            size = 3, color = "grey10") +
  scale_fill_gradient(
    low = "#F7F7F7", high = "#1F4E79",
    # Pass 6: explicit 0-100 range
    limits = c(0, 100),
    name = "% inh via\nAN05B035"
  ) +
  labs(x = "mAL subtype (ordered by AN05B035 share, both channels)",
       y = NULL) +
  theme_pub(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1,
                               size = rel(1.15), face = "bold"),
    legend.position = "right",
    panel.grid = element_blank()
  )
```

```{r compose-save, fig.width=12, fig.height=11}
# Pass 9: removed the redundant main title / subtitle that repeated info
# now printed inside the stacked bars + tile row themselves. Footer only
# carries the deprecation notice.
deprecation_note <- paste0(
  "DEPRECATED: 2-category subset of panel_M_inhibitory_sources ",
  "(see that figure for the 4-source decomposition)."
)

p_final <- p_bars / p_pct +
  plot_layout(heights = c(3.2, 1)) +
  plot_annotation(
    caption = deprecation_note,
    theme = theme(
      plot.caption = element_text(size = 9, color = "grey40",
                                  face = "italic", hjust = 0)
    )
  )

print(p_final)

out_pdf_dir <- panel_plot_dir("panel_M", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
out_pdf <- file.path(out_pdf_dir, "panel_M_E_inhibition_attribution.pdf")
ggsave(out_pdf, plot = p_final, width = 12, height = 11)
message(sprintf("  Saved: %s", out_pdf))

out_png_dir <- panel_plot_dir("panel_M", "png")
ggsave(file.path(out_png_dir, "panel_M_E_inhibition_attribution.png"),
       plot = p_final, width = 12, height = 11, dpi = 300)
```

```{r narrative, echo=FALSE}
narrative <- sprintf(
"[DEPRECATED] PANEL M extension: AN05B035-attributed inhibition on ppk23 AND ppk25
=================================================================================

This figure is a 2-category subset of the primary panel_M_inhibitory_sources
figure (AN05B035 vs not-via-AN05B035 only). Retained for reproducibility.

KEY NUMBERS
-----------
  Mean %%  AN05B035-attributable inhibition (ppk23):  %.1f %%
  Mean %%  AN05B035-attributable inhibition (ppk25):  %.1f %%
  Median %% (ppk23):                                  %.1f %%
  Median %% (ppk25):                                  %.1f %%
  Inhibitory paths (ppk23+ppk25, mAL_m targets):      %d
  Of those via AN05B035:                              %d (%.1f %%)
",
mean_pct_ppk23, mean_pct_ppk25,
median_pct_ppk23, median_pct_ppk25,
nrow(inh_paths),
sum(inh_paths$via_an05b035),
100 * mean(inh_paths$via_an05b035)
)

narr_dir <- file.path(fig5_dir, "plots_mspecific", "panel_M")
dir.create(narr_dir, recursive = TRUE, showWarnings = FALSE)
writeLines(narrative,
           file.path(narr_dir, "panel_M_E_inhibition_attribution_narrative.txt"))
cat(narrative)
```
