---
title: "Panel M: mAL_m subtypes ordered by SI input (strength recalculated from SI)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/panel_M/compute_si_paths.R")
library(viridisLite)
```

## Goal

Bar plot of the 16 mAL_m subtypes (x-axis, ordered descending by total
SI-driven input) stacked by GABA sign-inverter (SI) identity. Segment
heights are "strength *recalculated from the SI*" — meaning the
strongest-path computation is re-run with the SI as the source and the
mAL_m as the target, so the ORN -> SI prefix is dropped by
construction.

## Load SI -> mAL_m cache

```{r load}
si_to_mal <- load_paths_si_to_mal()
stopifnot(nrow(si_to_mal) > 0)

df <- si_to_mal %>%
  filter(start %in% gaba_sign_inverter_types, end %in% mal_subtypes) %>%
  group_by(mal = end, si = start) %>%
  summarise(strength = sum(strength, na.rm = TRUE), .groups = "drop")

# Totals for ordering
totals <- df %>%
  group_by(mal) %>%
  summarise(total = sum(strength), .groups = "drop") %>%
  arrange(desc(total))

# Fill in zero bars for mAL_m subtypes with no SI input (if any)
missing_mal <- setdiff(mal_subtypes, totals$mal)
if (length(missing_mal) > 0) {
  totals <- bind_rows(totals,
                      tibble(mal = missing_mal, total = 0))
}

mal_order <- totals$mal
df <- df %>% mutate(mal = factor(mal, levels = mal_order))
print(totals)
```

## SI ordering + palette (AN05B035 keyed to the SI-pool dark blue)

```{r palette}
si_ranking <- df %>%
  group_by(si) %>%
  summarise(mean_contribution = mean(strength), .groups = "drop") %>%
  arrange(desc(mean_contribution))
# Ensure every pool member appears even if absent from df
si_ranking <- bind_rows(
  si_ranking,
  tibble(si = setdiff(gaba_sign_inverter_types, si_ranking$si),
         mean_contribution = 0)
)
si_levels <- si_ranking$si

pal <- viridisLite::turbo(length(si_levels), begin = 0.05, end = 0.95)
names(pal) <- si_levels
if ("AN05B035" %in% names(pal)) {
  pal[["AN05B035"]] <- unname(panel_M_palette[["gaba_si"]])
}

df <- df %>% mutate(si = factor(si, levels = si_levels))
```

## Plot

```{r plot, fig.width=11, fig.height=6.5}
p <- ggplot(df, aes(x = mal, y = strength, fill = si)) +
  geom_col(position = "stack", color = "white", linewidth = 0.15) +
  scale_fill_manual(values = pal, name = "GABA sign-inverter") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = NULL, subtitle = NULL,
    x = NULL,
    y = "SI -> mAL_m path strength  (recalculated from SI)"
  ) +
  theme_pub(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")

print(p)
```

## Save

```{r save}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
png_dir <- panel_plot_dir("panel_M", "png")
data_dir <- panel_plot_dir("panel_M", "data")

pdf_path <- file.path(pdf_dir, "panel_M_mal_by_si_input.pdf")
png_path <- file.path(png_dir, "panel_M_mal_by_si_input.png")
csv_path <- file.path(data_dir, "panel_M_mal_by_si_input_data.csv")

pdf_device <- if (capabilities("cairo")) cairo_pdf else grDevices::pdf
ggsave(pdf_path, plot = p, width = 11, height = 6.5, device = pdf_device)
ggsave(png_path, plot = p, width = 11, height = 6.5, dpi = 300)
write.csv(df, csv_path, row.names = FALSE)

cat(sprintf("Saved PDF: %s  (%.1f KB)\n", pdf_path,
            file.info(pdf_path)$size / 1024))
cat(sprintf("Saved PNG: %s  (%.1f KB)\n", png_path,
            file.info(png_path)$size / 1024))
cat(sprintf("Saved CSV: %s\n", csv_path))
```

## Narrative

```{r narrative, echo=FALSE}
top3 <- head(totals, 3)
narr <- paste0(
"PANEL M: mAL_m subtypes ordered by GABA sign-inverter input
(strength recalculated from SI)
=============================================================

DESIGN
------
x-axis = 16 male-specific mAL_m subtypes, ordered descending by total
SI-driven path strength. Each bar is stacked by SI pool member. Segment
heights come from a top-50 strongest-path computation with the SI as
path source and the mAL_m as path target — so the ORN -> SI prefix is
dropped by construction and the bar height reflects the SI's
*downstream reach* onto the mAL_m.

", panel_M_pool_caption, "

TOP 3 mAL_m BY SI DRIVE
-----------------------
", paste(sprintf("  %-10s  total SI strength = %.4g", top3$mal, top3$total),
         collapse = "\n"), "

INTERPRETATION
--------------
Bar heights rank mAL_m subtypes by how strongly the SI pool projects
onto them. The stacked fills show which SI carries the drive — a bar
dominated by a single color is driven primarily by one SI, whereas a
multicolored bar receives converging input from several pool members.
This decomposition is channel-agnostic by construction (the ppk23/ppk25
ORN prefix has been dropped) and complements the
panel_M_si_channel_scatter plot, which attributes channel bias per SI.

METHOD NOTES
------------
- Cache: strongest.50.paths.gaba_si_pool.2.mAL_m.feather.
- Strength: multiplicative synapse-weight product from SI to mAL_m.
- AN05B035 keyed to the Panel M SI-pool dark blue; other SIs on a
  turbo palette ordered by mean contribution across mAL_m subtypes.
")

writeLines(narr, file.path(data_dir, "panel_M_mal_by_si_input_narrative.txt"))
cat(narr)
```
