---
title: "Panel M: path-strength contribution of SI-traversing paths vs all other paths"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
```

## Goal

Per male-specific mAL_m subtype, split the total ppk23/ppk25 path
strength into (a) paths whose interior passes through any GABA SI pool
member and (b) all other paths. Stacked bars faceted by channel show
what fraction of each subtype's drive actually flows through the SI
pool.

## Aggregate

```{r aggregate}
df <- mal_paths %>%
  filter(modality %in% c("ppk23", "ppk25"), end %in% mal_subtypes) %>%
  mutate(via_si = vapply(path, path_via_gaba_si, logical(1))) %>%
  group_by(mal = end, modality, via_si) %>%
  summarise(strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  mutate(bucket = factor(ifelse(via_si, "via SI pool", "other paths"),
                         levels = c("via SI pool", "other paths")))

totals <- df %>%
  group_by(mal, modality) %>%
  summarise(total = sum(strength), .groups = "drop")

# Order mAL_m by total strength (summed over both channels), descending.
mal_order <- totals %>%
  group_by(mal) %>%
  summarise(total = sum(total)) %>%
  arrange(desc(total)) %>%
  pull(mal)

df <- df %>% mutate(mal = factor(mal, levels = mal_order))

# Overall SI fraction per channel
channel_summary <- df %>%
  group_by(modality, via_si) %>%
  summarise(strength = sum(strength), .groups = "drop") %>%
  group_by(modality) %>%
  mutate(frac = strength / sum(strength)) %>%
  ungroup() %>%
  filter(via_si) %>%
  select(modality, si_frac = frac, si_strength = strength)
print(channel_summary)
```

## Plot

```{r plot, fig.width=12, fig.height=7}
pal <- c("via SI pool" = unname(panel_M_palette[["gaba_si"]]),
         "other paths" = "grey75")

subtitle_txt <- paste(
  sprintf("SI fraction of total drive: ppk23 = %.1f%%, ppk25 = %.1f%%",
          100 * channel_summary$si_frac[channel_summary$modality == "ppk23"],
          100 * channel_summary$si_frac[channel_summary$modality == "ppk25"]),
  "(SOURCE: mal_paths; top-50 strongest paths per (channel, mAL_m).)",
  sep = "\n"
)

per_bar <- df %>%
  group_by(mal, modality) %>%
  summarise(total = sum(strength),
            si    = sum(strength[via_si]),
            .groups = "drop") %>%
  mutate(si_pct = ifelse(total > 0, si / total, 0))

p <- ggplot(df, aes(x = mal, y = strength, fill = bucket)) +
  geom_col(position = "stack", color = "white", linewidth = 0.15) +
  geom_text(
    data = per_bar,
    mapping = aes(x = mal, y = total,
                  label = sprintf("%.0f%%", 100 * si_pct)),
    vjust = -0.4, size = 2.8, color = "grey15",
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = pal, name = NULL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  facet_wrap(~ modality, ncol = 1,
             labeller = panel_M_channel_labeller,
             scales = "free_y") +
  labs(
    title = NULL,
    x = NULL,
    y = "Total path strength to mAL_m  (% = SI share of that bar)"
  ) +
  theme_pub(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top",
        strip.text = element_text(size = rel(1.0), face = "bold"))

annot_df <- tibble(
  modality = c("ppk23", "ppk25"),
  label    = sprintf("SI fraction = %.1f%%",
                     100 * channel_summary$si_frac[match(c("ppk23","ppk25"),
                                                         channel_summary$modality)])
)

p <- p + geom_text(
  data = annot_df,
  mapping = aes(x = Inf, y = Inf, label = label),
  hjust = 1.05, vjust = 1.4, size = 3.2,
  color = "grey20", inherit.aes = FALSE
)

print(p)
```

## Fractional variant

Same aggregation, but each bar normalized to 1 — read off the per-mAL_m
SI fraction directly.

```{r plot_frac, fig.width=12, fig.height=7}
# Re-order mAL_m for the fraction plot by SI fraction summed across channels
frac_order <- df %>%
  group_by(mal) %>%
  summarise(frac = sum(strength[via_si]) / sum(strength), .groups = "drop") %>%
  arrange(desc(frac)) %>%
  pull(mal)
df_frac <- df %>% mutate(mal = factor(mal, levels = frac_order))

annot_df_frac <- tibble(
  modality = c("ppk23", "ppk25"),
  label    = sprintf("SI fraction = %.1f%%",
                     100 * channel_summary$si_frac[match(c("ppk23","ppk25"),
                                                         channel_summary$modality)])
)

p_frac <- ggplot(df_frac, aes(x = mal, y = strength, fill = bucket)) +
  geom_col(position = "fill", color = "white", linewidth = 0.15) +
  scale_fill_manual(values = pal, name = NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.03))) +
  facet_wrap(~ modality, ncol = 1,
             labeller = panel_M_channel_labeller) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "grey40", linewidth = 0.3) +
  geom_text(
    data = annot_df_frac,
    mapping = aes(x = Inf, y = Inf, label = label),
    hjust = 1.05, vjust = 1.4, size = 3.2,
    color = "grey20", inherit.aes = FALSE
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "Fraction of total path strength to mAL_m"
  ) +
  theme_pub(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top",
        strip.text = element_text(size = rel(1.0), face = "bold"))

print(p_frac)
```

## Save

```{r save}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
png_dir <- panel_plot_dir("panel_M", "png")
data_dir <- panel_plot_dir("panel_M", "data")

pdf_device <- if (capabilities("cairo")) cairo_pdf else grDevices::pdf

pdf_path <- file.path(pdf_dir, "panel_M_si_vs_other_contribution.pdf")
png_path <- file.path(png_dir, "panel_M_si_vs_other_contribution.png")
csv_path <- file.path(data_dir, "panel_M_si_vs_other_contribution_data.csv")
ggsave(pdf_path, plot = p, width = 12, height = 7, device = pdf_device)
ggsave(png_path, plot = p, width = 12, height = 7, dpi = 300)
write.csv(df, csv_path, row.names = FALSE)

pdf_path_f <- file.path(pdf_dir, "panel_M_si_vs_other_fraction.pdf")
png_path_f <- file.path(png_dir, "panel_M_si_vs_other_fraction.png")
ggsave(pdf_path_f, plot = p_frac, width = 12, height = 7, device = pdf_device)
ggsave(png_path_f, plot = p_frac, width = 12, height = 7, dpi = 300)

cat(sprintf("Saved PDF: %s  (%.1f KB)\n", pdf_path,
            file.info(pdf_path)$size / 1024))
cat(sprintf("Saved PDF: %s  (%.1f KB)\n", pdf_path_f,
            file.info(pdf_path_f)$size / 1024))
```

## Narrative

```{r narrative, echo=FALSE}
per_mal <- df %>%
  group_by(mal, modality) %>%
  summarise(
    total = sum(strength),
    si    = sum(strength[via_si]),
    .groups = "drop"
  ) %>%
  mutate(si_frac = ifelse(total > 0, si / total, 0))

top_si <- per_mal %>% arrange(desc(si_frac)) %>% head(5)
bot_si <- per_mal %>% arrange(si_frac)        %>% head(5)

narr <- paste0(
"PANEL M: SI-traversing path contribution vs other paths
========================================================

DESIGN
------
For each (channel, mAL_m) pair, paths in mal_paths (top-50 strongest
from ORN -> mAL_m per pair) are partitioned by whether any interior
node is a GABA sign-inverter pool member. Stacked bars show the
partitioned total strength; the SI fraction is (SI-traversing
strength) / (total strength).

", panel_M_pool_caption, "

CHANNEL-LEVEL SI FRACTION
-------------------------
ppk23: ", sprintf("%.1f%% of total drive routes via the SI pool (raw strength %.3f)",
                  100 * channel_summary$si_frac[channel_summary$modality == "ppk23"],
                  channel_summary$si_strength[channel_summary$modality == "ppk23"]), "
ppk25: ", sprintf("%.1f%% of total drive routes via the SI pool (raw strength %.3f)",
                  100 * channel_summary$si_frac[channel_summary$modality == "ppk25"],
                  channel_summary$si_strength[channel_summary$modality == "ppk25"]), "

TOP 5 mAL_m SUBTYPES BY SI FRACTION
-----------------------------------
", paste(sprintf("  %-12s  %-6s  SI=%.3f / total=%.3f  (%.1f%%)",
                 top_si$mal, top_si$modality,
                 top_si$si, top_si$total, 100 * top_si$si_frac),
         collapse = "\n"), "

BOTTOM 5 mAL_m SUBTYPES BY SI FRACTION
--------------------------------------
", paste(sprintf("  %-12s  %-6s  SI=%.3f / total=%.3f  (%.1f%%)",
                 bot_si$mal, bot_si$modality,
                 bot_si$si, bot_si$total, 100 * bot_si$si_frac),
         collapse = "\n"), "

INTERPRETATION
--------------
A stacked bar heavily blue means the mAL_m subtype gets most of its
channel drive through the SI pool; a mostly-grey bar means SI pool
routing is a minor slice. This complements
panel_M_inhibitory_sources (which restricts to inhibitory paths and
decomposes SI vs mAL-lateral vs other) by instead partitioning
TOTAL drive (excitatory + inhibitory) into SI-traversing vs not.
")

writeLines(narr, file.path(data_dir, "panel_M_si_vs_other_contribution_narrative.txt"))
cat(narr)
```
