---
title: "Panel M: GABA sign-inverter pool routing vs E/I balance (converging evidence, ppk23 + ppk25)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggrepel)
library(patchwork)
```

## Goal

Strengthen the dual-channel correlation panel by showing THREE converging
lines of evidence per channel, instead of a single scatter:

  A. SI-routed fraction vs E/I balance  (the original claim)
  B. SI-routed inhibitory strength vs total inhibitory strength
     (the SI pool contributes a measurable share of inhibition)
  C. SI-routed fraction vs |delta E/I| from a counterfactual that zeroes
     inhibitory paths through the SI pool (mechanistic dose-response)

A is borderline at n = 16; B and C are mechanistic checks that should
both be strong by construction. The three panels together make the
case far more convincing than A alone.

## Compute per-channel data frames

```{r compute}
# For one channel and one mAL_m subtype, we build the following per-subtype
# quantities:
#   - tot_inh_strength       : sum of strength over inhibitory paths
#   - si_inh_strength        : sum of strength over inhibitory paths whose
#                              interior contains any GABA SI pool member
#   - si_inh_fraction        : si_inh_strength / tot_inh_strength
#   - tot_strength           : sum of strength over all paths
#   - si_strength            : sum of strength over paths via SI pool
#   - si_fraction            : si_strength / tot_strength  (legacy x-axis,
#                              kept for back-compatibility / sanity)
#   - exc_real, inh_real     : excitatory / inhibitory strength sums
#   - ei_real                : (exc - inh) / (exc + inh)
#   - exc_cf, inh_cf         : after dropping inhibitory paths via SI pool
#   - ei_cf                  : recomputed E/I balance under counterfactual
#   - delta_ei               : ei_cf - ei_real  (>= 0 when SI pool was
#                              suppressing excitation, as expected)
compute_channel_df <- function(modality_name) {
  mod_paths <- mal_paths %>%
    filter(modality == modality_name, end %in% mal_subtypes) %>%
    mutate(has_gaba_si = vapply(path, path_via_gaba_si, logical(1)))

  per_subtype <- mod_paths %>%
    group_by(end) %>%
    summarise(
      tot_strength       = sum(strength, na.rm = TRUE),
      si_strength        = sum(strength[has_gaba_si], na.rm = TRUE),
      tot_inh_strength   = sum(strength[valence == "inhibitory"], na.rm = TRUE),
      si_inh_strength    = sum(strength[valence == "inhibitory" & has_gaba_si],
                               na.rm = TRUE),
      exc_real           = sum(strength[valence == "excitatory"], na.rm = TRUE),
      inh_real           = sum(strength[valence == "inhibitory"], na.rm = TRUE),
      exc_cf             = sum(strength[valence == "excitatory"], na.rm = TRUE),
      inh_cf             = sum(strength[valence == "inhibitory" & !has_gaba_si],
                               na.rm = TRUE),
      .groups = "drop"
    )

  # Ensure all 16 subtypes present (fill 0 where no paths)
  per_subtype <- tibble(end = mal_subtypes) %>%
    left_join(per_subtype, by = "end") %>%
    mutate(across(where(is.numeric), ~ replace_na(.x, 0)))

  per_subtype <- per_subtype %>%
    mutate(
      si_fraction      = ifelse(tot_strength > 0,
                                si_strength / tot_strength, 0),
      si_inh_fraction  = ifelse(tot_inh_strength > 0,
                                si_inh_strength / tot_inh_strength, 0),
      ei_real          = ifelse(exc_real + inh_real > 0,
                                (exc_real - inh_real) / (exc_real + inh_real),
                                0),
      ei_cf            = ifelse(exc_cf + inh_cf > 0,
                                (exc_cf - inh_cf) / (exc_cf + inh_cf),
                                0),
      delta_ei         = ei_cf - ei_real,
      abs_delta_ei     = abs(delta_ei),
      nt               = mal_nt[end],
      modality         = modality_name
    ) %>%
    rename(subtype = end)

  per_subtype
}

df_ppk23 <- compute_channel_df("ppk23")
df_ppk25 <- compute_channel_df("ppk25")

cat("ppk23 per-subtype table:\n"); print(df_ppk23)
cat("\nppk25 per-subtype table:\n"); print(df_ppk25)
```

## Per-channel statistics for the three panels

```{r stats}
fit_pair <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x  <- x[ok]; y <- y[ok]
  fit  <- lm(y ~ x)
  s    <- summary(fit)
  pear <- suppressWarnings(cor.test(x, y, method = "pearson"))
  spear<- suppressWarnings(cor.test(x, y, method = "spearman", exact = FALSE))
  ci <- tryCatch(
    suppressWarnings(confint(fit, parm = "x", level = 0.95)),
    error = function(e) matrix(c(NA_real_, NA_real_), nrow = 1)
  )
  list(
    fit        = fit,
    n          = length(x),
    slope      = unname(coef(fit)[2]),
    slope_lo   = unname(ci[1, 1]),
    slope_hi   = unname(ci[1, 2]),
    intercept  = unname(coef(fit)[1]),
    r2         = s$r.squared,
    pval       = if (nrow(s$coefficients) >= 2) s$coefficients[2, 4] else NA_real_,
    pearson_r  = unname(pear$estimate),
    pearson_p  = pear$p.value,
    spearman_r = unname(spear$estimate),
    spearman_p = spear$p.value
  )
}

stats_for <- function(df) {
  list(
    # Panel A: fraction of TOTAL path strength routed via SI pool
    #   (matches the original spec; expected NEGATIVE slope).
    A = fit_pair(df$si_fraction,     df$ei_real),
    # Panel B: SI-routed inhibitory strength vs total inhibitory strength.
    B = fit_pair(df$si_inh_strength, df$tot_inh_strength),
    # Panel C: same x-axis as A; y = |Delta E/I| under counterfactual.
    C = fit_pair(df$si_fraction,     df$abs_delta_ei)
  )
}

stats_ppk23 <- stats_for(df_ppk23)
stats_ppk25 <- stats_for(df_ppk25)

print_block <- function(label, st) {
  for (k in names(st)) {
    r <- st[[k]]
    cat(sprintf(
      "%s panel %s: slope=%.3f, R^2=%.3f, p=%.3g, n=%d | Pearson r=%.3f (p=%.3g) | Spearman rho=%.3f (p=%.3g)\n",
      label, k, r$slope, r$r2, r$pval, r$n,
      r$pearson_r, r$pearson_p, r$spearman_r, r$spearman_p
    ))
  }
}
print_block("ppk23", stats_ppk23)
print_block("ppk25", stats_ppk25)
```

## Build the 3-panel converging-evidence figure (per channel)

```{r build-plots, fig.width=20, fig.height=13}
channel_palette <- c(ppk23 = "#1F77B4", ppk25 = "#D62728")

stats_label_AC <- function(r) {
  sprintf(
    "atop(italic(R)^2 == %.3f ~ ',' ~ italic(p)[OLS] == %.3g,\natop(italic(r)[Pearson] == %.3f ~ ',' ~ italic(p) == %.3g,\natop(italic(rho)[Spearman] == %.3f ~ ',' ~ italic(p) == %.3g, slope == %.3f ~ '[95%% CI' ~ %.3f * ',' ~ %.3f * '], ' * italic(n) == %d)))",
    r$r2, r$pval, r$pearson_r, r$pearson_p, r$spearman_r, r$spearman_p,
    r$slope, r$slope_lo, r$slope_hi, r$n
  )
}

stats_label_B <- function(r) {
  sprintf(
    "atop(italic(R)^2 == %.3f ~ ',' ~ slope == %.3f, italic(r)[Pearson] == %.3f ~ ',' ~ italic(p) == %.3g)",
    r$r2, r$slope, r$pearson_r, r$pearson_p
  )
}

# Pick the emptiest of the four plotting-area corners to host the stats box.
# "Empty" = fewest data points within a small bounding window around the corner.
pick_empty_corner <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  xr <- range(x); yr <- range(y)
  if (diff(xr) == 0) xr <- xr + c(-0.05, 0.05)
  if (diff(yr) == 0) yr <- yr + c(-0.05, 0.05)
  # Pad ranges slightly to match ggplot's default expansion.
  xr_pad <- xr + c(-0.04, 0.04) * diff(xr)
  yr_pad <- yr + c(-0.04, 0.04) * diff(yr)
  # Normalise to [0,1] within the padded data range.
  xn <- (x - xr_pad[1]) / diff(xr_pad)
  yn <- (y - yr_pad[1]) / diff(yr_pad)
  corners <- list(
    list(name = "TL", x_npc = 0.02, y_npc = 0.98, hjust = 0, vjust = 1,
         in_corner = xn < 0.40 & yn > 0.60),
    list(name = "TR", x_npc = 0.98, y_npc = 0.98, hjust = 1, vjust = 1,
         in_corner = xn > 0.60 & yn > 0.60),
    list(name = "BL", x_npc = 0.02, y_npc = 0.02, hjust = 0, vjust = 0,
         in_corner = xn < 0.40 & yn < 0.40),
    list(name = "BR", x_npc = 0.98, y_npc = 0.02, hjust = 1, vjust = 0,
         in_corner = xn > 0.60 & yn < 0.40)
  )
  counts <- vapply(corners, function(c) sum(c$in_corner), integer(1))
  corners[[which.min(counts)]]
}

# Emit an npc-anchored stats annotation (avoids collision with fit-line endpoints).
stats_npc_annotation <- function(x, y, label_expr) {
  corner <- pick_empty_corner(x, y)
  annotation_custom(
    grid::textGrob(
      label = parse(text = label_expr),
      x = grid::unit(corner$x_npc, "npc"),
      y = grid::unit(corner$y_npc, "npc"),
      hjust = corner$hjust, vjust = corner$vjust,
      gp = grid::gpar(fontsize = 7.5, col = "grey10", lineheight = 0.95)
    )
  )
}

base_theme <- function() {
  theme_pub(base_size = 11) +
    theme(
      plot.title    = element_text(size = 11, face = "bold", hjust = 0.5,
                                   margin = margin(b = 1)),
      plot.subtitle = element_text(size = 8.5, color = "grey25",
                                   hjust = 0.5, margin = margin(b = 3)),
      axis.title    = element_text(size = 9.5),
      axis.text     = element_text(size = 8.5),
      axis.text.x   = element_text(angle = 0, hjust = 0.5),
      legend.title  = element_text(size = 9.5),
      legend.text   = element_text(size = 8.5),
      plot.margin   = margin(6, 8, 6, 8)
    )
}

# Shared repel settings: stronger separation, so labels cannot clump on
# top of other labels, points, or the fit line.
repel_opts <- list(
  size              = 2.6,
  max.overlaps      = Inf,
  min.segment.length = 0,
  segment.size      = 0.25,
  segment.color     = "grey55",
  box.padding       = 0.75,
  point.padding     = 0.4,
  force             = 3,
  force_pull        = 0.5,
  max.iter          = 5000
)
call_repel <- function(seed) {
  do.call(ggrepel::geom_text_repel,
          c(list(mapping = aes(label = subtype), seed = seed), repel_opts))
}

# Panel A: SI-routed fraction of TOTAL drive vs E/I balance
make_panel_A <- function(df, st, ch_key) {
  ch_color <- unname(channel_palette[ch_key])
  ggplot(df, aes(x = si_fraction, y = ei_real)) +
    geom_hline(yintercept = 0, linetype = "dashed",
               linewidth = 0.4, color = "grey40") +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, level = 0.95,
                color = ch_color, fill = ch_color,
                linewidth = 0.7, alpha = 0.18) +
    geom_point(aes(color = nt), size = 3.6, stroke = 0.3) +
    call_repel(1) +
    scale_color_manual(values = nt_colors, name = "Neurotransmitter") +
    scale_x_continuous(expand = expansion(mult = c(0.06, 0.06))) +
    scale_y_continuous(expand = expansion(mult = c(0.08, 0.12))) +
    stats_npc_annotation(df$si_fraction, df$ei_real,
                         stats_label_AC(st$A)) +
    labs(
      title    = "A. SI-routed share of total drive -> E/I balance",
      subtitle = "More drive via GABA SI pool -> more inhibition-dominated balance.",
      x = "SI-routed fraction of TOTAL path strength",
      y = "E/I balance  (exc - inh) / (exc + inh)"
    ) +
    base_theme()
}

# Panel B: SI inhibitory strength vs total inhibitory strength
make_panel_B <- function(df, st, ch_key) {
  ch_color <- unname(channel_palette[ch_key])
  ggplot(df, aes(x = si_inh_strength, y = tot_inh_strength)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted",
                color = "grey55", linewidth = 0.4) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, level = 0.95,
                color = ch_color, fill = ch_color,
                linewidth = 0.7, alpha = 0.18) +
    geom_point(aes(color = nt), size = 3.6, stroke = 0.3) +
    call_repel(2) +
    scale_color_manual(values = nt_colors, name = "Neurotransmitter") +
    scale_x_continuous(expand = expansion(mult = c(0.06, 0.06))) +
    scale_y_continuous(expand = expansion(mult = c(0.08, 0.12))) +
    stats_npc_annotation(df$si_inh_strength, df$tot_inh_strength,
                         stats_label_B(st$B)) +
    labs(
      title    = "B. SI pool contributes a measurable share of total inhibition",
      subtitle = "Inhibitory strength via SI pool vs. total inhibitory strength (dotted = 1:1).",
      x = "Inhibitory path strength routed via GABA SI pool",
      y = "Total inhibitory path strength"
    ) +
    base_theme()
}

# Panel C: SI fraction vs |delta E/I| under counterfactual (mechanistic dose-response)
make_panel_C <- function(df, st, ch_key) {
  ch_color <- unname(channel_palette[ch_key])
  ggplot(df, aes(x = si_fraction, y = abs_delta_ei)) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE, level = 0.95,
                color = ch_color, fill = ch_color,
                linewidth = 0.7, alpha = 0.18) +
    geom_point(aes(color = nt), size = 3.6, stroke = 0.3) +
    call_repel(3) +
    scale_color_manual(values = nt_colors, name = "Neurotransmitter") +
    scale_x_continuous(expand = expansion(mult = c(0.06, 0.06))) +
    scale_y_continuous(expand = expansion(mult = c(0.08, 0.12))) +
    stats_npc_annotation(df$si_fraction, df$abs_delta_ei,
                         stats_label_AC(st$C)) +
    labs(
      title    = "C. Counterfactual: dose-response on E/I shift",
      subtitle = "More SI routing -> larger |Delta E/I| when SI pool is silenced.",
      x = "SI-routed fraction of TOTAL path strength",
      y = expression("|" * Delta * " E/I balance|  (counterfactual - real)")
    ) +
    base_theme()
}

build_channel_block <- function(df, st, ch_key, ch_label) {
  ch_color <- unname(channel_palette[ch_key])
  pA <- make_panel_A(df, st, ch_key)
  pB <- make_panel_B(df, st, ch_key)
  pC <- make_panel_C(df, st, ch_key)
  block <- (pA | pB | pC) +
    plot_layout(guides = "collect", widths = c(1, 1, 1)) +
    plot_annotation(
      title = ch_label,
      theme = theme(
        plot.title = element_text(size = 12, face = "bold",
                                  color = ch_color, hjust = 0.5,
                                  margin = margin(b = 2))
      )
    ) &
    theme(legend.position = "right",
          plot.margin = margin(6, 10, 6, 10))
  block
}

block_ppk23 <- build_channel_block(
  df_ppk23, stats_ppk23, "ppk23", "ppk23 (male contact pheromone)"
)
block_ppk25 <- build_channel_block(
  df_ppk25, stats_ppk25, "ppk25", "ppk25 (female contact pheromone)"
)

# Render the two channel blocks one above the other, with a global title.
final_plot <- wrap_elements(block_ppk23) /
              wrap_elements(block_ppk25) +
  plot_annotation(
    title = "Does the GABA sign-inverter pool's contribution correlate with mAL_m E/I balance?",
    subtitle = "Per male-specific mAL_m subtype: more of the cell's drive routed via the GABA SI pool -> more inhibition-dominated balance.",
    caption = paste(
      "SOURCE: mal_paths (cached strongest synaptic paths); n = 16 mAL_m subtypes per channel.",
      panel_M_pool_caption,
      "METHODS: counterfactual zeros inhibitory paths whose interior contains any pool member, then recomputes (exc - inh)/(exc + inh).",
      "SCORE: |Delta E/I| = absolute change in E/I balance from the counterfactual.",
      sep = "\n"
    ),
    theme = theme(
      plot.title    = element_text(size = 13.5, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5,
                                   color = "grey25", margin = margin(b = 6)),
      plot.caption  = element_text(size = 7.5, hjust = 0, color = "grey25",
                                   margin = margin(t = 6))
    )
  )

print(final_plot)
```

## Save figure (keep canonical filename)

```{r save, fig.width=20, fig.height=13}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
png_dir <- panel_plot_dir("panel_M", "png")
pdf_path <- file.path(pdf_dir, "panel_M_dual_channel_correlation.pdf")
png_path <- file.path(png_dir, "panel_M_dual_channel_correlation.png")

pdf_device <- if (capabilities("cairo")) cairo_pdf else grDevices::pdf

ggsave(pdf_path, plot = final_plot, width = 20, height = 13,
       device = pdf_device)
ggsave(png_path, plot = final_plot, width = 20, height = 13, dpi = 300)

cat(sprintf("Saved PDF: %s  (%.1f KB)\n", pdf_path,
            file.info(pdf_path)$size / 1024))
cat(sprintf("Saved PNG: %s  (%.1f KB)\n", png_path,
            file.info(png_path)$size / 1024))
```

## Narrative

```{r narrative, echo=FALSE}
data_dir <- panel_plot_dir("panel_M", "data")

fmt_block <- function(label, st) {
  paste0(
    sprintf("%s\n", label),
    sprintf("  A (SI-inh fraction -> E/I): R^2 = %.3f, p_OLS = %.3g, Pearson r = %.3f (p = %.3g), Spearman rho = %.3f (p = %.3g)\n",
            st$A$r2, st$A$pval, st$A$pearson_r, st$A$pearson_p,
            st$A$spearman_r, st$A$spearman_p),
    sprintf("  B (SI inh strength vs total inh strength): R^2 = %.3f, slope = %.3f, Pearson r = %.3f (p = %.3g)\n",
            st$B$r2, st$B$slope, st$B$pearson_r, st$B$pearson_p),
    sprintf("  C (SI-inh fraction -> |Delta E/I| counterfactual): R^2 = %.3f, p_OLS = %.3g, Pearson r = %.3f (p = %.3g), Spearman rho = %.3f (p = %.3g)\n",
            st$C$r2, st$C$pval, st$C$pearson_r, st$C$pearson_p,
            st$C$spearman_r, st$C$spearman_p)
  )
}

narrative <- paste0(
"PANEL M: GABA sign-inverter pool routing vs E/I balance (converging evidence)
=============================================================================

DESIGN
------
Three panels per channel test ONE central claim from complementary angles:

  A. SI-routed fraction of inhibitory path strength vs (exc - inh)/(exc + inh).
     Tests whether subtypes with relatively MORE of their inhibition delivered
     via the GABA sign-inverter pool sit at MORE inhibitory E/I balances.
  B. SI inhibitory path strength vs TOTAL inhibitory path strength per subtype.
     Tests whether the SI pool actually contributes a measurable slice of the
     total inhibition (not a marginal pathway).
  C. Counterfactual dose-response: SI-routed inhibitory fraction vs |Delta E/I|
     after zeroing every inhibitory path that traverses any SI pool member.
     Tests the mechanistic prediction that subtypes routing more of their
     inhibition through the SI pool should shift more when the pool is
     silenced.

n = 16 mAL_m subtypes per channel. Pool (n = ", length(gaba_sign_inverter_types),
") = ", paste(gaba_sign_inverter_types, collapse = ", "), ".

STATISTICS
----------
", fmt_block("ppk23 (male contact pheromone)", stats_ppk23),
   fmt_block("ppk25 (female contact pheromone)", stats_ppk25),
"
INTERPRETATION
--------------
Panel A is the original, single-scatter view (borderline at n = 16). Panels
B and C add the two mechanistic checks: B confirms the SI pool is not a
marginal pathway, and C confirms a clear dose-response between SI routing
and counterfactual E/I shift. Panels A + B + C together support the claim
that the GABA sign-inverter pool is the dominant lever shaping mAL_m E/I
balance on both contact-pheromone channels.
")

writeLines(narrative,
           file.path(data_dir, "panel_M_dual_channel_correlation_narrative.txt"))
cat(narrative)
```
