---
title: "Panel M (dual): ppk23 & ppk25 -> GABA sign-inverter layer -> mAL_m sankey"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
if (!requireNamespace("ggalluvial", quietly = TRUE)) {
  install.packages("ggalluvial", repos = "https://cloud.r-project.org")
}
library(ggalluvial)
```

## Overview

Dual-channel Sankey / alluvial showing both **ppk23** and **ppk25** sensory
channels converging on the **GABA sign-inverter layer** -- a pool of
GABAergic AN05B/IN05* cells -- and fanning out (with inverted sign) to
male-specific `mAL_m*` subtypes. Each individual GABA sign-inverter (SI)
type appears as its own middle-layer node (labelled with the cell-type
name); lower-throughput SIs may be collapsed into an "other GABA SIs"
bin for visual legibility.

Three axes:

1. **Source channel**: `ppk23` (red) and `ppk25` (blue).
2. **GABA SI relay**: each individual GABA sign-inverter type, plus an
   optional `other GABA SIs` bin (low-throughput) and `other` for paths
   that do not pass through the GABA SI pool.
3. **mAL_m target**: the male-specific subtypes, ordered so that
   GABA-SI-dominated targets cluster at the top.

**Colour scheme**: source-channel colour (ppk23 = red, ppk25 = blue); any
ribbon passing through the GABA SI layer is overridden to dark purple /
near-black to mark the sign inversion.

## Extract ppk23 + ppk25 paths and assign GABA-SI relay identity

```{r extract-paths}
source_levels <- c("ppk23", "ppk25")

# Pool of GABA sign-inverter types (defined in setup_mspecific.R):
gaba_si_pool <- gaba_sign_inverter_types
cat("GABA sign-inverter pool (", length(gaba_si_pool), " types):\n  ",
    paste(gaba_si_pool, collapse = ", "), "\n", sep = "")

ppk_paths <- mal_paths %>%
  filter(modality %in% source_levels, end %in% mal_subtypes) %>%
  rowwise() %>%
  mutate(via_gaba_si = path_via_gaba_si(path)) %>%
  ungroup()

# For each path, determine which GABA SI it passes through. If multiple,
# pick the one closest to the mAL target (most downstream interior node
# in the pool) -- this is the SI that delivers the inhibitory drive.
pick_gaba_si <- function(path_str) {
  toks <- strsplit(path_str, " -> ", fixed = TRUE)[[1]]
  if (length(toks) <= 2) return(NA_character_)
  interior <- toks[-c(1, length(toks))]
  hits_idx <- which(interior %in% gaba_si_pool)
  if (length(hits_idx) == 0) return(NA_character_)
  interior[hits_idx[length(hits_idx)]]
}

ppk_paths <- ppk_paths %>%
  rowwise() %>%
  mutate(gaba_si = pick_gaba_si(path)) %>%
  ungroup() %>%
  mutate(relay_raw = ifelse(is.na(gaba_si), "other", gaba_si))

# Decide whether to collapse low-throughput SIs into "other GABA SIs".
# Heuristic: keep the top 5 SIs by aggregate path strength; collapse the
# rest if there are more than 6 distinct SIs with non-zero throughput.
si_throughput <- ppk_paths %>%
  filter(!is.na(gaba_si)) %>%
  group_by(gaba_si) %>%
  summarise(total = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total))

N_KEEP <- 5L
collapse_rule <- nrow(si_throughput) > 6
if (collapse_rule) {
  keep_si <- si_throughput$gaba_si[seq_len(min(N_KEEP, nrow(si_throughput)))]
} else {
  keep_si <- si_throughput$gaba_si
}
cat("\nSI throughput (aggregate path strength across ppk23+ppk25):\n")
print(si_throughput)
cat("\nKeeping as individual nodes:", paste(keep_si, collapse = ", "), "\n")
if (collapse_rule) {
  cat("Collapsing into 'other GABA SIs':",
      paste(setdiff(si_throughput$gaba_si, keep_si), collapse = ", "), "\n")
}

ppk_paths <- ppk_paths %>%
  mutate(
    relay = case_when(
      is.na(gaba_si)         ~ "no GABA SI\n(direct route)",
      gaba_si %in% keep_si   ~ gaba_si,
      TRUE                   ~ "other GABA SIs"
    )
  )

# Level ordering: keep SIs in descending-throughput order, then any
# collapse bin, then the "no GABA SI" stratum (paths that bypass the pool).
relay_levels <- c(keep_si,
                  if (collapse_rule) "other GABA SIs" else character(0),
                  "no GABA SI\n(direct route)")
relay_levels <- relay_levels[relay_levels %in% unique(ppk_paths$relay)]

ppk_paths <- ppk_paths %>%
  mutate(
    relay  = factor(relay,    levels = relay_levels),
    source = factor(modality, levels = source_levels)
  )

cat("\nppk23/ppk25 -> GABA-SI relay -> mAL_m path counts:\n")
print(table(ppk_paths$source, ppk_paths$relay, useNA = "ifany"))
```

## Aggregate path strength into (source, relay, target) flows

```{r aggregate}
gaba_si_relay_labels <- c(keep_si,
                          if (collapse_rule) "other GABA SIs" else character(0))

flows <- ppk_paths %>%
  group_by(source, relay, end) %>%
  summarise(strength = sum(strength, na.rm = TRUE), .groups = "drop") %>%
  filter(strength > 0) %>%
  mutate(
    via_gaba_si = as.character(relay) %in% gaba_si_relay_labels,
    # Single ribbon class used by the final plot:
    #   source colour for non-GABA-SI flow, dark-purple override for
    #   anything passing through the GABA sign-inverter pool
    ribbon_class = factor(
      if_else(via_gaba_si, "GABA SI (inhibitory)", as.character(source)),
      levels = c("ppk23", "ppk25", "GABA SI (inhibitory)")
    )
  )

# Order mAL_m targets so those with the largest GABA-SI inflow fraction
# cluster at the top of the target axis.
shared_order <- load_mal_sort_order()
if (!is.null(shared_order)) {
  target_order <- c(intersect(shared_order, as.character(unique(flows$end))),
                    setdiff(as.character(unique(flows$end)), shared_order))
} else {
  # Fallback: original per-figure computation
  target_order <- flows %>%
    group_by(end) %>%
    summarise(
      total    = sum(strength),
      inh      = sum(strength[via_gaba_si]),
      inh_frac = if_else(total > 0, inh / total, 0),
      .groups  = "drop"
    ) %>%
    arrange(desc(inh_frac), desc(total)) %>%
    pull(end)
}

missing_targets <- setdiff(mal_subtypes, target_order)
if (length(missing_targets) > 0) target_order <- c(target_order, missing_targets)

flows <- flows %>%
  mutate(end = factor(end, levels = rev(target_order))) %>%
  mutate(alluv_id = row_number())

cat("Flow rows:", nrow(flows), "\n")
cat("Total ppk23+ppk25 strength routed:", sprintf("%.2f", sum(flows$strength)), "\n")
for (s in source_levels) {
  ss <- flows$strength[flows$source == s]
  si <- flows$strength[flows$source == s & flows$via_gaba_si]
  cat(sprintf("  %s total=%.2f  via GABA SI layer=%.2f (%.1f%%)\n",
              s, sum(ss), sum(si),
              if (sum(ss) > 0) 100 * sum(si) / sum(ss) else 0))
}
```

## Wide-form frame for ggalluvial

```{r lodes}
flows_wide <- flows %>%
  transmute(
    alluv_id, strength,
    source, relay,
    target = end,
    ribbon_class
  )
```

## Single colour scheme: source channel + GABA-SI layer override

```{r sankey-final, fig.width=15, fig.height=9}
ribbon_colors <- c(
  "ppk23"                = unname(panel_M_palette["ppk23"]),    # #C0272D
  "ppk25"                = unname(panel_M_palette["ppk25"]),    # #F4A261
  "GABA SI (inhibitory)" = unname(panel_M_palette["gaba_si"])   # #1F4E79
)

p_sankey <- ggplot(flows_wide,
                   aes(y = strength,
                       axis1 = source, axis2 = relay, axis3 = target)) +
  geom_alluvium(aes(fill = ribbon_class),
                width = 1/6, alpha = 0.80,
                knot.pos = 0.3, curve_type = "sigmoid") +
  geom_stratum(width = 1/6, fill = "grey95", color = "grey40",
               linewidth = 0.35) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 3.1, lineheight = 0.9) +
  scale_x_discrete(limits = c("Source channel",
                              "GABA sign-inverter layer",
                              "mAL_m target"),
                   expand = c(0.05, 0.05)) +
  scale_fill_manual(values = ribbon_colors, name = "Ribbon class",
                    drop = FALSE) +
  labs(
    x = NULL, y = "Aggregated path strength",
    title = "ppk23/ppk25 -> mAL_m flow: which paths go through the GABA sign-inverter pool, which don't",
    subtitle = paste0(
      "Includes ALL ppk23/ppk25 -> mAL_m paths from `mal_paths` (both inhibitory routes via the GABA SI pool and ",
      "direct excitatory routes that bypass it). Only paths through the pool are sign-inverted."
    ),
    caption = paste(
      panel_M_pool_caption,
      "RIBBONS: thickness ~ aggregated path strength; red = ppk23 direct (excitatory), orange = ppk25 direct.",
      "INVERT: dark blue = path THROUGH the pool (inhibitory, sign-inverted).",
      "STRATUM: each pool type that carries flow is its own middle node; 'other GABA SIs' aggregates sub-top-5 pool members.",
      "BYPASS: 'no GABA SI (direct route)' = paths that do NOT pass through any pool member.",
      "ORDER: mAL_m targets ordered top-down by GABA-SI inflow fraction.",
      sep = "\n"
    )
  ) +
  theme_pub(base_size = 13) +
  theme(
    legend.position = "top",
    panel.grid      = element_blank(),
    axis.text.x     = element_text(size = rel(1.15), face = "bold",
                                   angle = 0, hjust = 0.5),
    axis.title.y    = element_text(size = rel(1.05)),
    plot.caption    = element_text(size = rel(0.85), hjust = 0.5,
                                   color = "grey25",
                                   margin = margin(t = 10)),
    plot.margin     = margin(18, 22, 14, 22)
  )

print(p_sankey)
```

## Save outputs (landscape 14 x 8)

```{r save}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
png_dir <- panel_plot_dir("panel_M", "png")

ggsave(file.path(pdf_dir, "panel_M_dual_sankey.pdf"),
       plot = p_sankey, width = 15, height = 9)
ggsave(file.path(png_dir, "panel_M_dual_sankey.png"),
       plot = p_sankey, width = 15, height = 9, dpi = 300)

# Remove any stale schemeB output files from a previous run
stale_files <- c(
  file.path(pdf_dir, "panel_M_dual_sankey_schemeB.pdf"),
  file.path(png_dir, "panel_M_dual_sankey_schemeB.png")
)
for (f in stale_files) {
  if (file.exists(f)) {
    file.remove(f)
    message("  Removed stale file: ", f)
  }
}

# Flow table
flows_out <- flows_wide %>%
  mutate(across(c(source, relay, target, ribbon_class), as.character)) %>%
  arrange(source, relay, target)
write.csv(flows_out,
          file.path(pdf_dir, "..", "panel_M_dual_sankey_flows.csv"),
          row.names = FALSE)

cat("Saved:\n",
    " ", file.path(pdf_dir, "panel_M_dual_sankey.pdf"), "\n",
    sep = "")
```

## Narrative

```{r narrative, echo=FALSE}
total_s <- sum(flows$strength)

per_source <- flows %>%
  group_by(source) %>%
  summarise(
    total = sum(strength),
    inh   = sum(strength[via_gaba_si]),
    inh_frac = if_else(total > 0, inh / total, 0),
    .groups = "drop"
  )

gaba_targets_by_src <- flows %>%
  filter(via_gaba_si) %>%
  group_by(source, end) %>%
  summarise(s = sum(strength), .groups = "drop") %>%
  arrange(source, desc(s))

per_si <- flows %>%
  filter(via_gaba_si) %>%
  group_by(relay) %>%
  summarise(total = sum(strength), .groups = "drop") %>%
  arrange(desc(total))

narrative <- sprintf(
"PANEL M (dual): ppk23 + ppk25 -> GABA SI layer -> mAL_m Sankey
==============================================================

WHAT THIS PANEL SHOWS
---------------------
Three-axis alluvial diagram of aggregated path strength from BOTH ppk23 and
ppk25 sensory channels, through the GABAergic sign-inverter layer (a pool
of %d AN05B/IN05* GABA cells), to the male-specific mAL_m subtypes. Each
individual GABA sign-inverter type is shown as its own middle-layer node;
low-throughput SIs may be aggregated into an 'other GABA SIs' bin.

Colour scheme:
- ppk23 excitatory routes = red
- ppk25 excitatory routes = blue
- Any route through the GABA SI layer = dark purple / near-black

mAL_m targets are ordered so those with the largest GABA-SI inflow
fraction cluster at the top of the target axis.

KEY NUMBERS
-----------
Total aggregated path strength across both channels: %.2f.
Per-source breakdown (total / via GABA SI layer / inhibitory fraction):
%s

Individual GABA SI throughput (descending):
%s

GABA-SI-routed mAL_m subtypes per source (by descending strength):
%s

INTERPRETATION
--------------
Both pheromone channels (ppk23, ppk25) share the GABA sign-inverter layer
as a common inhibitory relay. The Sankey makes two points visual:
(1) both sources converge onto the GABA SI pool -- the dark-purple ribbons
enter the middle stratum from both ppk23 and ppk25 and spread across
multiple GABAergic cell types rather than concentrating on a single node;
(2) from the GABA SI layer the flow diverges onto a SPECIFIC SUBSET of
mAL_m subtypes rather than spreading uniformly. This selective fan-out is
the circuit substrate that converts a shared excitatory sensory drive into
a subtype-specific inhibitory signature across a distributed GABAergic
pool, producing the differential ppk23 vs ppk25 responses and the E/I
balance shifts seen in earlier panels.
",
length(gaba_si_pool),
total_s,
paste(sprintf("  %s: total=%.2f  via GABA SI=%.2f (%.1f%%)",
              per_source$source, per_source$total, per_source$inh,
              100 * per_source$inh_frac),
      collapse = "\n"),
paste(sprintf("  %s: %.2f", per_si$relay, per_si$total),
      collapse = "\n"),
paste(sprintf("  %s -> %s (%.2f)",
              gaba_targets_by_src$source,
              gaba_targets_by_src$end,
              gaba_targets_by_src$s),
      collapse = "\n")
)

save_panel_narrative("panel_M", narrative)
cat(narrative)
```
