---
title: "Panel M: Census of interior nodes on ppk23/ppk25 inhibitory paths to mAL_m"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggplot2); library(dplyr); library(tidyr); library(patchwork)
```

## What this is

Every cell type that appears as an **interior** node on at least one
inhibitory ppk23 -> mAL_m path (from `mal_paths`). Each type is shown with
how many paths it sits on and how much total path strength it carries.
Bars are colored by whether the type is in the GABA sign-inverter pool
(`gaba_sign_inverter_types`), so the reader can see at a glance where the
inhibition is coming from and which types are doing the sign inversion.

## Build the census

```{r census}
inh <- mal_paths %>%
  filter(valence == "inhibitory",
         modality %in% c("ppk23","ppk25"),
         end %in% mal_subtypes) %>%
  mutate(nodes_list = strsplit(path, " -> ", fixed = TRUE))

# Extract interior nodes (drop source + target) per path; keep strength + modality
interior_long <- purrr::imap_dfr(inh$nodes_list, function(nodes, i) {
  if (length(nodes) <= 2) return(tibble())
  ints <- nodes[-c(1, length(nodes))]
  tibble(type     = ints,
         strength = inh$strength[i],
         end      = inh$end[i],
         modality = inh$modality[i])
})

# Attach consensus_nt so we can also show what transmitter each type uses
nt_lookup <- mba %>%
  dplyr::select(type, consensus_nt) %>%
  distinct() %>% group_by(type) %>% slice(1) %>% ungroup()

census_long <- interior_long %>%
  left_join(nt_lookup, by = "type") %>%
  group_by(type, consensus_nt, modality) %>%
  summarise(
    n_paths        = n(),
    n_mAL_targets  = dplyr::n_distinct(end),
    total_strength = sum(strength, na.rm = TRUE),
    .groups = "drop"
  )

# Wide: one row per type, one column per channel; plus combined totals for
# ordering.
census <- census_long %>%
  tidyr::pivot_wider(id_cols = c(type, consensus_nt),
                     names_from = modality,
                     values_from = c(n_paths, total_strength, n_mAL_targets),
                     values_fill = 0) %>%
  mutate(
    is_si  = type %in% gaba_sign_inverter_types,
    is_gaba = !is.na(consensus_nt) & consensus_nt == "gaba",
    role = case_when(
      is_si   ~ "GABA sign-inverter (pool)",
      is_gaba ~ "GABA (not in SI pool)",
      TRUE    ~ sprintf("passthrough (%s)", ifelse(is.na(consensus_nt), "unknown", consensus_nt))
    ),
    total_strength = total_strength_ppk23 + total_strength_ppk25,
    n_paths        = n_paths_ppk23 + n_paths_ppk25
  ) %>%
  arrange(desc(total_strength))

cat("n unique interior types on inhibitory ppk23->mAL paths:", nrow(census), "\n")
cat("of which GABA sign-inverter pool members:", sum(census$is_si), "\n")
cat("of which other GABA cells:", sum(census$is_gaba & !census$is_si), "\n")

# Export
pdf_dir <- panel_plot_dir("panel_M_inhibitory_census", "pdf")
data_dir <- panel_plot_dir("panel_M", "data")
write.csv(census,
          file.path(data_dir, "panel_M_inhibitory_census.csv"),
          row.names = FALSE)
print(head(census, 30))
```

## Plot: top types by total path strength

```{r plot-strength, fig.width=10, fig.height=10}
role_colors <- c(
  "GABA sign-inverter (pool)" = "#1F4E79",
  "GABA (not in SI pool)"     = "#7FA8D1",
  "passthrough (acetylcholine)" = "#C0272D",
  "passthrough (glutamate)"     = "#E69F00",
  "passthrough (gaba)"          = "#4B6FB5",  # fallback (shouldn't occur)
  "passthrough (unknown)"       = "grey55",
  "passthrough (histamine)"     = "#9467BD",
  "passthrough (dopamine)"      = "#8C564B",
  "passthrough (serotonin)"     = "#E377C2",
  "passthrough (octopamine)"    = "#BCBD22"
)

top_n <- 35
top_types <- census %>% slice_head(n = top_n) %>% pull(type)

top_long_strength <- census_long %>%
  filter(type %in% top_types) %>%
  left_join(census %>% select(type, role, is_si), by = "type") %>%
  mutate(type = factor(type, levels = rev(top_types)))

p_strength <- ggplot(top_long_strength,
       aes(x = total_strength, y = type, fill = role)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = ifelse(is_si, "SI", "")),
            x = Inf, hjust = 1.2, size = 3, fontface = "bold",
            color = "#1F4E79", check_overlap = TRUE) +
  facet_wrap(~ modality, nrow = 1,
             labeller = as_labeller(c(
               "ppk23" = "ppk23 (male contact pheromone)",
               "ppk25" = "ppk25 (female contact pheromone)"
             ))) +
  scale_fill_manual(values = role_colors, name = "Role") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.08))) +
  labs(
    x = "Total inhibitory path strength (summed over mAL_m targets)",
    y = NULL,
    title = sprintf("Top %d interior types on ppk23/ppk25 -> mAL_m inhibitory paths", top_n),
    subtitle = "Dark-blue = GABA sign-inverter pool (the actual sign inverters).\nOther colors = non-SI types that appear on these paths (GABA siblings or passthrough cells)."
  ) +
  theme_pub(base_size = 11) +
  theme(
    axis.text.y = element_text(size = rel(1), face = "plain"),
    legend.position = "bottom",
    plot.title.position = "plot",
    plot.subtitle = element_text(size = rel(0.95), color = "grey30")
  )
print(p_strength)
```

## Plot: same census by number of paths

```{r plot-paths, fig.width=10, fig.height=10}
top_types_pc <- census %>% arrange(desc(n_paths)) %>% slice_head(n = top_n) %>% pull(type)

top_long_paths <- census_long %>%
  filter(type %in% top_types_pc) %>%
  left_join(census %>% select(type, role, is_si), by = "type") %>%
  mutate(type = factor(type, levels = rev(top_types_pc)))

p_paths <- ggplot(top_long_paths,
       aes(x = n_paths, y = type, fill = role)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = ifelse(is_si, "SI", "")),
            x = Inf, hjust = 1.2, size = 3, fontface = "bold",
            color = "#1F4E79", check_overlap = TRUE) +
  facet_wrap(~ modality, nrow = 1,
             labeller = as_labeller(c(
               "ppk23" = "ppk23 (male contact pheromone)",
               "ppk25" = "ppk25 (female contact pheromone)"
             ))) +
  scale_fill_manual(values = role_colors, name = "Role") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.08))) +
  labs(
    x = "Number of inhibitory paths this type sits on",
    y = NULL,
    title = sprintf("Top %d interior types by path count", top_n)
  ) +
  theme_pub(base_size = 11) +
  theme(
    axis.text.y = element_text(size = rel(1), face = "plain"),
    legend.position = "bottom",
    plot.title.position = "plot"
  )
print(p_paths)
```

## Combined layout + save

```{r save, fig.width=18, fig.height=11}
p_combined <- (p_strength | p_paths) +
  plot_annotation(
    title    = "Where does the ppk23 inhibition onto mAL_m actually come from?",
    subtitle = paste0("Every cell type that appears as an interior node on at least one ",
                      "inhibitory ppk23 -> mAL_m path (from `mal_paths`). ",
                      "Dark-blue bars are the 11 GABA sign-inverter pool members; ",
                      "everything else is a non-SI type on the same paths."),
    theme = theme(
      plot.title    = element_text(face = "bold", size = 13, hjust = 0),
      plot.subtitle = element_text(size = 10, color = "grey25", hjust = 0,
                                   margin = margin(b = 6))
    )
  )

ggsave(file.path(pdf_dir, "panel_M_inhibitory_census.pdf"),
       plot = p_combined, width = 18, height = 11)

png_dir <- panel_plot_dir("panel_M_inhibitory_census", "png")
ggsave(file.path(png_dir, "panel_M_inhibitory_census.png"),
       plot = p_combined, width = 18, height = 11, dpi = 300)

print(p_combined)
```

## Narrative

```{r narrative, echo=FALSE}
per_chan <- census_long %>%
  mutate(is_si = type %in% gaba_sign_inverter_types) %>%
  group_by(modality) %>%
  summarise(
    si_frac_strength = sum(total_strength[is_si]) / sum(total_strength),
    si_frac_paths    = sum(n_paths[is_si])        / sum(n_paths),
    .groups = "drop"
  )

narrative <- sprintf(
"Panel M: ppk23/ppk25 inhibitory path census
============================================
%d unique cell types appear as interior nodes on inhibitory
ppk23 / ppk25 -> mAL_m paths. %d of them are in the GABA sign-inverter
pool: %s.

Pool contribution per channel:
  ppk23: %.0f%% of inhibitory path strength, %.0f%% of path instances
  ppk25: %.0f%% of inhibitory path strength, %.0f%% of path instances

Remaining non-SI types on these paths are either (a) other GABA cells that
did not make the pool (e.g. local mAL-lateral GABA, non-AN05B/IN05
classes), or (b) excitatory passthrough nodes -- inhibition on those paths
is still contributed by an SI pool member elsewhere in the chain.

Top 5 by combined total path strength (ppk23 + ppk25):
%s
",
nrow(census),
sum(census$is_si),
paste(sort(gaba_sign_inverter_types), collapse = ", "),
100 * per_chan$si_frac_strength[per_chan$modality == "ppk23"],
100 * per_chan$si_frac_paths[   per_chan$modality == "ppk23"],
100 * per_chan$si_frac_strength[per_chan$modality == "ppk25"],
100 * per_chan$si_frac_paths[   per_chan$modality == "ppk25"],
paste(sprintf("  %-14s  strength=%.3f  n_paths=%d  role=%s",
              head(census$type, 5),
              head(census$total_strength, 5),
              head(census$n_paths, 5),
              head(census$role, 5)),
      collapse = "\n")
)
writeLines(narrative,
           file.path(data_dir, "panel_M_inhibitory_census_narrative.txt"))
cat(narrative)
```
