---
title: "Panel M (PRIMARY): inhibitory source attribution (GABA sign-inverter pool vs mAL lateral vs other)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")
library(ggplot2); library(dplyr); library(tidyr); library(patchwork)
```

## Status

**PRIMARY inhibition-attribution figure for Panel M.**
Two sibling figures (`panel_M_E_inhibition_attribution`,
`panel_M_an05b035_fraction_bar`) have been marked deprecated -- they are
strict subsets of this decomposition. The remaining
sibling (`panel_M_an05_share_per_channel`) is kept as a supplementary
absolute-strength view.

## What this is

Panel M's central summary: ppk23 and ppk25 inhibition onto male-specific
mAL_m subtypes comes from TWO main sources Panel M focuses on:

1. **GABA sign-inverter pool** -- a distributed pool of 11 ascending/interneuron
   GABA relay types (AN05B021, AN05B023a-d, AN05B025, AN05B035, AN05B050_a,
   IN05B002, IN05B011a, IN05B011b). These are the cell types defined in
   `gaba_sign_inverter_types` (see `setup_mspecific.R`). Any inhibitory path
   whose interior passes through one of these types is attributed here.
2. **mAL lateral** -- mAL<->mAL GABA inhibition (covered in detail in Panel G).

One additional category is shown for completeness:
- **other** -- paths that reach the target via an unclassified GABA node,
  i.e. neither category 1 nor 2 was present. Kept dim to emphasise sources 1+2.

## Shared palette (used by all four panel_M inhibition-attribution figures)

```{r palette}
# Canonical palette -- identical across panel_M_inhibitory_sources,
# panel_M_E_inhibition_attribution (deprecated), panel_M_an05_share_per_channel,
# and panel_M_an05b035_fraction_bar (deprecated).
pal_inh <- c(
  "GABA sign-inverter" = "#1F4E79",  # deep blue = pool of GABA sign-inverters
  "mAL lateral"        = "grey55",   # grey = covered elsewhere (panel G)
  "other"              = "grey85"    # very dim = unclassified GABA
)

source_levels <- c("GABA sign-inverter", "mAL lateral", "other")

# Informative facet labels (Pass 8)
channel_labeller <- as_labeller(c(
  "ppk23" = "ppk23 channel (male contact pheromone)",
  "ppk25" = "ppk25 channel (female contact pheromone)"
))
```

## Classify paths

```{r classify}
inh_paths <- mal_paths %>%
  filter(valence == "inhibitory",
         modality %in% c("ppk23", "ppk25"),
         end %in% mal_subtypes)

nt_lookup <- mba %>% select(type, consensus_nt) %>% distinct() %>%
  group_by(type) %>% slice(1) %>% ungroup()

classify_relay <- function(node, nt) {
  if (node %in% gaba_sign_inverter_types) return("GABA sign-inverter")
  if (grepl("^mAL", node) && !is.na(nt) && nt == "gaba")
    return("mAL lateral")
  return("other")
}

path_source_weight <- function(path) {
  toks <- strsplit(path, " -> ", fixed = TRUE)[[1]]
  if (length(toks) <= 2) return(NULL)
  ints <- toks[-c(1, length(toks))]
  nts  <- nt_lookup$consensus_nt[match(ints, nt_lookup$type)]
  cats <- mapply(classify_relay, ints, nts)
  cats <- setdiff(unique(cats), "other")
  if (length(cats) == 0) cats <- "other"
  cats
}

inh_attrib <- inh_paths %>%
  rowwise() %>%
  mutate(cats = list(path_source_weight(path))) %>%
  ungroup()

attrib_long <- inh_attrib %>%
  mutate(n_cats = lengths(cats)) %>%
  tidyr::unnest(cats) %>%
  rename(source = cats) %>%
  mutate(contrib = strength / n_cats)

agg <- attrib_long %>%
  group_by(end, modality, source) %>%
  summarise(val = sum(contrib), .groups = "drop")

total_by_cell <- agg %>% group_by(end, modality) %>%
  summarise(tot = sum(val), .groups = "drop")
agg <- agg %>% left_join(total_by_cell, by = c("end", "modality")) %>%
  mutate(share = val / tot)

# Pass 2: sort mAL rows by GABA sign-inverter share (descending)
sort_order <- agg %>%
  filter(source == "GABA sign-inverter") %>%
  group_by(end) %>% summarise(s = sum(share)) %>%
  arrange(desc(s)) %>% pull(end)
# Any mALs missing from the GABA SI slice get appended at the bottom
sort_order <- c(sort_order, setdiff(mal_subtypes, sort_order))
agg$end <- factor(agg$end, levels = rev(sort_order))  # top = highest GABA SI share

agg$source <- factor(agg$source, levels = source_levels)

global_means <- agg %>% group_by(modality, source) %>%
  summarise(mean_share = mean(share) * 100, .groups = "drop")
cat("Mean share (% of inhibitory drive) per source, per channel:\n")
print(global_means %>% pivot_wider(names_from = modality, values_from = mean_share))

# Export shared sort order for supplementary figures (Pass 2 consistency)
saveRDS(sort_order,
        file.path(fig5_dir, "plots_mspecific", "panel_M",
                  "panel_M_mal_sort_order.rds"))
```

## Plot

```{r plot, fig.width=15, fig.height=9}
# Pass 3: percentage labels inside segments > 8% only
agg <- agg %>% mutate(
  lbl = ifelse(share > 0.08,
               sprintf("%.0f%%", share * 100),
               ""),
  lbl_col = ifelse(source %in% c("GABA sign-inverter"), "white", "grey15")
)

p_bars <- ggplot(agg, aes(x = share, y = end, fill = source)) +
  geom_col(position = "stack", width = 0.82) +
  geom_text(aes(label = lbl, color = lbl_col),
            position = position_stack(vjust = 0.5),
            size = 3, family = "sans", show.legend = FALSE) +
  scale_color_identity() +
  facet_wrap(~ modality, nrow = 1, labeller = channel_labeller) +
  scale_fill_manual(values = pal_inh, breaks = source_levels) +
  # Pass 6: explicit [0, 100%] x-axis range
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1.001),
                     breaks = c(0, 0.25, 0.5, 0.75, 1),
                     expand = expansion(mult = c(0, 0.01))) +
  labs(
    x = "Share of total inhibitory path strength (per mAL_m subtype, per channel)",
    y = NULL,
    fill = "Inhibitory source (GABAergic interior node type)",
    title = "Where does the inhibition in each ppk23/ppk25 -> mAL_m path come from?",
    subtitle = paste0(
      "Each ppk -> mAL_m path tagged as inhibitory (from mal_paths) is attributed by ",
      "which GABAergic interior node(s) sit on it.\nExcitatory passthrough nodes along the ",
      "path (e.g. AN09B017 glutamate relays) are NOT shown as a category -- ",
      "they are part of the path but don't contribute to the sign inversion."
    )
  ) +
  theme_pub(base_size = 12) +
  theme(
    # Pass 7: larger mAL subtype labels
    axis.text.y   = element_text(size = rel(1.15), face = "bold"),
    axis.text.x   = element_text(angle = 0, hjust = 0.5),
    strip.text    = element_text(face = "bold", size = rel(1.05)),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

# Footer: sectioned tags for scannability.
footer_note <- paste(
  "INPUT: every cached ppk23/ppk25 -> mAL_m path tagged inhibitory (mal_paths valence == \"inhibitory\"); excitatory paths not shown.",
  panel_M_pool_caption,
  "CATEGORIES: GABA sign-inverter = pool member; mAL lateral = mAL GABA cell (see Panel G); other = unclassified GABA interior.",
  "ATTRIBUTION: each inhibitory path split equally (1/k) across the GABAergic interior categories it traverses.",
  "EXCLUDE: excitatory passthrough nodes (ACh, Glu; e.g. AN09B017) are part of the path but not counted as inhibitory sources.",
  "CAVEAT: equal-split attribution is descriptive; no sensitivity analysis, not a claim about biological weighting.",
  sep = "\n"
)

# Attach the footer as a proper plot caption so there is no secondary
# coordinate system rendered underneath the bars.
p <- p_bars +
  plot_annotation(caption = footer_note,
                  theme = theme(plot.caption = element_text(
                    hjust = 0, size = 9, color = "grey25",
                    lineheight = 1.15, margin = margin(t = 10))))

save_fig5_plot(p, "panel_M_inhibitory_sources", width = 15, height = 9)
print(p)
```

## Narrative

```{r narrative, echo=FALSE}
sum_tab <- global_means %>%
  pivot_wider(names_from = modality, values_from = mean_share)

narrative <- paste0(
"PANEL M (PRIMARY): inhibitory source attribution\n",
"=================================================\n\n",
"For every male-specific mAL_m subtype and each of ppk23 / ppk25, every\n",
"cached inhibitory path is classified by the first relay category it\n",
"passes through:\n",
"  1. GABA sign-inverter pool (11 AN05B/IN05 GABA relay types:\n",
"     AN05B021, AN05B023a-d, AN05B025, AN05B035, AN05B050_a,\n",
"     IN05B002, IN05B011a, IN05B011b)\n",
"  2. mAL lateral (mAL <-> mAL GABA)\n",
"  3. other (unclassified GABA node; not in 1-2)\n",
"Paths crossing multiple categories are split equally across the categories\n",
"they touch (fractional attribution). Rows are sorted by GABA sign-inverter share.\n\n",
"Mean share of inhibitory drive (across ", length(mal_subtypes), " mAL_m subtypes):\n",
paste0(sprintf("  %-20s  ppk23 = %5.1f%%  |  ppk25 = %5.1f%%",
               sum_tab$source, sum_tab$ppk23, sum_tab$ppk25),
       collapse = "\n"),
"\n"
)
save_panel_narrative("panel_M_inhibitory_sources", narrative)
cat(narrative)
```
