---
title: "Figure 5 Panel M (v2): GABA sign-inverter pool morphology -- AN types: T1 input (dorsal) + SEZ input + SEZ output; IN types: VNC input + VNC output (dorsal)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
source("/Users/fkampf/Documents/pheromone.paper/figures/fig5/setup_mspecific.R")

suppressMessages({
  library(nat)
  library(nat.ggplot)
  library(neuprintr)
  library(malecns)
  options(malecns.dataset = "male-cns:v1.0")   # malecns helpers ignore MCNS_DATASET
  library(Rvcg)
  library(dplyr)
  library(tibble)
  library(ggplot2)
  library(patchwork)
  library(Matrix)
})

options(rgl.useNULL = TRUE)
```

## Panel goal

Anatomical "proof of role" panel for the 11-type GABA sign-inverter (SI) pool.

AN types (ascending, prefix "AN05"): three-panel row -- (1) where
ppk23/ppk25 contact input arrives on SI dendrites in the T1 leg neuromere
(DORSAL view, looking down on VNC), (2) where the AN neuron additionally
receives input in the SEZ (GNG + PRW, frontal view), (3) where the signal
is projected onto mAL_m in the SEZ (frontal view).

IN types (VNC-intrinsic interneurons, prefix "IN05"): two-panel row --
(1) inputs in VNC (DORSAL view), (2) outputs in VNC (DORSAL view). IN
neurons do not ascend to the SEZ, so the two SEZ panels are omitted.

Only pheromone-channel categories are plotted -- "other" is dropped. Neuron
morphology is rendered as smooth 3D surface meshes (fetched via
`malecns::read_mcns_meshes` from DVID and lightly decimated via
`Rvcg::vcgQEdecim` to 35% of faces, after `Rvcg::vcgClean` topology fix and
`Rvcg::vcgIsolated` orphan-component removal so thin neuron branches render
continuously).

## Pool bodyids

```{r ids}
si_meta <- mba %>%
  dplyr::filter(type %in% gaba_sign_inverter_types, !is.na(bodyid)) %>%
  dplyr::select(bodyid, type) %>%
  dplyr::distinct()
si_ids    <- si_meta$bodyid
n_neurons <- nrow(si_meta)
n_types   <- length(unique(si_meta$type))
cat(sprintf("SI pool: %d neurons across %d types\n", n_neurons, n_types))
print(si_meta %>% dplyr::count(type))
```

## Partner-type category lookups

```{r categorise-types}
ppk23_types <- channel_neurons[["ppk23"]]
ppk25_types <- channel_neurons[["ppk25"]]

an09_types <- c("AN09B017a","AN09B017b","AN09B017c","AN09B017d",
                "AN09B017e","AN09B017f","AN09B017g")

an09_bias <- sapply(an09_types, function(t) {
  p23 <- conn %>% filter(post_type == t, pre_type %in% ppk23_types) %>%
    summarise(w = sum(weight, na.rm = TRUE)) %>% pull(w)
  p25 <- conn %>% filter(post_type == t, pre_type %in% ppk25_types) %>%
    summarise(w = sum(weight, na.rm = TRUE)) %>% pull(w)
  if (length(p23) == 0) p23 <- 0
  if (length(p25) == 0) p25 <- 0
  log2((p23 + 1) / (p25 + 1))
})
relay_ppk23    <- names(an09_bias)[an09_bias >  1]
relay_ppk25    <- names(an09_bias)[an09_bias < -1]
relay_balanced <- names(an09_bias)[abs(an09_bias) <= 1]

categorise_input <- function(partner_type) {
  dplyr::case_when(
    partner_type %in% ppk23_types    ~ "ppk23 sensillum",
    partner_type %in% ppk25_types    ~ "ppk25 sensillum",
    partner_type %in% relay_ppk23    ~ "ppk23-biased relay",
    partner_type %in% relay_ppk25    ~ "ppk25-biased relay",
    partner_type %in% relay_balanced ~ "ppk-unbiased relay",
    TRUE ~ NA_character_
  )
}

# Unified output categorisation for BOTH AN and IN SI types. An output
# synapse onto `partner_type` is kept iff partner is the IMMEDIATE NEXT
# HOP after `si_type` on some ppk23/ppk25 -> mAL_m path in mal_paths AND
# either (a) partner is itself a male-specific mAL_m terminus or
# (b) partner has a direct 1-hop synapse onto a mAL_m subtype. mAL_m
# termini get colour-split by ppk23 vs ppk25 bias; non-terminal next hops
# are all "next hop that reaches mAL_m".
# (next_hop_sets / reaches_mAL are defined below and used inside this fn.)
next_hop_sets <- local({
  paths <- mal_paths %>%
    dplyr::filter(modality %in% c("ppk23","ppk25"),
                  end %in% mal_subtypes) %>%
    dplyr::pull(path)
  acc <- list()
  for (p in paths) {
    toks <- strsplit(p, " -> ", fixed = TRUE)[[1]]
    if (length(toks) < 2) next
    for (i in seq_len(length(toks) - 1)) {
      u <- toks[i]
      v <- toks[i + 1]
      if (u %in% gaba_sign_inverter_types) {
        acc[[u]] <- unique(c(acc[[u]], v))
      }
    }
  }
  acc
})

# Partners that directly synapse onto at least one mAL_m subtype.
# (1-hop adjacency from adj.matrix.raw.)
reaches_mAL <- local({
  types_row <- intersect(rownames(adj.matrix.raw), unique(unlist(next_hop_sets)))
  if (length(types_row) == 0) return(character(0))
  m <- adj.matrix.raw[types_row, intersect(mal_subtypes, colnames(adj.matrix.raw)),
                      drop = FALSE]
  types_row[Matrix::rowSums(m) > 0]
})

categorise_output_SI <- function(partner_type, si_type) {
  if (is.na(partner_type) || is.na(si_type)) return(NA_character_)
  hits <- next_hop_sets[[si_type]]
  if (is.null(hits) || !(partner_type %in% hits)) return(NA_character_)
  # Keep: partner is a male-specific mAL_m terminus, OR partner reaches mAL_m
  # directly at 1 hop.
  is_mAL_m_terminus <- partner_type %in% mal_subtypes
  if (!is_mAL_m_terminus && !(partner_type %in% reaches_mAL)) return(NA_character_)

  if (is_mAL_m_terminus) {
    p23 <- mal_activation_matrix[partner_type, "ppk23"]
    p25 <- mal_activation_matrix[partner_type, "ppk25"]
    if (is.na(p23)) p23 <- 0
    if (is.na(p25)) p25 <- 0
    if (p23 == 0 && p25 == 0) return("next hop that reaches mAL_m")
    if (p23 >= p25) return("mAL_m (ppk23-biased)")
    return("mAL_m (ppk25-biased)")
  }
  "next hop that reaches mAL_m"
}
```

## Shells (brain + VNC)

```{r fetch-shell}
scale_shell <- function(s) {
  s$Vertices$X <- s$Vertices$X / 8
  s$Vertices$Y <- s$Vertices$Y / 8
  s$Vertices$Z <- s$Vertices$Z / 8
  s
}
shell_brain <- scale_shell(malecns::malecns_shell.surf)
shell_vnc   <- scale_shell(malecns::malecnsvnc_shell.surf)
```

## ROI bounding boxes for cropping (no hull drawn)

```{r fetch-rois}
t1_rois  <- c("LegNp(T1)(L)", "LegNp(T1)(R)",
              "NTct(UTct-T1)(L)", "NTct(UTct-T1)(R)")
sez_rois <- c("GNG", "PRW")

fetch_roi_verts <- function(roi) {
  m <- tryCatch(neuprint_ROI_mesh(roi), error = function(e) NULL)
  if (is.null(m)) return(NULL)
  v <- as.data.frame(t(m$vb[1:3, , drop = FALSE]))
  colnames(v) <- c("x","y","z")
  v
}

# Frontal projection: axes (x, y). Used for brain-SEZ panels.
verts_to_bbox_xy <- function(verts, pad_frac = 0.08) {
  if (is.null(verts) || nrow(verts) == 0) return(NULL)
  xr <- range(verts$x, na.rm = TRUE)
  yr <- range(verts$y, na.rm = TRUE)
  px <- pad_frac * diff(xr); py <- pad_frac * diff(yr)
  list(x = c(xr[1] - px, xr[2] + px),
       y = c(yr[1] - py, yr[2] + py))
}

# Dorsal projection: axes (x, z). The 2D plot still uses variable names
# x / y but we populate "y" from the raw z component. Used for VNC panels.
verts_to_bbox_xz <- function(verts, pad_frac = 0.08) {
  if (is.null(verts) || nrow(verts) == 0) return(NULL)
  xr <- range(verts$x, na.rm = TRUE)
  zr <- range(verts$z, na.rm = TRUE)
  px <- pad_frac * diff(xr); pz <- pad_frac * diff(zr)
  list(x = c(xr[1] - px, xr[2] + px),
       y = c(zr[1] - pz, zr[2] + pz))
}

# SEZ: brain ROIs (GNG, PRW) come back from neuprint in 8-nm voxel coords,
# matching the synapse coordinate system. Use them directly.
sez_verts        <- dplyr::bind_rows(lapply(sez_rois, fetch_roi_verts))
bbox_sez_frontal <- verts_to_bbox_xy(sez_verts)

# T1: VNC ROI meshes from neuprint come back in non-voxel units in this
# dataset; instead, derive the VNC bounding box from the rescaled VNC shell
# (already in 8-nm voxel coords).
vnc_v <- data.frame(
  x = malecns::malecnsvnc_shell.surf$Vertices$X / 8,
  y = malecns::malecnsvnc_shell.surf$Vertices$Y / 8,
  z = malecns::malecnsvnc_shell.surf$Vertices$Z / 8
)
bbox_vnc_frontal <- verts_to_bbox_xy(vnc_v, pad_frac = 0.05)
bbox_vnc_dorsal  <- verts_to_bbox_xz(vnc_v, pad_frac = 0.05)

cat("VNC frontal bbox  x:", bbox_vnc_frontal$x, " y(=y):", bbox_vnc_frontal$y, "\n")
cat("VNC dorsal  bbox  x:", bbox_vnc_dorsal$x,  " y(=z):", bbox_vnc_dorsal$y,  "\n")
cat("SEZ frontal bbox  x:", bbox_sez_frontal$x, " y(=y):", bbox_sez_frontal$y, "\n")
```

## Meshes + synapses

```{r fetch}
# Fetch full 3D surface meshes per SI bodyid from DVID via malecns, in the
# same raw voxel coordinate system as synapses and the rescaled shells.
# Mesh is lightly decimated (keep 15% of faces) via Rvcg::vcgQEdecim so it
# stays smooth -- much less aggressive than the previous 0.3% attempt.
# Cached per bodyid.
mesh_cache_dir <- file.path(fig5_dir, "panel_M", "cache", "meshes")
dir.create(mesh_cache_dir, recursive = TRUE, showWarnings = FALSE)

DECIM_PERCENT  <- 0.35   # keep 35% of faces -- gentler, preserves thin branches
MIN_FACES_SKIP <- 20000  # skip decimation for small+medium meshes

decimate_mesh <- function(m, pct = DECIM_PERCENT, min_faces = MIN_FACES_SKIP) {
  if (is.null(m)) return(NULL)
  # 1. Clean topology: remove duplicate verts (0), non-manifold edges (1),
  #    small connected components (6), unreferenced verts (7). Fixes the
  #    gap artefacts where geom_polygon renders degenerate triangles as
  #    lines / outline-only fragments.
  m <- tryCatch(
    Rvcg::vcgClean(m, sel = c(0, 1, 6, 7), silent = TRUE),
    error = function(e) m
  )
  # 2. Drop orphan islands: keep only components with >= 50 faces so tiny
  #    disconnected fragments do not render as speckle/outlines.
  m <- tryCatch(
    Rvcg::vcgIsolated(m, facenum = 50, silent = TRUE),
    error = function(e) m
  )
  # 3. Decimate only if the mesh is big enough to warrant it.
  nfaces <- tryCatch(ncol(m$it), error = function(e) NA_integer_)
  if (is.na(nfaces) || nfaces < min_faces || pct >= 1) return(m)
  tryCatch(
    Rvcg::vcgQEdecim(m, percent = pct, topo = TRUE, silent = TRUE),
    error = function(e) { message("decimate failed: ", conditionMessage(e)); m }
  )
}

fetch_one_mesh <- function(id) {
  f <- file.path(mesh_cache_dir, paste0(id, ".rds"))
  if (file.exists(f)) {
    m <- tryCatch(readRDS(f), error = function(e) NULL)
    if (inherits(m, "mesh3d")) return(m)
    file.remove(f)
  }
  m <- tryCatch(
    malecns::read_mcns_meshes(id, units = "raw", type = "dvid"),
    error = function(e) { message("mesh fetch failed for ", id, ": ", conditionMessage(e)); NULL }
  )
  if (is.null(m)) return(NULL)
  # read_mcns_meshes returns either a single mesh3d or a neuronlist-like
  # collection. Normalise to a single mesh3d for this one bodyid.
  if (inherits(m, "mesh3d")) {
    mesh <- m
  } else if (is.list(m) && length(m) > 0) {
    cand <- m[[1]]
    mesh <- if (inherits(cand, "mesh3d")) cand else NULL
  } else mesh <- NULL
  if (is.null(mesh)) return(NULL)
  mesh <- decimate_mesh(mesh)
  saveRDS(mesh, f)
  mesh
}

mesh_list <- list()
for (id in si_ids) {
  m <- fetch_one_mesh(id)
  if (!is.null(m)) mesh_list[[as.character(id)]] <- m
}

have_meshes <- length(mesh_list) > 0
cat("mesh fetch:", have_meshes, "  (", if (have_meshes) length(mesh_list) else 0, " neurons)\n")
if (have_meshes) {
  face_counts <- sapply(mesh_list, function(m) tryCatch(ncol(m$it), error = function(e) NA_integer_))
  cat("face counts (post-decimation): min=", min(face_counts, na.rm = TRUE),
      " median=", stats::median(face_counts, na.rm = TRUE),
      " max=", max(face_counts, na.rm = TRUE), "\n", sep = "")
}

syn_tbl <- tryCatch(
  neuprint_get_synapses(si_ids),
  error = function(e) { message("synapse fetch failed: ", conditionMessage(e)); NULL }
)

have_data <- have_meshes && !is.null(syn_tbl) && nrow(syn_tbl) > 0
cat("meshes OK:", have_meshes,
    "  synapses OK:", !is.null(syn_tbl) && nrow(syn_tbl) > 0, "\n")
```

## Categorise synapses (drop "other" entirely)

```{r categorise-synapses}
bodyid_to_type <- setNames(mba$type, as.character(mba$bodyid))

if (have_data) {
  syn_df <- syn_tbl %>% as_tibble() %>%
    mutate(
      partner_type = bodyid_to_type[as.character(partner)],
      io           = ifelse(prepost == 1, "input", "output"),
      # SI-side cell type (which SI neuron this synapse is on)
      si_type      = bodyid_to_type[as.character(bodyid)]
    )

  syn_df$category <- NA_character_
  in_idx  <- which(syn_df$io == "input")
  out_idx <- which(syn_df$io == "output")

  # Input categorisation unchanged for both AN and IN cells.
  syn_df$category[in_idx]  <- categorise_input(syn_df$partner_type[in_idx])

  # Output categorisation: same rule for AN and IN -- keep only partners that
  # are a male-specific mAL_m terminus OR directly reach mAL_m at 1 hop,
  # AND are the immediate next hop after this SI type on some
  # ppk23/ppk25 -> mAL_m path.
  syn_df$category[out_idx] <- mapply(
    categorise_output_SI,
    syn_df$partner_type[out_idx],
    syn_df$si_type[out_idx]
  )

  # Drop "other": keep only synapses with a pheromone-channel category.
  syn_df <- syn_df %>% filter(!is.na(category))

  syn_df$category <- factor(
    syn_df$category,
    levels = c(
      "ppk23 sensillum", "ppk25 sensillum",
      "ppk23-biased relay", "ppk25-biased relay", "ppk-unbiased relay",
      "mAL_m (ppk23-biased)", "mAL_m (ppk25-biased)",
      "next hop that reaches mAL_m"
    )
  )

  cat("\n== INPUT breakdown (all) ==\n");  print(syn_df %>% filter(io == "input")  %>% count(category))
  cat("\n== OUTPUT breakdown (all) ==\n"); print(syn_df %>% filter(io == "output") %>% count(category))

  # Per-SI-type OUTPUT breakdown across the three output categories.
  cat("\n== Per-SI-type OUTPUT breakdown (counts) ==\n")
  .out_cats <- c("mAL_m (ppk23-biased)",
                 "mAL_m (ppk25-biased)",
                 "next hop that reaches mAL_m")
  .per_type_out <- syn_df %>%
    filter(io == "output", category %in% .out_cats) %>%
    count(si_type, category) %>%
    tidyr::pivot_wider(names_from = category, values_from = n, values_fill = 0)
  # Ensure every SI type shows up, even with zero outputs.
  .all_types <- sort(unique(si_meta$type))
  .per_type_out <- tibble::tibble(si_type = .all_types) %>%
    dplyr::left_join(.per_type_out, by = "si_type")
  for (.c in .out_cats) {
    if (!(.c %in% colnames(.per_type_out))) .per_type_out[[.c]] <- 0L
    .per_type_out[[.c]][is.na(.per_type_out[[.c]])] <- 0L
  }
  .per_type_out <- .per_type_out[, c("si_type", .out_cats)]
  print(as.data.frame(.per_type_out), row.names = FALSE)

  # Diagnostic: per-IN-type next-hop sets + output-synapse counts.
  cat("\n== IN next-hop sets (ppk23/ppk25 -> mAL_m, from mal_paths) ==\n")
  cat(sprintf("  reaches_mAL pool size: %d types\n", length(reaches_mAL)))
  for (.tp in sort(unique(si_meta$type[grepl("^IN", si_meta$type)]))) {
    .hits     <- next_hop_sets[[.tp]]
    .filtered <- if (is.null(.hits)) character(0)
                 else intersect(.hits, reaches_mAL)
    .n_syn <- syn_df %>%
      filter(io == "output", si_type == .tp,
             category == "next hop that reaches mAL_m") %>%
      nrow()
    cat(sprintf("  %-12s  n_next_hop_types = %d  n_reaches_mAL = %d  n_output_syn = %d\n",
                .tp,
                if (is.null(.hits)) 0L else length(.hits),
                length(.filtered),
                .n_syn))
    if (!is.null(.hits) && length(.hits) > 0) {
      cat(sprintf("     next-hop types: %s\n", paste(.hits, collapse = ", ")))
    }
    if (length(.filtered) > 0) {
      cat(sprintf("     filtered (reach mAL_m): %s\n",
                  paste(.filtered, collapse = ", ")))
    } else {
      cat("     filtered (reach mAL_m): <EMPTY> -- panel will have no next-hop synapses\n")
    }

    # Partner-type breakdown of passing output synapses (sample top 5).
    .passing_partners <- syn_df %>%
      filter(io == "output", si_type == .tp,
             category == "next hop that reaches mAL_m") %>%
      count(partner_type, sort = TRUE)
    if (nrow(.passing_partners) > 0) {
      .top <- head(.passing_partners, 5)
      cat(sprintf("     top partners by synapse count: %s\n",
                  paste(sprintf("%s (%d)", .top$partner_type, .top$n),
                        collapse = ", ")))
    }
  }
}
```

## Palette

```{r palette}
# Inputs: ppk23/ppk25 sensilla (red/orange from panel_M_palette); relays blue/green.
# Outputs (mAL_m targets): DISTINCT warm+cool duo so the SI -> mAL_m projection
# does not visually collide with the ppk23/ppk25 sensillum inputs.
cat_colors <- c(
  "ppk23 sensillum"              = unname(panel_M_palette["ppk23"]),
  "ppk25 sensillum"              = unname(panel_M_palette["ppk25"]),
  "ppk23-biased relay"           = "#377EB8",
  "ppk25-biased relay"           = "#4DAF4A",
  "ppk-unbiased relay"           = "#888888",
  "mAL_m (ppk23-biased)"         = "#8B2C7F",
  "mAL_m (ppk25-biased)"         = "#16A085",
  "next hop that reaches mAL_m"  = unname(panel_M_palette["gaba_si"])
)
```

## Per-panel synapse subsets (cropped to panel bbox)

```{r crop-to-bbox}
# Frontal projection: filter on (x, y) where bbox$y is the raw Y range.
in_bbox_frontal <- function(df, bbox) {
  df %>% filter(x >= bbox$x[1], x <= bbox$x[2],
                y >= bbox$y[1], y <= bbox$y[2])
}
# Dorsal projection: filter on (x, z) where bbox$y is actually the raw Z range.
in_bbox_dorsal <- function(df, bbox) {
  df %>% filter(x >= bbox$x[1], x <= bbox$x[2],
                z >= bbox$y[1], z <= bbox$y[2])
}
```

## Rotation matrices (frontal vs dorsal view)

```{r rotation-matrices}
# Frontal view: identity 4x4, plot (x, y).
R_frontal <- diag(4)

# Dorsal view: rotate -90 deg about X so y -> z and z -> -y. In the resulting
# 3D frame, what was raw z ends up in the plot's y slot. For the 2D projection
# we therefore plot a = x, b = raw z (and treat nat.ggplot::geom_neuron's
# rotation_matrix as doing the same transform to the mesh vertices).
R_dorsal <- matrix(c(1,  0, 0, 0,
                     0,  0, 1, 0,
                     0, -1, 0, 0,
                     0,  0, 0, 1), nrow = 4, byrow = TRUE)
```

## Render function: single panel for one SI cell type

```{r render-fn}
# Render one cell-type-specific panel. Uses 3D surface meshes fetched via
# read_mcns_meshes + decimated with Rvcg::vcgQEdecim, drawn by
# nat.ggplot::geom_neuron.mesh3d. Pheromone-channel synapses are layered on
# top as coloured points. The R argument selects frontal (XY) vs dorsal (XZ)
# projection; meshes and shells are rotated by R via
# nat.ggplot::geom_neuron(rotation_matrix = R), and raw 3D synapse coords
# are projected by picking y or z for the vertical plot axis accordingly.
render_panel <- function(meshes_this_type, syn_subset, title, bbox,
                         R = R_frontal,
                         scale_bar_um = 20, voxel_nm = 8,
                         syn_size = 1.2, syn_alpha = 0.5,
                         show_scalebar = TRUE,
                         show_title = TRUE,
                         show_legend = FALSE) {
  # Project raw 3D synapse coords onto the 2D plot plane. For the dorsal view
  # we feed the raw z into the plot's y slot so synapse positions stay
  # registered with the rotated meshes.
  is_dorsal <- identical(R, R_dorsal)
  syn_xy <- if (is_dorsal) {
    syn_subset %>% transmute(x = x, y = z, category = category)
  } else {
    syn_subset %>% transmute(x = x, y = y, category = category)
  }

  scale_units <- scale_bar_um * 1000 / voxel_nm
  pad_x <- 0.03 * diff(bbox$x)
  pad_y <- 0.03 * diff(bbox$y)
  sb_x0 <- bbox$x[2] - scale_units - pad_x
  sb_x1 <- bbox$x[2] - pad_x
  sb_y  <- bbox$y[2] - pad_y
  sb_lab <- sprintf("%d um", scale_bar_um)

  # Neuron bodies as decimated 3D meshes rendered through
  # nat.ggplot::geom_neuron.mesh3d. One layer per mesh so each neuron of
  # the type contributes its own filled shape.
  # Wrap nat.ggplot::geom_neuron so the scale_fill_gradient it returns has
  # `guide = guide_none()` -- that's the fix for the leftover "Z" colorbars
  # in the combined legend. show.legend can't suppress a scale-level guide,
  # but modifying $guide on the returned Scale object does.
  geom_neuron_nobar <- function(...) {
    layers <- nat.ggplot::geom_neuron(...)
    for (i in seq_along(layers)) {
      if (inherits(layers[[i]], "Scale")) {
        layers[[i]]$guide <- ggplot2::guide_none()
      }
    }
    layers
  }

  neuron_layers <- list()
  if (length(meshes_this_type) > 0) {
    for (nm in names(meshes_this_type)) {
      m <- meshes_this_type[[nm]]
      if (is.null(m)) next
      neuron_layers[[length(neuron_layers) + 1]] <-
        geom_neuron_nobar(
          x = m, rotation_matrix = R,
          cols = c("grey60","grey40"),
          alpha = 0.35,
          show.legend = FALSE
        )
    }
  }

  p <- ggplot() +
    # Faint shells for anatomical reference.
    geom_neuron_nobar(x = shell_brain, rotation_matrix = R,
                      cols = c("grey88","grey80"), size = 0.03, alpha = 0.08,
                      show.legend = FALSE) +
    geom_neuron_nobar(x = shell_vnc,   rotation_matrix = R,
                      cols = c("grey88","grey80"), size = 0.03, alpha = 0.08,
                      show.legend = FALSE) +
    neuron_layers +
    # Synapses on top.
    geom_point(data = syn_xy,
               aes(x = x, y = y, color = category),
               size = syn_size, alpha = syn_alpha, shape = 16, stroke = 0,
               show.legend = show_legend) +
    scale_color_manual(values = cat_colors, drop = FALSE, name = NULL) +
    coord_fixed(xlim = bbox$x, ylim = bbox$y) +
    scale_y_reverse() +
    theme_void(base_size = 9) +
    theme(
      plot.title      = if (show_title)
        element_text(size = 9.5, face = "plain",
                     margin = margin(b = 2), hjust = 0.5) else element_blank(),
      legend.position = "bottom",
      legend.text     = element_text(size = 8),
      legend.key.size = unit(0.35, "cm"),
      axis.text       = element_blank(),
      axis.ticks      = element_blank(),
      axis.title      = element_blank()
    ) +
    guides(fill = "none", size = "none", alpha = "none")

  if (show_title) p <- p + labs(title = title)
  if (show_scalebar) {
    p <- p +
      annotate("segment", x = sb_x0, xend = sb_x1, y = sb_y, yend = sb_y,
               linewidth = 0.8, color = "black") +
      annotate("text", x = (sb_x0 + sb_x1)/2, y = sb_y,
               label = sb_lab, vjust = -0.6, size = 2.4)
  }
  p
}

# Row-label "plot": a tiny ggplot showing just the cell-type name in bold.
row_label_plot <- function(type_name, n_cells) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.55,
             label = type_name,
             fontface = "bold", size = 3.4, hjust = 0.5) +
    annotate("text", x = 0.5, y = 0.35,
             label = sprintf("(n = %d)", n_cells),
             size = 2.6, color = "grey40", hjust = 0.5) +
    xlim(0, 1) + ylim(0, 1) +
    theme_void()
}
```

## Build + save: one PDF per SI cell type

```{r build-and-save}
pdf_dir <- panel_plot_dir("panel_M", "pdf")
png_dir <- panel_plot_dir("panel_M", "png")

# Remove any stale combined output from the previous one-big-PDF version.
old_pdf <- file.path(pdf_dir, "panel_M_morphology_v2.pdf")
old_png <- file.path(png_dir, "panel_M_morphology_v2.png")
if (file.exists(old_pdf)) file.remove(old_pdf)
if (file.exists(old_png)) file.remove(old_png)

pdf_dev <- if (capabilities("cairo")) cairo_pdf else pdf

pool_list_str <- paste(sort(gaba_sign_inverter_types), collapse = ", ")

per_type_caption <- function(tp, n_cells) {
  paste(
    sprintf("POOL: GABA sign-inverter pool (%d types) -- %s.",
            length(gaba_sign_inverter_types), pool_list_str),
    sprintf("THIS TYPE: %s (n = %d neurons).", tp, n_cells),
    panel_M_anatomy_legend,
    "METHODS: meshes via read_mcns_meshes (DVID), cleaned via Rvcg::vcgClean (sel=0,1,6,7) + Rvcg::vcgIsolated (facenum=50) to remove duplicate verts / non-manifold edges / orphan islands, then decimated to 35% (topo=TRUE, min 20000 faces) via Rvcg::vcgQEdecim, rendered through nat.ggplot::geom_neuron.mesh3d (cols=grey60/grey40, alpha=0.35). Meshes cached per bodyid in figures/fig5/panel_M/cache/meshes/.",
    "SHELLS: malecns::malecns_shell.surf + malecnsvnc_shell.surf (faint background), rescaled /8 to voxel coords.",
    "SYNAPSES: neuprint_get_synapses, classified by partner type; 'other' dropped. INPUTS: partner bias from channel_neurons + AN09B017 relay (from conn). OUTPUTS: partner bias from mal_activation_matrix (ppk23 vs ppk25).",
    "COLORS: ppk23 sensillum red, ppk25 sensillum orange, ppk23-biased relay blue #377EB8, ppk25-biased relay green #4DAF4A, mAL_m (ppk23-biased) magenta #8B2C7F, mAL_m (ppk25-biased) teal #16A085.",
    "PER-PANEL SYNAPSE STYLE: T1-in size=1.2 alpha=0.4 (dorsal, rotation_matrix = -90deg about X, plot x vs raw z); SEZ-in size=1.5 alpha=0.7 (frontal); SEZ-out size=1.3 alpha=0.6 (frontal). IN types: two dorsal VNC panels side-by-side, inputs (size=1.3 alpha=0.5) and outputs (size=1.3 alpha=0.6); no SEZ. SCALE: 20 um per panel.",
    sep = "\n")
}

written_pdfs <- character(0)
written_pngs <- character(0)

if (have_data) {
  types_sorted <- sort(unique(si_meta$type))

  for (tp in types_sorted) {
    ids_this_int <- si_meta %>% filter(type == tp) %>% pull(bodyid)
    ids_this     <- as.character(ids_this_int)
    n_cells      <- length(ids_this)

    # Meshes for this type only -- filtered by bodyid.
    meshes_this_type <- mesh_list[ids_this[ids_this %in% names(mesh_list)]]

    syn_this <- syn_df %>% filter(si_type == tp)

    is_IN <- grepl("^IN", tp)

    if (is_IN) {
      # ---------------------------------------------------------------
      # IN types: VNC-intrinsic -- no SEZ panels. Two side-by-side
      # dorsal VNC panels: inputs and outputs.
      # ---------------------------------------------------------------
      syn_in_vnc  <- syn_this %>% filter(io == "input")  %>%
        in_bbox_dorsal(bbox_vnc_dorsal)
      syn_out_vnc <- syn_this %>% filter(io == "output") %>%
        in_bbox_dorsal(bbox_vnc_dorsal)

      p_in_vnc  <- render_panel(meshes_this_type, syn_in_vnc,
                                "Inputs in VNC (dorsal view)",
                                bbox   = bbox_vnc_dorsal,
                                R      = R_dorsal,
                                syn_size = 1.3, syn_alpha = 0.5,
                                show_scalebar = TRUE,
                                show_title    = TRUE,
                                show_legend   = TRUE)
      p_out_vnc <- render_panel(meshes_this_type, syn_out_vnc,
                                "Outputs in VNC (dorsal view)",
                                bbox   = bbox_vnc_dorsal,
                                R      = R_dorsal,
                                syn_size = 1.3, syn_alpha = 0.6,
                                show_scalebar = TRUE,
                                show_title    = TRUE,
                                show_legend   = FALSE)

      combined <- wrap_plots(list(p_in_vnc, p_out_vnc), nrow = 1, guides = "collect") +
        plot_annotation(
          title    = sprintf("%s morphology  (n = %d)  --  VNC-intrinsic interneuron",
                             tp, n_cells),
          subtitle = "GABA sign-inverter pool member -- inputs and outputs in VNC",
          caption  = per_type_caption(tp, n_cells),
          theme    = theme(
            plot.title    = element_text(face = "bold", size = 12, hjust = 0,
                                         margin = margin(b = 2)),
            plot.subtitle = element_text(size = 9, color = "grey25", hjust = 0,
                                         margin = margin(b = 4)),
            plot.caption  = element_text(size = 7, color = "grey35",
                                         hjust = 0, lineheight = 1.2,
                                         margin = margin(t = 6))
          )
        ) &
        theme(legend.position = "bottom",
              legend.text     = element_text(size = 8),
              legend.key.size = unit(0.35, "cm")) &
        guides(
          color = guide_legend(
            override.aes = list(size = 3, alpha = 1, shape = 16),
            nrow = 1, byrow = TRUE),
          fill  = "none", alpha = "none", size = "none"
        )

      fig_w <- 12
      fig_h <- 12
    } else {
      # ---------------------------------------------------------------
      # AN types: ascending -- three-panel row.
      #   p1: input in T1 (dorsal view)
      #   p2: input in SEZ (frontal view)
      #   p3: output in SEZ -- mAL_m (frontal view)
      # Only panel 1 contributes the legend; panels 2/3 suppress it so
      # patchwork's guides = "collect" yields ONE shared legend at the
      # bottom.
      # ---------------------------------------------------------------
      syn_in_t1   <- syn_this %>% filter(io == "input")  %>%
        in_bbox_dorsal(bbox_vnc_dorsal)
      syn_in_sez  <- syn_this %>% filter(io == "input")  %>%
        in_bbox_frontal(bbox_sez_frontal)
      syn_out_sez <- syn_this %>% filter(io == "output") %>%
        in_bbox_frontal(bbox_sez_frontal)

      p1 <- render_panel(meshes_this_type, syn_in_t1,
                         "Inputs in T1 (leg neuromere, dorsal view)",
                         bbox   = bbox_vnc_dorsal,
                         R      = R_dorsal,
                         syn_size = 1.2, syn_alpha = 0.4,
                         show_scalebar = TRUE,
                         show_title = TRUE, show_legend = TRUE)
      p2 <- render_panel(meshes_this_type, syn_in_sez,
                         "Inputs in SEZ (GNG + PRW)",
                         bbox   = bbox_sez_frontal,
                         R      = R_frontal,
                         syn_size = 1.5, syn_alpha = 0.7,
                         show_scalebar = TRUE,
                         show_title = TRUE, show_legend = FALSE)
      p3 <- render_panel(meshes_this_type, syn_out_sez,
                         expression("Outputs in SEZ" %->% "mAL_m"),
                         bbox   = bbox_sez_frontal,
                         R      = R_frontal,
                         syn_size = 1.3, syn_alpha = 0.6,
                         show_scalebar = TRUE,
                         show_title = TRUE, show_legend = FALSE)

      combined <- wrap_plots(list(p1, p2, p3), nrow = 1, guides = "collect") +
        plot_annotation(
          title    = sprintf("%s morphology  (n = %d)", tp, n_cells),
          subtitle = "GABA sign-inverter pool member -- ppk23/ppk25 contact input and mAL_m projection",
          caption  = per_type_caption(tp, n_cells),
          theme    = theme(
            plot.title    = element_text(face = "bold", size = 12, hjust = 0,
                                         margin = margin(b = 2)),
            plot.subtitle = element_text(size = 9, color = "grey25", hjust = 0,
                                         margin = margin(b = 4)),
            plot.caption  = element_text(size = 7, color = "grey35",
                                         hjust = 0, lineheight = 1.2,
                                         margin = margin(t = 6))
          )
        ) &
        theme(legend.position = "bottom",
              legend.text     = element_text(size = 8),
              legend.key.size = unit(0.35, "cm")) &
        guides(
          color = guide_legend(
            override.aes = list(size = 3, alpha = 1, shape = 16),
            nrow = 1, byrow = TRUE),
          fill  = "none", alpha = "none", size = "none"
        )

      fig_w <- 14
      fig_h <- 6
    }

    pdf_path <- file.path(pdf_dir, sprintf("panel_M_morphology_%s.pdf", tp))
    png_path <- file.path(png_dir, sprintf("panel_M_morphology_%s.png", tp))

    ggsave(pdf_path, plot = combined, width = fig_w, height = fig_h,
           device = pdf_dev, limitsize = FALSE)
    ggsave(png_path, plot = combined, width = fig_w, height = fig_h,
           dpi = 200, limitsize = FALSE)

    written_pdfs <- c(written_pdfs, pdf_path)
    written_pngs <- c(written_pngs, png_path)

    cat(sprintf("wrote %s  (%.2f MB)  [%s, %dx%d]\n",
                pdf_path, file.info(pdf_path)$size / 1e6,
                if (is_IN) "IN / 2 panels" else "AN / 3 panels",
                fig_w, fig_h))
  }
} else {
  fallback <- ggplot() + theme_void() +
    annotate("text", x = 0, y = 0.5,
             label = "SI pool morphology -- data unavailable",
             size = 6, fontface = "bold") +
    xlim(-1, 1) + ylim(0, 1)
  pdf_path <- file.path(pdf_dir, "panel_M_morphology_NA.pdf")
  ggsave(pdf_path, plot = fallback, width = 14, height = 6,
         device = pdf_dev, limitsize = FALSE)
  cat("data unavailable -- wrote placeholder ", pdf_path, "\n", sep = "")
}

cat("\nTotal PDFs written:", length(written_pdfs), "\n")
```
